### 2013-01-17 docs/paper.v0.tex
add kathryn's draft to repo

%\documentclass{emulateapj}
%\documentclass[preprint]{aastex}
\documentclass{aastex}
%\usepackage{emulateapj5}
\usepackage{lscape}
\usepackage{natbib}

\begin{document}

\title{Spitzer, Gaia and the Potential of the Milky Way}

\author{Johnston, Price-Whelan, Madore, Majewski}

\begin{abstract}
sdfsdf
\end{abstract}

\section{Introduction}
\label{intro.sec}
 The existence of vast halos of unseen {\it dark} matter surrounding each galaxy has long been proposed \citep[e.g.,][]{zwicky37,rubin70} to explain the surprisingly large motions of the {\it baryonic} matter (i.e. stars, gas and dust) that we can see.
Dark-matter-only simulations of structure formation lead us to expect that these  halos should have a density distribution that can simply be described with a universal profile \citep{nfw96} with a variety of triaxial shapes \citep{jing02}.
The inclusion of baryons in the simulations tends to soften the triaxiality of the dark matter in the inner regions of the halo \citep[as the disk forms, see e.g.,][]{bailin05}, and can potentially alter the radial profile through a combination of adiabatic contraction and energetic feedback \citep[see recent discussion by][]{pontzen12}.
Hence, measurements of the shape, orientation, radial profile and extent of dark matter halos can shed light both on the formation of these vast structures and the role that the messy baryonic processes of gas inflow, star formation and feedback  have played in shaping them.

The  extent of the bulk of the visible matter in galaxies is roughly 10\% of that of the host dark matter halo. Hence the major components of a galaxy can be used to  trace the inner portion of its host halo's mass distribution.
The rotation curves of disk galaxies are particularly sensitive probes since matter in disks can be assumed to move on near circular orbits. 
This knowledge gives us extra information about the orbital distribution of the visible matter within the dark matter halo, which allows a more precise tracing of the potential than is feasible for objects such as elliptical galaxies where orbits are assumed to be random.
For example, rotation curves of low-surface-brightness galaxies can be used to examine whether dark matter halos have a cusped, power-law density profile in their centers as predicted by purely dark matter simulations of structure formation, or a constant density core perhaps induced by the influence of baryonic processes \citep[e.g.][]{kuzio09}.

% Quote recent Binney/McMillan and Bovy/Rix work here? In longer paper…. Really about large disk samples and disk df's?
% McMillan & Binney Tori work - but again assume fully populated DF

Around external galaxies, the  dark matter distribution  at larger radii has been explored using a variety of approaches \citep[see][for a a complete and detailed review]{courteau12}.
For example, the kinematics of tracers populations such as globular clusters or planetary nebulae  can be used to derive mass estimates under the assumptions that these satellite systems are relaxed and well-mixed \citep[early investigations include][]{cote03,mendez01}. 
The level of detail in these estimates is necessarily limited by the number of objects bright enough for spectroscopic measurements. The largest samples to date have of order several hundred members probing out to several 10's of kpc,  and allowing simple, parameterized models of the mass and orbit distribution to be simultaneously constrained \citep[e.g.][]{napolitano11,deason12a}. 
%\citep{romanowsky03,douglas07,romanowsky09,napolitano11,lee11}
Alternatively, the properties of gravitationally lensed background sources around a galaxy can be used to map the shape, orientation and radial profile of the projected mass distribution \citep[as done by the Lens Structure and Dynamics Survey described in][]{koopmans02}. 
The level of detail in these two-dimensional reconstructions is impressive:
recent work has even found some systems sensitive to the position and mass of a major subhalo orbiting with the main halo \citep[e.g.][]{vegetti10}.
Of course, lensing reconstructions can only be performed for galaxies which closely intersect our line of sight to background sources, but the advent of large photometric catalogues has allowed automatic searches for such chance alignments and significant increases in the number of objects studied \citep[e.g. the Sloan Lens ACS Survey, see][]{bolton06}.

The one place in the Universe where we might hope to dissect a dark matter halo in even more detail is locally for our own Milky Way Galaxy. 
Our proximity means that we can use individual stars as tracer populations and hence build much larger samples that probe deeper into the halo than the globular cluster and planetary nebula studies of external galaxies.
Moreover, our internal perspective allows us to build a three-dimensional picture of the mass distribution rather than measuring its two-dimensional projection, as is the case for gravitational lensing.
Hence we can hope to reconstruct not only the density, but also the shape and orientation of the dark matter halo as function of radius well beyond the disk.
For example, to date thousands of blue-horizontal-branch stars selected from the Sloan Digital Sky Survey \citep[SDSS][]{} have been used to probe mass out to tens of kpc \citep{deason12,kafle12}, and estimates with combined tracers extend to 150kpc \citep{deason12b}. 
\citet{loebman12} demonstrates how even larger stellar samples from current (i.e. SDSS) and future  (Gaia and LSST) stellar surveys will also be sensitive to any asphericity of the halo.

There is a caveat to this work using tracer populations: these studies all rely on the assumptions that the tracers  represent a random sampling of orbits drawn from a smooth distribution function and are well-mixed in orbital phase. 
However, the same surveys that provide large samples of tracers have also revealed the existence of large-scale spatial inhomogeneities in the form of giant stellar streams and clouds\citep{newberg01,majewski03,belokurov05}, clearly demonstrating that  a significant fraction of the stellar halo is neither randomly sampling a smooth orbit distribution nor is fully phase-mixed. 
The degree of substructure increases with Galactocentric radius \citep{bell08} in agreement with models where the stellar halo is formed purely from accretion events \citep{bullock05,cooper10,rashkov12}.
Moreover, even in the smoother, inner halo which is likely to be phase-mixed, significant clumpings in velocity \citep{schlauffman10} and orbital properties \citep{helmi99} have been observed, which indicate a non-smooth component to the orbit distribution.
By ``observing'' particles in a sample of simulated stellar halos, \citet{yencho06} showed that, if stellar halos are indeed entirely formed via accretion events, then the expected degree of non-randomness  could lead to systematic biases in mass estimates of order several tens of percent.
This encouraging result suggests that, if our expectations for the typical number and size of accretion events contributing to the stellar halo are correct, then the number of orbits effectively explored is sufficiently large that the assumption of random, fully phase-mixed orbits is reasonable for the current samples.

A complimentary method to measure the mass distribution is to instead take full advantage of the {\it non}-random nature of the stellar distribution in our stellar halo --- analogous to our exploitation of matter in disks moving on near-circular orbits.
While we do not know the exact orbits of  stellar debris from satellite destruction \citep[although][proposes a method of orbit reconstruction from limited dimensions of observations]{eyre08}, we do know something more about stars in a debris structure than in a random sample: we know that these stars were once all part of the same object.
This approach is very promising --- because tidal streams are dynamically cold systems, it requires orders of magnitude fewer tracers than a random sample to get constraints of comparable accuracy.
In the simplest case, we might {\it assume} that debris stars are actually still on the same orbit as their progenitor system.
Then we can imagine that measuring the full-space velocities ${\bf v}$ at different points ${\bf x}$ in the structure (e.g. along a stream) would actually give us a direct measure of differences in potential $\Phi$ (i.e. $\Phi({\bf x}_1)-\Phi({\bf x}_2)=(v_2^2-v_1^2)/2$).
In reality, we usually know at most four of the six phase-space co-ordinates for debris stars, often with significant error bars and potential measurements are made by fitting orbits to streams \citep[e.g.,][]{helmi05,johnston05,koposov10,law10,lux12}.
Moreover, our assumption of debris tracing a single orbit is actually incorrect \citep[see][for discussions of the orbit distribution in tidal debris]{johnston98,helmi99}, with systematic changes in orbital energy along debris streams that can lead to systematic biases in our measurements of the Galactic potential \citep{eyre09,varghese11}.

One way to address these systematic biases is to run fully self-consistent N-body simulations of satellite destruction in a variety of potentials (which naturally generate a distribution of orbital properties in the debris and along tidal streams) with the aim of simultaneously constraining both the properties of the satellite and the Milky Way. 
Many studies of the Sagittarius debris system (hereafter Sgr)  --- whose tidal streams entirely encircle the Milky Way --- have adopted this approach \citep[e.g.][]{law05,fellhauer06} with the most recent work for the first time attempting to place constraints on the triaxiality and orientation of the Milky Way's matter distribution \citep{law11}.
Studies which first fit the Sgr stream to a single orbit \citep{johnston05,law09} and were subsequently followed up with full N-body simulations \citep{law05,law12} offer another demonstration of the limitation of the former approach as potential parameters are revised when moving to the N-body approach.

Other approaches have been proposed that avoid the cost of full N-body simulations yet recognize the limitations of single orbit integration.
For example, using the results of N-body simulations as a guide, simple corrections around a single orbit can be adopted to make model predictions that more closely mimic expected debris behavior \citep[e.g.][]{johnston99a,varghese2011}.

The promise of near-future data sets including full (or nearly full) phase-space information has also inspired discussions that move beyond fitting observables along a stream. 
\citet{binney08} and \citet{penarrubia12} demonstrate conceptually that the distribution of energy and entropy (respectively) in debris will be minimized only for a correct assumption of the form of the Galactic potential (although neither tests their method with realistic observational errors???).
\citet{johnston99} (hereafter JZSH) showed that backwards integration of the positions of stars and their parent satellite could be used to distinguish Milky Way mass models: only in the correct potential would the stars' and satellite's  path coincide within a Hubble time.
The JZSH method was designed and tested for the proposed characteristics of the Space Interferometry Mission \citep{}, which was expected to be able to measure proper motions with $\mu$as/year accuracy for stars as faint as $V=20$.
Distances to stream members would only be poorly known and were estimated using anticipated properties of tidal debris.

In this paper we re-examine and update  the JZSH  in the context of current and near-future observational capabilities, and illustrates its power with the example of Sgr.
In Section \ref{context.sec} we outline the observational prospects and Sgr properties that motivated this re-examination.
In Section \ref{measure.sec} we present the updated potential measure and test it with synthetic observations of simulated Sgr debris.
In Section \ref{summary.sec} we summarize our results and outline prospects for combining constraints from Sgr with those from other debris systems.

\section{Context for the study}

The method  in Section \ref{measure.sec} takes advantage of three distinct developments in the observations of tidal streams: 
(i) the demonstration of a technique for deriving distances to individual RR Lyrae stars with 2\% accuracies (see Section \ref{spitzer.sec}); 
(ii) the prospect of proper motion measurements of the same stars with $\sim$10 mas/year  precision (see Section  \ref{gaia.sec});
and (iii) the tracing of debris associated with Sgr  entirely around the Galaxy (see Section \ref{sgr.sec})

\subsection{Spitzer RR Lyraes}

\label{spitzer.sec}

%WISE: benedict11? 'precision of spritzer will far exceed these' also 'serendipi…
%Text 'stolen' from proposal 1: the Carnegie RR Lyrae Program

Over twenty-five years ago Longmore, Fernley & Jameson (1986, MNRAS, 220, 279) demonstrated the existence of a near-infrared period-luminosity (PL) relation for RR Lyrae stars, using K-band observations of these stars in Galactic globular clusters. Catelan, Pritzl & Smith (2004, ApJS, 154, 633) showed that a PL relation for RR Lyrae stars is to be “expected” in the infrared (but vanishingly so in the optical), explained primarily by the changing bolometric correction for these stars as a function of the observed wavelength. Later, Madore & Freedman (2011, ApJ, 744, 132) demonstrated the inevitability of decreasing intrinsic scatter with increasing wavelength (and slope) of PL relations both for Cepheids and for RR Lyrae stars.
Moreover, both reddening effects and metallicity dependencies also decrease significantly for observations further into the IR.

Taken together, these studies suggest observations of RR Lyrae variables by NASA's Spitzer mission could provide us with a stellar distance indicator of unprecedented accuracy and precision. 
\citet{madore13} have demonstrated \citep[using 5 stars with trigonometric parallaxes measured by Hubble][]{benedict11} that the dispersion in the mid-IR PL relation at Spitzer wavelengths, is on the order of ±0.03 mag. The immediate implication is that it is now possible to determine individual distances that are good to better than ±2\% for single RR Lyrae stars, out to the limits of Spitzer's ability to detect and measure RR Lyrae variables (which is at least 80 kpc, for modest integration times).

\subsection{Gaia and Spitzer}
\label{gaia.sec}

ESA's Gaia satellite, planned for launch in June 2013, is an astrometric mission which aims to measure the positions of billions of stars with 10-100$\mu$as accuracies (depending on the color and magnitude of each star), scanning the entire sky multiple times over its 5-year lifetime \citep{perryman01}.
Gaia will revolutionize the field of Galactic Astronomy with sufficient accuracy to create full, 6-dimensional phase-space maps for distances of up to $\sim 10$kpc from the Sun \citep[see][for discussion]{}.

The left-hand panel of Figure \ref{data.fig} show Gaia's end-of-mission anticipated errors on the distance to  of an individual RR Lyrae star at different distances $d$ from the Sun.
\footnote{Taken from WWW site?.}
Note that within 2kpc of the Sun, Gaia will be able to measure distances to these stars to better than 2\%: RR Lyraes in this volume will be used to test and calibrate the Spitzer PL relation. 
Beyond 2kpc, the IR PL relation for RR Lyraes will provide more accurate distance estimates than Gaia.

The remaining panels of Figure \ref{data.fig} demonstrate that Gaia's anticipated proper motions errors for RR Lyraes (middle) are sufficiently accurate to distinguish tangential motions with better than $\sim$10 km/s precision out to many tens of kpc (right-panel).
Hence, the combination of Spitzer and Gaia data offers the exciting prospect of extending the ``horizon'' of where full six-dimensional phase-space maps of our Galaxy are possible from 10kpc to 80kpc.

\begin{figure}
\plotone{figs/data.eps}
\caption{Summary of data sensitivity: $\Delta d$, $\Delta \mu$ and $\Delta v_{\rm tan}$ as function of $d$.
{\it Also put Sgr and Orphan characteristics on this plot?}
\label{data.fig}
}
\end{figure}

\subsection{The Sagittarius debris system}

The Sagittarius Dwarf Galaxy (hereafter Sgr) was discovered
serendipitously during a radial velocity survey of the Galactic bulge
(Ibata et al. 1994, Nature, 370, 194). While the number of known MW
satellites has more than doubled recently (e.g.,
McConnachie, 2012, AJ, 144, 4), in many ways Sgr remains the most
studied and most intriguing --- it is the closest satellite, among the
largest (only the LMC and SMC are more luminous), and has the most
prominent tidal tails.
From the first maps of the Sgr core (Ibata et al. 1994; Ibata et
al. 1995, MNRAS, 277, 781) its elongated morphology suggested
destruction by tidal forces, as might be anticipated
from its proximity to the Galaxy.  Subsequent N-body models confirmed
the validity of this interpretation (Velazquez \& White, 1995, MNRAS,
275, L23) and showed that debris from Sgr should
form coherent streams of stars encircling our Galaxy (Johnston et
al. 1995, ApJ, 451, 598). Signatures of these giant star streams have
since been discovered in many studies, and comprehensively mapped
across the sky in carbon stars (Totten \& Irwin, 1998, MNRAS, 294, 1)
M giant stars selected from the Two Micron All Sky Survey (Majewski et
al.  2003, ApJ, 599, 1082) and main sequence turnoff stars selected
from the Sloan Digital Sky Survey (Belokurov et al. 2006, ApJ Lett, 642, L137). 
%\citep{belokurov06}.
%Distance estimates along the tail have been derived from these photometric surveys (e.g., Mart{\'{\i}}nez-Delgado et al. 2004, ApJ, 601, 241) and gradients in radial velocities and stellar populations measured with follow-up spectroscopy (e.g., Keller et al. 2010, ApJ, 720, 940).
%\citep{majewski04,martinez04,bellazzini06,chou07,chou10,keller10,carlin12}.
%These rich data sets have inspired equally rich interpretations of Sgr debris
% (e.g.,  Johnston et al. 1999, AJ, 118, 1719; Helmi, 2004, ApJL, 610, L97; Law et al. 2005, ApJ, 619, 807; Fellhauer et al. 2006, ApJ, 651, 167; Law \& Majewski, 2010, ApJ, 714, 229).
%(e.g.,  Law \& Majewski, 2010, ApJ, 714, 229 -- hereafter LM10).

Figure
\ref{sgr.fig}  illustrates results from LM10's N-body
simulation of dwarf satellite disruption along the expected Sgr orbit, with particles projected onto the plane perpendicular to the
Galactic disk containing both the Sun and Galactic center.
The simulation was run in a potential…..

LM10 demonstrate how a comparison of such simulations with the
observed debris enables a detailed reconstruction of Sgr's history ---
it's mass and stellar content, orbit, rate of destruction and even
original distribution in populations.  
(See also other refs ????)
Moreover, the phase-space
distribution of the debris offers a unique probe of the depth, shape
and extent of the Milky Way's dark matter halo that cannot be matched
using other techniques on our own or other galaxies (e.g., Ibata et
al. 2001, ApJ, 551, 294).  Most recently, LM10 have combined all
current data on Sgr debris to assess the triaxiality and orientation
of the outer Galaxy --- the first time that such a reconstruction of
the 3-dimensional mass distribution of a dark matter halo has been
feasible.  These constraints on Sgr's mass and orbit have
invigorated discussions of Sgr in a more cosmological context, the
mass and extent of the original dark matter halo that hosted it as well
as its effect on the MW itself (e.g., Michel-Dansac et
al. 2011, MNRAS, 414, L1).
%\citep{bailin03,purcell11,michel11,gomez12}.

The promise of the combined Spitzer and Gaia measurements of the distances to and proper motions of RR Lyraes in the Sgr debris opens up new approaches to interpretation.
With 2\% distances measurements, the thickness of Sgr's stream is actually greater
than the uncertainty.  Moreover,
with Gaia's final data release, there will be proper motion
measurements of accuracies less than the intrinsic velocity dispersion for much
of the stream (e.g., corresponding to a tangential velocity accuracy
of 10 km/s for an RR Lyrae at 35kpc) and ground-based radial
velocities of few km/s accuracies can easily be added. 
Hence, for the first time, the measured phase-space position of each individual RR Lyrae
relative to the stream is actually physically meaningful.
The next section outlines a method to take full advantage of this accuracy and use each RR Lyrae individually as a potential measure.

\begin{figure}
\plotone{figs/sgr.eps}
\caption{
\label{sgr.fig}
}
\end{figure}

\section{Proposed potential measure}

\subsection{Method}

Once the full
phase-space positions of these stars are known there is no need to
build a sophisticated model of the entire stream system to interpret
the data.  Instead, the knowledge that these stars were once part of
Sgr allows them each to be exploited as individual potential measurers.
If the orbits of an RR Lyrae and Sgr are simultaneously integrated
backwards in a potential that correctly models the MW, then
their orbits must intersect at some point within the last few Gyrs; in an
incorrect potential this will not happen.
This approach was first described by JZSH in anticipation of the
SIM survey.
The original algorithm does not assume that the distances to the debris stars are known, 
and instead estimates them using the expected properties for the debris.
A Spitzer survey of RR Lyraes would allow us to drop this approximate approach in favor of using the  measured distances.

At any point during the backwards integration, the proximity of a debris particle to the parent satellite is defined by the normalized phase-space distance, 
\begin{equation}
       D_{\rm ps}=\left[\left({d \over r_{\rm tide}}\right)^2+\left({v \over v_{\rm esc}}\right)^2\right],
\end{equation}
where $d$ and $v$ are the distance and speed separation between the star
and the parent respectively.  The one assumption in this method is what value to adopt for the satellite mass, $m_{\rm \sat}$,which sets the instantaneous tidal radius, $r_{\rm tide}=                                                                      
R (m_{\rm sat}/M_{\rm Gal})^{1/3}$ and escape velocity, $v_{\rm esc} =                                                               
G m_{\rm sat}/r_{\rm tide}$. The mass of the
Galaxy $M_{\rm Gal}$ enclosed within radius $R$ is calculated directly from the
assumed potential which is being tested.  The ``best'' potential parameters are those that
minimize the average $D_{\rm ps}$ for all stars.
{\it Or we could come up with some other statistic.}

\subsection{Test of method on synthetic Sgr data}

Figure \ref{boots.fig} illustrates potential power of the method for
particles from the Sgr disruption N-body simulation
 chosen to represent our proposed
RR Lyrae survey (yellow crosses in left panel).  The 53 particles within 35 kpc of
the Sun were ``observed'' with appropriate distance (assuming 2\%
accuracy), proper motion (assuming Gaia accuracies)\footnote{see
 http://www.rssd.esa.int/index.php?project=GAIA} and radial velocity
(assuming 5 km/s) errors.  The ``observed'' phase-space positions were
then integrated backwards in model potentials with axis ratios
($q_1/q_2$ and $q_z/q_2$) and orientation ($\phi$) of the dark matter
component systematically varied.  Sgr's phase space coordinates
%(whose position and motion were                                                                                                                         
(assumed to be exactly known, along with the distance to the Galactic
center) were simultaneously integrated backwards in the same
potentials, and the minimum normalized phase-space distance, $D_{\rm                                                              
 ps}$, between Sgr and each particle recorded.
he ``best'' potential parameters were those that
minimized the average $D_{\rm ps}$ for all stars.

Figure
\ref{boots.fig} contours confidence levels for the recovered parameters
(with a bold cross marking the true values used in the original
simulation), with levels estimated by repeating the analysis using
samples bootstrapped from our ``observations''.  The contours
demonstrate the power of this combination of Spitzer and Gaia data:
This small sample alone can broadly rule out significant areas of
parameter space.  Combining the constraints from these samples with
the extensive ground-based data sets already available would allow
quite clear measurements of the shape and strength of the Milky Way
potential.

\begin{figure}
\plotone{figs/boots.eps}
\caption{
\label{boots.fig}
}
\end{figure}

\subsection{Future prospects}

The great strength of the  JZSH ({\it can we come up with a catchy name?}) method is its simplicity:
it requires only the assumption a satellite mass $m_{\rm sat}$ and  backwards integration of orbits. 
There are no assumptions made about or N-body modeling required to follow the internal distribution of satellite stars or accurately assess the degree of alignment of debris with a single orbit.
In fact, the method relies on the ability to measure {\it differences} in orbital properties --- these differences result in different orbital time periods for the debris and the satellite which leads to their paths intersecting during the backwards integration.

There are two other apparent strengths of the method which require further investigation.
First it appears applicable to any debris that is known to come from a single object, and not restricted only to the very coldest tidal streams. For example, in principle the method could  applied to the vast debris {\it clouds} that have been discovered subtending hundreds-thousands square degrees \citep[e.g. the Triangulum Andromeda and Hercules-Aquila clouds]{rochapinto04,belokurov06}, or even stars that have only associations in orbital properties and do not form a coherent spatial structure \citep[such as the angular momentum groupings in local giants found by][]{helmi99}.
Second, it appears trivial to combine constraints from multiple debris systems at once by simply integrating all debris from several satellites simultaneously with $D_{\rm ps}$ defined appropriately for each star separately.

The limitations of the method also need to be understood. In particular, for the RR Lyraes associated with  the very coldest streams \citep[e.g. the globular clusters Pal5 and GD1][]{odenkirchen02,koposov10} the measurement errors are unlikely to be small enough for the 
the differences in orbital properties to between debris and satellite to be detected, and the method should fail.

Lastly, the current version of the algorithm relies on accurate knowledge of the phase-space position of the parent satellite, which may not be available in some cases \citep[for example, for the Orphan Stream][]{belokurov07}. Given sufficient data, position and motion of the original host could be solved for as additional free parameters, but the effectiveness of this approach has yet to be investigated.

\section{Summary}

This paper outlines an algorithm to measure the Galactic potential that takes maximal advantage of a combination of possible data from the Spitzer and Gaia satellite missions that, when combined,  promise full phase-space measurements of RR Lyrae stars around our Galaxy.
When applied to a synthetic data set generated from debris in an N-body simulation of the destruction of Sgr the method recovers the shape and orientation of the dark matter halo component of the potential with ??? accuracy using a sample of ??? stars.
This success provides strong motivation for: (i) further theoretical work to investigate the power of multiple debris structures with different morphologies and orientations in building a comprehensive picture of the Galactic mass distribution; and (ii) a  Spitzer survey of RR Lyrae stars in debris structures around the Milky Way as a ``legacy'' sample to combine with near-future Gaia data.

\bibliography{spitzer}
\bibliographystyle{apj}

\end{document}

### 2013-04-26 scripts/combine_catalogs.py
cleaned up script for plotting multiply rrlyr catalogs

@@ -19,13 +19,13 @@ __author__ = "adrn <adrn@astro.columbia.edu>"
 
 # Standard library
 import os, sys
+import logging
 
 # Third-party
 import numpy as np
 from matplotlib import cm
 import matplotlib.pyplot as plt
-import matplotlib.colors as col
-from scipy.stats import gaussian_kde
+from astropy.io import ascii
 import astropy.coordinates as coord
 from astropy.table import Table, Column
 import astropy.units as u
@@ -33,59 +33,188 @@ import astropy.units as u
 from streams.data import *
 from streams.coordinates import SgrCoordinates, OrphanCoordinates
 from streams.data.gaia import rr_lyrae_photometric_distance
+from streams.plot import discrete_cmap
 
-def discrete_cmap(N=8):
-    """create a colormap with N (N<15) discrete colors and register it"""
-    # define individual colors as hex values
-    cpool = ['#9E0142', '#D53E4F', '#F46D43', '#FDAE61', '#FEE08B',
-             '#E6F598', '#ABDDA4', '#66C2A5', '#3288BD', '#5E4FA2']
-    if N == 5:
-        cmap3 = col.ListedColormap(cpool[::2], 'nice_spectral')
-    else:
-        cmap3 = col.ListedColormap(cpool[0:N], 'nice_spectral')
-    cm.register_cmap(cmap=cmap3)
+# Create logger
+logger = logging.getLogger(__name__)
 
 discrete_cmap(5)
-css = read_catalina()
-quest = read_quest()
-asas = read_asas()
-nsvs = read_nsvs()
-sdss = read_stripe82()
-
-css.keep_columns(["ra","dec","V"])
-css.add_column(Column(1*np.ones(len(css)), name="survey"))
-css = np.array(css["ra","dec","V","survey"])
-
-quest.keep_columns(["ra","dec","V"])
-quest.add_column(Column(2*np.ones(len(quest)), name="survey"))
-quest = np.array(quest["ra","dec","V","survey"])
-
-asas.keep_columns(["ra","dec","V"])
-asas.add_column(Column(3*np.ones(len(asas)), name="survey"))
-asas = np.array(asas["ra","dec","V","survey"])
-
-nsvs.keep_columns(["ra","dec","V"])
-nsvs.add_column(Column(4*np.ones(len(nsvs)), name="survey"))
[…]

### 2013-04-26 streams/plot/core.py
added discrete colormap to general plotting

@@ -13,13 +13,27 @@ import cPickle as pickle
 # Third-party
 import numpy as np
 import matplotlib.pyplot as plt
+import matplotlib.colors as col
+import matplotlib.cm as cm
 import matplotlib.gridspec as gridspec
 import astropy.units as u
 
 from ..potential.lm10 import param_to_latex, param_ranges
 from ..potential.lm10 import halo_params as true_halo_params
 
-__all__ = ["emcee_plot", "plot_sampler_pickle"]
+__all__ = ["discrete_cmap", "emcee_plot", "plot_sampler_pickle"]
+
+def discrete_cmap(N=8):
+    """create a colormap with N (N<15) discrete colors and register it"""
+    # define individual colors as hex values
+    # #9E0142
+    cpool = ['#D085A4', '#D53E4F', '#F46D43', '#FDAE61', '#FEE08B',
+             '#E6F598', '#ABDDA4', '#66C2A5', '#3288BD', '#5E4FA2']
+    if N == 5:
+        cmap3 = col.ListedColormap(cpool[::2], 'nice_spectral')
+    else:
+        cmap3 = col.ListedColormap(cpool[0:N], 'nice_spectral')
+    cm.register_cmap(cmap=cmap3)
 
 # TODO: make this more general...
 def emcee_plot(sampler, params, converged_idx, 

### 2013-04-26 docs/mdm_proposal_may2013.tex
added may mdm proposal

\documentclass[preprint]{aastex}

\begin{document}

Galaxies, by mass, are mostly composed of \emph{Dark Matter}
\citep{}. Simulations of structure and galaxy formation -- both with
and without baryons -- lead us to believe that such \emph{halos} of
dark matter have universal density profiles \citep{} and are triaxial
in shape \citep{}. Including baryons in such simulations tends to
soften the triaxiality and alters the radial density profile. Thus,
precise measurements of the shape, orientation, and radial profile of
dark matter halos provides information about both dark matter physics
and the baryonic processes that helped shape them. Tidal streams from
infalling satellite galaxies have been used to constrain properties of
the Milky Way's halo \citep{}; typically found outwards of 20~kpc, the
orbits of associated stars are still largely under the influence of
the dark matter and thus serve as potentiometers \citep{}. Using a new
method that utilizes full phase-space (6D) kinematic information for a
small sample ($\sim$100s) of stars stripped from a known progenitor
(e.g., Sagittarius; Johnston\&Price-Whelan 2013, in prep), it is
possible to accurately constrain the axis ratios, total mass, and
orientation of the Milky Way's dark matter halo. We plan to use a
sample of RR Lyrae stars selected from catalogs of RRab stars from
time-domain, photometric surveys such as LINEAR \citep{}, PTF \citep{}
and CSS \citep{}. This method relies on a technique for deriving
distances to individual RR Lyrae stars with 2\% accuracy \citep{},
along with anticipated GAIA proper motion errors of
$\sim10-20~\mathrm{mas}/\mathrm{yr}$ for such stars at $\sim$
20-30~kpc. This leaves one remaining velocity component -- the radial
velocity -- which we propose to measure using the 2.4m Hiltner
telescope at MDM, equipped with the [Allyson ??].

\end{document}

### 2013-04-26 docs/mdm_proposal_may2013.tex
added kvj and allyson changes

@@
 \begin{document}
-
-Galaxies, by mass, are mostly composed of \emph{Dark Matter}
-\citep{}. Simulations of structure and galaxy formation -- both with
+\title{Probing the Milky Way's Dark Matter Halo with RR Lyraes in Tidal Streams I: The Sagittarius Leading Arm}
+\author{Adrian Price-Whelan\altaffilmark{1}, Kathryn Johnston\altaffilmark{1}, Allyson Sheffield\altaffilmark{1}}
+\altaffiltext{1}{Department of Astronomy, Columbia University, Mail Code 5246, New York, NY 10027}
+Galaxies, by mass, are mostly composed of \emph{Dark Matter}. Simulations of structure and galaxy formation -- both with
 and without baryons -- lead us to believe that such \emph{halos} of
-dark matter have universal density profiles \citep{} and are triaxial
-in shape \citep{}. Including baryons in such simulations tends to
+dark matter have universal density profiles \citep{nfw96} and are triaxial
+in shape \citep{jing02}. Including baryons in such simulations tends to
 soften the triaxiality and alters the radial density profile. Thus,
@@
 infalling satellite galaxies have been used to constrain properties of
-the Milky Way's halo \citep{}; typically found outwards of 20~kpc, the
+the Milky Way's halo \citep[e.g.][]{law10}; typically found outwards of 20~kpc, the
 orbits of associated stars are still largely under the influence of
-the dark matter and thus serve as potentiometers \citep{}. Using a new
+the dark matter and thus serve as potentiometers \citep[e.g.][]{johnston99}. Using an updated
 method that utilizes full phase-space (6D) kinematic information for a
 small sample ($\sim$100s) of stars stripped from a known progenitor
-(e.g., Sagittarius; Johnston\&Price-Whelan 2013, in prep), it is
+(e.g., Sagittarius; Price-Whelan et al 2013, in prep), it is
 possible to accurately constrain the axis ratios, total mass, and
-orientation of the Milky Way's dark matter halo. We plan to use a
-sample of RR Lyrae stars selected from catalogs of RRab stars from
-time-domain, photometric surveys such as LINEAR \citep{}, PTF \citep{}
-and CSS \citep{}. This method relies on a technique for deriving
+orientation of the Milky Way's dark matter halo. This method relies on a technique for deriving
 distances to individual RR Lyrae stars with 2\% accuracy \citep{},
 along with anticipated GAIA proper motion errors of
-$\sim10-20~\mathrm{mas}/\mathrm{yr}$ for such stars at $\sim$
+$\sim10-20~\mu\mathrm{as}/\mathrm{yr}$ for such stars at $\sim$
 20-30~kpc. This leaves one remaining velocity component -- the radial
-velocity -- which we propose to measure using the 2.4m Hiltner
-telescope at MDM, equipped with the [Allyson ??].
+velocity.  We plan to observe a
+sample of RR Lyrae stars selected to be likley members of Sgr's debris system 
+using catalogs of RRab stars from
+time-domain, photometric surveys such as LINEAR \citep{}, PTF \citep{}
+and CSS \citep{}. 
+Our observations will both confirm membership of the debris system in order to construct a target list for Spitzer (proposals are due in August 2013) as well
+as measure the missing velocity dimension.
+
+We propose to measure radial velocities of RR Lyraes in the Sgr leading arm using ModSpec on the 2.4m Hiltner
+telescope at MDM.
+Using the 600 groove mm$^{-1}$ grating blazed at 5000 \AA\ and a 2048x2048 detector with 15 micron pixels (Wilbur), we will have wavelength coverage of about 1000 \AA\, including the H$\beta$ line at 4861 \AA\ and the [OI] sky line at 5577 \AA\ (the latter to aid in the radial velocity determination by identifying any systematic offsets).
+We will collect bias and flat field frames at the beginning and end of the night, and ThAr arc lamp spectra for wavelength calibration will be taken throughout the night to account for telescope flexure.
+Radial velocity standards will be selected that are of a similar spectral type as the targets (e.g., HD122693  -- a dwarf of spectral type F8).
+Typical equatorial coordinates for the targets are ($\alpha,\delta$)=(13$^{h}-16^{h}$,-5$^{\circ}- +25^{\circ}$); stars in this range will be at airmass $<2$ all night.
+The typical $R$-band magnitude is 16.
+To achieve S/N of 30 per pixel with the proposed set-up, an integration time of 1200 seconds is needed for a target with $R$=16.
+Considering the total amount of observing time of 9 hours per night and the overhead needed for calibration frames, we anticipate observing approximately 40 RR Lyraes.
+Reductions will be carried out by Sheffield using a well-developed reduction pipeline for processing ModSpec data.
+
+\begin{thebibliography}[99]
+\bibitem[Jing 
+\& Suto(2002)]{jing02} Jing, Y.~P., \& Suto, Y.\ 2002, \apj, 574, 538 
+\bibitem[Navarro et al.(1996)]{nfw96} Navarro, J.~F., Frenk, 
+C.~S., \& White, S.~D.~M.\ 1996, \apj, 462, 563 
+\end{thebibliography}
 

### 2013-06-19 docs/paper1/paper.tex
reorganize docs

--- 2013-01-17 docs/paper.v0.tex
@@
-%\documentclass{emulateapj}
+\documentclass{emulateapj}
 %\documentclass[preprint]{aastex}
-\documentclass{aastex}
+%\documentclass{aastex}
 %\usepackage{emulateapj5}
-\usepackage{lscape}
+%\usepackage{lscape}
 \usepackage{natbib}
@@
 \label{intro.sec}
- The existence of vast halos of unseen {\it dark} matter surrounding each galaxy has long been proposed \citep[e.g.,][]{zwicky37,rubin70} to explain the surprisingly large motions of the {\it baryonic} matter (i.e.
-stars, gas and dust) that we can see.
-Dark-matter-only simulations of structure formation lead us to expect that these  halos should have a density distribution that can simply be described with a universal profile \citep{nfw96} with a variety of triaxial shapes \citep{jing02}.
-The inclusion of baryons in the simulations tends to soften the triaxiality of the dark matter in the inner regions of the halo \citep[as the disk forms, see e.g.,][]{bailin05}, and can potentially alter the radial profile through a combination of adiabatic contraction and energetic feedback \citep[see recent discussion by][]{pontzen12}.
-Hence, measurements of the shape, orientation, radial profile and extent of dark matter halos can shed light both on the formation of these vast structures and the role that the messy baryonic processes of gas inflow, star formation and feedback  have played in shaping them.
+ The existence of vast halos of unseen {\it dark} matter surrounding each galaxy has long been proposed \citep[e.g.,][]{zwicky37,rubin70} to explain the surprisingly large 
+motions of the {\it baryonic} matter (i.e. stars, gas and dust) that we can see.
+Dark-matter-only simulations of structure formation lead us to expect that these  halos should have a density distribution that can simply be described with a universal 
+profile \citep{nfw96} with a variety of triaxial shapes \citep{jing02}.
+The inclusion of baryons in the simulations tends to soften the triaxiality of the dark matter in the inner regions of the halo \citep[as the disk forms, see e.g.,][]{bailin05}, and 
+can potentially alter the radial profile through a combination of adiabatic contraction and energetic feedback \citep[see recent discussion by][]{pontzen12}.
+Hence, measurements of the shape, orientation, radial profile and extent of dark matter halos can shed light both on the formation of these vast structures and the role that 
+the messy baryonic processes of gas inflow, star formation and feedback  have played in shaping them.
 
-The  extent of the bulk of the visible matter in galaxies is roughly 10\% of that of the host dark matter halo.
-Hence the major components of a galaxy can be used to  trace the inner portion of its host halo's mass distribution.
+The  extent of the bulk of the visible matter in galaxies is roughly 10\% of that of the host dark matter halo. Hence the major components of a galaxy can be used to  trace 
+the inner portion of its host halo's mass distribution.
 The rotation curves of disk galaxies are particularly sensitive probes since matter in disks can be assumed to move on near circular orbits. 
-This knowledge gives us extra information about the orbital distribution of the visible matter within the dark matter halo, which allows a more precise tracing of the potential than is feasible for objects such as elliptical galaxies where orbits are assumed to be random.
-For example, rotation curves of low-surface-brightness galaxies can be used to examine whether dark matter halos have a cusped, power-law density profile in their centers as predicted by purely dark matter simulations of structure formation, or a constant density core perhaps induced by the influence of baryonic processes \citep[e.g.][]{kuzio09}.
+This knowledge gives us extra information about the orbital distribution of the visible matter within the dark matter halo, which allows a more precise tracing of the 
+potential than is feasible for objects such as elliptical galaxies where orbits are assumed to be random.
+For example, rotation curves of low-surface-brightness galaxies can be used to examine whether dark matter halos have a cusped, power-law density profile in their 
+centers as predicted by purely dark matter simulations of structure formation, or a constant density core perhaps induced by the influence of baryonic processes \citep[e.g.]
+[]{kuzio09}.
 
@@
 
-Around external galaxies, the  dark matter distribution  at larger radii has been explored using a variety of approaches \citep[see][for a a complete and detailed review]{courteau12}.
-For example, the kinematics of tracers populations such as globular clusters or planetary nebulae  can be used to derive mass estimates under the assumptions that these satellite systems are relaxed and well-mixed \citep[early investigations include][]{cote03,mendez01}.
-The level of detail in these estimates is necessarily limited by the number of objects bright enough for spectroscopic measurements.
-The largest samples to date have of order several hundred members probing out to several 10's of kpc,  and allowing simple, parameterized models of the mass and orbit distribution to be simultaneously constrained \citep[e.g.][]{napolitano11,deason12a}.
+Around external galaxies, the  dark matter distribution  at larger radii has been explored using a variety of approaches \citep[see][for a a complete and detailed review]
+{courteau12}.
+For example, the kinematics of tracers populations such as globular clusters or planetary nebulae  can be used to derive mass estimates under the assumptions that 
+these satellite systems are relaxed and well-mixed \citep[early investigations include][]{cote03,mendez01}. 
+The level of detail in these estimates is necessarily limited by the number of objects bright enough for spectroscopic measurements. The largest samples to date have of 
+order several hundred members probing out to several 10's of kpc,  and allowing simple, parameterized models of the mass and orbit distribution to be simultaneously 
+constrained \citep[e.g.][]{napolitano11,deason12a}. 
 %\citep{romanowsky03,douglas07,romanowsky09,napolitano11,lee11}
-Alternatively, the properties of gravitationally lensed background sources around a galaxy can be used to map the shape, orientation and radial profile of the projected mass distribution \citep[as done by the Lens Structure and Dynamics Survey described in][]{koopmans02}.
+Alternatively, the properties of gravitationally lensed background sources around a galaxy can be used to map the shape, orientation and radial profile of the projected 
+mass distribution \citep[as done by the Lens Structure and Dynamics Survey described in][]{koopmans02}. 
 The level of detail in these two-dimensional reconstructions is impressive:
 recent work has even found some systems sensitive to the position and mass of a major subhalo orbiting with the main halo \citep[e.g.][]{vegetti10}.
-Of course, lensing reconstructions can only be performed for galaxies which closely intersect our line of sight to background sources, but the advent of large photometric catalogues has allowed automatic searches for such chance alignments and significant increases in the number of objects studied \citep[e.g.
-the Sloan Lens ACS Survey, see][]{bolton06}.
+Of course, lensing reconstructions can only be performed for galaxies which closely intersect our line of sight to background sources, but the advent of large photometric 
+catalogues has allowed automatic searches for such chance alignments and significant increases in the number of objects studied \citep[e.g. the Sloan Lens ACS Survey, 
+see][]{bolton06}.
 
 The one place in the Universe where we might hope to dissect a dark matter halo in even more detail is locally for our own Milky Way Galaxy. 
-Our proximity means that we can use individual stars as tracer populations and hence build much larger samples that probe deeper into the halo than the globular cluster and planetary nebula studies of external galaxies.
-Moreover, our internal perspective allows us to build a three-dimensional picture of the mass distribution rather than measuring its two-dimensional projection, as is the case for gravitational lensing.
+Our proximity means that we can use individual stars as tracer populations and hence build much larger samples that probe deeper into the halo than the globular cluster 
+and planetary nebula studies of external galaxies.
+Moreover, our internal perspective allows us to build a three-dimensional picture of the mass distribution rather than measuring its two-dimensional projection, as is the 
+case for gravitational lensing.
 Hence we can hope to reconstruct not only the density, but also the shape and orientation of the dark matter halo as function of radius well beyond the disk.
-For example, to date thousands of blue-horizontal-branch stars selected from the Sloan Digital Sky Survey \citep[SDSS][]{} have been used to probe mass out to tens of kpc \citep{deason12,kafle12}, and estimates with combined tracers extend to 150kpc \citep{deason12b}.
-\citet{loebman12} demonstrates how even larger stellar samples from current (i.e. SDSS) and future  (Gaia and LSST) stellar surveys will also be sensitive to any asphericity of the halo.
+For example, to date thousands of blue-horizontal-branch stars selected from the Sloan Digital Sky Survey \citep[SDSS][]{} have been used to probe mass out to tens of 
+kpc \citep{deason12,kafle12}, and estimates with combined tracers extend to 150kpc \citep{deason12b}. 
+\citet{loebman12} demonstrates how even larger stellar samples from current (i.e. SDSS) and future  (Gaia and LSST) stellar surveys will also be sensitive to any 
+asphericity of the halo.
 
-There is a caveat to this work using tracer populations: these studies all rely on the assumptions that the tracers  represent a random sampling of orbits drawn from a smooth distribution function and are well-mixed in orbital phase.
-However, the same surveys that provide large samples of tracers have also revealed the existence of large-scale spatial inhomogeneities in the form of giant stellar streams and clouds\citep{newberg01,majewski03,belokurov05}, clearly demonstrating that  a significant fraction of the stellar halo is neither randomly sampling a smooth orbit distribution nor is fully phase-mixed.
-The degree of substructure increases with Galactocentric radius \citep{bell08} in agreement with models where the stellar halo is formed purely from accretion events \citep{bullock05,cooper10,rashkov12}.
-Moreover, even in the smoother, inner halo which is likely to be phase-mixed, significant clumpings in velocity \citep{schlauffman10} and orbital properties \citep{helmi99} have been observed, which indicate a non-smooth component to the orbit distribution.
-By ``observing'' particles in a sample of simulated stellar halos, \citet{yencho06} showed that, if stellar halos are indeed entirely formed via accretion events, then the expected degree of non-randomness  could lead to systematic biases in mass estimates of order several tens of percent.
-This encouraging result suggests that, if our expectations for the typical number and size of accretion events contributing to the stellar halo are correct, then the number of orbits effectively explored is sufficiently large that the assumption of random, fully phase-mixed orbits is reasonable for the current samples.
+There is a caveat to this work using tracer populations: these studies all rely on the assumptions that the tracers  represent a random sampling of orbits drawn from a 
+smooth distribution function and are well-mixed in orbital phase. 
+However, the same surveys that provide large samples of tracers have also revealed the existence of large-scale spatial inhomogeneities in the form of giant stellar 
+streams and clouds\citep{newberg01,majewski03,belokurov05}, clearly demonstrating that  a significant fraction of the stellar halo is neither randomly sampling a smooth 
+orbit distribution nor is fully phase-mixed. 
+The degree of substructure increases with Galactocentric radius \citep{bell08} in agreement with models where the stellar halo is formed purely from accretion events 
+\citep{bullock05,cooper10,rashkov12}.
+Moreover, even in the smoother, inner halo which is likely to be phase-mixed, significant clumpings in velocity \citep{schlauffman10} and orbital properties \citep{helmi99} 
+have been observed, which indicate a non-smooth component to the orbit distribution.
+By ``observing'' particles in a sample of simulated stellar halos, \citet{yencho06} showed that, if stellar halos are indeed entirely formed via accretion events, then the 
+expected degree of non-randomness  could lead to systematic biases in mass estimates of order several tens of percent.
+This encouraging result suggests that, if our expectations for the typical number and size of accretion events contributing to the stellar halo are correct, then the number of 
+orbits effectively explored is sufficiently large that the assumption of random, fully phase-mixed orbits is reasonable for the current samples.
 
-A complimentary method to measure the mass distribution is to instead take full advantage of the {\it non}-random nature of the stellar distribution in our stellar halo --- analogous to our exploitation of matter in disks moving on near-circular orbits.
-While we do not know the exact orbits of  stellar debris from satellite destruction \citep[although][proposes a method of orbit reconstruction from limited dimensions of observations]{eyre08}, we do know something more about stars in a debris structure than in a random sample: we know that these stars were once all part of the same object.
-This approach is very promising --- because tidal streams are dynamically cold systems, it requires orders of magnitude fewer tracers than a random sample to get constraints of comparable accuracy.
+A complimentary method to measure the mass distribution is to instead take full advantage of the {\it non}-random nature of the stellar distribution in our stellar halo --- 
+analogous to our exploitation of matter in disks moving on near-circular orbits.
+While we do not know the exact orbits of  stellar debris from satellite destruction \citep[although][proposes a method of orbit reconstruction from limited dimensions of 
+observations]{eyre08}, we do know something more about stars in a debris structure than in a random sample: we know that these stars were once all part of the same 
+object.
+This approach is very promising --- because tidal streams are dynamically cold systems, it requires orders of magnitude fewer tracers than a random sample to get 
+constraints of comparable accuracy.
 In the simplest case, we might {\it assume} that debris stars are actually still on the same orbit as their progenitor system.
-Then we can imagine that measuring the full-space velocities ${\bf v}$ at different points ${\bf x}$ in the structure (e.g.
-along a stream) would actually give us a direct measure of differences in potential $\Phi$ (i.e.
-$\Phi({\bf x}_1)-\Phi({\bf x}_2)=(v_2^2-v_1^2)/2$).
-In reality, we usually know at most four of the six phase-space co-ordinates for debris stars, often with significant error bars and potential measurements are made by fitting orbits to streams \citep[e.g.,][]{helmi05,johnston05,koposov10,law10,lux12}.
-Moreover, our assumption of debris tracing a single orbit is actually incorrect \citep[see][for discussions of the orbit distribution in tidal debris]{johnston98,helmi99}, with systematic changes in orbital energy along debris streams that can lead to systematic biases in our measurements of the Galactic potential \citep{eyre09,varghese11}.
+Then we can imagine that measuring the full-space velocities ${\bf v}$ at different points ${\bf x}$ in the structure (e.g. along a stream) would actually give us a direct 
+measure of differences in potential $\Phi$ (i.e. $\Phi({\bf x}_1)-\Phi({\bf x}_2)=(v_2^2-v_1^2)/2$).
+In reality, we usually know at most four of the six phase-space co-ordinates for debris stars, often with significant error bars and potential measurements are made by 
+fitting orbits to streams \citep[e.g.,][]{helmi05,johnston05,koposov10,law10,lux12}.
+Moreover, our assumption of debris tracing a single orbit is actually incorrect \citep[see][for discussions of the orbit distribution in tidal debris]{johnston98,helmi99}, with 
+systematic changes in orbital energy along debris streams that can lead to systematic biases in our measurements of the Galactic potential \citep{eyre09,varghese11}.
 
-One way to address these systematic biases is to run fully self-consistent N-body simulations of satellite destruction in a variety of potentials (which naturally generate a distribution of orbital properties in the debris and along tidal streams) with the aim of simultaneously constraining both the properties of the satellite and the Milky Way.
-Many studies of the Sagittarius debris system (hereafter Sgr)  --- whose tidal streams entirely encircle the Milky Way --- have adopted this approach \citep[e.g.][]{law05,fellhauer06} with the most recent work for the first time attempting to place constraints on the triaxiality and orientation of the Milky Way's matter distribution \citep{law11}.
-Studies which first fit the Sgr stream to a single orbit \citep{johnston05,law09} and were subsequently followed up with full N-body simulations \citep{law05,law12} offer another demonstration of the limitation of the former approach as potential parameters are revised when moving to the N-body approach.
+One way to address these systematic biases is to run fully self-consistent N-body simulations of satellite destruction in a variety of potentials (which naturally generate a 
+distribution of orbital properties in the debris and along tidal streams) with the aim of simultaneously constraining both the properties of the satellite and the Milky Way. 
+Many studies of the Sagittarius debris system (hereafter Sgr)  --- whose tidal streams entirely encircle the Milky Way --- have adopted this approach \citep[e.g.][]
+{law05,fellhauer06} with the most recent work for the first time attempting to place constraints on the triaxiality and orientation of the Milky Way's matter distribution 
+\citep{law11}.
+Studies which first fit the Sgr stream to a single orbit \citep{johnston05,law09} and were subsequently followed up with full N-body simulations \citep{law05,law12} offer 
+another demonstration of the limitation of the former approach as potential parameters are revised when moving to the N-body approach.
 
 Other approaches have been proposed that avoid the cost of full N-body simulations yet recognize the limitations of single orbit integration.
-For example, using the results of N-body simulations as a guide, simple corrections around a single orbit can be adopted to make model predictions that more closely mimic expected debris behavior \citep[e.g.][]{johnston99a,varghese2011}.
+For example, using the results of N-body simulations as a guide, simple corrections around a single orbit can be adopted to make model predictions that more closely 
+mimic expected debris behavior \citep[e.g.][]{johnston99a,varghese2011}.
 
-The promise of near-future data sets including full (or nearly full) phase-space information has also inspired discussions that move beyond fitting observables along a stream. 
-\citet{binney08} and \citet{penarrubia12} demonstrate conceptually that the distribution of energy and entropy (respectively) in debris will be minimized only for a correct assumption of the form of the Galactic potential (although neither tests their method with realistic observational errors???).
-\citet{johnston99} (hereafter JZSH) showed that backwards integration of the positions of stars and their parent satellite could be used to distinguish Milky Way mass models: only in the correct potential would the stars' and satellite's  path coincide within a Hubble time.
-The JZSH method was designed and tested for the proposed characteristics of the Space Interferometry Mission \citep{}, which was expected to be able to measure proper motions with $\mu$as/year accuracy for stars as faint as $V=20$.
+The promise of near-future data sets including full (or nearly full) phase-space information has also inspired discussions that move beyond fitting observables along a 
+stream. 
+\citet{binney08} and \citet{penarrubia12} demonstrate conceptually that the distribution of energy and entropy (respectively) in debris will be minimized only for a correct 
+assumption of the form of the Galactic potential (although neither tests their method with realistic observational errors???).
+\citet{johnston99} (hereafter JZSH) showed that backwards integration of the positions of stars and their parent satellite could be used to distinguish Milky Way mass 
+models: only in the correct potential would the stars' and satellite's  path coincide within a Hubble time.
+The JZSH method was designed and tested for the proposed characteristics of the Space Interferometry Mission \citep{}, which was expected to be able to measure 
+proper motions with $\mu$as/year accuracy for stars as faint as $V=20$.
 Distances to stream members would only be poorly known and were estimated using anticipated properties of tidal debris.
@@
 
-Over twenty-five years ago Longmore, Fernley & Jameson (1986, MNRAS, 220, 279) demonstrated the existence of a near-infrared period-luminosity (PL) relation for RR Lyrae stars, using K-band observations of these stars in Galactic globular clusters.
-Catelan, Pritzl & Smith (2004, ApJS, 154, 633) showed that a PL relation for RR Lyrae stars is to be “expected” in the infrared (but vanishingly so in the optical), explained primarily by the changing bolometric correction for these stars as a function of the observed wavelength.
-Later, Madore & Freedman (2011, ApJ, 744, 132) demonstrated the inevitability of decreasing intrinsic scatter with increasing wavelength (and slope) of PL relations both for Cepheids and for RR Lyrae stars.
+Over twenty-five years ago Longmore, Fernley & Jameson (1986, MNRAS, 220, 279) demonstrated the existence of a near-infrared period-luminosity (PL) relation for RR 
+Lyrae stars, using K-band observations of these stars in Galactic globular clusters. Catelan, Pritzl & Smith (2004, ApJS, 154, 633) showed that a PL relation for RR Lyrae 
+stars is to be “expected” in the infrared (but vanishingly so in the optical), explained primarily by the changing bolometric correction for these stars as a function of the 
+observed wavelength. Later, Madore & Freedman (2011, ApJ, 744, 132) demonstrated the inevitability of decreasing intrinsic scatter with increasing wavelength (and 
+slope) of PL relations both for Cepheids and for RR Lyrae stars.
 Moreover, both reddening effects and metallicity dependencies also decrease significantly for observations further into the IR.
 
-Taken together, these studies suggest observations of RR Lyrae variables by NASA's Spitzer mission could provide us with a stellar distance indicator of unprecedented accuracy and precision. 
-\citet{madore13} have demonstrated \citep[using 5 stars with trigonometric parallaxes measured by Hubble][]{benedict11} that the dispersion in the mid-IR PL relation at Spitzer wavelengths, is on the order of ±0.03 mag.
-The immediate implication is that it is now possible to determine individual distances that are good to better than ±2\% for single RR Lyrae stars, out to the limits of Spitzer's ability to detect and measure RR Lyrae variables (which is at least 80 kpc, for modest integration times).
+Taken together, these studies suggest observations of RR Lyrae variables by NASA's Spitzer mission could provide us with a stellar distance indicator of unprecedented 
+accuracy and precision. 
+\citet{madore13} have demonstrated \citep[using 5 stars with trigonometric parallaxes measured by Hubble][]{benedict11} that the dispersion in the mid-IR PL relation at 
+Spitzer wavelengths, is on the order of ±0.03 mag. The immediate implication is that it is now possible to determine individual distances that are good to better than ±2\% 
+for single RR Lyrae stars, out to the limits of Spitzer's ability to detect and measure RR Lyrae variables (which is at least 80 kpc, for modest integration times).
 
@@
 
-ESA's Gaia satellite, planned for launch in June 2013, is an astrometric mission which aims to measure the positions of billions of stars with 10-100$\mu$as accuracies (depending on the color and magnitude of each star), scanning the entire sky multiple times over its 5-year lifetime \citep{perryman01}.
-Gaia will revolutionize the field of Galactic Astronomy with sufficient accuracy to create full, 6-dimensional phase-space maps for distances of up to $\sim 10$kpc from the Sun \citep[see][for discussion]{}.
+ESA's Gaia satellite, planned for launch in June 2013, is an astrometric mission which aims to measure the positions of billions of stars with 10-100$\mu$as accuracies 
+(depending on the color and magnitude of each star), scanning the entire sky multiple times over its 5-year lifetime \citep{perryman01}.
+Gaia will revolutionize the field of Galactic Astronomy with sufficient accuracy to create full, 6-dimensional phase-space maps for distances of up to $\sim 10$kpc from the 
+Sun \citep[see][for discussion]{}.
 
-The left-hand panel of Figure \ref{data.fig} show Gaia's end-of-mission anticipated errors on the distance to  of an individual RR Lyrae star at different distances $d$ from the Sun.
+The left-hand panel of Figure \ref{data.fig} show Gaia's end-of-mission anticipated errors on the distance to  of an individual RR Lyrae star at different distances $d$ from 
+the Sun.
 \footnote{Taken from WWW site?.}
-Note that within 2kpc of the Sun, Gaia will be able to measure distances to these stars to better than 2\%: RR Lyraes in this volume will be used to test and calibrate the Spitzer PL relation. 
+Note that within 2kpc of the Sun, Gaia will be able to measure distances to these stars to better than 2\%: RR Lyraes in this volume will be used to test and calibrate the 
+Spitzer PL relation. 
 Beyond 2kpc, the IR PL relation for RR Lyraes will provide more accurate distance estimates than Gaia.
 
-The remaining panels of Figure \ref{data.fig} demonstrate that Gaia's anticipated proper motions errors for RR Lyraes (middle) are sufficiently accurate to distinguish tangential motions with better than $\sim$10 km/s precision out to many tens of kpc (right-panel).
-Hence, the combination of Spitzer and Gaia data offers the exciting prospect of extending the ``horizon'' of where full six-dimensional phase-space maps of our Galaxy are possible from 10kpc to 80kpc.
+The remaining panels of Figure \ref{data.fig} demonstrate that Gaia's anticipated proper motions errors for RR Lyraes (middle) are sufficiently accurate to distinguish 
+tangential motions with better than $\sim$10 km/s precision out to many tens of kpc (right-panel).
+Hence, the combination of Spitzer and Gaia data offers the exciting prospect of extending the ``horizon'' of where full six-dimensional phase-space maps of our Galaxy are 
+possible from 10kpc to 80kpc.
 
@@
 %\citep{belokurov06}.
-%Distance estimates along the tail have been derived from these photometric surveys (e.g., Mart{\'{\i}}nez-Delgado et al.
-2004, ApJ, 601, 241) and gradients in radial velocities and stellar populations measured with follow-up spectroscopy (e.g., Keller et al.
-2010, ApJ, 720, 940).
+%Distance estimates along the tail have been derived from these photometric surveys (e.g., Mart{\'{\i}}nez-Delgado et al. 2004, ApJ, 601, 241) and gradients in radial 
+velocities and stellar populations measured with follow-up spectroscopy (e.g., Keller et al. 2010, ApJ, 720, 940).
 %\citep{majewski04,martinez04,bellazzini06,chou07,chou10,keller10,carlin12}.
 %These rich data sets have inspired equally rich interpretations of Sgr debris
-% (e.g.,  Johnston et al. 1999, AJ, 118, 1719; Helmi, 2004, ApJL, 610, L97; Law et al. 2005, ApJ, 619, 807; Fellhauer et al. 2006, ApJ, 651, 167; Law \& Majewski, 2010, ApJ, 714, 229).
+% (e.g.,  Johnston et al. 1999, AJ, 118, 1719; Helmi, 2004, ApJL, 610, L97; Law et al. 2005, ApJ, 619, 807; Fellhauer et al. 2006, ApJ, 651, 167; Law \& Majewski, 2010, 
+ApJ, 714, 229).
 %(e.g.,  Law \& Majewski, 2010, ApJ, 714, 229 -- hereafter LM10).
@@
 
-The promise of the combined Spitzer and Gaia measurements of the distances to and proper motions of RR Lyraes in the Sgr debris opens up new approaches to interpretation.
+The promise of the combined Spitzer and Gaia measurements of the distances to and proper motions of RR Lyraes in the Sgr debris opens up new approaches to 
+interpretation.
 With 2\% distances measurements, the thickness of Sgr's stream is actually greater
@@
 where $d$ and $v$ are the distance and speed separation between the star
-and the parent respectively.
-The one assumption in this method is what value to adopt for the satellite mass, $m_{\rm \sat}$,which sets the instantaneous tidal radius, $r_{\rm tide}=
+and the parent respectively.  The one assumption in this method is what value to adopt for the satellite mass, $m_{\rm \sat}$,which sets the instantaneous tidal radius, 
+$r_{\rm tide}=                                                                      
 R (m_{\rm sat}/M_{\rm Gal})^{1/3}$ and escape velocity, $v_{\rm esc} =                                                               
@@
 it requires only the assumption a satellite mass $m_{\rm sat}$ and  backwards integration of orbits. 
-There are no assumptions made about or N-body modeling required to follow the internal distribution of satellite stars or accurately assess the degree of alignment of debris with a single orbit.
-In fact, the method relies on the ability to measure {\it differences} in orbital properties --- these differences result in different orbital time periods for the debris and the satellite which leads to their paths intersecting during the backwards integration.
+There are no assumptions made about or N-body modeling required to follow the internal distribution of satellite stars or accurately assess the degree of alignment of 
+debris with a single orbit.
+In fact, the method relies on the ability to measure {\it differences} in orbital properties --- these differences result in different orbital time periods for the debris and the 
+satellite which leads to their paths intersecting during the backwards integration.
 
 There are two other apparent strengths of the method which require further investigation.
-First it appears applicable to any debris that is known to come from a single object, and not restricted only to the very coldest tidal streams.
-For example, in principle the method could  applied to the vast debris {\it clouds} that have been discovered subtending hundreds-thousands square degrees \citep[e.g.
-the Triangulum Andromeda and Hercules-Aquila clouds]{rochapinto04,belokurov06}, or even stars that have only associations in orbital properties and do not form a coherent spatial structure \citep[such as the angular momentum groupings in local giants found by][]{helmi99}.
-Second, it appears trivial to combine constraints from multiple debris systems at once by simply integrating all debris from several satellites simultaneously with $D_{\rm ps}$ defined appropriately for each star separately.
+First it appears applicable to any debris that is known to come from a single object, and not restricted only to the very coldest tidal streams. For example, in principle the 
+method could  applied to the vast debris {\it clouds} that have been discovered subtending hundreds-thousands square degrees \citep[e.g. the Triangulum Andromeda 
+and Hercules-Aquila clouds]{rochapinto04,belokurov06}, or even stars that have only associations in orbital properties and do not form a coherent spatial structure 
+\citep[such as the angular momentum groupings in local giants found by][]{helmi99}.
+Second, it appears trivial to combine constraints from multiple debris systems at once by simply integrating all debris from several satellites simultaneously with $D_{\rm 
+ps}$ defined appropriately for each star separately.
 
-The limitations of the method also need to be understood.
-In particular, for the RR Lyraes associated with  the very coldest streams \citep[e.g.
-the globular clusters Pal5 and GD1][]{odenkirchen02,koposov10} the measurement errors are unlikely to be small enough for the
+The limitations of the method also need to be understood. In particular, for the RR Lyraes associated with  the very coldest streams \citep[e.g. the globular clusters Pal5 
+and GD1][]{odenkirchen02,koposov10} the measurement errors are unlikely to be small enough for the 
 the differences in orbital properties to between debris and satellite to be detected, and the method should fail.
 
-Lastly, the current version of the algorithm relies on accurate knowledge of the phase-space position of the parent satellite, which may not be available in some cases \citep[for example, for the Orphan Stream][]{belokurov07}.
-Given sufficient data, position and motion of the original host could be solved for as additional free parameters, but the effectiveness of this approach has yet to be investigated.
+Lastly, the current version of the algorithm relies on accurate knowledge of the phase-space position of the parent satellite, which may not be available in some cases 
+\citep[for example, for the Orphan Stream][]{belokurov07}. Given sufficient data, position and motion of the original host could be solved for as additional free parameters, 
+but the effectiveness of this approach has yet to be investigated.
 
@@
 
-This paper outlines an algorithm to measure the Galactic potential that takes maximal advantage of a combination of possible data from the Spitzer and Gaia satellite missions that, when combined,  promise full phase-space measurements of RR Lyrae stars around our Galaxy.
-When applied to a synthetic data set generated from debris in an N-body simulation of the destruction of Sgr the method recovers the shape and orientation of the dark matter halo component of the potential with ???
-accuracy using a sample of ???
-stars.
-This success provides strong motivation for: (i) further theoretical work to investigate the power of multiple debris structures with different morphologies and orientations in building a comprehensive picture of the Galactic mass distribution; and (ii) a  Spitzer survey of RR Lyrae stars in debris structures around the Milky Way as a ``legacy'' sample to combine with near-future Gaia data.
+This paper outlines an algorithm to measure the Galactic potential that takes maximal advantage of a combination of possible data from the Spitzer and Gaia satellite 
+missions that, when combined,  promise full phase-space measurements of RR Lyrae stars around our Galaxy.
+When applied to a synthetic data set generated from debris in an N-body simulation of the destruction of Sgr the method recovers the shape and orientation of the dark 
+matter halo component of the potential with ??? accuracy using a sample of ??? stars.
+This success provides strong motivation for: (i) further theoretical work to investigate the power of multiple debris structures with different morphologies and orientations in 
+building a comprehensive picture of the Galactic mass distribution; and (ii) a  Spitzer survey of RR Lyrae stars in debris structures around the Milky Way as a ``legacy'' 
+sample to combine with near-future Gaia data.
 

### 2013-06-19 streams/coordinates/core.py
add gsr to hel function to coordinates

@@ -1,0 +1,37 @@
+# coding: utf-8
+
+""" """
+
+from __future__ import division, print_function
+
+__author__ = "adrn <adrn@astro.columbia.edu>"
+
+# Standard library
+import os, sys
+
+# Third-party
+import numpy as np
+from numpy import cos, sin
+
+import astropy.coordinates as coord
+import astropy.units as u
+
+__all__ = ["gsr_to_hel"]
+
+def gsr_to_hel(l, b, v_gsr, v_sun_lsr=[10.,5.25,7.17]*u.km/u.s, v_circ=220*u.km/u.s):
+    """ Convert a velocity from the Galactic standard of rest (GSR) to 
+        heliocentric radial velocity. 
+    """
+    assert hasattr(v_gsr, 'unit')
+    assert hasattr(l, 'radian')
+    assert hasattr(b, 'radian')
+    
+    v_lsr = v_gsr - v_circ * sin(l.radian) * cos(b.radian)
+    
+    # velocity correction for Sun relative to LSR
+    v_correct = v_sun_lsr[0]*cos(b.radian)*cos(l.radian) + \
+                v_sun_lsr[1]*cos(b.radian)*sin(l.radian) + \
+                v_sun_lsr[2]*sin(b.radian)
+    v_hel = v_lsr - v_correct
+    
+    return v_hel

### 2013-06-19 streams/data/catalogs.py
attempt to deal with QUESt data better...

@@ -49,12 +49,22 @@ def read_quest():
         Vivas et al. 2004. 
         
         - Covers Stripe 82
+        - Spectral data from:
+            + http://iopscience.iop.org/1538-3881/129/1/189/fulltext/204289.tables.html
+            + Spectroscopy of bright QUEST RR Lyrae stars (Vivas+, 2008)
+            
     """
     fits_filename = os.path.join(project_root, "data", "catalog", \
                                 "quest_vivas2004_RRL.fits")
+    spec_filename = os.path.join(project_root, "data", "catalog", \
+                                "quest_spectroscopy.txt")
+                                
     hdulist = fits.open(fits_filename)
     tb_data = np.array(hdulist[1].data)
     data = Table(tb_data)
+    
+    spec_data = ascii.read(spec_filename, data_start=1)
+    spec_data = spec_data[(spec_data["sigma_fit"] > 0) & (spec_data["sigma_gamma"] > 0)]
     
     # Map RA/Dec columns to easier names
     data.add_column(Column(np.array(data["RAJ2000"]).astype(float), 
@@ -65,6 +75,23 @@ def read_quest():
                            name=str("V")))
     data.add_column(Column(rrl_photometric_distance(data['V'], -1.5), 
                            name="dist", units=u.kpc))
+    data.rename_column("__VZA2004_", "ID")
+    
+    vgsr = []
+    vgsr_err = []
+    for row in data:
+        if row["ID"] in spec_data["ID"]:
+            spec_row = spec_data[spec_data["ID"] == row["ID"]]
+            vgsr.append(spec_row["vgsr"])
+            vgsr_err.append(spec_row["sigma_fit"])
+        else:
+            vgsr.append(np.nan)
+            vgsr_err.append(np.nan)
+            
+        row["dist"] = spec_data["dist"]
+    
+    data.add_column(Column(vgsr, name="vgsr", units=u.km/u.s))
+    data.add_column(Column(vgsr_err, name="vgsr_err", units=u.km/u.s))
     
     data["ra"].units = u.degree
     data["dec"].units = u.degree

### 2013-06-19 docs/paper1/paper.tex
first block of text from me

@@
-\documentclass{emulateapj}
-%\documentclass[preprint]{aastex}
+%\documentclass{emulateapj}
+\documentclass[preprint]{aastex}
 %\documentclass{aastex}
@@
 \usepackage{natbib}
+\usepackage{amssymb,amsmath}
+
+% Convenience
+\newcommand{\bs}{\boldsymbol}
+\newcommand{\argmin}{\operatornamewithlimits{argmin}}
 
@@
 
-\begin{figure}
-\plotone{figs/data.eps}
-\caption{Summary of data sensitivity: $\Delta d$, $\Delta \mu$ and $\Delta v_{\rm tan}$ as function of $d$.
-{\it Also put Sgr and Orphan characteristics on this plot?}
-\label{data.fig}
-}
-\end{figure}
+%% \begin{figure}
+%% \plotone{figs/data.eps}
+%% \caption{Summary of data sensitivity: $\Delta d$, $\Delta \mu$ and $\Delta v_{\rm tan}$ as function of $d$.
+%% {\it Also put Sgr and Orphan characteristics on this plot?}
+%% \label{data.fig}
+%% }
+%% \end{figure}
 
@@
 
-\begin{figure}
-\plotone{figs/sgr.eps}
-\caption{
-\label{sgr.fig}
-}
-\end{figure}
+%% \begin{figure}
+%% \plotone{figs/sgr.eps}
+%% \caption{
+%% \label{sgr.fig}
+%% }
+%% \end{figure}
 
@@
 
-\begin{figure}
-\plotone{figs/boots.eps}
-\caption{
-\label{boots.fig}
-}
-\end{figure}
+%% \begin{figure}
+%% \plotone{figs/boots.eps}
+%% \caption{
+%% \label{boots.fig}
+%% }
+%% \end{figure}
 
@@
 
+% Above here is from KVJ. Below is APW.
+
+\section{Method}
+With access to full phase-space information for stars in an associated
+debris structure, it is no longer necessary to model the full orbit of
+the stream. Instead, it is possible to use the individual stars as
+potential measures by exploting the simple fact that the stars must
+have come from the same progenitor. If the present-day orbital
+parameters of the progenitor system are known, this places an
+intuitive constraint on the potential of the Milky Way: if the orbits
+of the stars and the orbit of the progenitor are integrated
+\emph{backwards} in time, the stars should `recombine' with the
+progenitor. Conversely, if the potential is incorrect, the orbits of
+the stars will diverge from that of the progenitor and thus will not
+be recaptured by the satellite system.
+
+The starting point for applying this method is a sample of stars with
+known full space kinematics $(\bs{x}_i, \bs{v}_i)|_{t=0}$, the orbital
+parameters for the progenitor system $(\bs{x}_p, \bs{v}_p)|_{t=0}$,
+and a functional form for the potential, $\Phi$\footnote{Future work
+  will investigate a more general ... basis function blah blah.}. For
+a given set of potential parameters, the orbits of the stars and
+progenitor are integrated backwards for some time (determined from an
+estimate of the interaction time of the system). [APW: explain that
+  this is not Nbody, just test particles!] At each timestep $t_j$, for
+each particle $i$, a set of normalized, relative phase-space
+coordinates, $(\bs{q}_i(t_j), \bs{p}_i(t_j))$, are computed using the
+instantaneous tidal radius, escape velocity, and phase-space position
+of the progenitor. At a particular timestep, these coordinates are
+defined as
+\begin{equation}
+  \bs{q}_i = \frac{\bs{x}_i -
+    \bs{x}_p}{R_{tide}}\,\,\,\,\,\,\,\,\bs{p}_i = \frac{\bs{v}_i -
+    \bs{v}_p}{v_{esc}}.
+\end{equation}
+These definitions require an estimate of the mass of the satellite,
+$m_{sat}$, which, combined with the computed enclosed mass of the
+potential, $m_{enc}$, sets the instantaneous tidal radius and escape
+velocity,
+\begin{equation}
+  R_{tide}=\Big(\frac{m_{sat}}{3m_{enc}}\Big)^{1/3}\,\,\,\,\,\,\,\,
+  v_{esc}=\sqrt{\frac{2Gm_{sat}}{R_{tide}}}.
+\end{equation}
+
+Qualitatively, when $D_{ps}=\sqrt{|\bs{q}_i|^2+|\bs{p}_i|^2} <
+\sqrt{2}$, the star is probably recaptured by the satellite, however
+this is not imposed as a hard constraint. Instead, we construct a 6D
+distribution of the minimum phase-space vectors for each star over the
+individual, integrated orbits. For each star, $i$, the phase-space
+distance, $D_{ps}$, is computed at each timestep $t_j$, and the vector
+with the minimum phase-space distance is stored
+\begin{align}
+  t^*_i &= \argmin_{t} D_{ps,i}\\
+  A_{ik} &= (\bs{q}_i(t^*_i),\bs{p}_i(t^*_i))
+\end{align}
+where $k\in\{1,6\}$.
+
+% Discussion?
+One of the assumptions of orbit fitting methods for measuring the
+potential is that the stars in a given stream are on the same
+orbit. \citep{binney??} show that this assumption leads to biased
+potential parameters, even without realistic observational errors on
+the kinematic information. Conversely, the method presented above is
+free of this assumption and instead \emph{relies} on the fact that
+each star is on an independent orbit. The power of this method is in
+the joint constraint that the best potential is the one that gets the
+most of these individual orbits to cross the orbit of the progenitor
+over the history of the system.
+
+% APW Note: The optimization, in a way, is getting a bunch of orbits (curves) to intersect another curve (the progenitor orbit) at some points (over time).
+
 \section{Summary}

### 2013-06-19 scripts/infer_log_potential.py
add overwrite option

@@ -12,6 +12,7 @@ __author__ = "adrn <adrn@astro.columbia.edu>"
 
 # Standard library
 import os, sys
+import shutil
 import copy
 import logging
 from datetime import datetime
@@ -140,7 +141,10 @@ def main(config_file):
             path = os.path.join(config["output_path"], iso_now)
         
         if os.path.exists(path):
-            raise IOError("Path {0} already exists!".format(path))
+            if config.get("overwrite", False):
+                shutil.rmtree(path)
+            else:
+                raise IOError("Path {0} already exists!".format(path))
             
         os.mkdir(path)
     

### 2013-06-19 streams/inference/backintegrate.py
use generalized variance not sum of variances

@@ -92,5 +92,5 @@ def generalized_variance(potential, particle_orbits, satellite_orbit):
     # cov_matrix -> (6 x 6) covariance matrix for particles
     w,v = np.linalg.eig(cov_matrix)
     
-    return np.sum(w)
-
+    #return np.sum(w)
+    return np.prod(w)

### 2013-06-19 streams/integrate/satellite_particles.py

@@ -57,7 +57,7 @@ class SatelliteParticleIntegrator(LeapfrogIntegrator):
             
         times = [t1]
         
-        Ntimesteps = int(5000.*resolution)
+        Ntimesteps = int(10000.*resolution)
         rs = np.zeros((Ntimesteps,) + self.r_im1.shape, dtype=float)
         vs = np.zeros((Ntimesteps,) + self.v_im1.shape, dtype=float)
         rs[0] = self.r_im1

### 2013-06-19 docs/proposals/mdmprop_2013F_ASheffield.tex

% MDMPROP.TEX -- MDM electronic proposal form.
% Revision: September 4, 2003 --- valid for the Fall 2010 observing semester.
% DEADLINE: 5:00 PM Friday, April 16, 2010
%
% General Instructions:
%
%   I. Where/how/when to submit this form:
%
%      Give a paper copy to Jules or his mailbox on the 10th floor
%      of Pupin by the above deadline, or email the postscript or
%      pdf file to jules@astro.columbia.edu.
%
% THE FORM STARTS HERE

% please don't modify or delete this line.
\documentstyle[mdmprop,11pt]{article}

\begin{document}

% Give a descriptive title for the proposal in the \title command.
%
%    \title{TEXT}
%
% Note that a title can be quite long; LaTeX will break the title into
% separate lines automatically.  If you wish to indicate line breaks
% yourself, do so with a `\\' command at the appropriate point in
% the title text.

\title{Probing the Milky Way's Dark Matter Halo with RR Lyraes in Tidal Streams I: The Triangulum-Andromeda Stellar Feature}

% Investigator's (PI and CoI) information blocks.
%
% The PI must be affiliated with Columbia University.
% Please give each investigator's name, email address.
% At least one investigator must be listed as definitely
% being present at the telescope.
%
%    \name{OBSERVER NAME}
%    \emailaddress{EMAIL ADDRESS}
%    \atthetelescope{Y/N}
%
% DO NOT remove the \begin{PI} and \end{PI} or the \begin{CoI} and
% \end{CoI} lines.

\begin{PI}
\name{Adrian Price-Whelan}			% REQUIRED
\emailaddress{adrn@astro.columbia.edu}		% REQUIRED
\atthetelescope{Y}       % REQUIRED
\end{PI}

\begin{CoI}
\name{Allyson Sheffield}			% REQUIRED
\emailaddress{asheffield@astro.columbia.edu}		% REQUIRED
\atthetelescope{Y}       % REQUIRED
\end{CoI}

\begin{CoI}
\name{Kathryn Johnston}			% REQUIRED
\emailaddress{kvj@astro.columbia.edu}		% REQUIRED
\atthetelescope{N}       % REQUIRED
\end{CoI}

\begin{CoI}
\name{Branimir Sesar}			% REQUIRED
\emailaddress{bsesar@astro.caltech.edu}		% REQUIRED
\atthetelescope{N}       % REQUIRED
\end{CoI}

% You can supply more CoI blocks, but only the first three will be
% printed on the form.

% Give a general abstract of the scientific justification appropriate for
% a non-specialist.  Write the abstract between the \begin{abstract} and
% \end{abstract} lines.  Limit yourself to approximately 175 words.

% DO NOT remove the \begin{abstract} and \end{abstract} lines.

\begin{abstract}
This proposal requests time to obtain radial velocity measurements for a sample of RR Lyrae in the Triangulum-Andromeda stellar cloud -- one of several new, amorphous stellar structures discovered recently in the Galactic halo. Models suggest that, like the more well-known stellar streams, these clouds could result from the disruption of satellites during the hierarchical formation of our Galaxy. Thus, this debris is equally suited for probing the shape, orientation, and radial profile of dark matter in the Galactic halo. At the distance of TriAnd ($\sim$ 25~kpc), we will soon have accurate proper motions from GAIA. Combined with accurate distances using the mid-IR period-luminosity relation for RR Lyrae (TODO: something about the Spitzer proposal?), we only need radial velocity measurements to obtain full 6-D phase-space information for this sample of stars. We will then use this 6D information to model the orbit of the progenitor and subsequently infer new constraints on the mass, axis ratios, and orientation of the Milky Way's dark matter halo.

\end{abstract}

% Indicate whether this proposal is part of your PhD thesis work by
% putting a Y or an N inside the \thesis{} curly braces.

\thesis{N}

% Indicate whether you are requesting long-term status by putting
% a Y or an N inside the \longterm{} curly braces.  If you do want
% long-term status (a "Y" answer), please tell us the number of nights,
% number of semesters, and what telescopes by filling in \longtermdetails{}.
% Please be brief; \longtermdetails is limited to one line.

\longterm{}
\longtermdetails{}

% List the details of the observing runs being requested, for
% UP TO THREE runs.  The parameters for each run are segregated
% between \begin{obsrun} and \end{obsrun} lines.  Please be sure
% that the information is isolated properly for each run.
%
%   \telescope{}	For example, \telescope{1.3m}
%   \instrument{}	For example, \instrument{Mk III + Charlotte}
%   \lunardays{}	For example, \lunardays{14}
%   \optimaldates{}	For example, \optimaldates{Sep 1 - Nov 30}
%   \acceptabledates{}	For example, \acceptabledates{Sep 1 - Dec 15}
%
%
% Instrument combinations may be specified with "+".
%
% \numnights should give the number of nights of the run
% \lunardays should specify the maximum number of nights from new moon
% which can be utilized to accomplish your scientific goals.  It should
% be a number from 1 (new moon) to 14 (full moon).
%
% \optimaldates should contain the range of OPTIMAL dates.
%
% \acceptabledates should give the range of ACCEPTABLE dates (i.e., you
% would not accept time outside those limits).
%
% To enter the acceptable and optimal date ranges, please use two
% dash-separated dates with 3-letter abbreviations for the month
% (Jan, Feb, Mar, Apr, May, Jun, Jul, Aug, Sep, Oct, Nov, Dec)
% followed by the day.  For example:  \optimaldates{Feb 15 - Apr 23}
%
% If you need to enter two or more ranges of acceptable or optimal dates
% for a single observing run, separate the ranges by commas.  For example:
% \acceptabledates{Oct 7 - Nov 24, Jan 7 - Jan 31}
% It is not necessary to give date ranges based on lunar phase information.
% For instance, if you wish to observe an object in March through May
% within five days of new moon, you may give \lunardays{5} and
% \optimaldates{Mar 1 - May 31} instead of multiple shorter date ranges.
%
% DO NOT remove the \begin{obsrun} and \end{obsrun} blocks.

\begin{obsrun}
\telescope{2.4-m}
\instrument{ModSpec, Wilbur, 600 l mm$^{-1}$ blazed at 5000 \AA}
\numnights{5}
\lunardays{grey}
\optimaldates{Oct 4 - Oct 8}
\acceptabledates{Oct 1 - Oct 20}
\end{obsrun}

\begin{obsrun}
\telescope{}
\instrument{}
\numnights{}
\lunardays{}
\optimaldates{}
\acceptabledates{}
\end{obsrun}

\begin{obsrun}
\telescope{}
\instrument{}
\numnights{}
\lunardays{}
\optimaldates{}
\acceptabledates{}
\end{obsrun}

% You may NOT supply more obsrun blocks.  Three is the limit.

% If there are dates that you cannot use for non-astronomical reasons,
% (i.e., other than moon phase or when your object is up)
% please give the dates by filling in the curly braces in \unusabledates{}.
% Please be brief; \unusabledates is LIMITED TO ONE LINE.

\unusabledates{}

% In the following "essay question" sections, the delimiting pieces of
% markup (\justification, \feasibility, etc.) act as LaTeX \section*{}
% commands.  If the author wanted to have numbered subsections within
% any of these, LaTeX's \subsection could be used.

% SCIENTIFIC JUSTIFICATION
%
% Give the scientific justification for the proposed observations.
% This section should consist of paragraphs of text (and may include
% EPS figures) that follow the \justification line.
% Try to include an explanation of the overall significance to astronomy.

% In order to include an EPS plot, you should use the LaTeX "figure"
% environment.  The plot file is included with the \plotone{FILENAME}
% command; two side-by-side plot files can be included by typing
% \plottwo{FILENAME1}{FILENAME2}.  Use \caption{} to specify a caption.
% The \epsscale{} command can be used to scale \plotone plots if they
% appear too large on the printed page.
%
% \begin{figure}
% \epsscale{0.85}
% \plotone{sample.eps}
% \caption{Sample figure showing important results.}
% \end{figure}
%
% If you need to rotate or make other transformations to a figure, you may
% use the \plotfiddle command:
% \plotfiddle{PSFILE}{VSIZE}{ROTANG}{HSCALE}{VSCALE}{HTRANS}{VTRANS}
% \plotfiddle{sample.eps}{2.6in}{-90.}{32.}{32.}{-250}{225}
% where HSCALE and VSCALE are percentages and HTRANS and VTRANS are
% in PostScript units, 72 PS units = 1 inch.
%
% If you wish to use the "reference" environment, follow
% the following example:
%
%\begin{references}
%\reference Armandroff \& Massey 1991 AJ 102, 927.
%\reference Berkhuijsen \& Humphreys 1989 A\&A 214, 68.
%\reference Massey 1993 in Massive Stars: Their Lives in the Interstellar
%  Medium (Review), ed. J. P. Cassinelli and E. B. Churchwell, p. 168.
%\reference Massey, Armandroff, \& Pyke 1993, in prep.
%\end{references}

\justification Galaxies, by mass, are mostly composed of \emph{Dark
  Matter}. Simulations of structure and galaxy formation -- both with
and without baryons -- lead us to believe that such \emph{halos} of
dark matter have universal density profiles \citep{nfw96} and are
triaxial in shape \citep{jing02}. Including baryons in such
simulations tends to soften the triaxiality and alters the radial
density profile. Thus, precise measurements of the shape, orientation,
and radial profile of dark matter halos provides information about
both dark matter physics and the baryonic processes that helped shape
them. Tidal streams from infalling satellite galaxies have been used
to constrain properties of the Milky Way's halo \citep[e.g.][]{law10};
typically found outwards of 20~kpc, the orbits of associated stars are
still largely under the influence of the dark matter and thus serve as
potentiometers \citep[e.g.][]{johnston99}. Using an updated method
that utilizes full phase-space (6D) kinematic information for a small
sample ($\sim$100s) of stars stripped from a known progenitor (e.g.,
Sagittarius; Price-Whelan et al 2013, in prep), it is possible to
accurately constrain the axis ratios, total mass, and orientation of
the Milky Way's dark matter halo. This method relies on a technique
for deriving distances to individual RR Lyrae stars with 2\% accuracy
\citep{madore12}, along with anticipated GAIA proper motion errors of
$\sim10-20~\mu\mathrm{as}/\mathrm{yr}$ for such stars at $\sim$
20-30~kpc. This leaves one remaining velocity component -- the radial
velocity.

We plan to observe a sample of RR Lyrae stars selected to be likley
members of the TriAnd structure using a catalogs of RRab stars from
the Palomar Transient Factory \citep[PTF][]{ptf}

%%AAS insert some background on TriAnd here, including recent (unpublished) results from KVJ models showing that the feature is part of a stream rather than a cloud

\begin{references}

\end{references}

%\begin{figure}
%\epsscale{1.00}
%\plotone{HA_hist.eps}
%\caption{Counts of 2MASS M giants. The left-hand panel shows M giant star counts in the latitude range $30^{\circ} < |b| < 50^{\circ}$ for the northern Galactic hemisphere at negative l (dotted bold line) and southern Galactic hemisphere at positive l (solid bold line). Gray lines indicate counts in the corresponding north/south positive/negative l regions. The right hand panels show the distance distribution for M giants in the bold histogram (south) in the left hand panel in the longitude ranges $15^{\circ} < l < 40^{\circ}$ (black) and $0^{\circ} < l < 15^{\circ}$ (gray).  The magnitude range for stars in the figure is $9.5 \le K_{S,0} \le 12.5$}
%\end{figure}

% FEASIBILITY
%
% Assess the technical and scientific feasibility of the observations.
% This section should consist of text and tables only (no figures)
% following the \feasibility line.
%
% List objects, coordinates, and magnitudes (or surface brightness,
% if appropriate), desired S/N, wavelength coverage and resolution.
% Justify the number of nights requested as well as the specific
% telescope(s), instruments, and lunar phase.  Indicate the optimal
% detector, as well as acceptable alternates.  If you've requested
% long-term status, explain why this is necessary for successful
% completion of the science.

\feasibility
Our program aims to map the velocity profile of the southern portion of the Hercules-Aquila (HerAq) stellar
cloud.
The RR Lyraes that we will target spans the sky in the region roughly $0^{h}<\alpha<4^{h}$ and $30^{\circ}<\delta<40^{\circ}$.
This range of equatorial coordinates is best observed in October 2013.

All targets are selected as RR Lyraes.
The range in magnitude is $16.2 \le R \le 18.2$.

Using the 600 groove mm$^{-1}$ grating blazed at 5000 \AA\ and a 2048x2048 detector with 15 micron pixels (Wilbur), we will have wavelength coverage of about 1300 \AA\, including the H$\beta$ line at 4861 \AA\ and the [OI] sky line at 5577 \AA\ (the latter to aid in the radial velocity determination by identifying any systematic offsets).
This set-up will give an expected resolution of 3-4 \AA.
We will collect bias and flat field frames at the beginning and end of the night, and ThAr arc lamp spectra for wavelength calibration will be taken throughout the night to account for telescope flexure.  Radial velocity standards will be selected that are of a similar spectral type as the targets (e.g., XXXX  -- a dwarf of spectral type F8).

We expect to
be able to get an RV precision of 5 km s$^{-1}$;
We prefer S/N $>30$ to get good metallicity estimates as well.
Using the Exposure Time Calculator we expect to be able to
achieve S/N=30 per pixel in XXX sec for
a $R=16$ RR Lyrae, and in XXX sec for a ....
To determine the wavelength solution, comparison lamp spectra will be taken throughout the night, specifically after a long telescope slew. In addition, quartz lamp spectra and bias frames will be obtained at the beginning of the night.
With 1-2 hours every night reserved for obtaining
RV standards and calibration frames, we can expect to observe about
10 targets per night.
We request 5 nights to obtain spectra of several dozen potential HerAq cloud members.

Tech Spec Summary:

Groove density: 600 lines mm$^{-1}$

Blaze $\lambda$=5000 \AA

CCD=``Wilbur'': 15 $\mu$m pixels

% OTHER FACILITIES
%
% Why MDM? If you are using other facilities for this project,
% explain how the MDM observations fit into the scheme of things.
%
% This section should consist of text and tables only (no figures)
% following the \feasibility line.

\whymdm
ModSpec on the 2.4-m Hiltner Telescope is a propitious choice for our HerAq cloud radial velocity survey -- the spectral band covered will provide not only precision RVs but also an estimate of the metallicities.
% PAST USE
%
% List your allocation of telescope time at MDM during the
% past 3 years, and describe the status of the project (cite
% publications where appropriate).  Mark any allocations of time
% related to the current proposal with a \relatedwork{} command.
% Are your MDM observations achieving their goals?

\thepast
I used Mark III on the 1.3-m telescope in January.
The run was very successful --- I obtained spectra for close to 200 M giants, and the RVs are preliminarily at the 10-20 km s$^{-1}$ precision level.

I used ModSpec on the 2.4-m telescope in May.
This was an excellent run.  The SkyCalc application made the acquisition of guide stars a very simple task (more time was needed to do this on the January 1.3-m run).

I have an upcoming run on the 2.4-m in November, which is aimed at targeting the TriAnd star cloud.

\end{document}

### 2013-06-21 docs/paper1/paper.tex
paper updates

@@
   t^*_i &= \argmin_{t} D_{ps,i}\\
-  A_{ik} &= (\bs{q}_i(t^*_i),\bs{p}_i(t^*_i))
+  A_{ik} &= (\bs{q}_i(t^*_i),\bs{p}_i(t^*_i)) % A_ik are really samples from a distribution, or we treat them as such
 \end{align}
-where $k\in\{1,6\}$.
+where $k\in\{1,6\}$. The variance of this distribution will be larger
+for orbits integrated in an incorrect potential relative to the
+distribution computed from the `true' orbital history of the
+stars. Thus, the \emph{generalized variance} of the distribution,
+computed for a given set of potential parameters, $\bs{\theta}$, is a
+natural choice for the scalar objective function used in constraining
+the potential of the Milky Way,
+\begin{equation}
+  f(\bs{\theta})=\det(\mathrm{Cov}(A_{ik}))
+\end{equation}
+
+\subsection{Application to Simulated Data}
 

### 2013-06-22 docs/paper1/paper.tex
sketch out structure, add Gaia info

@@
 
+% --------------------------------------------------------------------
 % Above here is from KVJ. Below is APW.
+%
+
+\section{Context and motivation}
+\subsection{2\% distance errors for RR Lyrae in the halo}
+
+\subsection{Gaia and the age of astrometry}
+The Gaia satellite \citep{gaia2001}, planned for launch in October
+2013, is an astrometric mission which, amongst a myriad of other
+science goals, aims to measure the positions of billions of stars with
+10-100~$\mu$as accuracies. With over 50 observations of the entire sky
+over the 5-year lifetime of the mission (cite?), Gaia will
+revolutionize the field of Galactic Astronomy by providing sufficient
+proper motion accuracy to create full, six-dimensional phase-space
+maps of the Galaxy for heliocentric distances of up to \sim 10~kpc for
+xxx stars.
+
+\subsection{The Sagittarius debris system}
 
 \section{Method}
-With access to full phase-space information for stars in an associated
-debris structure, it is no longer necessary to model the full orbit of
-the stream. Instead, it is possible to use the individual stars as
-potential measures by exploting the simple fact that the stars must
-have come from the same progenitor. If the present-day orbital
-parameters of the progenitor system are known, this places an
-intuitive constraint on the potential of the Milky Way: if the orbits
-of the stars and the orbit of the progenitor are integrated
-\emph{backwards} in time, the stars should `recombine' with the
-progenitor. Conversely, if the potential is incorrect, the orbits of
-the stars will diverge from that of the progenitor and thus will not
-be recaptured by the satellite system.
+With access to full phase-space information for stars in a tidal
+stream, it is no longer necessary to model the orbit of the stream
+itself. Instead, it is possible to use the individual stars as
+potential measures by exploting the fact that the stars must have come
+from the same progenitor. If the present-day orbital parameters of the
+progenitor system are known, this places an intuitive constraint on
+the potential of the Milky Way: if the orbits of the stars and the
+orbit of the progenitor are integrated \emph{backwards} in time, the
+stars should `recombine' with the progenitor. If the potential is
+incorrect, the orbits of the stars will diverge from that of the
+progenitor and thus will not be recaptured by the satellite system.
 
-The starting point for applying this method is a sample of stars with
-known full space kinematics $(\bs{x}_i, \bs{v}_i)|_{t=0}$, the orbital
-parameters for the progenitor system $(\bs{x}_p, \bs{v}_p)|_{t=0}$,
-and a functional form for the potential, $\Phi$\footnote{Future work
-  will investigate a more general ... basis function blah blah.}. For
-a given set of potential parameters, the orbits of the stars and
+Quantifying this method requires a sample of stars with known full
+space kinematics $(\bs{x}_i, \bs{v}_i)|_{t=0}$, the orbital parameters
+for the progenitor system $(\bs{x}_p, \bs{v}_p)|_{t=0}$, and a
+functional form for the potential, $\Phi$\footnote{Future work will
+  investigate a more general ... basis function blah blah.}. For a
+given set of potential parameters, the orbits of the stars and
 progenitor are integrated backwards for some time (determined from an
-estimate of the interaction time of the system). [APW: explain that
-  this is not Nbody, just test particles!] At each timestep $t_j$, for
-each particle $i$, a set of normalized, relative phase-space
-coordinates, $(\bs{q}_i(t_j), \bs{p}_i(t_j))$, are computed using the
-instantaneous tidal radius, escape velocity, and phase-space position
-of the progenitor. At a particular timestep, these coordinates are
-defined as
+estimate of the interaction time of the system). The stars and
+progenitor system are treated as test particles, so that the debris
+stars do not become re-bound to the satellite. Instead, at each
+timestep $t_j$, for each particle $i$, a set of normalized, relative
+phase-space coordinates, $(\bs{q}_i(t_j), \bs{p}_i(t_j))$, are
+computed using the instantaneous tidal radius, escape velocity, and
+phase-space position of the progenitor. At a particular timestep,
+these coordinates are defined as
 \begin{equation}
@@
 Qualitatively, when $D_{ps}=\sqrt{|\bs{q}_i|^2+|\bs{p}_i|^2} <
-\sqrt{2}$, the star is probably recaptured by the satellite, however
+\sqrt{2}$, the star is probably recaptured by the satellite, however,
 this is not imposed as a hard constraint. Instead, we construct a 6D
@@
 \end{align}
-where $k\in\{1,6\}$. The variance of this distribution will be larger
-for orbits integrated in an incorrect potential relative to the
-distribution computed from the `true' orbital history of the
-stars. Thus, the \emph{generalized variance} of the distribution,
-computed for a given set of potential parameters, $\bs{\theta}$, is a
-natural choice for the scalar objective function used in constraining
-the potential of the Milky Way,
+where $k\in\{1,6\}$. Intuitively, the variance of this distribution
+will be larger for orbits integrated in an incorrect potential
+relative to the distribution computed from the `true' orbital history
+of the stars. Thus, the \emph{generalized variance} of the
+distribution, computed for a given set of potential parameters,
+$\bs{\theta}$, is a natural choice for the scalar objective function
+used in constraining the potential of the Milky Way,
 \begin{equation}
-  f(\bs{\theta})=\det(\mathrm{Cov}(A_{ik}))
+  f(\bs{\theta})=\det(\mathrm{Cov}(A_{ik})).
 \end{equation}

### 2013-06-23 docs/paper1/paper.tex
paper updates

@@
 
-Over twenty-five years ago Longmore, Fernley & Jameson (1986, MNRAS, 220, 279) demonstrated the existence of a near-infrared period-luminosity (PL) relation for RR
-Lyrae stars, using K-band observations of these stars in Galactic globular clusters. Catelan, Pritzl & Smith (2004, ApJS, 154, 633) showed that a PL relation for RR Lyrae
+Over twenty-five years ago Longmore, Fernley \& Jameson (1986, MNRAS, 220, 279) demonstrated the existence of a near-infrared period-luminosity (PL) relation for RR
+Lyrae stars, using K-band observations of these stars in Galactic globular clusters. Catelan, Pritzl \& Smith (2004, ApJS, 154, 633) showed that a PL relation for RR Lyrae
 stars is to be “expected” in the infrared (but vanishingly so in the optical), explained primarily by the changing bolometric correction for these stars as a function of the
-observed wavelength. Later, Madore & Freedman (2011, ApJ, 744, 132) demonstrated the inevitability of decreasing intrinsic scatter with increasing wavelength (and
+observed wavelength. Later, Madore \& Freedman (2011, ApJ, 744, 132) demonstrated the inevitability of decreasing intrinsic scatter with increasing wavelength (and
 slope) of PL relations both for Cepheids and for RR Lyrae stars.
@@
 
-The Sagittarius Dwarf Galaxy (hereafter Sgr) was discovered
-serendipitously during a radial velocity survey of the Galactic bulge
-(Ibata et al. 1994, Nature, 370, 194). While the number of known MW
-satellites has more than doubled recently (e.g.,
-McConnachie, 2012, AJ, 144, 4), in many ways Sgr remains the most
-studied and most intriguing --- it is the closest satellite, among the
-largest (only the LMC and SMC are more luminous), and has the most
-prominent tidal tails.
-From the first maps of the Sgr core (Ibata et al. 1994; Ibata et
-al. 1995, MNRAS, 277, 781) its elongated morphology suggested
-destruction by tidal forces, as might be anticipated
-from its proximity to the Galaxy.  Subsequent N-body models confirmed
-the validity of this interpretation (Velazquez \& White, 1995, MNRAS,
-275, L23) and showed that debris from Sgr should
-form coherent streams of stars encircling our Galaxy (Johnston et
-al. 1995, ApJ, 451, 598). Signatures of these giant star streams have
-since been discovered in many studies, and comprehensively mapped
-across the sky in carbon stars (Totten \& Irwin, 1998, MNRAS, 294, 1)
-M giant stars selected from the Two Micron All Sky Survey (Majewski et
-al.  2003, ApJ, 599, 1082) and main sequence turnoff stars selected
-from the Sloan Digital Sky Survey (Belokurov et al. 2006, ApJ Lett, 642, L137).
-%\citep{belokurov06}.
-%Distance estimates along the tail have been derived from these photometric surveys (e.g., Mart{\'{\i}}nez-Delgado et al. 2004, ApJ, 601, 241) and gradients in radial
-velocities and stellar populations measured with follow-up spectroscopy (e.g., Keller et al. 2010, ApJ, 720, 940).
-%\citep{majewski04,martinez04,bellazzini06,chou07,chou10,keller10,carlin12}.
-%These rich data sets have inspired equally rich interpretations of Sgr debris
-% (e.g.,  Johnston et al. 1999, AJ, 118, 1719; Helmi, 2004, ApJL, 610, L97; Law et al. 2005, ApJ, 619, 807; Fellhauer et al. 2006, ApJ, 651, 167; Law \& Majewski, 2010,
-ApJ, 714, 229).
-%(e.g.,  Law \& Majewski, 2010, ApJ, 714, 229 -- hereafter LM10).
-
-Figure
-\ref{sgr.fig}  illustrates results from LM10's N-body
-simulation of dwarf satellite disruption along the expected Sgr orbit, with particles projected onto the plane perpendicular to the
-Galactic disk containing both the Sun and Galactic center.
-The simulation was run in a potential…..
-
-LM10 demonstrate how a comparison of such simulations with the
-observed debris enables a detailed reconstruction of Sgr's history ---
-it's mass and stellar content, orbit, rate of destruction and even
-original distribution in populations.
-(See also other refs ????)
-Moreover, the phase-space
-distribution of the debris offers a unique probe of the depth, shape
-and extent of the Milky Way's dark matter halo that cannot be matched
-using other techniques on our own or other galaxies (e.g., Ibata et
-al. 2001, ApJ, 551, 294).  Most recently, LM10 have combined all
-current data on Sgr debris to assess the triaxiality and orientation
-of the outer Galaxy --- the first time that such a reconstruction of
-the 3-dimensional mass distribution of a dark matter halo has been
-feasible.  These constraints on Sgr's mass and orbit have
-invigorated discussions of Sgr in a more cosmological context, the
-mass and extent of the original dark matter halo that hosted it as well
-as its effect on the MW itself (e.g., Michel-Dansac et
-al. 2011, MNRAS, 414, L1).
-%\citep{bailin03,purcell11,michel11,gomez12}.
-
-The promise of the combined Spitzer and Gaia measurements of the distances to and proper motions of RR Lyraes in the Sgr debris opens up new approaches to
-interpretation.
-With 2\% distances measurements, the thickness of Sgr's stream is actually greater
-than the uncertainty.  Moreover,
-with Gaia's final data release, there will be proper motion
-measurements of accuracies less than the intrinsic velocity dispersion for much
-of the stream (e.g., corresponding to a tangential velocity accuracy
-of 10 km/s for an RR Lyrae at 35kpc) and ground-based radial
-velocities of few km/s accuracies can easily be added.
-Hence, for the first time, the measured phase-space position of each individual RR Lyrae
-relative to the stream is actually physically meaningful.
-The next section outlines a method to take full advantage of this accuracy and use each RR Lyrae individually as a potential measure.
-
 %% \begin{figure}
@@
 where $d$ and $v$ are the distance and speed separation between the star
-and the parent respectively.  The one assumption in this method is what value to adopt for the satellite mass, $m_{\rm \sat}$,which sets the instantaneous tidal radius,
+and the parent respectively.  The one assumption in this method is what value to adopt for the satellite mass, $m_{\rm sat}$,which sets the instantaneous tidal radius,
 $r_{\rm tide}=
@@
 proper motion accuracy to create full, six-dimensional phase-space
-maps of the Galaxy for heliocentric distances of up to \sim 10~kpc for
-xxx stars.
+maps of the Galaxy for heliocentric distances of up to $\sim$ 10~kpc (xx kpc)
+for $V\sim xxx$ ($V\sim yyy$) stars.
+
+Figure~\ref{fig:gaia_errors} show the Gaia end-of-mission distance and
+tangential velocity error estimates for a typical halo RR Lyrae star
+\citep[{$M_V=0.45$, [Fe/H]$=-1.5$};][]{ivezic2008}. Within 2~kpc, Gaia
+will be able to measure distances to these stars with better than 2\%
+accuracy: RR Lyraes in this volume will be used to test and calibrate
+the Spitzer PL relation described above. Beyond the 2~kpc threshold,
+the mid-IR PL relation for RR Lyraes will provide more accurate
+distance estimates than Gaia.
+
+The anticipated proper motion errors for RR Lyrae correspond to
+tangential velocity errors of order $\sim$1~km/s at 30~kpc, and of
+order $\sim$10~km/s out to several tens of kiloparsecs. The
+combination of Spitzer and Gaia data will extend the `horizon' of
+where full six-dimensional phase-space maps of the Galaxy are possible
+from $<10$~kpc to 80~kpc. This enormous increase in volume will refine
+data on existing debris systems such as Sagittarius, the Orphan
+stream, and [...], and enable discovery of new substructure within the
+Galactic halo.
 
 \subsection{The Sagittarius debris system}
+The Sagittarius Dwarf Galaxy (hereafter Sgr) was discovered
+serendipitously during a radial velocity survey of the Galactic bulge
+\citep{ibata1994}. While the number of known Milky Way satellites has
+more than doubled recently \citep[e.g.,][]{mcconnachie2012}, in many
+ways Sgr remains the most studied and most intriguing: it is the
+closest satellite, among the largest (only the LMC and SMC are more
+luminous), and has the most prominent tidal tails.
+
+The first maps of the Sgr core \citep{ibata1994, ibata1995} revealed
+its elongated morphology and suggested tidal deformation, as would be
+anticipated from its proximity to the Galaxy. Subsequent N-body models
+confirmed this interpretation \citep{velazquez1995} and showed that
+debris from Sgr should form coherent streams of stars that encircle
+the Galaxy \citep{johnston1995}. Signatures of these [..] stellar
+streams have since been discovered in many studies and have been
+mapped acorss the sky in carbon stars \citep{totten1998}, M giants
+selected from the Two Micron All Sky Survey \citep{majewski2003}, main
+sequence turnoff stars from the Sloan Digital Sky Survey
+\citep{belokurov2006}, and RR Lyrae in the Catalina Sky Survey
+\citep{drake2013}.
+
+The distance of the stream has been estimated from photometric surveys \citep[e.g.,][]{martinezdelgado2004} and gradients in radial velocities and metallicity measured with follow-up spectroscopy \citep[e.g.,][]{vivas2005, keller2010}.
+
+(e.g., Keller et al. 2010, ApJ, 720, 940).
+%\citep{majewski04,martinez04,bellazzini06,chou07,chou10,keller10,carlin12}.
+%These rich data sets have inspired equally rich interpretations of Sgr debris
+% (e.g.,  Johnston et al. 1999, AJ, 118, 1719; Helmi, 2004, ApJL, 610, L97; Law et al. 2005, ApJ, 619, 807; Fellhauer et al. 2006, ApJ, 651, 167; Law \& Majewski, 2010,
+ApJ, 714, 229).
+%(e.g.,  Law \& Majewski, 2010, ApJ, 714, 229 -- hereafter LM10).
+
+Figure
+\ref{sgr.fig}  illustrates results from LM10's N-body
+simulation of dwarf satellite disruption along the expected Sgr orbit, with particles projected onto the plane perpendicular to the
+Galactic disk containing both the Sun and Galactic center.
+The simulation was run in a potential…..
+
+LM10 demonstrate how a comparison of such simulations with the
+observed debris enables a detailed reconstruction of Sgr's history ---
+it's mass and stellar content, orbit, rate of destruction and even
+original distribution in populations.
+(See also other refs ????)
+Moreover, the phase-space
+distribution of the debris offers a unique probe of the depth, shape
+and extent of the Milky Way's dark matter halo that cannot be matched
+using other techniques on our own or other galaxies (e.g., Ibata et
+al. 2001, ApJ, 551, 294).  Most recently, LM10 have combined all
+current data on Sgr debris to assess the triaxiality and orientation
+of the outer Galaxy --- the first time that such a reconstruction of
+the 3-dimensional mass distribution of a dark matter halo has been
+feasible.  These constraints on Sgr's mass and orbit have
+invigorated discussions of Sgr in a more cosmological context, the
+mass and extent of the original dark matter halo that hosted it as well
+as its effect on the MW itself (e.g., Michel-Dansac et
+al. 2011, MNRAS, 414, L1).
+%\citep{bailin03,purcell11,michel11,gomez12}.
+
+The promise of the combined Spitzer and Gaia measurements of the distances to and proper motions of RR Lyraes in the Sgr debris opens up new approaches to
+interpretation.
+With 2\% distances measurements, the thickness of Sgr's stream is actually greater
+than the uncertainty.  Moreover,
+with Gaia's final data release, there will be proper motion
+measurements of accuracies less than the intrinsic velocity dispersion for much
+of the stream (e.g., corresponding to a tangential velocity accuracy
+of 10 km/s for an RR Lyrae at 35kpc) and ground-based radial
+velocities of few km/s accuracies can easily be added.
+Hence, for the first time, the measured phase-space position of each individual RR Lyrae
+relative to the stream is actually physically meaningful.
+The next section outlines a method to take full advantage of this accuracy and use each RR Lyrae individually as a potential measure.
 
@@
 
-\bibliography{spitzer}
+\bibliography{refs}
 \bibliographystyle{apj}

### 2013-06-24 docs/paper1/paper.tex
updated references

@@
 
-The distance of the stream has been estimated from photometric surveys \citep[e.g.,][]{martinezdelgado2004} and gradients in radial velocities and metallicity measured with follow-up spectroscopy \citep[e.g.,][]{vivas2005, keller2010}.
+The distance of the stream has been estimated from photometric surveys
+\citep[e.g.,][]{martinezdelgado2004} and gradients in radial
+velocities and metallicity measured with follow-up spectroscopy
+\citep[e.g.,][]{majewski2004, vivas2005, bellazzini2006, chou2007,
+  chou2010, keller2010, carlin2012}. The multitude of Sgr stream data
+has inspired a rich set of interpretations and models of the debris
+\citep[e.g.,][]{johnston1999, helmi2004, law2005, fellhauer2006,
+  law2010}.
 
-(e.g., Keller et al. 2010, ApJ, 720, 940).
-%\citep{majewski04,martinez04,bellazzini06,chou07,chou10,keller10,carlin12}.
-%These rich data sets have inspired equally rich interpretations of Sgr debris
-% (e.g.,  Johnston et al. 1999, AJ, 118, 1719; Helmi, 2004, ApJL, 610, L97; Law et al. 2005, ApJ, 619, 807; Fellhauer et al. 2006, ApJ, 651, 167; Law \& Majewski, 2010,
-ApJ, 714, 229).
-%(e.g.,  Law \& Majewski, 2010, ApJ, 714, 229 -- hereafter LM10).
-
-Figure
-\ref{sgr.fig}  illustrates results from LM10's N-body
-simulation of dwarf satellite disruption along the expected Sgr orbit, with particles projected onto the plane perpendicular to the
-Galactic disk containing both the Sun and Galactic center.
-The simulation was run in a potential…..
+Figure \ref{fig:lm10} illustrates results from the \cite[][; hereafter
+  LM10]{law2010} N-body simulation of dwarf satellite disruption along
+the expected Sgr orbit, with particles projected onto the plane
+perpendicular to the Galactic disk containing both the Sun and
+Galactic center.  The simulation was run in a potential…..
 

### 2013-06-25 docs/paper1/paper.tex
paper update

@@
 
-\section{Context for the study}
-
-The method  in Section \ref{measure.sec} takes advantage of three distinct developments in the observations of tidal streams:
-(i) the demonstration of a technique for deriving distances to individual RR Lyrae stars with 2\% accuracies (see Section \ref{spitzer.sec});
-(ii) the prospect of proper motion measurements of the same stars with $\sim$10 mas/year  precision (see Section  \ref{gaia.sec});
-and (iii) the tracing of debris associated with Sgr  entirely around the Galaxy (see Section \ref{sgr.sec})
-
-\subsection{Spitzer RR Lyraes}
-
-\label{spitzer.sec}
-
-%WISE: benedict11? 'precision of spritzer will far exceed these' also 'serendipi…
-%Text 'stolen' from proposal 1: the Carnegie RR Lyrae Program
-
-Over twenty-five years ago Longmore, Fernley \& Jameson (1986, MNRAS, 220, 279) demonstrated the existence of a near-infrared period-luminosity (PL) relation for RR
-Lyrae stars, using K-band observations of these stars in Galactic globular clusters. Catelan, Pritzl \& Smith (2004, ApJS, 154, 633) showed that a PL relation for RR Lyrae
-stars is to be “expected” in the infrared (but vanishingly so in the optical), explained primarily by the changing bolometric correction for these stars as a function of the
-observed wavelength. Later, Madore \& Freedman (2011, ApJ, 744, 132) demonstrated the inevitability of decreasing intrinsic scatter with increasing wavelength (and
-slope) of PL relations both for Cepheids and for RR Lyrae stars.
-Moreover, both reddening effects and metallicity dependencies also decrease significantly for observations further into the IR.
-
-Taken together, these studies suggest observations of RR Lyrae variables by NASA's Spitzer mission could provide us with a stellar distance indicator of unprecedented
-accuracy and precision.
-\citet{madore13} have demonstrated \citep[using 5 stars with trigonometric parallaxes measured by Hubble][]{benedict11} that the dispersion in the mid-IR PL relation at
-Spitzer wavelengths, is on the order of ±0.03 mag. The immediate implication is that it is now possible to determine individual distances that are good to better than ±2\%
-for single RR Lyrae stars, out to the limits of Spitzer's ability to detect and measure RR Lyrae variables (which is at least 80 kpc, for modest integration times).
-
-\subsection{Gaia and Spitzer}
-\label{gaia.sec}
-
-ESA's Gaia satellite, planned for launch in June 2013, is an astrometric mission which aims to measure the positions of billions of stars with 10-100$\mu$as accuracies
-(depending on the color and magnitude of each star), scanning the entire sky multiple times over its 5-year lifetime \citep{perryman01}.
-Gaia will revolutionize the field of Galactic Astronomy with sufficient accuracy to create full, 6-dimensional phase-space maps for distances of up to $\sim 10$kpc from the
-Sun \citep[see][for discussion]{}.
-
-The left-hand panel of Figure \ref{data.fig} show Gaia's end-of-mission anticipated errors on the distance to  of an individual RR Lyrae star at different distances $d$ from
-the Sun.
-\footnote{Taken from WWW site?.}
-Note that within 2kpc of the Sun, Gaia will be able to measure distances to these stars to better than 2\%: RR Lyraes in this volume will be used to test and calibrate the
-Spitzer PL relation.
-Beyond 2kpc, the IR PL relation for RR Lyraes will provide more accurate distance estimates than Gaia.
-
-The remaining panels of Figure \ref{data.fig} demonstrate that Gaia's anticipated proper motions errors for RR Lyraes (middle) are sufficiently accurate to distinguish
-tangential motions with better than $\sim$10 km/s precision out to many tens of kpc (right-panel).
-Hence, the combination of Spitzer and Gaia data offers the exciting prospect of extending the ``horizon'' of where full six-dimensional phase-space maps of our Galaxy are
-possible from 10kpc to 80kpc.
-
-%% \begin{figure}
-%% \plotone{figs/data.eps}
-%% \caption{Summary of data sensitivity: $\Delta d$, $\Delta \mu$ and $\Delta v_{\rm tan}$ as function of $d$.
-%% {\it Also put Sgr and Orphan characteristics on this plot?}
-%% \label{data.fig}
-%% }
-%% \end{figure}
-
-\subsection{The Sagittarius debris system}
-
-%% \begin{figure}
-%% \plotone{figs/sgr.eps}
-%% \caption{
-%% \label{sgr.fig}
-%% }
-%% \end{figure}
-
 \section{Proposed potential measure}
@@
 \section{Context and motivation}
+The method presented in Section \ref{sec:method} takes advantage of
+three distinct developments in the observations of tidal streams: (i)
+the demonstration of a technique for deriving distances to individual
+RR Lyrae stars with 2\% accuracies (Section \ref{sec:spitzer}); (ii)
+the prospect of proper motion measurements of the same stars with
+$\sim$10~mas/yr precision (Section \ref{sec:gaia}); and (iii) the
+tracing of debris associated with Sgr around the entire Galaxy
+(Section \ref{sec:sgr})
+
 \subsection{2\% distance errors for RR Lyrae in the halo}
+\label{sec:spitzer}
+Over twenty-five years ago, \citet{longmore1986} demonstrated the
+existence of a near-infrared period-luminosity (PL) relation for RR
+Lyrae stars using $K$-band observations of these stars in Galactic
+globular clusters. More recently, \citet{catelan2004} showed that a PL
+relation for RR Lyrae stars is expected in the infrared -- but
+vanishingly so in the optical -- and is explained primarily by the
+changing bolometric correction for these stars as a function of the
+observed wavelength. Later, \citet{madore2012} demonstrated the
+inevitability of decreasing intrinsic scatter with increasing
+wavelength (and slope) of PL relations both for Cepheids and for RR
+Lyrae stars. Both reddening effects and metallicity dependencies also
+decrease significantly for observations further into the IR.
+
+These studies suggest that observations of RR Lyrae variables by
+NASA's Spitzer mission could provide us with a stellar distance
+indicator of unprecedented accuracy and precision. \citet{madore2012}
+have demonstrated \citep[using 5 stars with trigonometric parallaxes
+  measured by Hubble;][]{benedict2011} that the dispersion in the
+mid-IR PL relation--at Spitzer wavelengths--is of order
+$\pm$0.03~mag. This implies that it is possible to determine
+individual distances that are good to better than $\pm 2\%$ for RR
+Lyrae stars out to $\sim$80~kpc (the limit of Spitzer's ability to
+detect and measure RR Lyrae variables, for modest integration
+times).
 
 \subsection{Gaia and the age of astrometry}
+\label{sec:gaia}
 The Gaia satellite \citep{gaia2001}, planned for launch in October
@@
 \subsection{The Sagittarius debris system}
+\label{sec:sgr}
 The Sagittarius Dwarf Galaxy (hereafter Sgr) was discovered
@@
 streams have since been discovered in many studies and have been
-mapped acorss the sky in carbon stars \citep{totten1998}, M giants
+mapped across the sky in carbon stars \citep{totten1998}, M giants
 selected from the Two Micron All Sky Survey \citep{majewski2003}, main
@@
 
-Figure \ref{fig:lm10} illustrates results from the \cite[][; hereafter
-  LM10]{law2010} N-body simulation of dwarf satellite disruption along
-the expected Sgr orbit, with particles projected onto the plane
-perpendicular to the Galactic disk containing both the Sun and
-Galactic center.  The simulation was run in a potential…..
+Most recently, \citet[][hereafter LM10]{law2010} have combined all
+current data on Sgr debris to assess the triaxiality and orientation
+of the outer Galaxy --- the first time that such a reconstruction of
+the 3-dimensional mass distribution of a dark matter halo has been
+feasible. Figure \ref{fig:lm10} illustrates results from the LM10
+N-body simulation of dwarf satellite disruption along the expected Sgr
+orbit. The particle positions are shown in a coordinate system defined
+by the orbit of the satellite orbit \citep{majewski2004}. The
+simulation was run in a three-component potential, with a triaxial,
+logarithmic halo model of the form
+\begin{equation}
+  \Phi_{halo} = v^2 \ln(C_1 x^2 + C_2 y^2 + C_3 xy + (z/q_z)^2 + R_c^2).
+\end{equation}
 
@@
 observed debris enables a detailed reconstruction of Sgr's history ---
-it's mass and stellar content, orbit, rate of destruction and even
-original distribution in populations.
-(See also other refs ????)
-Moreover, the phase-space
-distribution of the debris offers a unique probe of the depth, shape
-and extent of the Milky Way's dark matter halo that cannot be matched
-using other techniques on our own or other galaxies (e.g., Ibata et
-al. 2001, ApJ, 551, 294).  Most recently, LM10 have combined all
-current data on Sgr debris to assess the triaxiality and orientation
-of the outer Galaxy --- the first time that such a reconstruction of
-the 3-dimensional mass distribution of a dark matter halo has been
-feasible.  These constraints on Sgr's mass and orbit have
-invigorated discussions of Sgr in a more cosmological context, the
-mass and extent of the original dark matter halo that hosted it as well
-as its effect on the MW itself (e.g., Michel-Dansac et
-al. 2011, MNRAS, 414, L1).
-%\citep{bailin03,purcell11,michel11,gomez12}.
+it's mass and stellar populations, orbit, rate of destruction, and
+even original distribution in populations (other refs??). Moreover,
+the phase-space distribution of the debris offers a unique probe of
+the depth, shape, extent, and orientation of the Milky Way's dark
+matter halo that cannot be matched using other techniques on our own
+or other galaxies \citep[e.g.,][]{ibata2001}. The LM10 constraints on
+Sgr's mass and orbit have invigorated discussions of Sgr in a more
+cosmological context, the mass and extent of the original dark matter
+halo that hosted it, as well as its effect on the Milky Way itself
+\citep{bailin2003,purcell2011,micheldansac2011,gomez2012}.
 
-The promise of the combined Spitzer and Gaia measurements of the distances to and proper motions of RR Lyraes in the Sgr debris opens up new approaches to
-interpretation.
-With 2\% distances measurements, the thickness of Sgr's stream is actually greater
-than the uncertainty.  Moreover,
-with Gaia's final data release, there will be proper motion
-measurements of accuracies less than the intrinsic velocity dispersion for much
-of the stream (e.g., corresponding to a tangential velocity accuracy
-of 10 km/s for an RR Lyrae at 35kpc) and ground-based radial
-velocities of few km/s accuracies can easily be added.
-Hence, for the first time, the measured phase-space position of each individual RR Lyrae
-relative to the stream is actually physically meaningful.
-The next section outlines a method to take full advantage of this accuracy and use each RR Lyrae individually as a potential measure.
+The combined Spitzer and Gaia measurements of distances and proper
+motions of RR Lyraes in the Sgr debris opens up new avenues of
+interpretation. A 2\% distance error at $\sim$30~kpc is smaller than
+the predicted thickness of the stream, ranging from
+$\sim2-10$~kpc. Combined with Gaia tangential velocity measurement
+accuracies ($\sim$2~km/s at 30~kpc) less than the intrinsic dispersion
+for much of the stream ($\sigma_v\sim ??$), and ground-based radial
+velocities with $\sim10$~km/s accuracy, the measured phase-space
+position of each individual RR Lyrae relative to the stream becomes a
+powerful probe of the potential: the next section outlines a new
+method to take advantage of this information and use each star as an
+individual potential measure.
 
 \section{Method}
+\label{sec:method}
 With access to full phase-space information for stars in a tidal
@@
 for the progenitor system $(\bs{x}_p, \bs{v}_p)|_{t=0}$, and a
-functional form for the potential, $\Phi$\footnote{Future work will
-  investigate a more general ... basis function blah blah.}. For a
+functional form for the potential, $\Phi$.\footnote{Future work will
+  investigate a more general ... basis function blah blah.} For a
 given set of potential parameters, the orbits of the stars and
@@
     \bs{x}_p}{R_{tide}}\,\,\,\,\,\,\,\,\bs{p}_i = \frac{\bs{v}_i -
-    \bs{v}_p}{v_{esc}}.
+    \bs{v}_p}{v_{esc}}
 \end{equation}
-These definitions require an estimate of the mass of the satellite,
-$m_{sat}$, which, combined with the computed enclosed mass of the
-potential, $m_{enc}$, sets the instantaneous tidal radius and escape
-velocity,
+where $\bs{x}_i$ and $\bs{v}_i$ are the 3D position and velocity of
+the $i$th particle, and $\bs{x}_p$ and $\bs{v}_p$ are the same for the
+satellite.  These definitions require an estimate of the mass of the
+satellite, $m_{sat}$, which, combined with the computed enclosed mass
+of the potential, $m_{enc}$, sets the instantaneous tidal radius and
+escape velocity,
 \begin{equation}
@@
 
-Qualitatively, when $D_{ps}=\sqrt{|\bs{q}_i|^2+|\bs{p}_i|^2} <
-\sqrt{2}$, the star is probably recaptured by the satellite, however,
-this is not imposed as a hard constraint. Instead, we construct a 6D
-distribution of the minimum phase-space vectors for each star over the
-individual, integrated orbits. For each star, $i$, the phase-space
-distance, $D_{ps}$, is computed at each timestep $t_j$, and the vector
-with the minimum phase-space distance is stored
+Qualitatively, when the distance in this normalized 6D space,
+$D_{ps}=\sqrt{|\bs{q}_i|^2+|\bs{p}_i|^2} < \sqrt{2}$, the star is
+probably recaptured by the satellite (but this is not imposed as a
+hard constraint). As a statistical measure over all stars, we
+construct a 6D distribution of the minimum phase-space vectors for
+each star over the individual, integrated orbits. For each star, $i$,
+the phase-space distance, $D_{ps}$, is computed at each timestep
+$t_j$, and the vector with the minimum phase-space distance is stored
 \begin{align}
@@
 relative to the distribution computed from the `true' orbital history
-of the stars. Thus, the \emph{generalized variance} of the
-distribution, computed for a given set of potential parameters,
-$\bs{\theta}$, is a natural choice for the scalar objective function
-used in constraining the potential of the Milky Way,
+of the stars; in an incorrect potential, small differences in the
+orbits of the stars relative to the orbit of the progenitor cause them
+to spread out in this relative, normalized phase space. Thus, the
+\emph{generalized variance} of the distribution, computed for a given
+set of potential parameters, $\bs{\theta}$, is a natural choice for
+the scalar objective function used in constraining the potential of
+the Milky Way,
 \begin{equation}
@@
 \subsection{Application to Simulated Data}
+The LM10 (see Section \ref{sec:sgr}) simulation data is a perfect
+test-bed for evaluating the effectiveness of this method. We start by
+extracting both particle data and the satellite orbital parameters
+from the present-day snapshot of the simulation
+data.\footnote{www.astro.virginia.edu/~srm4n/Sgr/data.html} We then
+`observe' a sample of 100 stars from the first leading and trailing
+wraps of the stream: the radial velocity errors are drawn from a
+Gaussian $\epsilon_{rv} \sim \mathcal{N}(\mu=0,\sigma=10~{\rm km/s})$,
+the tangential velocity errors depend on the distance to the star and
+are computed from the expected Gaia error curve (shown in
+Figure~\ref{fig:gaia}), and the distance errors on a star of distance
+$D$ are $\epsilon_{dist} \sim \mathcal{N}(0,0.02\times
+D)$. Figure~\ref{fig:} shows all 2D projections of the 6D minimum
+phase-space distance distribution for the observed particles in the
+correct and an incorrect potential. Visually, it is clear that the
+variance of the distribution increases in the wrong potential. The
+generalized variance defines a convex hypersurface over which we
+optimize four of the six logarithmic potential parameters:
+$v_{circ}$, $\phi$, $q_1$, and $q_z$. Figure~\ref{fig:} shows [...]
+curves produced by varying the true parameters of the potential by
+10\% around the true value.
 

### 2013-06-25 scripts/paper1_figures.py
script for generative paper figures

@@ -1,0 +1,55 @@
+# coding: utf-8
+
+""" A script for making figures for our Streams Paper 1 """
+
+from __future__ import division, print_function
+
+__author__ = "adrn <adrn@astro.columbia.edu>"
+
+# Standard library
+import os, sys
+
+# Third-party
+import astropy.units as u
+import numpy as np
+import matplotlib.pyplot as plt
+
+from streams.data.gaia import parallax_error, proper_motion_error, \
+                              rr_lyrae_V_minus_I, rr_lyrae_M_V, \
+                              apparent_magnitude
+from streams.potential import LawMajewski2010
+from streams.integrate.satellite_particles import SatelliteParticleIntegrator
+from streams.data import lm10_particles, lm10_satellite, lm10_time
+
+def normed_objective_plot():
+    """ Plot our objective function in each of the 4 parameters we vary """
+    
+    # Can change this to the true adaptive functions so I can compare
+    timestep2 = lambda *args,**kwargs: -1.
+    resolution = 1.
+    
+    # Read in the LM10 data
+    particles = lm10_particles(N=100, expr="(Pcol>0) & (abs(Lmflag)==1)")
+    satellite = lm10_satellite()
+    t1,t2 = lm10_time()
+    resolution = 3.
+    
+    variances = dict()
+    for param in ['q1','qz','v_halo','phi']:
+        if not variances.has_key(param):
+            variances[param] = []
+            
+        stats = np.linspace(true_params[param]*0.9,
+                            true_params[param]*1.1, 
+                            10)
+    
+        for stat in stats:
+            params = true_params.copy()
+            params[param] = stat
+            lm10 = LawMajewski2010(**params)
+            integrator = SatelliteParticleIntegrator(lm10, satellite, particles)
+            s_orbit,p_orbits = integrator.run(timestep_func=timestep2,
+                                      timestep_args=(lm10, satellite.m.value),
+                                      resolution=resolution,
+                                      t1=t1, t2=t2)
+            variances[param].append(generalized_variance_prod(lm10, p_orbits, s_orbit))

### 2013-06-25 scripts/quest_stripe82.py
script for selecting RR Lyrae in QUESt and Stripe 82

@@ -1,0 +1,141 @@
+# coding: utf-8
+
+""" Compile data from:
+    1) Photometric catalogs of RR Lyrae from Stripe 82 and QUEST
+        QUEST:
+        Stripe 82: 
+        
+
+"""
+
+from __future__ import division, print_function
+
+__author__ = "adrn <adrn@astro.columbia.edu>"
+
+# Standard library
+import os, sys
+
+# Third-party
+import numpy as np
+from astropy.table import Table, Column, vstack, join
+import astropy.io.fits as fits
+import astropy.io.ascii as ascii
+import astropy.coordinates as coord
+import astropy.units as u
+import matplotlib.pyplot as plt
+
+# Project
+from streams.util import project_root
+from streams.observation.rrlyrae import *
+from streams.data import add_sgr_coordinates
+from streams.data.simulation import lm10_particle_data
+
+def read_quest():
+    """ Read in the QUEST data -- RR Lyrae from the QUEST survey,
+        Vivas et al. 2004. 
+        
+        - Photometry from:
+            http://vizier.cfa.harvard.edu/viz-bin/VizieR?-source=J/AJ/127/1158
+        - Spectral data from:
+            http://iopscience.iop.org/1538-3881/129/1/189/fulltext/204289.tables.html
+            Spectroscopy of bright QUEST RR Lyrae stars (Vivas+, 2008)
+            
+    """
+    phot_filename = os.path.join(project_root, "data", "catalog", \
+                                 "quest_vivas2004_phot.tsv")
+    phot_data = ascii.read(phot_filename, delimiter="\t", data_start=3)
+    
+    # With more spectral data, add here
+    vivas2004_spec = ascii.read(os.path.join(project_root, "data", 
+                                             "catalog", "quest_vivas2004_spec.tsv"),
+                                delimiter="\t")
+                                
+    vivas2008_spec = ascii.read(os.path.join(project_root, "data", 
+                                             "catalog", "quest_vivas2008_spec.tsv"),
+                                delimiter="\t", data_start=3)
+    
+    spec_data = vstack((vivas2004_spec, vivas2008_spec))
+    all_data = join(left=phot_data, right=spec_data, keys=['[VZA2004]'], join_type='outer')
+    
+    new_columns = dict()
[…]

### 2013-06-25 text/proposals/mdm_proposal_may2013.tex
move docs to text

--- 2013-04-26 docs/mdm_proposal_may2013.tex
@@
 \begin{document}
-\title{Probing the Milky Way's Dark Matter Halo with RR Lyraes in Tidal Streams I: The Sagittarius Leading Arm}
-\author{Adrian Price-Whelan\altaffilmark{1}, Kathryn Johnston\altaffilmark{1}, Allyson Sheffield\altaffilmark{1}}
-\altaffiltext{1}{Department of Astronomy, Columbia University, Mail Code 5246, New York, NY 10027}
-Galaxies, by mass, are mostly composed of \emph{Dark Matter}. Simulations of structure and galaxy formation -- both with
+\title{Probing the Milky Way's Dark Matter Halo with RR Lyraes in
+  Tidal Streams I: The Sagittarius Leading Arm} \author{Adrian M.
+  Price-Whelan\altaffilmark{1}, Kathryn Johnston\altaffilmark{1},
+  Allyson Sheffield\altaffilmark{1}} \altaffiltext{1}{Department of
+  Astronomy, Columbia University, Mail Code 5246, New York, NY 10027}
+
+Galaxies, by mass, are mostly composed of \emph{Dark
+  Matter}. Simulations of structure and galaxy formation -- both with
 and without baryons -- lead us to believe that such \emph{halos} of
-dark matter have universal density profiles \citep{nfw96} and are triaxial
-in shape \citep{jing02}. Including baryons in such simulations tends to
-soften the triaxiality and alters the radial density profile. Thus,
-precise measurements of the shape, orientation, and radial profile of
-dark matter halos provides information about both dark matter physics
-and the baryonic processes that helped shape them. Tidal streams from
-infalling satellite galaxies have been used to constrain properties of
-the Milky Way's halo \citep[e.g.][]{law10}; typically found outwards of 20~kpc, the
-orbits of associated stars are still largely under the influence of
-the dark matter and thus serve as potentiometers \citep[e.g.][]{johnston99}. Using an updated
-method that utilizes full phase-space (6D) kinematic information for a
-small sample ($\sim$100s) of stars stripped from a known progenitor
-(e.g., Sagittarius; Price-Whelan et al 2013, in prep), it is
-possible to accurately constrain the axis ratios, total mass, and
-orientation of the Milky Way's dark matter halo. This method relies on a technique for deriving
-distances to individual RR Lyrae stars with 2\% accuracy \citep{},
-along with anticipated GAIA proper motion errors of
+dark matter have universal density profiles \citep{nfw96} and are
+triaxial in shape \citep{jing02}. Including baryons in such
+simulations tends to soften the triaxiality and alters the radial
+density profile. Thus, precise measurements of the shape, orientation,
+and radial profile of dark matter halos provides information about
+both dark matter physics and the baryonic processes that helped shape
+them. Tidal streams from infalling satellite galaxies have been used
+to constrain properties of the Milky Way's halo \citep[e.g.][]{law10};
+typically found outwards of 20~kpc, the orbits of associated stars are
+still largely under the influence of the dark matter and thus serve as
+potentiometers \citep[e.g.][]{johnston99}. Using an updated method
+that utilizes full phase-space (6D) kinematic information for a small
+sample ($\sim$100s) of stars stripped from a known progenitor (e.g.,
+Sagittarius; Price-Whelan et al 2013, in prep), it is possible to
+accurately constrain the axis ratios, total mass, and orientation of
+the Milky Way's dark matter halo. This method relies on a technique
+for deriving distances to individual RR Lyrae stars with 2\% accuracy
+\citep{madore12}, along with anticipated GAIA proper motion errors of
 $\sim10-20~\mu\mathrm{as}/\mathrm{yr}$ for such stars at $\sim$
 20-30~kpc. This leaves one remaining velocity component -- the radial
-velocity.  We plan to observe a
-sample of RR Lyrae stars selected to be likley members of Sgr's debris system 
-using catalogs of RRab stars from
-time-domain, photometric surveys such as LINEAR \citep{}, PTF \citep{}
-and CSS \citep{}. 
-Our observations will both confirm membership of the debris system in order to construct a target list for Spitzer (proposals are due in August 2013) as well
-as measure the missing velocity dimension.
+velocity.  We plan to observe a sample of RR Lyrae stars selected to
+be likley members of Sgr's debris system using catalogs of RRab stars
+from time-domain, photometric surveys such as LINEAR \citep{linear}, PTF
+\citep{ptf} and CSS \citep{catalina}.  Our observations will both confirm
+membership of the debris system in order to construct a target list
+for Spitzer (proposals are due in August 2013) as well as measure the
+missing velocity dimension.
 
-We propose to measure radial velocities of RR Lyraes in the Sgr leading arm using ModSpec on the 2.4m Hiltner
-telescope at MDM.
-Using the 600 groove mm$^{-1}$ grating blazed at 5000 \AA\ and a 2048x2048 detector with 15 micron pixels (Wilbur), we will have wavelength coverage of about 1000 \AA\, including the H$\beta$ line at 4861 \AA\ and the [OI] sky line at 5577 \AA\ (the latter to aid in the radial velocity determination by identifying any systematic offsets).
-We will collect bias and flat field frames at the beginning and end of the night, and ThAr arc lamp spectra for wavelength calibration will be taken throughout the night to account for telescope flexure.
-Radial velocity standards will be selected that are of a similar spectral type as the targets (e.g., HD122693  -- a dwarf of spectral type F8).
-Typical equatorial coordinates for the targets are ($\alpha,\delta$)=(13$^{h}-16^{h}$,-5$^{\circ}- +25^{\circ}$); stars in this range will be at airmass $<2$ all night.
-The typical $R$-band magnitude is 16.
-To achieve S/N of 30 per pixel with the proposed set-up, an integration time of 1200 seconds is needed for a target with $R$=16.
-Considering the total amount of observing time of 9 hours per night and the overhead needed for calibration frames, we anticipate observing approximately 40 RR Lyraes.
-Reductions will be carried out by Sheffield using a well-developed reduction pipeline for processing ModSpec data.
+We propose to measure radial velocities of RR Lyraes in the Sgr
+leading arm using ModSpec on the 2.4m Hiltner telescope at MDM.  Using
+the 600 groove mm$^{-1}$ grating blazed at 5000 \AA\ and a 2048x2048
+detector with 15 micron pixels (Wilbur), we will have wavelength
+coverage of about 1000 \AA\, including the H$\beta$ line at 4861
+\AA\ and the [OI] sky line at 5577 \AA\ (the latter to aid in the
+radial velocity determination by identifying any systematic offsets).
+We will collect bias and flat field frames at the beginning and end of
+the night, and ThAr arc lamp spectra for wavelength calibration will
+be taken throughout the night to account for telescope flexure.
+Radial velocity standards will be selected that are of a similar
+spectral type as the targets (e.g., HD122693 -- a dwarf of spectral
+type F8).  Typical equatorial coordinates for the targets are
+($\alpha,\delta$)=(13$^{h}-16^{h}$,-5$^{\circ}- +25^{\circ}$); stars
+in this range will be at airmass $<2$ all night.  The typical $R$-band
+magnitude is 16.  To achieve S/N of 30 per pixel with the proposed
+set-up, an integration time of 1200 seconds is needed for a target
+with $R$=16.  Considering the total amount of observing time of 9
+hours per night and the overhead needed for calibration frames, we
+anticipate observing approximately 40 RR Lyraes.  Reductions will be
+carried out by Sheffield using a well-developed reduction pipeline for
+processing ModSpec data.
 
-\begin{thebibliography}[99]
-\bibitem[Jing 
-\& Suto(2002)]{jing02} Jing, Y.~P., \& Suto, Y.\ 2002, \apj, 574, 538 
-\bibitem[Navarro et al.(1996)]{nfw96} Navarro, J.~F., Frenk, 
-C.~S., \& White, S.~D.~M.\ 1996, \apj, 462, 563 
+\begin{thebibliography}{ }
+\bibitem[Jing \& Suto(2002)]{jing02} Jing, Y.~P., \& Suto, Y.\ 2002, \apj, 574, 538
+\bibitem[Navarro et al.(1996)]{nfw96} Navarro, J.~F., Frenk, C.~S., \& White, S.~D.~M.\ 1996, \apj, 462, 563
+\bibitem[Law \& Majewski(2010)]{law10} Law, D.~R., \& Majewski, S.~R.\ 2010, \apj, 714, 229 
+\bibitem[Johnston et al.(1999)]{johnston99} Johnston, K.~V., Zhao, H., Spergel, D.~N., \& Hernquist, L.\ 1999, \apjl, 512, L109 
+\bibitem[Madore \& Freedman(2012)]{madore12} Madore, B.~F., \& Freedman, W.~L.\ 2012, \apj, 744, 132 
+\bibitem[Stokes et al.(2000)]{linear} Stokes, G.~H., Evans, J.~B., Viggh, H.~E.~M., Shelly, F.~C., \& Pearce, E.~C.\ 2000, \icarus, 148, 21 
+\bibitem[Law et al.(2009)]{ptf} Law, N.~M., Kulkarni, S.~R., Dekany, R.~G., et al.\ 2009, \pasp, 121, 1395 
+\bibitem[Drake et al.(2013)]{catalina} Drake, A.~J., Catelan, M., Djorgovski, S.~G., et al.\ 2013, \apj, 763, 32 
 \end{thebibliography}

### 2013-06-25 text/paper1/hogg_note.tex
start the note

\documentclass[12pt,letterpaper]{article}

\newcommand{\TT}{{\!\mathsf{T}}}

\begin{document}

\section*{to Adrian from Hogg}

Right now, I believe that you do \emph{something} like:
\begin{blockquote}
For each star,
evolve it's phase-space position $X$ (a 6-vector) backwards in time $t$.
Find the time $t_0$ at which it's 6-space distance $\Delta_0$ from the progenitor is minimized.
Compute the empirical variance tensor $V$ by taking the mean outer product $\Delta_0\cdot \Delta_0^\TT$.
Use $\ln\det V$ as the scalar objective by which different potentials are compared.
\end{blockquote}
I am not sure if this summary is correct.
For example, in the code snippet you tweeted, you are subtracting a mean from the 6-displacements.
That implies that perhaps you don't know where the progenitor really is?
But if so, then how do you find the time $t_0$ at which 6-displacement is minimized?

Here's my proposal:

\end{document}

### 2013-06-25 text/paper1/hogg_note.tex
my proposal written tersely

@@
 \newcommand{\TT}{{\!\mathsf{T}}}
+\newcommand{\given}{\,|\,}
 
@@
 Right now, I believe that you do \emph{something} like:
-\begin{blockquote}
+\begin{quote}
 For each star,
@@
 Use $\ln\det V$ as the scalar objective by which different potentials are compared.
-\end{blockquote}
+\end{quote}
 I am not sure if this summary is correct.
@@
 
-Here's my proposal:
+Here's my proposal.
+All I have to give you is a likelihood function%
+---a probability for the data given parameters---%
+and you can take it from there.
+Let me know if you want help with priors or sampling.
+I don't think this model would be hard to instantiate or sample.
+But it \emph{does} have a lot of nuisance parameters (one per star plus some extras).
+
+The progenitor has a 6-space position $X_p(t)$ which is time-dependent and potential-dependent.
+In addition to the potential parameters $\phi$
+there are 6 free parameters $\omega$ in the progenitor initial conditions (or orbit specification).
+In addition, the progenitor has an associated variance tensor $V$,
+which can be a general $6\times 6$ symmetric positive definite tensor or it can be required to be diagonal.
+I would suggest not only making it diagonal but making the spatial part and velocity part each be spherical
+(so it is specified by only two scalars: a spatial scale and a velocity scale).
+
+Each star $i$ has a \emph{perfectly measured} phase-space position $X_i(0)$ right now,
+and (given potential parameters $\phi$) can be integrated back in time to give any 6-position $X_i(t)$ in the past.
+Each star has a latent parameter $t_i$,
+which is the time at which it is considered to have departed the satellite.
+The likelihood is the evaluation of the Gaussian centered on the progenitor for each star at position $X_i(t_i)$;
+that is:
+\begin{eqnarray}
+p(d\given \phi,\omega,V,t_i) &=& N(y\given m,V)
+\\
+d &\equiv& X_i(0)
+\\
+y &\equiv& X_i(t_i)
+\\
+m &\equiv& X_p(t_i)
+\quad ,
+\end{eqnarray}
+where
+$N(y\given m,V)$ is the 6-dimensional Gaussian in $y$ with mean $m$ and variance tensor $V$,
+$d$ is the observed data point (a star's 6-vector position),
+$y$ is the 6-position of the star at past time $t_i$,
+and $m$ is the 6-position of the progenitor at that same time.
+
+[drops the mic]
 

### 2013-06-25 text/paper1/hogg_note.tex
small amount of explanation; new title

@@
 
-\section*{to Adrian from Hogg}
+\section*{to APW and KVJ from Hogg}
 
@@
 
-Here's my proposal.
+\paragraph{Here's my proposal:}
 All I have to give you is a likelihood function%
@@
 
+In my way of thinking about it,
+the parameters $\omega$, $V$, and all of the $t_i$ are nuisance parameters.
+That hurts a bit.
+However, the model will work well even with no serious priors on any of these.
+The reason is that the Gaussian $N(y\given m,V)$ delivers very high likelihood values
+when the data are well explained by small $\det V$.
+That's what you want.
+
+~
+
 [drops the mic]

### 2013-07-15 text/paper1/paper.tex
paper updates

--- 2013-06-25 docs/paper1/paper.tex
@@
 \usepackage{amssymb,amsmath}
+\usepackage{hyperref}
 
@@
 \newcommand{\argmin}{\operatornamewithlimits{argmin}}
+
+\setlength{\jot}{12pt}
 
@@
 
-\section{Proposed potential measure}
+% --------------------------------------------------------------------
+% Above here is from KVJ. Below is APW.
+%
 
-\subsection{Method}
+\section{Context and motivation}
+The method presented in Section \ref{sec:method} takes advantage of
+three distinct developments in the observations of tidal streams: (i)
+the demonstration of a technique for deriving distances to individual
+RR Lyrae stars with 2\% accuracies (Section \ref{sec:spitzer}); (ii)
+the prospect of proper motion measurements of the same stars with
+$\sim$10~mas/yr precision (Section \ref{sec:gaia}); and (iii) the
+tracing of debris associated with Sgr around the entire Galaxy
+(Section \ref{sec:sgr})
 
-Once the full
-phase-space positions of these stars are known there is no need to
-build a sophisticated model of the entire stream system to interpret
-the data.  Instead, the knowledge that these stars were once part of
-Sgr allows them each to be exploited as individual potential measurers.
-If the orbits of an RR Lyrae and Sgr are simultaneously integrated
-backwards in a potential that correctly models the MW, then
-their orbits must intersect at some point within the last few Gyrs; in an
-incorrect potential this will not happen.
-This approach was first described by JZSH in anticipation of the
-SIM survey.
-The original algorithm does not assume that the distances to the debris stars are known,
-and instead estimates them using the expected properties for the debris.
-A Spitzer survey of RR Lyraes would allow us to drop this approximate approach in favor of using the  measured distances.
+\subsection{2\% distance errors for RR Lyrae in the halo}
+\label{sec:spitzer}
+Over twenty-five years ago, \citet{longmore1986} demonstrated the
+existence of a near-infrared period-luminosity (PL) relation for RR
+Lyrae stars using $K$-band observations of stars in Galactic
+globular clusters. More recently, \citet{catelan2004} showed that a PL
+relation for RR Lyrae stars is expected in the infrared (but
+vanishingly so in the optical) and is explained primarily by the
+changing bolometric correction for these stars as a function of the
+observed wavelength. Later, \citet{madore2012} demonstrated the
+inevitability of decreasing intrinsic scatter with increasing
+wavelength (and slope) of PL relations both for Cepheids and for RR
+Lyrae stars. Both reddening effects and metallicity dependencies also
+decrease significantly for observations further into the IR.
 
-At any point during the backwards integration, the proximity of a debris particle to the parent satellite is defined by the normalized phase-space distance,
+These studies suggest that observations of RR Lyrae variables by
+NASA's Spitzer mission could provide us with a stellar distance
+indicator of unprecedented accuracy and precision. \citet{madore2012}
+have demonstrated \citep[using 5 stars with trigonometric parallaxes
+  measured by Hubble;][]{benedict2011} that the dispersion in the
+mid-IR PL relation -- at Spitzer wavelengths -- is of order
+$\pm$0.03~mag. This implies that it is possible to determine
+individual distances that are good to better than $\pm 2\%$ for RR
+Lyrae stars out to $\sim$80~kpc (the limit of Spitzer's ability to
+detect and measure RR Lyrae variables, for modest integration
+times). [Distance measurements of Blue Horizontal Branch and Blue Straggler stars in the halo typically achieve 15-25\% uncertainties.]
+
+\subsection{Gaia and the age of astrometry}
+\label{sec:gaia}
+The Gaia satellite \citep{gaia2001}, planned for launch in October
+2013, is an astrometric mission which, amongst a myriad other
+science goals, aims to measure the positions of billions of stars with
+10-100~$\mu$as accuracies. With over 50 observations of the entire sky
+over the 5-year lifetime of the mission (cite?), Gaia will
+revolutionize the field of Galactic Astronomy by providing sufficient
+proper motion accuracy to create full six-dimensional phase-space
+maps of the Galaxy for heliocentric distances of up to $\sim$ 10~kpc (xx kpc)
+for $V\sim xxx$ ($V\sim yyy$) stars.
+
+Figure~\ref{fig:gaia_errors} show the Gaia end-of-mission distance and
+tangential velocity error estimates for a typical halo RR Lyrae star
+\citep[{$M_V=0.45$, [Fe/H]$=-1.5$};][]{ivezic2008}. Within 2~kpc, Gaia
+will be able to measure distances to these stars with better than 2\%
+accuracy: RR Lyraes in this volume will be used to test and calibrate
+the Spitzer PL relation described above. Beyond the 2~kpc threshold,
+the mid-IR PL relation for RR Lyraes will provide more accurate
+distance measurements.
+
+The anticipated proper motion errors for RR Lyrae correspond to
+tangential velocity errors of order $\sim$1~km/s at 30~kpc, and of
+order $\sim$10~km/s out to several tens of kiloparsecs. The
+combination of Spitzer and Gaia data will extend the `horizon' of
+where full six-dimensional phase-space maps of the Galaxy are possible
+from $<10$~kpc to 80~kpc. This enormous increase in volume will refine
+data on existing debris systems such as Sagittarius, the Orphan
+stream, and [...], and enable discovery of new substructure within the
+Galactic halo.
+
+\subsection{The Sagittarius debris system}
+\label{sec:sgr}
+The Sagittarius Dwarf Galaxy (hereafter Sgr) was discovered
+serendipitously during a radial velocity survey of the Galactic bulge
+\citep{ibata1994}. While the number of known Milky Way satellites has
+more than doubled recently \citep[e.g.,][]{mcconnachie2012}, in many
+ways Sgr remains the most studied and most intriguing: it is the
+closest satellite, among the largest (only the LMC and SMC are more
+luminous), and has the most prominent tidal tails.
+
+The first maps of the Sgr core \citep{ibata1994, ibata1995} revealed
+its elongated morphology and suggested tidal deformation, as would be
+anticipated from its proximity to the Galaxy. Subsequent N-body models
+confirmed this interpretation \citep{velazquez1995} and showed that
+debris from Sgr should form coherent streams of stars that encircle
+the Galaxy \citep{johnston1995}. Signatures of these extensive stellar
+streams have since been discovered in many studies and have been
+mapped across the sky in carbon stars \citep{totten1998}, M giants
+selected from the Two Micron All Sky Survey \citep{majewski2003}, main
+sequence turnoff stars from the Sloan Digital Sky Survey
+\citep{belokurov2006}, and RR Lyrae in the Catalina Sky Survey
+\citep{drake2013}.
+
+The distance to the stream has been estimated from photometric surveys
+\citep[e.g.,][]{martinezdelgado2004} and gradients in radial
+velocities and metallicity measured with follow-up spectroscopy
+\citep[e.g.,][]{majewski2004, vivas2005, bellazzini2006, chou2007,
+  chou2010, keller2010, carlin2012}. The multitude of Sgr stream data
+has inspired a rich set of interpretations and models of the debris
+\citep[e.g.,][]{johnston1999, helmi2004, law2005, fellhauer2006,
+  law2010}.
+
+Most recently, \citet[][hereafter LM10]{law2010} have combined all
+current data on the Sgr debris to assess the triaxiality and orientation
+of the outer Galaxy, the first time that such a reconstruction of
+the 3-dimensional mass distribution of a dark matter halo has been
+feasible. Figure \ref{fig:lm10} illustrates results from the LM10
+N-body simulation of dwarf satellite disruption along the expected Sgr
+orbit. The particle positions are shown in a coordinate system defined
+by the orbit of the satellite orbit \citep{majewski2004}. The
+simulation was run in a three-component potential, with a triaxial,
+logarithmic halo model of the form
 \begin{equation}
-       D_{\rm ps}=\left[\left({d \over r_{\rm tide}}\right)^2+\left({v \over v_{\rm esc}}\right)^2\right],
+  \Phi_{halo} = v_{halo}^2 \ln(C_1 x^2 + C_2 y^2 + C_3 xy + (z/q_z)^2 + R_c^2).
 \end{equation}
-where $d$ and $v$ are the distance and speed separation between the star
-and the parent respectively.  The one assumption in this method is what value to adopt for the satellite mass, $m_{\rm sat}$,which sets the instantaneous tidal radius,
-$r_{\rm tide}=
-R (m_{\rm sat}/M_{\rm Gal})^{1/3}$ and escape velocity, $v_{\rm esc} =
-G m_{\rm sat}/r_{\rm tide}$. The mass of the
-Galaxy $M_{\rm Gal}$ enclosed within radius $R$ is calculated directly from the
-assumed potential which is being tested.  The ``best'' potential parameters are those that
-minimize the average $D_{\rm ps}$ for all stars.
-{\it Or we could come up with some other statistic.}
+where $C_1$, $C_2$, and $C_3$ are combinations of the $x$ and $y$ flattening ($q_1$, $q_2$) and orientation of the halo with respect to the baryonic disk:
+\begin{align}
+  C_1 &= \frac{\cos^2\phi}{q_1^2} + \frac{\sin^2\phi}{q_2^2}\\
+  C_2 &= \frac{\sin^2\phi}{q_1^2} + \frac{\cos^2\phi}{q_2^2}\\
+  C_3 &= 2\sin\phi\cos\phi \left(q_1^{-2} - q_2^{-2}\right)
+\end{align}
 
-\subsection{Test of method on synthetic Sgr data}
+LM10 demonstrate how a comparison of such simulations with the
+observed debris enables a detailed reconstruction of Sgr's history 
+(its mass and stellar populations, orbit, rate of destruction, and
+even original distribution in populations [other refs??]). The phase-space distribution of the debris offers a unique probe of
+the depth, shape, extent, and orientation of the Milky Way's dark
+matter halo that cannot be matched using other techniques on our own
+or other galaxies \citep[e.g.,][]{ibata2001}. The LM10 constraints on
+Sgr's mass and orbit have invigorated discussions of Sgr in a more
+cosmological context, the mass and extent of the original dark matter
+halo that hosted it, as well as its effect on the Milky Way itself
+\citep{bailin2003,purcell2011,micheldansac2011,gomez2012}.
 
-Figure \ref{boots.fig} illustrates potential power of the method for
-particles from the Sgr disruption N-body simulation
- chosen to represent our proposed
-RR Lyrae survey (yellow crosses in left panel).  The 53 particles within 35 kpc of
-the Sun were ``observed'' with appropriate distance (assuming 2\%
-accuracy), proper motion (assuming Gaia accuracies)\footnote{see
- http://www.rssd.esa.int/index.php?project=GAIA} and radial velocity
-(assuming 5 km/s) errors.  The ``observed'' phase-space positions were
-then integrated backwards in model potentials with axis ratios
-($q_1/q_2$ and $q_z/q_2$) and orientation ($\phi$) of the dark matter
-component systematically varied.  Sgr's phase space coordinates
-%(whose position and motion were
-(assumed to be exactly known, along with the distance to the Galactic
-center) were simultaneously integrated backwards in the same
-potentials, and the minimum normalized phase-space distance, $D_{\rm
- ps}$, between Sgr and each particle recorded.
-he ``best'' potential parameters were those that
-minimized the average $D_{\rm ps}$ for all stars.
+The combined Spitzer and Gaia measurements of distances and proper
+motions of RR Lyraes in the Sgr debris opens up new avenues of
+interpretation. A 2\% distance error at $\sim$30~kpc is smaller than
+the predicted thickness of the stream, ranging from
+$\sim$2-$10$~kpc. Combined with Gaia tangential velocity measurement
+accuracies ($\sim$2~km/s at 30~kpc) less than the intrinsic dispersion
+for much of the stream \citep[$\sigma_v\sim$10~km/s;][]{majewski2004}, and ground-based radial
+velocities with $\sim$10~km/s accuracy, the measured phase-space
+position of each individual RR Lyrae relative to the stream becomes a
+powerful probe of the potential: the next section outlines a new
+method to take advantage of this information.
 
-Figure
-\ref{boots.fig} contours confidence levels for the recovered parameters
-(with a bold cross marking the true values used in the original
-simulation), with levels estimated by repeating the analysis using
-samples bootstrapped from our ``observations''.  The contours
-demonstrate the power of this combination of Spitzer and Gaia data:
-This small sample alone can broadly rule out significant areas of
-parameter space.  Combining the constraints from these samples with
-the extensive ground-based data sets already available would allow
-quite clear measurements of the shape and strength of the Milky Way
-potential.
+\section{Method}
+\label{sec:method}
+With access to full phase-space information for stars in a tidal
+stream, it is no longer necessary to model the orbit of the stream
+itself. Instead, it is possible to use the individual stars as
+potential measures by exploting the fact that the stars must have come
+from the same progenitor. If the present-day orbital parameters of the
+progenitor system are known, this places an intuitive constraint on
+the potential of the Milky Way: if the orbits of the stars and the
+orbit of the progenitor are integrated \emph{backwards} in time, the
+stars should `recombine' with the progenitor. If the potential is
+incorrect, the orbits of the stars will diverge from that of the
+progenitor and thus will not be recaptured by the satellite system.
 
-%% \begin{figure}
-%% \plotone{figs/boots.eps}
-%% \caption{
-%% \label{boots.fig}
-%% }
-%% \end{figure}
+Quantifying this method requires a sample of stars with known full
+space kinematics $(\bs{x}_i, \bs{v}_i)|_{t=0}$ (e.g., measurements of all position and velocity components for these stars \emph{today} at $t=0$), the orbital parameters
+for the progenitor system $(\bs{x}_p, \bs{v}_p)|_{t=0}$, and a
+functional form for the potential, $\Phi$.\footnote{Future work will
+  investigate a more general ... basis function blah blah.} For a
+given set of potential parameters, the orbits of the stars and
+progenitor are integrated backwards for several Gigayears (determined from an
+estimate of the interaction time of the system). The stars and
+progenitor system are treated as test particles, so that the debris
+stars do not become re-bound to the satellite. Instead, at each
+timestep $t_j$, for each particle $i$, a set of normalized, relative
+phase-space coordinates, $(\bs{q}_i(t_j), \bs{p}_i(t_j))$, are
+computed using the instantaneous tidal radius, escape velocity, and
+phase-space position of the progenitor. At a particular timestep,
+these coordinates are defined as
+\begin{equation}
+  \bs{q}_i = \frac{\bs{x}_i -
+    \bs{x}_p}{R_{tide}}\,\,\,\,\,\,\,\,\bs{p}_i = \frac{\bs{v}_i -
+    \bs{v}_p}{v_{esc}}
+\end{equation}
+where $\bs{x}_i$ and $\bs{v}_i$ are the 3D position and velocity of
+the $i$th particle, and $\bs{x}_p$ and $\bs{v}_p$ are the same for the
+progenitor.  These definitions require an estimate of the mass of the
+satellite, $m_{sat}$, which, combined with the computed enclosed mass
+of the potential, $m_{enc}$, sets the instantaneous tidal radius and
+escape velocity,
+\begin{equation}
+  R_{tide}=\Big(\frac{m_{sat}}{3m_{enc}}\Big)^{1/3}\,\,\,\,\,\,\,\,
+  v_{esc}=\sqrt{\frac{2Gm_{sat}}{R_{tide}}}.
+\end{equation}
+
+Qualitatively, when the distance in this normalized 6D space,
+$D_{ps}=\sqrt{|\bs{q}_i|^2+|\bs{p}_i|^2} < \sqrt{2}$, the star is
+probably recaptured by the satellite, but this is not imposed as a
+hard constraint.  As a statistical measure of how many stars have recombined, we
+construct the 6D distribution of the minimum phase-space vectors for
+each star over each individual, integrated orbit. For each star, $i$,
+the phase-space distance, $D_{ps}$, is computed at each timestep
+$t_j$, and the vector with the minimum phase-space distance is stored
+\begin{align}
+  t^*_i &= \argmin_{t} D_{ps,i}\\
+  A_{ik} &= (\bs{q}_i(t^*_i),\bs{p}_i(t^*_i)) % A_ik are really samples from a distribution, or we treat them as such
+\end{align}
+where $k\in\{1,6\}$. Intuitively, the variance of this distribution
+will be larger for orbits integrated in an incorrect potential
+relative to the distribution computed from the `true' orbital history
+of the stars; in an incorrect potential, small differences in the
+orbits of the stars relative to the orbit of the progenitor cause them
+to spread out in this relative, normalized phase space. Thus, the
+\emph{generalized variance} of the distribution, computed for a given
+set of potential parameters, $\bs{\theta}$, is a natural choice for
+the scalar objective function used in constraining the potential of
+the Milky Way,
+\begin{equation}
+  f(\bs{\theta})=\det(\mathrm{Cov}(A_{ik})).
+\end{equation}
+[This objective function is an approximation to a more general likelihood function, described in detail in forthcoming work (Price-Whelan \& Johnston in prep.).
+The results shown in sections xx and yy are meant to illustrate the strength and concept of the method..]
+
+\subsection{Application to Simulated Data}
+The LM10 (see Section \ref{sec:sgr}) simulation data is a perfect
+test-bed for evaluating the effectiveness of this method. We start by
+extracting both particle data and the satellite orbital parameters
+from the present-day snapshot of the simulation
+data.\footnote{\url{www.astro.virginia.edu/~srm4n/Sgr/data.html}} We then
+`observe' a sample of 100 stars from the first leading and trailing
+wraps of the stream: the radial velocity errors are drawn from a
+Gaussian $\epsilon_{rv} \sim \mathcal{N}(\mu=0,\sigma=10~{\rm km/s})$,
+the tangential velocity errors depend on the distance to the star and
+are computed from the expected Gaia error curve (shown in
+Figure~\ref{fig:gaia}), and the distance errors on a star of distance
+$D$ are $\epsilon_{dist} \sim \mathcal{N}(0,0.02\times
+D)$. Figure~\ref{fig:} shows all 2D projections of the 6D minimum
+phase-space distance distribution for the observed particles in the
+correct and an incorrect potential. Qualitatively, it is clear that the
+variance of the distribution increases in the wrong potential. The objective function
+(generalized variance) defines a convex hypersurface over which we
+optimize four of the six logarithmic potential parameters:
+$v_{circ}$, $\phi$, $q_1$, and $q_z$. Figure~\ref{fig:} shows [...]
+curves produced by varying the true parameters of the potential by
+10\% around the true value.
+
+% Discussion?
+Many orbit fitting methods for measuring the Galactic potential assume that the stars in a given stream are on the same
+orbit. \cite{binney??} show that this assumption leads to biased
+potential parameters, even without realistic observational errors on
+the kinematic information. The method presented above is
+free of this assumption and instead \emph{relies} on the fact that
+each star is on a different orbit and is thus an independent potentiometer. The power of this method is in
+the joint constraint that the best potential is the one that gets the
+most of these individual orbits to intersect the progenitor
+over the history of interaction with the system.
+
+% APW Note: The optimization, in a way, is getting a bunch of orbits (curves) to intersect another curve (the progenitor orbit) at some points (over time).
 
@@
 
-% --------------------------------------------------------------------
-% Above here is from KVJ. Below is APW.
-%
-
-\section{Context and motivation}
-The method presented in Section \ref{sec:method} takes advantage of
-three distinct developments in the observations of tidal streams: (i)
-the demonstration of a technique for deriving distances to individual
-RR Lyrae stars with 2\% accuracies (Section \ref{sec:spitzer}); (ii)
-the prospect of proper motion measurements of the same stars with
-$\sim$10~mas/yr precision (Section \ref{sec:gaia}); and (iii) the
-tracing of debris associated with Sgr around the entire Galaxy
-(Section \ref{sec:sgr})
-
-\subsection{2\% distance errors for RR Lyrae in the halo}
-\label{sec:spitzer}
-Over twenty-five years ago, \citet{longmore1986} demonstrated the
-existence of a near-infrared period-luminosity (PL) relation for RR
-Lyrae stars using $K$-band observations of these stars in Galactic
-globular clusters. More recently, \citet{catelan2004} showed that a PL
-relation for RR Lyrae stars is expected in the infrared -- but
-vanishingly so in the optical -- and is explained primarily by the
-changing bolometric correction for these stars as a function of the
-observed wavelength. Later, \citet{madore2012} demonstrated the
-inevitability of decreasing intrinsic scatter with increasing
-wavelength (and slope) of PL relations both for Cepheids and for RR
-Lyrae stars. Both reddening effects and metallicity dependencies also
-decrease significantly for observations further into the IR.
-
-These studies suggest that observations of RR Lyrae variables by
-NASA's Spitzer mission could provide us with a stellar distance
-indicator of unprecedented accuracy and precision. \citet{madore2012}
-have demonstrated \citep[using 5 stars with trigonometric parallaxes
-  measured by Hubble;][]{benedict2011} that the dispersion in the
-mid-IR PL relation--at Spitzer wavelengths--is of order
-$\pm$0.03~mag. This implies that it is possible to determine
-individual distances that are good to better than $\pm 2\%$ for RR
-Lyrae stars out to $\sim$80~kpc (the limit of Spitzer's ability to
-detect and measure RR Lyrae variables, for modest integration
-times).
-
-\subsection{Gaia and the age of astrometry}
-\label{sec:gaia}
-The Gaia satellite \citep{gaia2001}, planned for launch in October
-2013, is an astrometric mission which, amongst a myriad of other
-science goals, aims to measure the positions of billions of stars with
-10-100~$\mu$as accuracies. With over 50 observations of the entire sky
-over the 5-year lifetime of the mission (cite?), Gaia will
-revolutionize the field of Galactic Astronomy by providing sufficient
-proper motion accuracy to create full, six-dimensional phase-space
-maps of the Galaxy for heliocentric distances of up to $\sim$ 10~kpc (xx kpc)
-for $V\sim xxx$ ($V\sim yyy$) stars.
-
-Figure~\ref{fig:gaia_errors} show the Gaia end-of-mission distance and
-tangential velocity error estimates for a typical halo RR Lyrae star
-\citep[{$M_V=0.45$, [Fe/H]$=-1.5$};][]{ivezic2008}. Within 2~kpc, Gaia
-will be able to measure distances to these stars with better than 2\%
-accuracy: RR Lyraes in this volume will be used to test and calibrate
-the Spitzer PL relation described above. Beyond the 2~kpc threshold,
-the mid-IR PL relation for RR Lyraes will provide more accurate
-distance estimates than Gaia.
-
-The anticipated proper motion errors for RR Lyrae correspond to
-tangential velocity errors of order $\sim$1~km/s at 30~kpc, and of
-order $\sim$10~km/s out to several tens of kiloparsecs. The
-combination of Spitzer and Gaia data will extend the `horizon' of
-where full six-dimensional phase-space maps of the Galaxy are possible
-from $<10$~kpc to 80~kpc. This enormous increase in volume will refine
-data on existing debris systems such as Sagittarius, the Orphan
-stream, and [...], and enable discovery of new substructure within the
-Galactic halo.
-
-\subsection{The Sagittarius debris system}
-\label{sec:sgr}
-The Sagittarius Dwarf Galaxy (hereafter Sgr) was discovered
-serendipitously during a radial velocity survey of the Galactic bulge
-\citep{ibata1994}. While the number of known Milky Way satellites has
-more than doubled recently \citep[e.g.,][]{mcconnachie2012}, in many
-ways Sgr remains the most studied and most intriguing: it is the
-closest satellite, among the largest (only the LMC and SMC are more
-luminous), and has the most prominent tidal tails.
-
-The first maps of the Sgr core \citep{ibata1994, ibata1995} revealed
-its elongated morphology and suggested tidal deformation, as would be
-anticipated from its proximity to the Galaxy. Subsequent N-body models
-confirmed this interpretation \citep{velazquez1995} and showed that
-debris from Sgr should form coherent streams of stars that encircle
-the Galaxy \citep{johnston1995}. Signatures of these [..] stellar
-streams have since been discovered in many studies and have been
-mapped across the sky in carbon stars \citep{totten1998}, M giants
-selected from the Two Micron All Sky Survey \citep{majewski2003}, main
-sequence turnoff stars from the Sloan Digital Sky Survey
-\citep{belokurov2006}, and RR Lyrae in the Catalina Sky Survey
-\citep{drake2013}.
-
-The distance of the stream has been estimated from photometric surveys
-\citep[e.g.,][]{martinezdelgado2004} and gradients in radial
-velocities and metallicity measured with follow-up spectroscopy
-\citep[e.g.,][]{majewski2004, vivas2005, bellazzini2006, chou2007,
-  chou2010, keller2010, carlin2012}. The multitude of Sgr stream data
-has inspired a rich set of interpretations and models of the debris
-\citep[e.g.,][]{johnston1999, helmi2004, law2005, fellhauer2006,
-  law2010}.
-
-Most recently, \citet[][hereafter LM10]{law2010} have combined all
-current data on Sgr debris to assess the triaxiality and orientation
-of the outer Galaxy --- the first time that such a reconstruction of
-the 3-dimensional mass distribution of a dark matter halo has been
-feasible. Figure \ref{fig:lm10} illustrates results from the LM10
-N-body simulation of dwarf satellite disruption along the expected Sgr
-orbit. The particle positions are shown in a coordinate system defined
-by the orbit of the satellite orbit \citep{majewski2004}. The
-simulation was run in a three-component potential, with a triaxial,
-logarithmic halo model of the form
-\begin{equation}
-  \Phi_{halo} = v^2 \ln(C_1 x^2 + C_2 y^2 + C_3 xy + (z/q_z)^2 + R_c^2).
-\end{equation}
-
-LM10 demonstrate how a comparison of such simulations with the
-observed debris enables a detailed reconstruction of Sgr's history ---
-it's mass and stellar populations, orbit, rate of destruction, and
-even original distribution in populations (other refs??). Moreover,
-the phase-space distribution of the debris offers a unique probe of
-the depth, shape, extent, and orientation of the Milky Way's dark
-matter halo that cannot be matched using other techniques on our own
-or other galaxies \citep[e.g.,][]{ibata2001}. The LM10 constraints on
-Sgr's mass and orbit have invigorated discussions of Sgr in a more
-cosmological context, the mass and extent of the original dark matter
-halo that hosted it, as well as its effect on the Milky Way itself
-\citep{bailin2003,purcell2011,micheldansac2011,gomez2012}.
-
-The combined Spitzer and Gaia measurements of distances and proper
-motions of RR Lyraes in the Sgr debris opens up new avenues of
-interpretation. A 2\% distance error at $\sim$30~kpc is smaller than
-the predicted thickness of the stream, ranging from
-$\sim2-10$~kpc. Combined with Gaia tangential velocity measurement
-accuracies ($\sim$2~km/s at 30~kpc) less than the intrinsic dispersion
-for much of the stream ($\sigma_v\sim ??$), and ground-based radial
-velocities with $\sim10$~km/s accuracy, the measured phase-space
-position of each individual RR Lyrae relative to the stream becomes a
-powerful probe of the potential: the next section outlines a new
-method to take advantage of this information and use each star as an
-individual potential measure.
-
-\section{Method}
-\label{sec:method}
-With access to full phase-space information for stars in a tidal
-stream, it is no longer necessary to model the orbit of the stream
-itself. Instead, it is possible to use the individual stars as
-potential measures by exploting the fact that the stars must have come
-from the same progenitor. If the present-day orbital parameters of the
-progenitor system are known, this places an intuitive constraint on
-the potential of the Milky Way: if the orbits of the stars and the
-orbit of the progenitor are integrated \emph{backwards} in time, the
-stars should `recombine' with the progenitor. If the potential is
-incorrect, the orbits of the stars will diverge from that of the
-progenitor and thus will not be recaptured by the satellite system.
-
-Quantifying this method requires a sample of stars with known full
-space kinematics $(\bs{x}_i, \bs{v}_i)|_{t=0}$, the orbital parameters
-for the progenitor system $(\bs{x}_p, \bs{v}_p)|_{t=0}$, and a
-functional form for the potential, $\Phi$.\footnote{Future work will
-  investigate a more general ... basis function blah blah.} For a
-given set of potential parameters, the orbits of the stars and
-progenitor are integrated backwards for some time (determined from an
-estimate of the interaction time of the system). The stars and
-progenitor system are treated as test particles, so that the debris
-stars do not become re-bound to the satellite. Instead, at each
-timestep $t_j$, for each particle $i$, a set of normalized, relative
-phase-space coordinates, $(\bs{q}_i(t_j), \bs{p}_i(t_j))$, are
-computed using the instantaneous tidal radius, escape velocity, and
-phase-space position of the progenitor. At a particular timestep,
-these coordinates are defined as
-\begin{equation}
-  \bs{q}_i = \frac{\bs{x}_i -
-    \bs{x}_p}{R_{tide}}\,\,\,\,\,\,\,\,\bs{p}_i = \frac{\bs{v}_i -
-    \bs{v}_p}{v_{esc}}
-\end{equation}
-where $\bs{x}_i$ and $\bs{v}_i$ are the 3D position and velocity of
-the $i$th particle, and $\bs{x}_p$ and $\bs{v}_p$ are the same for the
-satellite.  These definitions require an estimate of the mass of the
-satellite, $m_{sat}$, which, combined with the computed enclosed mass
-of the potential, $m_{enc}$, sets the instantaneous tidal radius and
-escape velocity,
-\begin{equation}
-  R_{tide}=\Big(\frac{m_{sat}}{3m_{enc}}\Big)^{1/3}\,\,\,\,\,\,\,\,
-  v_{esc}=\sqrt{\frac{2Gm_{sat}}{R_{tide}}}.
-\end{equation}
-
-Qualitatively, when the distance in this normalized 6D space,
-$D_{ps}=\sqrt{|\bs{q}_i|^2+|\bs{p}_i|^2} < \sqrt{2}$, the star is
-probably recaptured by the satellite (but this is not imposed as a
-hard constraint). As a statistical measure over all stars, we
-construct a 6D distribution of the minimum phase-space vectors for
-each star over the individual, integrated orbits. For each star, $i$,
-the phase-space distance, $D_{ps}$, is computed at each timestep
-$t_j$, and the vector with the minimum phase-space distance is stored
-\begin{align}
-  t^*_i &= \argmin_{t} D_{ps,i}\\
-  A_{ik} &= (\bs{q}_i(t^*_i),\bs{p}_i(t^*_i)) % A_ik are really samples from a distribution, or we treat them as such
-\end{align}
-where $k\in\{1,6\}$. Intuitively, the variance of this distribution
-will be larger for orbits integrated in an incorrect potential
-relative to the distribution computed from the `true' orbital history
-of the stars; in an incorrect potential, small differences in the
-orbits of the stars relative to the orbit of the progenitor cause them
-to spread out in this relative, normalized phase space. Thus, the
-\emph{generalized variance} of the distribution, computed for a given
-set of potential parameters, $\bs{\theta}$, is a natural choice for
-the scalar objective function used in constraining the potential of
-the Milky Way,
-\begin{equation}
-  f(\bs{\theta})=\det(\mathrm{Cov}(A_{ik})).
-\end{equation}
-
-\subsection{Application to Simulated Data}
-The LM10 (see Section \ref{sec:sgr}) simulation data is a perfect
-test-bed for evaluating the effectiveness of this method. We start by
-extracting both particle data and the satellite orbital parameters
-from the present-day snapshot of the simulation
-data.\footnote{www.astro.virginia.edu/~srm4n/Sgr/data.html} We then
-`observe' a sample of 100 stars from the first leading and trailing
-wraps of the stream: the radial velocity errors are drawn from a
-Gaussian $\epsilon_{rv} \sim \mathcal{N}(\mu=0,\sigma=10~{\rm km/s})$,
-the tangential velocity errors depend on the distance to the star and
-are computed from the expected Gaia error curve (shown in
-Figure~\ref{fig:gaia}), and the distance errors on a star of distance
-$D$ are $\epsilon_{dist} \sim \mathcal{N}(0,0.02\times
-D)$. Figure~\ref{fig:} shows all 2D projections of the 6D minimum
-phase-space distance distribution for the observed particles in the
-correct and an incorrect potential. Visually, it is clear that the
-variance of the distribution increases in the wrong potential. The
-generalized variance defines a convex hypersurface over which we
-optimize four of the six logarithmic potential parameters:
-$v_{circ}$, $\phi$, $q_1$, and $q_z$. Figure~\ref{fig:} shows [...]
-curves produced by varying the true parameters of the potential by
-10\% around the true value.
-
-% Discussion?
-One of the assumptions of orbit fitting methods for measuring the
-potential is that the stars in a given stream are on the same
-orbit. \citep{binney??} show that this assumption leads to biased
-potential parameters, even without realistic observational errors on
-the kinematic information. Conversely, the method presented above is
-free of this assumption and instead \emph{relies} on the fact that
-each star is on an independent orbit. The power of this method is in
-the joint constraint that the best potential is the one that gets the
-most of these individual orbits to cross the orbit of the progenitor
-over the history of the system.
-
-% APW Note: The optimization, in a way, is getting a bunch of orbits (curves) to intersect another curve (the progenitor orbit) at some points (over time).
-
 \section{Summary}

### 2013-07-16 scripts/paper1_figures.py
playing with figure code

@@ -85,7 +85,11 @@ def gaia_spitzer_errors():
                 'lines.color' : 'k',
                 'lines.marker' : None,
                 'text.usetex' : True,
-                'axes.edgecolor' : '#444444'}
+                'axes.edgecolor' : '#444444',
+                'axes.facecolor' : '#ffffff'}
+    
+    sgr_color = '#67A9CF'
+    orp_color = '#EF8A62'
     
     with rc_context(rc=rcparams):
         fig,axes = plt.subplots(2, 1, figsize=(8, 10), sharex=True)
@@ -125,20 +129,20 @@ def gaia_spitzer_errors():
     
     # Now add rectangles for Sgr, Orphan
     sgr_d = Rectangle((10., 0.15), 60., 0.15, 
-                      color='#67A9CF', alpha=0.75, label='Sgr thickness')
+                      color=sgr_color, alpha=1., label='Sgr thickness')
     axes[0].add_patch(sgr_d)
     
     # From fig. 3 in http://mnras.oxfordjournals.org/content/389/3/1391.full.pdf+html
     orp_d = Rectangle((10., 0.03), 35., 0.03,
-                      color='#EF8A62', alpha=0.75, label='Orp thickness')
+                      color=orp_color, alpha=1., label='Orp thickness')
     axes[0].add_patch(orp_d)
     
-    # ??
-    sgr_v = Rectangle((10., 12), 60., 1., color='#67A9CF', alpha=0.75,
+    # Dispersion from Majewski 2004: 10 km/s
+    sgr_v = Rectangle((10., 10), 60., 1., color=sgr_color, alpha=0.75,
                       label='Sgr dispersion')
     axes[1].add_patch(sgr_v)
     
-    orp_v = Rectangle((10., 10.), 35., 1., color='#EF8A62', alpha=0.75,
+    orp_v = Rectangle((10., 10.), 35., 1., color=orp_color, alpha=0.75,
                       label='Orp dispersion')
     axes[1].add_patch(orp_v)
     
@@ -153,12 +157,14 @@ def gaia_spitzer_errors():
     axes[0].set_yticklabels(["{:g}".format(yt) for yt in axes[0].get_yticks()])
     axes[1].set_yticklabels(["{:g}".format(yt) for yt in axes[1].get_yticks()])
     
-    # add Gaia and Spitzer to first plot
+    # add Gaia and Spitzer text to first plot
[…]

### 2013-07-17 text/paper1/paper.tex
text changes
some updates from KVJ comments

@@
 
-\author{Johnston, Price-Whelan, Madore, Majewski}
+\author{Adrian~M.~Price-Whelan\altaffilmark{1,2} Kathryn V. Johnston\altaffilmark{1}}
+
+% affiliations
+\altaffiltext{1}{Department of Astronomy, Columbia University, 550 W 120th St., New York, NY 10027, USA}
+\altaffiltext{2}{NSF Graduate Fellow}
 
 \begin{abstract}
-sdfsdf
+.
 \end{abstract}
@@
 detect and measure RR Lyrae variables, for modest integration
-times). [Distance measurements of Blue Horizontal Branch and Blue Straggler stars in the halo typically achieve 15-25\% uncertainties.]
+times). Distance measurements of Blue Horizontal Branch and Blue Straggler stars at these distances in the halo typically achieve 15-25\% uncertainties. [can cite Alis Deason if necessary].
 
@@
 The Gaia satellite \citep{gaia2001}, planned for launch in October
-2013, is an astrometric mission which, amongst a myriad other
-science goals, aims to measure the positions of billions of stars with
+2013, is an astrometric mission which aims to measure the positions of billions of stars with
 10-100~$\mu$as accuracies. With over 50 observations of the entire sky
-over the 5-year lifetime of the mission (cite?), Gaia will
+over the 5-year lifetime of the mission, Gaia will
 revolutionize the field of Galactic Astronomy by providing sufficient
 proper motion accuracy to create full six-dimensional phase-space
-maps of the Galaxy for heliocentric distances of up to $\sim$ 10~kpc (xx kpc)
-for $V\sim xxx$ ($V\sim yyy$) stars.
+maps of the Galaxy with $<$10\% distance errors for heliocentric distances of up to $\sim$10~kpc for ...
 
@@
 from $<10$~kpc to 80~kpc. This enormous increase in volume will refine
-data on existing debris systems such as Sagittarius, the Orphan
-stream, and [...], and enable discovery of new substructure within the
-Galactic halo.
+data on existing debris systems such as the Sagittarius stream.
 
@@
 \end{equation}
-where $C_1$, $C_2$, and $C_3$ are combinations of the $x$ and $y$ flattening ($q_1$, $q_2$) and orientation of the halo with respect to the baryonic disk:
+where $C_1$, $C_2$, and $C_3$ are combinations of the $x$ and $y$ flattening ($q_1$, $q_2$) and orientation of the halo with respect to the baryonic disk ($\phi$):
 \begin{align}
@@
 position of each individual RR Lyrae relative to the stream becomes a
-powerful probe of the potential: the next section outlines a new
+powerful probe of the potential (see \ref{fig:}): the next section outlines a new
 method to take advantage of this information.
 
-\section{Method}
+\section{Description and test of our algorithm}
 \label{sec:method}
@@
 
+\subsection{The algorithm: Rewinder}
 Quantifying this method requires a sample of stars with known full
-space kinematics $(\bs{x}_i, \bs{v}_i)|_{t=0}$ (e.g., measurements of all position and velocity components for these stars \emph{today} at $t=0$), the orbital parameters
+space kinematics $(\bs{x}_{i}, \bs{v}_{i})|_{t=0}$ (e.g., measurements of all position and velocity components for these stars \emph{today} at $t=0$), the orbital parameters
 for the progenitor system $(\bs{x}_p, \bs{v}_p)|_{t=0}$, and a
-functional form for the potential, $\Phi$.\footnote{Future work will
-  investigate a more general ... basis function blah blah.} For a
+functional form for the potential, $\Phi$. For a
 given set of potential parameters, the orbits of the stars and
@@
 timestep $t_j$, for each particle $i$, a set of normalized, relative
-phase-space coordinates, $(\bs{q}_i(t_j), \bs{p}_i(t_j))$, are
-computed using the instantaneous tidal radius, escape velocity, and
-phase-space position of the progenitor. At a particular timestep,
-these coordinates are defined as
+phase-space coordinates are
+computed 
 \begin{equation}
-  \bs{q}_i = \frac{\bs{x}_i -
-    \bs{x}_p}{R_{tide}}\,\,\,\,\,\,\,\,\bs{p}_i = \frac{\bs{v}_i -
-    \bs{v}_p}{v_{esc}}
+  \bs{q}_{i} = \frac{\bs{x}_{i} -
+    \bs{x}_{p}}{R_{\rm tide}}\,\,\,\,\,\,\,\,\bs{p}_{i} = \frac{\bs{v}_{i} -
+    \bs{v}_{p}}{v_{\rm esc}}
 \end{equation}
-where $\bs{x}_i$ and $\bs{v}_i$ are the 3D position and velocity of
-the $i$th particle, and $\bs{x}_p$ and $\bs{v}_p$ are the same for the
-progenitor.  These definitions require an estimate of the mass of the
+where $(\bs{x},\bs{v})_{i}$ and $(\bs{x},\bs{v})_{p}$ are the phase-space coordinates for the particles and progenitor, respectively. These definitions require an estimate of the mass of the
 satellite, $m_{sat}$, which, combined with the computed enclosed mass
@@
 \begin{equation}
-  R_{tide}=\Big(\frac{m_{sat}}{3m_{enc}}\Big)^{1/3}\,\,\,\,\,\,\,\,
-  v_{esc}=\sqrt{\frac{2Gm_{sat}}{R_{tide}}}.
+  R_{\rm tide}=\Big(\frac{m_{\rm sat}}{3m_{\rm enc}}\Big)^{1/3}\,\,\,\,\,\,\,\,
+  v_{\rm esc}=\sqrt{\frac{2Gm_{\rm sat}}{R_{\rm tide}}}.
 \end{equation}
@@
 Qualitatively, when the distance in this normalized 6D space,
-$D_{ps}=\sqrt{|\bs{q}_i|^2+|\bs{p}_i|^2} < \sqrt{2}$, the star is
+$D_{\rm ps}=\sqrt{|\bs{q}_{i}|^2+|\bs{p}_{i}|^2} < \sqrt{2}$, the star is
 probably recaptured by the satellite, but this is not imposed as a
-hard constraint.  As a statistical measure of how many stars have recombined, we
+hard constraint. [JZSH ...] As a statistical measure of how many stars have recombined, we
 construct the 6D distribution of the minimum phase-space vectors for
 each star over each individual, integrated orbit. For each star, $i$,
-the phase-space distance, $D_{ps}$, is computed at each timestep
-$t_j$, and the vector with the minimum phase-space distance is stored
+the phase-space distance, $D_{\rm ps}$, is computed at each timestep
+$t_{j}$, and the vector with the minimum phase-space distance is stored
 \begin{align}
-  t^*_i &= \argmin_{t} D_{ps,i}\\
-  A_{ik} &= (\bs{q}_i(t^*_i),\bs{p}_i(t^*_i)) % A_ik are really samples from a distribution, or we treat them as such
+  t^*_{i} &= \argmin_{t} D_{\rm ps,i}\\
+  \bs{A}_{i} &= (\bs{q}_{i}(t^*_{i}),\bs{p}_{i}(t^*_{i})).
 \end{align}
+Thus, the matrix $\bs{A}_{ik}$ contains these minimum phase-space distance vectors for each star,
 where $k\in\{1,6\}$. Intuitively, the variance of this distribution
@@
 \subsection{Application to Simulated Data}
-The LM10 (see Section \ref{sec:sgr}) simulation data is a perfect
+The LM10 simulation data (see Section \ref{sec:sgr}) is a perfect
 test-bed for evaluating the effectiveness of this method. We start by
@@
 
-% Discussion?
+% APW Note: The optimization, in a way, is getting a bunch of orbits (curves) to intersect another curve (the progenitor orbit) at some points (over time).
+
+\section{Discussion and future prospects}
+
 Many orbit fitting methods for measuring the Galactic potential assume that the stars in a given stream are on the same
@@
 potential parameters, even without realistic observational errors on
-the kinematic information. The method presented above is
+the kinematic data. The method presented above is
 free of this assumption and instead \emph{relies} on the fact that
@@
 
-% APW Note: The optimization, in a way, is getting a bunch of orbits (curves) to intersect another curve (the progenitor orbit) at some points (over time).
-
-\subsection{Future prospects}
+[merge these paragraphs (above, below)]
 
@@
 
+\acknowledgments
+We thank ...
+
+APW is supported by a National Science Foundation Graduate Research Fellowship under Grant No.\ 11-44155.
+
 \bibliography{refs}

### 2013-07-17 text/paper1/paper.tex
minor tweak

@@
 \begin{align}
-  t^*_{i} &= \argmin_{t} D_{\rm ps,i}\\
+  t^*_{i} &= \argmin_{t} D_{{\rm ps},i}\\
   \bs{A}_{i} &= (\bs{q}_{i}(t^*_{i}),\bs{p}_{i}(t^*_{i})).
@@
 Thus, the matrix $\bs{A}_{ik}$ contains these minimum phase-space distance vectors for each star,
-where $k\in\{1,6\}$. Intuitively, the variance of this distribution
+where $k\in\{1,6\}$. Intuitively, the variance of this distribution of minimum distance vectors
 will be larger for orbits integrated in an incorrect potential

### 2013-07-19 streams/potential/_lm10_acceleration.pyx
cast to array when done

@@ -94,4 +94,4 @@ def lm10_acceleration(double[:, ::1] r not None,
         data[ii,1] = fac1*r[ii,1] + fac2*r[ii,1] + fac3*(2.*C2*r[ii,1] + C3*r[ii,0])
         data[ii,2] = fac1*r[ii,2]*(1.+_tmp) + fac2*r[ii,2] + 2.*fac3*r[ii,2]/qz_sq
             
-    return data
+    return np.array(data)

### 2013-07-19 text/paper1/paper.tex

@@
 the prospect of proper motion measurements of the same stars with
-$\sim$10~mas/yr precision (Section \ref{sec:gaia}); and (iii) the
+$\sim$10~$\mu$as/yr precision (Section \ref{sec:gaia}); and (iii) the
 tracing of debris associated with Sgr around the entire Galaxy
@@
 proper motion accuracy to create full six-dimensional phase-space
-maps of the Galaxy with $<$10\% distance errors for heliocentric distances of up to $\sim$10~kpc for ...
+maps of the Galaxy with $<$10\% distance errors for heliocentric distances up to $\sim$6~kpc for RR Lyrae stars.
 
 Figure~\ref{fig:gaia_errors} show the Gaia end-of-mission distance and
-tangential velocity error estimates for a typical halo RR Lyrae star
-\citep[{$M_V=0.45$, [Fe/H]$=-1.5$};][]{ivezic2008}. Within 2~kpc, Gaia
+tangential velocity error estimates for RR Lyrae.
+The velocity uncertainty depends on the magnitude of the source, which depends on the metallicity: each line in Figure~\ref{fig:gaia_errors} is computed by Monte Carlo sampling from the empirical metallicity distribution of the Galactic halo from \cite{ivezic2008} (centered around [Fe/H] $\sim$ -1.5).
+Within 2~kpc, Gaia
 will be able to measure distances to these stars with better than 2\%
@@
 
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=1.\textwidth, trim=2 2 5 5, clip]{/Users/adrian/projects/streams/plots/paper1/gaia.pdf}
+\caption{ todo }\label{fig:gaia_errors}
+\end{center}
+\end{figure}
+
 The anticipated proper motion errors for RR Lyrae correspond to
@@
 where full six-dimensional phase-space maps of the Galaxy are possible
-from $<10$~kpc to 80~kpc. This enormous increase in volume will refine
+from $<10$~kpc to 80~kpc. This enormous increase in volume will greatly refine
 data on existing debris systems such as the Sagittarius stream.

### 2013-07-19 text/paper1/paper.tex
no time for talking

@@
 will be able to measure distances to these stars with better than 2\%
-accuracy: RR Lyraes in this volume will be used to test and calibrate
+accuracy---RR Lyraes in this volume will be used to test and calibrate
 the Spitzer PL relation described above. Beyond the 2~kpc threshold,
@@
 \begin{center}
-\includegraphics[width=1.\textwidth, trim=2 2 5 5, clip]{/Users/adrian/projects/streams/plots/paper1/gaia.pdf}
+\includegraphics[width=1.\textwidth]{/Users/adrian/projects/streams/plots/paper1/gaia.pdf}
 \caption{ todo }\label{fig:gaia_errors}
@@
 where full six-dimensional phase-space maps of the Galaxy are possible
-from $<10$~kpc to 80~kpc. This enormous increase in volume will greatly refine
+from $<$10~kpc to 80~kpc. This enormous increase in volume will greatly refine
 data on existing debris systems such as the Sagittarius stream.
@@
 \begin{equation}
-  \Phi_{halo} = v_{halo}^2 \ln(C_1 x^2 + C_2 y^2 + C_3 xy + (z/q_z)^2 + R_c^2).
+  \Phi_{halo} = v_{halo}^2 \ln(C_1 x^2 + C_2 y^2 + C_3 xy + (z/q_z)^2 + R_c^2)
 \end{equation}
-where $C_1$, $C_2$, and $C_3$ are combinations of the $x$ and $y$ flattening ($q_1$, $q_2$) and orientation of the halo with respect to the baryonic disk ($\phi$):
+where $C_1$, $C_2$, and $C_3$ are combinations of the $x$ and $y$ axis ratios ($q_1$, $q_2$) and orientation of the halo with respect to the baryonic disk ($\phi$):
 \begin{align}
@@
   C_2 &= \frac{\sin^2\phi}{q_1^2} + \frac{\cos^2\phi}{q_2^2}\\
-  C_3 &= 2\sin\phi\cos\phi \left(q_1^{-2} - q_2^{-2}\right)
+  C_3 &= 2\sin\phi\cos\phi \left(q_1^{-2} - q_2^{-2}\right).
 \end{align}
@@
 LM10 demonstrate how a comparison of such simulations with the
-observed debris enables a detailed reconstruction of Sgr's history 
-(its mass and stellar populations, orbit, rate of destruction, and
-even original distribution in populations [other refs??]). The phase-space distribution of the debris offers a unique probe of
+observed debris enables a detailed reconstruction of Sgr's history:  
+its mass and stellar populations, orbit, rate of destruction, and
+even original distribution in populations [other refs??]. The phase-space distribution of the debris offers a unique probe of
 the depth, shape, extent, and orientation of the Milky Way's dark
@@
 
-The combined Spitzer and Gaia measurements of distances and proper
+The prospect of combined Spitzer and Gaia measurements of distances and proper
 motions of RR Lyraes in the Sgr debris opens up new avenues of
@@
 velocities with $\sim$10~km/s accuracy, the measured phase-space
-position of each individual RR Lyrae relative to the stream becomes a
+position of each individual RR Lyrae becomes a
 powerful probe of the potential (see \ref{fig:}): the next section outlines a new
@@
 incorrect, the orbits of the stars will diverge from that of the
-progenitor and thus will not be recaptured by the satellite system.
+progenitor and thus will not be recaptured by the satellite system (Figure~\ref{fig:ps_distance}).
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=1.\textwidth]{/Users/adrian/projects/streams/plots/paper1/ps_distance.pdf}
+\caption{ todo }\label{fig:ps_distance}
+\end{center}
+\end{figure}
 

### 2013-07-23 text/paper1/paper.tex
revisions

@@
 indicator of unprecedented accuracy and precision. \citet{madore2012}
-have demonstrated \citep[using 5 stars with trigonometric parallaxes
+have demonstrated \citep[using five stars with trigonometric parallaxes
   measured by Hubble;][]{benedict2011} that the dispersion in the
 mid-IR PL relation -- at Spitzer wavelengths -- is of order
-$\pm$0.03~mag. This implies that it is possible to determine
-individual distances that are good to better than $\pm 2\%$ for RR
+$\sim$0.03~mag. This implies that it is possible to determine
+individual distances that are good to better than $2\%$ for RR
 Lyrae stars out to $\sim$80~kpc (the limit of Spitzer's ability to
 detect and measure RR Lyrae variables, for modest integration
-times). Distance measurements of Blue Horizontal Branch and Blue Straggler stars at these distances in the halo typically achieve 15-25\% uncertainties. [can cite Alis Deason if necessary].
+times). Distance measurements of Blue Horizontal Branch and Blue Straggler stars at these distances in the halo typically achieve $\sim$25\% uncertainties \citep[e.g.,][]{deason2012}.
 
@@
 \label{sec:gaia}
-The Gaia satellite \citep{gaia2001}, planned for launch in October
+The Gaia satellite \citep{gaia2001}, planned for launch in late
 2013, is an astrometric mission which aims to measure the positions of billions of stars with
 10-100~$\mu$as accuracies. With over 50 observations of the entire sky
-over the 5-year lifetime of the mission, Gaia will
+during the 5-year lifetime of the mission, Gaia will
 revolutionize the field of Galactic Astronomy by providing sufficient
@@
 the 3-dimensional mass distribution of a dark matter halo has been
-feasible. Figure \ref{fig:lm10} illustrates results from the LM10
-N-body simulation of dwarf satellite disruption along the expected Sgr
-orbit. The particle positions are shown in a coordinate system defined
-by the orbit of the satellite orbit \citep{majewski2004}. The
+feasible. Figure \ref{fig:lm10} shows particle positions from the final time-step of an N-body simulation of dwarf satellite disruption along the expected Sgr
+orbit in the best-fitting Milky Way halo model from LM10. [TODO: Maybe not true...] The positions are shown in a coordinate system defined
+by the orbit of the satellite \citep{majewski2004}. The
 simulation was run in a three-component potential, with a triaxial,
@@
 
-The prospect of combined Spitzer and Gaia measurements of distances and proper
-motions of RR Lyraes in the Sgr debris opens up new avenues of
-interpretation. A 2\% distance error at $\sim$30~kpc is smaller than
+Combined Spitzer and Gaia measurements of distances and proper
+motions of RR Lyraes in the Sgr debris will open up new avenues of
+interpretation. A 2\% distance error at $\sim$30~kpc ($\epsilon_{d}=0.6~{\rm kpc}$) is smaller than
 the predicted thickness of the stream, ranging from
 $\sim$2-$10$~kpc. Combined with Gaia tangential velocity measurement
-accuracies ($\sim$2~km/s at 30~kpc) less than the intrinsic dispersion
-for much of the stream \citep[$\sigma_v\sim$10~km/s;][]{majewski2004}, and ground-based radial
-velocities with $\sim$10~km/s accuracy, the measured phase-space
+accuracies ($\epsilon_{v}\sim2$~km/s at 30~kpc) less than the intrinsic dispersion
+for much of the stream \citep[$\sigma_v\sim10$~km/s;][]{majewski2004}, and ground-based radial
+velocities, the measured phase-space
 position of each individual RR Lyrae becomes a
-powerful probe of the potential (see \ref{fig:}): the next section outlines a new
+powerful probe of the potential (see \ref{fig:}). The next section outlines a new
 method to take advantage of this information.
@@
 for the progenitor system $(\bs{x}_p, \bs{v}_p)|_{t=0}$, and a
-functional form for the potential, $\Phi$. For a
-given set of potential parameters, the orbits of the stars and
+functional form for the potential, $\Phi({\boldsymbol\theta})$. For a
+given set of potential parameters, $\boldsymbol\theta$, the orbits of the stars and
 progenitor are integrated backwards for several Gigayears (determined from an
@@
 progenitor system are treated as test particles, so that the debris
-stars do not become re-bound to the satellite. Instead, at each
+stars do not become physically re-bound to the satellite. Instead, at each
 timestep $t_j$, for each particle $i$, a set of normalized, relative
@@
   \bs{q}_{i} = \frac{\bs{x}_{i} -
-    \bs{x}_{p}}{R_{\rm tide}}\,\,\,\,\,\,\,\,\bs{p}_{i} = \frac{\bs{v}_{i} -
+    \bs{x}_{p}}{R_{\rm tide}}\,\,\,\,;\,\,\,\,\bs{p}_{i} = \frac{\bs{v}_{i} -
     \bs{v}_{p}}{v_{\rm esc}}

### 2013-07-23 scripts/paper1_figures.py
update phase-space D vs. time plot

@@ -192,8 +192,8 @@ def phase_space_d_vs_time(N=10):
     selected_star_idx[randidx] = True
     
     wrong_params = true_params.copy()
-    wrong_params['qz'] = 1.1*wrong_params['qz']
-    wrong_params['v_halo'] = 1.1*wrong_params['v_halo']
+    wrong_params['qz'] = 1.25*wrong_params['qz']
+    #wrong_params['v_halo'] = wrong_params['v_halo']
     
     # define correct potential, and wrong potential
     true_potential = LawMajewski2010(**true_params)
@@ -242,21 +242,24 @@ def phase_space_d_vs_time(N=10):
     
     with rc_context(rc=rcparams):  
         fig,axes = plt.subplots(2,1, sharex=True, sharey=True, figsize=(10,10))
-        axes[0].axhline(2, linestyle='--', color='#444444')
-        axes[1].axhline(2, linestyle='--', color='#444444')
+        axes[0].axhline(1, linestyle='--', color='#444444')
+        axes[1].axhline(1, linestyle='--', color='#444444')
         
         for ii in randidx:
             for jj in range(2):
                 d = D_pses[jj][:,ii]
                 sR = sat_R[jj]
                 axes[jj].semilogy(ts[jj]/1000, d, alpha=0.5, color=rcparams['lines.color'])
-                axes[jj].semilogy(ts[jj]/1000, 1.05+0.9*(sR-sR.min())/(sR.max()-sR.min()), 
+                axes[jj].semilogy(ts[jj]/1000, 0.3*0.9*(sR-sR.min())/(sR.max()-sR.min()) + 0.45, 
                                   alpha=0.75, color='#CA0020')
                 axes[jj].semilogy(ts[jj][np.argmin(d)]/1000, np.min(d), marker='o',
                                   alpha=0.75, color=rcparams['lines.color'])
         
-        axes[1].set_ylim(1,100)
+        axes[1].set_ylim(0.4,20)
         axes[1].set_xlim(-6, 0)
+        
+        axes[0].xaxis.tick_bottom()
+        axes[1].xaxis.tick_bottom()
     
     axes[1].set_xlabel("Backwards integration time [Gyr]")
     axes[0].set_ylabel(r"$D_{ps}$", rotation='horizontal')
@@ -270,5 +273,5 @@ def phase_space_d_vs_time(N=10):
 
 if __name__ == '__main__':
     #gaia_spitzer_errors()
[…]

### 2013-07-24 scripts/spitzer_proposal_figures.py
more changes from kvj

@@ -144,16 +144,14 @@ def gaia_spitzer_errors():
         fig,axes = plt.subplots(1, 2, figsize=(12, 6), sharex=True)
         
         # vertical lines for dwarf satellites
-        for name,dist in [('Sagittarius', 24), ('Ursa Minor, Bootes', 65), 
-                          ('Sculptor', 80), ('Carina', 100)]:
-            axes[0].axvline(dist, zorder=-1, alpha=0.4, linestyle='--')
-            axes[1].axvline(dist, zorder=-1, alpha=0.4, linestyle='--')
+        axes[0].vlines([24,65,80,100], 0.001, 100, color='k', alpha=0.4, linestyles='--', zorder=-1)
+        axes[1].vlines([24,65,80,100], 0.1, 100, color='k', alpha=0.4, linestyles='--', zorder=-1)
         
         # label the vertical lines
-        axes[1].text(16., 0.1, 'Sgr-', alpha=0.6)
-        axes[1].text(27., 0.06, 'UMi, Boo-', alpha=0.6)
-        axes[1].text(57., 0.03, 'Scl-', alpha=0.6)
-        axes[1].text(68.5, 0.015, 'Car-', alpha=0.6)
+        axes[1].text(16.5, 0.07, 'Sgr -', alpha=0.6, rotation=45)
+        axes[1].text(32., 0.075, 'UMi, Boo -', alpha=0.6, rotation=45)
+        axes[1].text(57., 0.07, 'Scl -', alpha=0.6, rotation=45)
+        axes[1].text(71, 0.07, 'Car -', alpha=0.6, rotation=45)
         
         # Distance from 1kpc to ~100kpc
         D = np.logspace(0., 2., 50)*u.kpc
@@ -377,6 +375,6 @@ def bootstrapped_parameters_v1():
     fig.savefig(os.path.join(plot_path, "bootstrap.pdf"))
 
 if __name__ == '__main__':
-    #gaia_spitzer_errors()
+    gaia_spitzer_errors()
     #sgr()
-    bootstrapped_parameters_v1()
+    #bootstrapped_parameters_v1()

### 2013-07-24 text/paper1/paper.tex
suggestions and new paragraph from Kathryn

@@
 
-The  extent of the bulk of the visible matter in galaxies is roughly 10\% of that of the host dark matter halo. Hence the major components of a galaxy can be used to  trace
-the inner portion of its host halo's mass distribution.
-The rotation curves of disk galaxies are particularly sensitive probes since matter in disks can be assumed to move on near circular orbits.
-This knowledge gives us extra information about the orbital distribution of the visible matter within the dark matter halo, which allows a more precise tracing of the
-potential than is feasible for objects such as elliptical galaxies where orbits are assumed to be random.
-For example, rotation curves of low-surface-brightness galaxies can be used to examine whether dark matter halos have a cusped, power-law density profile in their
-centers as predicted by purely dark matter simulations of structure formation, or a constant density core perhaps induced by the influence of baryonic processes \citep[e.g.]
-[]{kuzio09}.
-
-% Quote recent Binney/McMillan and Bovy/Rix work here? In longer paper…. Really about large disk samples and disk df's?
-% McMillan & Binney Tori work - but again assume fully populated DF
-
-Around external galaxies, the  dark matter distribution  at larger radii has been explored using a variety of approaches \citep[see][for a a complete and detailed review]
-{courteau12}.
-For example, the kinematics of tracers populations such as globular clusters or planetary nebulae  can be used to derive mass estimates under the assumptions that
-these satellite systems are relaxed and well-mixed \citep[early investigations include][]{cote03,mendez01}.
-The level of detail in these estimates is necessarily limited by the number of objects bright enough for spectroscopic measurements. The largest samples to date have of
-order several hundred members probing out to several 10's of kpc,  and allowing simple, parameterized models of the mass and orbit distribution to be simultaneously
-constrained \citep[e.g.][]{napolitano11,deason12a}.
-%\citep{romanowsky03,douglas07,romanowsky09,napolitano11,lee11}
-Alternatively, the properties of gravitationally lensed background sources around a galaxy can be used to map the shape, orientation and radial profile of the projected
-mass distribution \citep[as done by the Lens Structure and Dynamics Survey described in][]{koopmans02}.
-The level of detail in these two-dimensional reconstructions is impressive:
-recent work has even found some systems sensitive to the position and mass of a major subhalo orbiting with the main halo \citep[e.g.][]{vegetti10}.
-Of course, lensing reconstructions can only be performed for galaxies which closely intersect our line of sight to background sources, but the advent of large photometric
-catalogues has allowed automatic searches for such chance alignments and significant increases in the number of objects studied \citep[e.g. the Sloan Lens ACS Survey,
-see][]{bolton06}.
-
 The one place in the Universe where we might hope to dissect a dark matter halo in even more detail is locally for our own Milky Way Galaxy.
@@
 orbit distribution nor is fully phase-mixed.
-The degree of substructure increases with Galactocentric radius \citep{bell08} in agreement with models where the stellar halo is formed purely from accretion events
-\citep{bullock05,cooper10,rashkov12}.
-Moreover, even in the smoother, inner halo which is likely to be phase-mixed, significant clumpings in velocity \citep{schlauffman10} and orbital properties \citep{helmi99}
-have been observed, which indicate a non-smooth component to the orbit distribution.
-By ``observing'' particles in a sample of simulated stellar halos, \citet{yencho06} showed that, if stellar halos are indeed entirely formed via accretion events, then the
-expected degree of non-randomness  could lead to systematic biases in mass estimates of order several tens of percent.
-This encouraging result suggests that, if our expectations for the typical number and size of accretion events contributing to the stellar halo are correct, then the number of
-orbits effectively explored is sufficiently large that the assumption of random, fully phase-mixed orbits is reasonable for the current samples.
 
@@
 \citep{law11}.
-Studies which first fit the Sgr stream to a single orbit \citep{johnston05,law09} and were subsequently followed up with full N-body simulations \citep{law05,law12} offer
-another demonstration of the limitation of the former approach as potential parameters are revised when moving to the N-body approach.
 
@@
 assumption of the form of the Galactic potential (although neither tests their method with realistic observational errors???).
-\citet{johnston99} (hereafter JZSH) showed that backwards integration of the positions of stars and their parent satellite could be used to distinguish Milky Way mass
-models: only in the correct potential would the stars' and satellite's  path coincide within a Hubble time.
-The JZSH method was designed and tested for the proposed characteristics of the Space Interferometry Mission \citep{}, which was expected to be able to measure
-proper motions with $\mu$as/year accuracy for stars as faint as $V=20$.
-Distances to stream members would only be poorly known and were estimated using anticipated properties of tidal debris.
 
@@
 With access to full phase-space information for stars in a tidal
-stream, it is no longer necessary to model the orbit of the stream
-itself. Instead, it is possible to use the individual stars as
+stream, it is possible to use the individual stars as
 potential measures by exploting the fact that the stars must have come
-from the same progenitor. If the present-day orbital parameters of the
-progenitor system are known, this places an intuitive constraint on
+from the same progenitor. If the present-day phase-space position of the
+progenitor system is also known, this places an intuitive constraint on
 the potential of the Milky Way: if the orbits of the stars and the
-orbit of the progenitor are integrated \emph{backwards} in time, the
+orbit of the progenitor are integrated \emph{backwards} in time
+(imagine watching a visualization of a simulation of satellite destruction "rewind"), the
 stars should `recombine' with the progenitor. If the potential is
@@
 progenitor and thus will not be recaptured by the satellite system (Figure~\ref{fig:ps_distance}).
+
+This approach was originally propose by \citep{jzsh1999} and designed and tested for the proposed characteristics of the Space Interferometry Mission \citep{unwin??}, which was expected to be able to measure proper motions with $\mu$as/year accuracy for stars as faint as $V=20$.
+Distances to stream members would only be poorly known and were estimated in the method by using anticipated properties of tidal debris.
+Below we present an updated version of the algorithm: the promise of 2\% distances to RR Lyrae stars (see Section \ref{sec:spitzer}) enables a direct measurement (rather than approximate estimate) of the position of a star within its debris structure; and the test statistic that quantifies how well stars recombine with the satellite has been rigorously redefined.
 

### 2013-07-24 text/paper1/paper.tex
add orbital radius to tidal radius definition

@@
 \begin{equation}
-  R_{\rm tide}=\Big(\frac{m_{\rm sat}}{3m_{\rm enc}}\Big)^{1/3}\,\,\,\,\,\,\,\,
+  R_{\rm tide}=R_{{\rm orbit}}\Big(\frac{m_{\rm sat}}{3m_{\rm enc}}\Big)^{1/3}\,\,\,\,;\,\,\,\,
   v_{\rm esc}=\sqrt{\frac{2Gm_{\rm sat}}{R_{\rm tide}}}.

### 2013-07-24 text/paper1/paper.tex
change _orbit to just R

--- 2013-07-24 text/paper1/paper.tex
@@
 feasible. Figure \ref{fig:lm10} shows particle positions from the final time-step of an N-body simulation of dwarf satellite disruption along the expected Sgr
-orbit in the best-fitting Milky Way halo model from LM10. [TODO: Maybe not true...] The positions are shown in a coordinate system defined
-by the orbit of the satellite \citep{majewski2004}. The
+orbit in the best-fitting Milky Way halo model from LM10. [TODO: Maybe not true...: The positions are shown in a coordinate system defined
+by the orbit of the satellite \citep{majewski2004}]. The
 simulation was run in a three-component potential, with a triaxial,
@@
 satellite, $m_{sat}$, which, combined with the computed enclosed mass
-of the potential, $m_{enc}$, sets the instantaneous tidal radius and
+of the potential, $m_{enc}$, and the orbital radius of the satellite, $R$, sets the instantaneous tidal radius and
 escape velocity,
 \begin{equation}
-  R_{\rm tide}=\Big(\frac{m_{\rm sat}}{3m_{\rm enc}}\Big)^{1/3}\,\,\,\,\,\,\,\,
+  R_{\rm tide}=R\Big(\frac{m_{\rm sat}}{3m_{\rm enc}}\Big)^{1/3}\,\,\,\,;\,\,\,\,
   v_{\rm esc}=\sqrt{\frac{2Gm_{\rm sat}}{R_{\rm tide}}}.
@@
 Qualitatively, when the distance in this normalized 6D space,
-$D_{\rm ps}=\sqrt{|\bs{q}_{i}|^2+|\bs{p}_{i}|^2} < \sqrt{2}$, the star is
-probably recaptured by the satellite, but this is not imposed as a
+$D_{{\rm ps},i}=\sqrt{|\bs{q}_{i}|^2+|\bs{p}_{i}|^2} < \sqrt{2}$, the star is
+likely recaptured by the satellite, but this is not imposed as a
 hard constraint. [JZSH ...] As a statistical measure of how many stars have recombined, we
@@
 Thus, the matrix $\bs{A}_{ik}$ contains these minimum phase-space distance vectors for each star,
-where $k\in\{1,6\}$. Intuitively, the variance of this distribution of minimum distance vectors
+where $k\in[1,6]$. Intuitively, the variance of this distribution of minimum distance vectors
 will be larger for orbits integrated in an incorrect potential

### 2013-07-24 scripts/paper1_figures.py
style updates and foo

@@ -30,10 +30,10 @@ from streams.potential.lm10 import true_params
 from streams.integrate.satellite_particles import SatelliteParticleIntegrator
 from streams.io.lm10 import particles_today, satellite_today, time
 
-matplotlib.rc('xtick', labelsize=14)
-matplotlib.rc('ytick', labelsize=14)
+matplotlib.rc('xtick', labelsize=18)
+matplotlib.rc('ytick', labelsize=18)
 matplotlib.rc('text', usetex=True)
-matplotlib.rc('axes', edgecolor='#444444', labelsize=20)
+matplotlib.rc('axes', edgecolor='#444444', labelsize=24)
 matplotlib.rc('lines', markeredgewidth=0)
 matplotlib.rc('font', family='sans-serif')
 rcParams['font.sans-serif'] = 'helvetica'
@@ -85,26 +85,64 @@ def normed_objective_plot():
 def variance_projections():
     """ Figure showing 2D projections of the 6D variance """
     
+    #particles = particles_today(N=100, expr="(Pcol>-1) & (abs(Lmflag) == 1) & (Pcol < 7)")
+    
     params = true_params.copy()
     params['qz'] = true_params['qz']*1.2
     params['v_halo'] = true_params['v_halo']*1.2
     
     # define both potentials
     correct_lm10 = LawMajewski2010(**true_params)
-    wrong_lm10 = LawMajewski2010(**true_params)
+    wrong_lm10 = LawMajewski2010(**params)
     
-    fig = None
-    for potential in [correct_lm10, wrong_lm10]:
-        integrator = SatelliteParticleIntegrator(potential, satellite, particles)
-        s_orbit,p_orbits = integrator.run(t1=t1, t2=t2, dt=-1.)
-        min_ps = minimum_distance_matrix(potential, p_orbits, s_orbit)
+    rcparams = {'lines.linestyle' : '-', 
+                'lines.linewidth' : 1.,
+                'lines.color' : 'k',
+                'lines.marker' : None,
+                'axes.facecolor' : '#ffffff'}
+    
+    with rc_context(rc=rcparams):
+        fig,axes = plt.subplots(2, 2, figsize=(10,10))
+        colors = ['#0571B0', '#D7191C']
+        for ii,potential in enumerate([correct_lm10, wrong_lm10]):
+            integrator = SatelliteParticleIntegrator(potential, satellite, particles)
[…]

### 2013-07-24 text/paper1/paper.tex
change numbers to text, additions

@@
 where $(\bs{x},\bs{v})_{i}$ and $(\bs{x},\bs{v})_{p}$ are the phase-space coordinates for the particles and progenitor, respectively. These definitions require an estimate of the mass of the
-satellite, $m_{sat}$, which, combined with the computed enclosed mass
-of the potential, $m_{enc}$, and the orbital radius of the satellite, $R$, sets the instantaneous tidal radius and
+satellite, $m_{sat}$, which, combined with the orbital radius of the satellite, $R$, and the computed enclosed mass
+of the potential within $R$, $m_{enc}$, sets the instantaneous tidal radius and
 escape velocity,
@@
 \end{equation}
+Qualitatively, when the distance in this normalized six-dimensional space,
+$D_{{\rm ps},i}=\sqrt{|\bs{q}_{i}|^2+|\bs{p}_{i}|^2} < \sqrt{2}$, the star is
+likely recaptured by the satellite.
+JZSH imposed a similar condition as a hard boundary and used the number of recaptured particles in a given backwards-integration as a cost function for optimization.
+What follows is a description of an updated procedure with a more natural choice for an objective function [TODO...].
 
-Qualitatively, when the distance in this normalized 6D space,
-$D_{{\rm ps},i}=\sqrt{|\bs{q}_{i}|^2+|\bs{p}_{i}|^2} < \sqrt{2}$, the star is
-likely recaptured by the satellite, but this is not imposed as a
-hard constraint. [JZSH ...] As a statistical measure of how many stars have recombined, we
-construct the 6D distribution of the minimum phase-space vectors for
-each star over each individual, integrated orbit. For each star, $i$,
+As a statistical measure of how many stars have recombined, we
+use the minimum phase-space vectors --- for
+each star over the individual, integrated orbits --- as samples from an underlying six-dimensional distribution. For each star, $i$,
 the phase-space distance, $D_{\rm ps}$, is computed at each timestep
@@
 \begin{equation}
-  f(\bs{\theta})=\det(\mathrm{Cov}(A_{ik})).
+  f(\bs{\theta})=\det(\mathrm{Cov}( \bs{A}_{ik})).
 \end{equation}
-[This objective function is an approximation to a more general likelihood function, described in detail in forthcoming work (Price-Whelan \& Johnston in prep.).
-The results shown in sections xx and yy are meant to illustrate the strength and concept of the method..]
 
@@
 are computed from the expected Gaia error curve (shown in
-Figure~\ref{fig:gaia}), and the distance errors on a star of distance
+Figure~\ref{fig:gaia_errors}), and the distance errors on a star of distance
 $D$ are $\epsilon_{dist} \sim \mathcal{N}(0,0.02\times
-D)$. Figure~\ref{fig:} shows all 2D projections of the 6D minimum
+D)$. Figure~\ref{fig:variance6d} shows all two-dimensional projections of the six-dimensional minimum
 phase-space distance distribution for the observed particles in the
-correct and an incorrect potential. Qualitatively, it is clear that the
-variance of the distribution increases in the wrong potential. The objective function
-(generalized variance) defines a convex hypersurface over which we
+correct and an incorrect potential: it is clear that the
+variance of the distribution increases in the wrong potential. The 
+generalized variance defines a convex hypersurface over which we
 optimize four of the six logarithmic potential parameters:
-$v_{circ}$, $\phi$, $q_1$, and $q_z$. Figure~\ref{fig:} shows [...]
-curves produced by varying the true parameters of the potential by
-10\% around the true value.
+$v_{circ}$, $\phi$, $q_1$, and $q_z$. We fix $q_2$ and $R_c$, which are degenerate with combinations of the other parameters. Figure~\ref{fig:} shows one-dimensional slices of the objective function
+produced by varying each of the potential parameters by
+$\pm20\%$ around the true values and holding all others fixed.
 
@@
 \section{Discussion and future prospects}
+
+% this needs to go somewhere down here...
+[This objective function is an approximation to a more general likelihood function, described in detail in forthcoming work (Price-Whelan \& Johnston in prep.).
+The results shown in sections xx and yy are meant to illustrate the strength and concept of the method..]
 

### 2013-07-26 text/paper1/paper.tex
not entirely sure..

@@
 
-This approach was originally propose by \citep{jzsh1999} and designed and tested for the proposed characteristics of the Space Interferometry Mission \citep{unwin??}, which was expected to be able to measure proper motions with $\mu$as/year accuracy for stars as faint as $V=20$.
+This approach was originally proposed by \citep{jzsh1999} and was designed and tested for the proposed characteristics of the Space Interferometry Mission \citep{unwin??}, which was expected to be able to measure proper motions with $\mu$as/year accuracy for stars as faint as $V=20$.
 Distances to stream members would only be poorly known and were estimated in the method by using anticipated properties of tidal debris.
-Below we present an updated version of the algorithm: the promise of 2\% distances to RR Lyrae stars (see Section \ref{sec:spitzer}) enables a direct measurement (rather than approximate estimate) of the position of a star within its debris structure; and the test statistic that quantifies how well stars recombine with the satellite has been rigorously redefined.
+Below we present an updated version of the algorithm: the promise of 2\% distances to RR Lyrae stars (see Section \ref{sec:spitzer}) enables a direct measurement (rather than approximate estimate) of the position of a star within its debris structure.
+The test statistic that quantifies how well stars recombine with the satellite has been rigorously redefined.
 
@@
 satellite, $m_{sat}$, which, combined with the orbital radius of the satellite, $R$, and the computed enclosed mass
-of the potential within $R$, $m_{enc}$, sets the instantaneous tidal radius and
+of the potential within $R$, $M_{enc}$, sets the instantaneous tidal radius and
 escape velocity,
 \begin{equation}
-  R_{\rm tide}=R\Big(\frac{m_{\rm sat}}{3m_{\rm enc}}\Big)^{1/3}\,\,\,\,;\,\,\,\,
+  R_{\rm tide}=R\Big(\frac{m_{\rm sat}}{3M_{\rm enc}}\Big)^{1/3}\,\,\,\,;\,\,\,\,
   v_{\rm esc}=\sqrt{\frac{2Gm_{\rm sat}}{R_{\rm tide}}}.
@@
 As a statistical measure of how many stars have recombined, we
-use the minimum phase-space vectors --- for
-each star over the individual, integrated orbits --- as samples from an underlying six-dimensional distribution. For each star, $i$,
+treat the minimum phase-space vectors (for
+each star over the individual, integrated orbits) as samples from an underlying six-dimensional distribution. For each star, $i$,
 the phase-space distance, $D_{\rm ps}$, is computed at each timestep
@@
 Thus, the matrix $\bs{A}_{ik}$ contains these minimum phase-space distance vectors for each star,
-where $k\in[1,6]$. Intuitively, the variance of this distribution of minimum distance vectors
+where $k\in[1,6]$. Intuitively, the variance of this distribution
 will be larger for orbits integrated in an incorrect potential
 relative to the distribution computed from the `true' orbital history
-of the stars; in an incorrect potential, small differences in the
+of the stars: in an incorrect potential, small differences in the
 orbits of the stars relative to the orbit of the progenitor cause them

### 2013-07-27 text/paper1/paper.tex
add note to write about bootstrap

@@
 JZSH imposed a similar condition as a hard boundary and used the number of recaptured particles in a given backwards-integration as a cost function for optimization.
-What follows is a description of an updated procedure with a more natural choice for an objective function [TODO...].
+What follows is a description of an updated procedure with a more natural, statistically-motivated choice for an objective function.
 
-As a statistical measure of how many stars have recombined, we
-treat the minimum phase-space vectors (for
-each star over the individual, integrated orbits) as samples from an underlying six-dimensional distribution. For each star, $i$,
+%As a measure of how many stars have recombined, we treat the minimum phase-space vectors (for each star over the individual, integrated orbits) as samples from an underlying six-dimensional distribution.
+
+For each star, $i$,
 the phase-space distance, $D_{\rm ps}$, is computed at each timestep
@@
 $\pm20\%$ around the true values and holding all others fixed.
+
+[Bootstrap]
 

### 2013-07-27 text/paper1/paper.tex
starting to write discsussion

@@
 
-% this needs to go somewhere down here...
-[This objective function is an approximation to a more general likelihood function, described in detail in forthcoming work (Price-Whelan \& Johnston in prep.).
-The results shown in sections xx and yy are meant to illustrate the strength and concept of the method..]
+Many orbit fitting methods for measuring the Galactic potential assume that the stars in a given stream are on the same
+orbit. \cite{binney??} show that this assumption leads to XXX\% biases in the 
+potential parameters, even without realistic observational errors on
+the kinematic data. The \emph{Rewinder} is
+free of this assumption and instead \emph{relies} on the fact that
+each star is on a different orbit, and is thus an independent potentiometer: the best potential is the one that gets the
+most of these orbits to intersect the progenitor
+over the history of interaction with the system.
+The strength of this method is in the simplicity: it requires only a rough estimate of the satellite mass $m_{\rm sat}$ combined with backwards integration of orbits.
+There are no assumptions made about, or N-body modeling of, the internal distribution of satellite stars.
 
-Many orbit fitting methods for measuring the Galactic potential assume that the stars in a given stream are on the same
-orbit. \cite{binney??} show that this assumption leads to biased
-potential parameters, even without realistic observational errors on
-the kinematic data. The method presented above is
-free of this assumption and instead \emph{relies} on the fact that
-each star is on a different orbit and is thus an independent potentiometer. The power of this method is in
-the joint constraint that the best potential is the one that gets the
-most of these individual orbits to intersect the progenitor
-over the history of interaction with the system.
-
-[merge these paragraphs (above, below)]
-
-The great strength of the  JZSH ({\it can we come up with a catchy name?}) method is its simplicity:
-it requires only the assumption a satellite mass $m_{\rm sat}$ and  backwards integration of orbits.
-There are no assumptions made about or N-body modeling required to follow the internal distribution of satellite stars or accurately assess the degree of alignment of
-debris with a single orbit.
-In fact, the method relies on the ability to measure {\it differences} in orbital properties --- these differences result in different orbital time periods for the debris and the
-satellite which leads to their paths intersecting during the backwards integration.
-
+[TODO: this paragraph needs work]
 There are two other apparent strengths of the method which require further investigation.
@@
 ps}$ defined appropriately for each star separately.
+
+% this needs to go somewhere down here...
+[This objective function is an approximation to a more general likelihood function, described in detail in forthcoming work (Price-Whelan \& Johnston in prep.).
+The results shown in sections xx and yy are meant to illustrate the strength and concept of the method..]
 

### 2013-07-27 scripts/optimize_likelihood.py
trying to optimize parameters instead of mcmc

@@ -34,9 +34,7 @@ x0 = fmin_l_bfgs_b(lambda *args,**kwargs: -ln_posterior(*args,**kwargs),
                   x0=[1.21, 1.21],
                   args=(['q1', 'qz'], particles, satellite, t1, t2, 3.),
                   bounds=[(1.2,1.4),(1.2,1.4)],
-                  maxiter=3,
-                  factr=1E5,
                   approx_grad=True,
-                  epsilon=1E-4)
+                  epsilon=1E-3)
 
 print(x0)

### 2013-07-27 text/paper1/paper.tex
revisions

@@
 There are two other apparent strengths of the method which require further investigation.
-First it appears applicable to any debris that is known to come from a single object, and not restricted only to the very coldest tidal streams. For example, in principle the
-method could  applied to the vast debris {\it clouds} that have been discovered subtending hundreds-thousands square degrees \citep[e.g. the Triangulum Andromeda
+\emph{The Rewinder} is applicable to any debris that is known to come from a single object, and not restricted only to the very coldest tidal streams. For example, in principle the
+method could applied to the vast debris {\it clouds} that have been discovered subtending large solid angles \citep[e.g. the Triangulum Andromeda
 and Hercules-Aquila clouds]{rochapinto04,belokurov06}, or even stars that have only associations in orbital properties and do not form a coherent spatial structure
 \citep[such as the angular momentum groupings in local giants found by][]{helmi99}.
-Second, it appears trivial to combine constraints from multiple debris systems at once by simply integrating all debris from several satellites simultaneously with $D_{\rm
-ps}$ defined appropriately for each star separately.
+Second, it appears trivial to combine constraints from multiple debris systems at once by simply integrating all debris from several satellites simultaneously with $D_{\rm ps}$ defined appropriately for each star separately.
 
-% this needs to go somewhere down here...
-[This objective function is an approximation to a more general likelihood function, described in detail in forthcoming work (Price-Whelan \& Johnston in prep.).
-The results shown in sections xx and yy are meant to illustrate the strength and concept of the method..]
+[present the limitations]
+The limitations of the method also need to be understood. Firstly, for the RR Lyraes associated with the very coldest streams \citep[e.g. the globular clusters Pal5
+and GD1][]{odenkirchen02,koposov10} the measurement errors are unlikely to be small enough for the
+the differences in orbital properties to between debris and satellite to be detected.
+[What will this cause?
+Maybe biases?
+One more sentence here].
+Second, the method requires specifying a functional form for the potential.
+[...].
+Lastly, the current version of the algorithm relies on accurate knowledge of the phase-space position of the parent satellite, which may not be available in some cases \citep[for example, for the Orphan Stream][]{belokurov07}.
 
-The limitations of the method also need to be understood. In particular, for the RR Lyraes associated with  the very coldest streams \citep[e.g. the globular clusters Pal5
-and GD1][]{odenkirchen02,koposov10} the measurement errors are unlikely to be small enough for the
-the differences in orbital properties to between debris and satellite to be detected, and the method should fail.
+[explain how some of them are handled automatically by the forthcoming machinery]
+The objective function presented above is an approximation to a more general likelihood function, described in detail in forthcoming work (Price-Whelan \& Johnston in prep.).
+The results shown in sections xx and yy are meant to illustrate the strength and concept of the method.
 
-Lastly, the current version of the algorithm relies on accurate knowledge of the phase-space position of the parent satellite, which may not be available in some cases
-\citep[for example, for the Orphan Stream][]{belokurov07}. Given sufficient data, position and motion of the original host could be solved for as additional free parameters,
+Given sufficient data, position and motion of the original host could be solved for as additional free parameters,
 but the effectiveness of this approach has yet to be investigated.

### 2013-07-27 text/paper1/paper.tex
line wrapping

@@
 indicator of unprecedented accuracy and precision. \citet{madore2012}
-have demonstrated \citep[using five stars with trigonometric parallaxes
-  measured by Hubble;][]{benedict2011} that the dispersion in the
-mid-IR PL relation -- at Spitzer wavelengths -- is of order
+have demonstrated \citep[using five stars with trigonometric
+  parallaxes measured by Hubble;][]{benedict2011} that the dispersion
+in the mid-IR PL relation -- at Spitzer wavelengths -- is of order
 $\sim$0.03~mag. This implies that it is possible to determine
-individual distances that are good to better than $2\%$ for RR
-Lyrae stars out to $\sim$80~kpc (the limit of Spitzer's ability to
-detect and measure RR Lyrae variables, for modest integration
-times). Distance measurements of Blue Horizontal Branch and Blue Straggler stars at these distances in the halo typically achieve $\sim$25\% uncertainties \citep[e.g.,][]{deason2012}.
+individual distances that are good to better than $2\%$ for RR Lyrae
+stars out to $\sim$80~kpc (the limit of Spitzer's ability to detect
+and measure RR Lyrae variables, for modest integration
+times). Distance measurements of Blue Horizontal Branch and Blue
+Straggler stars at these distances in the halo typically achieve
+$\sim$25\% uncertainties \citep[e.g.,][]{deason2012}.
 
@@
 \label{sec:gaia}
-The Gaia satellite \citep{gaia2001}, planned for launch in late
-2013, is an astrometric mission which aims to measure the positions of billions of stars with
-10-100~$\mu$as accuracies. With over 50 observations of the entire sky
-during the 5-year lifetime of the mission, Gaia will
-revolutionize the field of Galactic Astronomy by providing sufficient
-proper motion accuracy to create full six-dimensional phase-space
-maps of the Galaxy with $<$10\% distance errors for heliocentric distances up to $\sim$6~kpc for RR Lyrae stars.
+The Gaia satellite \citep{gaia2001}, planned for launch in late 2013,
+is an astrometric mission which aims to measure the positions of
+billions of stars with 10-100~$\mu$as accuracies. With over 50
+observations of the entire sky during the 5-year lifetime of the
+mission, Gaia will revolutionize the field of Galactic Astronomy by
+providing sufficient proper motion accuracy to create full
+six-dimensional phase-space maps of the Galaxy with $<$10\% distance
+errors for heliocentric distances up to $\sim$6~kpc for RR Lyrae
+stars.
 
 Figure~\ref{fig:gaia_errors} show the Gaia end-of-mission distance and
-tangential velocity error estimates for RR Lyrae.
-The velocity uncertainty depends on the magnitude of the source, which depends on the metallicity: each line in Figure~\ref{fig:gaia_errors} is computed by Monte Carlo sampling from the empirical metallicity distribution of the Galactic halo from \cite{ivezic2008} (centered around [Fe/H] $\sim$ -1.5).
-Within 2~kpc, Gaia
-will be able to measure distances to these stars with better than 2\%
-accuracy---RR Lyraes in this volume will be used to test and calibrate
-the Spitzer PL relation described above. Beyond the 2~kpc threshold,
-the mid-IR PL relation for RR Lyraes will provide more accurate
-distance measurements.
+tangential velocity error estimates for RR Lyrae. The velocity
+uncertainty depends on the magnitude of the source, which depends on
+the metallicity: each line in Figure~\ref{fig:gaia_errors} is computed
+by Monte Carlo sampling from the empirical metallicity distribution of
+the Galactic halo from \cite{ivezic2008} (centered around [Fe/H]
+$\sim$ -1.5). Within 2~kpc, Gaia will be able to measure distances to
+these stars with better than 2\% accuracy---RR Lyraes in this volume
+will be used to test and calibrate the Spitzer PL relation described
+above. Beyond the 2~kpc threshold, the mid-IR PL relation for RR
+Lyraes will provide more accurate distance measurements.
 
@@
 where full six-dimensional phase-space maps of the Galaxy are possible
-from $<$10~kpc to 80~kpc. This enormous increase in volume will greatly refine
-data on existing debris systems such as the Sagittarius stream.
+from $<$10~kpc to 80~kpc. This enormous increase in volume will
+greatly refine data on existing debris systems such as the Sagittarius
+stream.
 
@@
 Most recently, \citet[][hereafter LM10]{law2010} have combined all
-current data on the Sgr debris to assess the triaxiality and orientation
-of the outer Galaxy, the first time that such a reconstruction of
-the 3-dimensional mass distribution of a dark matter halo has been
-feasible. Figure \ref{fig:lm10} shows particle positions from the final time-step of an N-body simulation of dwarf satellite disruption along the expected Sgr
-orbit in the best-fitting Milky Way halo model from LM10. [TODO: Maybe not true...: The positions are shown in a coordinate system defined
-by the orbit of the satellite \citep{majewski2004}]. The
-simulation was run in a three-component potential, with a triaxial,
-logarithmic halo model of the form
+current data on the Sgr debris to assess the triaxiality and
+orientation of the outer Galaxy, the first time that such a
+reconstruction of the 3-dimensional mass distribution of a dark matter
+halo has been feasible. Figure \ref{fig:lm10} shows particle positions
+from the final time-step of an N-body simulation of dwarf satellite
+disruption along the expected Sgr orbit in the best-fitting Milky Way
+halo model from LM10. [TODO: Maybe not true...: The positions are
+  shown in a coordinate system defined by the orbit of the satellite
+  \citep{majewski2004}]. The simulation was run in a three-component
+potential, with a triaxial, logarithmic halo model of the form
 \begin{equation}
@@
 observed debris enables a detailed reconstruction of Sgr's history:
-its mass and stellar populations, orbit, rate of destruction, and
-even original distribution in populations [other refs??]. The phase-space distribution of the debris offers a unique probe of
-the depth, shape, extent, and orientation of the Milky Way's dark
-matter halo that cannot be matched using other techniques on our own
-or other galaxies \citep[e.g.,][]{ibata2001}. The LM10 constraints on
-Sgr's mass and orbit have invigorated discussions of Sgr in a more
-cosmological context, the mass and extent of the original dark matter
-halo that hosted it, as well as its effect on the Milky Way itself
+its mass and stellar populations, orbit, rate of destruction, and even
+original distribution in populations [other refs??]. The phase-space
+distribution of the debris offers a unique probe of the depth, shape,
+extent, and orientation of the Milky Way's dark matter halo that
+cannot be matched using other techniques on our own or other galaxies
+\citep[e.g.,][]{ibata2001}. The LM10 constraints on Sgr's mass and
+orbit have invigorated discussions of Sgr in a more cosmological
+context, the mass and extent of the original dark matter halo that
+hosted it, as well as its effect on the Milky Way itself
 \citep{bailin2003,purcell2011,micheldansac2011,gomez2012}.
 
-Combined Spitzer and Gaia measurements of distances and proper
-motions of RR Lyraes in the Sgr debris will open up new avenues of
-interpretation. A 2\% distance error at $\sim$30~kpc ($\epsilon_{d}=0.6~{\rm kpc}$) is smaller than
-the predicted thickness of the stream, ranging from
-$\sim$2-$10$~kpc. Combined with Gaia tangential velocity measurement
-accuracies ($\epsilon_{v}\sim2$~km/s at 30~kpc) less than the intrinsic dispersion
-for much of the stream \citep[$\sigma_v\sim10$~km/s;][]{majewski2004}, and ground-based radial
-velocities, the measured phase-space
-position of each individual RR Lyrae becomes a
-powerful probe of the potential (see \ref{fig:}). The next section outlines a new
-method to take advantage of this information.
+Combined Spitzer and Gaia measurements of distances and proper motions
+of RR Lyraes in the Sgr debris will open up new avenues of
+interpretation. A 2\% distance error at $\sim$30~kpc
+($\epsilon_{d}=0.6~{\rm kpc}$) is smaller than the predicted thickness
+of the stream, ranging from $\sim$2-$10$~kpc. Combined with Gaia
+tangential velocity measurement accuracies ($\epsilon_{v}\sim2$~km/s
+at 30~kpc) less than the intrinsic dispersion for much of the stream
+\citep[$\sigma_v\sim10$~km/s;][]{majewski2004}, and ground-based
+radial velocities, the measured phase-space position of each
+individual RR Lyrae becomes a powerful probe of the potential (see
+\ref{fig:}). The next section outlines a new method to take advantage
+of this information.
 
@@
 With access to full phase-space information for stars in a tidal
-stream, it is possible to use the individual stars as
-potential measures by exploting the fact that the stars must have come
-from the same progenitor. If the present-day phase-space position of the
-progenitor system is also known, this places an intuitive constraint on
-the potential of the Milky Way: if the orbits of the stars and the
+stream, it is possible to use the individual stars as potential
+measures by exploting the fact that the stars must have come from the
+same progenitor. If the present-day phase-space position of the
+progenitor system is also known, this places an intuitive constraint
+on the potential of the Milky Way: if the orbits of the stars and the
 orbit of the progenitor are integrated \emph{backwards} in time
-(imagine watching a visualization of a simulation of satellite destruction "rewind"), the
-stars should `recombine' with the progenitor. If the potential is
-incorrect, the orbits of the stars will diverge from that of the
-progenitor and thus will not be recaptured by the satellite system (Figure~\ref{fig:ps_distance}).
+(imagine watching a visualization of a simulation of satellite
+destruction "rewind"), the stars should `recombine' with the
+progenitor. If the potential is incorrect, the orbits of the stars
+will diverge from that of the progenitor and thus will not be
+recaptured by the satellite system (Figure~\ref{fig:ps_distance}).
 
-This approach was originally proposed by \citep{jzsh1999} and was designed and tested for the proposed characteristics of the Space Interferometry Mission \citep{unwin??}, which was expected to be able to measure proper motions with $\mu$as/year accuracy for stars as faint as $V=20$.
-Distances to stream members would only be poorly known and were estimated in the method by using anticipated properties of tidal debris.
-Below we present an updated version of the algorithm: the promise of 2\% distances to RR Lyrae stars (see Section \ref{sec:spitzer}) enables a direct measurement (rather than approximate estimate) of the position of a star within its debris structure.
-The test statistic that quantifies how well stars recombine with the satellite has been rigorously redefined.
+This approach was originally proposed by \citep{jzsh1999} and was
+designed and tested for the proposed characteristics of the Space
+Interferometry Mission \citep{unwin??}, which was expected to be able
+to measure proper motions with $\mu$as/year accuracy for stars as
+faint as $V=20$. Distances to stream members would only be poorly
+known and were estimated in the method by using anticipated properties
+of tidal debris. Below we present an updated version of the algorithm:
+the promise of 2\% distances to RR Lyrae stars (see Section
+\ref{sec:spitzer}) enables a direct measurement (rather than
+approximate estimate) of the position of a star within its debris
+structure. The test statistic that quantifies how well stars recombine
+with the satellite has been rigorously redefined.
 
@@
 Quantifying this method requires a sample of stars with known full
-space kinematics $(\bs{x}_{i}, \bs{v}_{i})|_{t=0}$ (e.g., measurements of all position and velocity components for these stars \emph{today} at $t=0$), the orbital parameters
-for the progenitor system $(\bs{x}_p, \bs{v}_p)|_{t=0}$, and a
-functional form for the potential, $\Phi({\boldsymbol\theta})$. For a
-given set of potential parameters, $\boldsymbol\theta$, the orbits of the stars and
-progenitor are integrated backwards for several Gigayears (determined from an
-estimate of the interaction time of the system). The stars and
+space kinematics $(\bs{x}_{i}, \bs{v}_{i})|_{t=0}$ (e.g., measurements
+of all position and velocity components for these stars \emph{today}
+at $t=0$), the orbital parameters for the progenitor system
+$(\bs{x}_p, \bs{v}_p)|_{t=0}$, and a functional form for the
+potential, $\Phi({\boldsymbol\theta})$. For a given set of potential
+parameters, $\boldsymbol\theta$, the orbits of the stars and
+progenitor are integrated backwards for several Gigayears (determined
+from an estimate of the interaction time of the system). The stars and
 progenitor system are treated as test particles, so that the debris
-stars do not become physically re-bound to the satellite. Instead, at each
-timestep $t_j$, for each particle $i$, a set of normalized, relative
-phase-space coordinates are
-computed 
+stars do not become physically re-bound to the satellite. Instead, at
+each timestep $t_j$, for each particle $i$, a set of normalized,
+relative phase-space coordinates are computed
 \begin{equation}
@@
 \end{equation}
-where $(\bs{x},\bs{v})_{i}$ and $(\bs{x},\bs{v})_{p}$ are the phase-space coordinates for the particles and progenitor, respectively. These definitions require an estimate of the mass of the
-satellite, $m_{sat}$, which, combined with the orbital radius of the satellite, $R$, and the computed enclosed mass
-of the potential within $R$, $M_{enc}$, sets the instantaneous tidal radius and
-escape velocity,
+where $(\bs{x},\bs{v})_{i}$ and $(\bs{x},\bs{v})_{p}$ are the
+phase-space coordinates for the particles and progenitor,
+respectively. These definitions require an estimate of the mass of the
+satellite, $m_{sat}$, which, combined with the orbital radius of the
+satellite, $R$, and the computed enclosed mass of the potential within
+$R$, $M_{enc}$, sets the instantaneous tidal radius and escape
+velocity,
 \begin{equation}
@@
 \end{equation}
-Qualitatively, when the distance in this normalized six-dimensional space,
-$D_{{\rm ps},i}=\sqrt{|\bs{q}_{i}|^2+|\bs{p}_{i}|^2} < \sqrt{2}$, the star is
-likely recaptured by the satellite.
-JZSH imposed a similar condition as a hard boundary and used the number of recaptured particles in a given backwards-integration as a cost function for optimization.
-What follows is a description of an updated procedure with a more natural, statistically-motivated choice for an objective function.
+Qualitatively, when the distance in this normalized six-dimensional
+space, $D_{{\rm ps},i}=\sqrt{|\bs{q}_{i}|^2+|\bs{p}_{i}|^2} <
+\sqrt{2}$, the star is likely recaptured by the satellite. JZSH
+imposed a similar condition as a hard boundary and used the number of
+recaptured particles in a given backwards-integration as a cost
+function for optimization. What follows is a description of an updated
+procedure with a more natural, statistically-motivated choice for an
+objective function.
 
@@
 \end{align}
-Thus, the matrix $\bs{A}_{ik}$ contains these minimum phase-space distance vectors for each star,
-where $k\in[1,6]$. Intuitively, the variance of this distribution
-will be larger for orbits integrated in an incorrect potential
-relative to the distribution computed from the `true' orbital history
-of the stars: in an incorrect potential, small differences in the
-orbits of the stars relative to the orbit of the progenitor cause them
-to spread out in this relative, normalized phase space. Thus, the
-\emph{generalized variance} of the distribution, computed for a given
-set of potential parameters, $\bs{\theta}$, is a natural choice for
-the scalar objective function used in constraining the potential of
-the Milky Way,
+Thus, the matrix $\bs{A}_{ik}$ contains these minimum phase-space
+distance vectors for each star, where $k\in[1,6]$. Intuitively, the
+variance of this distribution will be larger for orbits integrated in
+an incorrect potential relative to the distribution computed from the
+`true' orbital history of the stars: in an incorrect potential, small
+differences in the orbits of the stars relative to the orbit of the
+progenitor cause them to spread out in this relative, normalized phase
+space. Thus, the \emph{generalized variance} of the distribution,
+computed for a given set of potential parameters, $\bs{\theta}$, is a
+natural choice for the scalar objective function used in constraining
+the potential of the Milky Way,
 \begin{equation}
@@
 from the present-day snapshot of the simulation
-data.\footnote{\url{www.astro.virginia.edu/~srm4n/Sgr/data.html}} We then
-`observe' a sample of 100 stars from the first leading and trailing
-wraps of the stream: the radial velocity errors are drawn from a
-Gaussian $\epsilon_{rv} \sim \mathcal{N}(\mu=0,\sigma=10~{\rm km/s})$,
-the tangential velocity errors depend on the distance to the star and
-are computed from the expected Gaia error curve (shown in
-Figure~\ref{fig:gaia_errors}), and the distance errors on a star of distance
-$D$ are $\epsilon_{dist} \sim \mathcal{N}(0,0.02\times
-D)$. Figure~\ref{fig:variance6d} shows all two-dimensional projections of the six-dimensional minimum
-phase-space distance distribution for the observed particles in the
-correct and an incorrect potential: it is clear that the
-variance of the distribution increases in the wrong potential. The 
-generalized variance defines a convex hypersurface over which we
-optimize four of the six logarithmic potential parameters:
-$v_{circ}$, $\phi$, $q_1$, and $q_z$. We fix $q_2$ and $R_c$, which are degenerate with combinations of the other parameters. Figure~\ref{fig:} shows one-dimensional slices of the objective function
-produced by varying each of the potential parameters by
-$\pm20\%$ around the true values and holding all others fixed.
+data.\footnote{\url{www.astro.virginia.edu/~srm4n/Sgr/data.html}} We
+then `observe' a sample of 100 stars from the first leading and
+trailing wraps of the stream: the radial velocity errors are drawn
+from a Gaussian $\epsilon_{rv} \sim \mathcal{N}(\mu=0,\sigma=10~{\rm
+  km/s})$, the tangential velocity errors depend on the distance to
+the star and are computed from the expected Gaia error curve (shown in
+Figure~\ref{fig:gaia_errors}), and the distance errors on a star of
+distance $D$ are $\epsilon_{dist} \sim \mathcal{N}(0,0.02\times
+D)$. Figure~\ref{fig:variance6d} shows all two-dimensional projections
+of the six-dimensional minimum phase-space distance distribution for
+the observed particles in the correct and an incorrect potential: it
+is clear that the variance of the distribution increases in the wrong
+potential. The generalized variance defines a convex hypersurface over
+which we optimize four of the six logarithmic potential parameters:
+$v_{circ}$, $\phi$, $q_1$, and $q_z$. We fix $q_2$ and $R_c$, which
+are degenerate with combinations of the other
+parameters. Figure~\ref{fig:} shows one-dimensional slices of the
+objective function produced by varying each of the potential
+parameters by $\pm20\%$ around the true values and holding all others
+fixed.
 
@@
 
-Many orbit fitting methods for measuring the Galactic potential assume that the stars in a given stream are on the same
-orbit. \cite{binney??} show that this assumption leads to XXX\% biases in the 
-potential parameters, even without realistic observational errors on
-the kinematic data. The \emph{Rewinder} is
-free of this assumption and instead \emph{relies} on the fact that
-each star is on a different orbit, and is thus an independent potentiometer: the best potential is the one that gets the
-most of these orbits to intersect the progenitor
-over the history of interaction with the system.
-The strength of this method is in the simplicity: it requires only a rough estimate of the satellite mass $m_{\rm sat}$ combined with backwards integration of orbits.
-There are no assumptions made about, or N-body modeling of, the internal distribution of satellite stars.
+Many orbit fitting methods for measuring the Galactic potential assume
+that the stars in a given stream are on the same
+orbit. \cite{binney??} show that this assumption leads to XXX\% biases
+in the potential parameters, even without realistic observational
+errors on the kinematic data. The \emph{Rewinder} is free of this
+assumption and instead \emph{relies} on the fact that each star is on
+a different orbit, and is thus an independent potentiometer: the best
+potential is the one that gets the most of these orbits to intersect
+the progenitor over the history of interaction with the system. The
+strength of this method is in the simplicity: it requires only a rough
+estimate of the satellite mass $m_{\rm sat}$ combined with backwards
+integration of orbits. There are no assumptions made about, or N-body
+modeling of, the internal distribution of satellite stars.
 
-[TODO: this paragraph needs work]
-There are two other apparent strengths of the method which require further investigation.
-\emph{The Rewinder} is applicable to any debris that is known to come from a single object, and not restricted only to the very coldest tidal streams. For example, in principle the
-method could applied to the vast debris {\it clouds} that have been discovered subtending large solid angles \citep[e.g. the Triangulum Andromeda
-and Hercules-Aquila clouds]{rochapinto04,belokurov06}, or even stars that have only associations in orbital properties and do not form a coherent spatial structure
-\citep[such as the angular momentum groupings in local giants found by][]{helmi99}.
-Second, it appears trivial to combine constraints from multiple debris systems at once by simply integrating all debris from several satellites simultaneously with $D_{\rm ps}$ defined appropriately for each star separately.
+[TODO: this paragraph needs work] There are two other apparent
+strengths of the method which require further investigation.
+\emph{The Rewinder} is applicable to any debris that is known to come
+from a single object, and not restricted only to the very coldest
+tidal streams. For example, in principle the method could applied to
+the vast debris {\it clouds} that have been discovered subtending
+large solid angles \citep[e.g. the Triangulum Andromeda and
+  Hercules-Aquila clouds]{rochapinto04,belokurov06}, or even stars
+that have only associations in orbital properties and do not form a
+coherent spatial structure \citep[such as the angular momentum
+  groupings in local giants found by][]{helmi99}. Second, it appears
+trivial to combine constraints from multiple debris systems at once by
+simply integrating all debris from several satellites simultaneously
+with $D_{\rm ps}$ defined appropriately for each star separately.
 
-[present the limitations]
-The limitations of the method also need to be understood. Firstly, for the RR Lyraes associated with the very coldest streams \citep[e.g. the globular clusters Pal5
-and GD1][]{odenkirchen02,koposov10} the measurement errors are unlikely to be small enough for the
-the differences in orbital properties to between debris and satellite to be detected.
-[What will this cause?
-Maybe biases?
-One more sentence here].
-Second, the method requires specifying a functional form for the potential.
-[...].
-Lastly, the current version of the algorithm relies on accurate knowledge of the phase-space position of the parent satellite, which may not be available in some cases \citep[for example, for the Orphan Stream][]{belokurov07}.
+[present the limitations] The limitations of the method also need to
+be understood. Firstly, for the RR Lyraes associated with the very
+coldest streams \citep[e.g. the globular clusters Pal5 and
+  GD1][]{odenkirchen02,koposov10} the measurement errors are unlikely
+to be small enough for the the differences in orbital properties to
+between debris and satellite to be detected. [What will this cause?
+  Maybe biases? One more sentence here]. Second, the method requires
+specifying a functional form for the potential. [...]. Lastly, the
+current version of the algorithm relies on accurate knowledge of the
+phase-space position of the parent satellite, which may not be
+available in some cases \citep[for example, for the Orphan
+  Stream][]{belokurov07}.
 
-[explain how some of them are handled automatically by the forthcoming machinery]
-The objective function presented above is an approximation to a more general likelihood function, described in detail in forthcoming work (Price-Whelan \& Johnston in prep.).
-The results shown in sections xx and yy are meant to illustrate the strength and concept of the method.
+[explain how some of them are handled automatically by the forthcoming
+  machinery] The objective function presented above is an
+approximation to a more general likelihood function, described in
+detail in forthcoming work (Price-Whelan \& Johnston in prep.). The
+results shown in sections xx and yy are meant to illustrate the
+strength and concept of the method.
 
-Given sufficient data, position and motion of the original host could be solved for as additional free parameters,
-but the effectiveness of this approach has yet to be investigated.
+Given sufficient data, position and motion of the original host could
+be solved for as additional free parameters, but the effectiveness of
+this approach has yet to be investigated.
 
@@
 
-This paper outlines an algorithm to measure the Galactic potential that takes maximal advantage of a combination of possible data from the Spitzer and Gaia satellite
-missions that, when combined,  promise full phase-space measurements of RR Lyrae stars around our Galaxy.
-When applied to a synthetic data set generated from debris in an N-body simulation of the destruction of Sgr the method recovers the shape and orientation of the dark
-matter halo component of the potential with ??? accuracy using a sample of ??? stars.
-This success provides strong motivation for: (i) further theoretical work to investigate the power of multiple debris structures with different morphologies and orientations in
-building a comprehensive picture of the Galactic mass distribution; and (ii) a  Spitzer survey of RR Lyrae stars in debris structures around the Milky Way as a ``legacy''
-sample to combine with near-future Gaia data.
+This paper outlines an algorithm to measure the Galactic potential
+that takes maximal advantage of a combination of possible data from
+the Spitzer and Gaia satellite missions that, when combined, promise
+full phase-space measurements of RR Lyrae stars around our Galaxy.
+When applied to a synthetic data set generated from debris in an
+N-body simulation of the destruction of Sgr the method recovers the
+shape and orientation of the dark matter halo component of the
+potential with ??? accuracy using a sample of ??? stars.  This success
+provides strong motivation for: (i) further theoretical work to
+investigate the power of multiple debris structures with different
+morphologies and orientations in building a comprehensive picture of
+the Galactic mass distribution; and (ii) a Spitzer survey of RR Lyrae
+stars in debris structures around the Milky Way as a ``legacy'' sample
+to combine with near-future Gaia data.
 
@@
 
-APW is supported by a National Science Foundation Graduate Research Fellowship under Grant No.\ 11-44155.
+APW is supported by a National Science Foundation Graduate Research
+Fellowship under Grant No.\ 11-44155.
 

### 2013-07-27 scripts/optimize_likelihood.py
still tweaking optimizer params

@@ -35,6 +35,7 @@ x0 = fmin_l_bfgs_b(lambda *args,**kwargs: -ln_posterior(*args,**kwargs),
                   args=(['q1', 'qz'], particles, satellite, t1, t2, 3.),
                   bounds=[(1.2,1.4),(1.2,1.4)],
                   approx_grad=True,
-                  epsilon=1E-3)
+                  epsilon=1E-2,
+                  factr=1E8)
 
 print(x0)

### 2013-07-30 text/paper1/paper.tex
update references and kathryns comments

--- 2013-07-27 text/paper1/paper.tex
@@
 % affiliations
-\altaffiltext{1}{Department of Astronomy, Columbia University, 550 W 120th St., New York, NY 10027, USA}
+\altaffiltext{1}{Department of Astronomy, Columbia University, 550 W 120th St., New York, NY 027, USA}
 \altaffiltext{2}{NSF Graduate Fellow}
@@
 Dark-matter-only simulations of structure formation lead us to expect that these  halos should have a density distribution that can simply be described with a universal
-profile \citep{nfw96} with a variety of triaxial shapes \citep{jing02}.
+profile \citep{navarro96} with a variety of triaxial shapes \citep{jing02}.
 The inclusion of baryons in the simulations tends to soften the triaxiality of the dark matter in the inner regions of the halo \citep[as the disk forms, see e.g.,][]{bailin05}, and
@@
 
-The one place in the Universe where we might hope to dissect a dark matter halo in even more detail is locally for our own Milky Way Galaxy.
+The Milky Way is the best candidate  for dissecting a dark matter halo in great detail.
 Our proximity means that we can use individual stars as tracer populations and hence build much larger samples that probe deeper into the halo than the globular cluster
-and planetary nebula studies of external galaxies.
-Moreover, our internal perspective allows us to build a three-dimensional picture of the mass distribution rather than measuring its two-dimensional projection, as is the
-case for gravitational lensing.
-Hence we can hope to reconstruct not only the density, but also the shape and orientation of the dark matter halo as function of radius well beyond the disk.
-For example, to date thousands of blue-horizontal-branch stars selected from the Sloan Digital Sky Survey \citep[SDSS][]{} have been used to probe mass out to tens of
-kpc \citep{deason12,kafle12}, and estimates with combined tracers extend to 150kpc \citep{deason12b}.
+and planetary nebula studies of external galaxies \citep[e.g.][]{mendez01,cote03}.
+Moreover, our internal perspective allows us to build a three-dimensional picture of the mass distribution rather than measuring its two-dimensional projection.
+%Hence we can hope to reconstruct not only the density, but also the shape and orientation of the dark matter halo as function of radius well beyond the disk.
+For example, to date thousands of blue-horizontal-branch stars selected from the Sloan Digital Sky Survey  have been used to probe mass out to tens of
+kpc \citep[SDSS, see][]{deason12a,kafle12}, and estimates with combined tracers extend to 150kpc \citep{deason12b}.
 \citet{loebman12} demonstrates how even larger stellar samples from current (i.e. SDSS) and future  (Gaia and LSST) stellar surveys will also be sensitive to any
@@
 However, the same surveys that provide large samples of tracers have also revealed the existence of large-scale spatial inhomogeneities in the form of giant stellar
-streams and clouds\citep{newberg01,majewski03,belokurov05}, clearly demonstrating that  a significant fraction of the stellar halo is neither randomly sampling a smooth
+streams and clouds \citep{newberg02,majewski03,belokurov06}, clearly demonstrating that  a significant fraction of the stellar halo is neither randomly sampling a smooth
 orbit distribution nor is fully phase-mixed.
+%The degree of substructure increases with Galactocentric radius \citep{bell08} in agreement with models where the stellar halo is formed purely from accretion events \citep{bullock05,cooper10,rashkov12}.
+%Moreover, even in the smoother, inner halo which is likely to be phase-mixed, significant clumpings in velocity \citep{schlauffman10} and orbital properties \citep{helmi99} have been observed, which indicate a non-smooth component to the orbit distribution.
+%By ``observing'' particles in a sample of simulated stellar halos, \citet{yencho06} showed that, if stellar halos are indeed entirely formed via accretion events, then the expected degree of non-randomness  could lead to systematic biases in mass estimates of order several tens of percent.
+%This encouraging result suggests that, if our expectations for the typical number and size of accretion events contributing to the stellar halo are correct, then the number of orbits effectively explored is sufficiently large that the assumption of random, fully phase-mixed orbits is reasonable for the current samples.
 
-A complimentary method to measure the mass distribution is to instead take full advantage of the {\it non}-random nature of the stellar distribution in our stellar halo ---
-analogous to our exploitation of matter in disks moving on near-circular orbits.
+A complimentary method to measure the mass distribution is to instead take full advantage of the {\it non}-random nature of the stellar distribution in our stellar halo.
+% --- analogous to our exploitation of matter in disks moving on near-circular orbits.
 While we do not know the exact orbits of  stellar debris from satellite destruction \citep[although][proposes a method of orbit reconstruction from limited dimensions of
-observations]{eyre08}, we do know something more about stars in a debris structure than in a random sample: we know that these stars were once all part of the same
+observations]{eyre09b}, we do know something more about stars in a debris structure than in a random sample: we know that these stars were once all part of the same
 object.
@@
 measure of differences in potential $\Phi$ (i.e. $\Phi({\bf x}_1)-\Phi({\bf x}_2)=(v_2^2-v_1^2)/2$).
-In reality, we usually know at most four of the six phase-space co-ordinates for debris stars, often with significant error bars and potential measurements are made by
-fitting orbits to streams \citep[e.g.,][]{helmi05,johnston05,koposov10,law10,lux12}.
+In reality, we usually know at most four of the six phase-space co-ordinates for debris stars, often with significant error bars and potential measurements have been made by
+fitting orbits to streams \citep[e.g.,][]{helmi04,johnston05,koposov10,law10}.
 Moreover, our assumption of debris tracing a single orbit is actually incorrect \citep[see][for discussions of the orbit distribution in tidal debris]{johnston98,helmi99}, with
-systematic changes in orbital energy along debris streams that can lead to systematic biases in our measurements of the Galactic potential \citep{eyre09,varghese11}.
+systematic changes in orbital energy along debris streams that can lead to systematic biases in our measurements of the Galactic potential \citep{eyre09a,varghese11}.
+In particular, \citet{sanders13a} recently demonstrated that this bias is equally problematic for the very thinnest, coldest streams, whose observed properties may be indistinguishable from those of the parent orbit \citep[e.g.
+such as the globular cluster, GD1 --- see][]{koposov10}, as for the much more extended and hotter streams \citep[e.g.
+such as debris from the Sagittarius dwarf galaxy --- see][]{majewski03} where offsets from a single orbit are clearly apparent.
 
@@
 {law05,fellhauer06} with the most recent work for the first time attempting to place constraints on the triaxiality and orientation of the Milky Way's matter distribution
-\citep{law11}.
+\citep{law10}.
+%Studies which first fit the Sgr stream to a single orbit \citep{johnston05,law09} and were subsequently followed up with full N-body simulations \citep{law05,law12} offer another demonstration of the limitation of the former approach as potential parameters are revised when moving to the N-body approach.
 
@@
 For example, using the results of N-body simulations as a guide, simple corrections around a single orbit can be adopted to make model predictions that more closely
-mimic expected debris behavior \citep[e.g.][]{johnston99a,varghese2011}.
+mimic expected debris behavior \citep[e.g.][]{johnston99b,varghese11}.
 
@@
 \citet{binney08} and \citet{penarrubia12} demonstrate conceptually that the distribution of energy and entropy (respectively) in debris will be minimized only for a correct
-assumption of the form of the Galactic potential (although neither tests their method with realistic observational errors???).
+assumption of the form of the Galactic potential.
+\citet{sanders13b} examine the distribution of debris in action-angle co-ordinates, show stars that were once all members of the same object must lie along a single line in angle-frequency space and demonstrate how this constraint can be used as a potential measure.
+%\citet{johnston99a} (hereafter JZSH) showed that backwards integration of the positions of stars and their parent satellite could be used to distinguish Milky Way mass models: only in the correct potential would the stars' and satellite's  path coincide within a Hubble time.The JZSH method was designed and tested for the proposed characteristics of the Space Interferometry Mission \citep{}, which was expected to be able to measure proper motions with $\mu$as/year accuracy for stars as faint as $V=20$.
+Distances to stream members would only be poorly known and were estimated using anticipated properties of tidal debris.
 
-In this paper we re-examine and update  the JZSH  in the context of current and near-future observational capabilities, and illustrates its power with the example of Sgr.
+In this paper we re-examine and update  another approach to using tidal debris as a potential measure \citep[originially proposed by][]{johnston99a}  in the context of current and near-future observational capabilities, and illustrates its power with the example of Sgr.
 In Section \ref{context.sec} we outline the observational prospects and Sgr properties that motivated this re-examination.
@@
 
-\subsection{2\% distance errors for RR Lyrae in the halo}
+\subsection{{\it Spitzer} and 2\% distance errors for RR Lyrae in the halo}
 \label{sec:spitzer}
-Over twenty-five years ago, \citet{longmore1986} demonstrated the
-existence of a near-infrared period-luminosity (PL) relation for RR
-Lyrae stars using $K$-band observations of stars in Galactic
-globular clusters. More recently, \citet{catelan2004} showed that a PL
-relation for RR Lyrae stars is expected in the infrared (but
-vanishingly so in the optical) and is explained primarily by the
+
+There is a long tradition for using RR Lyrae stars in the Galaxy to study structure
+\citep[going back to first estimates of the distance to the Galactic center][]{shapley18},
+substructure \citep[e.g.][]{sesar10} and distances to satellite galaxies [e.g.][]{clementini03}.
+However, studies of RR Lyrae at optical wavelengths are limited by both metallicity affects on the intrinsic 
+brightness of these stars and variable extinction along the line of sight.
+Moreover, systematic differences between instruments make it difficult to tie observations across the sky
+to a common scale. 
+Hence, while among the best distance indicators available to us today, the absolute uncertainty in
+individual estimates are typically no better than 10\%
+
+At longer wavelengths, RR Lyrae promise tighter constraints on distances.
+The near-infrared period-luminosity (PL) relation for RR
+Lyrae stars was first mapped by \citet{longmore86} 
+using $K$-band observations of stars in Galactic
+globular clusters and later shown to be due to
+the
 changing bolometric correction for these stars as a function of the
-observed wavelength. Later, \citet{madore2012} demonstrated the
-inevitability of decreasing intrinsic scatter with increasing
-wavelength (and slope) of PL relations both for Cepheids and for RR
-Lyrae stars. Both reddening effects and metallicity dependencies also
+observed wavelength \citep{catelan04}.
+ \citet{madore12} has recently shown how the intrinsic scatter of the PL relation
+ decreases  with increasing
+wavelength. Both reddening effects and metallicity dependencies also
 decrease significantly for observations further into the IR.
@@
 NASA's Spitzer mission could provide us with a stellar distance
-indicator of unprecedented accuracy and precision. \citet{madore2012}
-have demonstrated \citep[using five stars with trigonometric parallaxes
-  measured by Hubble;][]{benedict2011} that the dispersion in the
-mid-IR PL relation -- at Spitzer wavelengths -- is of order
-$\sim$0.03~mag. This implies that it is possible to determine
-individual distances that are good to better than $2\%$ for RR
-Lyrae stars out to $\sim$80~kpc (the limit of Spitzer's ability to
+indicator of unprecedented accuracy and precision.
+\citet{madore12}
+For example, using 5 stars with trigonometric parallaxes
+  measured by Hubble \citep{benedict11}, \citet{madore12} measured  
+  the dispersion in the
+mid-IR PL relation -- at Spitzer wavelengths --  to be of order
+$\pm$0.03~mag. This implies that it is possible to determine
+individual distances that are good to better than $\pm 2\%$ for RR
+Lyrae stars out to $\sim$60~kpc (the limit of Spitzer's ability to
 detect and measure RR Lyrae variables, for modest integration
-times). Distance measurements of Blue Horizontal Branch and Blue Straggler stars at these distances in the halo typically achieve $\sim$25\% uncertainties \citep[e.g.,][]{deason2012}.
+times). Distance measurements of Blue Horizontal Branch and Blue Straggler stars at these distances in the halo typically achieve $\sim$25\% uncertainties \citep[e.g.,][]{deason12b}.
 
@@
 \label{sec:gaia}
-The Gaia satellite \citep{gaia2001}, planned for launch in late
-2013, is an astrometric mission which aims to measure the positions of billions of stars with
+The Gaia satellite \citep{gaia01}, planned for launch in late
+13, is an astrometric mission which aims to measure the positions of billions of stars with
 10-100~$\mu$as accuracies. With over 50 observations of the entire sky
@@
 tangential velocity error estimates for RR Lyrae.
-The velocity uncertainty depends on the magnitude of the source, which depends on the metallicity: each line in Figure~\ref{fig:gaia_errors} is computed by Monte Carlo sampling from the empirical metallicity distribution of the Galactic halo from \cite{ivezic2008} (centered around [Fe/H] $\sim$ -1.5).
+The velocity uncertainty depends on the magnitude of the source, which depends on the metallicity: each line in Figure~\ref{fig:gaia_errors} is computed by Monte Carlo sampling from the empirical metallicity distribution of the Galactic halo from \cite{ivezic08} (centered around [Fe/H] $\sim$ -1.5).
 Within 2~kpc, Gaia
@@
 serendipitously during a radial velocity survey of the Galactic bulge
-\citep{ibata1994}. While the number of known Milky Way satellites has
-more than doubled recently \citep[e.g.,][]{mcconnachie2012}, in many
+\citep{ibata94}. While the number of known Milky Way satellites has
+more than doubled recently \citep[e.g.,][]{mcconnachie12}, in many
 ways Sgr remains the most studied and most intriguing: it is the
@@
 
-The first maps of the Sgr core \citep{ibata1994, ibata1995} revealed
+The first maps of the Sgr core \citep{ibata94, ibata95} revealed
 its elongated morphology and suggested tidal deformation, as would be
 anticipated from its proximity to the Galaxy. Subsequent N-body models
-confirmed this interpretation \citep{velazquez1995} and showed that
+confirmed this interpretation \citep{velazquez95} and showed that
 debris from Sgr should form coherent streams of stars that encircle
-the Galaxy \citep{johnston1995}. Signatures of these extensive stellar
+the Galaxy \citep{johnston95}. Signatures of these extensive stellar
 streams have since been discovered in many studies and have been
-mapped across the sky in carbon stars \citep{totten1998}, M giants
-selected from the Two Micron All Sky Survey \citep{majewski2003}, main
+mapped across the sky in carbon stars \citep{totten98}, M giants
+selected from the Two Micron All Sky Survey \citep{majewski03}, main
 sequence turnoff stars from the Sloan Digital Sky Survey
-\citep{belokurov2006}, and RR Lyrae in the Catalina Sky Survey
-\citep{drake2013}.
+\citep{belokurov06}, and RR Lyrae in the Catalina Sky Survey
+\citep{drake13}.
 
 The distance to the stream has been estimated from photometric surveys
-\citep[e.g.,][]{martinezdelgado2004} and gradients in radial
+\citep[e.g.,][]{martinezdelgado04} and gradients in radial
 velocities and metallicity measured with follow-up spectroscopy
-\citep[e.g.,][]{majewski2004, vivas2005, bellazzini2006, chou2007,
-  chou2010, keller2010, carlin2012}. The multitude of Sgr stream data
+\citep[e.g.,][]{majewski04, vivas05, bellazzini06, chou07,
+  chou10, keller10, carlin12}. The multitude of Sgr stream data
 has inspired a rich set of interpretations and models of the debris
-\citep[e.g.,][]{johnston1999, helmi2004, law2005, fellhauer2006,
-  law2010}.
+\citep[e.g.,][]{johnston99b, helmi04, law05, fellhauer06,
+  law10}.
 
-Most recently, \citet[][hereafter LM10]{law2010} have combined all
+Most recently, \citet[][hereafter LM10]{law10} have combined all
 current data on the Sgr debris to assess the triaxiality and orientation
@@
 orbit in the best-fitting Milky Way halo model from LM10. [TODO: Maybe not true...: The positions are shown in a coordinate system defined
-by the orbit of the satellite \citep{majewski2004}]. The
+by the orbit of the satellite \citep{majewski04}]. The
 simulation was run in a three-component potential, with a triaxial,
@@
 matter halo that cannot be matched using other techniques on our own
-or other galaxies \citep[e.g.,][]{ibata2001}. The LM10 constraints on
+or other galaxies \citep[e.g.,][]{ibata01}. The LM10 constraints on
 Sgr's mass and orbit have invigorated discussions of Sgr in a more
@@
 halo that hosted it, as well as its effect on the Milky Way itself
-\citep{bailin2003,purcell2011,micheldansac2011,gomez2012}.
+\citep{bailin03,purcell11,micheldansac11,gomez12}.
 
@@
 accuracies ($\epsilon_{v}\sim2$~km/s at 30~kpc) less than the intrinsic dispersion
-for much of the stream \citep[$\sigma_v\sim10$~km/s;][]{majewski2004}, and ground-based radial
+for much of the stream \citep[$\sigma_v\sim10$~km/s;][]{majewski04}, and ground-based radial
 velocities, the measured phase-space
@@
 
-This approach was originally proposed by \citep{jzsh1999} and was designed and tested for the proposed characteristics of the Space Interferometry Mission \citep{unwin??}, which was expected to be able to measure proper motions with $\mu$as/year accuracy for stars as faint as $V=20$.
+This approach was originally proposed by \citep{johnston99a} and was designed and tested for the proposed characteristics of the Space Interferometry Mission \citep{unwin??}, which was expected to be able to measure proper motions with $\mu$as/year accuracy for stars as faint as $V=20$.
 Distances to stream members would only be poorly known and were estimated in the method by using anticipated properties of tidal debris.
@@
 Many orbit fitting methods for measuring the Galactic potential assume that the stars in a given stream are on the same
-orbit. \cite{binney??} show that this assumption leads to XXX\% biases in the 
+orbit. \cite{binney08} show that this assumption leads to XXX\% biases in the 
 potential parameters, even without realistic observational errors on

### 2013-07-31 text/paper1/paper.tex
revisions

@@
 %Hence we can hope to reconstruct not only the density, but also the shape and orientation of the dark matter halo as function of radius well beyond the disk.
-For example, to date thousands of blue-horizontal-branch stars selected from the Sloan Digital Sky Survey  have been used to probe mass out to tens of
+For example, to date thousands of blue-horizontal-branch stars selected from the Sloan Digital Sky Survey (SDSS) have been used to probe mass out to tens of
 kpc \citep[SDSS, see][]{deason12a,kafle12}, and estimates with combined tracers extend to 150kpc \citep{deason12b}.
@@
 In this paper we re-examine and update  another approach to using tidal debris as a potential measure \citep[originially proposed by][]{johnston99a}  in the context of current and near-future observational capabilities, and illustrates its power with the example of Sgr.
-In Section \ref{context.sec} we outline the observational prospects and Sgr properties that motivated this re-examination.
-In Section \ref{measure.sec} we present the updated potential measure and test it with synthetic observations of simulated Sgr debris.
-In Section \ref{summary.sec} we summarize our results and outline prospects for combining constraints from Sgr with those from other debris systems.
+In Section \ref{sec:context} we outline the observational prospects and Sgr properties that motivated this re-examination.
+In Section \ref{sec:method} we present the updated potential measure and test it with synthetic observations of simulated Sgr debris.
+In Section \ref{sec:discussion} we highlight the advantages and shortcomings of this method.
+We conclude in Section \ref{sec:conclusion}.
 
@@
 
-\section{Context and motivation}
+\section{Context and motivation} \label{sec:context}
 The method presented in Section \ref{sec:method} takes advantage of
@@
 mid-IR PL relation -- at Spitzer wavelengths --  to be of order
-$\pm$0.03~mag. This implies that it is possible to determine
-individual distances that are good to better than $\pm 2\%$ for RR
+$\sim$0.03~mag. This implies that it is possible to determine
+individual distances that are good to better than $2\%$ for RR
 Lyrae stars out to $\sim$60~kpc (the limit of Spitzer's ability to
@@
 \citep{drake13}.
+More recently, apogalactic debris has been traced to large distances using a combination of blue horizontal branch, main-sequence turn-off, and red giant stars from the SDSS \citep{belokurov13}.
 
@@
 the scalar objective function used in constraining the potential of
-the Milky Way,
+the Milky Way. If
 \begin{equation}
-  f(\bs{\theta})=\det(\mathrm{Cov}( \bs{A}_{ik})).
+  f(\bs{\theta})=\log \det(\mathrm{Cov}( \bs{A}_{ik})).
 \end{equation}
 
-\subsection{Application to Simulated Data}
+[Still working on this section...need to figure out best way to express what I'm trying to say.] Though this is not explicitly a \emph{likelihood}, optimizing over this statistic is a very simple approximation to performing a maximum-likelihood optimization.
+Assuming the samples in $\bs{A}_{ik}$ are approximately normal-distributed, the variance is closely related to the Shannon entropy \citep{shannon1948}, $H$,
+\begin{equation}
+  H \propto \log \sigma .
+\end{equation}
+Minimizing the entropy is then equivalent to minimizing the loss of \emph{information} for a given choice of parameters, $\bs{\theta}$, given the ``data,'' $\bs{A}_{ik}$, and is thus related to maximum-likelihood estimation \citep[see][for more detail]{mackay2002}.
+
+\subsection{Application to Simulated Data} \label{sec:results}
 The LM10 simulation data (see Section \ref{sec:sgr}) is a perfect
@@
 produced by varying each of the potential parameters by
-$\pm20\%$ around the true values and holding all others fixed.
+$\pm10\%$ around the true values and holding all others fixed.
 
-[Bootstrap]
+In anticipation of extending the above method to include a true likelihood function, we use a parallelized Markov Chain Monte Carlo (MCMC) algorithm \citep{foremanmackey2013} to sample from our objective function.\footnote{Though MCMC is typically not an efficient optimization tool, the stochasticity and easy parallelization of the algorithm mean it easily outperforms other optimizers on this problem, where the objective function is noisy and expensive to compute.} We use the median value of the samples as a point estimate for the potential parameters.
+To assess the uncertainty in the derived halo parameters, we bootstrap resample our 100 stars 100 times and estimate the potential parameters with each sample.
+Figure~\ref{fig:bootstrap} shows the recovered parameters for each bootstrapped sample and demonstrates the power of the combined Spitzer and Gaia data: a moderately sized sample of RR Lyrae alone places strong constraints on the shape and mass of the Milky Way potential.
 
@@
 \section{Discussion and future prospects} 
+\label{sec:discussion}
 
 Many orbit fitting methods for measuring the Galactic potential assume that the stars in a given stream are on the same
-orbit. \cite{binney08} show that this assumption leads to XXX\% biases in the 
-potential parameters, even without realistic observational errors on
+orbit. \cite{sanders13a} show that this assumption can lead to nearly 100\% errors in the 
+estimated potential parameters, even without realistic observational errors on
 the kinematic data. The \emph{Rewinder} is
 free of this assumption and instead \emph{relies} on the fact that
-each star is on a different orbit, and is thus an independent potentiometer: the best potential is the one that gets the
+each star is on a different orbit, and is thus an independent potentiometer; the best potential is the one that gets the
 most of these orbits to intersect the progenitor
 over the history of interaction with the system.
-The strength of this method is in the simplicity: it requires only a rough estimate of the satellite mass $m_{\rm sat}$ combined with backwards integration of orbits.
+
+The strengths of this method stem from the simplicity: it requires only a rough estimate of the satellite mass $m_{\rm sat}$ combined with backwards integration of orbits.
 There are no assumptions made about, or N-body modeling of, the internal distribution of satellite stars.
-
-[TODO: this paragraph needs work]
-There are two other apparent strengths of the method which require further investigation.
-\emph{The Rewinder} is applicable to any debris that is known to come from a single object, and not restricted only to the very coldest tidal streams. For example, in principle the
-method could applied to the vast debris {\it clouds} that have been discovered subtending large solid angles \citep[e.g. the Triangulum Andromeda
+Thus, \emph{The Rewinder} is applicable to any debris that is known to come from a single object, and not restricted only to the very coldest tidal streams.
+In principle, the
+method could applied to the vast debris {\it clouds} that have been discovered over large areas of the sky \citep[e.g. the Triangulum Andromeda
 and Hercules-Aquila clouds]{rochapinto04,belokurov06}, or even stars that have only associations in orbital properties and do not form a coherent spatial structure
 \citep[such as the angular momentum groupings in local giants found by][]{helmi99}.
-Second, it appears trivial to combine constraints from multiple debris systems at once by simply integrating all debris from several satellites simultaneously with $D_{\rm ps}$ defined appropriately for each star separately.
+The method trivially extends to combining constraints from multiple debris systems at once by simply integrating all debris from several satellites simultaneously, with $D_{\rm ps}$ defined appropriately for each star.
 
-[present the limitations]
-The limitations of the method also need to be understood. Firstly, for the RR Lyraes associated with the very coldest streams \citep[e.g. the globular clusters Pal5
-and GD1][]{odenkirchen02,koposov10} the measurement errors are unlikely to be small enough for the
-the differences in orbital properties to between debris and satellite to be detected.
-[What will this cause?
-Maybe biases?
-One more sentence here].
-Second, the method requires specifying a functional form for the potential.
-[...].
-Lastly, the current version of the algorithm relies on accurate knowledge of the phase-space position of the parent satellite, which may not be available in some cases \citep[for example, for the Orphan Stream][]{belokurov07}.
+It is also important to characterize the the limitations of this method.
+Firstly, the measurement errors for RR Lyraes associated with the very coldest streams \citep[e.g.
+the globular clusters Pal5 and GD1][]{odenkirchen02,koposov10} will likely be too large to resolve the minute differences in orbital properties between the debris and satellite to be detected.
+Second, the current version of the algorithm relies on accurate knowledge of the current position and velocity of the parent satellite, which may not be available in some cases \citep[for example, the Orphan Stream;][]{belokurov07}.
+Lastly, the present prescription treats the progenitor as a point-mass during orbit integration.
+Second-order effects in orbit evolution (e.g., dynamical friction) are currently ignored, leading to minor biases in the estimated potential parameters (for example, XXX in Figure~\ref{fig:bootstrap}).
 
-[explain how some of them are handled automatically by the forthcoming machinery]
-The objective function presented above is an approximation to a more general likelihood function, described in detail in forthcoming work (Price-Whelan \& Johnston in prep.).
-The results shown in sections xx and yy are meant to illustrate the strength and concept of the method.
+The objective function presented above is an approximation to a more general likelihood function, described in depth in forthcoming work (Price-Whelan \& Johnston in prep.).
+The results shown in section \ref{sec:results} are meant to illustrate the strength and concept behind the method, but [...] [I want to say one more sentence about the upcoming work, and possibly how it will solve some of the limitations presented above.]
 
-Given sufficient data, position and motion of the original host could be solved for as additional free parameters,
-but the effectiveness of this approach has yet to be investigated.
+%With a justified probabilistic framework, we will also be able to provide iven sufficient data, position and motion of the original host could be solved for as additional free parameters, but the effectiveness of this approach has yet to be investigated.
 
-\section{Summary}
+\section{Conclusions}
+\label{sec:conclusion}
 
-This paper outlines an algorithm to measure the Galactic potential that takes maximal advantage of a combination of possible data from the Spitzer and Gaia satellite
-missions that, when combined,  promise full phase-space measurements of RR Lyrae stars around our Galaxy.
-When applied to a synthetic data set generated from debris in an N-body simulation of the destruction of Sgr the method recovers the shape and orientation of the dark
-matter halo component of the potential with ??? accuracy using a sample of ??? stars.
-This success provides strong motivation for: (i) further theoretical work to investigate the power of multiple debris structures with different morphologies and orientations in
-building a comprehensive picture of the Galactic mass distribution; and (ii) a  Spitzer survey of RR Lyrae stars in debris structures around the Milky Way as a ``legacy''
-sample to combine with near-future Gaia data.
+This paper presents an algorithm for measuring the Galactic potential that anticipates combined data from the Spitzer and Gaia satellite missions that promise precise, full phase-space measurements of RR Lyrae stars in the halo of our Galaxy.
+When applied to a sample of 100 stars -- with realistic observational errors -- sampled from the \cite{law10} N-body simulation of the destruction of the Sgr dwarf satellite, the method recovers the shape and orientation of the dark matter potential XXX.
+This success provides strong motivation for 1) further theoretical work to [...], 2) an investigation of the power of using multiple, complimentary debris structures (e.g., with different morphologies and orientations) in building a comprehensive picture of the Galactic mass distribution, and 3) a Spitzer survey of RR Lyrae stars in debris structures around the Milky Way to get precise distances to combine with near-future Gaia velocity data.
 

### 2013-07-31 text/paper1/paper.tex
SIM reference, minor tweaks

@@
 \citep[going back to first estimates of the distance to the Galactic center][]{shapley18},
-substructure \citep[e.g.][]{sesar10} and distances to satellite galaxies [e.g.][]{clementini03}.
+substructure \citep[e.g.][]{sesar10}, and distances to satellite galaxies \citep[e.g.][]{clementini03}.
 However, studies of RR Lyrae at optical wavelengths are limited by both metallicity affects on the intrinsic 
@@
 Moreover, systematic differences between instruments make it difficult to tie observations across the sky
-to a common scale. 
-Hence, while among the best distance indicators available to us today, the absolute uncertainty in
-individual estimates are typically no better than 10\%
+to a common scale. While RR Lyrae are among the best distance indicators available to us today, the absolute uncertainty in
+individual estimates are typically $\gtrsim$10\%.
 
@@
 observed wavelength \citep{catelan04}.
- \citet{madore12} has recently shown how the intrinsic scatter of the PL relation
+ \citet{madore12} have recently shown that the intrinsic scatter of the PL relation
  decreases  with increasing
@@
 These studies suggest that observations of RR Lyrae variables by
-NASA's Spitzer mission could provide us with a stellar distance
-indicator of unprecedented accuracy and precision.
-\citet{madore12}
-For example, using 5 stars with trigonometric parallaxes
+NASA's Spitzer mission could provide stellar distance
+indicators of unprecedented accuracy and precision.
+Using five stars with trigonometric parallaxes
   measured by Hubble \citep{benedict11}, \citet{madore12} measured  
   the dispersion in the
-mid-IR PL relation -- at Spitzer wavelengths --  to be of order
-$\sim$0.03~mag. This implies that it is possible to determine
+mid-IR PL relation at Spitzer wavelengths to be $\sim$0.03~mag. This implies that it is possible to determine
 individual distances that are good to better than $2\%$ for RR
 Lyrae stars out to $\sim$60~kpc (the limit of Spitzer's ability to
-detect and measure RR Lyrae variables, for modest integration
+detect and measure RR Lyrae variables for modest integration
 times). Distance measurements of Blue Horizontal Branch and Blue Straggler stars at these distances in the halo typically achieve $\sim$25\% uncertainties \citep[e.g.,][]{deason12b}.
@@
 The Gaia satellite \citep{gaia01}, planned for launch in late
-13, is an astrometric mission which aims to measure the positions of billions of stars with
+2013, is an astrometric mission which aims to measure the positions of billions of stars with
 10-100~$\mu$as accuracies. With over 50 observations of the entire sky
-during the 5-year lifetime of the mission, Gaia will
+during the five-year lifetime of the mission, Gaia will
 revolutionize the field of Galactic Astronomy by providing sufficient
@@
 
-Figure~\ref{fig:gaia_errors} show the Gaia end-of-mission distance and
+Figure~\ref{fig:gaia_errors} shows the Gaia end-of-mission distance and
 tangential velocity error estimates for RR Lyrae.
@@
 Within 2~kpc, Gaia
-will be able to measure distances to these stars with better than 2\%
-accuracy---RR Lyraes in this volume will be used to test and calibrate
+will measure distances to these stars with better than 2\%
+accuracy -- RR Lyrae in this volume will be used to test and calibrate
 the Spitzer PL relation described above. Beyond the 2~kpc threshold,
-the mid-IR PL relation for RR Lyraes will provide more accurate
+the mid-IR PL relation for RR Lyrae will provide more accurate
 distance measurements.
@@
 tangential velocity errors of order $\sim$1~km/s at 30~kpc, and of
-order $\sim$10~km/s out to several tens of kiloparsecs. The
-combination of Spitzer and Gaia data will extend the `horizon' of
+order $\sim$10~km/s out to several tens of kiloparsecs.\footnote{\url{http://www.rssd.esa.int/index.php?page=Science_Performance&project=GAIA}} The
+combination of Spitzer and Gaia data will extend the ``horizon'' of
 where full six-dimensional phase-space maps of the Galaxy are possible
@@
 the 3-dimensional mass distribution of a dark matter halo has been
-feasible. Figure \ref{fig:lm10} shows particle positions from the final time-step of an N-body simulation of dwarf satellite disruption along the expected Sgr
-orbit in the best-fitting Milky Way halo model from LM10. [TODO: Maybe not true...: The positions are shown in a coordinate system defined
-by the orbit of the satellite \citep{majewski04}]. The
+feasible. Figure \ref{fig:lm10} shows particle positions (in Galactocentric coordinates) from the final time-step of an N-body simulation of dwarf satellite disruption along the expected Sgr
+orbit in the best-fitting Milky Way halo model from LM10. The
 simulation was run in a three-component potential, with a triaxial,
@@
 its mass and stellar populations, orbit, rate of destruction, and
-even original distribution in populations [other refs??]. The phase-space distribution of the debris offers a unique probe of
+even original distribution in populations. The phase-space distribution of the debris also offers a unique probe of
 the depth, shape, extent, and orientation of the Milky Way's dark
@@
 position of each individual RR Lyrae becomes a
-powerful probe of the potential (see \ref{fig:}). The next section outlines a new
+powerful probe of the potential. The next section outlines a new
 method to take advantage of this information.
@@
 orbit of the progenitor are integrated \emph{backwards} in time
-(imagine watching a visualization of a simulation of satellite destruction "rewind"), the
-stars should `recombine' with the progenitor. If the potential is
+(imagine watching satellite destruction in ``rewind"), the
+stars should recombine with the progenitor. If the potential is
 incorrect, the orbits of the stars will diverge from that of the
@@
 
-This approach was originally proposed by \citep{johnston99a} and was designed and tested for the proposed characteristics of the Space Interferometry Mission \citep{unwin??}, which was expected to be able to measure proper motions with $\mu$as/year accuracy for stars as faint as $V=20$.
+This approach was originally proposed by \citep{johnston99a} and was designed and tested for the proposed characteristics of the Space Interferometry Mission \citep{unwin08}, which was expected to be able to measure proper motions with $\mu$as/year accuracy for stars as faint as $V=20$.
 Distances to stream members would only be poorly known and were estimated in the method by using anticipated properties of tidal debris.
@@
   \bs{q}_{i} = \frac{\bs{x}_{i} -
-    \bs{x}_{p}}{R_{\rm tide}}\,\,\,\,;\,\,\,\,\bs{p}_{i} = \frac{\bs{v}_{i} -
+    \bs{x}_{p}}{R_{\rm tide}}\,\,\,\,,\,\,\,\,\bs{p}_{i} = \frac{\bs{v}_{i} -
     \bs{v}_{p}}{v_{\rm esc}}
@@
 \begin{equation}
-  R_{\rm tide}=R\Big(\frac{m_{\rm sat}}{3M_{\rm enc}}\Big)^{1/3}\,\,\,\,;\,\,\,\,
+  R_{\rm tide}=R\Big(\frac{m_{\rm sat}}{3M_{\rm enc}}\Big)^{1/3}\,\,\,\,,\,\,\,\,
   v_{\rm esc}=\sqrt{\frac{2Gm_{\rm sat}}{R_{\rm tide}}}.
@@
 likely recaptured by the satellite.
-JZSH imposed a similar condition as a hard boundary and used the number of recaptured particles in a given backwards-integration as a cost function for optimization.
+\citet{johnston99a} (hereafter JZSH) imposed a similar condition as a hard boundary and used the number of recaptured particles in a given backwards-integration as a cost function for optimization.
 What follows is a description of an updated procedure with a more natural, statistically-motivated choice for an objective function.
@@
 Thus, the matrix $\bs{A}_{ik}$ contains these minimum phase-space distance vectors for each star,
-where $k\in[1,6]$. Intuitively, the variance of this distribution
+where $k\in[1,6]$. Intuitively, the variance of the distribution of minimum phase-space vectors
 will be larger for orbits integrated in an incorrect potential
@@
 to spread out in this relative, normalized phase space. Thus, the
-\emph{generalized variance} of the distribution, computed for a given
-set of potential parameters, $\bs{\theta}$, is a natural choice for
+\emph{generalized variance} of the distribution -- computed for a given
+set of potential parameters, $\bs{\theta}$ -- is a natural choice for
 the scalar objective function used in constraining the potential of
-the Milky Way. If
-\begin{equation}
-  f(\bs{\theta})=\log \det(\mathrm{Cov}( \bs{A}_{ik})).
-\end{equation}
+the Milky Way 
+\begin{align}
+  \Sigma_n &= \mathrm{Cov}( \bs{A}_{ik}) \\
+  f(\bs{\theta}) &= \ln \det \Sigma.
+\end{align}
 
-[Still working on this section...need to figure out best way to express what I'm trying to say.] Though this is not explicitly a \emph{likelihood}, optimizing over this statistic is a very simple approximation to performing a maximum-likelihood optimization.
+[Still working on this bit...need to figure out best way to express what I'm trying to say.] Though this is not explicitly a \emph{likelihood}, optimizing over this statistic is a very simple approximation to performing a maximum-likelihood optimization.
 Assuming the samples in $\bs{A}_{ik}$ are approximately normal-distributed, the variance is closely related to the Shannon entropy \citep{shannon1948}, $H$,

### 2013-07-31 text/paper1/paper.tex
added figure placeholders

@@
-%\documentclass{emulateapj}
-\documentclass[preprint]{aastex}
+\documentclass{emulateapj}
+%\documentclass[preprint]{aastex}
+
 %\documentclass{aastex}
@@
 $D$ are $\epsilon_{dist} \sim \mathcal{N}(0,0.02\times
-D)$. Figure~\ref{fig:variance6d} shows all two-dimensional projections of the six-dimensional minimum
+D)$. Figure~\ref{fig:variance_proj} shows two-dimensional projections of the six-dimensional minimum
 phase-space distance distribution for the observed particles in the
@@
 optimize four of the six logarithmic potential parameters:
-$v_{circ}$, $\phi$, $q_1$, and $q_z$. We fix $q_2$ and $R_c$, which are degenerate with combinations of the other parameters. Figure~\ref{fig:} shows one-dimensional slices of the objective function
+$v_{circ}$, $\phi$, $q_1$, and $q_z$.
+We fix $q_2$ and $R_c$, which are degenerate with combinations of the other parameters.
+Figure~\ref{fig:objective} shows one-dimensional slices of the objective function
 produced by varying each of the potential parameters by
@@
 
-In anticipation of extending the above method to include a true likelihood function, we use a parallelized Markov Chain Monte Carlo (MCMC) algorithm \citep{foremanmackey2013} to sample from our objective function.\footnote{Though MCMC is typically not an efficient optimization tool, the stochasticity and easy parallelization of the algorithm mean it easily outperforms other optimizers on this problem, where the objective function is noisy and expensive to compute.} We use the median value of the samples as a point estimate for the potential parameters.
-To assess the uncertainty in the derived halo parameters, we bootstrap resample our 100 stars 100 times and estimate the potential parameters with each sample.
-Figure~\ref{fig:bootstrap} shows the recovered parameters for each bootstrapped sample and demonstrates the power of the combined Spitzer and Gaia data: a moderately sized sample of RR Lyrae alone places strong constraints on the shape and mass of the Milky Way potential.
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=1.\textwidth]{/Users/adrian/projects/streams/plots/paper1/variance_projections.pdf}
+\caption{ todo }\label{fig:variance_proj}
+\end{center}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=1.\textwidth]{/Users/adrian/projects/streams/plots/paper1/objective_function.pdf}
+\caption{ todo }\label{fig:objective}
+\end{center}
+\end{figure}
+
+In anticipation of extending the above method to include a true likelihood function, we use a parallelized Markov Chain Monte Carlo (MCMC) algorithm \citep{foremanmackey2013} to sample from our objective function.\footnote{Though MCMC is typically not an efficient optimization tool, in this case objective function is both noisy and expensive to compute.
+The stochasticity and easy parallelization of the algorithm mean it outperforms other optimizers on this problem.} We use the median value of the (burnt-in) sample distribution as a point estimate for the potential parameters.
+To assess the uncertainty in the derived halo parameters, we bootstrap resample our 100 stars 100 times and estimate the potential parameters with each resampling.
+Figure~\ref{fig:bootstrap} shows the recovered parameters for each bootstrapped sample and demonstrates the power of the combined Spitzer and Gaia data: a moderately sized sample of RR Lyrae alone places quite strong constraints on the shape and mass of the Galaxy's dark matter halo.
 
@@
 free of this assumption and instead \emph{relies} on the fact that
-each star is on a different orbit, and is thus an independent potentiometer; the best potential is the one that gets the
-most of these orbits to intersect the progenitor
-over the history of interaction with the system.
+each star is on a different orbit, and is thus an independent potentiometer. Qualitatively, the best potential is the one that gets the
+most of these orbits to intersect the progenitor over the history of interaction with the system.
 
-The strengths of this method stem from the simplicity: it requires only a rough estimate of the satellite mass $m_{\rm sat}$ combined with backwards integration of orbits.
+The strengths of this method stem from its simplicity: it requires only a rough estimate of the satellite mass $m_{\rm sat}$ combined with backwards integration of orbits.
 There are no assumptions made about, or N-body modeling of, the internal distribution of satellite stars.
-Thus, \emph{The Rewinder} is applicable to any debris that is known to come from a single object, and not restricted only to the very coldest tidal streams.
+Thus, the \emph{Rewinder} is applicable to any debris that is known to come from a single object and not restricted to the coldest tidal streams.
 In principle, the
-method could applied to the vast debris {\it clouds} that have been discovered over large areas of the sky \citep[e.g. the Triangulum Andromeda
+method could also be applied to the vast debris {\it clouds} that have been discovered over large areas of the sky \citep[e.g. the Triangulum Andromeda
 and Hercules-Aquila clouds]{rochapinto04,belokurov06}, or even stars that have only associations in orbital properties and do not form a coherent spatial structure
@@
 It is also important to characterize the the limitations of this method.
-Firstly, the measurement errors for RR Lyraes associated with the very coldest streams \citep[e.g.
-the globular clusters Pal5 and GD1][]{odenkirchen02,koposov10} will likely be too large to resolve the minute differences in orbital properties between the debris and satellite to be detected.
+Firstly, the measurement errors for RR Lyraes associated with the very coldest streams \citep[e.g., the globular clusters Pal5 and GD1;][]{odenkirchen02,koposov10} will likely be too large to resolve the minute differences in orbital properties between the debris and satellite to be detected.
 Second, the current version of the algorithm relies on accurate knowledge of the current position and velocity of the parent satellite, which may not be available in some cases \citep[for example, the Orphan Stream;][]{belokurov07}.
@@
 Second-order effects in orbit evolution (e.g., dynamical friction) are currently ignored, leading to minor biases in the estimated potential parameters (for example, XXX in Figure~\ref{fig:bootstrap}).
+
+[Ok, there is a question as to whether we even want to say something like the below...?]
 

### 2013-07-31 text/paper1/paper.tex
wrapping

@@
 
-There is a long tradition for using RR Lyrae stars in the Galaxy to study structure
-\citep[going back to first estimates of the distance to the Galactic center][]{shapley18},
-substructure \citep[e.g.][]{sesar10}, and distances to satellite galaxies \citep[e.g.][]{clementini03}.
-However, studies of RR Lyrae at optical wavelengths are limited by both metallicity affects on the intrinsic 
-brightness of these stars and variable extinction along the line of sight.
-Moreover, systematic differences between instruments make it difficult to tie observations across the sky
-to a common scale. While RR Lyrae are among the best distance indicators available to us today, the absolute uncertainty in
-individual estimates are typically $\gtrsim$10\%.
+There is a long tradition for using RR Lyrae stars in the Galaxy to
+study structure \citep[going back to first estimates of the distance
+  to the Galactic center][]{shapley18}, substructure
+\citep[e.g.][]{sesar10}, and distances to satellite galaxies
+\citep[e.g.][]{clementini03}.  However, studies of RR Lyrae at optical
+wavelengths are limited by both metallicity affects on the intrinsic
+brightness of these stars and variable extinction along the line of
+sight.  Moreover, systematic differences between instruments make it
+difficult to tie observations across the sky to a common scale. While
+RR Lyrae are among the best distance indicators available to us today,
+the absolute uncertainty in individual estimates are typically
+$\gtrsim$10\%.
 
-At longer wavelengths, RR Lyrae promise tighter constraints on distances.
-The near-infrared period-luminosity (PL) relation for RR
-Lyrae stars was first mapped by \citet{longmore86} 
-using $K$-band observations of stars in Galactic
-globular clusters and later shown to be due to
-the
-changing bolometric correction for these stars as a function of the
-observed wavelength \citep{catelan04}.
- \citet{madore12} have recently shown that the intrinsic scatter of the PL relation
- decreases  with increasing
-wavelength. Both reddening effects and metallicity dependencies also
-decrease significantly for observations further into the IR.
+At longer wavelengths, RR Lyrae promise tighter constraints on
+distances.  The near-infrared period-luminosity (PL) relation for RR
+Lyrae stars was first mapped by \citet{longmore86} using $K$-band
+observations of stars in Galactic globular clusters and later shown to
+be due to the changing bolometric correction for these stars as a
+function of the observed wavelength \citep{catelan04}.
+\citet{madore12} have recently shown that the intrinsic scatter of the
+PL relation decreases with increasing wavelength. Both reddening
+effects and metallicity dependencies also decrease significantly for
+observations further into the IR.
 
 These studies suggest that observations of RR Lyrae variables by
-NASA's Spitzer mission could provide stellar distance
-indicators of unprecedented accuracy and precision.
-Using five stars with trigonometric parallaxes
-  measured by Hubble \citep{benedict11}, \citet{madore12} measured  
-  the dispersion in the
-mid-IR PL relation at Spitzer wavelengths to be $\sim$0.03~mag. This implies that it is possible to determine
-individual distances that are good to better than $2\%$ for RR
-Lyrae stars out to $\sim$60~kpc (the limit of Spitzer's ability to
-detect and measure RR Lyrae variables for modest integration
-times). Distance measurements of Blue Horizontal Branch and Blue Straggler stars at these distances in the halo typically achieve $\sim$25\% uncertainties \citep[e.g.,][]{deason12b}.
+NASA's Spitzer mission could provide stellar distance indicators of
+unprecedented accuracy and precision.  Using five stars with
+trigonometric parallaxes measured by Hubble \citep{benedict11},
+\citet{madore12} measured the dispersion in the mid-IR PL relation at
+Spitzer wavelengths to be $\sim$0.03~mag. This implies that it is
+possible to determine individual distances that are good to better
+than $2\%$ for RR Lyrae stars out to $\sim$60~kpc (the limit of
+Spitzer's ability to detect and measure RR Lyrae variables for modest
+integration times). Distance measurements of Blue Horizontal Branch
+and Blue Straggler stars at these distances in the halo typically
+achieve $\sim$25\% uncertainties \citep[e.g.,][]{deason12b}.
 
@@
 \label{sec:gaia}
-The Gaia satellite \citep{gaia01}, planned for launch in late
-2013, is an astrometric mission which aims to measure the positions of billions of stars with
-10-100~$\mu$as accuracies. With over 50 observations of the entire sky
-during the five-year lifetime of the mission, Gaia will
+The Gaia satellite \citep{gaia01}, planned for launch in late 2013, is
+an astrometric mission which aims to measure the positions of billions
+of stars with 10-100~$\mu$as accuracies. With over 50 observations of
+the entire sky during the five-year lifetime of the mission, Gaia will
 revolutionize the field of Galactic Astronomy by providing sufficient
-proper motion accuracy to create full six-dimensional phase-space
-maps of the Galaxy with $<$10\% distance errors for heliocentric distances up to $\sim$6~kpc for RR Lyrae stars.
+proper motion accuracy to create full six-dimensional phase-space maps
+of the Galaxy with $<$10\% distance errors for heliocentric distances
+up to $\sim$6~kpc for RR Lyrae stars.
 
-Figure~\ref{fig:gaia_errors} shows the Gaia end-of-mission distance and
-tangential velocity error estimates for RR Lyrae.
-The velocity uncertainty depends on the magnitude of the source, which depends on the metallicity: each line in Figure~\ref{fig:gaia_errors} is computed by Monte Carlo sampling from the empirical metallicity distribution of the Galactic halo from \cite{ivezic08} (centered around [Fe/H] $\sim$ -1.5).
-Within 2~kpc, Gaia
-will measure distances to these stars with better than 2\%
-accuracy -- RR Lyrae in this volume will be used to test and calibrate
-the Spitzer PL relation described above. Beyond the 2~kpc threshold,
-the mid-IR PL relation for RR Lyrae will provide more accurate
-distance measurements.
+Figure~\ref{fig:gaia_errors} shows the Gaia end-of-mission distance
+and tangential velocity error estimates for RR Lyrae. The velocity
+uncertainty depends on the magnitude of the source, which depends on
+the metallicity: each line in Figure~\ref{fig:gaia_errors} is computed
+by Monte Carlo sampling from the empirical metallicity distribution of
+the Galactic halo from \cite{ivezic08} (centered around [Fe/H] $\sim$
+-1.5). Within 2~kpc, Gaia will measure distances to these stars with
+better than 2\% accuracy -- RR Lyrae in this volume will be used to
+test and calibrate the Spitzer PL relation described above. Beyond the
+2~kpc threshold, the mid-IR PL relation for RR Lyrae will provide more
+accurate distance measurements.
 
@@
 tangential velocity errors of order $\sim$1~km/s at 30~kpc, and of
-order $\sim$10~km/s out to several tens of kiloparsecs.\footnote{\url{http://www.rssd.esa.int/index.php?page=Science_Performance&project=GAIA}} The
-combination of Spitzer and Gaia data will extend the ``horizon'' of
-where full six-dimensional phase-space maps of the Galaxy are possible
-from $<$10~kpc to 80~kpc. This enormous increase in volume will greatly refine
-data on existing debris systems such as the Sagittarius stream.
+order $\sim$10~km/s out to several tens of
+kiloparsecs.\footnote{\url{http://www.rssd.esa.int/index.php?page=Science_Performance&project=GAIA}}
+The combination of Spitzer and Gaia data will extend the ``horizon''
+of where full six-dimensional phase-space maps of the Galaxy are
+possible from $<$10~kpc to 80~kpc. This enormous increase in volume
+will greatly refine data on existing debris systems such as the
+Sagittarius stream.
 
@@
 
-The first maps of the Sgr core \citep{ibata94, ibata95} revealed
-its elongated morphology and suggested tidal deformation, as would be
+The first maps of the Sgr core \citep{ibata94, ibata95} revealed its
+elongated morphology and suggested tidal deformation, as would be
 anticipated from its proximity to the Galaxy. Subsequent N-body models
@@
 \citep{belokurov06}, and RR Lyrae in the Catalina Sky Survey
-\citep{drake13}.
-More recently, apogalactic debris has been traced to large distances using a combination of blue horizontal branch, main-sequence turn-off, and red giant stars from the SDSS \citep{belokurov13}.
+\citep{drake13}. More recently, apogalactic debris has been traced to
+large distances using a combination of blue horizontal branch,
+main-sequence turn-off, and red giant stars from the SDSS
+\citep{belokurov13}.
 
@@
 Most recently, \citet[][hereafter LM10]{law10} have combined all
-current data on the Sgr debris to assess the triaxiality and orientation
-of the outer Galaxy, the first time that such a reconstruction of
-the 3-dimensional mass distribution of a dark matter halo has been
-feasible. Figure \ref{fig:lm10} shows particle positions (in Galactocentric coordinates) from the final time-step of an N-body simulation of dwarf satellite disruption along the expected Sgr
-orbit in the best-fitting Milky Way halo model from LM10. The
-simulation was run in a three-component potential, with a triaxial,
-logarithmic halo model of the form
+current data on the Sgr debris to assess the triaxiality and
+orientation of the outer Galaxy, the first time that such a
+reconstruction of the 3-dimensional mass distribution of a dark matter
+halo has been feasible. Figure \ref{fig:lm10} shows particle positions
+(in Galactocentric coordinates) from the final time-step of an N-body
+simulation of dwarf satellite disruption along the expected Sgr orbit
+in the best-fitting Milky Way halo model from LM10. The simulation was
+run in a three-component potential, with a triaxial, logarithmic halo
+model of the form
 \begin{equation}
@@
 \end{equation}
-where $C_1$, $C_2$, and $C_3$ are combinations of the $x$ and $y$ axis ratios ($q_1$, $q_2$) and orientation of the halo with respect to the baryonic disk ($\phi$):
+where $C_1$, $C_2$, and $C_3$ are combinations of the $x$ and $y$ axis
+ratios ($q_1$, $q_2$) and orientation of the halo with respect to the
+baryonic disk ($\phi$):
 \begin{align}
@@
 observed debris enables a detailed reconstruction of Sgr's history:
-its mass and stellar populations, orbit, rate of destruction, and
-even original distribution in populations. The phase-space distribution of the debris also offers a unique probe of
-the depth, shape, extent, and orientation of the Milky Way's dark
-matter halo that cannot be matched using other techniques on our own
-or other galaxies \citep[e.g.,][]{ibata01}. The LM10 constraints on
-Sgr's mass and orbit have invigorated discussions of Sgr in a more
-cosmological context, the mass and extent of the original dark matter
-halo that hosted it, as well as its effect on the Milky Way itself
+its mass and stellar populations, orbit, rate of destruction, and even
+original distribution in populations. The phase-space distribution of
+the debris also offers a unique probe of the depth, shape, extent, and
+orientation of the Milky Way's dark matter halo that cannot be matched
+using other techniques on our own or other galaxies
+\citep[e.g.,][]{ibata01}. The LM10 constraints on Sgr's mass and orbit
+have invigorated discussions of Sgr in a more cosmological context,
+the mass and extent of the original dark matter halo that hosted it,
+as well as its effect on the Milky Way itself
 \citep{bailin03,purcell11,micheldansac11,gomez12}.
 
-Combined Spitzer and Gaia measurements of distances and proper
-motions of RR Lyraes in the Sgr debris will open up new avenues of
-interpretation. A 2\% distance error at $\sim$30~kpc ($\epsilon_{d}=0.6~{\rm kpc}$) is smaller than
-the predicted thickness of the stream, ranging from
-$\sim$2-$10$~kpc. Combined with Gaia tangential velocity measurement
-accuracies ($\epsilon_{v}\sim2$~km/s at 30~kpc) less than the intrinsic dispersion
-for much of the stream \citep[$\sigma_v\sim10$~km/s;][]{majewski04}, and ground-based radial
-velocities, the measured phase-space
-position of each individual RR Lyrae becomes a
-powerful probe of the potential. The next section outlines a new
-method to take advantage of this information.
+Combined Spitzer and Gaia measurements of distances and proper motions
+of RR Lyraes in the Sgr debris will open up new avenues of
+interpretation. A 2\% distance error at $\sim$30~kpc
+($\epsilon_{d}=0.6~{\rm kpc}$) is smaller than the predicted thickness
+of the stream, ranging from $\sim$2-$10$~kpc. Combined with Gaia
+tangential velocity measurement accuracies ($\epsilon_{v}\sim2$~km/s
+at 30~kpc) less than the intrinsic dispersion for much of the stream
+\citep[$\sigma_v\sim10$~km/s;][]{majewski04}, and ground-based radial
+velocities, the measured phase-space position of each individual RR
+Lyrae becomes a powerful probe of the potential. The next section
+outlines a new method to take advantage of this information.
 
@@
 With access to full phase-space information for stars in a tidal
-stream, it is possible to use the individual stars as
-potential measures by exploting the fact that the stars must have come
-from the same progenitor. If the present-day phase-space position of the
-progenitor system is also known, this places an intuitive constraint on
-the potential of the Milky Way: if the orbits of the stars and the
+stream, it is possible to use the individual stars as potential
+measures by exploting the fact that the stars must have come from the
+same progenitor. If the present-day phase-space position of the
+progenitor system is also known, this places an intuitive constraint
+on the potential of the Milky Way: if the orbits of the stars and the
 orbit of the progenitor are integrated \emph{backwards} in time
-(imagine watching satellite destruction in ``rewind"), the
-stars should recombine with the progenitor. If the potential is
-incorrect, the orbits of the stars will diverge from that of the
-progenitor and thus will not be recaptured by the satellite system (Figure~\ref{fig:ps_distance}).
+(imagine watching satellite destruction in ``rewind"), the stars
+should recombine with the progenitor. If the potential is incorrect,
+the orbits of the stars will diverge from that of the progenitor and
+thus will not be recaptured by the satellite system
+(Figure~\ref{fig:ps_distance}).
 
-This approach was originally proposed by \citep{johnston99a} and was designed and tested for the proposed characteristics of the Space Interferometry Mission \citep{unwin08}, which was expected to be able to measure proper motions with $\mu$as/year accuracy for stars as faint as $V=20$.
-Distances to stream members would only be poorly known and were estimated in the method by using anticipated properties of tidal debris.
-Below we present an updated version of the algorithm: the promise of 2\% distances to RR Lyrae stars (see Section \ref{sec:spitzer}) enables a direct measurement (rather than approximate estimate) of the position of a star within its debris structure.
-The test statistic that quantifies how well stars recombine with the satellite has been rigorously redefined.
+This approach was originally proposed by \citep{johnston99a} and was
+designed and tested for the proposed characteristics of the Space
+Interferometry Mission \citep{unwin08}, which was expected to be able
+to measure proper motions with $\mu$as/year accuracy for stars as
+faint as $V=20$. Distances to stream members would only be poorly
+known and were estimated in the method by using anticipated properties
+of tidal debris. Below we present an updated version of the algorithm:
+the promise of 2\% distances to RR Lyrae stars (see Section
+\ref{sec:spitzer}) enables a direct measurement (rather than
+approximate estimate) of the position of a star within its debris
+structure. The test statistic that quantifies how well stars recombine
+with the satellite has been rigorously redefined.
 
@@
 \end{equation}
-Qualitatively, when the distance in this normalized six-dimensional space,
-$D_{{\rm ps},i}=\sqrt{|\bs{q}_{i}|^2+|\bs{p}_{i}|^2} < \sqrt{2}$, the star is
-likely recaptured by the satellite.
-\citet{johnston99a} (hereafter JZSH) imposed a similar condition as a hard boundary and used the number of recaptured particles in a given backwards-integration as a cost function for optimization.
-What follows is a description of an updated procedure with a more natural, statistically-motivated choice for an objective function.
+Qualitatively, when the distance in this normalized six-dimensional
+space, $D_{{\rm ps},i}=\sqrt{|\bs{q}_{i}|^2+|\bs{p}_{i}|^2} <
+\sqrt{2}$, the star is likely recaptured by the
+satellite. \citet{johnston99a} (hereafter JZSH) imposed a similar
+condition as a hard boundary and used the number of recaptured
+particles in a given backwards-integration as a cost function for
+optimization. What follows is a description of an updated procedure
+with a more natural, statistically-motivated choice for an objective
+function.
 
@@
 \end{align}
-Thus, the matrix $\bs{A}_{ik}$ contains these minimum phase-space distance vectors for each star,
-where $k\in[1,6]$. Intuitively, the variance of the distribution of minimum phase-space vectors
-will be larger for orbits integrated in an incorrect potential
-relative to the distribution computed from the `true' orbital history
-of the stars: in an incorrect potential, small differences in the
-orbits of the stars relative to the orbit of the progenitor cause them
-to spread out in this relative, normalized phase space. Thus, the
-\emph{generalized variance} of the distribution -- computed for a given
-set of potential parameters, $\bs{\theta}$ -- is a natural choice for
-the scalar objective function used in constraining the potential of
-the Milky Way 
+Thus, the matrix $\bs{A}_{ik}$ contains these minimum phase-space
+distance vectors for each star, where $k\in[1,6]$. Intuitively, the
+variance of the distribution of minimum phase-space vectors will be
+larger for orbits integrated in an incorrect potential relative to the
+distribution computed from the `true' orbital history of the stars: in
+an incorrect potential, small differences in the orbits of the stars
+relative to the orbit of the progenitor cause them to spread out in
+this relative, normalized phase space. Thus, the \emph{generalized
+  variance} of the distribution -- computed for a given set of
+potential parameters, $\bs{\theta}$ -- is a natural choice for the
+scalar objective function used in constraining the potential of the
+Milky Way
 \begin{align}
@@
 
-[Still working on this bit...need to figure out best way to express what I'm trying to say.] Though this is not explicitly a \emph{likelihood}, optimizing over this statistic is a very simple approximation to performing a maximum-likelihood optimization.
-Assuming the samples in $\bs{A}_{ik}$ are approximately normal-distributed, the variance is closely related to the Shannon entropy \citep{shannon1948}, $H$,
+[Still working on this bit...need to figure out best way to express
+  what I'm trying to say.] Though this is not explicitly a
+\emph{likelihood}, optimizing over this statistic is a very simple
+approximation to performing a maximum-likelihood
+optimization. Assuming the samples in $\bs{A}_{ik}$ are approximately
+normal-distributed, the variance is closely related to the Shannon
+entropy \citep{shannon1948}, $H$,
 \begin{equation}
@@
 \end{equation}
-Minimizing the entropy is then equivalent to minimizing the loss of \emph{information} for a given choice of parameters, $\bs{\theta}$, given the ``data,'' $\bs{A}_{ik}$, and is thus related to maximum-likelihood estimation \citep[see][for more detail]{mackay2002}.
+Minimizing the entropy is then equivalent to minimizing the loss of
+\emph{information} for a given choice of parameters, $\bs{\theta}$,
+given the ``data,'' $\bs{A}_{ik}$, and is thus related to
+maximum-likelihood estimation \citep[see][for more
+  detail]{mackay2002}.
 
@@
 from the present-day snapshot of the simulation
-data.\footnote{\url{www.astro.virginia.edu/~srm4n/Sgr/data.html}} We then
-`observe' a sample of 100 stars from the first leading and trailing
-wraps of the stream: the radial velocity errors are drawn from a
-Gaussian $\epsilon_{rv} \sim \mathcal{N}(\mu=0,\sigma=10~{\rm km/s})$,
-the tangential velocity errors depend on the distance to the star and
-are computed from the expected Gaia error curve (shown in
-Figure~\ref{fig:gaia_errors}), and the distance errors on a star of distance
-$D$ are $\epsilon_{dist} \sim \mathcal{N}(0,0.02\times
-D)$. Figure~\ref{fig:variance_proj} shows two-dimensional projections of the six-dimensional minimum
-phase-space distance distribution for the observed particles in the
-correct and an incorrect potential: it is clear that the
-variance of the distribution increases in the wrong potential. The 
-generalized variance defines a convex hypersurface over which we
-optimize four of the six logarithmic potential parameters:
-$v_{circ}$, $\phi$, $q_1$, and $q_z$.
-We fix $q_2$ and $R_c$, which are degenerate with combinations of the other parameters.
-Figure~\ref{fig:objective} shows one-dimensional slices of the objective function
-produced by varying each of the potential parameters by
-$\pm10\%$ around the true values and holding all others fixed.
+data.\footnote{\url{www.astro.virginia.edu/~srm4n/Sgr/data.html}} We
+then `observe' a sample of 100 stars from the first leading and
+trailing wraps of the stream: the radial velocity errors are drawn
+from a Gaussian $\epsilon_{rv} \sim \mathcal{N}(\mu=0,\sigma=10~{\rm
+  km/s})$, the tangential velocity errors depend on the distance to
+the star and are computed from the expected Gaia error curve (shown in
+Figure~\ref{fig:gaia_errors}), and the distance errors on a star of
+distance $D$ are $\epsilon_{dist} \sim \mathcal{N}(0,0.02\times
+D)$. Figure~\ref{fig:variance_proj} shows two-dimensional projections
+of the six-dimensional minimum phase-space distance distribution for
+the observed particles in the correct and an incorrect potential: it
+is clear that the variance of the distribution increases in the wrong
+potential. The generalized variance defines a convex hypersurface over
+which we optimize four of the six logarithmic potential parameters:
+$v_{circ}$, $\phi$, $q_1$, and $q_z$. We fix $q_2$ and $R_c$, which
+are degenerate with combinations of the other
+parameters. Figure~\ref{fig:objective} shows one-dimensional slices of
+the objective function produced by varying each of the potential
+parameters by $\pm10\%$ around the true values and holding all others
+fixed.
 
@@
 
-In anticipation of extending the above method to include a true likelihood function, we use a parallelized Markov Chain Monte Carlo (MCMC) algorithm \citep{foremanmackey2013} to sample from our objective function.\footnote{Though MCMC is typically not an efficient optimization tool, in this case objective function is both noisy and expensive to compute.
-The stochasticity and easy parallelization of the algorithm mean it outperforms other optimizers on this problem.} We use the median value of the (burnt-in) sample distribution as a point estimate for the potential parameters.
-To assess the uncertainty in the derived halo parameters, we bootstrap resample our 100 stars 100 times and estimate the potential parameters with each resampling.
-Figure~\ref{fig:bootstrap} shows the recovered parameters for each bootstrapped sample and demonstrates the power of the combined Spitzer and Gaia data: a moderately sized sample of RR Lyrae alone places quite strong constraints on the shape and mass of the Galaxy's dark matter halo.
+In anticipation of extending the above method to include a true
+likelihood function, we use a parallelized Markov Chain Monte Carlo
+(MCMC) algorithm \citep{foremanmackey2013} to sample from our
+objective function.\footnote{Though MCMC is typically not an efficient
+  optimization tool, in this case objective function is both noisy and
+  expensive to compute. The stochasticity and easy parallelization of
+  the algorithm mean it outperforms other optimizers on this problem.}
+We use the median value of the (burnt-in) sample distribution as a
+point estimate for the potential parameters. To assess the uncertainty
+in the derived halo parameters, we bootstrap resample our 100 stars
+100 times and estimate the potential parameters with each
+resampling. Figure~\ref{fig:bootstrap} shows the recovered parameters
+for each bootstrapped sample and demonstrates the power of the
+combined Spitzer and Gaia data: a moderately sized sample of RR Lyrae
+alone places quite strong constraints on the shape and mass of the
+Galaxy's dark matter halo.
 
@@
 
-Many orbit fitting methods for measuring the Galactic potential assume that the stars in a given stream are on the same
-orbit. \cite{sanders13a} show that this assumption can lead to nearly 100\% errors in the 
-estimated potential parameters, even without realistic observational errors on
-the kinematic data. The \emph{Rewinder} is
-free of this assumption and instead \emph{relies} on the fact that
-each star is on a different orbit, and is thus an independent potentiometer. Qualitatively, the best potential is the one that gets the
-most of these orbits to intersect the progenitor over the history of interaction with the system.
+Many orbit fitting methods for measuring the Galactic potential assume
+that the stars in a given stream are on the same
+orbit. \cite{sanders13a} show that this assumption can lead to nearly
+100\% errors in the estimated potential parameters, even without
+realistic observational errors on the kinematic data. The
+\emph{Rewinder} is free of this assumption and instead \emph{relies}
+on the fact that each star is on a different orbit, and is thus an
+independent potentiometer. Qualitatively, the best potential is the
+one that gets the most of these orbits to intersect the progenitor
+over the history of interaction with the system.
 
-The strengths of this method stem from its simplicity: it requires only a rough estimate of the satellite mass $m_{\rm sat}$ combined with backwards integration of orbits.
-There are no assumptions made about, or N-body modeling of, the internal distribution of satellite stars.
-Thus, the \emph{Rewinder} is applicable to any debris that is known to come from a single object and not restricted to the coldest tidal streams.
-In principle, the
-method could also be applied to the vast debris {\it clouds} that have been discovered over large areas of the sky \citep[e.g. the Triangulum Andromeda
-and Hercules-Aquila clouds]{rochapinto04,belokurov06}, or even stars that have only associations in orbital properties and do not form a coherent spatial structure
-\citep[such as the angular momentum groupings in local giants found by][]{helmi99}.
-The method trivially extends to combining constraints from multiple debris systems at once by simply integrating all debris from several satellites simultaneously, with $D_{\rm ps}$ defined appropriately for each star.
+The strengths of this method stem from its simplicity: it requires
+only a rough estimate of the satellite mass $m_{\rm sat}$ combined
+with backwards integration of orbits. There are no assumptions made
+about, or N-body modeling of, the internal distribution of satellite
+stars. Thus, the \emph{Rewinder} is applicable to any debris that is
+known to come from a single object and not restricted to the coldest
+tidal streams. In principle, the method could also be applied to the
+vast debris {\it clouds} that have been discovered over large areas of
+the sky \citep[e.g. the Triangulum Andromeda and Hercules-Aquila
+  clouds]{rochapinto04,belokurov06}, or even stars that have only
+associations in orbital properties and do not form a coherent spatial
+structure \citep[such as the angular momentum groupings in local
+  giants found by][]{helmi99}. The method trivially extends to
+combining constraints from multiple debris systems at once by simply
+integrating all debris from several satellites simultaneously, with
+$D_{\rm ps}$ defined appropriately for each star.
 
-It is also important to characterize the the limitations of this method.
-Firstly, the measurement errors for RR Lyraes associated with the very coldest streams \citep[e.g., the globular clusters Pal5 and GD1;][]{odenkirchen02,koposov10} will likely be too large to resolve the minute differences in orbital properties between the debris and satellite to be detected.
-Second, the current version of the algorithm relies on accurate knowledge of the current position and velocity of the parent satellite, which may not be available in some cases \citep[for example, the Orphan Stream;][]{belokurov07}.
-Lastly, the present prescription treats the progenitor as a point-mass during orbit integration.
-Second-order effects in orbit evolution (e.g., dynamical friction) are currently ignored, leading to minor biases in the estimated potential parameters (for example, XXX in Figure~\ref{fig:bootstrap}).
+It is also important to characterize the the limitations of this
+method. Firstly, the measurement errors for RR Lyraes associated with
+the very coldest streams \citep[e.g., the globular clusters Pal5 and
+  GD1;][]{odenkirchen02,koposov10} will likely be too large to resolve
+the minute differences in orbital properties between the debris and
+satellite to be detected. Second, the current version of the algorithm
+relies on accurate knowledge of the current position and velocity of
+the parent satellite, which may not be available in some cases
+\citep[for example, the Orphan Stream;][]{belokurov07}. Lastly, the
+present prescription treats the progenitor as a point-mass during
+orbit integration. Second-order effects in orbit evolution (e.g.,
+dynamical friction) are currently ignored, leading to minor biases in
+the estimated potential parameters (for example, XXX in
+Figure~\ref{fig:bootstrap}).
 
@@
 
-The objective function presented above is an approximation to a more general likelihood function, described in depth in forthcoming work (Price-Whelan \& Johnston in prep.).
-The results shown in section \ref{sec:results} are meant to illustrate the strength and concept behind the method, but [...] [I want to say one more sentence about the upcoming work, and possibly how it will solve some of the limitations presented above.]
+The objective function presented above is an approximation to a more
+general likelihood function, described in depth in forthcoming work
+(Price-Whelan \& Johnston in prep.). The results shown in section
+\ref{sec:results} are meant to illustrate the strength and concept
+behind the method, but [...] [I want to say one more sentence about
+  the upcoming work, and possibly how it will solve some of the
+  limitations presented above.]
 
@@
 
-This paper presents an algorithm for measuring the Galactic potential that anticipates combined data from the Spitzer and Gaia satellite missions that promise precise, full phase-space measurements of RR Lyrae stars in the halo of our Galaxy.
-When applied to a sample of 100 stars -- with realistic observational errors -- sampled from the \cite{law10} N-body simulation of the destruction of the Sgr dwarf satellite, the method recovers the shape and orientation of the dark matter potential XXX.
-This success provides strong motivation for 1) further theoretical work to [...], 2) an investigation of the power of using multiple, complimentary debris structures (e.g., with different morphologies and orientations) in building a comprehensive picture of the Galactic mass distribution, and 3) a Spitzer survey of RR Lyrae stars in debris structures around the Milky Way to get precise distances to combine with near-future Gaia velocity data.
+This paper presents an algorithm for measuring the Galactic potential
+that anticipates combined data from the Spitzer and Gaia satellite
+missions that promise precise, full phase-space measurements of RR
+Lyrae stars in the halo of our Galaxy. When applied to a sample of 100
+stars -- with realistic observational errors -- sampled from the
+\cite{law10} N-body simulation of the destruction of the Sgr dwarf
+satellite, the method recovers the shape and orientation of the dark
+matter potential XXX. This success provides strong motivation for 1)
+further theoretical work to [...], 2) an investigation of the power of
+using multiple, complimentary debris structures (e.g., with different
+morphologies and orientations) in building a comprehensive picture of
+the Galactic mass distribution, and 3) a Spitzer survey of RR Lyrae
+stars in debris structures around the Milky Way to get precise
+distances to combine with near-future Gaia velocity data.
 

### 2013-07-31 text/paper1/paper.tex
getting prose-y

@@
-\documentclass{emulateapj}
-%\documentclass[preprint]{aastex}
+%\documentclass{emulateapj}
+\documentclass[preprint]{aastex}
 
@@
 
-[Ok, there is a question as to whether we even want to say something like the below...?]
+[KVJ: Ok, there is a question as to whether we even want to say something like the below...?]
 
@@
 Lyrae stars in the halo of our Galaxy. When applied to a sample of 100
-stars -- with realistic observational errors -- sampled from the
+stars (with realistic observational errors) sampled from the
 \cite{law10} N-body simulation of the destruction of the Sgr dwarf
@@
 matter potential XXX. This success provides strong motivation for 1)
-further theoretical work to [...], 2) an investigation of the power of
+further theoretical work to develop a robust generative model (and therefore likelihood) that utilizes the concepts demonstrated by the \emph{Rewinder}, 2) an investigation of the power of
 using multiple, complimentary debris structures (e.g., with different
@@
 distances to combine with near-future Gaia velocity data.
+We conclude by noting that this method is complimentary to the orbit fitting algorithm presented in \citet{sanders13b} by favoring warmer, more massive debris structures like Sgr.
+By developing and comparing independent methods for measuring the potential of the Milky Way, we will begin to develop a deeper understanding of the vast structures of dark matter that permeate the universe.
+[Yea, the last sentence I wrote in the late night and am probably delirious...]
 

### 2013-07-31 text/paper1/paper.tex
grammar in intro - first edit

@@
 \label{intro.sec}
- The existence of vast halos of unseen {\it dark} matter surrounding each galaxy has long been proposed \citep[e.g.,][]{zwicky37,rubin70} to explain the surprisingly large
-motions of the {\it baryonic} matter (i.e. stars, gas and dust) that we can see.
-Dark-matter-only simulations of structure formation lead us to expect that these  halos should have a density distribution that can simply be described with a universal
-profile \citep{navarro96} with a variety of triaxial shapes \citep{jing02}.
-The inclusion of baryons in the simulations tends to soften the triaxiality of the dark matter in the inner regions of the halo \citep[as the disk forms, see e.g.,][]{bailin05}, and
+ The existence of vast halos of unseen {\it dark} matter surrounding each galaxy has long been proposed to explain the surprisingly large
+motions of the {\it baryonic} matter (i.e. stars, gas and dust) that we can see \citep[e.g.,][]{zwicky37,rubin70}.
+Dark-matter-only simulations of structure formation lead us to expect that these dark matter halos should have density distributions that are described by a universal radial profile \citep{navarro96} with a variety of triaxial shapes \citep{jing02}.
+The inclusion of baryons in the simulations tends to soften the triaxiality of the dark matter in the inner regions of the halo \citep[e.g., as the disk forms,][]{bailin05} and
 can potentially alter the radial profile through a combination of adiabatic contraction and energetic feedback \citep[see recent discussion by][]{pontzen12}.
-Hence, measurements of the shape, orientation, radial profile and extent of dark matter halos can shed light both on the formation of these vast structures and the role that
-the messy baryonic processes of gas inflow, star formation and feedback  have played in shaping them.
+Hence, measurements of the shape, orientation, radial profile, and extent of dark matter halos provides information about the formation of these vast structures, as well as the messy baryonic processes of gas inflow, star formation, and feedback that continue to shape them.
 
-The Milky Way is the best candidate  for dissecting a dark matter halo in great detail.
-Our proximity means that we can use individual stars as tracer populations and hence build much larger samples that probe deeper into the halo than the globular cluster
-and planetary nebula studies of external galaxies \citep[e.g.][]{mendez01,cote03}.
-Moreover, our internal perspective allows us to build a three-dimensional picture of the mass distribution rather than measuring its two-dimensional projection.
+The Milky Way is the best candidate for such a detailed study of a dark matter halo. 
+Our presence in the Galaxy means that we can resolve and use individual stars as tracers and hence build large samples that probe deeper into the halo than globular cluster
+or planetary nebula studies of external galaxies \citep[e.g.][]{mendez01,cote03}.
+Our internal perspective allows us to measure the three-dimensional mass distribution around the Galaxy rather than a two-dimensional projection, such as in lensing studies \citep{bolton06}.
 %Hence we can hope to reconstruct not only the density, but also the shape and orientation of the dark matter halo as function of radius well beyond the disk.
-For example, to date thousands of blue-horizontal-branch stars selected from the Sloan Digital Sky Survey (SDSS) have been used to probe mass out to tens of
+To date, thousands of blue horizontal branch stars selected from the Sloan Digital Sky Survey (SDSS) have been used to probe the Milky Way mass out to tens of
 kpc \citep[SDSS, see][]{deason12a,kafle12}, and estimates with combined tracers extend to 150kpc \citep{deason12b}.
-\citet{loebman12} demonstrates how even larger stellar samples from current (i.e. SDSS) and future  (Gaia and LSST) stellar surveys will also be sensitive to any
+\citet{loebman12} demonstrates that even larger stellar samples from current (e.g., SDSS) and future  (e.g., Gaia and LSST) stellar surveys will also be sensitive to any
 asphericity of the halo.
 
-There is a caveat to this work using tracer populations: these studies all rely on the assumptions that the tracers  represent a random sampling of orbits drawn from a
-smooth distribution function and are well-mixed in orbital phase.
-However, the same surveys that provide large samples of tracers have also revealed the existence of large-scale spatial inhomogeneities in the form of giant stellar
-streams and clouds \citep{newberg02,majewski03,belokurov06}, clearly demonstrating that  a significant fraction of the stellar halo is neither randomly sampling a smooth
-orbit distribution nor is fully phase-mixed.
+The caveat to working with tracer populations is the assumption that the tracer stars represent a random sampling of orbits drawn from a smooth distribution function and are well-mixed in orbital phase.
+However, the same surveys that provide large samples of tracers have also revealed the existence of large-scale spatial inhomogeneities in the form of giant stellar streams and clouds \citep{newberg02,majewski03,belokurov06}, clearly demonstrating that a significant fraction of the stellar halo is neither randomly sampled from a smooth distribution function nor is fully phase-mixed.
 %The degree of substructure increases with Galactocentric radius \citep{bell08} in agreement with models where the stellar halo is formed purely from accretion events \citep{bullock05,cooper10,rashkov12}.
@@
 
-A complimentary method to measure the mass distribution is to instead take full advantage of the {\it non}-random nature of the stellar distribution in our stellar halo.
 % --- analogous to our exploitation of matter in disks moving on near-circular orbits.
-While we do not know the exact orbits of  stellar debris from satellite destruction \citep[although][proposes a method of orbit reconstruction from limited dimensions of
-observations]{eyre09b}, we do know something more about stars in a debris structure than in a random sample: we know that these stars were once all part of the same
-object.
-This approach is very promising --- because tidal streams are dynamically cold systems, it requires orders of magnitude fewer tracers than a random sample to get
-constraints of comparable accuracy.
+A complimentary method to measure the mass distribution is to instead take full advantage of the {\it non}-random nature of the stellar distribution in the Galactic halo.
+While we do not know the exact orbits of  stellar debris from satellite destruction -- though \citet{eyre09b} propose a method of full orbit reconstruction from incomplete kinematic data -- we do know something more about stars in a debris structure: the stars were once all part of the same object.
+Methods that utilize this knowledge on dynamically cold systems (e.g, tidal streams) require orders of magnitude fewer tracers than a randomly sampled population, but achieve comparable accuracy on derived constraints.
 In the simplest case, we might {\it assume} that debris stars are actually still on the same orbit as their progenitor system.

### 2013-07-31 text/paper1/paper.tex
apw intro revisions take 2

@@
 A complimentary method to measure the mass distribution is to instead take full advantage of the {\it non}-random nature of the stellar distribution in the Galactic halo.
-While we do not know the exact orbits of  stellar debris from satellite destruction -- though \citet{eyre09b} propose a method of full orbit reconstruction from incomplete kinematic data -- we do know something more about stars in a debris structure: the stars were once all part of the same object.
+While we do not know the exact orbits of  stellar debris from satellite destruction --- though \citet{eyre09b} propose a method of full orbit reconstruction from incomplete kinematic data --- we do know something more about stars in a debris structure: the stars were once all part of the same object.
 Methods that utilize this knowledge on dynamically cold systems (e.g, tidal streams) require orders of magnitude fewer tracers than a randomly sampled population, but achieve comparable accuracy on derived constraints.
-In the simplest case, we might {\it assume} that debris stars are actually still on the same orbit as their progenitor system.
+[KVJ: can we cut some of this?
+Anything from here to end of the paragraph?] In the simplest case, we might {\it assume} that debris stars are actually still on the same orbit as their progenitor system.
 Then we can imagine that measuring the full-space velocities ${\bf v}$ at different points ${\bf x}$ in the structure (e.g. along a stream) would actually give us a direct
 measure of differences in potential $\Phi$ (i.e. $\Phi({\bf x}_1)-\Phi({\bf x}_2)=(v_2^2-v_1^2)/2$).
-In reality, we usually know at most four of the six phase-space co-ordinates for debris stars, often with significant error bars and potential measurements have been made by
-fitting orbits to streams \citep[e.g.,][]{helmi04,johnston05,koposov10,law10}.
+[KVJ: incomplete sentence?
+Not sure what the point you were making is?] In reality, we usually know at most four of the six phase-space co-ordinates for debris stars, often with significant error bars and potential measurements have been made by fitting orbits to streams \citep[e.g.,][]{helmi04,johnston05,koposov10,law10}.
 Moreover, our assumption of debris tracing a single orbit is actually incorrect \citep[see][for discussions of the orbit distribution in tidal debris]{johnston98,helmi99}, with
@@
 Other approaches have been proposed that avoid the cost of full N-body simulations yet recognize the limitations of single orbit integration.
-For example, using the results of N-body simulations as a guide, simple corrections around a single orbit can be adopted to make model predictions that more closely
+For example, using the results of N-body simulations as a guide, simple corrections around a single orbit can be adopted to make model predictions that closely
 mimic expected debris behavior \citep[e.g.][]{johnston99b,varghese11}.
@@
 assumption of the form of the Galactic potential.
-\citet{sanders13b} examine the distribution of debris in action-angle co-ordinates, show stars that were once all members of the same object must lie along a single line in angle-frequency space and demonstrate how this constraint can be used as a potential measure.
+\citet{sanders13b} examine the distribution of debris in action-angle co-ordinates and show that stars stripped from the same disrupted object must lie along a single line in angle-frequency space, providing a constraint that can be used as a potential measure.
 %\citet{johnston99a} (hereafter JZSH) showed that backwards integration of the positions of stars and their parent satellite could be used to distinguish Milky Way mass models: only in the correct potential would the stars' and satellite's  path coincide within a Hubble time.The JZSH method was designed and tested for the proposed characteristics of the Space Interferometry Mission \citep{}, which was expected to be able to measure proper motions with $\mu$as/year accuracy for stars as faint as $V=20$.
@@
 
-In this paper we re-examine and update  another approach to using tidal debris as a potential measure \citep[originially proposed by][]{johnston99a}  in the context of current and near-future observational capabilities, and illustrates its power with the example of Sgr.
+In this paper we re-examine and update a complimentary approach to using tidal debris as a potential measure \citep[originially proposed by][]{johnston99a}  in the context of current and near-future observational capabilities.
+We illustrate the power of this method by applying it to a simulation of the Sgr debris system.
 In Section \ref{sec:context} we outline the observational prospects and Sgr properties that motivated this re-examination.
@@
 -1.5). Within 2~kpc, Gaia will measure distances to these stars with
-better than 2\% accuracy -- RR Lyrae in this volume will be used to
+better than 2\% accuracy --- RR Lyrae in this volume will be used to
 test and calibrate the Spitzer PL relation described above. Beyond the
@@
 this relative, normalized phase space. Thus, the \emph{generalized
-  variance} of the distribution -- computed for a given set of
-potential parameters, $\bs{\theta}$ -- is a natural choice for the
+  variance} of the distribution --- computed for a given set of
+potential parameters, $\bs{\theta}$ --- is a natural choice for the
 scalar objective function used in constraining the potential of the

### 2013-07-31 streams/inference/core.py
use logdet

@@ -91,9 +91,9 @@ def generalized_variance(potential, particle_orbits, satellite_orbit):
     cov_matrix = np.cov(np.fabs(min_ps.T))
     
     # cov_matrix -> (6 x 6) covariance matrix for particles
-    w,v = np.linalg.eig(cov_matrix)
+    #w,v = np.linalg.eig(cov_matrix)
     #return np.sum(w)
-    return np.log(np.prod(w))
+    #return np.log(np.prod(w))
     
     sign,logdet = np.linalg.slogdet(cov_matrix)
     return logdet

### 2013-07-31 text/paper1/paper.tex
add bootstrap figure placeholder
kvj suggestions

@@
 proper motion accuracy to create full six-dimensional phase-space maps
-of the Galaxy with $<$10\% distance errors for heliocentric distances
+of the Galaxy with $<$10\% distance errors for heliocentric distances of
 up to $\sim$6~kpc for RR Lyrae stars.
@@
 and tangential velocity error estimates for RR Lyrae. The velocity
-uncertainty depends on the magnitude of the source, which depends on
+uncertainty depends on the magnitude (and color) of the source, which depends on
 the metallicity: each line in Figure~\ref{fig:gaia_errors} is computed
@@
 
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=1.\textwidth]{/Users/adrian/projects/streams/plots/spitzer_proposal/lm10.png}
+\caption{ todo }\label{fig:lm10}
+\end{center}
+\end{figure}
+
 LM10 demonstrate how a comparison of such simulations with the
@@
 ($\epsilon_{d}=0.6~{\rm kpc}$) is smaller than the predicted thickness
-of the stream, ranging from $\sim$2-$10$~kpc. Combined with Gaia
+of the stream, ranging from $\sim$2-$10$~kpc. When combined with Gaia proper motion errors,
 tangential velocity measurement accuracies ($\epsilon_{v}\sim2$~km/s
-at 30~kpc) less than the intrinsic dispersion for much of the stream
-\citep[$\sigma_v\sim10$~km/s;][]{majewski04}, and ground-based radial
-velocities, the measured phase-space position of each individual RR
+at 30~kpc; see Figure~\ref{fig:gaia_errors}) are less than the intrinsic dispersion for much of the stream
+\citep[$\sigma_v\sim10$~km/s;][]{majewski04} and the measured phase-space position of each individual RR
 Lyrae becomes a powerful probe of the potential. The next section
@@
 on the potential of the Milky Way: if the orbits of the stars and the
-orbit of the progenitor are integrated \emph{backwards} in time
+orbit of the progenitor are integrated \emph{backwards} in a a potential that accurately models the Milky Way
 (imagine watching satellite destruction in ``rewind"), the stars
@@
 
-This approach was originally proposed by \citep{johnston99a} and was
+This approach was originally proposed by \citet{johnston99a} and was
 designed and tested for the proposed characteristics of the Space
@@
 approximate estimate) of the position of a star within its debris
-structure. The test statistic that quantifies how well stars recombine
+structure; the test statistic that quantifies how well stars recombine
 with the satellite has been rigorously redefined.
@@
 \sqrt{2}$, the star is likely recaptured by the
-satellite. \citet{johnston99a} (hereafter JZSH) imposed a similar
+satellite. \citet{johnston99a} imposed a similar
 condition as a hard boundary and used the number of recaptured
@@
 potential parameters, $\bs{\theta}$ --- is a natural choice for the
-scalar objective function used in constraining the potential of the
+scalar objective function, $f(\bs{\theta})$, used in constraining the potential of the
 Milky Way
@@
 
-[Still working on this bit...need to figure out best way to express
-  what I'm trying to say.] Though this is not explicitly a
-\emph{likelihood}, optimizing over this statistic is a very simple
-approximation to performing a maximum-likelihood
-optimization. Assuming the samples in $\bs{A}_{ik}$ are approximately
-normal-distributed, the variance is closely related to the Shannon
-entropy \citep{shannon1948}, $H$,
-\begin{equation}
-  H \propto \log \sigma .
-\end{equation}
-Minimizing the entropy is then equivalent to minimizing the loss of
-\emph{information} for a given choice of parameters, $\bs{\theta}$,
-given the ``data,'' $\bs{A}_{ik}$, and is thus related to
-maximum-likelihood estimation \citep[see][for more
-  detail]{mackay2002}.
+%[Still working on this bit...need to figure out best way to express
+%  what I'm trying to say.] Though this is not explicitly a
+%\emph{likelihood}, optimizing over this statistic is a very simple
+%approximation to performing a maximum-likelihood
+%optimization. Assuming the samples in $\bs{A}_{ik}$ are approximately
+%normal-distributed, the variance is closely related to the Shannon
+%entropy \citep{shannon1948}, $H$,
+%\begin{equation}
+%  H \propto \log \sigma .
+%\end{equation}
+%Minimizing the entropy is then equivalent to minimizing the loss of
+%\emph{information} for a given choice of parameters, $\bs{\theta}$,
+%given the ``data,'' $\bs{A}_{ik}$, and is thus related to
+%maximum-likelihood estimation \citep[see][for more
+%  detail]{mackay2002}.
 
@@
 then `observe' a sample of 100 stars from the first leading and
-trailing wraps of the stream: the radial velocity errors are drawn
-from a Gaussian $\epsilon_{rv} \sim \mathcal{N}(\mu=0,\sigma=10~{\rm
-  km/s})$, the tangential velocity errors depend on the distance to
-the star and are computed from the expected Gaia error curve (shown in
-Figure~\ref{fig:gaia_errors}), and the distance errors on a star of
-distance $D$ are $\epsilon_{dist} \sim \mathcal{N}(0,0.02\times
-D)$. Figure~\ref{fig:variance_proj} shows two-dimensional projections
+trailing wraps of the stream; the radial velocity and distance errors are drawn % APW HERE
+from Gaussians ($\varepsilon_{\rm RV} \sim \mathcal{N}(\mu=0,\sigma=10~{\rm
+  km/s})$ and $\varepsilon_{\rm D} \sim \mathcal{N}(0,0.02\times
+D)$) and the proper motion errors are computed from the expected Gaia error curve. Figure~\ref{fig:variance_proj} shows two-dimensional projections
 of the six-dimensional minimum phase-space distance distribution for
@@
 alone places quite strong constraints on the shape and mass of the
-Galaxy's dark matter halo.
+Galaxy's dark matter halo. [APW: explicitly state limits here, 1-sigma bounds]
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=1.\textwidth]{/Users/adrian/projects/streams/plots/spitzer_proposal/bootstrap.pdf}
+\caption{ todo }\label{fig:bootstrap}
+\end{center}
+\end{figure}
 
@@
 
-\section{Discussion and future prospects}
+\section{Discussion, strengths, and limitations}
 \label{sec:discussion}
 
-Many orbit fitting methods for measuring the Galactic potential assume
+The strengths of this method stem from its simplicity: it requires
+only a rough estimate of the satellite mass $m_{\rm sat}$ combined
+with backwards integration of orbits. Many orbit fitting methods for measuring the Galactic potential assume
 that the stars in a given stream are on the same
@@
 on the fact that each star is on a different orbit, and is thus an
-independent potentiometer. Qualitatively, the best potential is the
-one that gets the most of these orbits to intersect the progenitor
-over the history of interaction with the system.
-
-The strengths of this method stem from its simplicity: it requires
-only a rough estimate of the satellite mass $m_{\rm sat}$ combined
-with backwards integration of orbits. There are no assumptions made
+independent potentiometer. There are also no assumptions made
 about, or N-body modeling of, the internal distribution of satellite
@@
 \citep[for example, the Orphan Stream;][]{belokurov07}. Lastly, the
-present prescription treats the progenitor as a point-mass during
-orbit integration. Second-order effects in orbit evolution (e.g.,
-dynamical friction) are currently ignored, leading to minor biases in
-the estimated potential parameters (for example, XXX in
-Figure~\ref{fig:bootstrap}).
+present prescription neglects second-order effects in orbit evolution (e.g.,
+dynamical friction) which may lead to biases in
+the estimated potential parameters.
 
-[KVJ: Ok, there is a question as to whether we even want to say something like the below...?]
+%With a justified probabilistic framework, we will also be able to provide iven sufficient data, position and motion of the original host could be solved for as additional free parameters, but the effectiveness of this approach has yet to be investigated.
+
+\section{Conclusions and motivation for future work}
+\label{sec:conclusion}
+
+This paper presents an algorithm for measuring the Galactic potential
+that anticipates combined data from the Spitzer and Gaia satellite
+missions that promise precise, full phase-space measurements of RR
+Lyrae stars in the halo of our Galaxy. When applied to a sample of 100
+stars (with realistic observational errors) sampled from the
+\cite{law10} N-body simulation of the destruction of the Sgr dwarf
+satellite, the method recovers the shape and orientation of the dark
+matter potential XXX.
 
@@
 
-%With a justified probabilistic framework, we will also be able to provide iven sufficient data, position and motion of the original host could be solved for as additional free parameters, but the effectiveness of this approach has yet to be investigated.
-
-\section{Conclusions}
-\label{sec:conclusion}
-
-This paper presents an algorithm for measuring the Galactic potential
-that anticipates combined data from the Spitzer and Gaia satellite
-missions that promise precise, full phase-space measurements of RR
-Lyrae stars in the halo of our Galaxy. When applied to a sample of 100
-stars (with realistic observational errors) sampled from the
-\cite{law10} N-body simulation of the destruction of the Sgr dwarf
-satellite, the method recovers the shape and orientation of the dark
-matter potential XXX. This success provides strong motivation for 1)
-further theoretical work to develop a robust generative model (and therefore likelihood) that utilizes the concepts demonstrated by the \emph{Rewinder}, 2) an investigation of the power of
-using multiple, complimentary debris structures (e.g., with different
+This success provides strong motivation for 1)
+further theoretical work to develop a robust generative model (and therefore likelihood) that utilizes the concepts demonstrated by the \emph{Rewinder}, 2) an investigation of the power of using multiple, complimentary debris structures (e.g., with different
 morphologies and orientations) in building a comprehensive picture of
@@
 distances to combine with near-future Gaia velocity data.
-We conclude by noting that this method is complimentary to the orbit fitting algorithm presented in \citet{sanders13b} by favoring warmer, more massive debris structures like Sgr.
-By developing and comparing independent methods for measuring the potential of the Milky Way, we will begin to develop a deeper understanding of the vast structures of dark matter that permeate the universe.
-[Yea, the last sentence I wrote in the late night and am probably delirious...]
 

### 2013-08-01 text/paper1/paper.tex
remove kvj note

@@
 Methods that utilize this knowledge on dynamically cold systems (e.g, tidal streams) require orders of magnitude fewer tracers than a randomly sampled population, but achieve comparable accuracy on derived constraints.
-[KVJ: can we cut some of this?
-Anything from here to end of the paragraph?] In the simplest case, we might {\it assume} that debris stars are actually still on the same orbit as their progenitor system.
+In the simplest case, we might {\it assume} that debris stars are actually still on the same orbit as their progenitor system.
 Then we can imagine that measuring the full-space velocities ${\bf v}$ at different points ${\bf x}$ in the structure (e.g. along a stream) would actually give us a direct

### 2013-08-01 scripts/paper1_figures.py
one legend

@@ -219,21 +219,19 @@ def gaia_spitzer_errors():
     
     # Now add rectangles for Sgr, Orphan
     sgr_d = Rectangle((10., 0.15), 60., 0.15, 
-                      color=sgr_color, alpha=1., label='Sgr stream width')
+                      color=sgr_color, alpha=1., label='Sgr stream')
     axes[0].add_patch(sgr_d)
     
     # From fig. 3 in http://mnras.oxfordjournals.org/content/389/3/1391.full.pdf+html
     orp_d = Rectangle((10., 0.03), 35., 0.03,
-                      color=orp_color, alpha=1., label='Orp stream width')
+                      color=orp_color, alpha=1., label='Orp stream')
     axes[0].add_patch(orp_d)
     
     # Dispersion from Majewski 2004: 10 km/s
-    sgr_v = Rectangle((10., 10), 60., 1., color=sgr_color, alpha=0.75,
-                      label='Sgr stream dispersion')
+    sgr_v = Rectangle((10., 10), 60., 1., color=sgr_color, alpha=0.75)
     axes[1].add_patch(sgr_v)
     
-    orp_v = Rectangle((10., 8.), 35., 1., color=orp_color, alpha=0.75,
-                      label='Orp stream dispersion')
+    orp_v = Rectangle((10., 8.), 35., 1., color=orp_color, alpha=0.75)
     axes[1].add_patch(orp_v)
     
     axes[0].set_ylim(top=10.)
@@ -255,7 +253,7 @@ def gaia_spitzer_errors():
     
     # add legends
     axes[0].legend(loc='upper left', fancybox=True)
-    axes[1].legend(loc='upper left', fancybox=True)
+    #axes[1].legend(loc='upper left', fancybox=True)
     
     fig.subplots_adjust(hspace=0.0)
     plt.tight_layout()

### 2013-08-01 text/paper1/paper.tex
working out conclusions

@@
 
-\author{Adrian~M.~Price-Whelan\altaffilmark{1,2} Kathryn V. Johnston\altaffilmark{1}}
+\author{Adrian~M.~Price-Whelan\altaffilmark{1} Kathryn V. Johnston\altaffilmark{1}}
 
@@
 \altaffiltext{1}{Department of Astronomy, Columbia University, 550 W 120th St., New York, NY 027, USA}
-\altaffiltext{2}{NSF Graduate Fellow}
+%\altaffiltext{2}{NSF Graduate Fellow}
 
@@
   \Sigma_n &= \mathrm{Cov}( \bs{A}_{ik}) \\
-  f(\bs{\theta}) &= \ln \det \Sigma.
+  f(\bs{\theta}) &= \ln \det \Sigma_n.
 \end{align}
@@
 satellite, the method recovers the shape and orientation of the dark
-matter potential XXX.
+matter potential [APW: again, quote errors].
 
-The objective function presented above is an approximation to a more
-general likelihood function, described in depth in forthcoming work
-(Price-Whelan \& Johnston in prep.). The results shown in section
-\ref{sec:results} are meant to illustrate the strength and concept
-behind the method, but [...] [I want to say one more sentence about
-  the upcoming work, and possibly how it will solve some of the
-  limitations presented above.]
+The success of this method provides strong motivation for 1) further theoretical work to develop a robust generative model (and therefore likelihood) that utilizes the concepts demonstrated by the \emph{Rewinder} and 2) an observational campaign with \emph{Spitzer} to survey RR Lyrae stars in debris structures around the Milky Way to get precise
+distances to combine with near-future Gaia velocity data.
+Forthcoming work will develop the \emph{Rewinder} in a probabilistically rigorous regime; this method will then naturally incorporate observational errors, missing data, and more complex dynamical models.
+This work also calls for an investigation of the power of using multiple, complimentary debris structures (e.g., with different morphologies and orientations) in building a comprehensive picture of
+the Galactic mass distribution.
 
-This success provides strong motivation for 1)
-further theoretical work to develop a robust generative model (and therefore likelihood) that utilizes the concepts demonstrated by the \emph{Rewinder}, 2) an investigation of the power of using multiple, complimentary debris structures (e.g., with different
-morphologies and orientations) in building a comprehensive picture of
-the Galactic mass distribution, and 3) a Spitzer survey of RR Lyrae
-stars in debris structures around the Milky Way to get precise
-distances to combine with near-future Gaia velocity data.
+[KVJ: Some concluding thoughts?]
 

### 2013-08-01 scripts/paper1_figures.py
add sgr figure

@@ -15,7 +15,7 @@ import astropy.units as u
 import numpy as np
 import matplotlib.pyplot as plt
 import matplotlib
-from matplotlib import rc_context, rcParams
+from matplotlib import rc_context, rcParams, cm
 from matplotlib.patches import Rectangle
 import triangle
 
@@ -29,12 +29,12 @@ from streams.inference.lm10 import timestep
 from streams.potential import LawMajewski2010
 from streams.potential.lm10 import true_params, param_to_latex
 from streams.integrate.satellite_particles import SatelliteParticleIntegrator
-from streams.io.lm10 import particles_today, satellite_today, time
+from streams.io.lm10 import particle_table, particles_today, satellite_today, time
 
 matplotlib.rc('xtick', labelsize=18)
 matplotlib.rc('ytick', labelsize=18)
 #matplotlib.rc('text', usetex=True)
-matplotlib.rc('axes', edgecolor='#444444', labelsize=24, labelweight=400)
+matplotlib.rc('axes', edgecolor='#444444', labelsize=24, labelweight=400, linewidth=2.0)
 matplotlib.rc('lines', markeredgewidth=0)
 matplotlib.rc('font', family='Source Sans Pro')
 #matplotlib.rc('savefig', bbox='standard')
@@ -345,8 +345,65 @@ def phase_space_d_vs_time(N=10):
     
     return 
 
+def sgr():
+    """ Top-down plot of Sgr particles, with selected stars and then 
+        re-observed
+    """
+    
+    cached_data_file = os.path.join(plot_path, 'sgr_kde.pickle')
+    
+    fig,ax = plt.subplots(1, 1, figsize=(8,8))
+    
+    # read in all particles as a table
+    pdata = particle_table(N=0, expr="(Pcol<7) & (abs(Lmflag)==1)")
+    extent = {'x':(-90,55), 
+              'y':(-58,68)}
+    
+    if not os.path.exists(cached_data_file):    
+        Z = sgr_kde(pdata, extent=extent)
+        
+        with open(cached_data_file, 'w') as f:
+            pickle.dump(Z, f)
+    
+    with open(cached_data_file, 'r') as f:
+        Z = pickle.load(f)
+    
+    ax.imshow(Z**0.5, interpolation="nearest", 
+              extent=extent['x']+extent['y'],
+              cmap=cm.Blues, aspect=1)
[…]

### 2013-08-01 text/paper1/paper.tex
use this version of lm10 figure
use new bootstrap figure placeholder

@@
-%\documentclass{emulateapj}
-\documentclass[preprint]{aastex}
+\documentclass{emulateapj}
+%\documentclass[preprint]{aastex}
 
@@
 \begin{center}
-\includegraphics[width=1.\textwidth]{/Users/adrian/projects/streams/plots/paper1/gaia.pdf}
+\includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/gaia.pdf}
 \caption{ todo }\label{fig:gaia_errors}
@@
 \begin{center}
-\includegraphics[width=1.\textwidth]{/Users/adrian/projects/streams/plots/spitzer_proposal/lm10.png}
+\includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/lm10.png}
 \caption{ todo }\label{fig:lm10}
@@
 \begin{center}
-\includegraphics[width=1.\textwidth]{/Users/adrian/projects/streams/plots/paper1/ps_distance.pdf}
+\includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/ps_distance.pdf}
 \caption{ todo }\label{fig:ps_distance}
@@
 then `observe' a sample of 100 stars from the first leading and
-trailing wraps of the stream; the radial velocity and distance errors are drawn % APW HERE
+trailing wraps of the stream; the radial velocity and distance errors are drawn 
 from Gaussians ($\varepsilon_{\rm RV} \sim \mathcal{N}(\mu=0,\sigma=10~{\rm
   km/s})$ and $\varepsilon_{\rm D} \sim \mathcal{N}(0,0.02\times
-D)$) and the proper motion errors are computed from the expected Gaia error curve. Figure~\ref{fig:variance_proj} shows two-dimensional projections
-of the six-dimensional minimum phase-space distance distribution for
-the observed particles in the correct and an incorrect potential: it
-is clear that the variance of the distribution increases in the wrong
-potential. The generalized variance defines a convex hypersurface over
+D)$) and the proper motion errors are computed from the expected Gaia error curve.
+
+The variance of the six-dimensional minimum phase-space distance distribution increases in the wrong potential. Hence, the generalized variance defines a convex hypersurface over
 which we optimize four of the six logarithmic potential parameters:
@@
 
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=1.\textwidth]{/Users/adrian/projects/streams/plots/paper1/variance_projections.pdf}
-\caption{ todo }\label{fig:variance_proj}
-\end{center}
-\end{figure}
+%\begin{figure}[h]
+%\begin{center}
+%\includegraphics[width=1.\textwidth]{/Users/adrian/projects/streams/plots/paper1/variance_projections.pdf}
+%\caption{ todo }\label{fig:variance_proj}
+%\end{center}
+%\end{figure}
 
@@
 \begin{center}
-\includegraphics[width=1.\textwidth]{/Users/adrian/projects/streams/plots/paper1/objective_function.pdf}
+\includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/objective_function.pdf}
 \caption{ todo }\label{fig:objective}
@@
 \begin{center}
-\includegraphics[width=1.\textwidth]{/Users/adrian/projects/streams/plots/spitzer_proposal/bootstrap.pdf}
+\includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/bootstrap.pdf}
 \caption{ todo }\label{fig:bootstrap}

### 2013-08-02 scripts/paper1_figures.py
fixed colors in gaia error plot

@@ -19,6 +19,7 @@ from matplotlib import rc_context, rcParams, cm
 from matplotlib.patches import Rectangle
 import triangle
 
+from streams.util import project_root
 from streams.observation import apparent_magnitude
 from streams.observation.gaia import parallax_error, proper_motion_error, \
                                      add_uncertainties_to_particles
@@ -85,7 +86,7 @@ def normed_objective_plot():
     fig,ax = plt.subplots(1,1,figsize=(8,8))
     
     nstats = np.linspace(p_range[0], p_range[1], Nbins)
-    linestyles = [(1,'-'), (2,'-.'), (2,'--'), (2,':')]
+    linestyles = [(2,'-'), (3,'-.'), (3,'--'), (3,':')]
     for ii, name in enumerate(['q1','qz','phi','v_halo']):
         vals = variances[name]
         ls = linestyles[ii] 
@@ -178,8 +179,8 @@ def gaia_spitzer_errors():
                 'lines.marker' : None,
                 'axes.facecolor' : '#ffffff'}
     
-    sgr_color = '#67A9CF'
-    orp_color = '#EF8A62'
+    sgr_color = '#2B83BA'
+    orp_color = '#ABDDA4'
     
     with rc_context(rc=rcparams):
         fig,axes = plt.subplots(2, 1, figsize=(8, 10), sharex=True)
@@ -209,10 +210,11 @@ def gaia_spitzer_errors():
             dVtan = (dpm*D).to(u.km*u.radian/u.s).value
         
             # Plot Gaia distance errors
-            axes[0].loglog(D.kiloparsec, (dD/D).decompose(), color='k', alpha=0.1)
+            axes[0].loglog(D.kiloparsec, (dD/D).decompose(), color='#666666', 
+                           alpha=0.1)
                 
             # Plot tangential velocity errors
-            axes[1].loglog(D.kiloparsec, dVtan, color='k', alpha=0.1)
+            axes[1].loglog(D.kiloparsec, dVtan, color='#666666', alpha=0.1)
     
         # Add spitzer 2% line to distance plot
         axes[0].axhline(0.02, linestyle='--', linewidth=4, color='k', alpha=0.75)
@@ -228,10 +230,10 @@ def gaia_spitzer_errors():
     axes[0].add_patch(orp_d)
     
[…]

### 2013-08-02 scripts/paper1_figures.py
minor tweaks

@@ -97,7 +97,7 @@ def normed_objective_plot():
     ax.legend(loc='upper right', fancybox=True)
     
     ax.set_yticks([])
-    ax.set_xlabel('Normalized potential parameter value', labelpad=15)
+    ax.set_xlabel('Normalized parameter value', labelpad=15)
     ax.set_ylabel('Generalized variance', labelpad=15)
     
     fig.savefig(os.path.join(plot_path, "objective_function.pdf"))
@@ -218,6 +218,9 @@ def gaia_spitzer_errors():
     
         # Add spitzer 2% line to distance plot
         axes[0].axhline(0.02, linestyle='--', linewidth=4, color='k', alpha=0.75)
+        
+        # Add photometric 20% line to distance plot
+        axes[0].axhline(0.2, linestyle='-.', linewidth=4, color='k', alpha=0.75)
     
     # Now add rectangles for Sgr, Orphan
     sgr_d = Rectangle((10., 0.15), 60., 0.15, 
@@ -250,7 +253,7 @@ def gaia_spitzer_errors():
     axes[1].set_yticklabels(["{:g}".format(yt) for yt in axes[1].get_yticks()])
     
     # add Gaia and Spitzer text to first plot
-    axes[0].text(4., 0.12, 'Gaia', fontsize=16, rotation=45, fontweight=500)
+    axes[0].text(4., 0.11, 'Gaia', fontsize=16, rotation=32, fontweight=500)
     axes[0].text(4., 0.011, 'Spitzer', fontsize=16, alpha=0.75, fontweight=500)
     
     # add legends

### 2013-08-02 text/paper1/paper_kvj_0802.tex
kathryns recent hack of intro

--- 2013-08-01 text/paper1/paper.tex
@@
  The existence of vast halos of unseen {\it dark} matter surrounding each galaxy has long been proposed to explain the surprisingly large
-motions of the {\it baryonic} matter (i.e. stars, gas and dust) that we can see \citep[e.g.,][]{zwicky37,rubin70}.
+motions of the {\it baryonic} matter (i.e. stars, gas and dust) that we can see \citep[e.g.,][]{rubin70}.
 Dark-matter-only simulations of structure formation lead us to expect that these dark matter halos should have density distributions that are described by a universal radial profile \citep{navarro96} with a variety of triaxial shapes \citep{jing02}.
@@
 The Milky Way is the best candidate for such a detailed study of a dark matter halo. 
-Our presence in the Galaxy means that we can resolve and use individual stars as tracers and hence build large samples that probe deeper into the halo than globular cluster
-or planetary nebula studies of external galaxies \citep[e.g.][]{mendez01,cote03}.
+Our presence in the Galaxy means that we can resolve and use individual stars as tracers and hence build large samples that probe deep into the halo.
+%er into the halo than globular cluster or planetary nebula studies of external galaxies \citep[e.g.][]{mendez01,cote03}.
 Our internal perspective allows us to measure the three-dimensional mass distribution around the Galaxy rather than a two-dimensional projection, such as in lensing studies \citep{bolton06}.
@@
 kpc \citep[SDSS, see][]{deason12a,kafle12}, and estimates with combined tracers extend to 150kpc \citep{deason12b}.
-\citet{loebman12} demonstrates that even larger stellar samples from current (e.g., SDSS) and future  (e.g., Gaia and LSST) stellar surveys will also be sensitive to any
-asphericity of the halo.
+%\citet{loebman12} demonstrates that even larger stellar samples from current (e.g., SDSS) and future  (e.g., Gaia and LSST) stellar surveys will also be sensitive to any asphericity of the halo.
 
@@
 A complimentary method to measure the mass distribution is to instead take full advantage of the {\it non}-random nature of the stellar distribution in the Galactic halo.
-While we do not know the exact orbits of  stellar debris from satellite destruction --- though \citet{eyre09b} propose a method of full orbit reconstruction from incomplete kinematic data --- we do know something more about stars in a debris structure: the stars were once all part of the same object.
-Methods that utilize this knowledge on dynamically cold systems (e.g, tidal streams) require orders of magnitude fewer tracers than a randomly sampled population, but achieve comparable accuracy on derived constraints.
-In the simplest case, we might {\it assume} that debris stars are actually still on the same orbit as their progenitor system.
-Then we can imagine that measuring the full-space velocities ${\bf v}$ at different points ${\bf x}$ in the structure (e.g. along a stream) would actually give us a direct
-measure of differences in potential $\Phi$ (i.e. $\Phi({\bf x}_1)-\Phi({\bf x}_2)=(v_2^2-v_1^2)/2$).
+While we do not know the exact orbits of  stellar debris from satellite destruction 
+%--- though \citet{eyre09b} propose a method of full orbit reconstruction from incomplete kinematic data --- 
+we do know something more about stars in a debris structure: the stars were once all part of the same object.
+Methods that utilize this knowledge on dynamically cold  tidal streams can require orders of magnitude fewer tracers than a randomly sampled population to achieve comparable accuracy on derived constraints.
+%In the simplest case, we might {\it assume} that debris stars are actually still on the same orbit as their progenitor system.
+%Then we can imagine that measuring the full-space velocities ${\bf v}$ at different points ${\bf x}$ in the structure (e.g.
+along a stream) would actually give us a direct measure of differences in potential $\Phi$ (i.e.
+$\Phi({\bf x}_1)-\Phi({\bf x}_2)=(v_2^2-v_1^2)/2$).
 [KVJ: incomplete sentence?
-Not sure what the point you were making is?] In reality, we usually know at most four of the six phase-space co-ordinates for debris stars, often with significant error bars and potential measurements have been made by fitting orbits to streams \citep[e.g.,][]{helmi04,johnston05,koposov10,law10}.
-Moreover, our assumption of debris tracing a single orbit is actually incorrect \citep[see][for discussions of the orbit distribution in tidal debris]{johnston98,helmi99}, with
+Not sure what the point you were making is?] In reality, With current data, we usually know at most four of the six phase-space co-ordinates for debris stars, often with significant error bars and
+One approach to potential measurement is to simply fit orbits to observations of streams 
+%\citep[e.g.,][]{helmi04,johnston05,koposov10}.
+\citep[e.g.,][]{koposov10}.
+However, the assumption that debris traces a single orbit is actually incorrect \citep[see][for discussions of the orbit distribution in tidal debris]{johnston98,helmi99}, with
 systematic changes in orbital energy along debris streams that can lead to systematic biases in our measurements of the Galactic potential \citep{eyre09a,varghese11}.
@@
 distribution of orbital properties in the debris and along tidal streams) with the aim of simultaneously constraining both the properties of the satellite and the Milky Way.
-Many studies of the Sagittarius debris system (hereafter Sgr) --- whose tidal streams entirely encircle the Milky Way --- have adopted this approach \citep[e.g.][]
-{law05,fellhauer06} with the most recent work for the first time attempting to place constraints on the triaxiality and orientation of the Milky Way's matter distribution
+Many studies of the Sagittarius debris system (hereafter Sgr) --- whose tidal streams entirely encircle the Milky Way --- have adopted this approach 
+%\citep[e.g.][]{law05,fellhauer06} 
+with the most recent work for the first time attempting to place constraints on the triaxiality and orientation of the Milky Way's matter distribution
 \citep{law10}.
@@
 There is a long tradition for using RR Lyrae stars in the Galaxy to
-study structure \citep[going back to first estimates of the distance
-  to the Galactic center][]{shapley18}, substructure
+study structure 
+%citep[going back to first estimates of the distanceto the Galactic center]
+\citep[e.g.][]{shapley18}, substructure
 \citep[e.g.][]{sesar10}, and distances to satellite galaxies
 \citep[e.g.][]{clementini03}.  However, studies of RR Lyrae at optical
-wavelengths are limited by both metallicity affects on the intrinsic
+wavelengths are limited by both metallicity effects on the intrinsic
 brightness of these stars and variable extinction along the line of
@@
 Lyrae stars was first mapped by \citet{longmore86} using $K$-band
-observations of stars in Galactic globular clusters and later shown to
-be due to the changing bolometric correction for these stars as a
-function of the observed wavelength \citep{catelan04}.
-\citet{madore12} have recently shown that the intrinsic scatter of the
-PL relation decreases with increasing wavelength. Both reddening
-effects and metallicity dependencies also decrease significantly for
-observations further into the IR.
-
-These studies suggest that observations of RR Lyrae variables by
-NASA's Spitzer mission could provide stellar distance indicators of
-unprecedented accuracy and precision.  Using five stars with
+observations of stars in Galactic globular clusters 
+%and later shown to be due to the changing bolometric correction for these stars as a function of the observed wavelength \citep{catelan04}.
+\citet{madore12} have recently shown, 
+using five stars with
 trigonometric parallaxes measured by Hubble \citep{benedict11},
-\citet{madore12} measured the dispersion in the mid-IR PL relation at
-Spitzer wavelengths to be $\sim$0.03~mag. This implies that it is
-possible to determine individual distances that are good to better
-than $2\%$ for RR Lyrae stars out to $\sim$60~kpc (the limit of
+that the dispersion in the mid-IR PL relation at
+wavelengths measurable by NASA's {\it Spitzer} mission is $\sim$0.03~mag.
+This implies that it is
+possible to use {\it Spitzer} to determine distances that are good to better
+than $2\%$ for individual RR Lyrae stars out to $\sim$60~kpc (the limit of
 Spitzer's ability to detect and measure RR Lyrae variables for modest
-integration times). Distance measurements of Blue Horizontal Branch
-and Blue Straggler stars at these distances in the halo typically
-achieve $\sim$25\% uncertainties \citep[e.g.,][]{deason12b}.
+integration times). 
+For comparison, distance measurements of Blue Horizontal Branch
+and Blue Straggler stars at typically
+achieve $\sim$10-15\% uncertainties, and then only if appropriate color measurements are available \citep[e.g.,][]{deason12b}.
 
@@
 -1.5). Within 2~kpc, Gaia will measure distances to these stars with
-better than 2\% accuracy --- RR Lyrae in this volume will be used to
+better than 2\% accuracy --- RR Lyrae in this volume can be used to
 test and calibrate the Spitzer PL relation described above. Beyond the
@@
 \includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/gaia.pdf}
-\caption{ todo }\label{fig:gaia_errors}
+\caption{Expected \emph{Gaia} distance and tangential velocity errors as a function of heliocentric distance for RR Lyrae stars.
+Errors are a function of color and magnitude of the source, which depend on the metallicity.
+Each grey line  is computed by Monte Carlo sampling from the empirical metallicity distribution of
+the Galactic halo from \cite{ivezic08} (centered around [Fe/H] $\sim$ -1.5), and using [...] to compute the absolute magnitude of the star.
+Note in top panel, parallax distance errors from \emph{Gaia} are larger than the line-of-sight size of both Sgr and Orphan (Orp), but photometric distance errors are comparable to the the Sgr scale.
+Bottom panel shows that the \emph{Gaia} tangential velocity errors (proper motion error $\times$ distance) are smaller than the internal velocity dispersion of both Sgr and Orp.
+}\label{fig:gaia_errors}
 \end{center}
@@
 of where full six-dimensional phase-space maps of the Galaxy are
-possible from $<$10~kpc to 80~kpc. This enormous increase in volume
+possible from $<$10~kpc to 60~kpc. This enormous increase in volume
 will greatly refine data on existing debris systems such as the
@@
 serendipitously during a radial velocity survey of the Galactic bulge
-\citep{ibata94}. While the number of known Milky Way satellites has
-more than doubled recently \citep[e.g.,][]{mcconnachie12}, in many
-ways Sgr remains the most studied and most intriguing: it is the
-closest satellite, among the largest (only the LMC and SMC are more
-luminous), and has the most prominent tidal tails.
-
-The first maps of the Sgr core \citep{ibata94, ibata95} revealed its
-elongated morphology and suggested tidal deformation, as would be
-anticipated from its proximity to the Galaxy. Subsequent N-body models
-confirmed this interpretation \citep{velazquez95} and showed that
-debris from Sgr should form coherent streams of stars that encircle
-the Galaxy \citep{johnston95}. Signatures of these extensive stellar
-streams have since been discovered in many studies and have been
+\citep{ibata94}. 
+%While the number of known Milky Way satellites has more than doubled recently \citep[e.g.,][]{mcconnachie12}, in many ways Sgr remains the most studied and most intriguing: it is the closest satellite, among the largest, and has the most prominent tidal tails.
+%The first maps of the Sgr core \citep{ibata94, ibata95} revealed its elongated morphology and suggested tidal deformation, as would be anticipated from its proximity to the Galaxy.
+Subsequent N-body models confirmed this interpretation \citep{velazquez95} and showed that debris from Sgr should form coherent streams of stars that encircle the Galaxy \citep{johnston95}.
+Signatures of extensive stellar
+streams associated with Sgr have since been discovered in many studies and have been
 mapped across the sky in carbon stars \citep{totten98}, M giants
@@
 \citep{belokurov06}, and RR Lyrae in the Catalina Sky Survey
-\citep{drake13}. More recently, apogalactic debris has been traced to
-large distances using a combination of blue horizontal branch,
+\citep{drake13}. Apogalactic debris has been traced out to $\sim 100$kpc 
+using a combination of blue horizontal branch,
 main-sequence turn-off, and red giant stars from the SDSS
 \citep{belokurov13}.
+The distance to the stream has been estimated from photometric surveys \citep[e.g.,][]{martinezdelgado04} and gradients in radial velocities and metallicity measured with follow-up spectroscopy\citep[e.g.,][]{majewski04, vivas05, bellazzini06, chou10, keller10}.
 
-The distance to the stream has been estimated from photometric surveys
-\citep[e.g.,][]{martinezdelgado04} and gradients in radial
-velocities and metallicity measured with follow-up spectroscopy
-\citep[e.g.,][]{majewski04, vivas05, bellazzini06, chou07,
-  chou10, keller10, carlin12}. The multitude of Sgr stream data
-has inspired a rich set of interpretations and models of the debris
-\citep[e.g.,][]{johnston99b, helmi04, law05, fellhauer06,
-  law10}.
-
+This multitude of Sgr stream data has inspired a rich set of interpretations and models of the debris
+\citep[e.g.,][]{johnston99b, fellhauer06}.
 Most recently, \citet[][hereafter LM10]{law10} have combined all
-current data on the Sgr debris to assess the triaxiality and
-orientation of the outer Galaxy, the first time that such a
-reconstruction of the 3-dimensional mass distribution of a dark matter
-halo has been feasible. Figure \ref{fig:lm10} shows particle positions
-(in Galactocentric coordinates) from the final time-step of an N-body
+current data on the Sgr debris to constrain both a model of its evolution
+and the potential in which it orbits.
+Figure \ref{fig:lm10} shows particle positions
+(in Galactocentric coordinates) from the final time-step of the LM10 N-body
 simulation of dwarf satellite disruption along the expected Sgr orbit
-in the best-fitting Milky Way halo model from LM10. The simulation was
+in the best-fitting Milky Way halo model. The simulation was
 run in a three-component potential, with a triaxial, logarithmic halo
@@
 \includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/lm10.png}
-\caption{ todo }\label{fig:lm10}
+\caption{ Particle density (blue) of the first leading and trailing wraps from the final time-step of the \citet{law10} simulation of the Sgr stream.
+The density is estimated using Kernel Density Estimation (KDE) with a Gaussian kernel on $\sim$40,000 particle positions.
+Point markers (black) show positions of a representative sample of 100 stars drawn from this density distribution.
+The position of the Sun is shown with the solar symbol.
+}\label{fig:lm10}
 \end{center}
@@
 its mass and stellar populations, orbit, rate of destruction, and even
-original distribution in populations. The phase-space distribution of
-the debris also offers a unique probe of the depth, shape, extent, and
-orientation of the Milky Way's dark matter halo that cannot be matched
-using other techniques on our own or other galaxies
-\citep[e.g.,][]{ibata01}. The LM10 constraints on Sgr's mass and orbit
-have invigorated discussions of Sgr in a more cosmological context,
-the mass and extent of the original dark matter halo that hosted it,
-as well as its effect on the Milky Way itself
-\citep{bailin03,purcell11,micheldansac11,gomez12}.
+original distribution in populations. 
+%The LM10 constraints on Sgr's mass and orbit have invigorated discussions of Sgr in a more cosmological context, the mass and extent of the original dark matter halo that hosted it, as well as its effect on the Milky Way itself \citep{bailin03,purcell11,micheldansac11,gomez12}.
+Moreover, this comparison enabled the first 
+reconstruction of the 3-dimensional mass distribution of a dark matter
+halo through the constraints on the potential parameters $v_{\rm halo}$,$q_1$, $q_z$ and $\phi$.
 
@@
 \begin{center}
-\includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/ps_distance.pdf}
-\caption{ todo }\label{fig:ps_distance}
+\includegraphics[width=0.5\textwidth]{/Users/adrian/projects/streams/plots/paper1/ps_distance.pdf}
+\caption{Phase space distance ($D_{\rm ps}$) for 10 stars integrated backwards in the correct potential (top) and a potential where $q_z$ is 25\% larger (bottom).
+The same 10 particles are used in both figures, so the initial conditions are identical.
+Red (thick, solid line) shows the galactocentric distance of the progenitor (in arbitrary, normalized units) over the integration.
+Horizontal (dashed) line shows $D_{\rm ps}=\sqrt{2}$.}\label{fig:ps_distance}
 \end{center}
@@
 \includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/objective_function.pdf}
-\caption{ todo }\label{fig:objective}
+\caption{ 1D slices of the objective function (generalized variance) for each halo potential parameter.
+Curves are normalized to have the same value at the true parameter value (1.00), and show the effect of varying each parameter by $\pm10$\%.
+}\label{fig:objective}
 \end{center}
@@
 \includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/bootstrap.pdf}
-\caption{ todo }\label{fig:bootstrap}
+\caption{ Black points show the ``best-fit'' parameters resulting from each bootstrap resample of 100 stars from the Sgr stream particle density shown in Figure~\ref{fig:lm10}.
+Blue (vertical and horizontal) lines show the true values of the parameters.
+}\label{fig:bootstrap}
 \end{center}

### 2013-08-02 scripts/paper1_figures.py
sexy error margins

@@ -17,7 +17,6 @@ import matplotlib.pyplot as plt
 import matplotlib
 from matplotlib import rc_context, rcParams, cm
 from matplotlib.patches import Rectangle, Ellipse
-from scipy.stats import gaussian_kde
 
 from streams.util import project_root
 from streams.observation import apparent_magnitude
@@ -482,42 +481,32 @@ def bootstrapped_parameters():
 
     y_param = 'v_halo'
     x_params = ['q1', 'qz', 'phi']
-    
-    # for KDE
-    X,Y = np.mgrid[-0.1:0.1:50j, -0.1:0.1:50j] 
-    positions = np.vstack([X.ravel(),Y.ravel()])
-    
+
     for ii,x_param in enumerate(x_params):
         ydata = (np.array(data[y_param])-_true_params[y_param]) / _true_params[y_param]
         xdata = (np.array(data[x_param])-_true_params[x_param]) / _true_params[x_param]
-    
-        #values = np.vstack([ydata, xdata])
-        #kernel = gaussian_kde(values)
-        #Z = np.reshape(kernel(positions).T, X.shape).T
-        #Z = np.flipud(Z)
         
-        #print(kernel.covariance)
-        #axes[ii].imshow(Z, interpolation="nearest", 
-        #                extent=[-0.1,0.1,-0.1,0.1],
-        #                cmap=cm.Blues, aspect=1)
+        axes[ii].axhline(0., linewidth=2, color='#ABDDA4', alpha=0.75, zorder=-1)
+        axes[ii].axvline(0., linewidth=2, color='#ABDDA4', alpha=0.75, zorder=-1)
+        
         points = np.vstack([ydata, xdata]).T
         plot_point_cov(points, nstd=2, ax=axes[ii], alpha=0.25, color='#777777')
         plot_point_cov(points, nstd=1, ax=axes[ii], alpha=0.5, color='#777777')
         plot_point_cov(points, nstd=2, ax=axes[ii], color='#000000', fill=False)
         plot_point_cov(points, nstd=1, ax=axes[ii], color='#000000', fill=False)
         
-        axes[ii].axhline(0., linewidth=1, color='#2B8CBE', alpha=0.5)
-        axes[ii].axvline(0., linewidth=1, color='#2B8CBE', alpha=0.5)
-        axes[ii].plot(ydata, xdata, marker='o', alpha=0.75, linestyle='none')
[…]

### 2013-08-02 text/paper1/paper.tex
make figure bigger in text
figure captions

@@
 \includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/bootstrap.pdf}
-\caption{ todo }\label{fig:bootstrap}
+\caption{ Blue points show the ``best-fit'' parameters resulting from each bootstrap resample of 100 stars from the Sgr stream particle density shown in Figure~\ref{fig:lm10}.
+Green (vertical and horizontal) lines show the true values of the parameters.
+Grey ellipses show one- and two-sigma margins, assuming the points are normally distributed.
+}\label{fig:bootstrap}
 \end{center}
@@
 
-%With a justified probabilistic framework, we will also be able to provide iven sufficient data, position and motion of the original host could be solved for as additional free parameters, but the effectiveness of this approach has yet to be investigated.
+%With a justified probabilistic framework, we will also be able to provide ...
+given sufficient data, position and motion of the original host could be solved for as additional free parameters, but the effectiveness of this approach has yet to be investigated.
 

### 2013-08-07 text/paper1/paper.tex
god damn oxford comma
transpose bootstrap figure

@@
 
-\title{Spitzer, Gaia and the Potential of the Milky Way}
+\title{Spitzer, Gaia, and the Potential of the Milky Way}
 
@@
 Each grey line  is computed by Monte Carlo sampling from the empirical metallicity distribution of
-the Galactic halo from \cite{ivezic08} (centered around [Fe/H] $\sim$ -1.5), and using [...] to compute the absolute magnitude of the star.
-Note in top panel, parallax distance errors from \emph{Gaia} are larger than the line-of-sight size of both Sgr and Orphan (Orp), but photometric distance errors are comparable to the the Sgr scale.
+the Galactic halo from \cite{ivezic08} (centered around [Fe/H] $\sim$ -1.5), and using eq.~14 from \citet{benedict11} to compute the absolute magnitude of the star.
+Note in top panel, parallax distance errors from \emph{Gaia} are larger than the line-of-sight size of both Sgr and Orphan (Orp), but photometric distance errors are comparable to the the Sgr scale (assuming 10\% errors, dotted line).
 Bottom panel shows that the \emph{Gaia} tangential velocity errors (proper motion error $\times$ distance) are smaller than the internal velocity dispersion of both Sgr and Orp.
@@
 
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/bootstrap.pdf}
+%\begin{figure}[h]
+%\begin{center}
+%\includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/bootstrap.pdf}
+%\caption{ Blue points show the ``best-fit'' parameters resulting from each bootstrap resample of 100 stars from the Sgr stream particle density shown in Figure~\ref{fig:lm10}.
+Green (vertical and horizontal) lines show the true values of the parameters.
+Grey ellipses show one- and two-sigma margins, assuming the points are normally distributed.
+}\label{fig:bootstrap}
+%\end{center}
+%\end{figure}
+
+\begin{figure*}[!t]
+\centering\includegraphics[width=0.9\textwidth,trim=0 0 0 0, clip]{/Users/adrian/projects/streams/plots/paper1/bootstrap.pdf}
 \caption{ Blue points show the ``best-fit'' parameters resulting from each bootstrap resample of 100 stars from the Sgr stream particle density shown in Figure~\ref{fig:lm10}.
@@
 }\label{fig:bootstrap}
-\end{center}
-\end{figure}
+\end{figure*}
 

### 2013-08-07 scripts/paper1_figures.py

@@ -219,7 +219,7 @@ def gaia_spitzer_errors():
         axes[0].axhline(0.02, linestyle='--', linewidth=4, color='k', alpha=0.75)
         
         # Add photometric 20% line to distance plot
-        axes[0].axhline(0.2, linestyle='-.', linewidth=4, color='k', alpha=0.75)
+        axes[0].axhline(0.1, linestyle=':', linewidth=4, color='k', alpha=0.75)
     
     # Now add rectangles for Sgr, Orphan
     sgr_d = Rectangle((10., 0.15), 60., 0.15, 
@@ -522,10 +522,64 @@ def bootstrapped_parameters():
     fig.subplots_adjust(hspace=0., wspace=0.)
     fig.savefig(os.path.join(plot_path, "bootstrap.pdf"))
 
+def bootstrapped_parameters_transpose():
+    data_file = os.path.join(project_root, "plots", "hotfoot", 
+                             "SMASH", "all_best_parameters.pickle")
+    
+    with open(data_file) as f:
+        data = pickle.load(f)
+    
+    rcparams = {'axes.linewidth' : 3.,
+                'xtick.major.size' : 8.}
+    with rc_context(rc=rcparams): 
+        fig,axes = plt.subplots(1,3,figsize=(12,5))
+
+    y_param = 'v_halo'
+    x_params = ['q1', 'qz', 'phi']
+
+    for ii,x_param in enumerate(x_params):
+        
+        xdata = (np.array(data[x_param])-_true_params[x_param]) / _true_params[x_param]
+        ydata = (np.array(data[y_param])-_true_params[y_param]) / _true_params[y_param]
+        
+        axes[ii].axhline(0., linewidth=2, color='#ABDDA4', alpha=0.75, zorder=-1)
+        axes[ii].axvline(0., linewidth=2, color='#ABDDA4', alpha=0.75, zorder=-1)
+        
+        points = np.vstack([xdata, ydata]).T
+        plot_point_cov(points, nstd=2, ax=axes[ii], alpha=0.25, color='#777777')
+        plot_point_cov(points, nstd=1, ax=axes[ii], alpha=0.5, color='#777777')
+        plot_point_cov(points, nstd=2, ax=axes[ii], color='#000000', fill=False)
+        plot_point_cov(points, nstd=1, ax=axes[ii], color='#000000', fill=False)
+        
[…]

### 2013-08-07 text/paper1/paper.tex
the rewinder -> rewinder

@@
 about, or N-body modeling of, the internal distribution of satellite
-stars. Thus, the \emph{Rewinder} is applicable to any debris that is
+stars. Thus, \emph{Rewinder} is applicable to any debris that is
 known to come from a single object and not restricted to the coldest
@@
 
-The success of this method provides strong motivation for 1) further theoretical work to develop a robust generative model (and therefore likelihood) that utilizes the concepts demonstrated by the \emph{Rewinder} and 2) an observational campaign with \emph{Spitzer} to survey RR Lyrae stars in debris structures around the Milky Way to get precise
+The success of this method provides strong motivation for 1) further theoretical work to develop a robust generative model (and therefore likelihood) that utilizes the concepts demonstrated by \emph{Rewinder} and 2) an observational campaign with \emph{Spitzer} to survey RR Lyrae stars in debris structures around the Milky Way to get precise
 distances to combine with near-future Gaia velocity data.
-Forthcoming work will develop the \emph{Rewinder} in a probabilistically rigorous regime; this method will then naturally incorporate observational errors, missing data, and more complex dynamical models.
+Forthcoming work will develop \emph{Rewinder} in a probabilistically rigorous regime; this method will then naturally incorporate observational errors, missing data, and more complex dynamical models.
 This work also calls for an investigation of the power of using multiple, complimentary debris structures (e.g., with different morphologies and orientations) in building a comprehensive picture of

### 2013-08-07 scripts/paper1_figures.py
add error estimate

@@ -560,7 +560,7 @@ def bootstrapped_parameters_transpose():
         axes[ii].set_yticks([-0.1, -0.05, 0., 0.05, 0.1])
         axes[ii].xaxis.tick_bottom()
     
-    axes[2].set_xlabel(r"$\delta v_{\rm halo}$", 
+    axes[0].set_ylabel(r"$\delta v_{\rm halo}$", 
                        fontsize=26, rotation='horizontal')
     
     axes[0].set_yticks([-0.1, -0.05, 0., 0.05, 0.1])
@@ -575,6 +575,23 @@ def bootstrapped_parameters_transpose():
     fig.subplots_adjust(hspace=0., wspace=0.)
     fig.savefig(os.path.join(plot_path, "bootstrap.pdf"))
 
+def parameter_errors():
+    data_file = os.path.join(project_root, "plots", "hotfoot", 
+                             "SMASH", "all_best_parameters.pickle")
+    
+    with open(data_file) as f:
+        data = pickle.load(f)
+    
+    params = ['q1', 'qz', 'phi', 'v_halo']
+    d = np.vstack(tuple([data[p] for p in params]))
+    cov = np.cov(d)
+    
+    errors = np.sqrt(np.diag(cov))
+    print("q1", errors[0])
+    print("qz", errors[1])
+    print("phi", np.degrees(errors[2]))
+    print("v_halo", (errors[3]*u.kpc/u.Myr).to(u.km/u.s).value)
+
 if __name__ == '__main__':
     #gaia_spitzer_errors()
     #sgr()
@@ -582,4 +599,5 @@ if __name__ == '__main__':
     #normed_objective_plot()
     #variance_projections()
     #bootstrapped_parameters()
-    bootstrapped_parameters_transpose()
+    #bootstrapped_parameters_transpose()
+    parameter_errors()

### 2013-08-07 text/paper1/paper.tex
add kathryn's suggestion

@@
 We conclude in Section \ref{sec:conclusion}.
-
-% --------------------------------------------------------------------
-% Above here is from KVJ. Below is APW.
-%
 
@@
 
-The success of this method provides strong motivation for 1) further theoretical work to develop a robust generative model (and therefore likelihood) that utilizes the concepts demonstrated by \emph{Rewinder} and 2) an observational campaign with \emph{Spitzer} to survey RR Lyrae stars in debris structures around the Milky Way to get precise
-distances to combine with near-future Gaia velocity data.
-Forthcoming work will develop \emph{Rewinder} in a probabilistically rigorous regime; this method will then naturally incorporate observational errors, missing data, and more complex dynamical models.
-This work also calls for an investigation of the power of using multiple, complimentary debris structures (e.g., with different morphologies and orientations) in building a comprehensive picture of
-the Galactic mass distribution.
+While the tests presented in this paper are very simple, the accuracy of potential recovery promised by such a small sample of stars 
+provides strong motivation for further theoretical work to: 1) develop a robust generative model (and therefore likelihood) that utilizes the concepts demonstrated by \emph{Rewinder}; 2) investigate the power of using multiple debris structures; and 3) examine how \emph{Rewinder} might work with less accurate measurements or missing dimensions.
 
-[KVJ: Some concluding thoughts?]
+Our results also motivate an observational campaign with \emph{Spitzer} to survey RR Lyrae stars in debris structures around the Milky Way to get precise distances to combine with near-future Gaia velocity data.
+If just 100 stars in a single stellar stream allow us to measure the depth, shape, and orientation of the Milky Way potential, larger samples in multiple structures offer the prospect of assessing these quantities as a function of Galactocentric radius.
+Tracing the mass in a dark matter halo with this level of detail is impossible for any other galaxy in the Universe.
 

### 2013-08-07 text/paper1/paper.tex
with bib, pre-chopping

@@
 \begin{abstract}
-.
+
+Near-future data from ESA's Gaia mission will provide precise, full phase-space information for hundreds of millions of stars out to heliocentric distances of a few kiloparsecs.
+This ``horizon'' can be extended with more accurate distances, allowing for more precise tangential velocity measurements out to several tens of kiloparsecs if we limit studies to samples of RR Lyrae stars.
+Recent work has demonstrated how Spitzer observations of RR Lyrae stars can be used to make distance estimates accurate to 2\%, effectively extending the Gaia, precise-data horizon by a factor of ten in distance and a factor of 1000 in volume.
+In this \emph{Letter} we develop and demonstrate a new method that uses small samples of stars associated with debris from satellite destruction to measure the Galactic potential.
+We then apply this method to simulated observations of stars from the end point of a simulation of satellite destruction and discuss the success of the algorithm in recovering the true potential used for the simulation.
+The results suggest that we will be able to probe the three-dimensional structure of the Galaxy's dark matter halo, constraining its shape, orientation, and depth as a function of radius.
+The opportunity to examine a dark matter halo in such great detail is truly unique and is enabled by the combination of Spitzer and Gaia with our intimate perspective on our own Galaxy.
+
 \end{abstract}
@@
 Our presence in the Galaxy means that we can resolve and use individual stars as tracers and hence build large samples that probe deep into the halo.
-%er into the halo than globular cluster or planetary nebula studies of external galaxies \citep[e.g.][]{mendez01,cote03}.
-Our internal perspective allows us to measure the three-dimensional mass distribution around the Galaxy rather than a two-dimensional projection, such as in lensing studies \citep{bolton06}.
-%Hence we can hope to reconstruct not only the density, but also the shape and orientation of the dark matter halo as function of radius well beyond the disk.
 To date, thousands of blue horizontal branch stars selected from the Sloan Digital Sky Survey (SDSS) have been used to probe the Milky Way mass out to tens of
 kpc \citep[SDSS, see][]{deason12a,kafle12}, and estimates with combined tracers extend to 150kpc \citep{deason12b}.
-%\citet{loebman12} demonstrates that even larger stellar samples from current (e.g., SDSS) and future  (e.g., Gaia and LSST) stellar surveys will also be sensitive to any asphericity of the halo.
 
@@
 However, the same surveys that provide large samples of tracers have also revealed the existence of large-scale spatial inhomogeneities in the form of giant stellar streams and clouds \citep{newberg02,majewski03,belokurov06}, clearly demonstrating that a significant fraction of the stellar halo is neither randomly sampled from a smooth distribution function nor is fully phase-mixed.
-%The degree of substructure increases with Galactocentric radius \citep{bell08} in agreement with models where the stellar halo is formed purely from accretion events \citep{bullock05,cooper10,rashkov12}.
-%Moreover, even in the smoother, inner halo which is likely to be phase-mixed, significant clumpings in velocity \citep{schlauffman10} and orbital properties \citep{helmi99} have been observed, which indicate a non-smooth component to the orbit distribution.
-%By ``observing'' particles in a sample of simulated stellar halos, \citet{yencho06} showed that, if stellar halos are indeed entirely formed via accretion events, then the expected degree of non-randomness  could lead to systematic biases in mass estimates of order several tens of percent.
-%This encouraging result suggests that, if our expectations for the typical number and size of accretion events contributing to the stellar halo are correct, then the number of orbits effectively explored is sufficiently large that the assumption of random, fully phase-mixed orbits is reasonable for the current samples.
 
-% --- analogous to our exploitation of matter in disks moving on near-circular orbits.
 A complimentary method to measure the mass distribution is to instead take full advantage of the {\it non}-random nature of the stellar distribution in the Galactic halo.
 While we do not know the exact orbits of  stellar debris from satellite destruction 
-%--- though \citet{eyre09b} propose a method of full orbit reconstruction from incomplete kinematic data --- 
 we do know something more about stars in a debris structure: the stars were once all part of the same object.
 Methods that utilize this knowledge on dynamically cold  tidal streams can require orders of magnitude fewer tracers than a randomly sampled population to achieve comparable accuracy on derived constraints.
-%In the simplest case, we might {\it assume} that debris stars are actually still on the same orbit as their progenitor system.
-%Then we can imagine that measuring the full-space velocities ${\bf v}$ at different points ${\bf x}$ in the structure (e.g.
-along a stream) would actually give us a direct measure of differences in potential $\Phi$ (i.e.
-$\Phi({\bf x}_1)-\Phi({\bf x}_2)=(v_2^2-v_1^2)/2$).
-[KVJ: incomplete sentence?
-Not sure what the point you were making is?] In reality, With current data, we usually know at most four of the six phase-space co-ordinates for debris stars, often with significant error bars and
 One approach to potential measurement is to simply fit orbits to observations of streams 
-%\citep[e.g.,][]{helmi04,johnston05,koposov10}.
 \citep[e.g.,][]{koposov10}.
@@
 Many studies of the Sagittarius debris system (hereafter Sgr) --- whose tidal streams entirely encircle the Milky Way --- have adopted this approach 
-%\citep[e.g.][]{law05,fellhauer06} 
 with the most recent work for the first time attempting to place constraints on the triaxiality and orientation of the Milky Way's matter distribution
 \citep{law10}.
-%Studies which first fit the Sgr stream to a single orbit \citep{johnston05,law09} and were subsequently followed up with full N-body simulations \citep{law05,law12} offer another demonstration of the limitation of the former approach as potential parameters are revised when moving to the N-body approach.
 
@@
 \citet{sanders13b} examine the distribution of debris in action-angle co-ordinates and show that stars stripped from the same disrupted object must lie along a single line in angle-frequency space, providing a constraint that can be used as a potential measure.
-%\citet{johnston99a} (hereafter JZSH) showed that backwards integration of the positions of stars and their parent satellite could be used to distinguish Milky Way mass models: only in the correct potential would the stars' and satellite's  path coincide within a Hubble time.The JZSH method was designed and tested for the proposed characteristics of the Space Interferometry Mission \citep{}, which was expected to be able to measure proper motions with $\mu$as/year accuracy for stars as faint as $V=20$.
-Distances to stream members would only be poorly known and were estimated using anticipated properties of tidal debris.
 
@@
 study structure 
-%citep[going back to first estimates of the distanceto the Galactic center]
 \citep[e.g.][]{shapley18}, substructure
@@
 observations of stars in Galactic globular clusters 
-%and later shown to be due to the changing bolometric correction for these stars as a function of the observed wavelength \citep{catelan04}.
 \citet{madore12} have recently shown, 
@@
 \citep{ibata94}.
-%While the number of known Milky Way satellites has more than doubled recently \citep[e.g.,][]{mcconnachie12}, in many ways Sgr remains the most studied and most intriguing: it is the closest satellite, among the largest, and has the most prominent tidal tails.
-%The first maps of the Sgr core \citep{ibata94, ibata95} revealed its elongated morphology and suggested tidal deformation, as would be anticipated from its proximity to the Galaxy.
-Subsequent N-body models confirmed this interpretation \citep{velazquez95} and showed that debris from Sgr should form coherent streams of stars that encircle the Galaxy \citep{johnston95}.
+
 Signatures of extensive stellar
@@
 original distribution in populations. 
-%The LM10 constraints on Sgr's mass and orbit have invigorated discussions of Sgr in a more cosmological context, the mass and extent of the original dark matter halo that hosted it, as well as its effect on the Milky Way itself \citep{bailin03,purcell11,micheldansac11,gomez12}.
 Moreover, this comparison enabled the first 
@@
 the orbits of the stars will diverge from that of the progenitor and
-thus will not be recaptured by the satellite system
-(Figure~\ref{fig:ps_distance}).
+thus will not be recaptured by the satellite system.
+
+%(Figure~\ref{fig:ps_distance}).
 
@@
 \begin{center}
-\includegraphics[width=0.5\textwidth]{/Users/adrian/projects/streams/plots/paper1/ps_distance.pdf}
+\includegraphics[width=0.47\textwidth,trim=5 0 0 0, clip]{/Users/adrian/projects/streams/plots/paper1/ps_distance.pdf}
 \caption{Phase space distance ($D_{\rm ps}$) for 10 stars integrated backwards in the correct potential (top) and a potential where $q_z$ is 25\% larger (bottom).
@@
 
-%As a measure of how many stars have recombined, we treat the minimum phase-space vectors (for each star over the individual, integrated orbits) as samples from an underlying six-dimensional distribution.
-
 For each star, $i$,
@@
 
-%[Still working on this bit...need to figure out best way to express
-%  what I'm trying to say.] Though this is not explicitly a
-%\emph{likelihood}, optimizing over this statistic is a very simple
-%approximation to performing a maximum-likelihood
-%optimization. Assuming the samples in $\bs{A}_{ik}$ are approximately
-%normal-distributed, the variance is closely related to the Shannon
-%entropy \citep{shannon1948}, $H$,
-%\begin{equation}
-%  H \propto \log \sigma .
-%\end{equation}
-%Minimizing the entropy is then equivalent to minimizing the loss of
-%\emph{information} for a given choice of parameters, $\bs{\theta}$,
-%given the ``data,'' $\bs{A}_{ik}$, and is thus related to
-%maximum-likelihood estimation \citep[see][for more
-%  detail]{mackay2002}.
-
 \subsection{Application to Simulated Data} \label{sec:results}
@@
 
-%\begin{figure}[h]
-%\begin{center}
-%\includegraphics[width=1.\textwidth]{/Users/adrian/projects/streams/plots/paper1/variance_projections.pdf}
-%\caption{ todo }\label{fig:variance_proj}
-%\end{center}
-%\end{figure}
-
 \begin{figure}[h]
@@
 point estimate for the potential parameters. To assess the uncertainty
-in the derived halo parameters, we bootstrap resample our 100 stars
+in the derived halo parameters, we sample 100 stars
 100 times and estimate the potential parameters with each
 resampling. Figure~\ref{fig:bootstrap} shows the recovered parameters
-for each bootstrapped sample and demonstrates the power of the
+for each sample and demonstrates the power of the
 combined Spitzer and Gaia data: a moderately sized sample of RR Lyrae
 alone places quite strong constraints on the shape and mass of the
-Galaxy's dark matter halo. [APW: explicitly state limits here, 1-sigma bounds]
+Galaxy's dark matter halo. From the covariance matrix derived from the 
+distribution of points in Figure~\ref{fig:bootstrap}, we find the mean 
+recovered parameters and one-sigma deviation to be 
+$q_1 = 1.36 \pm 0.02$, $q_z = 1.35 \pm 0.03$, $\phi = 96.6 \pm 0.6$ degrees,
+and $v_{\rm halo} = 123.0 \pm 1.5$ km/s.
 
@@
 \centering\includegraphics[width=0.9\textwidth,trim=0 0 0 0, clip]{/Users/adrian/projects/streams/plots/paper1/bootstrap.pdf}
-\caption{ Blue points show the ``best-fit'' parameters resulting from each bootstrap resample of 100 stars from the Sgr stream particle density shown in Figure~\ref{fig:lm10}.
+\caption{ Blue points show the ``best-fit'' parameters resulting from each resample of 100 stars from the Sgr stream particle density shown in Figure~\ref{fig:lm10}.
 Green (vertical and horizontal) lines show the true values of the parameters.
@@
 \end{figure*}
-
-% APW Note: The optimization, in a way, is getting a bunch of orbits (curves) to intersect another curve (the progenitor orbit) at some points (over time).
 
@@
 present prescription neglects second-order effects in orbit evolution (e.g.,
-dynamical friction) which may lead to biases in
+dynamical friction) and N-body interactions which may lead to biases in
 the estimated potential parameters.
-
-%With a justified probabilistic framework, we will also be able to provide ...
-given sufficient data, position and motion of the original host could be solved for as additional free parameters, but the effectiveness of this approach has yet to be investigated.
 
@@
 \cite{law10} N-body simulation of the destruction of the Sgr dwarf
-satellite, the method recovers the shape and orientation of the dark
-matter potential [APW: again, quote errors].
+satellite, \emph{Rewinder} recovers the shape and orientation of the dark
+matter potential.
 
@@
 \acknowledgments
-We thank ...
+We thank Barry Madore for providing the inspiration for this work. Thanks to 
+David Hogg for statistical advice and many useful discussions. Thanks also to
+Steve Majewski, David Law, and David Nidever.
 
 APW is supported by a National Science Foundation Graduate Research
-Fellowship under Grant No.\ 11-44155.
+Fellowship under Grant No.\ 11-44155. This work was supported in part by 
+the National Science Foundation under Grant No. PHYS-1066293 and the 
+hospitality of the Aspen Center for Physics.
 
-\bibliography{refs}
+%\bibliography{refs}
 \bibliographystyle{apj}
 
+\begin{thebibliography}{47}
+\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
+
+\bibitem[{{Bailin} {et~al.}(2005){Bailin}, {Kawata}, {Gibson}, {Steinmetz},
+  {Navarro}, {Brook}, {Gill}, {Ibata}, {Knebe}, {Lewis}, \&
+  {Okamoto}}]{bailin05}
+{Bailin}, J., {Kawata}, D., {Gibson}, B.~K., {et~al.} 2005, \apjl, 627, L17
+
+\bibitem[{{Bellazzini} {et~al.}(2006){Bellazzini}, {Newberg}, {Correnti},
+  {Ferraro}, \& {Monaco}}]{bellazzini06}
+{Bellazzini}, M., {Newberg}, H.~J., {Correnti}, M., {Ferraro}, F.~R., \&
+  {Monaco}, L. 2006, \aap, 457, L21
+
+\bibitem[{{Belokurov} {et~al.}(2006){Belokurov}, {Zucker}, {Evans}, {Gilmore},
+  {Vidrih}, {Bramich}, {Newberg}, {Wyse}, {Irwin}, {Fellhauer}, {Hewett},
+  {Walton}, {Wilkinson}, {Cole}, {Yanny}, {Rockosi}, {Beers}, {Bell},
+  {Brinkmann}, {Ivezi{\'c}}, \& {Lupton}}]{belokurov06}
+{Belokurov}, V., {Zucker}, D.~B., {Evans}, N.~W., {et~al.} 2006, \apjl, 642,
+  L137
+
+\bibitem[{{Belokurov} {et~al.}(2007){Belokurov}, {Evans}, {Irwin},
+  {Lynden-Bell}, {Yanny}, {Vidrih}, {Gilmore}, {Seabroke}, {Zucker},
+  {Wilkinson}, {Hewett}, {Bramich}, {Fellhauer}, {Newberg}, {Wyse}, {Beers},
+  {Bell}, {Barentine}, {Brinkmann}, {Cole}, {Pan}, \& {York}}]{belokurov07}
+{Belokurov}, V., {Evans}, N.~W., {Irwin}, M.~J., {et~al.} 2007, \apj, 658, 337
+
+\bibitem[{{Belokurov} {et~al.}(2013){Belokurov}, {Koposov}, {Evans},
+  {Pe{\~n}arrubia}, {Irwin}, {Smith}, {Lewis}, {Gieles}, {Wilkinson},
+  {Gilmore}, {Olszewski}, \& {Niederste-Ostholt}}]{belokurov13}
+{Belokurov}, V., {Koposov}, S.~E., {Evans}, N.~W., {et~al.} 2013, ArXiv
+  e-prints
+
+\bibitem[{{Benedict} {et~al.}(2011){Benedict}, {McArthur}, {Feast}, {Barnes},
+  {Harrison}, {Bean}, {Menzies}, {Chaboyer}, {Fossati}, {Nesvacil}, {Smith},
+  {Kolenberg}, {Laney}, {Kochukhov}, {Nelan}, {Shulyak}, {Taylor}, \&
+  {Freedman}}]{benedict11}
+{Benedict}, G.~F., {McArthur}, B.~E., {Feast}, M.~W., {et~al.} 2011, \aj, 142,
+  187
+
+\bibitem[{{Binney}(2008)}]{binney08}
+{Binney}, J. 2008, \mnras, 386, L47
+
+\bibitem[{{Chou} {et~al.}(2010){Chou}, {Cunha}, {Majewski}, {Smith},
+  {Patterson}, {Mart{\'{\i}}nez-Delgado}, \& {Geisler}}]{chou10}
+{Chou}, M.-Y., {Cunha}, K., {Majewski}, S.~R., {et~al.} 2010, \apj, 708, 1290
+
+\bibitem[{{Clementini} {et~al.}(2003){Clementini}, {Gratton}, {Bragaglia},
+  {Carretta}, {Di Fabrizio}, \& {Maio}}]{clementini03}
+{Clementini}, G., {Gratton}, R., {Bragaglia}, A., {et~al.} 2003, \aj, 125, 1309
+
+\bibitem[{{Deason} {et~al.}(2012{\natexlab{a}}){Deason}, {Belokurov}, {Evans},
+  \& {An}}]{deason12a}
+{Deason}, A.~J., {Belokurov}, V., {Evans}, N.~W., \& {An}, J.
+  2012{\natexlab{a}}, \mnras, 424, L44
+
+\bibitem[{{Deason} {et~al.}(2012{\natexlab{b}}){Deason}, {Belokurov}, {Evans},
+  {Koposov}, {Cooke}, {Pe{\~n}arrubia}, {Laporte}, {Fellhauer}, {Walker}, \&
+  {Olszewski}}]{deason12b}
+{Deason}, A.~J., {Belokurov}, V., {Evans}, N.~W., {et~al.} 2012{\natexlab{b}},
+  \mnras, 425, 2840
+
+\bibitem[{{Drake} {et~al.}(2013){Drake}, {Catelan}, {Djorgovski}, {Torrealba},
+  {Graham}, {Belokurov}, {Koposov}, {Mahabal}, {Prieto}, {Donalek}, {Williams},
+  {Larson}, {Christensen}, \& {Beshore}}]{drake13}
+{Drake}, A.~J., {Catelan}, M., {Djorgovski}, S.~G., {et~al.} 2013, \apj, 763,
+  32
+
+\bibitem[{{Eyre} \& {Binney}(2009)}]{eyre09a}
+{Eyre}, A., \& {Binney}, J. 2009, \mnras, 400, 548
+
+\bibitem[{{Fellhauer} {et~al.}(2006){Fellhauer}, {Belokurov}, {Evans},
+  {Wilkinson}, {Zucker}, {Gilmore}, {Irwin}, {Bramich}, {Vidrih}, {Wyse},
+  {Beers}, \& {Brinkmann}}]{fellhauer06}
+{Fellhauer}, M., {Belokurov}, V., {Evans}, N.~W., {et~al.} 2006, \apj, 651, 167
+
+\bibitem[{{Foreman-Mackey} {et~al.}(2013){Foreman-Mackey}, {Hogg}, {Lang}, \&
+  {Goodman}}]{foremanmackey2013}
+{Foreman-Mackey}, D., {Hogg}, D.~W., {Lang}, D., \& {Goodman}, J. 2013, \pasp,
+  125, 306
+
+\bibitem[{{Helmi} \& {White}(1999)}]{helmi99}
+{Helmi}, A., \& {White}, S.~D.~M. 1999, \mnras, 307, 495
+
+\bibitem[{{Ibata} {et~al.}(1994){Ibata}, {Gilmore}, \& {Irwin}}]{ibata94}
+{Ibata}, R.~A., {Gilmore}, G., \& {Irwin}, M.~J. 1994, \nat, 370, 194
+
+\bibitem[{{Ivezi{\'c}} {et~al.}(2008){Ivezi{\'c}}, {Sesar}, {Juri{\'c}},
+  {Bond}, {Dalcanton}, {Rockosi}, {Yanny}, {Newberg}, {Beers}, {Allende
+  Prieto}, {Wilhelm}, {Lee}, {Sivarani}, {Norris}, {Bailer-Jones}, {Re
+  Fiorentin}, {Schlegel}, {Uomoto}, {Lupton}, {Knapp}, {Gunn}, {Covey},
+  {Smith}, {Miknaitis}, {Doi}, {Tanaka}, {Fukugita}, {Kent}, {Finkbeiner},
+  {Munn}, {Pier}, {Quinn}, {Hawley}, {Anderson}, {Kiuchi}, {Chen}, {Bushong},
+  {Sohi}, {Haggard}, {Kimball}, {Barentine}, {Brewington}, {Harvanek},
+  {Kleinman}, {Krzesinski}, {Long}, {Nitta}, {Snedden}, {Lee}, {Harris},
+  {Brinkmann}, {Schneider}, \& {York}}]{ivezic08}
+{Ivezi{\'c}}, {\v Z}., {Sesar}, B., {Juri{\'c}}, M., {et~al.} 2008, \apj, 684,
+  287
+
+\bibitem[{{Jing} \& {Suto}(2002)}]{jing02}
+{Jing}, Y.~P., \& {Suto}, Y. 2002, \apj, 574, 538
+
+\bibitem[{{Johnston}(1998)}]{johnston98}
+{Johnston}, K.~V. 1998, \apj, 495, 297
+
+\bibitem[{{Johnston} {et~al.}(1999{\natexlab{a}}){Johnston}, {Majewski},
+  {Siegel}, {Reid}, \& {Kunkel}}]{johnston99b}
+{Johnston}, K.~V., {Majewski}, S.~R., {Siegel}, M.~H., {Reid}, I.~N., \&
+  {Kunkel}, W.~E. 1999{\natexlab{a}}, \aj, 118, 1719
+
+\bibitem[{{Johnston} {et~al.}(1999{\natexlab{b}}){Johnston}, {Zhao}, {Spergel},
+  \& {Hernquist}}]{johnston99a}
+{Johnston}, K.~V., {Zhao}, H., {Spergel}, D.~N., \& {Hernquist}, L.
+  1999{\natexlab{b}}, \apjl, 512, L109
+
+\bibitem[{{Kafle} {et~al.}(2012){Kafle}, {Sharma}, {Lewis}, \&
+  {Bland-Hawthorn}}]{kafle12}
+{Kafle}, P.~R., {Sharma}, S., {Lewis}, G.~F., \& {Bland-Hawthorn}, J. 2012,
+  \apj, 761, 98
+
+\bibitem[{{Keller} {et~al.}(2010){Keller}, {Yong}, \& {Da Costa}}]{keller10}
+{Keller}, S.~C., {Yong}, D., \& {Da Costa}, G.~S. 2010, \apj, 720, 940
+
+\bibitem[{{Koposov} {et~al.}(2010){Koposov}, {Rix}, \& {Hogg}}]{koposov10}
+{Koposov}, S.~E., {Rix}, H.-W., \& {Hogg}, D.~W. 2010, \apj, 712, 260
+
+\bibitem[{{Law} \& {Majewski}(2010)}]{law10}
+{Law}, D.~R., \& {Majewski}, S.~R. 2010, \apj, 714, 229
+
+\bibitem[{{Longmore} {et~al.}(1986){Longmore}, {Fernley}, \&
+  {Jameson}}]{longmore86}
+{Longmore}, A.~J., {Fernley}, J.~A., \& {Jameson}, R.~F. 1986, \mnras, 220, 279
+
+\bibitem[{{Madore} \& {Freedman}(2012)}]{madore12}
+{Madore}, B.~F., \& {Freedman}, W.~L. 2012, \apj, 744, 132
+
+\bibitem[{{Majewski} {et~al.}(2003){Majewski}, {Skrutskie}, {Weinberg}, \&
+  {Ostheimer}}]{majewski03}
+{Majewski}, S.~R., {Skrutskie}, M.~F., {Weinberg}, M.~D., \& {Ostheimer}, J.~C.
+  2003, \apj, 599, 1082
+
+\bibitem[{{Majewski} {et~al.}(2004){Majewski}, {Kunkel}, {Law}, {Patterson},
+  {Polak}, {Rocha-Pinto}, {Crane}, {Frinchaboy}, {Hummels}, {Johnston}, {Rhee},
+  {Skrutskie}, \& {Weinberg}}]{majewski04}
+{Majewski}, S.~R., {Kunkel}, W.~E., {Law}, D.~R., {et~al.} 2004, \aj, 128, 245
+
+\bibitem[{{Mart{\'{\i}}nez-Delgado} {et~al.}(2004){Mart{\'{\i}}nez-Delgado},
+  {G{\'o}mez-Flechoso}, {Aparicio}, \& {Carrera}}]{martinezdelgado04}
+{Mart{\'{\i}}nez-Delgado}, D., {G{\'o}mez-Flechoso}, M.~{\'A}., {Aparicio}, A.,
+  \& {Carrera}, R. 2004, \apj, 601, 242
+
+\bibitem[{{Navarro} {et~al.}(1996){Navarro}, {Frenk}, \& {White}}]{navarro96}
+{Navarro}, J.~F., {Frenk}, C.~S., \& {White}, S.~D.~M. 1996, \apj, 462, 563
+
+\bibitem[{{Newberg} {et~al.}(2002){Newberg}, {Yanny}, {Rockosi}, {Grebel},
+  {Rix}, {Brinkmann}, {Csabai}, {Hennessy}, {Hindsley}, {Ibata}, {Ivezi{\'c}},
+  {Lamb}, {Nash}, {Odenkirchen}, {Rave}, {Schneider}, {Smith}, {Stolte}, \&
+  {York}}]{newberg02}
+{Newberg}, H.~J., {Yanny}, B., {Rockosi}, C., {et~al.} 2002, \apj, 569, 245
+
+\bibitem[{{Odenkirchen} {et~al.}(2002){Odenkirchen}, {Grebel}, {Dehnen}, {Rix},
+  \& {Cudworth}}]{odenkirchen02}
+{Odenkirchen}, M., {Grebel}, E.~K., {Dehnen}, W., {Rix}, H.-W., \& {Cudworth},
+  K.~M. 2002, \aj, 124, 1497
+
+\bibitem[{{Pe{\~n}arrubia} {et~al.}(2012){Pe{\~n}arrubia}, {Koposov}, \&
+  {Walker}}]{penarrubia12}
+{Pe{\~n}arrubia}, J., {Koposov}, S.~E., \& {Walker}, M.~G. 2012, \apj, 760, 2
+
+\bibitem[{{Perryman} {et~al.}(2001){Perryman}, {de Boer}, {Gilmore}, {H{\o}g},
+  {Lattanzi}, {Lindegren}, {Luri}, {Mignard}, {Pace}, \& {de Zeeuw}}]{gaia01}
+{Perryman}, M.~A.~C., {de Boer}, K.~S., {Gilmore}, G., {et~al.} 2001, \aap,
+  369, 339
+
+\bibitem[{{Pontzen} \& {Governato}(2012)}]{pontzen12}
+{Pontzen}, A., \& {Governato}, F. 2012, \mnras, 421, 3464
+
+\bibitem[{{Rocha-Pinto} {et~al.}(2004){Rocha-Pinto}, {Majewski}, {Skrutskie},
+  {Crane}, \& {Patterson}}]{rochapinto04}
+{Rocha-Pinto}, H.~J., {Majewski}, S.~R., {Skrutskie}, M.~F., {Crane}, J.~D., \&
+  {Patterson}, R.~J. 2004, \apj, 615, 732
+
+\bibitem[{{Rubin} \& {Ford}(1970)}]{rubin70}
+{Rubin}, V.~C., \& {Ford}, Jr., W.~K. 1970, \apj, 159, 379
+
+\bibitem[{{Sanders} \& {Binney}(2013{\natexlab{a}})}]{sanders13a}
+{Sanders}, J.~L., \& {Binney}, J. 2013{\natexlab{a}}, \mnras, 433, 1813
+
+\bibitem[{{Sanders} \& {Binney}(2013{\natexlab{b}})}]{sanders13b}
+---. 2013{\natexlab{b}}, \mnras, 433, 1826
+
+\bibitem[{{Sesar} {et~al.}(2010){Sesar}, {Ivezi{\'c}}, {Grammer}, {Morgan},
+  {Becker}, {Juri{\'c}}, {De Lee}, {Annis}, {Beers}, {Fan}, {Lupton}, {Gunn},
+  {Knapp}, {Jiang}, {Jester}, {Johnston}, \& {Lampeitl}}]{sesar10}
+{Sesar}, B., {Ivezi{\'c}}, {\v Z}., {Grammer}, S.~H., {et~al.} 2010, \apj, 708,
+  717
+
+\bibitem[{{Shapley}(1918)}]{shapley18}
+{Shapley}, H. 1918, \apj, 48, 154
+
+\bibitem[{{Totten} \& {Irwin}(1998)}]{totten98}
+{Totten}, E.~J., \& {Irwin}, M.~J. 1998, \mnras, 294, 1
+
+\bibitem[{{Unwin} {et~al.}(2008){Unwin}, {Shao}, {Tanner}, {Allen}, {Beichman},
+  {Boboltz}, {Catanzarite}, {Chaboyer}, {Ciardi}, {Edberg}, {Fey}, {Fischer},
+  {Gelino}, {Gould}, {Grillmair}, {Henry}, {Johnston}, {Johnston}, {Jones},
+  {Kulkarni}, {Law}, {Majewski}, {Makarov}, {Marcy}, {Meier}, {Olling}, {Pan},
+  {Patterson}, {Pitesky}, {Quirrenbach}, {Shaklan}, {Shaya}, {Strigari},
+  {Tomsick}, {Wehrle}, \& {Worthey}}]{unwin08}
+{Unwin}, S.~C., {Shao}, M., {Tanner}, A.~M., {et~al.} 2008, \pasp, 120, 38
+
+\bibitem[{{Varghese} {et~al.}(2011){Varghese}, {Ibata}, \&
+  {Lewis}}]{varghese11}
+{Varghese}, A., {Ibata}, R., \& {Lewis}, G.~F. 2011, \mnras, 417, 198
+
+\bibitem[{{Vivas} {et~al.}(2005){Vivas}, {Zinn}, \& {Gallart}}]{vivas05}
+{Vivas}, A.~K., {Zinn}, R., \& {Gallart}, C. 2005, \aj, 129, 189
+
+\end{thebibliography}
+
 \end{document}

### 2013-08-07 text/paper1/paper.tex
chop one

@@
 The inclusion of baryons in the simulations tends to soften the triaxiality of the dark matter in the inner regions of the halo \citep[e.g., as the disk forms,][]{bailin05} and
-can potentially alter the radial profile through a combination of adiabatic contraction and energetic feedback \citep[see recent discussion by][]{pontzen12}.
+can potentially alter the radial profile through a combination of adiabatic contraction and energetic feedback \citep[e.g.][]{pontzen12}.
 Hence, measurements of the shape, orientation, radial profile, and extent of dark matter halos provides information about the formation of these vast structures, as well as the messy baryonic processes of gas inflow, star formation, and feedback that continue to shape them.
 
-The Milky Way is the best candidate for such a detailed study of a dark matter halo. 
-Our presence in the Galaxy means that we can resolve and use individual stars as tracers and hence build large samples that probe deep into the halo.
+The Milky Way is the best candidate for such a detailed study of a dark matter halo
+since we can resolve large samples of individual stars to use as tracers.
 To date, thousands of blue horizontal branch stars selected from the Sloan Digital Sky Survey (SDSS) have been used to probe the Milky Way mass out to tens of
@@
 
-The caveat to working with tracer populations is the assumption that the tracer stars represent a random sampling of orbits drawn from a smooth distribution function and are well-mixed in orbital phase.
-However, the same surveys that provide large samples of tracers have also revealed the existence of large-scale spatial inhomogeneities in the form of giant stellar streams and clouds \citep{newberg02,majewski03,belokurov06}, clearly demonstrating that a significant fraction of the stellar halo is neither randomly sampled from a smooth distribution function nor is fully phase-mixed.
+This approach assumes that the tracers represent a random sampling of orbits drawn from a smooth distribution function and are well-mixed in orbital phase.
+However, the same surveys that provide large samples have also revealed the existence of large-scale spatial inhomogeneities in the form of giant stellar streams \citep{newberg02,majewski03,belokurov06}, clearly demonstrating that a significant fraction of the stellar halo is neither randomly sampled from a smooth distribution function nor is fully phase-mixed.
 
-A complimentary method to measure the mass distribution is to instead take full advantage of the {\it non}-random nature of the stellar distribution in the Galactic halo.
-While we do not know the exact orbits of  stellar debris from satellite destruction 
-we do know something more about stars in a debris structure: the stars were once all part of the same object.
-Methods that utilize this knowledge on dynamically cold  tidal streams can require orders of magnitude fewer tracers than a randomly sampled population to achieve comparable accuracy on derived constraints.
-One approach to potential measurement is to simply fit orbits to observations of streams 
+A complimentary approach to measuring the mass distribution is to instead take full advantage of the {\it non}-random nature of the Galaxy's stellar distribution, i.e.
+utilizing the knowledge that stars in streams were once all part of the same object.
+Such approaches can require orders of magnitude fewer tracers than a randomly sampled population to achieve comparable accuracy.
+One method is to simply fit orbits to observations of streams 
 \citep[e.g.,][]{koposov10}.
-However, the assumption that debris traces a single orbit is actually incorrect \citep[see][for discussions of the orbit distribution in tidal debris]{johnston98,helmi99}, with
-systematic changes in orbital energy along debris streams that can lead to systematic biases in our measurements of the Galactic potential \citep{eyre09a,varghese11}.
+However, the assumption that debris traces a single orbit is actually incorrect \citep[see][]{johnston98,helmi99}, with
+systematic changes in orbital properties along debris streams that can lead to systematic biases in our measurements of the Galactic potential \citep{eyre09a,varghese11}.
 In particular, \citet{sanders13a} recently demonstrated that this bias is equally problematic for the very thinnest, coldest streams, whose observed properties may be indistinguishable from those of the parent orbit \citep[e.g.
@@
 
-One way to address these systematic biases is to run fully self-consistent N-body simulations of satellite destruction in a variety of potentials (which naturally generate a
-distribution of orbital properties in the debris and along tidal streams) with the aim of simultaneously constraining both the properties of the satellite and the Milky Way.
+One way to address these systematic biases is to run fully self-consistent N-body simulations of satellite destruction in a variety of potentials with the aim of simultaneously constraining both the properties of the satellite and the Milky Way.
 Many studies of the Sagittarius debris system (hereafter Sgr) --- whose tidal streams entirely encircle the Milky Way --- have adopted this approach 
-with the most recent work for the first time attempting to place constraints on the triaxiality and orientation of the Milky Way's matter distribution
+with the most recent work for the first time attempting to place constraints on the triaxiality and orientation of the dark matter halo
 \citep{law10}.
@@
 
-The promise of near-future data sets including full (or nearly full) phase-space information has also inspired discussions that move beyond fitting observables along a
-stream.
+The promise of near-future data sets including full phase-space information has also inspired other approaches.
 \citet{binney08} and \citet{penarrubia12} demonstrate conceptually that the distribution of energy and entropy (respectively) in debris will be minimized only for a correct
@@
 
-In this paper we re-examine and update a complimentary approach to using tidal debris as a potential measure \citep[originially proposed by][]{johnston99a}  in the context of current and near-future observational capabilities.
-We illustrate the power of this method by applying it to a simulation of the Sgr debris system.
+In this paper we re-examine and update a complimentary approach to using tidal debris as a potential measure \citep[originally proposed by][]{johnston99a}  in the context of current and near-future observational capabilities, and apply it to a simulation of the Sgr debris system.
 In Section \ref{sec:context} we outline the observational prospects and Sgr properties that motivated this re-examination.
@@
 The method presented in Section \ref{sec:method} takes advantage of
-three distinct developments in the observations of tidal streams: (i)
+three distinct developments: (i)
 the demonstration of a technique for deriving distances to individual
@@
 At longer wavelengths, RR Lyrae promise tighter constraints on
-distances.  The near-infrared period-luminosity (PL) relation for RR
-Lyrae stars was first mapped by \citet{longmore86} using $K$-band
-observations of stars in Galactic globular clusters 
+distances.  
 \citet{madore12} have recently shown, 
@@
 trigonometric parallaxes measured by Hubble \citep{benedict11},
-that the dispersion in the mid-IR PL relation at
+that the dispersion in the mid-IR PL relation 
+\citep[first mapped by][]{longmore86}
+at
 wavelengths measurable by NASA's {\it Spitzer} mission is $\sim$0.03~mag.
@@
 For comparison, distance measurements of Blue Horizontal Branch
-and Blue Straggler stars at typically
-achieve $\sim$10-15\% uncertainties, and then only if appropriate color measurements are available \citep[e.g.,][]{deason12b}.
+stars typically
+achieve $\sim$10-15\% uncertainties  \citep[if appropriate color measurements are available. e.g.,][]{deason12b}.
 
@@
 \label{sec:gaia}
-The Gaia satellite \citep{gaia01}, planned for launch in late 2013, is
+The Gaia satellite \citep{gaia01} is
 an astrometric mission which aims to measure the positions of billions
-of stars with 10-100~$\mu$as accuracies. With over 50 observations of
-the entire sky during the five-year lifetime of the mission, Gaia will
-revolutionize the field of Galactic Astronomy by providing sufficient
+of stars with 10-100~$\mu$as accuracies and sufficient
 proper motion accuracy to create full six-dimensional phase-space maps
@@
 and tangential velocity error estimates for RR Lyrae. The velocity
-uncertainty depends on the magnitude (and color) of the source, which depends on
+uncertainty depends on the magnitude (and color) of the source, which depends in turn on
 the metallicity: each line in Figure~\ref{fig:gaia_errors} is computed
@@
 Errors are a function of color and magnitude of the source, which depend on the metallicity.
-Each grey line  is computed by Monte Carlo sampling from the empirical metallicity distribution of
-the Galactic halo from \cite{ivezic08} (centered around [Fe/H] $\sim$ -1.5), and using eq.~14 from \citet{benedict11} to compute the absolute magnitude of the star.
-Note in top panel, parallax distance errors from \emph{Gaia} are larger than the line-of-sight size of both Sgr and Orphan (Orp), but photometric distance errors are comparable to the the Sgr scale (assuming 10\% errors, dotted line).
+Parallax distance errors from \emph{Gaia} are larger than the line-of-sight size of both Sgr and Orphan (Orp), but photometric distance errors are comparable to the the Sgr scale (assuming 10\% errors, dotted line).
 Bottom panel shows that the \emph{Gaia} tangential velocity errors (proper motion error $\times$ distance) are smaller than the internal velocity dispersion of both Sgr and Orp.
@@
 possible from $<$10~kpc to 60~kpc. This enormous increase in volume
-will greatly refine data on existing debris systems such as the
-Sagittarius stream.
+will greatly refine data on existing debris systems.
 
@@
 \citep{ibata94}. 
-
 Signatures of extensive stellar
-streams associated with Sgr have since been discovered in many studies and have been
+streams associated with Sgr have since  been
 mapped across the sky in carbon stars \citep{totten98}, M giants
 selected from the Two Micron All Sky Survey \citep{majewski03}, main
-sequence turnoff stars from the Sloan Digital Sky Survey
+sequence turnoff stars from SDSS
 \citep{belokurov06}, and RR Lyrae in the Catalina Sky Survey
-\citep{drake13}. Apogalactic debris has been traced out to $\sim 100$kpc 
-using a combination of blue horizontal branch,
-main-sequence turn-off, and red giant stars from the SDSS
-\citep{belokurov13}.
-The distance to the stream has been estimated from photometric surveys \citep[e.g.,][]{martinezdelgado04} and gradients in radial velocities and metallicity measured with follow-up spectroscopy\citep[e.g.,][]{majewski04, vivas05, bellazzini06, chou10, keller10}.
+\citep{drake13}.
 
-This multitude of Sgr stream data has inspired a rich set of interpretations and models of the debris
+Sgr stream data has inspired a rich set of  models 
 \citep[e.g.,][]{johnston99b, fellhauer06}.
@@
 Figure \ref{fig:lm10} shows particle positions
-(in Galactocentric coordinates) from the final time-step of the LM10 N-body
+from the final time-step of the LM10 N-body
 simulation of dwarf satellite disruption along the expected Sgr orbit
@@
 \caption{ Particle density (blue) of the first leading and trailing wraps from the final time-step of the \citet{law10} simulation of the Sgr stream.
-The density is estimated using Kernel Density Estimation (KDE) with a Gaussian kernel on $\sim$40,000 particle positions.
 Point markers (black) show positions of a representative sample of 100 stars drawn from this density distribution.
@@
 Moreover, this comparison enabled the first 
-reconstruction of the 3-dimensional mass distribution of a dark matter
+assessment of the 3-dimensional mass distribution of a dark matter
 halo through the constraints on the potential parameters $v_{\rm halo}$,$q_1$, $q_z$ and $\phi$.
@@
 stream, it is possible to use the individual stars as potential
-measures by exploting the fact that the stars must have come from the
-same progenitor. If the present-day phase-space position of the
-progenitor system is also known, this places an intuitive constraint
-on the potential of the Milky Way: if the orbits of the stars and the
+measures by exploiting the fact that the stars must have come from the
+same progenitor: if the orbits of the stars and the
 orbit of the progenitor are integrated \emph{backwards} in a a potential that accurately models the Milky Way
@@
 designed and tested for the proposed characteristics of the Space
-Interferometry Mission \citep{unwin08}, which was expected to be able
-to measure proper motions with $\mu$as/year accuracy for stars as
-faint as $V=20$. Distances to stream members would only be poorly
-known and were estimated in the method by using anticipated properties
-of tidal debris. Below we present an updated version of the algorithm:
+Interferometry Mission \citep{unwin08}. Below we present an updated version of the algorithm:
 the promise of 2\% distances to RR Lyrae stars (see Section
 \ref{sec:spitzer}) enables a direct measurement (rather than
-approximate estimate) of the position of a star within its debris
-structure; the test statistic that quantifies how well stars recombine
+approximate estimate, as previously assumed) of the position of a star within its debris
+structure; and the test statistic that quantifies how well stars recombine
 with the satellite has been rigorously redefined.
@@
 parameters, $\boldsymbol\theta$, the orbits of the stars and
-progenitor are integrated backwards for several Gigayears (determined
-from an estimate of the interaction time of the system). The stars and
-progenitor system are treated as test particles, so that the debris
-stars do not become physically re-bound to the satellite. Instead, at
+progenitor are integrated backwards for several Gigayears. At
 each timestep $t_j$, for each particle $i$, a set of normalized,
@@
 data.\footnote{\url{www.astro.virginia.edu/~srm4n/Sgr/data.html}} We
-then `observe' a sample of 100 stars from the first leading and
+then ``observe'' a sample of 100 stars from the first leading and
 trailing wraps of the stream; the radial velocity and distance errors are drawn 
@@
 only a rough estimate of the satellite mass $m_{\rm sat}$ combined
-with backwards integration of orbits. Many orbit fitting methods for measuring the Galactic potential assume
-that the stars in a given stream are on the same
-orbit. \cite{sanders13a} show that this assumption can lead to nearly
-100\% errors in the estimated potential parameters, even without
-realistic observational errors on the kinematic data. The
-\emph{Rewinder} is free of this assumption and instead \emph{relies}
+with backwards integration of orbits. 
+The \emph{Rewinder} is free of the assumption that stars in streams follow
+a single orbit and instead \emph{relies}
 on the fact that each star is on a different orbit, and is thus an
 independent potentiometer. There are also no assumptions made
-about, or N-body modeling of, the internal distribution of satellite
+about of the internal distribution of satellite
 stars. Thus, \emph{Rewinder} is applicable to any debris that is
 known to come from a single object and not restricted to the coldest
-tidal streams. In principle, the method could also be applied to the
-vast debris {\it clouds} that have been discovered over large areas of
-the sky \citep[e.g. the Triangulum Andromeda and Hercules-Aquila
+tidal streams. In principle, it could also be applied to the
+vast stellar debris {\it clouds} that have been discovered 
+\citep[e.g. the Triangulum Andromeda and Hercules-Aquila
   clouds]{rochapinto04,belokurov06}, or even stars that have only
 associations in orbital properties and do not form a coherent spatial
-structure \citep[such as the angular momentum groupings in local
-  giants found by][]{helmi99}. The method trivially extends to
+structure \citep[e.g.][]{helmi99}. The method trivially extends to
 combining constraints from multiple debris systems at once by simply
@@
 that anticipates combined data from the Spitzer and Gaia satellite
-missions that promise precise, full phase-space measurements of RR
+missions which promise precise, full phase-space measurements of RR
 Lyrae stars in the halo of our Galaxy. When applied to a sample of 100
-stars (with realistic observational errors) sampled from the
+stars (with realistic observational errors) drawn from the
 \cite{law10} N-body simulation of the destruction of the Sgr dwarf
-satellite, \emph{Rewinder} recovers the shape and orientation of the dark
-matter potential.
+satellite, \emph{Rewinder} recovers the depth, shape and orientation of the dark
+matter potential to within a few percent.
 

### 2013-08-08 text/paper1/paper.tex
apw chopping

@@
 
-Near-future data from ESA's Gaia mission will provide precise, full phase-space information for hundreds of millions of stars out to heliocentric distances of a few kiloparsecs.
-This ``horizon'' can be extended with more accurate distances, allowing for more precise tangential velocity measurements out to several tens of kiloparsecs if we limit studies to samples of RR Lyrae stars.
+Near-future data from ESA's Gaia mission will provide precise, full phase-space information for hundreds of millions of stars out to heliocentric distances of $\sim$ 10 kpc.
+This ``horizon'' for full phase-space measurements is imposed by the Gaia parallax errors degrading to worse than 10\%, and could be significantly extended by an accurate distance indicator.
 Recent work has demonstrated how Spitzer observations of RR Lyrae stars can be used to make distance estimates accurate to 2\%, effectively extending the Gaia, precise-data horizon by a factor of ten in distance and a factor of 1000 in volume.
-In this \emph{Letter} we develop and demonstrate a new method that uses small samples of stars associated with debris from satellite destruction to measure the Galactic potential.
-We then apply this method to simulated observations of stars from the end point of a simulation of satellite destruction and discuss the success of the algorithm in recovering the true potential used for the simulation.
-The results suggest that we will be able to probe the three-dimensional structure of the Galaxy's dark matter halo, constraining its shape, orientation, and depth as a function of radius.
-The opportunity to examine a dark matter halo in such great detail is truly unique and is enabled by the combination of Spitzer and Gaia with our intimate perspective on our own Galaxy.
+This \emph{Letter} presents one approach to exploit data of such accuracy to measure the Galactic potential using small samples of stars associated with debris from satellite destruction.
+The method is tested with synthetic observations of 100 stars from the end point of a simulation of satellite destruction: the shape, orientation, and depth of the potential used in the simulation are recovered to within a few percent.
+The success of this simple test with such a small sample in a single debris stream suggests that constraints from multiple streams could be combined to examine the Galaxy's  dark matter halo in even more  detail --- a truly unique opportunity that is enabled by the combination of Spitzer and Gaia with our intimate perspective on our own Galaxy.
 
@@
  The existence of vast halos of unseen {\it dark} matter surrounding each galaxy has long been proposed to explain the surprisingly large
-motions of the {\it baryonic} matter (i.e. stars, gas and dust) that we can see \citep[e.g.,][]{rubin70}.
+motions of the {\it baryonic} matter that we can see \citep[e.g.,][]{rubin70}.
 Dark-matter-only simulations of structure formation lead us to expect that these dark matter halos should have density distributions that are described by a universal radial profile \citep{navarro96} with a variety of triaxial shapes \citep{jing02}.
 The inclusion of baryons in the simulations tends to soften the triaxiality of the dark matter in the inner regions of the halo \citep[e.g., as the disk forms,][]{bailin05} and
-can potentially alter the radial profile through a combination of adiabatic contraction and energetic feedback \citep[e.g.][]{pontzen12}.
-Hence, measurements of the shape, orientation, radial profile, and extent of dark matter halos provides information about the formation of these vast structures, as well as the messy baryonic processes of gas inflow, star formation, and feedback that continue to shape them.
+can alter the radial profile through a combination of adiabatic contraction and energetic feedback \citep[e.g.][]{pontzen12}.
+Hence, measurements of the shape, orientation, radial profile, and extent of dark matter halos provides information about the formation of these vast structures, as well as the messy baryonic processes that continue to shape them.
 
 The Milky Way is the best candidate for such a detailed study of a dark matter halo
-since we can resolve large samples of individual stars to use as tracers.
+since we can resolve large samples of stellar tracers.
 To date, thousands of blue horizontal branch stars selected from the Sloan Digital Sky Survey (SDSS) have been used to probe the Milky Way mass out to tens of
@@
 This approach assumes that the tracers represent a random sampling of orbits drawn from a smooth distribution function and are well-mixed in orbital phase.
-However, the same surveys that provide large samples have also revealed the existence of large-scale spatial inhomogeneities in the form of giant stellar streams \citep{newberg02,majewski03,belokurov06}, clearly demonstrating that a significant fraction of the stellar halo is neither randomly sampled from a smooth distribution function nor is fully phase-mixed.
+However, the same surveys that provide large samples have also revealed the existence of large-scale spatial inhomogeneities in the form of giant stellar streams \citep{newberg02,majewski03,belokurov06}, demonstrating that a significant fraction of the stellar halo is neither randomly sampled nor is fully phase-mixed.
 
-A complimentary approach to measuring the mass distribution is to instead take full advantage of the {\it non}-random nature of the Galaxy's stellar distribution, i.e.
-utilizing the knowledge that stars in streams were once all part of the same object.
+A complimentary approach to measuring the mass distribution is to instead take full advantage of the {\it non}-random nature of the Galaxy's stellar distribution and utilize the knowledge that stars in streams were once all part of the same object.
 Such approaches can require orders of magnitude fewer tracers than a randomly sampled population to achieve comparable accuracy.
@@
 
-One way to address these systematic biases is to run fully self-consistent N-body simulations of satellite destruction in a variety of potentials with the aim of simultaneously constraining both the properties of the satellite and the Milky Way.
+One way to address these biases is to run self-consistent N-body simulations of satellite destruction in a variety of potentials with the aim of simultaneously constraining both the properties of the satellite and the Milky Way.
 Many studies of the Sagittarius debris system (hereafter Sgr) --- whose tidal streams entirely encircle the Milky Way --- have adopted this approach 
@@
 The promise of near-future data sets including full phase-space information has also inspired other approaches.
-\citet{binney08} and \citet{penarrubia12} demonstrate conceptually that the distribution of energy and entropy (respectively) in debris will be minimized only for a correct
+\citet{binney08} and \citet{penarrubia12} demonstrate conceptually that the distribution of energy and entropy in debris, respectively, will be minimized only for a correct
 assumption of the form of the Galactic potential.
@@
 
-\subsection{{\it Spitzer} and 2\% distance errors for RR Lyrae in the halo}
+\subsection{{\it Spitzer} and 2\% distance errors to RR Lyrae in the halo}
 \label{sec:spitzer}
@@
 sight.  Moreover, systematic differences between instruments make it
-difficult to tie observations across the sky to a common scale. While
-RR Lyrae are among the best distance indicators available to us today,
-the absolute uncertainty in individual estimates are typically
-$\gtrsim$10\%.
+difficult to tie observations across the sky to a common scale.
 
@@
 trigonometric parallaxes measured by Hubble \citep{benedict11},
-that the dispersion in the mid-IR PL relation 
+that the dispersion in the mid-IR Period-Luminosity relation 
 \citep[first mapped by][]{longmore86}
@@
 stars typically
-achieve $\sim$10-15\% uncertainties  \citep[if appropriate color measurements are available. e.g.,][]{deason12b}.
+achieve $\sim$10-15\% uncertainties  \citep[if appropriate color measurements are available, e.g.,][]{deason12b}.
+
+\begin{figure}[h!]
+\begin{center}
+\includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/gaia.pdf}
+\caption{Expected \emph{Gaia} distance and tangential velocity errors as a function of heliocentric distance for RR Lyrae stars.
+Errors are a function of color and magnitude of the source, and hence the metallicity.
+Parallax distance errors from \emph{Gaia} are larger than the line-of-sight size of both Sgr and Orphan (Orp), but photometric distance errors are comparable to the the Sgr scale (assuming 10\% errors, dotted line).
+Bottom panel shows that the \emph{Gaia} tangential velocity errors are smaller than the internal velocity dispersion of both Sgr and Orp.\footnote{\url{http://www.rssd.esa.int/index.php?page=Science_Performance&project=GAIA}} }\label{fig:gaia_errors}
+\end{center}
+\end{figure}
 
@@
 an astrometric mission which aims to measure the positions of billions
-of stars with 10-100~$\mu$as accuracies and sufficient
-proper motion accuracy to create full six-dimensional phase-space maps
-of the Galaxy with $<$10\% distance errors for heliocentric distances of
+of stars with 10-100~$\mu$as accuracies. Combined with expected 
+proper motion accuracies, this will enable full six-dimensional phase-space 
+maps of the Galaxy with $<$10\% distance errors to heliocentric distances of
 up to $\sim$6~kpc for RR Lyrae stars.
@@
 test and calibrate the Spitzer PL relation described above. Beyond the
-2~kpc threshold, the mid-IR PL relation for RR Lyrae will provide more
-accurate distance measurements.
+2~kpc threshold, the mid-IR PL relation for RR Lyrae will provide better 
+distance measurements.
 
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/gaia.pdf}
-\caption{Expected \emph{Gaia} distance and tangential velocity errors as a function of heliocentric distance for RR Lyrae stars.
-Errors are a function of color and magnitude of the source, which depend on the metallicity.
-Parallax distance errors from \emph{Gaia} are larger than the line-of-sight size of both Sgr and Orphan (Orp), but photometric distance errors are comparable to the the Sgr scale (assuming 10\% errors, dotted line).
-Bottom panel shows that the \emph{Gaia} tangential velocity errors (proper motion error $\times$ distance) are smaller than the internal velocity dispersion of both Sgr and Orp.
-}\label{fig:gaia_errors}
-\end{center}
-\end{figure}
-
-The anticipated proper motion errors for RR Lyrae correspond to
-tangential velocity errors of order $\sim$1~km/s at 30~kpc, and of
-order $\sim$10~km/s out to several tens of
-kiloparsecs.\footnote{\url{http://www.rssd.esa.int/index.php?page=Science_Performance&project=GAIA}}
-The combination of Spitzer and Gaia data will extend the ``horizon''
-of where full six-dimensional phase-space maps of the Galaxy are
-possible from $<$10~kpc to 60~kpc. This enormous increase in volume
-will greatly refine data on existing debris systems.
+The combination of Spitzer and Gaia data will 
+extend the ``horizon'' of where precise, six-dimensional phase-space 
+maps of the Galaxy are possible from $<$10~kpc to 60~kpc. This enormous 
+increase in volume will greatly refine data on debris systems in the halo.
 
@@
 Moreover, this comparison enabled the first 
-assessment of the 3-dimensional mass distribution of a dark matter
-halo through the constraints on the potential parameters $v_{\rm halo}$,$q_1$, $q_z$ and $\phi$.
+assessment of the three-dimensional mass distribution of a dark matter
+halo through the constraints on the potential parameters $v_{\rm halo}$, $q_1$, $q_z$, and $\phi$.
 
@@
 \label{sec:method}
-With access to full phase-space information for stars in a tidal
+With access to 6D information for stars in a tidal
 stream, it is possible to use the individual stars as potential
 measures by exploiting the fact that the stars must have come from the
-same progenitor: if the orbits of the stars and the
-orbit of the progenitor are integrated \emph{backwards} in a a potential that accurately models the Milky Way
+same progenitor: if the orbits of the stars and progenitor are integrated 
+\emph{backwards} in a a potential that accurately models the Milky Way
 (imagine watching satellite destruction in ``rewind"), the stars
@@
 the orbits of the stars will diverge from that of the progenitor and
-thus will not be recaptured by the satellite system.
+thus will not be recaptured by the satellite system (Figure~\ref{fig:ps_distance}).
 
-%(Figure~\ref{fig:ps_distance}).
-
-This approach was originally proposed by \citet{johnston99a} and was
-designed and tested for the proposed characteristics of the Space
+This approach was originally proposed by \citet{johnston99a} and was tested on the proposed characteristics of the Space
 Interferometry Mission \citep{unwin08}. Below we present an updated version of the algorithm:
@@
 then ``observe'' a sample of 100 stars from the first leading and
-trailing wraps of the stream; the radial velocity and distance errors are drawn 
+trailing wraps of the stream. The radial velocity and distance errors are drawn 
 from Gaussians ($\varepsilon_{\rm RV} \sim \mathcal{N}(\mu=0,\sigma=10~{\rm
@@
 
-The variance of the six-dimensional minimum phase-space distance distribution increases in the wrong potential. Hence, the generalized variance defines a convex hypersurface over
+The generalized variance defines a convex hypersurface over
 which we optimize four of the six logarithmic potential parameters:
@@
   expensive to compute. The stochasticity and easy parallelization of
-  the algorithm mean it outperforms other optimizers on this problem.}
-We use the median value of the (burnt-in) sample distribution as a
+  the algorithm outperforms other optimizers on this problem.}
+We use the median value of the converged sample distribution as a
 point estimate for the potential parameters. To assess the uncertainty
@@
 resampling. Figure~\ref{fig:bootstrap} shows the recovered parameters
-for each sample and demonstrates the power of the
-combined Spitzer and Gaia data: a moderately sized sample of RR Lyrae
-alone places quite strong constraints on the shape and mass of the
+for each sample and demonstrates the power of this method:
+a moderately sized sample of RR Lyrae
+alone places strong constraints on the shape and mass of the
 Galaxy's dark matter halo. From the covariance matrix derived from the 
 distribution of points in Figure~\ref{fig:bootstrap}, we find the mean 
-recovered parameters and one-sigma deviation to be 
+recovered parameters and one-sigma deviations to be 
 $q_1 = 1.36 \pm 0.02$, $q_z = 1.35 \pm 0.03$, $\phi = 96.6 \pm 0.6$ degrees,
@@
 with backwards integration of orbits. 
-The \emph{Rewinder} is free of the assumption that stars in streams follow
+\emph{Rewinder} does not assume that stream stars follow
 a single orbit and instead \emph{relies}
-on the fact that each star is on a different orbit, and is thus an
-independent potentiometer. There are also no assumptions made
+on the fact that each star is on a different orbit. There are also no assumptions made
 about of the internal distribution of satellite
@@
 the minute differences in orbital properties between the debris and
-satellite to be detected. Second, the current version of the algorithm
+satellite. Second, the current version of the algorithm
 relies on accurate knowledge of the current position and velocity of
-the parent satellite, which may not be available in some cases
+the parent satellite, which may not be available 
 \citep[for example, the Orphan Stream;][]{belokurov07}. Lastly, the
-present prescription neglects second-order effects in orbit evolution (e.g.,
+present prescription neglects orbital evolution (e.g.,
 dynamical friction) and N-body interactions which may lead to biases in
@@
 \cite{law10} N-body simulation of the destruction of the Sgr dwarf
-satellite, \emph{Rewinder} recovers the depth, shape and orientation of the dark
+satellite, \emph{Rewinder} recovers the depth, shape, and orientation of the dark
 matter potential to within a few percent.
@@
 Our results also motivate an observational campaign with \emph{Spitzer} to survey RR Lyrae stars in debris structures around the Milky Way to get precise distances to combine with near-future Gaia velocity data.
-If just 100 stars in a single stellar stream allow us to measure the depth, shape, and orientation of the Milky Way potential, larger samples in multiple structures offer the prospect of assessing these quantities as a function of Galactocentric radius.
+If just 100 stars in a single stellar stream allow us to measure the depth, shape, and orientation of the Milky Way potential, larger samples in multiple structures (e.g., the Orphan Stream, already mapped in RR Lyrae by \citet{sesar13}) offer the prospect of assessing these quantities as a function of Galactocentric radius.
 Tracing the mass in a dark matter halo with this level of detail is impossible for any other galaxy in the Universe.
@@
 
-\begin{thebibliography}{47}
+\begin{thebibliography}{42}
 \expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
@@
 {Bailin}, J., {Kawata}, D., {Gibson}, B.~K., {et~al.} 2005, \apjl, 627, L17
-
-\bibitem[{{Bellazzini} {et~al.}(2006){Bellazzini}, {Newberg}, {Correnti},
-  {Ferraro}, \& {Monaco}}]{bellazzini06}
-{Bellazzini}, M., {Newberg}, H.~J., {Correnti}, M., {Ferraro}, F.~R., \&
-  {Monaco}, L. 2006, \aap, 457, L21
 
@@
 
-\bibitem[{{Belokurov} {et~al.}(2013){Belokurov}, {Koposov}, {Evans},
-  {Pe{\~n}arrubia}, {Irwin}, {Smith}, {Lewis}, {Gieles}, {Wilkinson},
-  {Gilmore}, {Olszewski}, \& {Niederste-Ostholt}}]{belokurov13}
-{Belokurov}, V., {Koposov}, S.~E., {Evans}, N.~W., {et~al.} 2013, ArXiv
-  e-prints
-
 \bibitem[{{Benedict} {et~al.}(2011){Benedict}, {McArthur}, {Feast}, {Barnes},
@@
 {Binney}, J. 2008, \mnras, 386, L47
-
-\bibitem[{{Chou} {et~al.}(2010){Chou}, {Cunha}, {Majewski}, {Smith},
-  {Patterson}, {Mart{\'{\i}}nez-Delgado}, \& {Geisler}}]{chou10}
-{Chou}, M.-Y., {Cunha}, K., {Majewski}, S.~R., {et~al.} 2010, \apj, 708, 1290
 
@@
 
-\bibitem[{{Keller} {et~al.}(2010){Keller}, {Yong}, \& {Da Costa}}]{keller10}
-{Keller}, S.~C., {Yong}, D., \& {Da Costa}, G.~S. 2010, \apj, 720, 940
-
 \bibitem[{{Koposov} {et~al.}(2010){Koposov}, {Rix}, \& {Hogg}}]{koposov10}
@@
 {Majewski}, S.~R., {Kunkel}, W.~E., {Law}, D.~R., {et~al.} 2004, \aj, 128, 245
-
-\bibitem[{{Mart{\'{\i}}nez-Delgado} {et~al.}(2004){Mart{\'{\i}}nez-Delgado},
-  {G{\'o}mez-Flechoso}, {Aparicio}, \& {Carrera}}]{martinezdelgado04}
-{Mart{\'{\i}}nez-Delgado}, D., {G{\'o}mez-Flechoso}, M.~{\'A}., {Aparicio}, A.,
-  \& {Carrera}, R. 2004, \apj, 601, 242
 
@@
 
+\bibitem[{{Sesar} {et~al.}(2013){Sesar}, {Grillmair}, {Cohen}, {Bellm},
+  {Bhalerao}, {Levitan}, {Laher}, {Ofek}, {Surace}, {Tang}, {Waszczak},
+  {Kulkarni}, \& {Prince}}]{sesar13}
+{Sesar}, B., {Grillmair}, C.~J., {Cohen}, J.~G., {et~al.} 2013, ArXiv e-prints
+
 \bibitem[{{Shapley}(1918)}]{shapley18}
@@
 
-\bibitem[{{Vivas} {et~al.}(2005){Vivas}, {Zinn}, \& {Gallart}}]{vivas05}
-{Vivas}, A.~K., {Zinn}, R., \& {Gallart}, C. 2005, \aj, 129, 189
-
 \end{thebibliography}

### 2013-08-08 text/paper1/paper.tex

@@
 
-Near-future data from ESA's Gaia mission will provide precise, full phase-space information for hundreds of millions of stars out to heliocentric distances of $\sim$ 10 kpc.
+Near-future data from ESA's Gaia mission will provide precise, full phase-space information for hundreds of millions of stars out to heliocentric distances of $\sim$10 kpc.
 This ``horizon'' for full phase-space measurements is imposed by the Gaia parallax errors degrading to worse than 10\%, and could be significantly extended by an accurate distance indicator.
@@
 since we can resolve large samples of stellar tracers.
-To date, thousands of blue horizontal branch stars selected from the Sloan Digital Sky Survey (SDSS) have been used to probe the Milky Way mass out to tens of
+Thousands of blue horizontal branch stars selected from the Sloan Digital Sky Survey (SDSS) have been used to probe the Milky Way mass out to tens of
 kpc \citep[SDSS, see][]{deason12a,kafle12}, and estimates with combined tracers extend to 150kpc \citep{deason12b}.
@@
 
-A complimentary approach to measuring the mass distribution is to instead take full advantage of the {\it non}-random nature of the Galaxy's stellar distribution and utilize the knowledge that stars in streams were once all part of the same object.
+A complimentary approach to measuring the mass distribution is to instead take advantage of the {\it non}-random nature of the Galaxy's stellar distribution and utilize the knowledge that stars in streams were once all part of the same object.
 Such approaches can require orders of magnitude fewer tracers than a randomly sampled population to achieve comparable accuracy.
@@
 \citep[e.g.,][]{koposov10}.
-However, the assumption that debris traces a single orbit is actually incorrect \citep[see][]{johnston98,helmi99}, with
-systematic changes in orbital properties along debris streams that can lead to systematic biases in our measurements of the Galactic potential \citep{eyre09a,varghese11}.
+However, the assumption that debris traces a single orbit is actually incorrect \citep[see][]{johnston98,helmi99}.
+Changes in orbital properties along debris streams that can lead to systematic biases in measurements of the Galactic potential \citep{eyre09a,varghese11}.
 In particular, \citet{sanders13a} recently demonstrated that this bias is equally problematic for the very thinnest, coldest streams, whose observed properties may be indistinguishable from those of the parent orbit \citep[e.g.
@@
 
-Other approaches have been proposed that avoid the cost of full N-body simulations yet recognize the limitations of single orbit integration.
-For example, using the results of N-body simulations as a guide, simple corrections around a single orbit can be adopted to make model predictions that closely
-mimic expected debris behavior \citep[e.g.][]{johnston99b,varghese11}.
-
 The promise of near-future data sets including full phase-space information has also inspired other approaches.
-\citet{binney08} and \citet{penarrubia12} demonstrate conceptually that the distribution of energy and entropy in debris, respectively, will be minimized only for a correct
+\citet{binney08} and \citet{penarrubia12} demonstrate that the distribution of energy and entropy in debris, respectively, will be minimized only for a correct
 assumption of the form of the Galactic potential.
@@
 
-In this paper we re-examine and update a complimentary approach to using tidal debris as a potential measure \citep[originally proposed by][]{johnston99a}  in the context of current and near-future observational capabilities, and apply it to a simulation of the Sgr debris system.
+In this \emph{Letter} we re-examine and update a complimentary approach to using tidal debris as a potential measure \citep[originally proposed by][]{johnston99a}  in the context of current and near-future observational capabilities, and apply it to a simulation of the Sgr debris system.
 In Section \ref{sec:context} we outline the observational prospects and Sgr properties that motivated this re-examination.
@@
 trigonometric parallaxes measured by Hubble \citep{benedict11},
-that the dispersion in the mid-IR Period-Luminosity relation 
+that the dispersion in the mid-IR Period-Luminosity (PL) relation 
 \citep[first mapped by][]{longmore86}
@@
 This implies that it is
-possible to use {\it Spitzer} to determine distances that are good to better
-than $2\%$ for individual RR Lyrae stars out to $\sim$60~kpc (the limit of
-Spitzer's ability to detect and measure RR Lyrae variables for modest
-integration times). 
+possible to use {\it Spitzer} to determine distances that are good to $2\%$ for 
+individual RR Lyrae stars out to $\sim$60~kpc ({\it Spitzer}'s limit for detecting and measuring RR Lyrae).
 For comparison, distance measurements of Blue Horizontal Branch
@@
 \includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/gaia.pdf}
+
 \caption{Expected \emph{Gaia} distance and tangential velocity errors as a function of heliocentric distance for RR Lyrae stars.
-Errors are a function of color and magnitude of the source, and hence the metallicity.
+Errors are a function of color and magnitude of the source, and hence the metallicity: each line is computed
+by Monte Carlo sampling from the empirical metallicity distribution of
+the Galactic halo from \cite{ivezic08}.
 Parallax distance errors from \emph{Gaia} are larger than the line-of-sight size of both Sgr and Orphan (Orp), but photometric distance errors are comparable to the the Sgr scale (assuming 10\% errors, dotted line).
@@
 proper motion accuracies, this will enable full six-dimensional phase-space 
-maps of the Galaxy with $<$10\% distance errors to heliocentric distances of
+maps of the Galaxy with $<$10\% distance errors for heliocentric distances of
 up to $\sim$6~kpc for RR Lyrae stars.
@@
 Figure~\ref{fig:gaia_errors} shows the Gaia end-of-mission distance
-and tangential velocity error estimates for RR Lyrae. The velocity
-uncertainty depends on the magnitude (and color) of the source, which depends in turn on
-the metallicity: each line in Figure~\ref{fig:gaia_errors} is computed
-by Monte Carlo sampling from the empirical metallicity distribution of
-the Galactic halo from \cite{ivezic08} (centered around [Fe/H] $\sim$
--1.5). Within 2~kpc, Gaia will measure distances to these stars with
+and tangential velocity error estimates for RR Lyrae. Within 2~kpc, Gaia will measure distances to these stars with
 better than 2\% accuracy --- RR Lyrae in this volume can be used to
@@
 mapped across the sky in carbon stars \citep{totten98}, M giants
-selected from the Two Micron All Sky Survey \citep{majewski03}, main
+selected from 2MASS \citep{majewski03}, main
 sequence turnoff stars from SDSS
@@
 original distribution in populations. 
-Moreover, this comparison enabled the first 
+This comparison enabled the first 
 assessment of the three-dimensional mass distribution of a dark matter
@@
 
-Combined Spitzer and Gaia measurements of distances and proper motions
-of RR Lyraes in the Sgr debris will open up new avenues of
+Combined {\it Spitzer} and {\it Gaia} measurements of distances and proper motions
+of RR Lyrae in the Sgr debris will open up new avenues of
 interpretation. A 2\% distance error at $\sim$30~kpc
 ($\epsilon_{d}=0.6~{\rm kpc}$) is smaller than the predicted thickness
-of the stream, ranging from $\sim$2-$10$~kpc. When combined with Gaia proper motion errors,
+of the stream ($\sim$2-$10$~kpc). When combined with Gaia proper motion errors,
 tangential velocity measurement accuracies ($\epsilon_{v}\sim2$~km/s
 at 30~kpc; see Figure~\ref{fig:gaia_errors}) are less than the intrinsic dispersion for much of the stream
-\citep[$\sigma_v\sim10$~km/s;][]{majewski04} and the measured phase-space position of each individual RR
-Lyrae becomes a powerful probe of the potential. The next section
+\citep[$\sigma_v\sim10$~km/s;][]{majewski04}. The next section
 outlines a new method to take advantage of this information.
@@
 With access to 6D information for stars in a tidal
-stream, it is possible to use the individual stars as potential
-measures by exploiting the fact that the stars must have come from the
+stream, each star becomes a powerful potential
+measure by exploiting the fact that the stars must have come from the
 same progenitor: if the orbits of the stars and progenitor are integrated 
@@
 
+\begin{figure*}[ht!]
+\centering\includegraphics[width=0.9\textwidth,trim=0 0 0 0, clip]{/Users/adrian/projects/streams/plots/paper1/bootstrap.pdf}
+\caption{ Blue points show the ``best-fit'' parameters resulting from each resample of 100 stars from the Sgr stream particle density shown in Figure~\ref{fig:lm10}.
+Green (vertical and horizontal) lines show the true values of the parameters.
+Grey ellipses show one- and two-sigma margins, assuming the points are normally distributed.
+}\label{fig:bootstrap}
+\end{figure*}
+
 In anticipation of extending the above method to include a true
@@
 and $v_{\rm halo} = 123.0 \pm 1.5$ km/s.
-
-%\begin{figure}[h]
-%\begin{center}
-%\includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/bootstrap.pdf}
-%\caption{ Blue points show the ``best-fit'' parameters resulting from each bootstrap resample of 100 stars from the Sgr stream particle density shown in Figure~\ref{fig:lm10}.
-Green (vertical and horizontal) lines show the true values of the parameters.
-Grey ellipses show one- and two-sigma margins, assuming the points are normally distributed.
-}\label{fig:bootstrap}
-%\end{center}
-%\end{figure}
-
-\begin{figure*}[!t]
-\centering\includegraphics[width=0.9\textwidth,trim=0 0 0 0, clip]{/Users/adrian/projects/streams/plots/paper1/bootstrap.pdf}
-\caption{ Blue points show the ``best-fit'' parameters resulting from each resample of 100 stars from the Sgr stream particle density shown in Figure~\ref{fig:lm10}.
-Green (vertical and horizontal) lines show the true values of the parameters.
-Grey ellipses show one- and two-sigma margins, assuming the points are normally distributed.
-}\label{fig:bootstrap}
-\end{figure*}
 

### 2013-08-09 text/paper1/paper.tex
with Kathryn's final suggestions

@@
 
-\author{Adrian~M.~Price-Whelan\altaffilmark{1} Kathryn V. Johnston\altaffilmark{1}}
+\author{Adrian~M.~Price-Whelan\altaffilmark{1,2} Kathryn V. Johnston\altaffilmark{1}}
 
@@
 %\altaffiltext{2}{NSF Graduate Fellow}
+\altaffiltext{2}{\email{adrn@astro.columia.edu}}
 
@@
 
-Near-future data from ESA's Gaia mission will provide precise, full phase-space information for hundreds of millions of stars out to heliocentric distances of $\sim$10 kpc.
-This ``horizon'' for full phase-space measurements is imposed by the Gaia parallax errors degrading to worse than 10\%, and could be significantly extended by an accurate distance indicator.
-Recent work has demonstrated how Spitzer observations of RR Lyrae stars can be used to make distance estimates accurate to 2\%, effectively extending the Gaia, precise-data horizon by a factor of ten in distance and a factor of 1000 in volume.
+Near-future data from ESA's {\it Gaia} mission will provide precise, full phase-space information for hundreds of millions of stars out to heliocentric distances of $\sim$10 kpc.
+This ``horizon'' for full phase-space measurements is imposed by the {\it Gaia} parallax errors degrading to worse than 10\%, and could be significantly extended by an accurate distance indicator.
+Recent work has demonstrated how {\it Spitzer} observations of RR Lyrae stars can be used to make distance estimates accurate to 2\%, effectively extending the {\it Gaia}, precise-data horizon by a factor of ten in distance and a factor of 1000 in volume.
 This \emph{Letter} presents one approach to exploit data of such accuracy to measure the Galactic potential using small samples of stars associated with debris from satellite destruction.
 The method is tested with synthetic observations of 100 stars from the end point of a simulation of satellite destruction: the shape, orientation, and depth of the potential used in the simulation are recovered to within a few percent.
-The success of this simple test with such a small sample in a single debris stream suggests that constraints from multiple streams could be combined to examine the Galaxy's  dark matter halo in even more  detail --- a truly unique opportunity that is enabled by the combination of Spitzer and Gaia with our intimate perspective on our own Galaxy.
+The success of this simple test with such a small sample in a single debris stream suggests that constraints from multiple streams could be combined to examine the Galaxy's  dark matter halo in even more  detail --- a truly unique opportunity that is enabled by the combination of {\it Spitzer} and {\it Gaia} with our intimate perspective on our own Galaxy.
 
@@
 \citep[e.g.,][]{koposov10}.
-However, the assumption that debris traces a single orbit is actually incorrect \citep[see][]{johnston98,helmi99}.
-Changes in orbital properties along debris streams that can lead to systematic biases in measurements of the Galactic potential \citep{eyre09a,varghese11}.
+However, the assumption that debris traces a single orbit is actually incorrect \citep[see][]{johnston98,helmi99} and
+changes in orbital properties along debris streams can lead to systematic biases in measurements of the Galactic potential \citep{eyre09a,varghese11}.
 In particular, \citet{sanders13a} recently demonstrated that this bias is equally problematic for the very thinnest, coldest streams, whose observed properties may be indistinguishable from those of the parent orbit \citep[e.g.
@@
 
-\caption{Expected \emph{Gaia} distance and tangential velocity errors as a function of heliocentric distance for RR Lyrae stars.
+\caption{Expected {\it Gaia} distance and tangential velocity errors as a function of heliocentric distance for RR Lyrae stars.
 Errors are a function of color and magnitude of the source, and hence the metallicity: each line is computed
@@
 the Galactic halo from \cite{ivezic08}.
-Parallax distance errors from \emph{Gaia} are larger than the line-of-sight size of both Sgr and Orphan (Orp), but photometric distance errors are comparable to the the Sgr scale (assuming 10\% errors, dotted line).
-Bottom panel shows that the \emph{Gaia} tangential velocity errors are smaller than the internal velocity dispersion of both Sgr and Orp.\footnote{\url{http://www.rssd.esa.int/index.php?page=Science_Performance&project=GAIA}} }\label{fig:gaia_errors}
+Parallax distance errors from {\it Gaia} are larger than the line-of-sight size of both Sgr and Orphan (Orp), but photometric distance errors are comparable to the the Sgr scale (assuming 10\% errors, dotted line).
+Bottom panel shows that the {\it Gaia} tangential velocity errors are smaller than the internal velocity dispersion of nearer regions of both Sgr and Orp.\footnote{\url{http://www.rssd.esa.int/index.php?page=Science_Performance&project=GAIA}} }\label{fig:gaia_errors}
 \end{center}
@@
 \label{sec:gaia}
-The Gaia satellite \citep{gaia01} is
+The {\it Gaia} satellite \citep{gaia01} is
 an astrometric mission which aims to measure the positions of billions
@@
 
-Figure~\ref{fig:gaia_errors} shows the Gaia end-of-mission distance
-and tangential velocity error estimates for RR Lyrae. Within 2~kpc, Gaia will measure distances to these stars with
+Figure~\ref{fig:gaia_errors} shows the {\it Gaia} end-of-mission distance
+and tangential velocity error estimates for RR Lyrae. Within 2~kpc, {\it Gaia} will measure distances to these stars with
 better than 2\% accuracy --- RR Lyrae in this volume can be used to
-test and calibrate the Spitzer PL relation described above. Beyond the
+test and calibrate the {\it Spitzer} PL relation described above. Beyond the
 2~kpc threshold, the mid-IR PL relation for RR Lyrae will provide better 
@@
 
-The combination of Spitzer and Gaia data will 
+The combination of {\it Spitzer} and {\it Gaia} data will 
 extend the ``horizon'' of where precise, six-dimensional phase-space 
@@
 \label{sec:sgr}
-The Sagittarius Dwarf Galaxy (hereafter Sgr) was discovered
-serendipitously during a radial velocity survey of the Galactic bulge
-\citep{ibata94}. 
+Sgr was discovered serendipitously during a radial velocity 
+survey of the Galactic bulge \citep{ibata94}. 
 Signatures of extensive stellar
@@
 of RR Lyrae in the Sgr debris will open up new avenues of
-interpretation. A 2\% distance error at $\sim$30~kpc
-($\epsilon_{d}=0.6~{\rm kpc}$) is smaller than the predicted thickness
-of the stream ($\sim$2-$10$~kpc). When combined with Gaia proper motion errors,
-tangential velocity measurement accuracies ($\epsilon_{v}\sim2$~km/s
-at 30~kpc; see Figure~\ref{fig:gaia_errors}) are less than the intrinsic dispersion for much of the stream
-\citep[$\sigma_v\sim10$~km/s;][]{majewski04}. The next section
-outlines a new method to take advantage of this information.
+interpretation.
+
+Figure~\ref{fig:gaia_errors} shows that a 2\% distance error is smaller than the 
+distance range in the stream (top panel). Similarly, {\it Gaia} proper motion
+error estimates correspond to tangential velocity errors less than the velocity
+dispersion for much of the stream. The next section outlines a new method to 
+take advantage of this information.
+
+%A 2\% distance error at $\sim$30~kpc
+%($\epsilon_{d}=0.6~{\rm kpc}$) is smaller than the predicted thickness
+%of the stream ($\sim$2-$10$~kpc). When combined with {\it Gaia} proper motion errors,
+%tangential velocity measurement accuracies ($\epsilon_{v}\sim2$~km/s
+%at 30~kpc; see Figure~\ref{fig:gaia_errors}) are less than the intrinsic dispersion for much of the stream
+%\citep[$\sigma_v\sim10$~km/s;][]{majewski04}. The next section
+%outlines a new method to take advantage of this information.
 
@@
 (imagine watching satellite destruction in ``rewind"), the stars
-should recombine with the progenitor. If the potential is incorrect,
+should recombine with the progenitor.\footnote{\url{http://dx.doi.org/10.6084/m9.figshare.767305}} If the potential is incorrect,
 the orbits of the stars will diverge from that of the progenitor and
-thus will not be recaptured by the satellite system (Figure~\ref{fig:ps_distance}).
+thus will not be recaptured by the satellite system (Figure~\ref{fig:ps_distance}).\footnote{\url{http://dx.doi.org/10.6084/m9.figshare.767306}}
 
@@
   km/s})$ and $\varepsilon_{\rm D} \sim \mathcal{N}(0,0.02\times
-D)$) and the proper motion errors are computed from the expected Gaia error curve.
+D)$) and the proper motion errors are computed from the expected {\it Gaia} error curve.
 
@@
 recovered parameters and one-sigma deviations to be 
-$q_1 = 1.36 \pm 0.02$, $q_z = 1.35 \pm 0.03$, $\phi = 96.6 \pm 0.6$ degrees,
-and $v_{\rm halo} = 123.0 \pm 1.5$ km/s.
+$q_1 = 1.36 \pm 0.02$, $q_z = 1.36 \pm 0.03$, $\phi = 96.0 \pm 1.5$ degrees,
+and $v_{\rm halo} = 123.2 \pm 1.6$ km/s.
 
@@
 vast stellar debris {\it clouds} that have been discovered 
-\citep[e.g. the Triangulum Andromeda and Hercules-Aquila
-  clouds]{rochapinto04,belokurov06}, or even stars that have only
+\citep[e.g., the Triangulum-Andromeda and Hercules-Aquila
+  clouds;][]{rochapinto04,belokurov06}, or even stars that have only
 associations in orbital properties and do not form a coherent spatial
@@
 This paper presents an algorithm for measuring the Galactic potential
-that anticipates combined data from the Spitzer and Gaia satellite
+that anticipates combined data from the {\it Spitzer} and {\it Gaia} satellite
 missions which promise precise, full phase-space measurements of RR
@@
 
-Our results also motivate an observational campaign with \emph{Spitzer} to survey RR Lyrae stars in debris structures around the Milky Way to get precise distances to combine with near-future Gaia velocity data.
+Our results also motivate an observational campaign with {\it Spitzer} to survey RR Lyrae stars in debris structures around the Milky Way to get precise distances to combine with near-future {\it Gaia} velocity data.
 If just 100 stars in a single stellar stream allow us to measure the depth, shape, and orientation of the Milky Way potential, larger samples in multiple structures (e.g., the Orphan Stream, already mapped in RR Lyrae by \citet{sesar13}) offer the prospect of assessing these quantities as a function of Galactocentric radius.

### 2013-08-09 scripts/paper1_figures.py
bootstrap figure

@@ -472,7 +472,7 @@ def plot_cov_ellipse(cov, pos, nstd=2, ax=None, **kwargs):
 
 def bootstrapped_parameters():
     data_file = os.path.join(project_root, "plots", "hotfoot", 
-                             "SMASH", "all_best_parameters.pickle")
+                             "SMASH_aspen", "all_best_parameters.pickle")
     
     with open(data_file) as f:
         data = pickle.load(f)
@@ -524,7 +524,7 @@ def bootstrapped_parameters():
 
 def bootstrapped_parameters_transpose():
     data_file = os.path.join(project_root, "plots", "hotfoot", 
-                             "SMASH", "all_best_parameters.pickle")
+                             "SMASH_aspen", "all_best_parameters.pickle")
     
     with open(data_file) as f:
         data = pickle.load(f)
@@ -597,7 +597,7 @@ def bootstrapped_parameters_transpose():
 
 def parameter_errors():
     data_file = os.path.join(project_root, "plots", "hotfoot", 
-                             "SMASH", "all_best_parameters.pickle")
+                             "SMASH_aspen", "all_best_parameters.pickle")
     
     with open(data_file) as f:
         data = pickle.load(f)
@@ -621,4 +621,4 @@ if __name__ == '__main__':
     #variance_projections()
     #bootstrapped_parameters()
     bootstrapped_parameters_transpose()
-    #parameter_errors()
+    parameter_errors()

### 2013-08-09 scripts/paper1_figures.py
make pdf lm10

@@ -13,8 +13,8 @@ import cPickle as pickle
 # Third-party
 import astropy.units as u
 import numpy as np
-import matplotlib.pyplot as plt
 import matplotlib
+import matplotlib.pyplot as plt
 from matplotlib import rc_context, rcParams, cm
 from matplotlib.patches import Rectangle, Ellipse
 
@@ -391,7 +391,7 @@ def sgr():
         ax.plot(pdata['x'], pdata['z'], marker='.', alpha=0.85, ms=9)
         
         # add solar symbol
-        ax.text(-8., 0., s=r"$\odot$")
+        ax.text(-8., 0., s=r"$\odot$", fontsize=32)
 
         ax.set_xlabel("$X_{GC}$ [kpc]")
         ax.set_xlabel("$X_{GC}$ [kpc]")
@@ -407,7 +407,7 @@ def sgr():
     
     ax.set_aspect('equal')
     
-    fig.savefig(os.path.join(plot_path, "lm10.png"))
+    fig.savefig(os.path.join(plot_path, "lm10.pdf"))
 
 def plot_point_cov(points, nstd=2, ax=None, **kwargs):
     """
@@ -615,10 +615,10 @@ def parameter_errors():
 
 if __name__ == '__main__':
     #gaia_spitzer_errors()
-    #sgr()
+    sgr()
     #phase_space_d_vs_time()
     #normed_objective_plot()
     #variance_projections()
     #bootstrapped_parameters()
-    bootstrapped_parameters_transpose()
-    parameter_errors()
+    #bootstrapped_parameters_transpose()
+    #parameter_errors()

### 2013-08-09 text/paper1/ms.tex
rename to ms

--- 2013-08-09 text/paper1/paper.tex
@@
 \begin{center}
-\includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/gaia.pdf}
+\includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/gaia.eps}
 
@@
 \begin{center}
-\includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/lm10.png}
+\includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/lm10.eps}
 \caption{ Particle density (blue) of the first leading and trailing wraps from the final time-step of the \citet{law10} simulation of the Sgr stream.
@@
 error estimates correspond to tangential velocity errors less than the velocity
-dispersion for much of the stream. The next section outlines a new method to 
-take advantage of this information.
+dispersion for much of the stream (bottom panel). The next section outlines a 
+new method to take advantage of this information.
 
@@
 \begin{center}
-\includegraphics[width=0.47\textwidth,trim=5 0 0 0, clip]{/Users/adrian/projects/streams/plots/paper1/ps_distance.pdf}
+\includegraphics[width=0.47\textwidth,trim=5 0 0 0, clip]{/Users/adrian/projects/streams/plots/paper1/ps_distance.eps}
 \caption{Phase space distance ($D_{\rm ps}$) for 10 stars integrated backwards in the correct potential (top) and a potential where $q_z$ is 25\% larger (bottom).
@@
 \begin{center}
-\includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/objective_function.pdf}
+\includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/objective_function.eps}
 \caption{ 1D slices of the objective function (generalized variance) for each halo potential parameter.
@@
 \begin{figure*}[ht!]
-\centering\includegraphics[width=0.9\textwidth,trim=0 0 0 0, clip]{/Users/adrian/projects/streams/plots/paper1/bootstrap.pdf}
+\centering\includegraphics[width=0.9\textwidth,trim=0 0 0 0, clip]{/Users/adrian/projects/streams/plots/paper1/bootstrap.eps}
 \caption{ Blue points show the ``best-fit'' parameters resulting from each resample of 100 stars from the Sgr stream particle density shown in Figure~\ref{fig:lm10}.

### 2013-08-09 text/paper1/ms.tex
added keywords

--- 2013-08-09 text/paper1/paper.tex
@@
 \end{abstract}
+
+\keywords{
+  Galaxy: structure
+  ---
+  Galaxy: halo
+  ---
+  cosmology: dark matter
+}
 
@@
 \begin{center}
+%\includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/gaia.eps}
 \includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/gaia.pdf}
@@
 \begin{center}
-\includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/lm10.png}
+%\includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/lm10.eps}
+\includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/lm10.pdf}
 \caption{ Particle density (blue) of the first leading and trailing wraps from the final time-step of the \citet{law10} simulation of the Sgr stream.
@@
 \begin{center}
+%\includegraphics[width=0.47\textwidth,trim=5 0 0 0, clip]{/Users/adrian/projects/streams/plots/paper1/ps_distance.eps}
 \includegraphics[width=0.47\textwidth,trim=5 0 0 0, clip]{/Users/adrian/projects/streams/plots/paper1/ps_distance.pdf}
@@
 \begin{center}
+%\includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/objective_function.eps}
 \includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/objective_function.pdf}
@@
 \begin{figure*}[ht!]
+%\centering\includegraphics[width=0.9\textwidth,trim=0 0 0 0, clip]{/Users/adrian/projects/streams/plots/paper1/bootstrap.eps}
 \centering\includegraphics[width=0.9\textwidth,trim=0 0 0 0, clip]{/Users/adrian/projects/streams/plots/paper1/bootstrap.pdf}

### 2013-08-13 streams/inference/lm10.py
updated objective function and added notebook for testing

@@ -108,6 +108,45 @@ def timestep(r, v):
     
     return dt
 
+def objective(potential, satellite_orbit, particle_orbits):
+    """ This is a new objective function, motivated by the fact that what 
+        I was doing before doesn't really make sense...
+    """
+    
+    # get numbers for any relevant loops below
+    Ntimesteps, Nparticles, Ndim = particle_orbits._r.shape
+    
+    r_tide = potential._tidal_radius(m=satellite_orbit._m,
+                                     r=satellite_orbit._r)[:,:,np.newaxis]
+    v_esc = potential._escape_velocity(m=satellite_orbit._m,
+                                       r_tide=r_tide)
+    
+    # compute relative, normalized coordinates and then phase-space distance
+    R = particle_orbits._r - satellite_orbit._r
+    V = particle_orbits._v - satellite_orbit._v
+    Q = R / r_tide
+    P = V / v_esc
+    D_ps = np.sqrt(np.sum(Q**2, axis=-1) + np.sum(P**2, axis=-1))
+    
+    # Find the index of the time of the minimum D_ps for each particle
+    min_time_idx = D_ps.argmin(axis=0)
+    
+    cov = np.zeros((6,6))
+    b = np.vstack((R.T, V.T)).T
+    for ii in range(Nparticles):
+        idx = min_time_idx[ii]
+        r_t = np.squeeze(r_tide[idx])
+        v_e = np.squeeze(v_esc[idx])
+        cov += np.outer(b[idx,ii], b[idx,ii].T)
+    cov /= Nparticles
+    
+    # velocity dispersion from measuring the dispersion of the still-bound
+    #   particles from LM10
+    r_disp = r_tide
+    v_disp = 0.0133 # kpc/Myr
+    eig_vals = np.linalg.eigvals(cov) / np.array([r_disp]*3+[v_disp]*3)**2
+    return np.sum(eig_vals)
+
 def ln_likelihood(p, param_names, particles, satellite, t1, t2, resolution):
     """ Evaluate the likelihood function for a given set of halo 
         parameters.
@@ -128,7 +167,7 @@ def ln_likelihood(p, param_names, particles, satellite, t1, t2, resolution):
                                       resolution=resolution,
                                       t1=t1, t2=t2)
     
-    return -generalized_variance(lm10, p_orbits, s_orbit)
+    return -objective(lm10, p_orbits, s_orbit)
 
 def ln_posterior(p, *args):
     param_names, particles, satellite, t1, t2, resolution = args

### 2013-08-13 streams/observation/gaia.py
fix size of proper motion error array

@@ -159,9 +159,13 @@ def rr_lyrae_add_observational_uncertainties(x,y,z,vx,vy,vz,**kwargs):
         dmu = proper_motion_error(V, rrl_V_minus_I)            
     
     dmu = (dmu.to(u.rad/u.s).value / u.s).to(u.km / (u.kpc*u.s))
+    try:
+        size = len(dmu)
+    except TypeError:
+        size = len(vr)
     
-    mul += np.random.normal(0., dmu.value, size=len(dmu))*dmu.unit
-    mub += np.random.normal(0., dmu.value, size=len(dmu))*dmu.unit
+    mul += np.random.normal(0., dmu.value, size=size)*dmu.unit
+    mub += np.random.normal(0., dmu.value, size=size)*dmu.unit
     
     # compute tangential velocities as distance times proper motion
     v_l = d*mul

### 2013-09-19 text/paper1/ms.tex
typo in email address
clarified figure 3

@@
 %\altaffiltext{2}{NSF Graduate Fellow}
-\altaffiltext{2}{\email{adrn@astro.columia.edu}}
+\altaffiltext{2}{\email{adrn@astro.columbia.edu}}
 
@@
 \includegraphics[width=0.47\textwidth,trim=5 0 0 0, clip]{/Users/adrian/projects/streams/plots/paper1/ps_distance.pdf}
-\caption{Phase space distance ($D_{\rm ps}$) for 10 stars integrated backwards in the correct potential (top) and a potential where $q_z$ is 25\% larger (bottom).
+\caption{Phase space distance ($D_{\rm ps}$) for 10 randomly selected stars integrated backwards in the correct potential (top) and a potential where $q_z$ is 25\% larger (bottom).
 The same 10 particles are used in both figures, so the initial conditions are identical.

### 2013-09-19 text/paper1/ms.tex
added sentence to clarify changing tidal radius and vesc with time

@@
 \end{equation}
+These quantities are computed at each time step to take the time 
+dependence into account, neglecting mass-loss from the satellite.
 Qualitatively, when the distance in this normalized six-dimensional

### 2013-09-26 text/paper1/ms.tex
updated paper and ref response

@@
 current data on the Sgr debris to constrain both a model of its evolution
-and the potential in which it orbits.
+and the potential in which it orbits.\footnote{However, new observational work by 
+\cite{belokurov13} suggest that the trailing tail of Sgr debris does not
+match the LM10 model.}
 Figure \ref{fig:lm10} shows particle positions
@@
 original distribution in populations. 
-This comparison enabled the first 
-assessment of the three-dimensional mass distribution of a dark matter
+This comparison enabled an assessment of the three-dimensional 
+mass distribution of the Milky Way's dark matter
 halo through the constraints on the potential parameters $v_{\rm halo}$, $q_1$, $q_z$, and $\phi$.
@@
 Qualitatively, when the distance in this normalized six-dimensional
-space, $D_{{\rm ps},i}=\sqrt{|\bs{q}_{i}|^2+|\bs{p}_{i}|^2} <
-\sqrt{2}$, the star is likely recaptured by the
-satellite. \citet{johnston99a} imposed a similar
-condition as a hard boundary and used the number of recaptured
-particles in a given backwards-integration as a cost function for
-optimization. What follows is a description of an updated procedure
-with a more natural, statistically-motivated choice for an objective
-function.
+space, $D_{{\rm ps},i}=\sqrt{|\bs{q}_{i}|^2+|\bs{p}_{i}|^2} \lesssim 2$, 
+the star is likely recaptured by the satellite (in the absence of errors,
+ we find that $\sim$90\% of the initially bound particles come within
+ this limit when integrating all orbits backwards). \citet{johnston99a} 
+ imposed a similar condition as a hard boundary and used the number 
+of recaptured particles in a given backwards-integration as a cost 
+function for optimization. What follows is a description of an updated 
+procedure with a more natural, statistically-motivated choice for an 
+objective function.
 
@@
 
+\bibliographystyle{apj}
 %\bibliography{refs}
-\bibliographystyle{apj}
 
@@
 {Belokurov}, V., {Evans}, N.~W., {Irwin}, M.~J., {et~al.} 2007, \apj, 658, 337
+
+\bibitem[{{Belokurov} {et~al.}(2013){Belokurov}, {Koposov}, {Evans},
+  {Pe{\~n}arrubia}, {Irwin}, {Smith}, {Lewis}, {Gieles}, {Wilkinson},
+  {Gilmore}, {Olszewski}, \& {Niederste-Ostholt}}]{belokurov13}
+{Belokurov}, V., {Koposov}, S.~E., {Evans}, N.~W., {et~al.} 2013, ArXiv
+  e-prints
 

### 2013-09-26 text/paper1/ms.tex

@@
 Qualitatively, when the distance in this normalized six-dimensional
-space, $D_{{\rm ps},i}=\sqrt{|\bs{q}_{i}|^2+|\bs{p}_{i}|^2} \lesssim 2$, 
-the star is likely recaptured by the satellite (in the absence of errors,
- we find that $\sim$90\% of the initially bound particles come within
- this limit when integrating all orbits backwards). \citet{johnston99a} 
- imposed a similar condition as a hard boundary and used the number 
-of recaptured particles in a given backwards-integration as a cost 
-function for optimization. What follows is a description of an updated 
-procedure with a more natural, statistically-motivated choice for an 
-objective function.
+space, $D_{{\rm ps},i}=\sqrt{|\bs{q}_{i}|^2+|\bs{p}_{i}|^2} <
+\sqrt{2}$, the star is likely recaptured by the
+satellite. \citet{johnston99a} imposed a similar
+condition as a hard boundary and used the number of recaptured
+particles in a given backwards-integration as a cost function for
+optimization. What follows is a description of an updated procedure
+with a more natural, statistically-motivated choice for an objective
+function.
 

### 2013-09-27 text/paper1/ms.tex
remove sgr orbit from fig 3 and update caption

--- 2013-09-26 text/paper1/ms.tex
@@
 \citep[e.g.,][]{johnston99b, fellhauer06}.
-Most recently, \citet[][hereafter LM10]{law10} have combined all
-current data on the Sgr debris to constrain both a model of its evolution
-and the potential in which it orbits.\footnote{However, new observational work by 
+Most recently, \citet[][hereafter LM10]{law10} combined all
+the (then) current data on the Sgr debris to constrain both a model of its evolution
+and the potential in which it orbits. (Note, new observational work by 
 \cite{belokurov13} suggest that the trailing tail of Sgr debris does not
-match the LM10 model.}
+match the LM10 model.)
 Figure \ref{fig:lm10} shows particle positions
@@
 
-LM10 demonstrate how a comparison of such simulations with the
-observed debris enables a detailed reconstruction of Sgr's history:
-its mass and stellar populations, orbit, rate of destruction, and even
-original distribution in populations. 
-This comparison enabled an assessment of the three-dimensional 
+%LM10 demonstrate how a comparison of such simulations with the observed debris enables a detailed reconstruction of Sgr's history: its mass and stellar populations, orbit, rate of destruction, and even original distribution in populations.
+A comparison of simulations and data enabled LM10 to make an assessment of the three-dimensional 
 mass distribution of the Milky Way's dark matter
-halo through the constraints on the potential parameters $v_{\rm halo}$, $q_1$, $q_z$, and $\phi$.
-
-Combined {\it Spitzer} and {\it Gaia} measurements of distances and proper motions
-of RR Lyrae in the Sgr debris will open up new avenues of
-interpretation.
-
-Figure~\ref{fig:gaia_errors} shows that a 2\% distance error is smaller than the 
+halo through the constraints on the potential parameters $v_{\rm halo}$, $q_1$, $q_z$, and $\phi$. Combined {\it Spitzer} and {\it Gaia} measurements of distances and proper motions
+of RR Lyrae in the Sgr debris will open up new avenues for potential constraints. Figure~\ref{fig:gaia_errors} shows that a 2\% distance error is smaller than the 
 distance range in the stream (top panel). Similarly, {\it Gaia} proper motion
@@
 The same 10 particles are used in both figures, so the initial conditions are identical.
-Red (thick, solid line) shows the galactocentric distance of the progenitor (in arbitrary, normalized units) over the integration.
-Horizontal (dashed) line shows $D_{\rm ps}=\sqrt{2}$.}\label{fig:ps_distance}
+Horizontal (dashed) line shows $D_{\rm ps}=2$, for reference.}\label{fig:ps_distance}
 \end{center}
@@
 \caption{ 1D slices of the objective function (generalized variance) for each halo potential parameter.
-Curves are normalized to have the same value at the true parameter value (1.00), and show the effect of varying each parameter by $\pm10$\%.
-}\label{fig:objective}
+The parameter values are normalized by the true values show the effect of varying each parameter by $\pm10$\%.
+The values of the objective function (vertical axis) are not interesting but note the minima around the truth (1.0).}\label{fig:objective}
 \end{center}
@@
 the minute differences in orbital properties between the debris and
-satellite. Second, the current version of the algorithm
+satellite. Second, the present prescription neglects orbital evolution (e.g.,
+dynamical friction) and scattering of stream stars due to the potential
+of the satellite itself. Preliminary simulations (to be fully explored in 
+forthcoming work) suggest that these first two points can be neglected 
+for satellite masses between $\sim$$10^7$ and $\sim$$10^9~\mathrm{M}_{\odot}$. Lastly, the 
+current version of the algorithm
 relies on accurate knowledge of the current position and velocity of
 the parent satellite, which may not be available 
-\citep[for example, the Orphan Stream;][]{belokurov07}. Lastly, the
-present prescription neglects orbital evolution (e.g.,
-dynamical friction) and N-body interactions which may lead to biases in
-the estimated potential parameters.
+\citep[for example, the Orphan Stream;][]{belokurov07}.
 

### 2013-09-27 scripts/figures/paper1_figures.py
removed sgr orbit from D_ps plot

@@ -210,21 +210,16 @@ def gaia_spitzer_errors(**kwargs):
 def phase_space_d_vs_time(**kwargs):
     """ Plot the PSD for 10 stars vs. back-integration time. """
     
+    N = int(kwargs.get("N", 10))
+    seed = int(kwargs.get("seed", 112358))
+
     # Read in the LM10 data
-    np.random.seed(142)
-    particles = particles_today(N=100, expr="(Pcol>-1) & (abs(Lmflag) == 1) & (Pcol < 7)")
+    np.random.seed(seed)
+    particles = particles_today(N=N, expr="(Pcol>-1) & (abs(Lmflag) == 1) & (Pcol < 7)")
     satellite = satellite_today()
     t1,t2 = time()
+    t2 = -6000.
 
-    N = int(kwargs.get("N", 10))
-    seed = int(kwargs.get("seed", 41))
-
-    #np.random.seed(112)
-    np.random.seed(seed)
-    randidx = np.random.randint(100, size=N)
-    selected_star_idx = np.zeros(100).astype(bool)
-    selected_star_idx[randidx] = True
-    
     wrong_params = true_params.copy()
     wrong_params['qz'] = 1.25*wrong_params['qz']
     #wrong_params['v_halo'] = wrong_params['v_halo']
@@ -243,8 +238,7 @@ def phase_space_d_vs_time(**kwargs):
         s_orbit,p_orbits = integrator.run(t1=t1, t2=t2, dt=-1.)
         
         R,V = relative_normalized_coordinates(potential, s_orbit, p_orbits) 
-        #D_ps = np.sqrt(np.sum(R**2, axis=-1)/2 + np.sum(V**2, axis=-1))
-        D_ps = np.sqrt(np.sum(R**2, axis=-1)/2)
+        D_ps = np.sqrt(np.sum(R**2, axis=-1) + np.sum(V**2, axis=-1))
         D_pses.append(D_ps)
         sat_R.append(np.sqrt(np.sum(s_orbit._r**2, axis=-1)))
         ts.append(s_orbit._t)
@@ -271,16 +265,16 @@ def phase_space_d_vs_time(**kwargs):
     
     with rc_context(rc=rcparams):  
         fig,axes = plt.subplots(2,1, sharex=True, sharey=True, figsize=(8,12))
-        axes[0].axhline(np.sqrt(2), linestyle='--', color='#444444')
[…]

### 2013-10-08 text/proposals/sigma_xi_fall2013.tex
adding sigma Xi proposal

\documentclass[preprint]{aastex}

\begin{document}
\title{Probing the Milky Way's Dark Matter Halo with RR Lyraes} \author{Adrian M.
  Price-Whelan\altaffilmark{1}, Kathryn Johnston\altaffilmark{1},
  Allyson Sheffield\altaffilmark{1}} \altaffiltext{1}{Department of
  Astronomy, Columbia University, Mail Code 5246, New York, NY 10027}

Galaxies, by mass, are mostly composed of \emph{Dark
  Matter}. Simulations of structure and galaxy formation -- both with
and without baryons -- lead us to believe that galaxies live in halos of
dark matter that have universal density profiles \citep{nfw96} and are
triaxial in shape \citep{jing02}. Including baryons in such
simulations tends to soften the triaxiality and alters the radial
density profile of the dark matter. Thus, precise measurements of the 
shape, orientation, and radial profile of dark matter halos provide 
constraints on both dark matter physics and the baryonic processes 
that helped shape them.

Tidal streams and substructure in the Milky Way's 
halo (from disrupted, infalling satellite galaxies) have been used
to constrain properties of the mass distribution around the Milky Way \citep[e.g.,][]{law10}.
Typically found outwards of 20~kpc from the Galactic center, the orbits 
of associated stars are still largely under the influence of the dark matter 
and thus serve as tracers of mass \citep[e.g.,][]{johnston99, pricewhelan13}. 
Tens of stellar 
streams and spatial overdensities have been identified in the Sloan
Digital Sky Survey (SDSS) and Two-Micron All Sky Survey (2MASS)
and a variety of other deep, photometric studies \citep[e.g.,][]{majewski03,belokurov06}. 
For most of these interesting relics of the Milky Way's merger history, 
only sky positions and rough photometric distances have been measured. 
A combination of better distance measurements with follow-up spectroscopy
would allow for a more detailed reconstruction of these stellar overdensities 
by enabling direct measurements of radial velocities and metallicities --- 
both powerful tools for distinguishing populations of stars formed in satellite 
galaxies from the background of Milky Way halo stars.

Measuring distances to stars is notoriously difficult. Recently, \citet{madore12}
have shown that a class of variable star --- RR Lyrae stars --- have a 
tight period-luminosity relation in the mid-infrared. By comparing distances
measured using this relation to five stars stars with trigonometric parallaxes 
measured by Hubble, they demonstrate that it is possible to use observations with
the Spitzer space telescope to determine distances that are good to 2\% for individual RR Lyrae 
stars out to $\sim$60~kpc. As a part of a much
larger collaboration, we have applied for observing time with Spitzer to measure
mid-infrared light curves for hundreds of stars in the Sagittarius stream \citep[Sgr;][]{majewski03},
the Orphan stream \citep[Orp;][]{belokurov07}, and the Triangulum-Andromeda 
overdensity \citep[TriAnd;][]{rochapinto04}.

We have already received time on the 2.4m Hiltner telescope at MDM 
observatory (August 2013 and October 2013) to obtain spectra for stars
associated with TriAnd. We are applying for i) support for travel for three
observing runs to be conducted this Spring (2014) to follow up both 
calibration stars and Sgr and Orp stream stars targeted with Spitzer, and ii) 
equipment for data reduction and storage.
Combined with future data (proper motions) from the ESA's Gaia mission,
we will for the first time have precise, 6D kinematic information
for samples of stars at large distances in the Galactic halo. This data will 
enable us to model the large-scale mass distribution around the Milky Way 
with unprecedented detail.

\begin{thebibliography}{ }
\bibitem[{{Belokurov} {et~al.}(2006){Belokurov}, {Zucker}, {Evans}, {Gilmore},
  {Vidrih}, {Bramich}, {Newberg}, {Wyse}, {Irwin}, {Fellhauer}, {Hewett},
  {Walton}, {Wilkinson}, {Cole}, {Yanny}, {Rockosi}, {Beers}, {Bell},
  {Brinkmann}, {Ivezi{\'c}}, \& {Lupton}}]{belokurov06}
{Belokurov}, V., {Zucker}, D.~B., {Evans}, N.~W., {et~al.} 2006, \apjl, 642,
  L137
\bibitem[{{Belokurov} {et~al.}(2007){Belokurov}, {Evans}, {Irwin},
  {Lynden-Bell}, {Yanny}, {Vidrih}, {Gilmore}, {Seabroke}, {Zucker},
  {Wilkinson}, {Hewett}, {Bramich}, {Fellhauer}, {Newberg}, {Wyse}, {Beers},
  {Bell}, {Barentine}, {Brinkmann}, {Cole}, {Pan}, \& {York}}]{belokurov07}
{Belokurov}, V., {Evans}, N.~W., {Irwin}, M.~J., {et~al.} 2007, \apj, 658, 337
\bibitem[Jing \& Suto(2002)]{jing02} Jing, Y.~P., \& Suto, Y.\ 2002, \apj, 574, 538
\bibitem[Johnston et al.(1999)]{johnston99} Johnston, K.~V., Zhao, H., Spergel, D.~N., \& Hernquist, L.\ 1999, \apjl, 512, L109 
\bibitem[Law \& Majewski(2010)]{law10} Law, D.~R., \& Majewski, S.~R.\ 2010, \apj, 714, 229 
\bibitem[Madore \& Freedman(2012)]{madore12} Madore, B.~F., \& Freedman, W.~L.\ 2012, \apj, 744, 132 
\bibitem[{{Majewski} {et~al.}(2003){Majewski}, {Skrutskie}, {Weinberg}, \&
  {Ostheimer}}]{majewski03}
{Majewski}, S.~R., {Skrutskie}, M.~F., {Weinberg}, M.~D., \& {Ostheimer}, J.~C.
  2003, \apj, 599, 1082
\bibitem[Navarro et al.(1996)]{nfw96} Navarro, J.~F., Frenk, C.~S., \& White, S.~D.~M.\ 1996, \apj, 462, 563
\bibitem[Price-Whelan \& Johnston(2013)]{pricewhelan13} {Price-Whelan}, A.~M. \& {Johnston}, K.~V. {arXiv:1308.2670}, August 2013.
\bibitem[{{Rocha-Pinto} {et~al.}(2004){Rocha-Pinto}, {Majewski}, {Skrutskie},
  {Crane}, \& {Patterson}}]{rochapinto04}
{Rocha-Pinto}, H.~J., {Majewski}, S.~R., {Skrutskie}, M.~F., {Crane}, J.~D., \&
  {Patterson}, R.~J. 2004, \apj, 615, 732
\end{thebibliography}

\end{document}

### 2013-10-10 scripts/figures/paper1_figures.py
tub figure

@@ -534,63 +534,59 @@ def when_particles_recombine(**kwargs):
     tub = (data['tub']*usys['time']).to(u.Myr).value
     R_orbit = np.sqrt(np.sum(xs**2, axis=-1))
     
-    tub_file = os.path.join(_full_path, "tub_back{0}_{1}.pickle".format(N,D_ps_limit))
-    if overwrite and os.path.exists(tub_file):
-        os.remove(tub_file)
-
-    if not os.path.exists(tub_file):
-        # read full orbits of particles from running on hotfoot
-        p_x = np.load(os.path.join(project_root, "data", "{0}particles_p.npy".format(N)))
-        s_x = np.load(os.path.join(project_root, "data", "{0}particles_s.npy".format(N)))
-
-        r_tide = potential._tidal_radius(m=satellite._m,
-                                         r=s_x[...,:3])[:,:,np.newaxis]
-
-        v_esc = potential._escape_velocity(m=satellite._m,
-                                           r_tide=r_tide)
-    
-        R,V = (p_x[...,:3] - s_x[...,:3]) / r_tide, (p_x[...,3:] - s_x[...,3:]) / v_esc
-        D_ps = np.sqrt(np.sum(R**2, axis=-1) + np.sum(V**2, axis=-1))
-        tub_back = []
-        for ii in range(D_ps.shape[1]):
-            idx = D_ps[:,ii] < D_ps_limit
-            if np.any(ts[idx] == ts[0]):
-                tub_back.append(ts[0])
-            elif np.any(idx):
-                tub_back.append(np.median(ts[idx]))
-                #tub_back.append(ts[idx][-1])
-            else:
-                #print("min Dps:",np.min(D_ps[:,ii]))
-                #print(ts[D_ps[:,ii].argmin()])
-                tub_back.append(np.nan)
-        tub_back = np.array(tub_back)
-        fnpickle(tub_back, tub_file)
-
-    tub_back = fnunpickle(tub_file)
-    bound_stars_back = [np.sum(tub_back > t) for t in ts]
-
     from scipy.signal import argrelextrema
     apos,xxx = argrelextrema(R_orbit, np.greater)
     peris,xxx = argrelextrema(R_orbit, np.less)    
 
     fig,axes = plt.subplots(2,1,sharex=True,figsize=(18,12))
     fig.suptitle("Capture condition: $D_{{ps}}$<{0}".format(D_ps_limit), fontsize=32)
-        
[…]

### 2013-10-10 scripts/back_integrate_movies.py
back integrate movies

@@ -195,6 +195,12 @@ def plot_xz_animation(potential, s, p, filename=""):
     print("Fraction bound at end of run: {0}".format(sum(~idx)/p.nparticles))
 
 if __name__ == "__main__":
+    print(ffmpeg_cmd.format(os.path.join(plot_path, "correct_%4d.png"),
+                                os.path.join(plot_path, "correct.mp4")))
+    print(ffmpeg_cmd.format(os.path.join(plot_path, "wrong_%4d.png"),
+                                os.path.join(plot_path, "wrong.mp4")))
+    sys.exit(0)
+
     N = 10000
     
     # read in lm10 stuffs

### 2013-10-10 text/paper1/ms.tex

@@
 
-This approach assumes that the tracers represent a random sampling of orbits drawn from a smooth distribution function and are well-mixed in orbital phase.
-However, the same surveys that provide large samples have also revealed the existence of large-scale spatial inhomogeneities in the form of giant stellar streams \citep{newberg02,majewski03,belokurov06}, demonstrating that a significant fraction of the stellar halo is neither randomly sampled nor is fully phase-mixed.
+This approach assumes that the tracers represent a random sampling of phase-mixed orbits drawn from a smooth distribution function, however large area surveys have revealed the existence of large-scale spatial inhomogeneities in the form of giant stellar streams \citep{newberg02,majewski03,belokurov06}, demonstrating that a significant fraction of the stellar halo is neither randomly sampled nor is fully phase-mixed.
 
@@
 changes in orbital properties along debris streams can lead to systematic biases in measurements of the Galactic potential \citep{eyre09a,varghese11}.
-In particular, \citet{sanders13a} recently demonstrated that this bias is equally problematic for the very thinnest, coldest streams, whose observed properties may be indistinguishable from those of the parent orbit \citep[e.g.
+\citet{sanders13a} recently demonstrated that this bias is equally problematic for the very thinnest, coldest streams, whose observed properties may be indistinguishable from those of the parent orbit \citep[e.g.
 such as the globular cluster, GD1 --- see][]{koposov10}, as for the much more extended and hotter streams \citep[e.g.
@@
 One way to address these biases is to run self-consistent N-body simulations of satellite destruction in a variety of potentials with the aim of simultaneously constraining both the properties of the satellite and the Milky Way.
-Many studies of the Sagittarius debris system (hereafter Sgr) --- whose tidal streams entirely encircle the Milky Way --- have adopted this approach 
-with the most recent work for the first time attempting to place constraints on the triaxiality and orientation of the dark matter halo
+Many studies of the Sagittarius debris system (hereafter Sgr) have adopted this approach,
+with the most recent work attempting to place constraints on the triaxiality and orientation of the dark matter halo
 \citep{law10}.
@@
 \begin{center}
-%\includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/gaia.eps}
-\includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/gaia.pdf}
+\includegraphics[width=0.45\textwidth]{fig1.eps}
+%\includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/gaia.pdf}
 
@@
 the (then) current data on the Sgr debris to constrain both a model of its evolution
-and the potential in which it orbits. (Note, new observational work by 
-\cite{belokurov13} suggest that the trailing tail of Sgr debris does not
+and the potential in which it orbits. (Note that new observational work by 
+\citet{belokurov13} suggest that the trailing tail of Sgr debris does not
 match the LM10 model.)
@@
 \begin{center}
-%\includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/lm10.eps}
-\includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/lm10.pdf}
+\includegraphics[width=0.45\textwidth]{fig2.eps}
+%\includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/lm10.pdf}
 \caption{ Particle density (blue) of the first leading and trailing wraps from the final time-step of the \citet{law10} simulation of the Sgr stream.
-Point markers (black) show positions of a representative sample of 100 stars drawn from this density distribution.
+Point markers (black) show positions of a random sample of 100 stars drawn from this density distribution.
 The position of the Sun is shown with the solar symbol.
@@
 mass distribution of the Milky Way's dark matter
-halo through the constraints on the potential parameters $v_{\rm halo}$, $q_1$, $q_z$, and $\phi$. Combined {\it Spitzer} and {\it Gaia} measurements of distances and proper motions
+halo through constraints on the potential parameters $v_{\rm halo}$, $q_1$, $q_z$, and $\phi$. Combined {\it Spitzer} and {\it Gaia} measurements of distances and proper motions
 of RR Lyrae in the Sgr debris will open up new avenues for potential constraints. Figure~\ref{fig:gaia_errors} shows that a 2\% distance error is smaller than the 
@@
 same progenitor: if the orbits of the stars and progenitor are integrated 
-\emph{backwards} in a a potential that accurately models the Milky Way
-(imagine watching satellite destruction in ``rewind"), the stars
-should recombine with the progenitor.\footnote{\url{http://dx.doi.org/10.6084/m9.figshare.767305}} If the potential is incorrect,
+\emph{backwards} in a a potential that accurately models the Milky Way, the stars
+should recombine with the progenitor (imagine watching satellite destruction in ``rewind").\footnote{\url{http://dx.doi.org/10.6084/m9.figshare.767305}} If the potential is incorrect,
 the orbits of the stars will diverge from that of the progenitor and
@@
 approximate estimate, as previously assumed) of the position of a star within its debris
-structure; and the test statistic that quantifies how well stars recombine
-with the satellite has been rigorously redefined.
+structure. The test statistic that quantifies how well stars recombine
+with the satellite has also been rigorously redefined.
 
@@
 \begin{center}
-%\includegraphics[width=0.47\textwidth,trim=5 0 0 0, clip]{/Users/adrian/projects/streams/plots/paper1/ps_distance.eps}
-\includegraphics[width=0.47\textwidth,trim=5 0 0 0, clip]{/Users/adrian/projects/streams/plots/paper1/ps_distance.pdf}
+\includegraphics[width=0.47\textwidth,trim=5 0 0 0, clip]{fig3.eps}
+%\includegraphics[width=0.47\textwidth,trim=5 0 0 0, clip]{/Users/adrian/projects/streams/plots/paper1/ps_distance.pdf}
 \caption{Phase space distance ($D_{\rm ps}$) for 10 randomly selected stars integrated backwards in the correct potential (top) and a potential where $q_z$ is 25\% larger (bottom).
@@
  this limit when integrating all orbits backwards). \citet{johnston99a} 
- imposed a similar condition as a hard boundary and used the number 
-of recaptured particles in a given backwards-integration as a cost 
-function for optimization. What follows is a description of an updated 
-procedure with a more natural, statistically-motivated choice for an 
-objective function.
+ imposed a similar condition as a hard boundary and maximized the 
+number of recaptured particles in a given backwards-integration.
+What follows is a description of an updated procedure with a 
+statistically-motivated choice for an objective function.
 
@@
 distribution computed from the `true' orbital history of the stars: in
-an incorrect potential, small differences in the orbits of the stars
-relative to the orbit of the progenitor cause them to spread out in
-this relative, normalized phase space. Thus, the \emph{generalized
+an incorrect potential, the orbits of the stars relative to the orbit of 
+the progenitor spread out in phase space. Thus, the \emph{generalized
  variance} of the distribution --- computed for a given set of
@@
   km/s})$ and $\varepsilon_{\rm D} \sim \mathcal{N}(0,0.02\times
-D)$) and the proper motion errors are computed from the expected {\it Gaia} error curve.
+D)$) and the proper motion errors are computed from the expected {\it Gaia} error curve.\footnote{\url{http://www.rssd.esa.int/index.php?page=Science_Performance&project=GAIA}}
 
-The generalized variance defines a convex hypersurface over
+The generalized variance defines a convex function over
 which we optimize four of the six logarithmic potential parameters:
-$v_{circ}$, $\phi$, $q_1$, and $q_z$. We fix $q_2$ and $R_c$, which
-are degenerate with combinations of the other
-parameters. Figure~\ref{fig:objective} shows one-dimensional slices of
+$v_{circ}$, $\phi$, $q_1$, and $q_z$ ($q_2$ and $R_c$
+are degenerate with combinations of the other parameters). 
+Figure~\ref{fig:objective} shows one-dimensional slices of
 the objective function produced by varying each of the potential
@@
 \begin{center}
-%\includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/objective_function.eps}
-\includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/objective_function.pdf}
+\includegraphics[width=0.45\textwidth]{fig4.eps}
+%\includegraphics[width=0.45\textwidth]{/Users/adrian/projects/streams/plots/paper1/objective_function.pdf}
 \caption{ 1D slices of the objective function (generalized variance) for each halo potential parameter.
@@
 \begin{figure*}[ht!]
-%\centering\includegraphics[width=0.9\textwidth,trim=0 0 0 0, clip]{/Users/adrian/projects/streams/plots/paper1/bootstrap.eps}
-\centering\includegraphics[width=0.9\textwidth,trim=0 0 0 0, clip]{/Users/adrian/projects/streams/plots/paper1/bootstrap.pdf}
+\centering\includegraphics[width=0.9\textwidth,trim=0 0 0 0, clip]{fig5.eps}
+%\centering\includegraphics[width=0.9\textwidth,trim=0 0 0 0, clip]{/Users/adrian/projects/streams/plots/paper1/bootstrap.pdf}
 \caption{ Blue points show the ``best-fit'' parameters resulting from each resample of 100 stars from the Sgr stream particle density shown in Figure~\ref{fig:lm10}.
@@
 dynamical friction) and scattering of stream stars due to the potential
-of the satellite itself. Preliminary simulations (to be fully explored in 
-forthcoming work) suggest that these first two points can be neglected 
+of the satellite. Preliminary simulations (to be fully explored in 
+forthcoming work) suggest that these two points can be neglected 
 for satellite masses between $\sim$$10^7$ and $\sim$$10^9~\mathrm{M}_{\odot}$. Lastly, the 
-current version of the algorithm
-relies on accurate knowledge of the current position and velocity of
-the parent satellite, which may not be available 
-\citep[for example, the Orphan Stream;][]{belokurov07}.
+current version of the algorithm relies on knowledge of the current 
+position and velocity of the parent satellite, which may not be available 
+\citep[e.g., the Orphan Stream;][]{belokurov07}.
 
@@
 While the tests presented in this paper are very simple, the accuracy of potential recovery promised by such a small sample of stars 
-provides strong motivation for further theoretical work to: 1) develop a robust generative model (and therefore likelihood) that utilizes the concepts demonstrated by \emph{Rewinder}; 2) investigate the power of using multiple debris structures; and 3) examine how \emph{Rewinder} might work with less accurate measurements or missing dimensions.
+provides strong motivation for further theoretical work to: 1) develop a robust generative model that utilizes the concepts demonstrated by \emph{Rewinder}; 2) investigate the power of using multiple debris structures; and 3) examine how \emph{Rewinder} might work with less accurate measurements or missing dimensions.
 
 Our results also motivate an observational campaign with {\it Spitzer} to survey RR Lyrae stars in debris structures around the Milky Way to get precise distances to combine with near-future {\it Gaia} velocity data.
-If just 100 stars in a single stellar stream allow us to measure the depth, shape, and orientation of the Milky Way potential, larger samples in multiple structures (e.g., the Orphan Stream, already mapped in RR Lyrae by \citet{sesar13}) offer the prospect of assessing these quantities as a function of Galactocentric radius.
+If just 100 stars in a single stellar stream allow us to study the depth, shape, and orientation of the Milky Way potential, larger samples in multiple structures \citep[e.g., the Orphan Stream;][]{sesar13} offer the prospect of assessing these quantities as a function of Galactocentric radius.
 Tracing the mass in a dark matter halo with this level of detail is impossible for any other galaxy in the Universe.
@@
 We thank Barry Madore for providing the inspiration for this work. Thanks to 
-David Hogg for statistical advice and many useful discussions. Thanks also to
-Steve Majewski, David Law, and David Nidever.
+David Hogg for statistical advice. Thanks also to Steve Majewski, David 
+Law, and David Nidever. We also thank the anonymous referee for useful 
+suggestions.
 

### 2013-10-13 streams/inference/back_integrate.py
empty objective function

@@ -19,6 +19,8 @@ from ..integrate.satellite_particles import satellite_particles_integrate
 
 __all__ = ["back_integrate_likelihood", "variance_likelihood"]
 
+def objective():
+    pass
 
 def back_integrate_likelihood(p, param_names, satellite, particles, 
                               Potential, t1, t2):

### 2013-10-13 text/notes/likelihood.tex
add note describing likelihood

\documentclass[letterpaper,12pt,preprint]{aastex}

% packages
\usepackage{amssymb,amsmath}

% commands
\newcommand{\given}{\,|\,}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\transpose}[1]{{#1}^{\mathsf{T}}}
\newcommand{\inverse}[1]{{#1}^{-1}}

\begin{document}

\title{Back-integration likelihood}
\author{Adrian M. Price-Whelan}

Starting with notation, take ${\bf x}_i$ to be the true 6D ($x$, $y$, and $z$ positions along with velocity components in the same coordinate system) of the star $i$. Take ${\bf D}_i$ to be the observed 6D vector -- that is, sky coordinates, distance, proper motion, and radial velocity. For now, we assume the star was drawn at some time $t_i$ from a spherical Gaussian (progenitor) described by a single spatial scale (the tidal radius, $r_{\mathrm{tide}}$) and a single velocity scale (the velocity dispersion, $\sigma_\mathrm{v}$). These scales are specified in the covariance matrix ${\bf \Sigma}_p$, where the subscript $p$ refers to the progenitor.

\begin{align}
	p({\bf D}_i, {\bf x}_i^{(0)} \given {\bf x}_p^{(0)}, {\bf \Sigma}_p, \Phi, t_i) &= 
	p({\bf D}_i | {\bf x}_i^{(0)}) \mathcal{N}({\bf x}_i(t_i\given {\bf x}_i^{(0)}, \Phi ) \given {\bf x}_p(t_i\given {\bf x}_p^{(0)}, \Phi ), {\bf \Sigma}_p)\\
	p({\bf D}_i \given {\bf x}_p^{(0)}, {\bf \Sigma}_p, \Phi, t_i) &= \int p({\bf D}_i, {\bf x}_i^{(0)} \given {\bf x}_p^{(0)}, {\bf \Sigma}_p, \Phi, t_i) \dd {\bf x}_i^{(0)}
\end{align}

\section{Tests}
Try for a single star, relative to a perfectly known progenitor. Increase the observational errors on the star, make sure the likelihood gets broader.

\end{document}

### 2013-10-14 text/notes/likelihood.tex
updated note on likelihood

@@
 
+I've already implemented the case assuming we have perfect data -- that looks like this:
+\begin{align}
+	p({\bf x}_i^{(0)} \given {\bf x}_p^{(0)}, {\bf \Sigma}_p, \Phi, t_i) &= 
+        \mathcal{N}({\bf x}_i(t_i\given {\bf x}_i^{(0)}, \Phi ) \given {\bf x}_p(t_i\given {\bf x}_p^{(0)}, \Phi ), {\bf \Sigma}_p)\\
+	p({\bf x}_i^{(0)} \given {\bf x}_p^{(0)}, {\bf \Sigma}_p, \Phi) &= 
+	\int p({\bf x}_i^{(0)} \given {\bf x}_p^{(0)}, {\bf \Sigma}_p, \Phi, t_i) \dd t_i\\
+\end{align}
+then if we assume we know the position and shape of the progenitor exactly, it becomes a marginal likelihood I can handle:
+\begin{align}
+	p({\bf x}_i^{(0)} \given \Phi, t_i) &= 
+        \mathcal{N}({\bf x}_i(t_i\given {\bf x}_i^{(0)}, \Phi ) \given {\bf x}_p(t_i\given {\bf x}_p^{(0)}, \Phi ), {\bf \Sigma}_p)\\
+	p({\bf x}_i^{(0)} \given \Phi) &= 
+	\int p({\bf x}_i^{(0)} \given \Phi, t_i) \dd t_i\\
+\end{align}
+which, numerically, is just:
+\begin{align}
+	p({\bf x}_i^{(0)} \given \Phi) &= 
+	\sum_k \mathcal{N}({\bf x}_i(t_i\given {\bf x}_i^{(0)}, \Phi ) \given {\bf x}_p(t_i\given {\bf x}_p^{(0)}, \Phi ), {\bf \Sigma}_p) \Delta t_{i,k}\\
+\end{align}
+
+Now, I want to include the observational errors directly in this model. I think that looks like this:
 \begin{align}
@@
 	p({\bf D}_i | {\bf x}_i^{(0)}) \mathcal{N}({\bf x}_i(t_i\given {\bf x}_i^{(0)}, \Phi ) \given {\bf x}_p(t_i\given {\bf x}_p^{(0)}, \Phi ), {\bf \Sigma}_p)\\
-	p({\bf D}_i \given {\bf x}_p^{(0)}, {\bf \Sigma}_p, \Phi, t_i) &= \int p({\bf D}_i, {\bf x}_i^{(0)} \given {\bf x}_p^{(0)}, {\bf \Sigma}_p, \Phi, t_i) \dd {\bf x}_i^{(0)}
+	p({\bf D}_i \given {\bf x}_p^{(0)}, {\bf \Sigma}_p, \Phi, t_i) &= \int p({\bf D}_i, {\bf x}_i^{(0)} \given {\bf x}_p^{(0)}, {\bf \Sigma}_p, \Phi, t_i)p({\bf D}_i | {\bf x}_i^{(0)}) \dd {\bf x}_i^{(0)}
 \end{align}
+again, assuming we know the progenitor shape and position today perfectly well, then this is:
+\begin{align}
+	p({\bf D}_i \given \Phi) &= \iint p({\bf D}_i, {\bf x}_i^{(0)} \given {\bf x}_p^{(0)}, {\bf \Sigma}_p, \Phi, t_i)p({\bf D}_i | {\bf x}_i^{(0)}) \dd {\bf x}_i^{(0)} \dd t_i\\
+\end{align}
+all along assuming a uniform prior on $t_i$.
 
-\section{Tests}
-Try for a single star, relative to a perfectly known progenitor. Increase the observational errors on the star, make sure the likelihood gets broader.
+But I've totally confused myself on how to implement marginalizing over the error distributions...
 

### 2013-10-14 streams/io/sgr.py
add tub as attribute to particle class

@@ -67,8 +67,11 @@ def mass_selector(m):
         pc = table_to_particles(data, usys,
                                 position_columns=["x","y","z"],
                                 velocity_columns=["vx","vy","vz"])
-        
-        return pc.to(u_galactic)
+        pc = pc.to(u_galactic)
+
+        t_unit = filter(lambda x: x.is_equivalent(u.s), usys)[0]
+        pc.tub = (np.array(data["tub"])*t_unit).to(u.Myr).value
+        return pc
     
     def satellite_orbit():
         """ Read in the position and velocity of the Orphan satellite center 

### 2013-10-16 text/paper1/ms.tex
removed video footnotes

@@
 \emph{backwards} in a a potential that accurately models the Milky Way, the stars
-should recombine with the progenitor (imagine watching satellite destruction in ``rewind").\footnote{\url{http://dx.doi.org/10.6084/m9.figshare.767305}} If the potential is incorrect,
+should recombine with the progenitor (imagine watching satellite destruction in ``rewind"). If the potential is incorrect,
 the orbits of the stars will diverge from that of the progenitor and
-thus will not be recaptured by the satellite system (Figure~\ref{fig:ps_distance}).\footnote{\url{http://dx.doi.org/10.6084/m9.figshare.767306}}
+thus will not be recaptured by the satellite system (Figure~\ref{fig:ps_distance}).
 

### 2013-10-16 scripts/figures/paper1_figures.py
fix yaxis label and scaling for fig 1 and 3

@@ -29,10 +29,8 @@ from streams.observation.gaia import parallax_error, proper_motion_error, \
 from streams.observation.rrlyrae import rrl_M_V, rrl_V_minus_I
 
 from streams.inference import relative_normalized_coordinates, generalized_variance, minimum_distance_matrix
-from streams.inference.lm10 import timestep
-from streams.potential import LawMajewski2010
-from streams.potential.lm10 import true_params, _true_params, param_to_latex
-from streams.integrate.satellite_particles import SatelliteParticleIntegrator
+from streams.potential.lm10 import LawMajewski2010, true_params, _true_params, param_to_latex
+from streams.integrate.satellite_particles import satellite_particles_integrate
 from streams.io.lm10 import particle_table, particles_today, satellite_today, time
 from streams.io import read_table
 from streams.plot import *
@@ -54,7 +52,8 @@ def normed_objective_plot(**kwargs):
     
     # Read in the LM10 data
     np.random.seed(142)
-    particles = particles_today(N=100, expr="(Pcol>-1) & (abs(Lmflag) == 1) & (Pcol < 7)")
+    Nparticles = 100
+    particles = particles_today(N=Nparticles, expr="(Pcol>-1) & (abs(Lmflag) == 1) & (Pcol < 7)")
     satellite = satellite_today()
     t1,t2 = time()
 
@@ -72,6 +71,7 @@ def normed_objective_plot(**kwargs):
     
     if not os.path.exists(data_file):
         variances = dict()
+        acc = np.zeros((Nparticles+1,3)) # placeholder
         for param in ['q1','qz','v_halo','phi']:
             if not variances.has_key(param):
                 variances[param] = []
@@ -84,8 +84,10 @@ def normed_objective_plot(**kwargs):
                 params = true_params.copy()
                 params[param] = stat
                 lm10 = LawMajewski2010(**params)
-                integrator = SatelliteParticleIntegrator(lm10, satellite, particles)
-                s_orbit,p_orbits = integrator.run(t1=t1, t2=t2, dt=-1.)
+                s_orbit,p_orbits = satellite_particles_integrate(satellite, particles,
+                                                     potential,
+                                                     potential_args=(Nparticles+1, acc), \
[…]

### 2013-10-18 scripts/back_integrate_movies.py
way old

@@ -158,10 +158,15 @@ def plot_xz_animation(potential, s, p, filename=""):
     ax.set_ylabel("$Z_{GC}$", color='w', fontsize=28, rotation='horizontal')
     fig.subplots_adjust(left=0.08, right=0.98, top=0.98, bottom=0.06)
     
+    D_ps_limit = 2.2
+    
+    _diff = sum(all_D_ps[0] < D_ps_limit)
+    _N = p.nparticles - _diff
+
     jj = 0
     for ii in range(0,len(t),10):
         D_ps = all_D_ps[ii]
-        idx = idx & (D_ps > 2.2)
+        idx = idx & (D_ps > D_ps_limit)
         
         offsets = p._r[ii]
         offsets[np.logical_not(idx)] = np.ones_like(offsets[np.logical_not(idx)])*len(p._r)
@@ -170,29 +175,36 @@ def plot_xz_animation(potential, s, p, filename=""):
         try:
             scatter_map[(1,0)].set_offsets(np.vstack((offsets[:,0],offsets[:,2])).T)
             scatter_map_sat[(1,0)].center = (sat_r[0], sat_r[2])
-            scatter_map_sat[(1,0)].set_radius(r_tide[ii,0,0].value)
+            scatter_map_sat[(1,0)].set_radius(r_tide[ii,0,0].value*2.)
         except NameError:
             scatter_map = dict()
             scatter_map_sat = dict()
             
-            c = Circle((sat_r[0], sat_r[2]), radius=r_tide[ii,0,0].value*1.4, 
+            c = Circle((sat_r[0], sat_r[2]), radius=r_tide[ii,0,0].value*2., 
                        facecolor='#CA0020', alpha=0.3, edgecolor='none')
             ax.add_patch(c)
             scatter_map_sat[(1,0)] = c
             
             scatter_map[(1,0)] = ax.scatter(offsets[:,0], offsets[:,2], 
-                                            marker='.', color='#ABD9E9', 
-                                            alpha=0.4, s=8)
+                                            marker='.', color='#7FCDBB', 
+                                            alpha=0.15, s=10)
             
         ax.set_xlim(-81,81)
         ax.set_ylim(-81,81)
         ax.set_xticklabels([])
         ax.set_yticklabels([])
+        xx = "{0:.2f}% recaptured".format((sum(~idx)-_diff) / _N * 100)
+        try:
+            recap_text.set_text(xx)
+        except NameError:
[…]

### 2013-10-18 text/paper1/ms.tex

@@
 Parallax distance errors from {\it Gaia} are larger than the line-of-sight size of both Sgr and Orphan (Orp), but photometric distance errors are comparable to the the Sgr scale (assuming 10\% errors, dotted line).
-Bottom panel shows that the {\it Gaia} tangential velocity errors are smaller than the internal velocity dispersion of nearer regions of both Sgr and Orp.\footnote{\url{http://www.rssd.esa.int/index.php?page=Science_Performance&project=GAIA}} }\label{fig:gaia_errors}
+Bottom panel shows that the {\it Gaia} tangential velocity errors are smaller than the internal velocity dispersion of nearer regions of both Sgr and Orp.
+}\label{fig:gaia_errors}
 \end{center}

### 2013-10-23 text/paper1/ms.tex
removed majewski 2004 reference, updated sesar 2013

@@
 
-\bibitem[{{Majewski} {et~al.}(2004){Majewski}, {Kunkel}, {Law}, {Patterson},
-  {Polak}, {Rocha-Pinto}, {Crane}, {Frinchaboy}, {Hummels}, {Johnston}, {Rhee},
-  {Skrutskie}, \& {Weinberg}}]{majewski04}
-{Majewski}, S.~R., {Kunkel}, W.~E., {Law}, D.~R., {et~al.} 2004, \aj, 128, 245
-
 \bibitem[{{Navarro} {et~al.}(1996){Navarro}, {Frenk}, \& {White}}]{navarro96}
@@
   {Kulkarni}, \& {Prince}}]{sesar13}
-{Sesar}, B., {Grillmair}, C.~J., {Cohen}, J.~G., {et~al.} 2013, ArXiv e-prints
+{Sesar}, B., {Grillmair}, C.~J., {Cohen}, J.~G., {et~al.} 2013, \apj, 776, 26
 

### 2013-10-27 misc/ldss_redux/cookbook.tex
add marlas reduction code to repo

\documentclass[11pt,letterpaper,dvips]{article}

\pretolerance=10000
\textwidth=7.0in
\textheight=10.0in
\voffset = -0.3in
\topmargin=0.0in
\headheight=0.00in
\hoffset = -0.3in
\headsep=0.00in
\oddsidemargin=0in
\evensidemargin=0in
\parindent=2em
\parskip=1.5ex

\newcommand{\nnn}{ccd\#\#\#}

\special{papersize=8.5in,11in}

\renewcommand{\theenumi}{\Roman{enumi}}

\begin{document}
%\pagestyle{empty}

\begin{center}
{{\Huge \bf LDSS3 Long-Slit Reductions}}\\
{{\Huge \bf [v1.0; March 2005]}}
\end{center}

\begin{enumerate}
{\Large \bf \item LDSS3 General Info}

This code reduces data from Magellan/LDSS3 and is written for
single-slit observations only.  Raw LDSS3 data consists of four fits
files corresponding to each of the four CCD chip amplifiers. Raw data
files are named ccd***c[1-4].fits.  The CCD chip layout is as follows:
c1 = lower left CCD quadrant, c2 = lower right, c3 = upper right, c4 =
upper left.  The code stitches the four quadrants together and
operates on combined frames for all reduction operations after bias
subtraction.  First time users should run the invidual steps listed
below, however the data can be reduced in a single step by running the
wrapper script xxxx.  Example fits files of intermediate processing
steps can be found in LDSS3/examples.  Comments and questions should
be sent to M. Geha at mgeha@ociw.edu.

{\Large  \bf \item LDSS3 Suggested Long-Slit Calibrations}

   \begin{itemize}
	\item 10 bias (zero exposure) frames.
	\item 1 Arc at position of science frames.
	\item 2 QTZ Lamps at position of science frames (especially if working
	  redward of $6700\mbox{\AA}$).
	\item Bright imaging flats to determined gain factor.
	\item All calibrations should be taken with the same readout
	  speed (slow/fast/turbo) as the science data.

   \end{itemize}

{\Large \bf \item Required Software } 
  \begin{itemize}
     \item The LDSS3 code requires a copy of IDL version 5.4 or higher
      and relies on several IDL software libraries.  Download and
      installation information are available at the websites listed below.
	\item  IDLUTILS/IDLSPEC2d mantained by SDSS available at\\
	http://spectro.princeton.edu/idlspec2d\_install.html

	\item XIDL package mantained by JXP available at\\
	http://www.ucolick.org/$\sim$xavier/IDL/index.html
   \end{itemize}

\vskip 0.5cm
{\Large \bf \item Initial Setup (Repeat for each night) }

  \begin{itemize}

    \item Create a new directory for the night's data (eg. obs030205/).

    \item In the night directory, create a rawdata/ directory and put all
    the raw data in it.  The code assumes data is gziped and is named
    ccd***.fits.gz.

    \item Create a setcrds file in directory above rawdata/
      (eg. setcrds\_obs030205.pro).  This file sets which exposures
      are flats/arcs/science frames.  An initial guess at this
      information is generated by ldss\_setup.pro (see below) from
      information in the headers.  The setcrds routine explicitly sets
      which exposures are reduced and which calibrations are
      associated with a given science exposure.  This file is called
      by ldss\_setup.

    \item An example setcrds file is available in
      {\tt /examples/setcrds\_obs030205.pro}.  For a given science object,
      the following lines are required:\\

        \begin{itemize}
	\item Note the IDL routine {\bf fnd\_indx} converts
	the frame number of an image to the index in the structure.

        strct[fnd\_indx(strct,59):fnd\_indx(strct,67)].objname   = 'galaxy\_name'    ; name of galaxy\\
	strct[fnd\_indx(strct,59):fnd\_indx(strct,67)].objnum  = 1   ; unique ID for exposures associated with science target\\
	strct[fnd\_indx(strct,59):fnd\_indx(strct,67)].flg\_anly= 1   ; analyze frames\\
	strct[fnd\_indx(strct,63):fnd\_indx(strct,64)].type    = 'OBJ'~~~~; science frames\\
	strct[fnd\_indx(strct,65)].type = 'ARC'~~~~~~~~~~~~~~~~~~~~~~~~~~~; arc lamp\\
	strct[fnd\_indx(strct,66):fnd\_indx(strct,67)].type    = 'FLT'~~~~; flat field\\

	\item If strct[*].flg\_anly = 0 an exposure is not analyzed (eg. a bad flat field or alignment
	  exposures.)

        \end{itemize}

	 \item The routine ldss\_procnight runs all of the routine
	   below for a given night.  I suggest first running the
	   routines individually before using this routine.

     \end{itemize}

{\Large \bf \item Calibrations }

{\Large \bf Create LDSS3 Structure }

   \begin{itemize}
    \item Launch idl in the directory above rawdata/

    \item {\bf ldss\_setup} :: Create the ldss structure.  This routine
      organizes the night's data and is used in all subsequent processing steps.
      This routine creates several things:
         \begin{enumerate}
	   \item  Empty directories (Arcs/ Flats/ Final/) where intermediate
	     processing steps will be stored.
	   \item  An IDl structure in memory with whatever name you chose.
	   \item  An IDl structure written to a .fits file.
         \end{enumerate}
    \item This routine first creates the directories, then searches for data
      files in rawdata/.  Header information is used to create the ldss
      structure.  This information is then modified by calling the 
      setcrds file created above.

    \item See Section xx for the tags set in the ldss structure.

    \item This routine outputs a .fits copy of the ldss structure
      named ldss\_ + night + .fits (eg. ldss_obs030205.fits).  If you
      re-start IDL, you will need to read this structure into memory:
      IDL$>$ strct = mrdfits('ldss_obs030205.fits',1)

         \quad Example: IDL$> \;$ {\bf ldss\_setup}, 'obs030205', strct \\  
   \end{itemize}

{\Large \bf Create Bias Frame }

	\begin{itemize}
	  \item {\bf ldss\_mkbias} :: Make bias frame for each of 4
	    LDSS3 chips. Median combines all bias frames as defined in
	    the structure with strct.tpye = 'ZRO'

	  \item There is two-dimensional structure in the bias frame
	    that cannot be removed via overscan subtraction alone.

	  \item  This routine outputs four images in the Bias/biasc* directory.

	  \item To check the Bias frame : {\bf xatv}, 'Bias/Biasc*.fits'

         \quad Example: IDL$> \;$ {\bf ldss\_mkbias}, strct, /plot \\

	\end{itemize}

{\Large \bf Set Gain Factor }
     \begin{itemize}
       \item {\bf ldss\_gaincalc} :: Determine the gain factors
	 relaticve to chip 1.
       \item As of March 2005, the relative gain factors for the four LDSS3
	 was not constant between runs and depended on the readout
	 speed (eg. slow, fast, turbo).  As the instrument is used
	 more these values could probably be hardwired into the code.
	 For now, we need to determine this empirically and add them
	 to the ldss structure.

       \item In setcrds file, define a few bright imaging flat fields to determine
	 gain factor (must be same readout speed as science exposures).  These will
	 have stct.type = 'GFT'.

         \quad Example: IDL$> \;$ {\bf ldss\_gaincalc}, 'obs030205', strct, /plot \\

     \end{itemize}

{\Large \bf Create Rough Arc Solution }

	\begin{itemize}
	  \item {\bf ldss\_idarc} :: Create a rough wavelength
           solution based on a stacked arc image.  This solution will
           serve a starting guess for reducing arc frames associated with
           the science exposures.

	  \item This routine requires the file ldss\_henear.dat
	    included in the ldss tarfile.  The user is required to
	    identify the wavelength and corresponding pixel value for
	    3-4 arc lines as input to this routine

	  \item Writes a rough wavelength solution to Arc/arcsol.fits

         \quad Example: IDL$> \,$  hand_lambda = [4471.47, 5015.675, 5944.834, 6402.246]\\ 
	                IDL$> \,$  hand_pix    = [342,     1172,     2507,      3146]\\
			IDL$> \,$ {\bf ldss\_idarc}, strct, hand\_lambda= hand\_lambda, hand\_pix= hand\_pix, /plot \\

	\end{itemize}

{\Large \bf \item Process Science Frames }

	\begin{itemize}
	\item {\bf ldss\_procobj} :: This routine processes raw
	  science frames.  Science frames are bias subtracted,
	  overscan subtracted, flat fielded, flaged for cosmic rays
	  and sky subtracted.  At the moment, it is not possible to
	  turn off any of these steps, but this could be modified.

         \quad Example: IDL$> \;$ {\bf ldss\_procobj}, strct, objnum, imred \\

	 \item This routine calls the following subroutines:

	\item {\bf ldss\_stitch4} :: Bias subtract and stitch the four LDSS3
	  CCD chips together.
	    \begin{itemize}
	      \item This code can also be run as a stand alone routine
                 to stitch a single LDSS3 frame together.  Data
                 reduction proceeds on stitched frames
		 \quad Example: IDL$> \;$ {\bf ldss\_stitch4}, im, inroot='rawdata/ccd0030'
	    \end{itemize}

	  \begin{enumerate}
	  \item {\bf ldss\_mkflat} :: Process flat field images
	  associated with given science target. 
	    \begin{itemize}
	      \item This routine processes and combines the flat field
		images associated with the science exposure
		(preferably taken at the same sky position).  The
		combined flat field to normalized by fitting a
		polynomial (order = 7) to the spatially collapsed flat.
	      \item  Output .fits file Flats/flat\_[objnum].fits
	    \end{itemize}

	\item Create inverse varience image ivar = im * gain + RN$^2$

	\item Divide by normalized flat field and trim image (default
	trim section = [*,1300:2500])

	\item {\bf ldss\_mkmask} :: Mask regions of zero flux (intended
	  for the bridge regions of the LDSS3 longslits).

	\item {\bf ldss\_crflag} :: Flag cosmic rays, outputs indices
	  of all cr pixels which are to be set to zero (these pixels not used
	  in subsequent processing steps).
	    \begin{itemize}
	      \item Cosmic rays are identified as being sharper than
		the image PSF.  Based on the IDLUTILS routine {\bf
		reject\_cr}.
	    \end{itemize}

	\item {\bf ldss\_fixcol} :: Interpolate over bad columns are
	defined in the file ldss\_badcol.dat.

	\item{\bf ldss\_arcsol} :: Create a 2D wavelength array.
	    \begin{itemize}
	      \item Processes arc lamp image associated with science
                frame (preferably taken at same sky position).  For
                each row, solve for wavelength solution.  Initial fit
                taken from results of ldss\_idarc (see above).
	      \item For each row of arc
	      \item Writes to Arc/arc\_[objnum].fits
	    \end{itemize}

	\item {\bf ldss\_skysub} :: Sky subtract single image via
	bspline fitting.
	    \begin{itemize}
	      \item Fits a bspline to the collapsed sky region (sky can
               be defined in ldss\_procobj, default sky = [*,0:199] and
               [*,900:*]).  Based on the IDLUTILS routine {\bf bspline\_iterfit}.
	    \end{itemize}

	\item Write reduced image to Final/f\_ccd**.fits.  The final images
	  are .fits files with three image planes/axes.  These are
	        img  = mrdfits('Final/f\_ccd*.fits', 1)~~~ reduced image  
	        ivar = mrdfits('Final/f\_ccd*.fits', 2)~~~ inverse varience  
	        wave = mrdfits('Final/f\_ccd*.fits', 3)~~~ wavelength at each pixel

        \end{enumerate}
	\end{itemize}

{\Large \bf \item Combine and Extract Science Spectra }

    \begin{itemize}
	\item {\bf ldss\_combspec} :: This routine combines multiple object
	structures (primarily their spectra) into one Final structure. 
	For objects with multiple
	spectra, the spectra are matched to the same flux and summed.  At
	the moment this routine only handles two spectra at a time.  \\
	\item Output :: A structure containg all of the fluxed, coadded spectra
         \quad Example: IDL$> \;$ {\bf ldss\_combspec}, strct, objnum, imcomb\\
  \end{itemize}

\clearpage
{\Large \bf \item LDSS3 Reduction Structure }
	\begin{center}
	{\Large {\bf ldssstr}}
	\end{center}

	{\small
	\begin{tabular}{lcl}
	  \hline
	  Tag & Type & Comment \\
	  \hline
         ccdframe   & ' '   & CCD FRAME name (in /rawdata) \\
         ccdnum     & 0     & CCD FRAME number\\
         flg\_anly & 0      &  Analysis flag 0=Don't Analyse \\
         type      & ' '    & ObjTyp : OBJ, STD, DRK, ZRO, FLT, ARC \\
	 objname   & ' '    & Science object name \\
	 objnum    & ' '    & Science object number (unique for a objects in a given night)\\
         exp       & 0.d    & Exposure time \\
         dateobs   & 0.0d   & Date of Obs \\  
         UT        & ' '    & UT \\
         RA        & 0.0    & RA \\
         DEC       & 0.0    & DEC \\
         SRA       & ' '    & RA \\
         SDEC      & ' '    & DEC \\
         Equinox   & 0.     & EQUINOX \\
	 airmass   & 0.0    & airmass\\
	 rotang    & 0.0    & Rotation angle\\
         grism     & ' '    & Grism : VPHB, VPHR, MEDR \\
         slitwidth & 0      & Aperture position \\
         gain     & fltarr(4) & Gain for each chip\\
         readno   & fltarr(4) & Read Noise for each chip \\

	  \hline
	\end{tabular}
	}

\end{enumerate}

\hline
This software was written and is mantained by M.Geha (OCIW).  Please
send questions, software bugs or comments to mgeha@ociw.edu.

\end{document}

### 2013-11-04 text/notes/spitzer_target_selection/spitzer_sgr.tex
move to subdir

\documentclass[letterpaper,12pt,preprint]{aastex}

% packages
\usepackage{amssymb,amsmath}

\begin{document}

\title{Spitzer Sgr target selection}
\author{Adrian M. Price-Whelan}

In this document, I'll explain how we selecting RR Lyrae (RRL) targets associated with the Sagittarius stream for Spitzer observation. Because the stream extends out beyond 60~kpc, many of these targets are quite expensive in terms of required Spitzer integration time. We've decided to only select RRLs that match the stream in both distance and radial velocity (at a given phase along the orbit), as compared to the Law \& Majewski 2010 simulation of the stream. The Catalina Sky Survey (CSS) recently published a catalog of $\sim$12,000 type-ab RRL stars identified from time-domain photometry, along with $\sim$1500 radial velocities. We use this catalog as our sample.

As an initial cut, we use the Sgr coordinate system from Majewski et al. (2003) to select all CSS RRLs within 20~kpc of the $B=0$ plane, and a Galactocentric distance $D_{\rm gc} > 15~{\rm kpc}$ (the inner halo is a mess). Figure~\ref{fig:css_all} shows (in Galactocentric cartesian coordinates) a) the distribution of all CSS RRLs, b) all RRLs that match the above cut, and c) the LM10 particles in the same coordinates. The leading arm is distinctly visible in the northern CSS data, and hints of the trailing arm are seen in the south. Note that there appears to be some offset between the CSS leading arm and the simulation --- this could be physical, but could also be an error in their distance calibration. I've also included a figure (Fig.~\ref{fig:css_vgsr}) that shows the radial velocities against Sgr longitude ($\Lambda$) for LM10 particles (grey) and the CSS RRLs (red), but no clear overdensities are visible.

Next we take 30 bins in $\Lambda$ and compute the median and $\pm$3-$\sigma$ distances for each bin from the LM10 particle data for the leading and trailing arm separately. We then find any star that could feasibly be within $\pm$3-$\sigma$ of the median distance to each arm, taking into account the distance errors. Figure~\ref{fig:lm10_selection} shows these (soft) margins overlaid with the LM10 particles, while figure~\ref{fig:css_selection} shows the CSS RRLs with RVs selected with this method. We ignore trailing arm data with $\Lambda>180$ since Belokurov et al. 2013 show with BHB stars that the trailing arm may not match the LM10 model past apocenter.

We then do the same thing with $v_{\rm gsr}$ by taking 30 bins in $\Lambda$ and repeating the above algorithm. Figure~\ref{fig:lm10_vgsr} shows the median and $\pm$3-$\sigma$ velocities for these bins, and figure~\ref{fig:css_vgsr} shows the CSS RRLs selected with this window. The remaining sample consists of 193 RR Lyrae, with a total required integration time of $\sim$1000 hr --- way over the budgeted time. We 1) want to get precise distances along many phases of the stream and 2) don't want to spend all of our time on distant targets which require the most exposure time. To select a subset, we take 5 stars at random from 15 degree bins in $\Lambda$. The final sample is shown in red in Figure~\ref{fig:final_sample}, with the other likely stream members in black, overplotted on the LM10 particles.

\begin{figure}
\begin{center}
\includegraphics[width=0.9\textwidth]{catalina_all.pdf}
\caption{ From left to right: a) All CSS RRLs, b) CSS RRLs $<$ 20 kpc from the midplane of the Sgr coordinate system, and c) LM10 simulation particles. }\label{fig:css_all}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=0.75\textwidth]{catalina_all_vgsr.pdf}
\caption{ Galactocentric radial velocity for LM10 particles (grey) and CSS RRLs (red) for both the leading (left) and trailing (right) arms of Sgr debris. }\label{fig:css_vgsr}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{lm10_selection.pdf}
\caption{ LM10 particles from each Sgr arm, with median distance (red) and $\pm$3-$\sigma$ (green) distance values for 30 bins in $\Lambda$. }\label{fig:lm10_selection}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{css_selection_vgsr.pdf}
\caption{ CSS RRLs with measured radial velocities selected to be within $\pm$3-$\sigma$ + distance error from the median LM10 distance in each bin from fig.~\ref{fig:lm10_selection}. }\label{fig:css_selection}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{lm10_vgsr_selection.pdf}
\caption{ LM10 particles from each Sgr arm, with median radial velocity (red) and $\pm$3-$\sigma$ (green) values for 30 bins in $\Lambda$. }\label{fig:lm10_vgsr}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{css_selection_from_vgsr.pdf}
\caption{ CSS RRLs selected to be within $\pm$3-$\sigma$ from the median LM10 radial velocity in each bin from fig.~\ref{fig:lm10_vgsr}. }\label{fig:css_vgsr}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{final_sample.pdf}
\caption{ Final sample of CSS RRLs selected to be observed with Spitzer (red) and other RRLs that meet the criteria, but are unable to observe due to time constraints. Concentric circles with labels show required total integration time (over 12 visits) for a typical RR Lyrae at that distance. }\label{fig:final_sample}
\end{center}
\end{figure}

\end{document}

### 2013-11-08 text/notes/spitzer_target_selection/spitzer_sgr.tex
target selection

@@
 
+\section{An overview of the sample}
+\begin{itemize}
+	\item Three clumps (29 stars total) along the leading tail (sampling in orbital phase);
+	\item a clump of 8 stars selected to be associated with the bifurcated part of the stream;
+	\item a clump of 10 stars selected to be associated with the nearby part of the trailing arm;
+	\item 16 stars selected from the trailing arm with $\Lambda$ $<$ 180.
+\end{itemize}
+Though the bifurcated stream targets and nearby trailing targets are selected based on distance, radial velocity, and $B$, it is possible that they will not reveal anything (still interesting) -- these two components only account for $\sim$15\% of the time, so we think this is acceptable.
+
+\section{Details about target selection}
+
 As an initial cut, we use the Sgr coordinate system from Majewski et al.
@@
 We then find any star that could feasibly be within $\pm$3-$\sigma$ of the median distance to each arm, taking into account the distance errors.
-Figure~\ref{fig:lm10_selection} shows these (soft) margins overlaid with the LM10 particles, while figure~\ref{fig:css_selection} shows the CSS RRLs with RVs selected with this method.
-We ignore trailing arm data with $\Lambda>180$ since Belokurov et al.
-2013 show with BHB stars that the trailing arm may not match the LM10 model past apocenter.
+Figure~\ref{fig:dist_selection} shows these margins overlaid with the CSS RRLs selected from both the leading and trailing arm to match in distance.
+We manually offset the leading arm by $-$5~kpc where necessary to account for the apparent shift in distance in the CSS overdensity.
+We then do the same thing with $v_{\rm gsr}$ by taking 30 bins in $\Lambda$ and repeating the above algorithm.
+Figure~\ref{fig:vgsr_selection} shows the CSS RRLs selected with this window.
 
-We then do the same thing with $v_{\rm gsr}$ by taking 30 bins in $\Lambda$ and repeating the above algorithm.
-Figure~\ref{fig:lm10_vgsr} shows the median and $\pm$3-$\sigma$ velocities for these bins, and figure~\ref{fig:css_vgsr} shows the CSS RRLs selected with this window.
-The remaining sample consists of 193 RR Lyrae, with a total required integration time of $\sim$1000 hr --- way over the budgeted time.
-We 1) want to get precise distances along many phases of the stream and 2) don't want to spend all of our time on distant targets which require the most exposure time.
-To select a subset, we take 5 stars at random from 15 degree bins in $\Lambda$.
-The final sample is shown in red in Figure~\ref{fig:final_sample}, with the other likely stream members in black, overplotted on the LM10 particles.
+Figure~\ref{fig:koposov} (Figure~1 from Koposov et al.
+(2012)) shows the bifurcated part of the stream and main overdensity observed in SDSS and how they each vary in $\Lambda$ vs.
+$B$.
+For $180 < \Lambda < 300$, the main leading arm is actually below $B=0$ and the bifurcation is at or slightly above the midplane.
+The opposite is true for $\Lambda < 180$.
+Given this, we only select leading arm targets with $B<0$ for large $\Lambda$, and $B<5$ elsewhere.
+Figure~\ref{fig:LB} shows in green the targets selected to be associated with the bifurcation, and in blue all others from the far wraps.
+
+The remaining sample consists of 66 RR Lyrae, with a total required integration time of $\sim$280 hr and is shown in Figure~\ref{fig:final_sample} (green), along with other RR Lyrae that meet the above cuts but are not in the chosen clumps (red), overlaid with the LM10 particles.
 
@@
 \begin{center}
-\includegraphics[width=\textwidth]{lm10_selection.pdf}
-\caption{ LM10 particles from each Sgr arm, with median distance (red) and $\pm$3-$\sigma$ (green) distance values for 30 bins in $\Lambda$. }\label{fig:lm10_selection}
+\includegraphics[width=\textwidth]{dist_selection.pdf}
+\caption{  CSS RRLs selected to be within $\pm$3-$\sigma$ from the median LM10 distance in each bin. }\label{fig:dist_selection}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\textwidth]{css_selection_vgsr.pdf}
-\caption{ CSS RRLs with measured radial velocities selected to be within $\pm$3-$\sigma$ + distance error from the median LM10 distance in each bin from fig.~\ref{fig:lm10_selection}.
-}\label{fig:css_selection}
+\includegraphics[width=\textwidth]{vgsr_selection.pdf}
+\caption{ CSS RRLs selected to be within $\pm$3-$\sigma$ from the median LM10 radial velocity in each bin. }\label{fig:vgsr_selection}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\textwidth]{lm10_vgsr_selection.pdf}
-\caption{ LM10 particles from each Sgr arm, with median radial velocity (red) and $\pm$3-$\sigma$ (green) values for 30 bins in $\Lambda$. }\label{fig:lm10_vgsr}
+\includegraphics[width=\textwidth]{koposov.png}
+\caption{ Figure~1 from Koposov et al. (2012), showing Sgr in the SDSS footprint. The bifurcation is near $B=0$ for $\Lambda>180$ and near $B=10$ for $\Lambda<180$. }\label{fig:koposov}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\textwidth]{css_selection_from_vgsr.pdf}
-\caption{ CSS RRLs selected to be within $\pm$3-$\sigma$ from the median LM10 radial velocity in each bin from fig.~\ref{fig:lm10_vgsr}. }\label{fig:css_vgsr}
+\includegraphics[width=\textwidth]{L_vs_B.pdf}
+\caption{ $\Lambda$ vs.
+$B$ for LM10 particles (black) and selected particles (excluding the nearby trailing stars, for now).
+Blue are selected to not be associated with the bifurcation, green are the clump selected to be associated with the bifurcation.
+}\label{fig:LB}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\textwidth]{final_sample.pdf}
-\caption{ Final sample of CSS RRLs selected to be observed with Spitzer (red) and other RRLs that meet the criteria, but are unable to observe due to time constraints.
-Concentric circles with labels show required total integration time (over 12 visits) for a typical RR Lyrae at that distance.
+\includegraphics[width=\textwidth]{with_near_trailing_xz.pdf}
+\caption{ Final sample of CSS RRLs selected to be observed with Spitzer (green) and other RRLs that meet the criteria, but are unable to observe due to time constraints (red).
 }\label{fig:final_sample}
+%Concentric circles with labels show required total integration time (over 12 visits) for a typical RR Lyrae at that distance.
 \end{center}

### 2013-11-08 scripts/spitzer_targets.py

@@ -349,7 +349,7 @@ def sgr(overwrite=False, seed=42):
 
     if not os.path.exists(_selection_cache):
         lmflag_dist_idx = sgr_dist(sgr_catalina_rv, lm10,
-                                 Nbins=40, sigma_cut=2.3)
+                                 Nbins=40, sigma_cut=2.5)
         fnpickle(lmflag_dist_idx, _selection_cache)
     else:
         lmflag_dist_idx = fnunpickle(_selection_cache)
@@ -399,7 +399,7 @@ def sgr(overwrite=False, seed=42):
 
     # deselect stars possibly associated with the bifurcation
     no_bif = (L > 180) & (L < 360) & (B > -40) & (B < 0)
-    no_bif |= ((L <= 180) & (B < 0))
+    no_bif |= ((L <= 180) & (B < 0) & (L > 80))
 
     # draw a box around some possible bifurcation members
     bifurcation_box  = (L > 210) & (L < 225) & (B > 0) & (B < 15) # OR
@@ -426,7 +426,11 @@ def sgr(overwrite=False, seed=42):
         this_ix = lead_ix & (clump_dist <= xxx)
         ix_lead_clumps |= this_ix #select_only(this_ix, 10)
     ix_lead_clumps &= lead_ix
-    ix_lead_clumps |= ((L > 80) & (L < 180) & lead_ix) # all southern ones
+
+    # all southern leading
+    lll = ((L > 80) & (L < 180) & lead_ix)
+    print("southern leading", sum(lll))
+    ix_lead_clumps |= lll
 
     # select all trailing stuff in south > 90
     # ix_trail_clumps_with = np.zeros_like(trail_ix).astype(bool)
@@ -451,7 +455,11 @@ def sgr(overwrite=False, seed=42):
         this_ix = trail_ix_with & (clump_dist <= xxx)
         ix_trail_clumps_with |= this_ix #select_only(this_ix, 10)
     ix_trail_clumps_with &= trail_ix_with
-    ix_trail_clumps_with |= (L > 90) & (L < 180) & (trail_ix_with)
+
+    # all trailing southern
+    ttt = (L > 90) & (L < 180) & (trail_ix_with)
+    print("southern leading", sum(ttt))
+    ix_trail_clumps_with |= ttt
 
     i1 = integration_time(sgr_catalina_rv[ix_bif]["dist"])
     i2 = integration_time(sgr_catalina_rv[ix_lead_clumps]["dist"])
[…]

### 2013-11-09 text/notes/spitzer_target_selection/spitzer_sgr.tex
fixed koposovs flipped beta

@@
 \begin{itemize}
-	\item Three clumps (29 stars total) along the leading tail (sampling in orbital phase);
+	\item Three clumps (31 stars total) along the leading tail (sampling in orbital phase);
 	\item a clump of 8 stars selected to be associated with the bifurcated part of the stream;
-	\item a clump of 10 stars selected to be associated with the nearby part of the trailing arm;
+	\item a clump of   stars selected to be associated with the nearby part of the trailing arm;
 	\item 16 stars selected from the trailing arm with $\Lambda$ $<$ 180.
@@
 
-The remaining sample consists of 66 RR Lyrae, with a total required integration time of $\sim$280 hr and is shown in Figure~\ref{fig:final_sample} (green), along with other RR Lyrae that meet the above cuts but are not in the chosen clumps (red), overlaid with the LM10 particles.
+The remaining sample consists of 65 RR Lyrae, with a total required integration time of $\sim$280 hr and is shown in Figure~\ref{fig:final_sample} (green), along with other RR Lyrae that meet the above cuts but are not in the chosen clumps (red), overlaid with the LM10 particles.
 
@@
 \begin{center}
-\includegraphics[width=\textwidth]{L_vs_B.pdf}
+\includegraphics[width=\textwidth]{without_near_trailing_LB.pdf}
 \caption{ $\Lambda$ vs.

### 2013-11-09 scripts/spitzer_targets.py

@@ -389,7 +389,7 @@ def sgr(overwrite=False, seed=42):
 
     lead_ix = lmflag_dist_idx[1] & lmflag_rv_idx[1]
     trail_ix = lmflag_dist_idx[-1] & lmflag_rv_idx[-1]
-    trail_ix[L < 180] &= B[L < 180] < 5
+    trail_ix[L < 180] &= B[L < 180] > -5
 
     trail_ix &= np.logical_not((L > 50) & (L < 100) & (sgr_catalina_rv["dist"] < 40))
 
@@ -398,11 +398,11 @@ def sgr(overwrite=False, seed=42):
     trail_ix_with = trail_ix & ( ((L > 230) & (L < 315)) | (L < 180) )
 
     # deselect stars possibly associated with the bifurcation
-    no_bif = (L > 180) & (L < 360) & (B > -40) & (B < 0)
-    no_bif |= ((L <= 180) & (B < 0) & (L > 80))
+    no_bif = (L > 180) & (L < 360) & (B < 15) & (B > 0)
+    no_bif |= ((L <= 180) & (B < 10) & (B > -5) & (L > 80))
 
     # draw a box around some possible bifurcation members
-    bifurcation_box  = (L > 210) & (L < 225) & (B > 0) & (B < 15) # OR
+    bifurcation_box  = (L > 205) & (L < 225) & (B < 2) & (B > -10) # OR
     print(sum(bifurcation_box), "bifurcation stars")
     print(sum(lead_ix), "leading arm stars")
     print(sum(trail_ix_with), "trailing arm stars (with nearby)")
@@ -467,9 +467,9 @@ def sgr(overwrite=False, seed=42):
     i3_with = integration_time(sgr_catalina_rv[ix_trail_clumps_with]["dist"])
     i3_without = integration_time(sgr_catalina_rv[ix_trail_clumps_without]["dist"])
 
-    print(len(sgr_catalina_rv[ix_bif]))
-    print(len(sgr_catalina_rv[ix_lead_clumps]))
-    print(len(sgr_catalina_rv[ix_trail_clumps_with]))
+    print("final num bifurcation", len(sgr_catalina_rv[ix_bif]))
+    print("final num leading arm", len(sgr_catalina_rv[ix_lead_clumps]))
+    print("final num trailing arm", len(sgr_catalina_rv[ix_trail_clumps_with]))
     targets_with = vstack((sgr_catalina_rv[ix_bif],
[…]

### 2013-12-19 text/notes/likelihood/likelihood.tex
note on likelihood

\documentclass[letterpaper,12pt,preprint]{aastex}

% packages
\usepackage{amssymb,amsmath, amsbsy}

% commands
\newcommand{\given}{\,|\,}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\transpose}[1]{{#1}^{\mathsf{T}}}
\newcommand{\inverse}[1]{{#1}^{-1}}

% bold X and D
\newcommand{\D}{{\bf D}}
\newcommand{\X}{{\bf X}}
\newcommand{\bSigma}{{\bf \Sigma}}
\newcommand{\bsigma}{\boldsymbol\sigma}

\begin{document}

\title{Back-integration likelihood}
\author{Adrian M. Price-Whelan}

Starting with notation, take ${\bf D}^{(i)}$ to be the observed heliocentric 6D position of the $i$th star -- e.g., the measured $l$, $b$, $D$, $\mu_l$, $\mu_b$, $v_r$ -- and ${\bf X}^{(i)}$ to be the corresponding \emph{true} 6D position of the same star. We assume that this star was once part of a progenitor system -- e.g., satellite galaxy -- presently at observed position ${\bf D}_s$ (with true position ${\bf X}_s$) that is tidally disrupting and forming a cold debris structure in the potential, $\Phi$, of some parent galaxy. We then model the measured position of a star today as the result of sampling from a time-dependent 6D distribution -- the progenitor system -- at some time $\tau^{(i)}$ in the past, and evolving that sampled 6D position to present day in the parent potential, neglecting the effect of self-interaction with the halo of the progenitor. We model the progenitor as a 6D Gaussian with mean vector given by the position of the satellite at any given time, ${\bf X}_s(t)$, and with shape defined by a time-dependent covariance matrix $\bSigma_s(t)$ (the subscript $s$ refers to satellite or progenitor). With these assumptions we can write the posterior for a single star as:
\begin{align}
	p(\Phi, \X^{(i)}, \X_s, \bSigma_s, \tau^{(i)} \given \D^{(i)}, \D_s) &\propto 
		p(\D^{(i)} \given \X^{(i)}) p(\D_s \given \X_s) p(\X^{(i)} \given \X_s, \bSigma_s, \tau^{(i)})\\
	p(\D^{(i)} \given \X^{(i)}) &= \mathcal{N}(\X^{(i)} \given \D^{(i)}, \bsigma^{(i)})\\
	p(\D_s \given \X_s) &= \mathcal{N}(\X_s \given \D_s, \bsigma_s)\\
	p(\X^{(i)} \given \X_s, \bSigma_s, \tau^{(i)}) &= \mathcal{N}(\widetilde{\X}^{(i)}(\tau^{(i)}) \given \widetilde{\X}_s(\tau^{(i)}), \bSigma_s(\tau^{(i)}))
\end{align}
where the vectors $\bsigma^{(i)}$ and $\bsigma_s$ represent the (assumed) Gaussian, uncorrelated observational errors on the observed 6D position of star $i$ and the satellite, respectively, and the tildes denote a conversion from spherical, heliocentric to cartesian, Galactocentric coordinates. 
%Thus, 
%\begin{align}
%	p(\Phi, \X, \X_s, \bSigma_s, \tau \given \D, \D_s) &= \prod_i p(\Phi, \X^{(i)}, \X_s, \bSigma_s, \tau^{(i)} \given \D^{(i)}, \D_s)\\
%	p &= \prod_i p^{(i)}\\
%	\ln p &= \sum_i \ln p^{(i)}\\
%	\ln p = \sum_i [\ln\mathcal{N}(\X^{(i)} \given \D^{(i)}, \bsigma^{(i)}) &+ \ln\mathcal{N}(\X_s \given \D_s, \bsigma_s) + \ln\mathcal{N}(\X^{(i)}(\tau^{(i)}) \given \X_s(\tau^{(i)}), \bSigma_s(\tau^{(i)}))]
%\end{align}

We first consider the case of a spherical Gaussian progenitor described by a single (time-dependent) spatial scale -- the tidal radius, $r_{\mathrm{tide}}(t)$ -- and a single velocity scale -- the velocity dispersion, $\sigma_\mathrm{v}$, with a \emph{perfectly} observed position. Under these assumptions, the posterior can be written
\begin{align}
	p(\Phi, \X^{(i)}, \tau^{(i)} \given \D^{(i)}) &\propto 
		p(\D^{(i)} \given \X^{(i)}) p(\X^{(i)} \given \X_s, \bSigma_s, \tau^{(i)})\\
\end{align}
and thus the $\ln$-posterior is
\begin{align}
	\ln p(\Phi, \X^{(i)}, \tau^{(i)} \given \D^{(i)}) &= \ln p(\D^{(i)} \given \X^{(i)}) + \ln p(\X^{(i)} \given \X_s, \bSigma_s, \tau^{(i)}).
\end{align}
The first term on the RHS -- from the observed positions of the stars	-- can be expanded to read
\begin{align}
	\ln p(\D^{(i)} \given \X^{(i)}) &= K - \frac{1}{2}\sum_j \left[ 2\ln\sigma_j + \left(\frac{X^{(i)}_j-D^{(i)}_j}{\sigma^{(i)}_j}\right)^2\right].
\end{align}
The second term -- quantifying how well the potential ``returns'' the stars to the progenitor -- reads as follows for the case of a spherical progenitor:
\begin{align}
	\ln p(\X^{(i)} \given \X_s, \bSigma_s, \tau^{(i)}) &= M - \frac{1}{2}\ln(\left\vert\bSigma_s(\tau^{(i)})\right\vert) - \frac{1}{2}\langle \widetilde{\X}-\widetilde{\X}_s, \bSigma_s^{-1}(\widetilde{\X}-\widetilde{\X}_s) \rangle\\
	&= \left[M - 3(\ln r_{\rm tide} + \ln \sigma_v) - \frac{1}{2}\left( \frac{({\boldsymbol r}^{(i)}-{\boldsymbol r}_s)^2}{r_{\rm tide}^2} + \frac{ ({\boldsymbol v}^{(i)}-{\boldsymbol v}_s)^2}{\sigma_v^2} \right)\right]_{t=\tau^{(i)}}\label{eq:ln_like}
\end{align}
where $\boldsymbol r$,$\boldsymbol v$ are the (cartesian) positions and velocities integrated \emph{backwards} in potential $\Phi$ from their present-day positions $\widetilde{\X}$. The whole expression Eq.~\ref{eq:ln_like} is evaluated at the time unbound, $\tau^{(i)}$.

%\begin{align}
%	= K - \frac{1}{2}\sum_j \left[ 2\ln\sigma_j + \left(\frac{X^{(i)}_j-D^{(i)}_j}{\sigma^{(i)}_j}\right)^2\right]&\notag \\
%	- \frac{1}{2}\ln(\left\vert\bSigma_s(\tau^{(i)})\right\vert) - \frac{1}{2}&\langle \X-\X_s, \bSigma_s^{-1}(\X-\X_s) \rangle\\
%	= K - \frac{1}{2}\sum_j \left[ 2\ln\sigma_j + \left(\frac{X^{(i)}_j-D^{(i)}_j}{\sigma^{(i)}_j}\right)^2\right]&\notag \\
%	- \frac{1}{2}\ln(\left\vert\bSigma_s(\tau^{(i)})\right\vert) - \frac{1}{2}&\langle \X-\X_s, \bSigma_s^{-1}(\X-\X_s) \rangle
%\end{align}

\end{document}

### 2014-01-10 text/paper2/ms.tex
started writing paper 2

\documentclass[letterpaper,12pt,preprint]{aastex}

% packages
\usepackage{amssymb,amsmath, amsbsy}

\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\given}{\,|\,}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\transpose}[1]{{#1}^{\mathsf{T}}}
\newcommand{\inverse}[1]{{#1}^{-1}}

% bold X and D
\newcommand{\D}{{\bf D}}
\newcommand{\X}{{\bf X}}
\newcommand{\bSigma}{{\bf \Sigma}}
\newcommand{\bsigma}{\boldsymbol\sigma}

\begin{document}

\title{Inferring the gravitational potential of the Milky Way with measurements of \emph{just two stars}}
\author{Adrian M. Price-Whelan, David W. Hogg, Kathryn V. Johnston, others}

\begin{abstract}
Informative phase-space structures, and in particular those created by
disrupting stellar systems or revealed by chemical tags, may deliver
very precise measures of the gravitational potential in the Milky Way
Halo.  Here we show that even a single pair of stars---two stars that
are known (for some non-kinematic reason) to be likely to be
associated with one another at birth---could provide a significant
constraint on the potential.  The inference is based on a
probabilistic generative model---a likelihood function and priors over
nuisance parameters---that evolves a past putative origin disruption
event forward in time to the present day.  The time, six-volume, and
phase-space location of the origin event are nuisance parameters in
the model and marginalized away.  The method makes no assumption of
integrability, works with finite or even large observational
uncertainties, does not require all dimensions of phase space to be
observed, handles non-zero probability that the two stars are not in
fact associated, and generalizes naturally to larger numbers of stars
and multiple independent structures.  Applications to the GD-1 cold
stellar stream and current surveys for RR Lyrae stars in the Halo are
discussed.
\end{abstract}

\keywords{
  this ---
  that ---
  Milky Way
}

\section{Introduction}

[Cold streams contain tons of information.]

[GD-1 was fit as if it highlighted an orbit.  That is known to be
  wrong.]

[In the halo, any phase-space structure might be long-lived.  Chemical
  tagging could in principle illuminate it.]

[Chemical tagging will always only give probabilistic information.
  But what if it delivers somewhat confident information about small
  stellar families in the halo?]

[Actions, angles, integrable, not, etc.  Can integrate in any
  potential, even a time-varying one.  APW \& KVJ have used this to
  ``run the clock back''.  Here we only let ourselves \emph{generate
    the data}.  Why is an objective based on past collisions not the
  same as assuming that the stars are on the same orbit?]

[Sometimes stellar samples might get extremely small.  For example,
  \gaia\ might find some very low-mass cold stellar streams.  For
  another, there might only be a coupld RRL stars in a particular
  stream, but those RRL stars might be individually extremely well
  measured in six-space.]

[Obviously a small stellar sample can only tell you a small number of
  things about the Milky Way potential.  The idea is---in the long
  run---to construct likelihood functions for many small stellar
  samples and in the end do inference by multiplying them together.
  Duh!]

\section{The model}

Take ${\bf D}^{(i)}$ to be the observed heliocentric 6D position of the $i$th star -- e.g., the measured $l$, $b$, $D$, $\mu_l$, $\mu_b$, $v_r$ -- and ${\bf X}^{(i)}$ to be the corresponding \emph{true} 6D position of the same star. We assume that each star was once part of a progenitor system---e.g., satellite galaxy---presently at observed position ${\bf D}_s$ (with true position ${\bf X}_s$) that is tidally disrupting and forming a cold debris structure in the potential, $\Phi$, of some parent galaxy. We then model the measured position of a star today as the result of sampling from a time-dependent 6D distribution -- the progenitor system -- at some time $\tau^{(i)}$ in the past, and evolving that sampled 6D position to present day in the parent potential, neglecting the effect of self-interaction with the halo of the progenitor. We want to perform inference \emph{without} building and running---within an MCMC chain---a fully dynamical model of a disrupting satellite in the Milky Way. We therefore approximate the progenitor system as a 6D Gaussian with mean vector given by the position of the satellite at any given time, ${\bf X}_s(t)$, and with shape defined by a time-dependent covariance matrix $\bSigma_s(t)$ (the subscript $s$ refers to satellite or progenitor). With these assumptions, the generative model for the data is written:
\begin{align}
	p(\Phi, \X^{(i)}, \X_s, \bSigma_s, \tau^{(i)} \given \D^{(i)}, \D_s) &\propto 
		p(\D^{(i)} \given \X^{(i)}) p(\D_s \given \X_s) p(\X^{(i)} \given \X_s, \bSigma_s, \tau^{(i)})
\end{align}

Taking $\mathcal{N}(x \given \mu, \Sigma)$ to be the normal distribution over x with mean $\mu$ and covariance tensor $\Sigma$, the individual components are simply:
\begin{align}
	p(\D^{(i)} \given \X^{(i)}) &= \mathcal{N}(\X^{(i)} \given \D^{(i)}, \bsigma^{(i)})\\
	p(\D_s \given \X_s) &= \mathcal{N}(\X_s \given \D_s, \bsigma_s)\\
	p(\X^{(i)} \given \X_s, \bSigma_s, \tau^{(i)}) &= \mathcal{N}(\widetilde{\X}^{(i)}(\tau^{(i)}) \given \widetilde{\X}_s(\tau^{(i)}), \bSigma_s(\tau^{(i)}))\label{eq:likelihood}
\end{align}
where the vectors $\bsigma^{(i)}$ and $\bsigma_s$ represent the (assumed) Gaussian, uncorrelated observational errors on the observed 6D position of star $i$ and the satellite, respectively, and the tildes denote a conversion from spherical, heliocentric to cartesian, Galactocentric coordinates. The Gaussian in Eq.~\ref{eq:likelihood} is evaluated at the time at which the $i$th star becomes unbound, $\tau^{(i)}$.

We first consider the case of a progenitor described only by a single (time-dependent) spatial scale---the tidal radius, $r_{\mathrm{tide}}(t)$---and a single velocity scale---the velocity dispersion, $\sigma_\mathrm{v}$---with \emph{perfectly} observed position and velocity components. Under these assumptions, the posterior is greatly simplified:
\begin{align}
	p(\Phi, \X^{(i)}, \tau^{(i)} \given \D^{(i)}) &\propto 
		p(\D^{(i)} \given \X^{(i)}) p(\X^{(i)} \given \X_s, r_{\mathrm{tide}}, \sigma_\mathrm{v}, \tau^{(i)})\label{eq:posterior}
\end{align}
and thus the $\ln$-posterior is
\begin{align}
	\ln p(\Phi, \X^{(i)}, \tau^{(i)} \given \D^{(i)}) &= \ln p(\D^{(i)} \given \X^{(i)}) + \ln p(\X^{(i)} \given \X_s, r_{\mathrm{tide}}, \sigma_\mathrm{v}, \tau^{(i)}).
\end{align}
The first term on the RHS -- from the observed positions of the stars	-- can be expanded to:
\begin{align}
	\ln p(\D^{(i)} \given \X^{(i)}) &= K - \frac{1}{2}\sum_j \left[ 2\ln\sigma_j + \left(\frac{X^{(i)}_j-D^{(i)}_j}{\sigma^{(i)}_j}\right)^2\right]
\end{align}
where $K$ is a constant related to the dimensionality.

The second term -- quantifying how well the potential ``returns'' the stars to the progenitor -- reads as follows (for the case of a spherical progenitor):
\begin{align}
	\ln p(\X^{(i)} \given \X_s, \bSigma_s, \tau^{(i)}) &= M - \frac{1}{2}\ln(\det \bSigma_s(\tau^{(i)})) - \frac{1}{2}\langle \widetilde{\X}-\widetilde{\X}_s, \bSigma_s^{-1}(\widetilde{\X}-\widetilde{\X}_s) \rangle\\
	&= M - \left[3(\ln r_{\rm tide} + \ln \sigma_v) - \frac{1}{2}\left( \frac{({\boldsymbol r}^{(i)}-{\boldsymbol r}_s)^2}{r_{\rm tide}^2} + \frac{ ({\boldsymbol v}^{(i)}-{\boldsymbol v}_s)^2}{\sigma_v^2} \right)\right]_{t=\tau^{(i)}}\label{eq:ln_like}
\end{align}
where $\boldsymbol r$,$\boldsymbol v$ are the (cartesian) positions and velocities integrated \emph{backwards} in potential $\Phi$ from their present-day positions $\widetilde{\X}$, and $M$ is another additive constant. The whole expression Eq.~\ref{eq:ln_like} is evaluated at the time unbound, $\tau^{(i)}$.

%The marginalization integrals require prior pdfs for the
%parameters---or at least the nuisance parameters.  Here the nuisance
%parameters include the 6-vector position $X_k^{(0)}$ of the
%progenitor, the 6-tensor size $\Sigma_k$ of the progenitor, and all
%the times $t_{nk}$:

Eq.~\ref{eq:posterior} does not include the prior distributions needed to perform marginalizations (integrals) over nuisance parameters of the model. In the most naive model, the unbinding times, $\tau^{(i)}$, could be uniformly drawn from the lifetime of the progenitor; that is, the times could be drawn from a uniform pdf extending from some time, $T$, several Gyr in the past to the present day $t=0$. In what follows we adopt for simplicity the flat prior $p(\tau)=U(\tau\given T, 0)$, where $U(x\given a,b)$ is the uniform distribution for $x$ between limits $a<x<b$.

%A more sophisticated model might have the
%release times be more likely to occur near pericenter passages; that
%is, if the progenitor has finite eccentricity (as it surely will) then
%the probability of releasing a star is probably a strong function of
%Galactocentric distance.

[Question for Hogg: I don't understand this paragraph -- can't we just ignore a prior on the true star positions?]
The prior on the phase-space position $X_k$ could be built from some
kind of maximum-entropy considerations in some baseline potential or
in the current setting of the model potential parameters.  Or it could
be generated from empirical distributions found for known substructures or
satellites found in data or simulations of galaxy formation.  For
simplicity we choose the 3-vector position part of $X_k$ to be drawn
from the singular isothermal sphere $p(x)\propto |x|^{-2}$ in the
range [HOGG]$<|x|<$[HOGG]~kpc and the 3-vector velocity part to be
drawn from a spherical (diagonal) Gaussian of velocity dispersion
[HOGG].  These priors are not really justified by prior knowledge, but
they also don't matter enormously when the data are informative.  When
we test the method (below), we will explicitly test in cases where
these priors are inappropriate, to demonstrate that they don't
substantially affect the results in realistic situations.

The prior on the progenitor 6-tensor size $\Sigma_k$---which really
has only two degrees of freedom, $\sigma_{kx}^2$ and $\sigma_{kv}^2$, the
diagonal elements of the 3-tensor spatial extent and the diagonal
elements of the 3-tensor velocity extent---must enforce positivity and
(relatively speaking) ``coldness'' of the structure.  We relatively
arbitrarily put [HOGG] priors on these with parameters [HOGG].

[Milky Way ``background'' model.  Mixture model.]

[Generalization to multiple phase-space structures. Mixture of
  mixtures.]

[Comment on bow-tie structure, Kuepper work, and so on;
  might want to generalize the Gaussian brutality.]

\section{Experiments}

[Each experiment requires a description of how the fake data were
  generated, what the likelihood function is, and posterior potential
  information obtained.]

[basic experiment: $N=2$, $K=1$, high $P_{nk}$, all dimensions
  measured at best-present-day quality]

[\gaia\ quality]

[low $P_{nk}$, MW bg model]

[unknown $P_{nk}$]

[missing data]

[$N>2$]

[$K>1$]

[Can we do a real experiment with two real RRL stars?]

\section{Discussion}

[What's unrealistic about what we have done?  The kinematic model of
  the progenitor is exceedingly simplistic; why are we okay with it
  nonetheless but what could we do to improve it?  Related to this,
  the progenitor doesn't shrink with time (as it should).  What's up
  with that?]

[We did not assume the potential is static in time.  At least not in
  general; nothing in the method requires or suggests this.]

[Issues about computation and scaling.]

[Don't be afraid of nuisance parameters!]

[Nothing special about streams; will work on any structures, including
  shells and the like.]

[Don't eschew low-dimensional constraints on the potential!]

\acknowledgements
Binney, Rix, Sanders

\end{document}

### 2014-01-15 streams/potential/common.py
added spherical nfw potential and try to make acc faster

@@ -23,7 +23,8 @@ from .core import CartesianPotential
 __all__ = ["PointMassPotential", "MiyamotoNagaiPotential",\
            "HernquistPotential", "LogarithmicPotentialLJ",\
            "PlummerPotential", "IsochronePotential",\
-           "AxisymmetricNFWPotential", "AxisymmetricLogarithmicPotential"]
+           "AxisymmetricNFWPotential", "AxisymmetricLogarithmicPotential",
+           "SphericalNFWPotential"]
 
 ############################################################
 #    Potential due to a point mass at a given position
@@ -491,6 +492,68 @@ class AxisymmetricNFWPotential(CartesianPotential):
                                                  f=f, f_prime=df,
                                                  latex=latex,
                                                  parameters=parameters)
+
+##############################################################################
+#    Spherical NFW potential
+#
+def _cartesian_spherical_nfw_model(bases):
+    """ Generates functions to evaluate an NFW potential and its
+        derivatives at a specified position.
+
+        Physical parameters for this potential are:
+            log_m : total mass in the core of the potential
+            a : scale-length
+    """
+
+    # scale G to be in this unit system
+    _G = G.decompose(bases=bases).value
+
+    def f(r,r_0,log_m,a):
+        rr = r-r_0
+        x,y,z = rr.T
+
+        m = np.exp(log_m)
+        R2 = x**2 + y**2 + z**2
+
+        return -3 * _G * m / R2 * np.log(1. + R2/a)
+
+    def df(r,r_0,log_m,a):
+        rr = r-r_0
+        x,y,z = rr.T
+
+        m = np.exp(log_m)
+        R = np.sqrt(x*x + y*y + z*z)
+
+        dphi_dr = 3*_G*m/R * (np.log(1+R/a)/R - 1/(R+a))
+        _x = dphi_dr * x/R
+        _y = dphi_dr * y/R
+        _z = dphi_dr * z/R
+
+        return -np.array([_x,_y,_z]).T
+
+    return (f, df)
+
+class SphericalNFWPotential(CartesianPotential):
+
+    def __init__(self, units, **parameters):
+
+        latex = "$\sigma$"
+
+        assert "log_m" in parameters.keys(), "You must specify a log-mass."
+        assert "a" in parameters.keys(), "You must specify the parameter 'a'."
+
+        # get functions for evaluating potential and derivatives
+        f,df = _cartesian_spherical_nfw_model(units)
+        super(SphericalNFWPotential, self).__init__(units,
+                                                 f=f, f_prime=df,
[…]

### 2014-01-15 streams/potential/lm10.py

@@ -91,10 +91,10 @@ class LawMajewski2010(CompositePotential):
         #                                       halo=halo)
         self.units = usys
         self._G = G.decompose(bases=usys).value
+        self._parameter_dict = dict([(k,v._value) for k,v in self.parameters.items()])
 
     def _acceleration_at(self, r, n_particles, acc):
-        p_dict = dict([(k,v._value) for k,v in self.parameters.items()])
-        return lm10_acceleration(r, n_particles, acc, **p_dict)
+        return lm10_acceleration(r, n_particles, acc, **self._parameter_dict)
 
     def _enclosed_mass(self, R):
         """ Compute the enclosed mass at the position r. Assumes it's far from

### 2014-01-15 text/notes/likelihood/likelihood.tex
minor tweak
forgot phi some places

@@
 Starting with notation, take ${\bf D}^{(i)}$ to be the observed heliocentric 6D position of the $i$th star -- e.g., the measured $l$, $b$, $D$, $\mu_l$, $\mu_b$, $v_r$ -- and ${\bf X}^{(i)}$ to be the corresponding \emph{true} 6D position of the same star.
-We assume that this star was once part of a progenitor system -- e.g., satellite galaxy -- presently at observed position ${\bf D}_s$ (with true position ${\bf X}_s$) that is tidally disrupting and forming a cold debris structure in the potential, $\Phi$, of some parent galaxy.
-We then model the measured position of a star today as the result of sampling from a time-dependent 6D distribution -- the progenitor system -- at some time $\tau^{(i)}$ in the past, and evolving that sampled 6D position to present day in the parent potential, neglecting the effect of self-interaction with the halo of the progenitor.
+We assume that this star was once part of a progenitor system -- e.g., satellite galaxy -- presently at observed position ${\bf D}_s$ (with true position, ${\bf X}_s$ and shape described by a covariance matrix, $\bSigma_s$) that is tidally disrupting and forming a cold debris structure in the potential, $\Phi$, of some parent galaxy.
+We then model the measured position of a star today as the result of sampling from a time-dependent 6D distribution -- the progenitor system -- at some time $\tau^{(i)}$ in the past, evolved to present day in the parent potential, neglecting the effect of self-interaction with the halo of the progenitor.
 We model the progenitor as a 6D Gaussian with mean vector given by the position of the satellite at any given time, ${\bf X}_s(t)$, and with shape defined by a time-dependent covariance matrix $\bSigma_s(t)$ (the subscript $s$ refers to satellite or progenitor).
@@
 \begin{align}
-	p(\Phi, \X^{(i)}, \X_s, \bSigma_s, \tau^{(i)} \given \D^{(i)}, \D_s) &\propto 
-		p(\D^{(i)} \given \X^{(i)}) p(\D_s \given \X_s) p(\X^{(i)} \given \X_s, \bSigma_s, \tau^{(i)})\\
+	p(\Phi, \X^{(i)}, \X_s, \bSigma_s, \tau^{(i)} \given \D^{(i)}, \D_s) &= 
+		\frac{1}{Z} p(\D^{(i)} \given \X^{(i)}) p(\D_s \given \X_s) p(\X^{(i)} \given \X_s, \bSigma_s, \tau^{(i)}, \Phi)\\
 	p(\D^{(i)} \given \X^{(i)}) &= \mathcal{N}(\X^{(i)} \given \D^{(i)}, \bsigma^{(i)})\\
 	p(\D_s \given \X_s) &= \mathcal{N}(\X_s \given \D_s, \bsigma_s)\\
-	p(\X^{(i)} \given \X_s, \bSigma_s, \tau^{(i)}) &= \mathcal{N}(\widetilde{\X}^{(i)}(\tau^{(i)}) \given \widetilde{\X}_s(\tau^{(i)}), \bSigma_s(\tau^{(i)}))
+	p(\X^{(i)} \given \X_s, \bSigma_s, \tau^{(i)}, \Phi) &= \mathcal{N}(\widetilde{\X}^{(i)}(\tau^{(i)}) \given \widetilde{\X}_s(\tau^{(i)}), \bSigma_s(\tau^{(i)}))
 \end{align}
-where the vectors $\bsigma^{(i)}$ and $\bsigma_s$ represent the (assumed) Gaussian, uncorrelated observational errors on the observed 6D position of star $i$ and the satellite, respectively, and the tildes denote a conversion from spherical, heliocentric to cartesian, Galactocentric coordinates.
+where the vectors $\bsigma^{2}$ and $\bsigma_s^2$ represent the (assumed) Gaussian, uncorrelated observational variances on the observed 6D position of star and the satellite, respectively.
+%, and the tildes denote a conversion from spherical, heliocentric to cartesian, Galactocentric coordinates. 
 %Thus, 
@@
 	\ln p(\X^{(i)} \given \X_s, \bSigma_s, \tau^{(i)}) &= M - \frac{1}{2}\ln(\left\vert\bSigma_s(\tau^{(i)})\right\vert) - \frac{1}{2}\langle \widetilde{\X}-\widetilde{\X}_s, \bSigma_s^{-1}(\widetilde{\X}-\widetilde{\X}_s) \rangle\\
-	&= \left[M - 3(\ln r_{\rm tide} + \ln \sigma_v) - \frac{1}{2}\left( \frac{({\boldsymbol r}^{(i)}-{\boldsymbol r}_s)^2}{r_{\rm tide}^2} + \frac{ ({\boldsymbol v}^{(i)}-{\boldsymbol v}_s)^2}{\sigma_v^2} \right)\right]_{t=\tau^{(i)}}\label{eq:ln_like}
+	&= M - \left[3(\ln r_{\rm tide} + \ln \sigma_v) - \frac{1}{2}\left( \frac{({\boldsymbol r}^{(i)}-{\boldsymbol r}_s)^2}{r_{\rm tide}^2} + \frac{ ({\boldsymbol v}^{(i)}-{\boldsymbol v}_s)^2}{\sigma_v^2} \right)\right]_{t=\tau^{(i)}}\label{eq:ln_like}
 \end{align}
@@
 
+\section{Marginalize over $\tau$}
+\begin{align}
+	p(\Phi, \X^{(i)}, \tau^{(i)} \given \D^{(i)}) &\propto 
+		p(\D^{(i)} \given \X^{(i)}) p(\X^{(i)} \given \X_s, \bSigma_s, \tau^{(i)})\\
+	p(\Phi, \X^{(i)} \given \D^{(i)}) &= \int p(\Phi, \X^{(i)}, \tau^{(i)} \given \D^{(i)}) p(\tau^{(i)}) d\tau^{(i)}
+\end{align}	
+\begin{align}
+	p(\Phi, \X^{(i)} \given \D^{(i)}) &\propto p(\D^{(i)} \given \X^{(i)}) \int p(\X^{(i)} \given \X_s, \bSigma_s, \tau^{(i)}) p(\tau^{(i)}) d\tau^{(i)}\\
+	&\propto p(\D^{(i)} \given \X^{(i)}) \sum_j p(\X^{(i)} \given \X_s, \bSigma_s)|_{\tau^{(i)}_j} \Delta t
+\end{align}
+\begin{align}
+	\ln p^{(i)} &\propto \ln p(\D^{(i)} \given \X^{(i)}) + \ln \sum_j p(\X^{(i)} \given \X_s, \bSigma_s, \tau^{(i)}_j) \Delta t\\
+	&\propto \ln p(\D^{(i)} \given \X^{(i)}) + \ln \sum_j \mathcal{N}(\X^{(i)}( \tau^{(i)}_j) \given \X_s( \tau^{(i)}_j), \bSigma_s( \tau^{(i)}_j)) \Delta t\\
+\end{align}
+
 \end{document}

### 2014-01-20 text/notes/likelihood/likelihood.tex
change to W

@@
 \newcommand{\D}{{\bf D}}
+\newcommand{\W}{{\bf W}}
 \newcommand{\X}{{\bf X}}
@@
 
-Starting with notation, take ${\bf D}^{(i)}$ to be the observed heliocentric 6D position of the $i$th star -- e.g., the measured $l$, $b$, $D$, $\mu_l$, $\mu_b$, $v_r$ -- and ${\bf X}^{(i)}$ to be the corresponding \emph{true} 6D position of the same star.
-We assume that this star was once part of a progenitor system -- e.g., satellite galaxy -- presently at observed position ${\bf D}_s$ (with true position, ${\bf X}_s$ and shape described by a covariance matrix, $\bSigma_s$) that is tidally disrupting and forming a cold debris structure in the potential, $\Phi$, of some parent galaxy.
+Starting with notation, take ${\bf D}^{(i)}$ to be the observed, heliocentric 6D position of the $i$th star -- e.g., the measured position on the sky, ($l$, $b$), distance, $D$, proper motions, ($\mu_l$, $\mu_b$), and radial velocity, $v_r$ -- and ${\bf W}^{(i)}$ to be the corresponding \emph{true} 6D position of the same star in the same coordinate and reference frame.
+We assume that this star was once part of a progenitor system -- e.g., satellite galaxy -- presently at observed position ${\bf D}_s$ (with true position, ${\bf W}_s$ and shape described by a covariance matrix, $\bSigma_s$) that is tidally disrupting and forming a cold debris structure in the potential, $\Phi$, of some parent galaxy.
+The true 6D \emph{spherical, heliocentric} positions of the stars, $\W^{(i)}$, and satellite, $\W_s$, have corresponding 6D \emph{cartesian, Galactocentric} coordinate vectors $\X^{(i)}$ and $\X_s$.
 We then model the measured position of a star today as the result of sampling from a time-dependent 6D distribution -- the progenitor system -- at some time $\tau^{(i)}$ in the past, evolved to present day in the parent potential, neglecting the effect of self-interaction with the halo of the progenitor.
-We model the progenitor as a 6D Gaussian with mean vector given by the position of the satellite at any given time, ${\bf X}_s(t)$, and with shape defined by a time-dependent covariance matrix $\bSigma_s(t)$ (the subscript $s$ refers to satellite or progenitor).
+We model the progenitor as a 6D Gaussian with mean vector given by the position of the satellite at any given time, ${\bf W}_s(t)$, and with shape defined by a time-dependent covariance matrix $\bSigma_s(t)$ (the subscript $s$ refers to satellite or progenitor).
 With these assumptions we can write the posterior for a single star as:
 \begin{align}
-	p(\Phi, \X^{(i)}, \X_s, \bSigma_s, \tau^{(i)} \given \D^{(i)}, \D_s) &= 
-		\frac{1}{Z} p(\D^{(i)} \given \X^{(i)}) p(\D_s \given \X_s) p(\X^{(i)} \given \X_s, \bSigma_s, \tau^{(i)}, \Phi)\\
-	p(\D^{(i)} \given \X^{(i)}) &= \mathcal{N}(\X^{(i)} \given \D^{(i)}, \bsigma^{(i)})\\
-	p(\D_s \given \X_s) &= \mathcal{N}(\X_s \given \D_s, \bsigma_s)\\
-	p(\X^{(i)} \given \X_s, \bSigma_s, \tau^{(i)}, \Phi) &= \mathcal{N}(\widetilde{\X}^{(i)}(\tau^{(i)}) \given \widetilde{\X}_s(\tau^{(i)}), \bSigma_s(\tau^{(i)}))
+	p(\Phi, \W^{(i)}, \W_s, \bSigma_s, \tau^{(i)} \given \D^{(i)}, \D_s) &= 
+		\frac{1}{Z} p(\D^{(i)} \given \W^{(i)}) p(\D_s \given \W_s) p(\W^{(i)} \given \W_s, \bSigma_s, \tau^{(i)}, \Phi)\\
+	p(\D^{(i)} \given \W^{(i)}) &= \mathcal{N}(\W^{(i)} \given \D^{(i)}, \bsigma^{(i)})\\
+	p(\D_s \given \W_s) &= \mathcal{N}(\W_s \given \D_s, \bsigma_s)\\
+	p(\W^{(i)} \given \W_s, \bSigma_s, \tau^{(i)}, \Phi) &= \mathcal{N}(\widetilde{\W}^{(i)}(\tau^{(i)}) \given \widetilde{\W}_s(\tau^{(i)}), \bSigma_s(\tau^{(i)}))
 \end{align}
@@
 %\begin{align}
-%	p(\Phi, \X, \X_s, \bSigma_s, \tau \given \D, \D_s) &= \prod_i p(\Phi, \X^{(i)}, \X_s, \bSigma_s, \tau^{(i)} \given \D^{(i)}, \D_s)\\
+%	p(\Phi, \W, \W_s, \bSigma_s, \tau \given \D, \D_s) &= \prod_i p(\Phi, \W^{(i)}, \W_s, \bSigma_s, \tau^{(i)} \given \D^{(i)}, \D_s)\\
 %	p &= \prod_i p^{(i)}\\
 %	\ln p &= \sum_i \ln p^{(i)}\\
-%	\ln p = \sum_i [\ln\mathcal{N}(\X^{(i)} \given \D^{(i)}, \bsigma^{(i)}) &+ \ln\mathcal{N}(\X_s \given \D_s, \bsigma_s) + \ln\mathcal{N}(\X^{(i)}(\tau^{(i)}) \given \X_s(\tau^{(i)}), \bSigma_s(\tau^{(i)}))]
+%	\ln p = \sum_i [\ln\mathcal{N}(\W^{(i)} \given \D^{(i)}, \bsigma^{(i)}) &+ \ln\mathcal{N}(\W_s \given \D_s, \bsigma_s) + \ln\mathcal{N}(\W^{(i)}(\tau^{(i)}) \given \W_s(\tau^{(i)}), \bSigma_s(\tau^{(i)}))]
 %\end{align}
@@
 \begin{align}
-	p(\Phi, \X^{(i)}, \tau^{(i)} \given \D^{(i)}) &\propto 
-		p(\D^{(i)} \given \X^{(i)}) p(\X^{(i)} \given \X_s, \bSigma_s, \tau^{(i)})\\
+	p(\Phi, \W^{(i)}, \tau^{(i)} \given \D^{(i)}) &\propto 
+		p(\D^{(i)} \given \W^{(i)}) p(\W^{(i)} \given \W_s, \bSigma_s, \tau^{(i)})\\
 \end{align}
@@
 \begin{align}
-	\ln p(\Phi, \X^{(i)}, \tau^{(i)} \given \D^{(i)}) &= \ln p(\D^{(i)} \given \X^{(i)}) + \ln p(\X^{(i)} \given \X_s, \bSigma_s, \tau^{(i)}).
+	\ln p(\Phi, \W^{(i)}, \tau^{(i)} \given \D^{(i)}) &= \ln p(\D^{(i)} \given \W^{(i)}) + \ln p(\W^{(i)} \given \W_s, \bSigma_s, \tau^{(i)}).
 \end{align}
@@
 \begin{align}
-	\ln p(\D^{(i)} \given \X^{(i)}) &= K - \frac{1}{2}\sum_j \left[ 2\ln\sigma_j + \left(\frac{X^{(i)}_j-D^{(i)}_j}{\sigma^{(i)}_j}\right)^2\right].
+	\ln p(\D^{(i)} \given \W^{(i)}) &= K - \frac{1}{2}\sum_j \left[ 2\ln\sigma_j + \left(\frac{X^{(i)}_j-D^{(i)}_j}{\sigma^{(i)}_j}\right)^2\right].
 \end{align}
@@
 \begin{align}
-	\ln p(\X^{(i)} \given \X_s, \bSigma_s, \tau^{(i)}) &= M - \frac{1}{2}\ln(\left\vert\bSigma_s(\tau^{(i)})\right\vert) - \frac{1}{2}\langle \widetilde{\X}-\widetilde{\X}_s, \bSigma_s^{-1}(\widetilde{\X}-\widetilde{\X}_s) \rangle\\
+	\ln p(\W^{(i)} \given \W_s, \bSigma_s, \tau^{(i)}) &= M - \frac{1}{2}\ln(\left\vert\bSigma_s(\tau^{(i)})\right\vert) - \frac{1}{2}\langle \widetilde{\W}-\widetilde{\W}_s, \bSigma_s^{-1}(\widetilde{\W}-\widetilde{\W}_s) \rangle\\
 	&= M - \left[3(\ln r_{\rm tide} + \ln \sigma_v) - \frac{1}{2}\left( \frac{({\boldsymbol r}^{(i)}-{\boldsymbol r}_s)^2}{r_{\rm tide}^2} + \frac{ ({\boldsymbol v}^{(i)}-{\boldsymbol v}_s)^2}{\sigma_v^2} \right)\right]_{t=\tau^{(i)}}\label{eq:ln_like}
 \end{align}
-where $\boldsymbol r$,$\boldsymbol v$ are the (cartesian) positions and velocities integrated \emph{backwards} in potential $\Phi$ from their present-day positions $\widetilde{\X}$.
+where $\boldsymbol r$,$\boldsymbol v$ are the (cartesian) positions and velocities integrated \emph{backwards} in potential $\Phi$ from their present-day positions $\widetilde{\W}$.
 The whole expression Eq.~\ref{eq:ln_like} is evaluated at the time unbound, $\tau^{(i)}$.
@@
 %	= K - \frac{1}{2}\sum_j \left[ 2\ln\sigma_j + \left(\frac{X^{(i)}_j-D^{(i)}_j}{\sigma^{(i)}_j}\right)^2\right]&\notag \\
-%	- \frac{1}{2}\ln(\left\vert\bSigma_s(\tau^{(i)})\right\vert) - \frac{1}{2}&\langle \X-\X_s, \bSigma_s^{-1}(\X-\X_s) \rangle\\
+%	- \frac{1}{2}\ln(\left\vert\bSigma_s(\tau^{(i)})\right\vert) - \frac{1}{2}&\langle \W-\W_s, \bSigma_s^{-1}(\W-\W_s) \rangle\\
 %	= K - \frac{1}{2}\sum_j \left[ 2\ln\sigma_j + \left(\frac{X^{(i)}_j-D^{(i)}_j}{\sigma^{(i)}_j}\right)^2\right]&\notag \\
-%	- \frac{1}{2}\ln(\left\vert\bSigma_s(\tau^{(i)})\right\vert) - \frac{1}{2}&\langle \X-\X_s, \bSigma_s^{-1}(\X-\X_s) \rangle
+%	- \frac{1}{2}\ln(\left\vert\bSigma_s(\tau^{(i)})\right\vert) - \frac{1}{2}&\langle \W-\W_s, \bSigma_s^{-1}(\W-\W_s) \rangle
 %\end{align}
@@
 \begin{align}
-	p(\Phi, \X^{(i)}, \tau^{(i)} \given \D^{(i)}) &\propto 
-		p(\D^{(i)} \given \X^{(i)}) p(\X^{(i)} \given \X_s, \bSigma_s, \tau^{(i)})\\
-	p(\Phi, \X^{(i)} \given \D^{(i)}) &= \int p(\Phi, \X^{(i)}, \tau^{(i)} \given \D^{(i)}) p(\tau^{(i)}) d\tau^{(i)}
+	p(\Phi, \W^{(i)}, \tau^{(i)} \given \D^{(i)}) &\propto 
+		p(\D^{(i)} \given \W^{(i)}) p(\W^{(i)} \given \W_s, \bSigma_s, \tau^{(i)})\\
+	p(\Phi, \W^{(i)} \given \D^{(i)}) &= \int p(\Phi, \W^{(i)}, \tau^{(i)} \given \D^{(i)}) p(\tau^{(i)}) d\tau^{(i)}
 \end{align}	
 \begin{align}
-	p(\Phi, \X^{(i)} \given \D^{(i)}) &\propto p(\D^{(i)} \given \X^{(i)}) \int p(\X^{(i)} \given \X_s, \bSigma_s, \tau^{(i)}) p(\tau^{(i)}) d\tau^{(i)}\\
-	&\propto p(\D^{(i)} \given \X^{(i)}) \sum_j p(\X^{(i)} \given \X_s, \bSigma_s)|_{\tau^{(i)}_j} \Delta t
+	p(\Phi, \W^{(i)} \given \D^{(i)}) &\propto p(\D^{(i)} \given \W^{(i)}) \int p(\W^{(i)} \given \W_s, \bSigma_s, \tau^{(i)}) p(\tau^{(i)}) d\tau^{(i)}\\
+	&\propto p(\D^{(i)} \given \W^{(i)}) \sum_j p(\W^{(i)} \given \W_s, \bSigma_s)|_{\tau^{(i)}_j} \Delta t
 \end{align}
 \begin{align}
-	\ln p^{(i)} &\propto \ln p(\D^{(i)} \given \X^{(i)}) + \ln \sum_j p(\X^{(i)} \given \X_s, \bSigma_s, \tau^{(i)}_j) \Delta t\\
-	&\propto \ln p(\D^{(i)} \given \X^{(i)}) + \ln \sum_j \mathcal{N}(\X^{(i)}( \tau^{(i)}_j) \given \X_s( \tau^{(i)}_j), \bSigma_s( \tau^{(i)}_j)) \Delta t\\
+	\ln p^{(i)} &\propto \ln p(\D^{(i)} \given \W^{(i)}) + \ln \sum_j p(\W^{(i)} \given \W_s, \bSigma_s, \tau^{(i)}_j) \Delta t\\
+	&\propto \ln p(\D^{(i)} \given \W^{(i)}) + \ln \sum_j \mathcal{N}(\W^{(i)}( \tau^{(i)}_j) \given \W_s( \tau^{(i)}_j), \bSigma_s( \tau^{(i)}_j)) \Delta t\\
 \end{align}

### 2014-01-20 scripts/observe_simulation.py
Change example

@@ -160,13 +160,15 @@ if __name__ == "__main__":
 
     """
         e.g.:
-        python scripts/observe_simulation.py -v --class_name=SgrSimulation --expr='tub!=0' --N=32 --file=/Users/adrian/projects/streams/data/observed_particles/N32_no_errors.hdf5
+python scripts/observe_simulation.py -v --class_name=SgrSimulation --expr='tub!=0' \
+--N=128 --file=/Users/adrian/projects/streams/data/observed_particles/N128.hdf5 \
+--seed=84
     """
 
     # TODO: class kwargs
     # particle_error_model=gaia_spitzer_errors, satellite_error_model=gaia_spitzer_errors,
     # particle_error_model=gaia_spitzer_errors, satellite_error_model=None,
     observe_simulation(args.class_name,
+        particle_error_model=gaia_spitzer_errors, satellite_error_model=gaia_spitzer_errors,
         selection_expr=args.expr, N=args.N, output_file=args.output_file,
-        particle_error_model=gaia_spitzer_errors, satellite_error_model=None,
         overwrite=args.overwrite, seed=args.seed, class_kwargs=dict(mass="2.5e8"))

### 2014-01-20 text/notes/likelihood/likelihood.tex
changed some notation

@@
 \newcommand{\X}{{\bf X}}
+\newcommand{\V}{{\bf V}}
 \newcommand{\bSigma}{{\bf \Sigma}}
@@
 
-Starting with notation, take ${\bf D}^{(i)}$ to be the observed, heliocentric 6D position of the $i$th star -- e.g., the measured position on the sky, ($l$, $b$), distance, $D$, proper motions, ($\mu_l$, $\mu_b$), and radial velocity, $v_r$ -- and ${\bf W}^{(i)}$ to be the corresponding \emph{true} 6D position of the same star in the same coordinate and reference frame.
-We assume that this star was once part of a progenitor system -- e.g., satellite galaxy -- presently at observed position ${\bf D}_s$ (with true position, ${\bf W}_s$ and shape described by a covariance matrix, $\bSigma_s$) that is tidally disrupting and forming a cold debris structure in the potential, $\Phi$, of some parent galaxy.
-The true 6D \emph{spherical, heliocentric} positions of the stars, $\W^{(i)}$, and satellite, $\W_s$, have corresponding 6D \emph{cartesian, Galactocentric} coordinate vectors $\X^{(i)}$ and $\X_s$.
-We then model the measured position of a star today as the result of sampling from a time-dependent 6D distribution -- the progenitor system -- at some time $\tau^{(i)}$ in the past, evolved to present day in the parent potential, neglecting the effect of self-interaction with the halo of the progenitor.
-We model the progenitor as a 6D Gaussian with mean vector given by the position of the satellite at any given time, ${\bf W}_s(t)$, and with shape defined by a time-dependent covariance matrix $\bSigma_s(t)$ (the subscript $s$ refers to satellite or progenitor).
-With these assumptions we can write the posterior for a single star as:
+Starting with notation, take ${\bf D}^{(i)}$ to be the observed, heliocentric 6D position of the $i$th star -- e.g., the measured position on the sky, ($l$, $b$), distance, $D$, proper motions, ($\mu_l$, $\mu_b$), and radial velocity, $v_r$ -- and ${\bf W}^{(i)}$ to be the corresponding \emph{true}, 6D position of the same star in the same coordinate and reference frame.
+We assume that this star was once part of a progenitor system -- e.g., satellite galaxy -- tidally disrupting and forming a cold debris structure in the potential, $\Phi$, of some parent galaxy.
+The satellite is observed to be at heliocentric 6D position ${\bf D}_s$, with true position ${\bf W}_s$, and shape described by a covariance matrix, $\V_s$ (the subscript $s$ refers to the satellite).
+The true, 6D spherical, heliocentric positions of the stars and satellite -- $\W^{(i)}$ and $\W_s$ -- have corresponding 6D cartesian, Galactocentric coordinate vectors $\X^{(i)}$ and $\X_s$.
+
+We model the observed, present-day position of a star as the result of sampling from a time-dependent 6D distribution -- the progenitor system -- at some time $\tau^{(i)}$ in the past, evolved to present day in the parent potential, neglecting the effect of self-interaction with the halo of the progenitor.
+We model the progenitor as a 6D Gaussian with mean vector given by the position of the satellite at any given time, ${\bf W}_s(t)$, and with shape defined by the time-dependent covariance matrix $\V_s(t)$.
+With these assumptions we can write the joint posterior probability for the observed position of a star, properties of the disrupting satellite, and parent potential given the data as:
 \begin{align}
-	p(\Phi, \W^{(i)}, \W_s, \bSigma_s, \tau^{(i)} \given \D^{(i)}, \D_s) &= 
-		\frac{1}{Z} p(\D^{(i)} \given \W^{(i)}) p(\D_s \given \W_s) p(\W^{(i)} \given \W_s, \bSigma_s, \tau^{(i)}, \Phi)\\
-	p(\D^{(i)} \given \W^{(i)}) &= \mathcal{N}(\W^{(i)} \given \D^{(i)}, \bsigma^{(i)})\\
-	p(\D_s \given \W_s) &= \mathcal{N}(\W_s \given \D_s, \bsigma_s)\\
-	p(\W^{(i)} \given \W_s, \bSigma_s, \tau^{(i)}, \Phi) &= \mathcal{N}(\widetilde{\W}^{(i)}(\tau^{(i)}) \given \widetilde{\W}_s(\tau^{(i)}), \bSigma_s(\tau^{(i)}))
+	%p(\Phi, \W^{(i)}, \W_s, \V_s, \tau^{(i)} \given \D^{(i)}, \D_s) &= 
+	%	\frac{1}{\mathcal{Z}} p(\D^{(i)} \given \W^{(i)}) p(\D_s \given \W_s) p(\W^{(i)} \given \W_s, \V_s, \tau^{(i)}, \Phi)
+	p(\Phi, \W^{(i)}, \W_s, \V_s, \tau^{(i)} \given \D^{(i)}, \D_s) &= 
+		\frac{1}{\mathcal{Z}} p(\D^{(i)} \given \W^{(i)}) p(\D_s \given \W_s) p(\D^{(i)} \given \D_s, \V_s, \tau^{(i)}, \Phi)
 \end{align}
-where the vectors $\bsigma^{2}$ and $\bsigma_s^2$ represent the (assumed) Gaussian, uncorrelated observational variances on the observed 6D position of star and the satellite, respectively.
+where the factor $\mathcal{Z}$ only depends on the properties of the data (the \emph{evidence}), and is thus constant when varying the model parameters.
+We assume that the observational errors are Gaussian:
+\begin{align}
+	p(\D^{(i)} \given \W^{(i)}) &= \mathcal{N}(\W^{(i)} \given \D^{(i)}, \bSigma^{(i)})\\
+	p(\D_s \given \W_s) &= \mathcal{N}(\W_s \given \D_s, \bSigma_s)
+\end{align}
+where the covariance matrices $\bSigma^{(i)}$ and $\bSigma_s$ represent the observational variances on the observed 6D position of star and the satellite, respectively. [...] hmm...
+\begin{align}
+	p(\D^{(i)} \given \D_s, \V_s, \tau^{(i)}, \Phi) &= p(\W^{(i)} \given \W_s, \V_s, \tau^{(i)}, \Phi) \times \det{\bf J}^{(i)}\\
+	&= \left(\mathcal{N}(\W^{(i)} \given \W_s, \V_s) \times \det{\bf J}^{(i)}\right)\rvert_{t=\tau^{(i)}}
+\end{align}
+
 %, and the tildes denote a conversion from spherical, heliocentric to cartesian, Galactocentric coordinates. 

### 2014-01-26 text/notes/likelihood/likelihood.tex

@@
 
-Starting with notation, take ${\bf D}^{(i)}$ to be the observed, heliocentric 6D position of the $i$th star -- e.g., the measured position on the sky, ($l$, $b$), distance, $D$, proper motions, ($\mu_l$, $\mu_b$), and radial velocity, $v_r$ -- and ${\bf W}^{(i)}$ to be the corresponding \emph{true}, 6D position of the same star in the same coordinate and reference frame.
-We assume that this star was once part of a progenitor system -- e.g., satellite galaxy -- tidally disrupting and forming a cold debris structure in the potential, $\Phi$, of some parent galaxy.
-The satellite is observed to be at heliocentric 6D position ${\bf D}_s$, with true position ${\bf W}_s$, and shape described by a covariance matrix, $\V_s$ (the subscript $s$ refers to the satellite).
-The true, 6D spherical, heliocentric positions of the stars and satellite -- $\W^{(i)}$ and $\W_s$ -- have corresponding 6D cartesian, Galactocentric coordinate vectors $\X^{(i)}$ and $\X_s$.
+Take ${\bf D}$ to be the observed, heliocentric 6D position of a star --- e.g., the measured position on the sky, ($l$, $b$); distance, $D$; proper motions, ($\mu_l$, $\mu_b$); and line-of-sight velocity, $v_r$ --- and take ${\bf W}$ to be the corresponding \emph{true}, 6D position of the same star in the same coordinate and reference frame.
+We have determined that this star was once part of a progenitor system (e.g., satellite galaxy) that is tidally disrupting and forming a cold debris structure in the potential, $\Phi$, of some parent galaxy.
+The position of the satellite is observed to be at heliocentric 6D position ${\bf D}_s$, with true position ${\bf W}_s$, where the subscript $s$ refers to the satellite.
+The true, 6D spherical, heliocentric positions of the star and satellite -- $\W$ and $\W_s$ -- have corresponding 6D cartesian, Galactocentric coordinate vectors $\X$ and $\X_s$.
 
-We model the observed, present-day position of a star as the result of sampling from a time-dependent 6D distribution -- the progenitor system -- at some time $\tau^{(i)}$ in the past, evolved to present day in the parent potential, neglecting the effect of self-interaction with the halo of the progenitor.
-We model the progenitor as a 6D Gaussian with mean vector given by the position of the satellite at any given time, ${\bf W}_s(t)$, and with shape defined by the time-dependent covariance matrix $\V_s(t)$.
+We model the present-day position and velocity of the disrupted star as the result of sampling from isotropic, log-normal distributions centered on the tidal radius, $R_{\rm tide}$, and escape velocity, $V_{\rm esc}$ of the satellite, respectively, at some time $\tau$ in the past, evolved to present day in the parent potential, $\Phi$.
+In this model, we neglect the effect of self-interaction with the halo of the progenitor [TODO: maybe we need to justify this, add APW's figure from paper 1 ref.
+response?].
+
 With these assumptions we can write the joint posterior probability for the observed position of a star, properties of the disrupting satellite, and parent potential given the data as:

### 2014-01-26 text/paper2/ms.tex

@@
 \newcommand{\project}[1]{\textsl{#1}}
-\newcommand{\gaia}{\project{Gaia}}
+\newcommand{\gaia}{\project{Gaia}~}
+\newcommand{\spitzer}{\project{Spitzer}~}
 \newcommand{\given}{\,|\,}
@@
 
-\title{Inferring the gravitational potential of the Milky Way with measurements of \emph{just two stars}}
-\author{Adrian M. Price-Whelan, David W. Hogg, Kathryn V. Johnston, others}
+\title{Inferring the gravitational potential of the Milky Way with few, precisely measured stars}
+\author{Adrian M. Price-Whelan, DWH, KVJ, others?}
 
 \begin{abstract}
-Informative phase-space structures, and in particular those created by
-disrupting stellar systems or revealed by chemical tags, may deliver
-very precise measures of the gravitational potential in the Milky Way
-Halo.  Here we show that even a single pair of stars---two stars that
-are known (for some non-kinematic reason) to be likely to be
-associated with one another at birth---could provide a significant
-constraint on the potential.  The inference is based on a
-probabilistic generative model---a likelihood function and priors over
-nuisance parameters---that evolves a past putative origin disruption
-event forward in time to the present day.  The time, six-volume, and
-phase-space location of the origin event are nuisance parameters in
-the model and marginalized away.  The method makes no assumption of
-integrability, works with finite or even large observational
-uncertainties, does not require all dimensions of phase space to be
-observed, handles non-zero probability that the two stars are not in
-fact associated, and generalizes naturally to larger numbers of stars
-and multiple independent structures.  Applications to the GD-1 cold
-stellar stream and current surveys for RR Lyrae stars in the Halo are
-discussed.
+%Informative phase-space structures, and in particular those created by
+%disrupting stellar systems or revealed by chemical tags, may deliver
+%very precise measures of the gravitational potential in the Milky Way
+%Halo.  Here we show that even a single pair of stars---two stars that
+%are known (for some non-kinematic reason) to be likely to be
+%associated with one another at birth---could provide a significant
+%constraint on the potential.  The inference is based on a
+%probabilistic generative model---a likelihood function and priors over
+%nuisance parameters---that evolves a past putative origin disruption
+%event forward in time to the present day.  The time, six-volume, and
+%phase-space location of the origin event are nuisance parameters in
+%the model and marginalized away.  The method makes no assumption of
+%integrability, works with finite or even large observational
+%uncertainties, does not require all dimensions of phase space to be
+%observed, handles non-zero probability that the two stars are not in
+%fact associated, and generalizes naturally to larger numbers of stars
+%and multiple independent structures.  Applications to the GD-1 cold
+%stellar stream and current surveys for RR Lyrae stars in the Halo are
+%discussed.
+TODO...
 \end{abstract}
@@
 \keywords{
-  this ---
-  that ---
-  Milky Way
+  Galaxy: structure
+  ---
+  Galaxy: halo
+  ---
+  cosmology: dark matter
 }
@@
 [Actions, angles, integrable, not, etc.  Can integrate in any
-  potential, even a time-varying one.  APW \& KVJ have used this to
-  ``run the clock back''.  Here we only let ourselves \emph{generate
-    the data}.  Why is an objective based on past collisions not the
+  potential, even a time-varying one. We use this to
+  ``run the clock back''.  Why is an objective based on past collisions not the
   same as assuming that the stars are on the same orbit?]
@@
 
-\section{Experiments}
+\section{Tests on simulated data}
 
@@
 
-[basic experiment: $N=2$, $K=1$, high $P_{nk}$, all dimensions
-  measured at best-present-day quality]
+\subsection{\gaia + \spitzer}
 
-[\gaia\ quality]
-
-[low $P_{nk}$, MW bg model]
-
-[unknown $P_{nk}$]
-
-[missing data]
-
-[$N>2$]
-
-[$K>1$]
-
-[Can we do a real experiment with two real RRL stars?]
+\subsection{\spitzer + current data}
 

### 2014-01-26 plot_super_test.sh
old

@@ -1,23 +1,0 @@
-#!/bin/sh
-
-# Directives
-#PBS -N infer_potential
-#PBS -W group_list=yetiastro
-#PBS -l nodes=1:ppn=4,walltime=16:00:00,mem=64gb
-#PBS -M amp2217@columbia.edu
-#PBS -m abe
-#PBS -V
-
-# Set output and error directories
-#PBS -o localhost:/vega/astro/users/amp2217/pbs_output
-#PBS -e localhost:/vega/astro/users/amp2217/pbs_output
-
-# print date and time to file
-date
-
-#Command to execute Python program
-mpiexec -n 4 /vega/astro/users/amp2217/yt-x86_64/bin/python /vega/astro/users/amp2217/projects/streams/scripts/super_test.py --mpi -v --machine=yeti
-
-date
-
-#End of script

### 2014-01-27 text/notes/likelihood/likelihood.tex
some paper notes

@@
 
-We model the present-day position and velocity of the disrupted star as the result of sampling from isotropic, log-normal distributions centered on the tidal radius, $R_{\rm tide}$, and escape velocity, $V_{\rm esc}$ of the satellite, respectively, at some time $\tau$ in the past, evolved to present day in the parent potential, $\Phi$.
+We model the present-day position and velocity of the disrupted star as the result of sampling from isotropic, log-normal distributions centered on the tidal radius, $R_{\rm tide}$, and escape velocity, $V_{\rm esc}$, of the satellite, respectively, at some time $\tau$ in the past, evolved to present day in the parent potential, $\Phi$.
 In this model, we neglect the effect of self-interaction with the halo of the progenitor [TODO: maybe we need to justify this, add APW's figure from paper 1 ref.

### 2014-01-29 text/notes/likelihood/likelihood.tex
some minor tweaks

--- 2014-01-26 text/notes/likelihood/likelihood.tex
@@
 Take ${\bf D}$ to be the observed, heliocentric 6D position of a star --- e.g., the measured position on the sky, ($l$, $b$); distance, $D$; proper motions, ($\mu_l$, $\mu_b$); and line-of-sight velocity, $v_r$ --- and take ${\bf W}$ to be the corresponding \emph{true}, 6D position of the same star in the same coordinate and reference frame.
-We have determined that this star was once part of a progenitor system (e.g., satellite galaxy) that is tidally disrupting and forming a cold debris structure in the potential, $\Phi$, of some parent galaxy.
+Assume that we have determined that this star was once part of a progenitor system (e.g., satellite galaxy) tidally disrupting and forming a cold debris structure in the potential, $\Phi$, of some parent galaxy.
 The position of the satellite is observed to be at heliocentric 6D position ${\bf D}_s$, with true position ${\bf W}_s$, where the subscript $s$ refers to the satellite.
 The true, 6D spherical, heliocentric positions of the star and satellite -- $\W$ and $\W_s$ -- have corresponding 6D cartesian, Galactocentric coordinate vectors $\X$ and $\X_s$.
+Initially, we assume the present-day mass of the satellite, $M_s$, and velocity dispersion, $\sigma_v$, have been measured perfectly (but we will return to this later).
 
-We model the present-day position and velocity of the disrupted star as the result of sampling from isotropic, log-normal distributions centered on the tidal radius, $R_{\rm tide}$, and escape velocity, $V_{\rm esc}$ of the satellite, respectively, at some time $\tau$ in the past, evolved to present day in the parent potential, $\Phi$.
+We model the present-day position and velocity of the disrupted star as the result of sampling from isotropic, log-normal distributions centered on the instantaneous tidal radius, $R_{\rm tide}$, and velocity dispersion, $\sigma_v$, of the satellite at some time $\tau$ in the past, evolved to present day in the parent potential, $\Phi$.
 In this model, we neglect the effect of self-interaction with the halo of the progenitor [TODO: maybe we need to justify this, add APW's figure from paper 1 ref.
-response?].
-
-With these assumptions we can write the joint posterior probability for the observed position of a star, properties of the disrupting satellite, and parent potential given the data as:
+response?
+refer to another section].
+With these assumptions we can write the joint posterior probability for the observed position of a star, satellite, and parent potential given the data as:
 \begin{align}

### 2014-01-30 streams/io/core.py
fixed huge bug in reading data

@@ -9,6 +9,7 @@ __author__ = "adrn <adrn@astro.columbia.edu>"
 # Standard library
 import os, sys
 import gc
+from random import sample
 
 # Third-party
 import h5py
@@ -110,8 +111,8 @@ def read_table(filename, expr=None, N=None):
         _table = _table[idx]
 
     if N is not None and N > 0:
-        np.random.shuffle(_table)
-        _table = _table[:min(N,len(_table))]
+        idx = np.array(sample(xrange(len(_table)), min(N,len(_table))))
+        _table = _table[idx]
 
     return _table
 

### 2014-01-30 streams/inference/core.py
whitespace

@@ -196,7 +196,8 @@ class StreamModel(object):
 
                     ### HACK TO INITIALIZE WALKERS NEAR true tub!
                     if param_name == "tub":
-                        p0[ii,ix1:ix1+param.size] = np.random.normal(self.true_particles.tub, 100.)
+                        p0[ii,ix1:ix1+param.size] = np.random.normal(self.true_particles.tub,
+                                                                     100.)
 
                     ix1 += param.size
 

### 2014-01-30 text/notes/likelihood/likelihood.tex
writing

@@
 \newcommand{\inverse}[1]{{#1}^{-1}}
+\newcommand{\jac}[1]{\left\vert \J{#1} \right\vert}
 
@@
 \newcommand{\V}{{\bf V}}
+\newcommand{\J}{{\boldsymbol J}}
 \newcommand{\bSigma}{{\bf \Sigma}}
@@
 
-Take ${\bf D}$ to be the observed, heliocentric 6D position of a star --- e.g., the measured position on the sky, ($l$, $b$); distance, $D$; proper motions, ($\mu_l$, $\mu_b$); and line-of-sight velocity, $v_r$ --- and take ${\bf W}$ to be the corresponding \emph{true}, 6D position of the same star in the same coordinate and reference frame.
+Take $\D$ to be the observed, heliocentric 6D position of a star --- e.g., the measured position on the sky, ($l$, $b$); distance, $D$; proper motions, ($\mu_l$, $\mu_b$); and line-of-sight velocity, $v_r$ --- and take $\W$ to be the corresponding \emph{true}, 6D position of the same star in the same coordinate and reference frame.
 Assume that we have determined that this star was once part of a progenitor system (e.g., satellite galaxy) tidally disrupting and forming a cold debris structure in the potential, $\Phi$, of some parent galaxy.
-The position of the satellite is observed to be at heliocentric 6D position ${\bf D}_s$, with true position ${\bf W}_s$, where the subscript $s$ refers to the satellite.
+The position of the satellite is observed to be at heliocentric 6D position $\D_s$, with true position $\W_s$, where the subscript $s$ refers to the satellite.
 The true, 6D spherical, heliocentric positions of the star and satellite -- $\W$ and $\W_s$ -- have corresponding 6D cartesian, Galactocentric coordinate vectors $\X$ and $\X_s$.
@@
 \begin{align}
-	%p(\Phi, \W^{(i)}, \W_s, \V_s, \tau^{(i)} \given \D^{(i)}, \D_s) &= 
-	%	\frac{1}{\mathcal{Z}} p(\D^{(i)} \given \W^{(i)}) p(\D_s \given \W_s) p(\W^{(i)} \given \W_s, \V_s, \tau^{(i)}, \Phi)
-	p(\Phi, \W^{(i)}, \W_s, \V_s, \tau^{(i)} \given \D^{(i)}, \D_s) &= 
-		\frac{1}{\mathcal{Z}} p(\D^{(i)} \given \W^{(i)}) p(\D_s \given \W_s) p(\D^{(i)} \given \D_s, \V_s, \tau^{(i)}, \Phi)
+	p(\Phi, \W, \W_s, \tau \given \D, \D_s) &= \frac{1}{\mathcal{Z}} p(\D \given \W) p(\D_s \given \W_s) 
+												       p(\W \given \W_s, \tau, \Phi) 
+												       p(\Phi) p(\tau)
 \end{align}
 where the factor $\mathcal{Z}$ only depends on the properties of the data (the \emph{evidence}), and is thus constant when varying the model parameters.
-We assume that the observational errors are Gaussian:
+We assume that the observational errors are Gaussian in heliocentric coordinates:
 \begin{align}
-	p(\D^{(i)} \given \W^{(i)}) &= \mathcal{N}(\W^{(i)} \given \D^{(i)}, \bSigma^{(i)})\\
+	p(\D \given \W) &= \mathcal{N}(\W \given \D, \bSigma)\\
 	p(\D_s \given \W_s) &= \mathcal{N}(\W_s \given \D_s, \bSigma_s)
 \end{align}
-where the covariance matrices $\bSigma^{(i)}$ and $\bSigma_s$ represent the observational variances on the observed 6D position of star and the satellite, respectively. [...] hmm...
+where the covariance matrices $\bSigma^{(i)}$ and $\bSigma_s$ represent the observational variances on the observed 6D position of star and the satellite, respectively.
+The likelihood for a given star is given by  [...] hmm...
 \begin{align}
-	p(\D^{(i)} \given \D_s, \V_s, \tau^{(i)}, \Phi) &= p(\W^{(i)} \given \W_s, \V_s, \tau^{(i)}, \Phi) \times \det{\bf J}^{(i)}\\
+	p(\W \given \W_s, \tau, \Phi) &= p(\X \given \X_s, \tau, \Phi) \jac{(x,y,z,v_x,v_y,v_z)}\\
 	&= \left(\mathcal{N}(\W^{(i)} \given \W_s, \V_s) \times \det{\bf J}^{(i)}\right)\rvert_{t=\tau^{(i)}}
 \end{align}
+
+% , but the vectors $\D$ and $\W$ [exist] in different coordinate systems --- heliocentric spherical and Galactocentric cartesian, respectively --- so we must also track the Jacobian that [describes] this transformation:
+
+In general, $p(\tau)$ is likely conditional on the observed orbit of the progenitor but this is neglected and assumed to be uniform over the entire interaction history between the satellite and parent.
+
+% \jac{(x,y,z)}
 

### 2014-01-30 scripts/figures/paper2_figures.py
looking at angle of tub

@@ -192,7 +192,7 @@ def rel_dist(**kwargs):
 def costheta(**kwargs):
 
     nbins = 50
-    fig,axes = plt.subplots(4,1,figsize=(8.5,11), sharex=True, sharey=True)
+    fig,axes = plt.subplots(4,2,figsize=(8.5,11), sharex='col', sharey=True)
     fig2,axes2 = plt.subplots(4,1,figsize=(8.5,11), sharex=True, sharey=True)
 
     #_m = kwargs["mass"]
@@ -229,27 +229,27 @@ def costheta(**kwargs):
         normed_vs = s_x[...,3:] / np.sqrt(np.sum(s_x[...,3:]**2, axis=-1))[:,np.newaxis]
 
         costheta = np.sum(normed_rel_r*normed_vs, axis=-1)
-        sintheta = np.sum(normed_rel_r*normed_rs, axis=-1)
+        cosphi = np.sum(normed_rel_r*normed_rs, axis=-1)
 
-        cosphi = np.sum(normed_rel_v*normed_vs, axis=-1)
+        cos = np.sum(normed_rel_v*normed_vs, axis=-1)
         sinphi = np.sum(normed_rel_v*normed_rs, axis=-1)
 
-        #mu_r, sig_r = norm.fit(np.log(rel_r))
+        n,bins,patches = axes[kk,0].hist(costheta, bins=nbins, alpha=0.25)
+        n,bins,patches = axes[kk,1].hist(cosphi, bins=nbins, alpha=0.25)
+        print(np.sum(cosphi < 0.) / np.sum(cosphi > 0.))
 
-        n,bins,patches = axes[kk].hist(costheta, bins=nbins, alpha=0.25)
-        n,bins,patches = axes[kk].hist(cosphi, bins=bins, alpha=0.25)
+        #n,bins,patches = axes2[kk].hist(sintheta, bins=nbins, alpha=0.25)
+        #n,bins,patches = axes2[kk].hist(sinphi, bins=bins, alpha=0.25)
 
-        n,bins,patches = axes2[kk].hist(sintheta, bins=nbins, alpha=0.25)
-        n,bins,patches = axes2[kk].hist(sinphi, bins=bins, alpha=0.25)
-
-        for ax in [axes[kk],axes2[kk]]:
+        for ax in [axes[kk,0],axes2[kk]]:
             xlim = ax.get_xlim()
             ylim = ax.get_ylim()
[…]

### 2014-01-30 text/notes/likelihood/likelihood.tex
more writing

@@
 Initially, we assume the present-day mass of the satellite, $M_s$, and velocity dispersion, $\sigma_v$, have been measured perfectly (but we will return to this later).
-
-We model the present-day position and velocity of the disrupted star as the result of sampling from isotropic, log-normal distributions centered on the instantaneous tidal radius, $R_{\rm tide}$, and velocity dispersion, $\sigma_v$, of the satellite at some time $\tau$ in the past, evolved to present day in the parent potential, $\Phi$.
-In this model, we neglect the effect of self-interaction with the halo of the progenitor [TODO: maybe we need to justify this, add APW's figure from paper 1 ref.
-response?
-refer to another section].
-With these assumptions we can write the joint posterior probability for the observed position of a star, satellite, and parent potential given the data as:
+With this notation, we can write the joint posterior probability for the observed position of a star, satellite, and parent potential given the data as:
 \begin{align}
@@
 where the covariance matrices $\bSigma^{(i)}$ and $\bSigma_s$ represent the observational variances on the observed 6D position of star and the satellite, respectively.
-The likelihood for a given star is given by  [...] hmm...
+
+We model the present-day position and velocity of the disrupted star as the result of sampling from isotropic, log-normal distributions centered on the instantaneous tidal radius, $R_{\rm tide}$, and velocity dispersion, $\sigma_v$, of the satellite at some time $\tau$ in the past, evolved to present day in the parent potential, $\Phi$.
+In this model, we neglect the effect of self-interaction with the halo of the progenitor [TODO: maybe we need to justify this, add APW's figure from paper 1 ref.
+response?
+refer to another section].
+
+The disruption of a satellite galaxy due to tidal shocking [...].
+Thus, we model the present as disrupting The likelihood of a star's position today is computed by first integrating the orbit of the satellite and star backwards for several Gigayears in the parent potential, $\Phi$, by treating both as test particles.
+
 \begin{align}

### 2014-01-30 streams/inference/back_integrate.py
forgot 1.6

@@ -104,7 +104,7 @@ def back_integration_likelihood(t1, t2, dt, potential, p_hel, s_hel, tub):
     #J2 = lnRV_xyz_jac(rel_x)
     #jac2 = np.array([np.linalg.slogdet(np.linalg.inv(j))[1] for j in J2])
 
-    r_tide = potential._tidal_radius(2.5e8, s_x)
+    r_tide = potential._tidal_radius(2.5e8, s_x)*1.6
     #v_esc = potential._escape_velocity(2.5e8, r_tide=r_tide)
     v_disp = 0.017198632325
 

### 2014-01-31 text/notes/likelihood/likelihood.tex
writing

--- 2014-01-30 text/notes/likelihood/likelihood.tex
@@
 \newcommand{\X}{{\bf X}}
-\newcommand{\V}{{\bf V}}
 \newcommand{\J}{{\boldsymbol J}}
@@
 \newcommand{\bsigma}{\boldsymbol\sigma}
+\newcommand{\rtide}{R_{\rm tide}}
 
@@
 Initially, we assume the present-day mass of the satellite, $M_s$, and velocity dispersion, $\sigma_v$, have been measured perfectly (but we will return to this later).
-
-We model the present-day position and velocity of the disrupted star as the result of sampling from isotropic, log-normal distributions centered on the instantaneous tidal radius, $R_{\rm tide}$, and velocity dispersion, $\sigma_v$, of the satellite at some time $\tau$ in the past, evolved to present day in the parent potential, $\Phi$.
-In this model, we neglect the effect of self-interaction with the halo of the progenitor [TODO: maybe we need to justify this, add APW's figure from paper 1 ref.
-response?
-refer to another section].
-With these assumptions we can write the joint posterior probability for the observed position of a star, satellite, and parent potential given the data as:
+With this notation, we write the joint posterior probability for the observed position of a star, satellite, and parent potential given the data as:
 \begin{align}
@@
 where the covariance matrices $\bSigma^{(i)}$ and $\bSigma_s$ represent the observational variances on the observed 6D position of star and the satellite, respectively.
-The likelihood for a given star is given by  [...] hmm...
+
+The disruption of a satellite galaxy due to tidal shocking [...] Thus, we model the present-day position and velocity of the disrupted star as the result of sampling from isotropic, log-normal distributions centered on the instantaneous tidal radius, $R_{\rm tide}$, and velocity dispersion, $\sigma_v$, of the satellite at some time $\tau$ in the past, evolved to present day in the parent potential, $\Phi$.
+In this model, we neglect the effect of self-interaction with the halo of the progenitor [TODO: maybe we need to justify this, add APW's figure from paper 1 ref.
+response?
+refer to another section].
+The likelihood of a star's position today is computed by first transforming the positions to Galactocentric, cartesian coordinates and integrating the orbits of the satellite and star backwards for several Gigayears in the parent potential, $\Phi$, treating both as test particles.
+The distance, $R$, and velocity, $V$, of the star relative to the satellite are computed at each time-step, and the likelihood is computed as the product of an isotropic, log-normal centered on the tidal radius with an isotropic, log-normal centered on the velocity dispersion of the satellite:
 \begin{align}
-	p(\W \given \W_s, \tau, \Phi) &= p(\X \given \X_s, \tau, \Phi) \jac{(x,y,z,v_x,v_y,v_z)}\\
-	&= \left(\mathcal{N}(\W^{(i)} \given \W_s, \V_s) \times \det{\bf J}^{(i)}\right)\rvert_{t=\tau^{(i)}}
+	p(\W \given \W_s, \tau, \Phi) &= p(\X \given \X_s, \tau, \Phi) \jac{(l,b,D,\mu_l,\mu_b,v_r)}\\
+	p(\X \given \X_s, \tau, \Phi) &= p(\ln R, \ln V \given \tau, \Phi) \jac{(x,y,z,v_x,v_y,v_z)}\label{eq:int} \\
+	p(\ln R, \ln V \given \tau, \Phi) &= \mathcal{N}(\ln R(\tau) | \ln \rtide(\tau))\mathcal{N}(\ln V(\tau) | \ln \sigma_v(\tau))
 \end{align}
+where in eq.~\ref{eq:int} we have implicitly integrated over the two angle coordinates associated with $\ln R$ and the two angles associated with $\ln V$.
 

### 2014-01-31 text/paper2/ms.tex
added thing to address
started experiments section

@@
+%\documentclass{emulateapj}
 \documentclass[letterpaper,12pt,preprint]{aastex}
@@
 
+% commands
 \newcommand{\project}[1]{\textsl{#1}}
@@
 \newcommand{\inverse}[1]{{#1}^{-1}}
+\newcommand{\jac}[1]{\left\vert \J{#1} \right\vert}
 
-% bold X and D
 \newcommand{\D}{{\bf D}}
+\newcommand{\W}{{\bf W}}
 \newcommand{\X}{{\bf X}}
+\newcommand{\J}{{\boldsymbol J}}
 \newcommand{\bSigma}{{\bf \Sigma}}
 \newcommand{\bsigma}{\boldsymbol\sigma}
+\newcommand{\rtide}{R_{\rm tide}}
 
@@
 
-\title{Inferring the gravitational potential of the Milky Way with few, precisely measured stars}
+\title{Inferring the gravitational potential of the Milky Way with few precisely measured stars}
 \author{Adrian M. Price-Whelan, DWH, KVJ, others?}
@@
 
+Things to address:
+- Density is hard to match -- need background model, way to go from luminosity to density, needs assumptions about the mass function and structure of the satellite
+-
+
 \section{The model}
 
-Take ${\bf D}^{(i)}$ to be the observed heliocentric 6D position of the $i$th star -- e.g., the measured $l$, $b$, $D$, $\mu_l$, $\mu_b$, $v_r$ -- and ${\bf X}^{(i)}$ to be the corresponding \emph{true} 6D position of the same star.
-We assume that each star was once part of a progenitor system---e.g., satellite galaxy---presently at observed position ${\bf D}_s$ (with true position ${\bf X}_s$) that is tidally disrupting and forming a cold debris structure in the potential, $\Phi$, of some parent galaxy.
-We then model the measured position of a star today as the result of sampling from a time-dependent 6D distribution -- the progenitor system -- at some time $\tau^{(i)}$ in the past, and evolving that sampled 6D position to present day in the parent potential, neglecting the effect of self-interaction with the halo of the progenitor.
-We want to perform inference \emph{without} building and running---within an MCMC chain---a fully dynamical model of a disrupting satellite in the Milky Way.
-We therefore approximate the progenitor system as a 6D Gaussian with mean vector given by the position of the satellite at any given time, ${\bf X}_s(t)$, and with shape defined by a time-dependent covariance matrix $\bSigma_s(t)$ (the subscript $s$ refers to satellite or progenitor).
-With these assumptions, the generative model for the data is written:
+Take $\D$ to be the observed, heliocentric 6D position of a star --- e.g., the measured position on the sky, ($l$, $b$); distance, $D$; proper motions, ($\mu_l$, $\mu_b$); and line-of-sight velocity, $v_r$ --- and take $\W$ to be the corresponding \emph{true}, 6D position of the same star in the same coordinate and reference frame.
+Assume that we have determined that this star was once part of a progenitor system (e.g., satellite galaxy) tidally disrupting and forming a cold debris structure in the potential, $\Phi$, of some parent galaxy.
+The position of the satellite is observed to be at heliocentric 6D position $\D_s$, with true position $\W_s$, where the subscript $s$ refers to the satellite.
+The true, 6D spherical, heliocentric positions of the star and satellite -- $\W$ and $\W_s$ -- have corresponding 6D cartesian, Galactocentric coordinate vectors $\X$ and $\X_s$.
+Initially, we assume the present-day mass of the satellite, $M_s$, and velocity dispersion, $\sigma_v$, have been measured perfectly (but we will return to this later).
+With this notation, we write the joint posterior probability for the observed position of a star, satellite, and parent potential given the data as:
 \begin{align}
-	p(\Phi, \X^{(i)}, \X_s, \bSigma_s, \tau^{(i)} \given \D^{(i)}, \D_s) &\propto 
-		p(\D^{(i)} \given \X^{(i)}) p(\D_s \given \X_s) p(\X^{(i)} \given \X_s, \bSigma_s, \tau^{(i)})
+	p(\Phi, \W, \W_s, \tau \given \D, \D_s) &= \frac{1}{\mathcal{Z}} p(\D \given \W) p(\D_s \given \W_s) 
+												       p(\W \given \W_s, \tau, \Phi) 
+												       p(\Phi) p(\tau)
 \end{align}
+where the factor $\mathcal{Z}$ only depends on the properties of the data (the \emph{evidence}), and is thus constant when varying the model parameters.
+We assume that the observational errors are Gaussian in heliocentric coordinates:
+\begin{align}
+	p(\D \given \W) &= \mathcal{N}(\W \given \D, \bSigma)\\
+	p(\D_s \given \W_s) &= \mathcal{N}(\W_s \given \D_s, \bSigma_s)
+\end{align}
+where the covariance matrices $\bSigma$ and $\bSigma_s$ specify the observational variances on the observed 6D position of star and the satellite, respectively.
 
-Taking $\mathcal{N}(x \given \mu, \Sigma)$ to be the normal distribution over x with mean $\mu$ and covariance tensor $\Sigma$, the individual components are simply:
+The disruption of a satellite galaxy due to tidal shocking [...] Thus, we model the present-day position and velocity of the disrupted star as the result of sampling from isotropic, log-normal distributions centered on the instantaneous tidal radius, $R_{\rm tide}$, and velocity dispersion, $\sigma_v$, of the satellite at some time $\tau$ in the past, evolved to present day in the parent potential, $\Phi$.
+In this model, we neglect the effect of self-interaction with the halo of the progenitor [TODO: maybe we need to justify this, add APW's figure from paper 1 ref.
+response?
+refer to another section].
+The likelihood of a star's position today is computed by first transforming the positions to Galactocentric, cartesian coordinates and integrating the orbits of the satellite and star backwards for several Gigayears in the parent potential, $\Phi$, treating both as test particles.
+The distance, $R$, and velocity, $V$, of the star relative to the satellite are computed at each time-step, and the likelihood is computed as the product of an isotropic, log-normal centered on the tidal radius with an isotropic, log-normal centered on the velocity dispersion of the satellite:
 \begin{align}
-	p(\D^{(i)} \given \X^{(i)}) &= \mathcal{N}(\X^{(i)} \given \D^{(i)}, \bsigma^{(i)})\\
-	p(\D_s \given \X_s) &= \mathcal{N}(\X_s \given \D_s, \bsigma_s)\\
-	p(\X^{(i)} \given \X_s, \bSigma_s, \tau^{(i)}) &= \mathcal{N}(\widetilde{\X}^{(i)}(\tau^{(i)}) \given \widetilde{\X}_s(\tau^{(i)}), \bSigma_s(\tau^{(i)}))\label{eq:likelihood}
+	p(\W \given \W_s, \tau, \Phi) &= p(\X \given \X_s, \tau, \Phi) \jac{(l,b,D,\mu_l,\mu_b,v_r)}\\
+	p(\X \given \X_s, \tau, \Phi) &= p(\ln R, \ln V \given \tau, \Phi) \jac{(x,y,z,v_x,v_y,v_z)}\label{eq:int} \\
+	p(\ln R, \ln V \given \tau, \Phi) &= \mathcal{N}(\ln R(\tau) | \ln \rtide(\tau))\mathcal{N}(\ln V(\tau) | \ln \sigma_v(\tau))
 \end{align}
-where the vectors $\bsigma^{(i)}$ and $\bsigma_s$ represent the (assumed) Gaussian, uncorrelated observational errors on the observed 6D position of star $i$ and the satellite, respectively, and the tildes denote a conversion from spherical, heliocentric to cartesian, Galactocentric coordinates.
-The Gaussian in Eq.~\ref{eq:likelihood} is evaluated at the time at which the $i$th star becomes unbound, $\tau^{(i)}$.
+where in eq.~\ref{eq:int} we have implicitly integrated over and neglected the two angle coordinates associated with $\ln R$ and the two angles associated with $\ln V$.
 
-We first consider the case of a progenitor described only by a single (time-dependent) spatial scale---the tidal radius, $r_{\mathrm{tide}}(t)$---and a single velocity scale---the velocity dispersion, $\sigma_\mathrm{v}$---with \emph{perfectly} observed position and velocity components.
-Under these assumptions, the posterior is greatly simplified:
+% , but the vectors $\D$ and $\W$ [exist] in different coordinate systems --- heliocentric spherical and Galactocentric cartesian, respectively --- so we must also track the Jacobian that [describes] this transformation:
+
+In general, $p(\tau)$ is likely conditional on the observed orbit of the progenitor but this is neglected and assumed to be uniform over the entire interaction history between the satellite and parent.
+Similarly, the priors on potential parameters, $p(\Phi)$, are taken to be uniform.
+
+\section{Experiments}
+We conduct several tests of this method [...].
+For all tests in this article, we model the potential of the parent galaxy as a three-component sum of a Miyamoto-Nagai disk \citep{}, Hernquist bulge \citep{}, and a triaxial, logarithmic halo \citep[e.g.,][]{law10}:
 \begin{align}
-	p(\Phi, \X^{(i)}, \tau^{(i)} \given \D^{(i)}) &\propto 
-		p(\D^{(i)} \given \X^{(i)}) p(\X^{(i)} \given \X_s, r_{\mathrm{tide}}, \sigma_\mathrm{v}, \tau^{(i)})\label{eq:posterior}
+	\Phi_{\rm disk} &= -\frac{GM_{\rm disk}}{\sqrt{R^2 + (a + \sqrt{z^2 + b^2})^2}}\\
+	\Phi_{\rm bulge} &= -\frac{GM_{\rm bulge}}{r + c}\\
+	\Phi_{halo} &= \frac{1}{2}v_{0}^2 \ln(C_1 x^2 + C_2 y^2 + C_3 xy + (z/q_z)^2 + R_c^2)
 \end{align}
-and thus the $\ln$-posterior is
+where $C_1$, $C_2$, and $C_3$ are combinations of the $x$ and $y$ axis
+ratios ($q_1$, $q_2$) and orientation of the halo with respect to the
+baryonic disk ($\phi$):
 \begin{align}
-	\ln p(\Phi, \X^{(i)}, \tau^{(i)} \given \D^{(i)}) &= \ln p(\D^{(i)} \given \X^{(i)}) + \ln p(\X^{(i)} \given \X_s, r_{\mathrm{tide}}, \sigma_\mathrm{v}, \tau^{(i)}).
+  C_1 &= \frac{\cos^2\phi}{q_1^2} + \frac{\sin^2\phi}{q_2^2}\\
+  C_2 &= \frac{\sin^2\phi}{q_1^2} + \frac{\cos^2\phi}{q_2^2}\\
+  C_3 &= 2\sin\phi\cos\phi \left(q_1^{-2} - q_2^{-2}\right).
 \end{align}
-The first term on the RHS -- from the observed positions of the stars	-- can be expanded to:
-\begin{align}
-	\ln p(\D^{(i)} \given \X^{(i)}) &= K - \frac{1}{2}\sum_j \left[ 2\ln\sigma_j + \left(\frac{X^{(i)}_j-D^{(i)}_j}{\sigma^{(i)}_j}\right)^2\right]
-\end{align}
-where $K$ is a constant related to the dimensionality.
-
-The second term -- quantifying how well the potential ``returns'' the stars to the progenitor -- reads as follows (for the case of a spherical progenitor):
-\begin{align}
-	\ln p(\X^{(i)} \given \X_s, \bSigma_s, \tau^{(i)}) &= M - \frac{1}{2}\ln(\det \bSigma_s(\tau^{(i)})) - \frac{1}{2}\langle \widetilde{\X}-\widetilde{\X}_s, \bSigma_s^{-1}(\widetilde{\X}-\widetilde{\X}_s) \rangle\\
-	&= M - \left[3(\ln r_{\rm tide} + \ln \sigma_v) - \frac{1}{2}\left( \frac{({\boldsymbol r}^{(i)}-{\boldsymbol r}_s)^2}{r_{\rm tide}^2} + \frac{ ({\boldsymbol v}^{(i)}-{\boldsymbol v}_s)^2}{\sigma_v^2} \right)\right]_{t=\tau^{(i)}}\label{eq:ln_like}
-\end{align}
-where $\boldsymbol r$,$\boldsymbol v$ are the (cartesian) positions and velocities integrated \emph{backwards} in potential $\Phi$ from their present-day positions $\widetilde{\X}$, and $M$ is another additive constant.
-The whole expression Eq.~\ref{eq:ln_like} is evaluated at the time unbound, $\tau^{(i)}$.
-
-%The marginalization integrals require prior pdfs for the
-%parameters---or at least the nuisance parameters.  Here the nuisance
-%parameters include the 6-vector position $X_k^{(0)}$ of the
-%progenitor, the 6-tensor size $\Sigma_k$ of the progenitor, and all
-%the times $t_{nk}$:
-
-Eq.~\ref{eq:posterior} does not include the prior distributions needed to perform marginalizations (integrals) over nuisance parameters of the model.
-In the most naive model, the unbinding times, $\tau^{(i)}$, could be uniformly drawn from the lifetime of the progenitor; that is, the times could be drawn from a uniform pdf extending from some time, $T$, several Gyr in the past to the present day $t=0$.
-In what follows we adopt for simplicity the flat prior $p(\tau)=U(\tau\given T, 0)$, where $U(x\given a,b)$ is the uniform distribution for $x$ between limits $a<x<b$.
-
-%A more sophisticated model might have the
-%release times be more likely to occur near pericenter passages; that
-%is, if the progenitor has finite eccentricity (as it surely will) then
-%the probability of releasing a star is probably a strong function of
-%Galactocentric distance.
-
-[Question for Hogg: I don't understand this paragraph -- can't we just ignore a prior on the true star positions?]
-The prior on the phase-space position $X_k$ could be built from some
-kind of maximum-entropy considerations in some baseline potential or
-in the current setting of the model potential parameters.  Or it could
-be generated from empirical distributions found for known substructures or
-satellites found in data or simulations of galaxy formation.  For
-simplicity we choose the 3-vector position part of $X_k$ to be drawn
-from the singular isothermal sphere $p(x)\propto |x|^{-2}$ in the
-range [HOGG]$<|x|<$[HOGG]~kpc and the 3-vector velocity part to be
-drawn from a spherical (diagonal) Gaussian of velocity dispersion
-[HOGG].  These priors are not really justified by prior knowledge, but
-they also don't matter enormously when the data are informative.  When
-we test the method (below), we will explicitly test in cases where
-these priors are inappropriate, to demonstrate that they don't
-substantially affect the results in realistic situations.
-
-The prior on the progenitor 6-tensor size $\Sigma_k$---which really
-has only two degrees of freedom, $\sigma_{kx}^2$ and $\sigma_{kv}^2$, the
-diagonal elements of the 3-tensor spatial extent and the diagonal
-elements of the 3-tensor velocity extent---must enforce positivity and
-(relatively speaking) ``coldness'' of the structure.  We relatively
-arbitrarily put [HOGG] priors on these with parameters [HOGG].
-
-[Milky Way ``background'' model.  Mixture model.]
 
@@
 
+\bibliographystyle{apj}
+\bibliography{refs}
+
 \acknowledgements

### 2014-01-31 text/paper2/ms.tex
added potential stuff

@@
 \usepackage{amssymb,amsmath, amsbsy}
+\usepackage{booktabs}
 
@@
 Initially, we assume the present-day mass of the satellite, $M_s$, and velocity dispersion, $\sigma_v$, have been measured perfectly (but we will return to this later).
-With this notation, we write the joint posterior probability for the observed position of a star, satellite, and parent potential given the data as:
+With this notation, we write the joint posterior probability for the observed position of a star, satellite, and parameters for the parent potential as:
 \begin{align}
@@
 We conduct several tests of this method [...].
-For all tests in this article, we model the potential of the parent galaxy as a three-component sum of a Miyamoto-Nagai disk \citep{}, Hernquist bulge \citep{}, and a triaxial, logarithmic halo \citep[e.g.,][]{law10}:
+For all tests in this article, we model the potential of the parent galaxy as a three-component sum of a Miyamoto-Nagai disk \citep{}, Hernquist bulge \citep[spheroid][]{}, and a triaxial, logarithmic halo \citep[e.g.,][]{law10}:
 \begin{align}
 	\Phi_{\rm disk} &= -\frac{GM_{\rm disk}}{\sqrt{R^2 + (a + \sqrt{z^2 + b^2})^2}}\\
-	\Phi_{\rm bulge} &= -\frac{GM_{\rm bulge}}{r + c}\\
-	\Phi_{halo} &= \frac{1}{2}v_{0}^2 \ln(C_1 x^2 + C_2 y^2 + C_3 xy + (z/q_z)^2 + R_c^2)
+	\Phi_{\rm spher} &= -\frac{GM_{\rm spher}}{r + c}\\
+	\Phi_{halo} &= v_{h}^2 \ln(C_1 x^2 + C_2 y^2 + C_3 xy + (z/q_z)^2 + r_h^2)
 \end{align}
@@
 \end{align}
+
+\begin{table}[h]
+	\begin{tabular}{l c c c} \toprule
+		{\bf Component} & {\bf Parameter} & {\bf Value} & {\bf Prior} \\\toprule
+		disk & $M_{\rm disk}$ & $1.0\times10^{11}M_\odot$ & -- \\ 
+		& $a$ & 6.5 kpc & --\\
+		& $b$ & 0.26 kpc & --\\
+		\midrule
+		spheroid & $M_{\rm spher}$ & $3.4\times10^{10}M_\odot$ & --\\ 
+		& $c$ & 0.7 kpc & --\\
+		\midrule
+		halo & $v_h$ & -- & $\mathcal{U}(100,200)$ km $\mathrm{s}^{-1}$ \\
+		& $q_1$ & -- & $\mathcal{U}(0.7,2)$\\
+		& $q_2$ & 1.0 & --\\
+		& $q_z$ & -- & $\mathcal{U}(0.7,2)$\\
+		& $\phi$ & -- & $\mathcal{U}(45,135)$ deg\\
+		& $r_h$ & 12 kpc & --\\
+		\bottomrule
+		\hline
+	\end{tabular}
+	\caption{Parameter values used ...}
+\end{table}
 

### 2014-01-31 text/paper2/ms.tex
more description of params

@@
 
-In general, $p(\tau)$ is likely conditional on the observed orbit of the progenitor but this is neglected and assumed to be uniform over the entire interaction history between the satellite and parent.
+In general, $p(\tau)$ is likely conditional on the observed orbit of the progenitor---stars are preferentially stripped at pericentric passages \citep{??}---but this is neglected and assumed to be uniform over the entire interaction history between the satellite and parent.
 Similarly, the priors on potential parameters, $p(\Phi)$, are taken to be uniform.
@@
 \begin{align}
-	\Phi_{\rm disk} &= -\frac{GM_{\rm disk}}{\sqrt{R^2 + (a + \sqrt{z^2 + b^2})^2}}\\
-	\Phi_{\rm spher} &= -\frac{GM_{\rm spher}}{r + c}\\
-	\Phi_{halo} &= v_{h}^2 \ln(C_1 x^2 + C_2 y^2 + C_3 xy + (z/q_z)^2 + r_h^2)
+	&\Phi_{\rm disk}(R,z) = -\frac{GM_{\rm disk}}{\sqrt{R^2 + (a + \sqrt{z^2 + b^2})^2}}\\
+	&\Phi_{\rm spher}(r) = -\frac{GM_{\rm spher}}{r + c}\\
+	&\Phi_{halo}(x,y,z) = v_{h}^2 \ln(C_1 x^2 + C_2 y^2 + C_3 xy + (z/q_z)^2 + r_h^2)
 \end{align}
@@
 
+[TODO: explain simulated data we use - KVJ are you ok releasing data from your 2.5e6-2.5e9 Sgr sims? Explain that we use the LM10 potential in Table~\ref{tbl:potential} for the sims].
+
+When recovering the potential, we hold fixed the disk and spheroid parameters (see Table~\ref{tbl:potential}), along with two halo parameters: the scale radius, $r_h$, and one of the flattening components in the x-y plane, $q_2$.
+These halo parameters are degenerate with combinations of the other parameters --- e.g., setting $q_1=2.0$, $q_2=1.0$, $\phi=0^\circ$ is equivalent to setting $q_1=1.0$, $q_2=2.0$, $\phi=90^\circ$.
+The priors on the remaining halo parameters are taken to be uniform over a conservative domain of realistic values: for $v_h$, 100-200~kms$^{-1}$ corresponds to a range in solar circular velocities from $\sim$210-250 kms$^{-1}$ (holding other parameters fixed); the axis ratios [...]; and $\phi$ is restricted to $\pm45^\circ$ around the value measured by LM10 ($\phi = 97^\circ$).
+
 \begin{table}[h]
+\begin{center}
 	\begin{tabular}{l c c c} \toprule
@@
 		halo & $v_h$ & -- & $\mathcal{U}(100,200)$ km $\mathrm{s}^{-1}$ \\
-		& $q_1$ & -- & $\mathcal{U}(0.7,2)$\\
+		& $q_1$ & -- & $\mathcal{U}(1,2)$\\
 		& $q_2$ & 1.0 & --\\
-		& $q_z$ & -- & $\mathcal{U}(0.7,2)$\\
-		& $\phi$ & -- & $\mathcal{U}(45,135)$ deg\\
+		& $q_z$ & -- & $\mathcal{U}(1,2)$\\
+		& $\phi$ & -- & $\mathcal{U}(52,142)$ deg\\
 		& $r_h$ & 12 kpc & --\\
@@
 	\end{tabular}
-	\caption{Parameter values used ...}
+	\caption{Parameter values used ...\label{tbl:potential}}
+\end{center}
 \end{table}
+
+\pagebreak
+
+(Hogg comments)
 

### 2014-02-01 text/paper2/ms.tex

@@
 
-Things to address:
-- Density is hard to match -- need background model, way to go from luminosity to density, needs assumptions about the mass function and structure of the satellite
--
+[Densities are hard to match -- need a background model, way to go from luminosity to density, needs assumptions about the mass function and structure of the satellite\
+
+[Cold vs. Hot streams in the context of observational errors]
+
+[We properly handle observational errors]
 
@@
 
-\pagebreak
-
-(Hogg comments)
-
-[Generalization to multiple phase-space structures. Mixture of
-  mixtures.]
-
-[Comment on bow-tie structure, Kuepper work, and so on;
-  might want to generalize the Gaussian brutality.]
-
-\section{Tests on simulated data}
-
-[Each experiment requires a description of how the fake data were
-  generated, what the likelihood function is, and posterior potential
-  information obtained.]
-
 \subsection{\gaia + \spitzer}
+64 stars in Sgr-like stream
 
 \subsection{\spitzer + current data}
+64 stars in Sgr-like stream
+
+\subsection{\spitzer + current data, no progenitor}
+32 stars in Orphan-like stream, XX stars in GD-1 -like stream
 
 \section{Discussion}
+
+[Cold vs. Hot streams]
+
+[bow-tie structure, Kuepper work, and so on]
 
@@
 [Issues about computation and scaling.]
-
-[Don't be afraid of nuisance parameters!]
 

### 2014-02-07 text/paper2/ms.tex
notes from hoggs office

@@
 
+[From KVJ:]
+- Broader comparison to other methods
+	- 'generative methods' more generally (ie N-body, plus Varghese, plus Kuepper plus new possible action-angle models....
+	- statistical methods
+	- Sanders \& Binney 2013
+- Then a short note on worries
+	- for generative models
+	- for any methods based on action/angle 
+- Then how rewinder gets round these????
+	- Reference Deg \& Widrow to say they illustrate the idea of the power of multiple streams to back up the bit where you say that's the long-term aim
+
 [Cold streams contain tons of information.]
@@
 
+Even if you have a ton of stars, some might be orders of magnitude more constraining (Bovy, ..., Hogg)
+
 [Obviously a small stellar sample can only tell you a small number of
@@
 
+[Are there integrable triaxial potentials?]
+
 \section{The model}
 
-Take $\D$ to be the observed, heliocentric 6D position of a star --- e.g., the measured position on the sky, ($l$, $b$); distance, $D$; proper motions, ($\mu_l$, $\mu_b$); and line-of-sight velocity, $v_r$ --- and take $\W$ to be the corresponding \emph{true}, 6D position of the same star in the same coordinate and reference frame.
+Take $\D$ to be the observed, heliocentric 6D position of a star --- e.g., the measured position on the sky, ($l$, $b$); distance, $D$; proper motions, ($\mu_l$, $\mu_b$); and line-of-sight velocity, $v_r$ --- and take $\W$ to be the corresponding \emph{true}, 6D position of the star in the same coordinate and reference frame.
 Assume that we have determined that this star was once part of a progenitor system (e.g., satellite galaxy) tidally disrupting and forming a cold debris structure in the potential, $\Phi$, of some parent galaxy.
-The position of the satellite is observed to be at heliocentric 6D position $\D_s$, with true position $\W_s$, where the subscript $s$ refers to the satellite.
-The true, 6D spherical, heliocentric positions of the star and satellite -- $\W$ and $\W_s$ -- have corresponding 6D cartesian, Galactocentric coordinate vectors $\X$ and $\X_s$.
-Initially, we assume the present-day mass of the satellite, $M_s$, and velocity dispersion, $\sigma_v$, have been measured perfectly (but we will return to this later).
+The position of the satellite is observed to be at heliocentric 6D position $\D_s$, with true position $\W_s$ (where any subscript $s$ refers to the satellite).
+The true 6D spherical, heliocentric positions of the star and satellite -- $\W$ and $\W_s$ -- have corresponding 6D cartesian, Galactocentric coordinate vectors $\X$ and $\X_s$.
+Initially we assume the present-day mass of the satellite, $M_s$, and velocity dispersion, $\sigma_v$, have been measured perfectly and neglect mass-loss from the satellite.
 With this notation, we write the joint posterior probability for the observed position of a star, satellite, and parameters for the parent potential as:
@@
 Similarly, the priors on potential parameters, $p(\Phi)$, are taken to be uniform.
+
+Hogg:
+
+If interested in oblateness and mass, there are close integrable potentials
+
+- if want to ask the cosmological question (connection to lambda-cdm), need to face non-integrable potentials
+
+- No even approximate way to answer questions about substructure or severe triaxiality that changes with radius
 

### 2014-02-07 text/paper2/ms.tex
minor outline revisions

@@
 
-\title{Inferring the gravitational potential of the Milky Way with few precisely measured stars}
-\author{Adrian M. Price-Whelan, DWH, KVJ, others?}
+\title{Inferring the gravitational potential of the Milky Way with a few precisely measured stars}
+\author{Adrian M. Price-Whelan, David W. Hogg, Kathryn V. Johnston, [TODO: order?]}
 
@@
 
-[From KVJ:]
-- Broader comparison to other methods
-	- 'generative methods' more generally (ie N-body, plus Varghese, plus Kuepper plus new possible action-angle models....
-	- statistical methods
-	- Sanders \& Binney 2013
-- Then a short note on worries
-	- for generative models
-	- for any methods based on action/angle 
-- Then how rewinder gets round these????
-	- Reference Deg \& Widrow to say they illustrate the idea of the power of multiple streams to back up the bit where you say that's the long-term aim
+[TODO: A historical introduction to dark matter, potential measures?]
 
-[Cold streams contain tons of information.]
+Major open questions in near-field cosmology: (1) is the Milky Way potential triaxial and if so, how much?
+(2) How does the potential vary with radius?
+(3) Is there graininess to the potential?
+e.g., is there substructure?
 
-[GD-1 was fit as if it highlighted an orbit.  That is known to be
-  wrong.]
+Cosmological simulations of galaxy formation in the $\Lambda$CDM paradigm seem to produce triaxial dark matter halos \citep{dubinski91, jing02, kuhlen07, veraciro11}.
+Halos tend to be more strongly triaxial in dark-matter-only simulations, but still present with the inclusion of baryons.
+Baryonic processes tend to modify the dark matter structure in the inner galaxy, leading possibly to radially-dependent axis ratios and halo orientation.
+[statements about integrability].
+[if want to connect lambda-CDM to the milky way, we need to face non-integrable potentials]
 
-[In the halo, any phase-space structure might be long-lived.  Chemical
-  tagging could in principle illuminate it.]
+[something about measuring halo shape with lensing].
+
+[why use tidal streams?] \cite{lm10} used N-body models to ... triaxial.
+
+Another prediction of hierarchical structure formation in $\Lambda$CDM is the presence of substructure.
+
+- Broader comparison to other methods\\
+\indent\indent- 'generative methods' more generally (ie N-body, plus Varghese, plus Kuepper plus new possible action-angle models....\\
+\indent\indent- statistical methods\\
+\indent\indent- Sanders \& Binney 2013\\
+- Then a short note on worries\\
+\indent\indent- for generative models\\
+\indent\indent- for any methods based on action/angle \\
+\indent\indent- GD-1 was fit as if it highlighted an orbit.  That is known to be wrong.\\
+\indent\indent- Densities are hard to match: need a background model, way to go from luminosity to density, needs assumptions about the mass function and structure of the satellite\\
+\indent\indent- Triaxiality / radially-dependent shapes and substructure present problems for integral-based methods\\
+- Then how rewinder gets round these????\\
+\indent\indent- Reference Deg \& Widrow to say they illustrate the idea of the power of multiple streams to back up the bit where you say that's the long-term aim\\
 
@@
   \gaia\ might find some very low-mass cold stellar streams.  For
-  another, there might only be a coupld RRL stars in a particular
+  another, there might only be a couple RRL stars in a particular
   stream, but those RRL stars might be individually extremely well
@@
 
-Even if you have a ton of stars, some might be orders of magnitude more constraining (Bovy, ..., Hogg)
-
-[Obviously a small stellar sample can only tell you a small number of
-  things about the Milky Way potential.  The idea is---in the long
-  run---to construct likelihood functions for many small stellar
-  samples and in the end do inference by multiplying them together.
-  Duh!]
-
-[Densities are hard to match -- need a background model, way to go from luminosity to density, needs assumptions about the mass function and structure of the satellite\
+[Even if you have many stars, some might be orders of magnitude more constraining (Bovy, ..., Hogg)]
 
@@
 [We properly handle observational errors]
-
-[Are there integrable triaxial potentials?]
 
@@
 Similarly, the priors on potential parameters, $p(\Phi)$, are taken to be uniform.
-
-Hogg:
-
-If interested in oblateness and mass, there are close integrable potentials
-
-- if want to ask the cosmological question (connection to lambda-cdm), need to face non-integrable potentials
-
-- No even approximate way to answer questions about substructure or severe triaxiality that changes with radius
 

### 2014-02-07 scripts/test_infer_potential.py
test case

@@ -14,6 +14,7 @@ import time
 
 # Third-party
 import astropy.units as u
+from astropy.utils import isiterable
 import h5py
 import matplotlib.pyplot as plt
 import numpy as np
@@ -25,6 +26,7 @@ import streams.io as io
 from streams.io.sgr import SgrSimulation
 import streams.inference as si
 import streams.potential as sp
+from streams.util import project_root
 
 # Create logger
 logger = logging.getLogger(__name__)
@@ -32,9 +34,15 @@ logging.basicConfig(level=logging.DEBUG)
 
 minimum_config = """
 name: test
-data_file: data/observed_particles/N1024.hdf5
+data_file: data/observed_particles/2.5e8_N1024.hdf5
+# nparticles: 4
+# particle_idx: [217, 276, 358, 470]
 nparticles: 4
-particle_idx: [0, 1, 512, 513]
+particle_idx: [77, 132, 658, 847]
+# nparticles: 8
+# particle_idx: [25, 77, 132, 140, 359, 658, 693, 847]
+# nparticles: 32
+# particle_idx: [28, 55, 65, 142, 144, 154, 243, 277, 339, 362, 374, 376, 395, 396, 434, 439, 464, 467, 467, 536, 536, 553, 584, 706, 715, 730, 734, 734, 822, 852, 862, 939]
 
 potential:
     class_name: LawMajewski2010
@@ -51,10 +59,10 @@ pot_params = """
     parameters:
         q1:
             a: 1.
-            b: 1.7
+            b: 1.6
         qz:
             a: 1.
-            b: 1.7
+            b: 1.6
         phi:
             a: 80. degree
             b: 120. degree
@@ -65,7 +73,7 @@ pot_params = """
 
 ptc_params = """
     parameters:
-        _X:
+#        _X:
         tub:
 """
 
@@ -78,52 +86,68 @@ lm10_c = minimum_config.format(potential_params=pot_params,
                                particles_params="",
                                satellite_params="")
 
-_configs = dict()
-_configs[0] = minimum_config.format(potential_params=pot_params,
-                                    particles_params="",
-                                    satellite_params="")
-# _configs[1] = minimum_config.format(potential_params="",
-#                                     particles_params=ptc_params,
-#                                     satellite_params="")
-# _configs.append(minimum_config.format(potential_params="",
-#                                       particles_params="",
-#                                       satellite_params=sat_params))
-# _configs.append(minimum_config.format(potential_params=pot_params,
[…]

### 2014-02-08 text/paper2/ms.tex
lots of writin'

@@
 
-[TODO: A historical introduction to dark matter, potential measures?]
+Cosmological simulations of galaxy formation in the $\Lambda$CDM paradigm predict dark matter halos that (1) are permeated with substructure on many scales, (2) are triaxial in shape, and (3) have density profiles and shapes that vary with radius \citep{dubinski91, jing02, kuhlen07, veraciro11}.
+Dark-matter-only simulations seem to produce strongly triaxial halos, but including baryons tends to soften the triaxiality in the inner galaxy [through a combination of adiabatic contraction and energetic feedback \citep[see recent discussion by][]{pontzen12} TODO: modify previous sentence bc in PW\&J 2013].
+These [...] may lead to radially-dependent axis ratios, halo orientation, and density profiles.
 
-Major open questions in near-field cosmology: (1) is the Milky Way potential triaxial and if so, how much?
-(2) How does the potential vary with radius?
-(3) Is there graininess to the potential?
-e.g., is there substructure?
+The bulk of the baryonic matter in galaxies spans roughly 5-10\% of the extent of the host dark matter halo.
+Hence, the major components of a galaxy are sensitive to the inner portion of its host halo's mass distribution.
+For example, the rotation curves of disk galaxies are particularly sensitive probes of the inner mass since matter in disks can be assumed to move on near circular orbits.
+This extra knowledge about the orbital distribution of the baryonic matter within a disk allows a more precise tracing of the potential than is feasible for objects such as elliptical galaxies where orbits are assumed to be random.
+[Potential is most likely smoothed out (energetic feedback?), closely integrable allowing beautiful dynamical ...]
 
-Cosmological simulations of galaxy formation in the $\Lambda$CDM paradigm seem to produce triaxial dark matter halos \citep{dubinski91, jing02, kuhlen07, veraciro11}.
-Halos tend to be more strongly triaxial in dark-matter-only simulations, but still present with the inclusion of baryons.
-Baryonic processes tend to modify the dark matter structure in the inner galaxy, leading possibly to radially-dependent axis ratios and halo orientation.
-[statements about integrability].
-[if want to connect lambda-CDM to the milky way, we need to face non-integrable potentials]
+Conversely, measuring the dark matter distribution at large radii is difficult as the density of visible tracers becomes small.
+Around external galaxies, the extended mass distribution has been explored using a variety of approaches \citep[see][for a a complete and detailed review]{courteau12}.
+For example, the kinematics of tracer populations such as globular clusters or planetary nebulae can be used to derive mass estimates under the assumptions that these satellite systems are relaxed and well-mixed \citep[early investigations include][]{mendez01,cote03}.
+These estimates are necessarily limited by the number of objects bright enough for spectroscopic measurements and thus the largest samples to date have roughly several hundred members probing out to several 10's of kpc.
+Simple, parameterized models of both the mass and orbit distribution have been simultaneously constrained using such data \citep[e.g.][]{napolitano11,deason12a}.
+%\citep{romanowsky03,douglas07,romanowsky09,napolitano11,lee11}
+Alternatively, the statistical properties of gravitationally lensed background sources around a galaxy can be used to constrain the \emph{projected} shape, orientation, and radial profile of mass \citep[as done by the Lens Structure and Dynamics Survey described in][]{koopmans02}.
+Of course, lensing reconstructions can only be performed for galaxies which closely intersect our line of sight to background sources, but the advent of large photometric catalogues has allowed automatic searches for such chance alignments and significant increases in the number of objects studied in this way \citep[e.g.
+the Sloan Lens ACS Survey, see][]{bolton06}.
 
-[something about measuring halo shape with lensing].
+Within the Milky Way we have a unique vantage point that allows us to measure the three-dimensional structure of our own dark matter halo.
+One approach to ...
+[explain ``random'' methods: kinematic tracers].
+Another approach is to exploit the non-random nature of the stellar populations in the halo.
+A consequence of hierarchical structure formation [...] and such structures have been observed [e.g., ...cite belokurov?].
+Tidal streams are dynamically cold systems --- debris typically have small distributions of energy and angular momenta --- and thus require orders of magnitude fewer tracers than a random sample to get constraints of comparable accuracy.
+For example, in the simplest case we might {\it assume} that debris stars are actually still on the same orbit as their progenitor system \citep[a \emph{wrong} assumption, see e.g.][]{eyre}.
+Then we can imagine measuring the full-space velocities ${\bf v}$ at different points ${\bf x}$ in the structure (e.g.
+along a stream), which would give us a direct measure of differences in a potential, $\Phi$.
 
-[why use tidal streams?] \cite{lm10} used N-body models to ... triaxial.
-
-Another prediction of hierarchical structure formation in $\Lambda$CDM is the presence of substructure.
+Since the initial discovery of tidal debris from Sgr, myriad other streams and debris systems have been discovered in large photometric surveys probing Galactocentric distances from $\sim$15-100 kpc.
+This has motivated numerous efforts for developing methods to constrain properties of the Galactic potential with cold debris structures in the Milky Way halo.
 
 - Broader comparison to other methods\\
+\cite[][hereafter LM10]{lm10} ran N-body models of the disruption of the Sagittarius (Sgr) dwarf spheroidal galaxy and used these simulations to constrain the shape of the Milky Way potential out to $\sim$70 kpc [check number], finding that the halo ...
+(though, see ...
+for a discussion of unexplained features in the Sgr debris).
 \indent\indent- 'generative methods' more generally (ie N-body, plus Varghese, plus Kuepper plus new possible action-angle models....\\
@@
 \indent\indent- Reference Deg \& Widrow to say they illustrate the idea of the power of multiple streams to back up the bit where you say that's the long-term aim\\
+
+[Another prediction of hierarchical structure formation in $\Lambda$CDM is the presence of substructure. ]
 

### 2014-02-08 text/paper2/ms.tex
model section

@@
 \newcommand{\rtide}{R_{\rm tide}}
+
+\newcommand{\bsr}{\boldsymbol r}
+\newcommand{\bsv}{\boldsymbol v}
 
@@
 
+[Conceptual description of the model]
+
 Take $\D$ to be the observed, heliocentric 6D position of a star --- e.g., the measured position on the sky, ($l$, $b$); distance, $D$; proper motions, ($\mu_l$, $\mu_b$); and line-of-sight velocity, $v_r$ --- and take $\W$ to be the corresponding \emph{true}, 6D position of the star in the same coordinate and reference frame.
@@
 The position of the satellite is observed to be at heliocentric 6D position $\D_s$, with true position $\W_s$ (where any subscript $s$ refers to the satellite).
-The true 6D spherical, heliocentric positions of the star and satellite -- $\W$ and $\W_s$ -- have corresponding 6D cartesian, Galactocentric coordinate vectors $\X$ and $\X_s$.
+The true 6D heliocentric spherical positions of the star and satellite -- $\W$ and $\W_s$ -- have corresponding 6D Galactocentric cartesian coordinate vectors $\X$ and $\X_s$ comprised of position and velocity vectors ($\bsr$,$\bsv$) and ($\bsr_s$,$\bsv_s$).
 Initially we assume the present-day mass of the satellite, $M_s$, and velocity dispersion, $\sigma_v$, have been measured perfectly and neglect mass-loss from the satellite.
@@
 
-The disruption of a satellite galaxy due to tidal shocking [...] Thus, we model the present-day position and velocity of the disrupted star as the result of sampling from isotropic, log-normal distributions centered on the instantaneous tidal radius, $R_{\rm tide}$, and velocity dispersion, $\sigma_v$, of the satellite at some time $\tau$ in the past, evolved to present day in the parent potential, $\Phi$.
+The disruption of a satellite galaxy due to tidal shocking [...connect to Choi work some how?] We model the present-day position and velocity of a disrupted star as the result of sampling from isotropic, log-normal distributions in relative distance, $R=\left\vert \bsr-\bsr_s \right\vert$, and relative velocity, $V=\left\vert \bsv-\bsv_s \right\vert$, at the time the star comes unbound from the satellite, evolved to present day in the parent potential, $\Phi$.
+The distributions over $\ln R$ and $\ln V$ are centered on the logarithm of the instantaneous tidal radius, $R_{\rm tide}$, and velocity dispersion, $\sigma_v$, of the satellite at time $\tau$: $\mathcal{N}(\ln R \given \ln\rtide)\vert_{t=\tau}$ and $\mathcal{N}(\ln V \given \ln \sigma_v)\vert_{t=\tau}$.
 In this model, we neglect the effect of self-interaction with the halo of the progenitor [TODO: maybe we need to justify this, add APW's figure from paper 1 ref.
@@
 refer to another section].
-The likelihood of a star's position today is computed by first transforming the positions to Galactocentric, cartesian coordinates and integrating the orbits of the satellite and star backwards for several Gigayears in the parent potential, $\Phi$, treating both as test particles.
-The distance, $R$, and velocity, $V$, of the star relative to the satellite are computed at each time-step, and the likelihood is computed as the product of an isotropic, log-normal centered on the tidal radius with an isotropic, log-normal centered on the velocity dispersion of the satellite:
+
+The likelihood of a star's position today is computed by first transforming the observed position to Galactocentric cartesian coordinates and integrating the orbits of the satellite and star backwards for several Gigayears in the parent potential, $\Phi$, treating both as test particles.
+The distance, $R$, and velocity, $V$, of the star relative to the satellite are computed at each time-step, and the likelihood is computed as the product of the isotropic, log-normal centered on the tidal radius with the isotropic, log-normal centered on the velocity dispersion of the satellite:
 \begin{align}
@@
 	p(\X \given \X_s, \tau, \Phi) &= p(\ln R, \ln V \given \tau, \Phi) \jac{(x,y,z,v_x,v_y,v_z)}\label{eq:int} \\
-	p(\ln R, \ln V \given \tau, \Phi) &= \mathcal{N}(\ln R(\tau) | \ln \rtide(\tau))\mathcal{N}(\ln V(\tau) | \ln \sigma_v(\tau))
+	p(\ln R, \ln V \given \tau, \Phi) &= \mathcal{N}(\ln R | \ln \rtide)\mathcal{N}(\ln V | \ln \sigma_v)\vert_{t=\tau}
 \end{align}
+% , but the vectors $\D$ and $\W$ [exist] in different coordinate systems --- heliocentric spherical and Galactocentric cartesian, respectively --- so we must also track the Jacobian that [describes] this transformation:
 where in eq.~\ref{eq:int} we have implicitly integrated over and neglected the two angle coordinates associated with $\ln R$ and the two angles associated with $\ln V$.
-
-% , but the vectors $\D$ and $\W$ [exist] in different coordinate systems --- heliocentric spherical and Galactocentric cartesian, respectively --- so we must also track the Jacobian that [describes] this transformation:
-
-In general, $p(\tau)$ is likely conditional on the observed orbit of the progenitor---stars are preferentially stripped at pericentric passages \citep{??}---but this is neglected and assumed to be uniform over the entire interaction history between the satellite and parent.
-Similarly, the priors on potential parameters, $p(\Phi)$, are taken to be uniform.
+We then compute the marginal likelihood, $p(\W \given \W_s, \Phi)$, by integrating over all possible unbinding times,
+\begin{align}
+	p(\W \given \W_s, \Phi) = \int p(\W \given \W_s, \tau, \Phi)p(\tau) d\tau.
+\end{align}
+In general, $p(\tau)$ is likely conditional on the orbit of the progenitor --- stars are preferentially stripped at pericentric passages --- but this is neglected and is taken to be uniform over the entire interaction history between the satellite and parent.
 

### 2014-02-10 text/paper2/ms.tex
added more intro text

@@
 This extra knowledge about the orbital distribution of the baryonic matter within a disk allows a more precise tracing of the potential than is feasible for objects such as elliptical galaxies where orbits are assumed to be random.
-[Potential is most likely smoothed out (energetic feedback?), closely integrable allowing beautiful dynamical ...]
+[Potential is most likely smoothed out (energetic feedback), can approximate as closely integrable allowing beautiful dynamical methods ...]
 
@@
 A consequence of hierarchical structure formation [...] and such structures have been observed [e.g., ...cite belokurov?].
-Tidal streams are dynamically cold systems --- debris typically have small distributions of energy and angular momenta --- and thus require orders of magnitude fewer tracers than a random sample to get constraints of comparable accuracy.
+Tidal streams are dynamically cold systems --- debris typically have small distributions of energy and angular momentum --- and thus require orders of magnitude fewer tracers than a random sample to get constraints of comparable accuracy.
 For example, in the simplest case we might {\it assume} that debris stars are actually still on the same orbit as their progenitor system \citep[a \emph{wrong} assumption, see e.g.][]{eyre}.
@@
 along a stream), which would give us a direct measure of differences in a potential, $\Phi$.
+In practice, simple methods such as orbit fitting lead to systematic biases in inferred properties of the underlying potential because streams do not strictly delineate orbits [cite eyre binney work].
 
 Since the initial discovery of tidal debris from Sgr, myriad other streams and debris systems have been discovered in large photometric surveys probing Galactocentric distances from $\sim$15-100 kpc.
-This has motivated numerous efforts for developing methods to constrain properties of the Galactic potential with cold debris structures in the Milky Way halo.
+The known debris structures are $\gtrsim$10 kpc from the Sun, thus while overdensities in the halo are less contaminated by background, the constituent stars often have incomplete or poorly measured phase-space information.
+Perhaps the most rigorous method of modeling streams and the potential is to run full N-body simulations of satellite disruption and compare the density of observed stars to simulated stars.
+Such a model could naturally incorporate observational uncertainties, missing dimensions, and would, in principle, allow for studying any arbitrary time-dependent or non-integrable potential.
+However, such a model is presently computationally intractable [though see milkyway@home stuff...].
+\cite[][hereafter LM10]{lm10} ran a grid of N-body models of the disruption of the Sagittarius (Sgr) dwarf spheroidal galaxy and matched their simulated streams to all Sgr data available at that time.
+By comparing the phase-space positions of their stream particles with the data, they were able to use these simulations to constrain the shape of the Milky Way potential out to $\sim$70 kpc [check number], finding that the halo ...
+(though, see ...
+for a discussion of unexplained features in the Sgr debris).
+
+The difficulty and computational costs associated with proper N-body modeling has motivated numerous efforts for developing approximate methods for constraining properties of the Galactic potential with cold debris structures.
 
 - Broader comparison to other methods\\
-\cite[][hereafter LM10]{lm10} ran N-body models of the disruption of the Sagittarius (Sgr) dwarf spheroidal galaxy and used these simulations to constrain the shape of the Milky Way potential out to $\sim$70 kpc [check number], finding that the halo ...
-(though, see ...
-for a discussion of unexplained features in the Sgr debris).
 \indent\indent- 'generative methods' more generally (ie N-body, plus Varghese, plus Kuepper plus new possible action-angle models....\\

### 2014-02-10 scripts/figures/paper2_figures.py
added a figure showing potential contours

@@ -28,7 +28,7 @@ from scipy.stats import norm
 from streams.util import project_root
 from streams.io.sgr import SgrSimulation
 from streams.integrate import LeapfrogIntegrator
-from streams.potential.lm10 import LawMajewski2010
+from streams.potential.lm10 import LawMajewski2010, LawMajewski2010Py
 
 matplotlib.rc('xtick', labelsize=18)
 matplotlib.rc('ytick', labelsize=18)
@@ -74,6 +74,47 @@ def graphical_model(**kwargs):
     pgm.render()
     #pgm.figure.savefig("classic.pdf")
     pgm.figure.savefig(filename, dpi=150)
+
+def potential_contours(**kwargs):
+
+    filename = os.path.join(plot_path, "potentials.pdf")
+    fig,axes = plt.subplots(2,2,figsize=(8,8), sharex=True, sharey=True)
+
+    potentials = [LawMajewski2010Py(),
+                  LawMajewski2010Py(q1=0.7),
+                  LawMajewski2010Py(qz=0.7),
+                  LawMajewski2010Py(phi=45*u.degree)]
+
+    grid = np.linspace(-75, 75, 250)
+    X1, X2 = np.meshgrid(grid,grid)
+
+    r = np.array([np.zeros_like(X1.ravel()).tolist() \
+                   for xx in range(3)])
+    r[1] = X1.ravel()
+    r[2] = X2.ravel()
+
+    levels = None
+    for ii,potential in enumerate(potentials):
+        Z = potential._value_at(r.T).reshape(X1.shape)
+        if levels is None:
+            cs = axes.flat[ii].contourf(X1, X2, Z, cmap=cm.Blues_r)
+            levels = cs.levels
+        else:
+            cs = axes.flat[ii].contourf(X1, X2, Z, cmap=cm.Blues_r, levels=levels)
+
+    axes[0,0].text(-60, 50, "LM10", fontweight=400)
+    axes[0,1].text(-60, 50, r"$q_1=0.7$")
+    axes[1,0].text(-60, 50, r"$q_z=0.7$")
+    axes[1,1].text(-60, 50, r"$\phi=45^\circ$")
+
+    axes[0,0].set_ylabel("Z [kpc]")
+    axes[1,0].set_ylabel("Z [kpc]")
+
+    axes[1,0].set_xlabel("Y [kpc]")
+    axes[1,1].set_xlabel("Y [kpc]")
+
[…]

### 2014-02-10 streams/potential/lm10.py

@@ -212,3 +212,69 @@ class LawMajewski2010(CompositePotential):
         r_unit = filter(lambda x: x.is_equivalent(u.km), self.units)[0]
         t_unit = filter(lambda x: x.is_equivalent(u.s), self.units)[0]
         return v_esc * r_unit/t_unit
+
+class LawMajewski2010Py(CompositePotential):
+
+    def __init__(self, **parameters):
+        """ Represents the functional form of the Galaxy potential used by
+            Law and Majewski 2010.
+
+            Miyamoto-Nagai disk
+            Hernquist bulge
+            Logarithmic halo
+
+            Model parameters: q1, qz, phi, v_halo
+
+            Parameters
+            ----------
+            parameters : dict
+                A dictionary of parameters for the potential definition.
+        """
+
+        # v_halo range comes from 5E11 < M < 5E12, current range of MW mass @ 200 kpc
+        lm10_parameters = { 'q1' : PotentialParameter(truth=1.38,
+                                              range=(1.1, 1.7),
+                                              latex=r"$q_1$"),
+                    'q2' : PotentialParameter(truth=1.,
+                                              range=(0.7, 1.3),
+                                              latex=r"$q_2$"),
+                    'qz' : PotentialParameter(truth=1.36,
+                                              range=(1.1, 1.7),
+                                              latex=r"$q_z$"),
+                    'phi' : PotentialParameter(truth=97.*u.deg,
+                                               range=(85.*u.deg,115.*u.deg),
+                                               latex=r"$\phi$"),
+                    'v_halo' : PotentialParameter(truth=121.858*u.km/u.s,
+                                                  range=(100.*u.km/u.s,
+                                                         150.*u.km/u.s),
+                                                  latex=r"$v_{\rm halo}$"),
+                    'R_halo' : PotentialParameter(truth=12.*u.kpc,
+                                                  range=(8.*u.kpc, 16*u.kpc),
+                                                  latex=r"$R_{\rm halo}$")}
+
+        self.parameters = dict(lm10_parameters)
+        for k,v in self.parameters.items():
+            self.__dict__[k] = v
+
+        for p_name in parameters.keys():
+            if hasattr(parameters[p_name], "unit"):
+                self.parameters[p_name].value = parameters[p_name]
+            else:
+                self.parameters[p_name]._value = parameters[p_name]
+
+        bulge = HernquistPotential(usys,
+                                   m=3.4E10*u.M_sun,
[…]

### 2014-02-10 text/paper2/ms.tex
experiments

@@
 The position of the satellite is observed to be at heliocentric 6D position $\D_s$, with true position $\W_s$ (where any subscript $s$ refers to the satellite).
-The true 6D heliocentric spherical positions of the star and satellite -- $\W$ and $\W_s$ -- have corresponding 6D Galactocentric cartesian coordinate vectors $\X$ and $\X_s$ comprised of position and velocity vectors ($\bsr$,$\bsv$) and ($\bsr_s$,$\bsv_s$).
+The true 6D heliocentric spherical positions of the star and satellite -- $\W$ and $\W_s$ -- have corresponding 6D Galactocentric cartesian coordinate vectors $\X$ and $\X_s$ comprised of 3D position and velocity vectors ($\bsr$,$\bsv$) and ($\bsr_s$,$\bsv_s$).
 Initially we assume the present-day mass of the satellite, $M_s$, and velocity dispersion, $\sigma_v$, have been measured perfectly and neglect mass-loss from the satellite.
-With this notation, we write the joint posterior probability for the observed position of a star, satellite, and parameters for the parent potential as:
+With this notation, we write the joint posterior probability for the observed position of a star, satellite, and parent potential parameters as:
 \begin{align}
@@
 \end{align}
-where the covariance matrices $\bSigma$ and $\bSigma_s$ specify the observational variances on the observed 6D position of star and the satellite, respectively.
+where the covariance matrices $\bSigma$ and $\bSigma_s$ specify the observational uncertainties on the observed 6D position of star and the satellite, respectively.
 
@@
 The distributions over $\ln R$ and $\ln V$ are centered on the logarithm of the instantaneous tidal radius, $R_{\rm tide}$, and velocity dispersion, $\sigma_v$, of the satellite at time $\tau$: $\mathcal{N}(\ln R \given \ln\rtide)\vert_{t=\tau}$ and $\mathcal{N}(\ln V \given \ln \sigma_v)\vert_{t=\tau}$.
-In this model, we neglect the effect of self-interaction with the halo of the progenitor [TODO: maybe we need to justify this, add APW's figure from paper 1 ref.
-response?
-refer to another section].
+In this model, we neglect the effect of self-interaction with the halo of the progenitor.
 
@@
 	p(\X \given \X_s, \tau, \Phi) &= p(\ln R, \ln V \given \tau, \Phi) \jac{(x,y,z,v_x,v_y,v_z)}\label{eq:int} \\
-	p(\ln R, \ln V \given \tau, \Phi) &= \mathcal{N}(\ln R | \ln \rtide)\mathcal{N}(\ln V | \ln \sigma_v)\vert_{t=\tau}
+	p(\ln R, \ln V \given \tau, \Phi) &= \left[\mathcal{N}(\ln R | \ln \rtide)\mathcal{N}(\ln V | \ln \sigma_v)\right]_{t=\tau}
 \end{align}
@@
 \section{Experiments}
-We conduct several tests of this method [...].
-For all tests in this article, we model the potential of the parent galaxy as a three-component sum of a Miyamoto-Nagai disk \citep{}, Hernquist bulge \citep[spheroid][]{}, and a triaxial, logarithmic halo \citep[e.g.,][]{law10}:
+We conduct several tests of this method using N-body simulations run with an SCF code [...]. [KVJ: describe simulations here?]
+
+For all tests in this article, we model the potential of the parent galaxy as a three-component sum of a Miyamoto-Nagai disk \citep{}, Hernquist bulge \citep[spheroid;][]{}, and a triaxial, logarithmic halo \citep[e.g.,][]{law10}:
 \begin{align}
@@
 
-[TODO: explain simulated data we use - KVJ are you ok releasing data from your 2.5e6-2.5e9 Sgr sims? Explain that we use the LM10 potential in Table~\ref{tbl:potential} for the sims].
-
 When recovering the potential, we hold fixed the disk and spheroid parameters (see Table~\ref{tbl:potential}), along with two halo parameters: the scale radius, $r_h$, and one of the flattening components in the x-y plane, $q_2$.
 These halo parameters are degenerate with combinations of the other parameters --- e.g., setting $q_1=2.0$, $q_2=1.0$, $\phi=0^\circ$ is equivalent to setting $q_1=1.0$, $q_2=2.0$, $\phi=90^\circ$.
-The priors on the remaining halo parameters are taken to be uniform over a conservative domain of realistic values: for $v_h$, 100-200~kms$^{-1}$ corresponds to a range in solar circular velocities from $\sim$210-250 kms$^{-1}$ (holding other parameters fixed); the axis ratios [...]; and $\phi$ is restricted to $\pm45^\circ$ around the value measured by LM10 ($\phi = 97^\circ$).
+The priors on the remaining halo parameters are taken to be uniform over a conservative domain of realistic values: for $v_h$, 100-200~kms$^{-1}$ corresponds to a range in solar circular velocities from $\sim$210-250 kms$^{-1}$ (holding other parameters fixed); the range in axis ratios allow for prolate, oblate, and mixed triaxiality; and $\phi$ is restricted to $\pm45^\circ$ around the value measured by LM10 ($\phi = 97^\circ$).
+Figure~\ref{fig:potential} shows equipotential slices in the Galactic Y-Z plane (X=0) for various choices of [...].
 

### 2014-02-10 text/paper2/ms.tex
added figure

@@
 The priors on the remaining halo parameters are taken to be uniform over a conservative domain of realistic values: for $v_h$, 100-200~kms$^{-1}$ corresponds to a range in solar circular velocities from $\sim$210-250 kms$^{-1}$ (holding other parameters fixed); the range in axis ratios allow for prolate, oblate, and mixed triaxiality; and $\phi$ is restricted to $\pm45^\circ$ around the value measured by LM10 ($\phi = 97^\circ$).
-Figure~\ref{fig:potential} shows equipotential slices in the Galactic Y-Z plane (X=0) for various choices of [...].
+Figure~\ref{fig:potential} shows equipotential slices in the Galactic X-Z plane (Y=0) and Y-Z plane (X=0) for a few choices of  $q_1$,  $q_z$, and $\phi$ while holding all other parameters fixed, as described in the figure caption.
+
+\begin{figure}[h]
+\begin{center}
+%\includegraphics[width=0.45\textwidth]{fig4.eps}
+\includegraphics[width=\textwidth]{/Users/adrian/projects/streams/plots/paper2/potentials.pdf}
+\caption{ TODO }\label{fig:potential}
+\end{center}
+\end{figure}
 

### 2014-02-10 scripts/figures/paper2_figures.py

@@ -78,42 +78,89 @@ def graphical_model(**kwargs):
 def potential_contours(**kwargs):
 
     filename = os.path.join(plot_path, "potentials.pdf")
-    fig,axes = plt.subplots(2,2,figsize=(8,8), sharex=True, sharey=True)
+    fig,axes = plt.subplots(2,4,figsize=(14,7))
 
-    potentials = [LawMajewski2010Py(),
-                  LawMajewski2010Py(q1=0.7),
-                  LawMajewski2010Py(qz=0.7),
-                  LawMajewski2010Py(phi=45*u.degree)]
+    base_params = dict(q1=1., qz=1., q2=1., phi=0.)
+    potentials = []
+    potentials.append(LawMajewski2010Py(**base_params))
+
+    pp = base_params.copy()
+    pp['q1'] = 1.5
+    potentials.append(LawMajewski2010Py(**pp))
+    axes[0,1].text(0.5, 1.04, r"$q_1=1.5$",
+                   horizontalalignment='center',
+                   fontsize=20,
+                   transform=axes[0,1].transAxes)
+
+    pp = base_params.copy()
+    pp['qz'] = 1.5
+    potentials.append(LawMajewski2010Py(**pp))
+    axes[0,2].text(0.5, 1.04, r"$q_z=1.5$",
+                   horizontalalignment='center',
+                   fontsize=20,
+                   transform=axes[0,2].transAxes)
+
+    pp = base_params.copy()
+    pp['phi'] = 45*u.degree
+    pp['q1'] = 1.5
+    potentials.append(LawMajewski2010Py(**pp))
+    axes[0,3].text(0.5, 1.04, r"$q_1=1.5$, $\phi=45^\circ$",
+                   horizontalalignment='center',
+                   fontsize=20,
+                   transform=axes[0,3].transAxes)
 
     grid = np.linspace(-75, 75, 250)
     X1, X2 = np.meshgrid(grid,grid)
 
+    # top row:
     r = np.array([np.zeros_like(X1.ravel()).tolist() \
                    for xx in range(3)])
-    r[1] = X1.ravel()
-    r[2] = X2.ravel()
+    r[0] = X1.ravel()
+    r[1] = X2.ravel()
 
     levels = None
     for ii,potential in enumerate(potentials):
[…]

### 2014-02-10 streams/inference/core.py
missing progenitor

@@ -127,7 +127,9 @@ class StreamModel(object):
 
             logger.debug("Satellite properties added as parameters:")
             if config['satellite']['parameters'].has_key('_X'):
-                priors = [LogNormalPrior(satellite._X[0],satellite._error_X[0])]
+                #priors = [LogNormalPrior(satellite._X[0],satellite._error_X[0])]
+                priors=[LogUniformPrior(np.array([-np.pi,-np.pi/2.,10,-0.02424,-0.02424,-0.4091]),
+                                        np.array([np.pi,np.pi/2.,100,0.02424,0.02424,0.4091]))]
                 s_X = ModelParameter('_X', value=satellite._X, prior=priors,
                                         truth=true_satellite._X)
                 model.add_parameter('satellite', s_X)
@@ -199,10 +201,10 @@ class StreamModel(object):
                         std = np.ravel([pr.sigma for pr in param._prior])
                         p0[ii,ix1:ix1+param.size] = np.random.normal(_X, std)
 
-                    if group_name == "satellite" and param_name == "_X":
-                        _X = self.true_satellite._X.ravel()
-                        std = np.ravel([pr.sigma for pr in param._prior])
-                        p0[ii,ix1:ix1+param.size] = np.random.normal(_X, std)
+                    # if group_name == "satellite" and param_name == "_X":
+                    #     _X = self.true_satellite._X.ravel()
+                    #     std = np.ravel([pr.sigma for pr in param._prior])
+                    #     p0[ii,ix1:ix1+param.size] = np.random.normal(_X, std)
 
                     ### HACK TO INITIALIZE WALKERS NEAR TRUTH
                     #######################################################

### 2014-02-10 text/paper2/ms.tex
cleanup

@@
 \begin{align}
-	p(\W \given \W_s, \tau, \Phi) &= p(\X \given \X_s, \tau, \Phi) \jac{(l,b,D,\mu_l,\mu_b,v_r)}\\
-	p(\X \given \X_s, \tau, \Phi) &= p(\ln R, \ln V \given \tau, \Phi) \jac{(x,y,z,v_x,v_y,v_z)}\label{eq:int} \\
+	p(\W \given \W_s, \tau, \Phi) &= p(\X \given \X_s, \tau, \Phi) \jac{(l,b,D,\mu_l,\mu_b,v_r)}_{t=\tau}\\
+	p(\X \given \X_s, \tau, \Phi) &= p(\ln R, \ln V \given \tau, \Phi) \jac{(x,y,z,v_x,v_y,v_z)}_{t=\tau}\label{eq:int} \\
 	p(\ln R, \ln V \given \tau, \Phi) &= \left[\mathcal{N}(\ln R | \ln \rtide)\mathcal{N}(\ln V | \ln \sigma_v)\right]_{t=\tau}
@@
 
-When recovering the potential, we hold fixed the disk and spheroid parameters (see Table~\ref{tbl:potential}), along with two halo parameters: the scale radius, $r_h$, and one of the flattening components in the x-y plane, $q_2$.
-These halo parameters are degenerate with combinations of the other parameters --- e.g., setting $q_1=2.0$, $q_2=1.0$, $\phi=0^\circ$ is equivalent to setting $q_1=1.0$, $q_2=2.0$, $\phi=90^\circ$.
-The priors on the remaining halo parameters are taken to be uniform over a conservative domain of realistic values: for $v_h$, 100-200~kms$^{-1}$ corresponds to a range in solar circular velocities from $\sim$210-250 kms$^{-1}$ (holding other parameters fixed); the range in axis ratios allow for prolate, oblate, and mixed triaxiality; and $\phi$ is restricted to $\pm45^\circ$ around the value measured by LM10 ($\phi = 97^\circ$).
-Figure~\ref{fig:potential} shows equipotential slices in the Galactic X-Z plane (Y=0) and Y-Z plane (X=0) for a few choices of  $q_1$,  $q_z$, and $\phi$ while holding all other parameters fixed, as described in the figure caption.
-
 \begin{figure}[h]
@@
 \end{figure}
+
+When recovering the potential, we hold fixed the disk and spheroid parameters (see Table~\ref{tbl:potential}), along with two halo parameters: the scale radius, $r_h$, and one of the flattening components in the x-y plane, $q_2$.
+These halo parameters are degenerate with combinations of the other parameters --- e.g., setting $q_1=2.0$, $q_2=1.0$, $\phi=0^\circ$ is equivalent to setting $q_1=1.0$, $q_2=2.0$, $\phi=90^\circ$.
+The priors on the remaining halo parameters are taken to be uniform over a conservative domain of realistic values: for $v_h$, 100-200~kms$^{-1}$ corresponds to a range in solar circular velocities from $\sim$210-250 kms$^{-1}$ (holding other parameters fixed); the range in axis ratios allow for prolate, oblate, and mixed triaxiality; and $\phi$ is restricted to $\pm45^\circ$ around the value measured by LM10 ($\phi = 97^\circ$).
+Figure~\ref{fig:potential} shows equipotential slices in the Galactic X-Z plane (Y=0) and Y-Z plane (X=0) for a few choices of  $q_1$,  $q_z$, and $\phi$ while holding all other parameters fixed, as described in the figure caption.
 
@@
 \subsection{\gaia + \spitzer}
-64 stars in Sgr-like stream
+4 stars in Sgr-like stream
 
 \subsection{\spitzer + current data}
-64 stars in Sgr-like stream
+4 stars in Sgr-like stream
 
 \subsection{\spitzer + current data, no progenitor}
-32 stars in Orphan-like stream, XX stars in GD-1 -like stream
+4 stars in Orphan-like stream, XX stars in GD-1 -like stream
 
@@
 \acknowledgements
-Binney, Rix, Sanders
 

### 2014-02-11 text/paper2/ms.tex
qualitative

@@
 \keywords{
+  cosmology: dark matter
+  ---
   Galaxy: structure
@@
   Galaxy: halo
-  ---
-  cosmology: dark matter
 }
@@
 However, such a model is presently computationally intractable [though see milkyway@home stuff...].
-\cite[][hereafter LM10]{lm10} ran a grid of N-body models of the disruption of the Sagittarius (Sgr) dwarf spheroidal galaxy and matched their simulated streams to all Sgr data available at that time.
+\cite[][hereafter LM10]{lm10} ran a grid of N-body models of the disruption of the Sagittarius (Sgr) dwarf spheroidal galaxy and qualitatively matched their simulated streams to all Sgr data available at that time.
 By comparing the phase-space positions of their stream particles with the data, they were able to use these simulations to constrain the shape of the Milky Way potential out to $\sim$70 kpc [check number], finding that the halo ...

### 2014-02-12 text/paper2/ms.tex
change to _p for progenitor

@@
 \newcommand{\bsigma}{\boldsymbol\sigma}
-\newcommand{\rtide}{R_{\rm tide}}
+\newcommand{\rtide}{r_{\rm tide}}
 
@@
 \begin{abstract}
+Context:
+Aims: 
+Methods:
+Results:
+Conclusions:
+
 %Informative phase-space structures, and in particular those created by
@@
 However, such a model is presently computationally intractable [though see milkyway@home stuff...].
-\cite[][hereafter LM10]{lm10} ran a grid of N-body models of the disruption of the Sagittarius (Sgr) dwarf spheroidal galaxy and qualitatively matched their simulated streams to all Sgr data available at that time.
-By comparing the phase-space positions of their stream particles with the data, they were able to use these simulations to constrain the shape of the Milky Way potential out to $\sim$70 kpc [check number], finding that the halo ...
+\cite{law10} ran several N-body models of the disruption of the Sagittarius (Sgr) dwarf spheroidal galaxy and qualitatively matched their simulated streams to all Sgr data available at that time.
+By comparing the phase-space positions of their stream particles with the data, they were able to use these simulations to constrain the shape of the Milky Way potential out to $\sim$70 kpc [check number] and found that the best-fitting halo has a peculiar triaxial shape elongated in the Galactic Z direction, orthogonal to the plane of the disk.
 (though, see ...
-for a discussion of unexplained features in the Sgr debris).
+for possible complications in the Sgr debris belokurov 2012?).
 
@@
 
-[Conceptual description of the model]
+\begin{figure}[h]
+\begin{center}
+%\includegraphics[width=0.45\textwidth]{fig4.eps}
+\includegraphics[width=\textwidth]{/Users/adrian/projects/streams/plots/paper2/rel_dists.pdf}
+\caption{ TODO }\label{fig:reldist}
+\end{center}
+\end{figure}
+
+[Conceptual description of the model] Figure~\ref{fig:reldist}
 
@@
 Assume that we have determined that this star was once part of a progenitor system (e.g., satellite galaxy) tidally disrupting and forming a cold debris structure in the potential, $\Phi$, of some parent galaxy.
-The position of the satellite is observed to be at heliocentric 6D position $\D_s$, with true position $\W_s$ (where any subscript $s$ refers to the satellite).
-The true 6D heliocentric spherical positions of the star and satellite -- $\W$ and $\W_s$ -- have corresponding 6D Galactocentric cartesian coordinate vectors $\X$ and $\X_s$ comprised of 3D position and velocity vectors ($\bsr$,$\bsv$) and ($\bsr_s$,$\bsv_s$).
-Initially we assume the present-day mass of the satellite, $M_s$, and velocity dispersion, $\sigma_v$, have been measured perfectly and neglect mass-loss from the satellite.
-With this notation, we write the joint posterior probability for the observed position of a star, satellite, and parent potential parameters as:
+The position of the progenitor is observed to be at heliocentric 6D position $\D_p$, with true position $\W_p$ (where any subscript $p$ refers to the progenitor).
+The true 6D heliocentric spherical positions of the star and progenitor -- $\W$ and $\W_p$ -- have corresponding 6D Galactocentric cartesian coordinate vectors $\X$ and $\X_p$ comprised of 3D position and velocity vectors ($\bsr$,$\bsv$) and ($\bsr_p$,$\bsv_p$).
+Initially we assume the present-day mass of the progenitor, $M_p$, and velocity dispersion, $\sigma_v$, have been measured perfectly and neglect mass-loss from the progenitor.
+With this notation, we write the joint posterior probability for the observed position of a star, progenitor, and parent potential parameters as:
 \begin{align}
-	p(\Phi, \W, \W_s, \tau \given \D, \D_s) &= \frac{1}{\mathcal{Z}} p(\D \given \W) p(\D_s \given \W_s) 
-												       p(\W \given \W_s, \tau, \Phi) 
+	p(\Phi, \W, \W_p, \tau \given \D, \D_p) &= \frac{1}{\mathcal{Z}} p(\D \given \W) p(\D_p \given \W_p) 
+												       p(\W \given \W_p, \tau, \Phi) 
 												       p(\Phi) p(\tau)
@@
 	p(\D \given \W) &= \mathcal{N}(\W \given \D, \bSigma)\\
-	p(\D_s \given \W_s) &= \mathcal{N}(\W_s \given \D_s, \bSigma_s)
+	p(\D_p \given \W_p) &= \mathcal{N}(\W_p \given \D_p, \bSigma_p)
 \end{align}
-where the covariance matrices $\bSigma$ and $\bSigma_s$ specify the observational uncertainties on the observed 6D position of star and the satellite, respectively.
+where the covariance matrices $\bSigma$ and $\bSigma_p$ specify the observational uncertainties on the observed 6D position of star and the progenitor, respectively.
 
-The disruption of a satellite galaxy due to tidal shocking [...connect to Choi work some how?] We model the present-day position and velocity of a disrupted star as the result of sampling from isotropic, log-normal distributions in relative distance, $R=\left\vert \bsr-\bsr_s \right\vert$, and relative velocity, $V=\left\vert \bsv-\bsv_s \right\vert$, at the time the star comes unbound from the satellite, evolved to present day in the parent potential, $\Phi$.
-The distributions over $\ln R$ and $\ln V$ are centered on the logarithm of the instantaneous tidal radius, $R_{\rm tide}$, and velocity dispersion, $\sigma_v$, of the satellite at time $\tau$: $\mathcal{N}(\ln R \given \ln\rtide)\vert_{t=\tau}$ and $\mathcal{N}(\ln V \given \ln \sigma_v)\vert_{t=\tau}$.
+The disruption of a satellite galaxy due to tidal shocking [...connect to Choi work some how?] We model the present-day position and velocity of a disrupted star as the result of sampling from isotropic, log-normal distributions in relative distance, $R=\left\vert \bsr-\bsr_p \right\vert$, and relative velocity, $V=\left\vert \bsv-\bsv_p \right\vert$, at the time the star comes unbound from the progenitor, evolved to present day in the parent potential, $\Phi$.
+The distributions over $\ln R$ and $\ln V$ are centered on the logarithm of the instantaneous tidal radius, $\rtide$, and velocity dispersion, $\sigma_v$, of the progenitor at time $\tau$: $\mathcal{N}(\ln R \given \ln\rtide)\vert_{t=\tau}$ and $\mathcal{N}(\ln V \given \ln \sigma_v)\vert_{t=\tau}$.
 In this model, we neglect the effect of self-interaction with the halo of the progenitor.
 
-The likelihood of a star's position today is computed by first transforming the observed position to Galactocentric cartesian coordinates and integrating the orbits of the satellite and star backwards for several Gigayears in the parent potential, $\Phi$, treating both as test particles.
-The distance, $R$, and velocity, $V$, of the star relative to the satellite are computed at each time-step, and the likelihood is computed as the product of the isotropic, log-normal centered on the tidal radius with the isotropic, log-normal centered on the velocity dispersion of the satellite:
+The likelihood of a star's position today is computed by first transforming the observed position to Galactocentric cartesian coordinates and integrating the orbits of the progenitor and star backwards for several Gigayears in the parent potential, $\Phi$, treating both as test particles.
+The distance, $R$, and velocity, $V$, of the star relative to the progenitor are computed at each time-step, and the likelihood is computed as the product of the isotropic, log-normal centered on the logarithm of the instantaneous tidal radius with the isotropic, log-normal centered on the logarithm of the velocity dispersion of the progenitor:
 \begin{align}
-	p(\W \given \W_s, \tau, \Phi) &= p(\X \given \X_s, \tau, \Phi) \jac{(l,b,D,\mu_l,\mu_b,v_r)}_{t=\tau}\\
-	p(\X \given \X_s, \tau, \Phi) &= p(\ln R, \ln V \given \tau, \Phi) \jac{(x,y,z,v_x,v_y,v_z)}_{t=\tau}\label{eq:int} \\
-	p(\ln R, \ln V \given \tau, \Phi) &= \left[\mathcal{N}(\ln R | \ln \rtide)\mathcal{N}(\ln V | \ln \sigma_v)\right]_{t=\tau}
+	p(\W \given \W_p, \tau, \Phi) &= p(\X \given \X_p, \tau, \Phi) \jac{(l,b,D,\mu_l,\mu_b,v_r)}_{t=\tau}\\
+	p(\X \given \X_p, \tau, \Phi) &= p(\ln R, \ln V \given \tau, \Phi) \jac{(x,y,z,v_x,v_y,v_z)}_{t=\tau}\label{eq:int} \\
+	p(\ln R, \ln V \given \tau, \Phi) &= \left[\mathcal{N}(\ln R | \ln \rtide, \sigma_R)\mathcal{N}(\ln V | \ln \sigma_v, \sigma_V)\right]_{t=\tau}
 \end{align}
@@
 where in eq.~\ref{eq:int} we have implicitly integrated over and neglected the two angle coordinates associated with $\ln R$ and the two angles associated with $\ln V$. 
-We then compute the marginal likelihood, $p(\W \given \W_s, \Phi)$, by integrating over all possible unbinding times,
 \begin{align}
-	p(\W \given \W_s, \Phi) = \int p(\W \given \W_s, \tau, \Phi)p(\tau) d\tau.
+	\ln p(\ln R, \ln V \given \tau, \Phi) = 
 \end{align}
-In general, $p(\tau)$ is likely conditional on the orbit of the progenitor --- stars are preferentially stripped at pericentric passages --- but this is neglected and is taken to be uniform over the entire interaction history between the satellite and parent.
+We then compute the marginal likelihood, $p(\W \given \W_p, \Phi)$, by integrating over all possible unbinding times,
+\begin{align}
+	p(\W \given \W_p, \Phi) = \int p(\W \given \W_p, \tau, \Phi)p(\tau) d\tau.
+\end{align}
+In general, $p(\tau)$ is likely conditional on the orbit of the progenitor --- stars are preferentially stripped at pericentric passages --- but this is neglected and is taken to be uniform over the entire interaction history between the progenitor and parent.
 

### 2014-02-12 text/paper2/ms.tex
kvj structure

@@
 
+% Generate approx. to N-body -> single Orbit --> Orbits + offsets --> Bovy
+% method, problem, etc. then address what we fix
+
 - Broader comparison to other methods\\
-\indent\indent- 'generative methods' more generally (ie N-body, plus Varghese, plus Kuepper plus new possible action-angle models....\\
-\indent\indent- statistical methods\\
+\indent\indent- 'generative methods' more generally (ie N-body, Bovy, plus Varghese, plus Kuepper plus new possible action-angle models....\\
 \indent\indent- Sanders \& Binney 2013\\
+\indent\indent- statistical methods, Penarrubia, Sanderson, Rewinder\\
 - Then a short note on worries\\
@@
 
-[Chemical tagging will always only give probabilistic information.
-  But what if it delivers somewhat confident information about small
-  stellar families in the halo?]
+%[Chemical tagging will always only give probabilistic information.
+%  But what if it delivers somewhat confident information about small
+%  stellar families in the halo?]
 
@@
 
-\section{The model}
+\section{Simulations}\label{sec:sims}
+
+We [...] N-body simulations run with an SCF code [...]. [KVJ: describe simulations here?]
+
+\section{...}
+
+\subsection{Notation}
+Take $\D$ to be the observed, heliocentric 6D position of a star --- e.g., the measured position on the sky, ($l$, $b$); distance, $D$; proper motions, ($\mu_l$, $\mu_b$); and line-of-sight velocity, $v_r$ --- and take $\W$ to be the corresponding \emph{true}, 6D position of the star in the same coordinate and reference frame.
+Assume that we have determined that this star was once part of a progenitor system (e.g., satellite galaxy) tidally disrupting and forming a cold debris structure in the potential, $\Phi$, of some parent galaxy.
+The position of the progenitor is observed to be at heliocentric 6D position $\D_p$, with true position $\W_p$ (where any subscript $p$ refers to the progenitor).
+The true 6D heliocentric spherical positions of the star and progenitor -- $\W$ and $\W_p$ -- have corresponding 6D Galactocentric cartesian coordinate vectors $\X$ and $\X_p$ comprised of 3D position and velocity vectors ($\bsr$,$\bsv$) and ($\bsr_p$,$\bsv_p$).
+Initially we assume the present-day mass of the progenitor, $M_p$, and velocity dispersion, $\sigma_v$, have been measured perfectly and neglect mass-loss from the progenitor.
+
+\subsection{Motivation}
+In \cite{apw13}, we introduced the \emph{Rewinder} algorithm [...].
+In that work, we define a phase-space distance, $D_{ps}$, as the sum of the normalized, relative position, $\frac{\bsr-\bsr_p}{\rtide}$, and velocity, $\frac{\bsv-\bsv_p}{v_{\rm esc}}$, for a stream star.
+We minimized the generalized variance of the 6D distribution over each normalized coordinate evaluated at the time that each star has minimum $D_ps$.
+This simple idea proved quite powerful for constraining properties of the LM10 potential used simulated observations of their N-body simulation.
+However, this method does not properly incorporate observational uncertainties.
+
+Using the simulations described in Section~\ref{sec:sims}
 
@@
 
-[Conceptual description of the model] Figure~\ref{fig:reldist}
+%The core of this method provides a basis for the probabilistic model presented below.
 
-Take $\D$ to be the observed, heliocentric 6D position of a star --- e.g., the measured position on the sky, ($l$, $b$); distance, $D$; proper motions, ($\mu_l$, $\mu_b$); and line-of-sight velocity, $v_r$ --- and take $\W$ to be the corresponding \emph{true}, 6D position of the star in the same coordinate and reference frame.
-Assume that we have determined that this star was once part of a progenitor system (e.g., satellite galaxy) tidally disrupting and forming a cold debris structure in the potential, $\Phi$, of some parent galaxy.
-The position of the progenitor is observed to be at heliocentric 6D position $\D_p$, with true position $\W_p$ (where any subscript $p$ refers to the progenitor).
-The true 6D heliocentric spherical positions of the star and progenitor -- $\W$ and $\W_p$ -- have corresponding 6D Galactocentric cartesian coordinate vectors $\X$ and $\X_p$ comprised of 3D position and velocity vectors ($\bsr$,$\bsv$) and ($\bsr_p$,$\bsv_p$).
-Initially we assume the present-day mass of the progenitor, $M_p$, and velocity dispersion, $\sigma_v$, have been measured perfectly and neglect mass-loss from the progenitor.
+\subsection{Probabilistic model}
 With this notation, we write the joint posterior probability for the observed position of a star, progenitor, and parent potential parameters as:
@@
 \begin{align}
-	\ln p(\ln R, \ln V \given \tau, \Phi) = 
+	\ln p(\ln R, \ln V \given \tau, \Phi) = -\ln(2\pi)-\ln(\sigma_R\sigma_V)-
+		\ln\left[\left(\frac{R}{\rtide}\right)^{1/\sigma_R} \left(\frac{V}{\sigma_v}\right)^{1/\sigma_V}\right]
 \end{align}
@@
 \section{Experiments}
-We conduct several tests of this method using N-body simulations run with an SCF code [...]. [KVJ: describe simulations here?]
 

### 2014-02-12 text/paper2/ms.tex
more refs and text

@@
 The difficulty and computational costs associated with proper N-body modeling has motivated numerous efforts for developing approximate methods for constraining properties of the Galactic potential with cold debris structures.
+The simplest approximation is to fit a single orbit to observed debris \citep[e.g.,][]{koposov10, deg13}.
+This has the advantage of naturally incorporating observational errors, but is known to be invalid as debris does not exactly follow an orbit \citep[e.g.,][]{binney08, eyre11, sanders13a}.
+To account for this, methods have been proposed that model the debris as an orbit plus some dispersion or offset in phase space \citep[e.g.,][]{eyre09a, varghese11, kuepper12} and action-angle coordinates \citep{eyre11, sanders13b, bovy14}.
+Other statistical methods have been proposed \citep[][]{penarrubia12, sanderson14} that may prove powerful when applied to, e.g., data from the \gaia mission, where full 6D coordinates will be known for large samples of stars in the halo --- and therefore many debris structures --- but membership is not necessarily known.
+However, these methods do not account for observational errors.
 
-% Generate approx. to N-body -> single Orbit --> Orbits + offsets --> Bovy
-% method, problem, etc. then address what we fix
+In methods that rely on action-angle coordinates, [...].
 
-- Broader comparison to other methods\\
-\indent\indent- 'generative methods' more generally (ie N-body, Bovy, plus Varghese, plus Kuepper plus new possible action-angle models....\\
-\indent\indent- Sanders \& Binney 2013\\
-\indent\indent- statistical methods, Penarrubia, Sanderson, Rewinder\\
 - Then a short note on worries\\

### 2014-02-12 text/paper2/ms.tex
filling in comparison

@@
 \newcommand{\bsv}{\boldsymbol v}
+
+\newcommand{\paperone}{Paper 1}
 
@@
 This has the advantage of naturally incorporating observational errors, but is known to be invalid as debris does not exactly follow an orbit \citep[e.g.,][]{binney08, eyre11, sanders13a}.
-To account for this, methods have been proposed that model the debris as an orbit plus some dispersion or offset in phase space \citep[e.g.,][]{eyre09a, varghese11, kuepper12} and action-angle coordinates \citep{eyre11, sanders13b, bovy14}.
+To account for this, methods have been proposed that model the debris as an orbit plus some dispersion or offset in phase space \citep[e.g.,][]{eyre09a, varghese11, kuepper12} or action-angle coordinates \citep{eyre11, sanders13b, bovy14}.
 Other statistical methods have been proposed \citep[][]{penarrubia12, sanderson14} that may prove powerful when applied to, e.g., data from the \gaia mission, where full 6D coordinates will be known for large samples of stars in the halo --- and therefore many debris structures --- but membership is not necessarily known.
@@
 
-In methods that rely on action-angle coordinates, [...].
+Methods that rely on action-angle coordinates, [...].
+Any generative model that relies on matching densities --- whether in phase-space, action-space, or observed coordinates --- must also include details about the progenitor and a model for the background of stars.
+[...] In \citet[][hereafter \paperone]{apw13}, we introduced a simple method for using individual stars combined with knowledge about the mass and orbit of the progenitor to constrain properties of the host galaxy potential (\emph{Rewinder}).
+[...]
 
@@
 
-\section{...}
+\section{Method}
 

### 2014-02-13 text/paper2/ms.tex
sharpen intro paragraph

@@
 Cosmological simulations of galaxy formation in the $\Lambda$CDM paradigm predict dark matter halos that (1) are permeated with substructure on many scales, (2) are triaxial in shape, and (3) have density profiles and shapes that vary with radius \citep{dubinski91, jing02, kuhlen07, veraciro11}.
-Dark-matter-only simulations seem to produce strongly triaxial halos, but including baryons tends to soften the triaxiality in the inner galaxy [through a combination of adiabatic contraction and energetic feedback \citep[see recent discussion by][]{pontzen12} TODO: modify previous sentence bc in PW\&J 2013].
-These [...] may lead to radially-dependent axis ratios, halo orientation, and density profiles.
+Dark-matter-only simulations produce strongly triaxial halos \citep{} with large density variations \citep{zemp09}.
+Including baryons tends to soften the triaxiality in the inner galaxy through energetic feedback \citep[e.g.,][]{pontzen12} but does not erase the clumpy, triaxial nature of the outer halo.
+This leads to radially-dependent axis ratios and orientation, density profiles, and smoothness, thus predicting complicated, non-intregrable potential forms for Milky Way-like galaxies.
 

### 2014-02-13 text/paper2/ms.tex
fix notation

@@
 Take $\D$ to be the observed, heliocentric 6D position of a star --- e.g., the measured position on the sky, ($l$, $b$); distance, $D$; proper motions, ($\mu_l$, $\mu_b$); and line-of-sight velocity, $v_r$ --- and take $\W$ to be the corresponding \emph{true}, 6D position of the star in the same coordinate and reference frame.
-Assume that we have determined that this star was once part of a progenitor system (e.g., satellite galaxy) tidally disrupting and forming a cold debris structure in the potential, $\Phi$, of some parent galaxy.
+Assume that we have determined that this star was once part of a progenitor system (e.g., satellite galaxy) with mass $M_p(t)$ and velocity dispersion $\sigma_p$ tidally disrupting and forming a cold debris structure in the potential, $\Phi$, of some parent galaxy.
 The position of the progenitor is observed to be at heliocentric 6D position $\D_p$, with true position $\W_p$ (where any subscript $p$ refers to the progenitor).
 The true 6D heliocentric spherical positions of the star and progenitor -- $\W$ and $\W_p$ -- have corresponding 6D Galactocentric cartesian coordinate vectors $\X$ and $\X_p$ comprised of 3D position and velocity vectors ($\bsr$,$\bsv$) and ($\bsr_p$,$\bsv_p$).
-Initially we assume the present-day mass of the progenitor, $M_p$, and velocity dispersion, $\sigma_v$, have been measured perfectly and neglect mass-loss from the progenitor.
 
@@
 \section{Experiments}
+Initially we assume the present-day mass of the progenitor, $M_p$, and velocity dispersion, $\sigma_v$, have been measured perfectly and neglect mass-loss from the progenitor.
 

### 2014-02-13 text/paper2/ms.tex
rambling

@@
 Dark-matter-only simulations produce strongly triaxial halos \citep{} with large density variations \citep{zemp09}.
-Including baryons tends to soften the triaxiality in the inner galaxy through energetic feedback \citep[e.g.,][]{pontzen12} but does not erase the clumpy, triaxial nature of the outer halo.
+Including baryons tends to soften the triaxiality in the inner galaxy \citep[e.g.,][]{dubinski94, pontzen12} but does not erase the clumpy, triaxial nature of the outer halo.
 This leads to radially-dependent axis ratios and orientation, density profiles, and smoothness, thus predicting complicated, non-intregrable potential forms for Milky Way-like galaxies.
@@
 The bulk of the baryonic matter in galaxies spans roughly 5-10\% of the extent of the host dark matter halo.
-Hence, the major components of a galaxy are sensitive to the inner portion of its host halo's mass distribution.
-For example, the rotation curves of disk galaxies are particularly sensitive probes of the inner mass since matter in disks can be assumed to move on near circular orbits.
-This extra knowledge about the orbital distribution of the baryonic matter within a disk allows a more precise tracing of the potential than is feasible for objects such as elliptical galaxies where orbits are assumed to be random.
-[Potential is most likely smoothed out (energetic feedback), can approximate as closely integrable allowing beautiful dynamical methods ...]
+Hence, the major observable components of a galaxy are sensitive to the inner portion of its host halo's mass distribution.
+For example, the rotation curves of disk galaxies are excellent probes of the inner mass, especially since matter in disks can be assumed to move on near circular orbits.
+In the inner regions of the Galaxy, the potential is likely modified due to past dissipative infall \citep{dubinski94} or cooling \citep{bryan13} of baryons, which act to reduce triaxiality and smooth out density fluctuations.
+The inner potential is likely well-described by a closely-integrable potential, thus allowing [...]
 
-Conversely, measuring the dark matter distribution at large radii is difficult as the density of visible tracers becomes small.
-Around external galaxies, the extended mass distribution has been explored using a variety of approaches \citep[see][for a a complete and detailed review]{courteau12}.
+%This extra knowledge about the orbital distribution of the baryonic matter within a disk allows a more precise tracing of the potential than is feasible for objects such as elliptical galaxies where orbits are assumed to be random.
+
+Conversely, measuring the dark matter distribution at large radii is complicated by the low density of visible tracers and [...something about assumptions about integrability and smoothness].
+Around external galaxies, the extended mass distribution has been probed using a variety of approaches \citep[see][for a a complete and detailed review]{courteau12}.
 For example, the kinematics of tracer populations such as globular clusters or planetary nebulae can be used to derive mass estimates under the assumptions that these satellite systems are relaxed and well-mixed \citep[early investigations include][]{mendez01,cote03}.
@@
 
-Within the Milky Way we have a unique vantage point that allows us to measure the three-dimensional structure of our own dark matter halo.
-One approach to ...
-[explain ``random'' methods: kinematic tracers].
+Within the Milky Way our unique vantage point allows us a three-dimensional view of the structure in own dark matter halo.
+Deason et al.
+(201?) used BHB stars selected to be in the halo from the Sloan Digital Sky Survey \cite[SDSS;][]{} as random tracers [...reconstruct velocity trends?].
 Another approach is to exploit the non-random nature of the stellar populations in the halo.

### 2014-02-13 text/paper2/ms.tex
comment out figures
intro

@@
 \title{Inferring the gravitational potential of the Milky Way with a few precisely measured stars}
-\author{Adrian M. Price-Whelan, David W. Hogg, Kathryn V. Johnston, [TODO: order?]}
+\author{Adrian M. Price-Whelan, David W. Hogg, Kathryn V. Johnston}
 
@@
 Dark-matter-only simulations produce strongly triaxial halos \citep{} with large density variations \citep{zemp09}.
-Including baryons tends to soften the triaxiality in the inner galaxy \citep[e.g.,][]{dubinski94, pontzen12} but does not erase the clumpy, triaxial nature of the outer halo.
+Including baryons tends to soften the triaxiality and graininess in the inner galaxy \citep[e.g.,][]{dubinski94, pontzen12} but does not erase the clumpy, triaxial nature of the outer halo.
 This leads to radially-dependent axis ratios and orientation, density profiles, and smoothness, thus predicting complicated, non-intregrable potential forms for Milky Way-like galaxies.

### 2014-02-13 text/paper2/ms.tex
intro enhancemens

@@
 Dark-matter-only simulations produce strongly triaxial halos \citep{} with large density variations \citep{zemp09}.
-Including baryons tends to soften the triaxiality and graininess in the inner galaxy \citep[e.g.,][]{dubinski94, pontzen12} but does not erase the clumpy, triaxial nature of the outer halo.
-This leads to radially-dependent axis ratios and orientation, density profiles, and smoothness, thus predicting complicated, non-intregrable potential forms for Milky Way-like galaxies.
+Including baryons tends to soften the triaxiality and graininess in the inner galaxy through a combination of dissipative infall \citep{dubinski94} or cooling \citep{bryan13}.
+These processes combined with the gravity from a baryonic disk (ellipsoid) can act to make the inner halo more oblate (spherical), however these processes do not erase the clumpy, triaxial nature of the outer halo \citep[e.g.,][]{pontzen12}.
+This leads to radially-dependent axis ratios and orientation, density profile, and smoothness, thus suggesting complicated, non-intregrable potential forms for Milky Way-like galaxies.
 
@@
 For example, the rotation curves of disk galaxies are excellent probes of the inner mass, especially since matter in disks can be assumed to move on near circular orbits.
-In the inner regions of the Galaxy, the potential is likely modified due to past dissipative infall \citep{dubinski94} or cooling \citep{bryan13} of baryons, which act to reduce triaxiality and smooth out density fluctuations.
-The inner potential is likely well-described by a closely-integrable potential, thus allowing [...]
-
-%This extra knowledge about the orbital distribution of the baryonic matter within a disk allows a more precise tracing of the potential than is feasible for objects such as elliptical galaxies where orbits are assumed to be random.
-
-Conversely, measuring the dark matter distribution at large radii is complicated by the low density of visible tracers and [...something about assumptions about integrability and smoothness].
-Around external galaxies, the extended mass distribution has been probed using a variety of approaches \citep[see][for a a complete and detailed review]{courteau12}.
+The inner potential is likely well-described by a closely-integrable potential, thus allowing [...] Conversely, measuring the dark matter distribution at large radii is complicated by the low density of visible tracers and [...something about assumptions about integrability and smoothness].
+Around external galaxies, the extended mass distribution has been probed using a variety of approaches \citep[see][for a a complete and detailed review]{courteau13}.
 For example, the kinematics of tracer populations such as globular clusters or planetary nebulae can be used to derive mass estimates under the assumptions that these satellite systems are relaxed and well-mixed \citep[early investigations include][]{mendez01,cote03}.
-These estimates are necessarily limited by the number of objects bright enough for spectroscopic measurements and thus the largest samples to date have roughly several hundred members probing out to several 10's of kpc.
-Simple, parameterized models of both the mass and orbit distribution have been simultaneously constrained using such data \citep[e.g.][]{napolitano11,deason12a}.
+Simple, parameterized models of both the mass and orbit distribution have been simultaneously constrained using such data \citep[e.g.][]{napolitano11,deason12c}.
 %\citep{romanowsky03,douglas07,romanowsky09,napolitano11,lee11}
@@
 Within the Milky Way our unique vantage point allows us a three-dimensional view of the structure in own dark matter halo.
-Deason et al.
-(201?) used BHB stars selected to be in the halo from the Sloan Digital Sky Survey \cite[SDSS;][]{} as random tracers [...reconstruct velocity trends?].
-Another approach is to exploit the non-random nature of the stellar populations in the halo.
-A consequence of hierarchical structure formation [...] and such structures have been observed [e.g., ...cite belokurov?].
+Our proximity allows us to use individual stars as kinematic tracers and hence build much larger samples that probe deeper into the halo than the globular cluster and planetary nebula studies of external galaxies.
+For example, \cite{deason12a} used halo BHB stars selected from the Sloan Digital Sky Survey \cite[SDSS;][]{york00} as a random tracer population to measure the mass and slope of a simple, power-law fit to the Galactic potential.
+Such studies assume that the tracer orbits are randomly sampled from a smooth distribution function and phase mixed, however the existence of large-scale substructure in stellar streams and clouds is known to bias mass and velocity inferences that ignore this fact.
+Another approach then is to exploit the non-random nature of the stellar populations in the halo.
 Tidal streams are dynamically cold systems --- debris typically have small distributions of energy and angular momentum --- and thus require orders of magnitude fewer tracers than a random sample to get constraints of comparable accuracy.

### 2014-02-13 text/paper2/ms.tex
some reorganizing

@@
 Cosmological simulations of galaxy formation in the $\Lambda$CDM paradigm predict dark matter halos that (1) are permeated with substructure on many scales, (2) are triaxial in shape, and (3) have density profiles and shapes that vary with radius \citep{dubinski91, jing02, kuhlen07, veraciro11}.
-Dark-matter-only simulations produce strongly triaxial halos \citep{} with large density variations \citep{zemp09}.
+Dark-matter-only simulations produce triaxial halos \citep{jing02} with large density fluctuations \citep{zemp09}.
 Including baryons tends to soften the triaxiality and graininess in the inner galaxy through a combination of dissipative infall \citep{dubinski94} or cooling \citep{bryan13}.
@@
 Such studies assume that the tracer orbits are randomly sampled from a smooth distribution function and phase mixed, however the existence of large-scale substructure in stellar streams and clouds is known to bias mass and velocity inferences that ignore this fact.
-Another approach then is to exploit the non-random nature of the stellar populations in the halo.
+Another approach is to exploit the non-random nature of the stellar populations in the halo.
 Tidal streams are dynamically cold systems --- debris typically have small distributions of energy and angular momentum --- and thus require orders of magnitude fewer tracers than a random sample to get constraints of comparable accuracy.
-For example, in the simplest case we might {\it assume} that debris stars are actually still on the same orbit as their progenitor system \citep[a \emph{wrong} assumption, see e.g.][]{eyre}.
-Then we can imagine measuring the full-space velocities ${\bf v}$ at different points ${\bf x}$ in the structure (e.g.
-along a stream), which would give us a direct measure of differences in a potential, $\Phi$.
-In practice, simple methods such as orbit fitting lead to systematic biases in inferred properties of the underlying potential because streams do not strictly delineate orbits [cite eyre binney work].
+For example, in the simplest case we might assume that debris stars are actually still on the same orbit as their progenitor system \citep[a \emph{wrong} assumption, see e.g.,][]{}.
+Then we can imagine measuring the full-space velocities ${\bf v}$ at different points ${\bf x}$ in the structure (e.g., along a stream), which would give us a direct measure of differences in a potential, $\Phi$.
+%In practice, simple methods such as orbit fitting lead to systematic biases in inferred properties of the underlying potential because streams do not strictly delineate orbits [cite eyre binney work].
 
-Since the initial discovery of tidal debris from Sgr, myriad other streams and debris systems have been discovered in large photometric surveys probing Galactocentric distances from $\sim$15-100 kpc.
+Since the initial discovery of tidal debris from Sgr, myriad other streams and debris systems have been identified in large photometric surveys, probing Galactocentric distances from $\sim$15-100 kpc.
 The known debris structures are $\gtrsim$10 kpc from the Sun, thus while overdensities in the halo are less contaminated by background, the constituent stars often have incomplete or poorly measured phase-space information.
-Perhaps the most rigorous method of modeling streams and the potential is to run full N-body simulations of satellite disruption and compare the density of observed stars to simulated stars.
-Such a model could naturally incorporate observational uncertainties, missing dimensions, and would, in principle, allow for studying any arbitrary time-dependent or non-integrable potential.
-However, such a model is presently computationally intractable [though see milkyway@home stuff...].
-\cite{law10} ran several N-body models of the disruption of the Sagittarius (Sgr) dwarf spheroidal galaxy and qualitatively matched their simulated streams to all Sgr data available at that time.
-By comparing the phase-space positions of their stream particles with the data, they were able to use these simulations to constrain the shape of the Milky Way potential out to $\sim$70 kpc [check number] and found that the best-fitting halo has a peculiar triaxial shape elongated in the Galactic Z direction, orthogonal to the plane of the disk.
-(though, see ...
-for possible complications in the Sgr debris belokurov 2012?).
+Perhaps the most rigorous method of modeling streams and the potential is to run full N-body simulations of satellite disruption and compare the density of observed stars to simulated stars in whatever coordinates are measured.
+A generative N-body model could naturally incorporate observational uncertainties, missing dimensions, and would, in principle, allow for studying any arbitrary time-dependent or non-integrable potential.
+However, such a model is presently computationally intractable.
+\cite{law10} demonstrated the power of this idea by running a grid of N-body models of the disruption of the Sagittarius (Sgr) dwarf spheroidal galaxy and matched their simulated streams to all Sgr data available at that time.
+By comparing the phase-space positions of their stream particles with the data, they were able to constrain the shape of the Milky Way potential out to $\sim$70 kpc and found that the best-fitting halo has a prolate, triaxial shape elongated in the Galactic Z direction, orthogonal to the plane of the disk.
+% (though, see ... for possible complications in the Sgr debris belokurov 2012?).
 
@@
 The simplest approximation is to fit a single orbit to observed debris \citep[e.g.,][]{koposov10, deg13}.
-This has the advantage of naturally incorporating observational errors, but is known to be invalid as debris does not exactly follow an orbit \citep[e.g.,][]{binney08, eyre11, sanders13a}.
-To account for this, methods have been proposed that model the debris as an orbit plus some dispersion or offset in phase space \citep[e.g.,][]{eyre09a, varghese11, kuepper12} or action-angle coordinates \citep{eyre11, sanders13b, bovy14}.
-Other statistical methods have been proposed \citep[][]{penarrubia12, sanderson14} that may prove powerful when applied to, e.g., data from the \gaia mission, where full 6D coordinates will be known for large samples of stars in the halo --- and therefore many debris structures --- but membership is not necessarily known.
-However, these methods do not account for observational errors.
+This has the advantage of naturally incorporating observational errors, but is known to be invalid as debris does not exactly follow an orbit \citep[e.g.,][]{eyre11, sanders13a}.
+To account for this, methods have been proposed that model the debris as an orbit plus some dispersion or offset, either in phase space \citep[e.g.,][]{eyre09a, varghese11, kuepper12} or action-angle coordinates \citep{eyre11, sanders13b, bovy14}.
+Other statistical methods have been proposed \citep[][]{penarrubia12, sanderson14} that may prove powerful when applied to, e.g., data from the \gaia mission, where full 6D coordinates will be known for large samples of stars in the halo --- and therefore many debris structures --- but stream membership is not known for all stars (however these statistical methods do not account for observational errors).
 
-Methods that rely on action-angle coordinates, [...].
+[Here we need to talk about integrable potentials and shortcomings of the methods above?] Methods that rely on action-angle coordinates, [...].
 Any generative model that relies on matching densities --- whether in phase-space, action-space, or observed coordinates --- must also include details about the progenitor and a model for the background of stars.
-[...] In \citet[][hereafter \paperone]{apw13}, we introduced a simple method for using individual stars combined with knowledge about the mass and orbit of the progenitor to constrain properties of the host galaxy potential (\emph{Rewinder}).
 [...]
-
 - Then a short note on worries\\
@@
 - Then how rewinder gets round these????\\
-\indent\indent- Reference Deg \& Widrow to say they illustrate the idea of the power of multiple streams to back up the bit where you say that's the long-term aim\\
-
-[Another prediction of hierarchical structure formation in $\Lambda$CDM is the presence of substructure. ]
-
-%[Chemical tagging will always only give probabilistic information.
-%  But what if it delivers somewhat confident information about small
-%  stellar families in the halo?]
+\indent\indent- Reference Deg \& Widrow to say they illustrate the idea of the power of multiple streams to back up the bit where you say that's the long-term aim
 
@@
 
-[Cold vs. Hot streams in the context of observational errors]
+In \citet[][hereafter \paperone]{apw13}, we introduced a simple method for using individual stars associated with tidal debris combined with knowledge about the mass and orbit of the progenitor to constrain properties of the host galaxy potential (\emph{Rewinder}).
+[...]
 
@@
 \subsection{Motivation}
-In \cite{apw13}, we introduced the \emph{Rewinder} algorithm [...].
-In that work, we define a phase-space distance, $D_{ps}$, as the sum of the normalized, relative position, $\frac{\bsr-\bsr_p}{\rtide}$, and velocity, $\frac{\bsv-\bsv_p}{v_{\rm esc}}$, for a stream star.
+
+[Cold vs. Hot streams in the context of observational errors]
+
+In \paperone, we define a phase-space distance, $D_{ps}$, as the sum of the normalized, relative position, $\frac{\bsr-\bsr_p}{\rtide}$, and velocity, $\frac{\bsv-\bsv_p}{v_{\rm esc}}$, for a stream star.
 We minimized the generalized variance of the 6D distribution over each normalized coordinate evaluated at the time that each star has minimum $D_ps$.
@@
 \begin{center}
-%\includegraphics[width=0.45\textwidth]{fig4.eps}
-\includegraphics[width=\textwidth]{/Users/adrian/projects/streams/plots/paper2/rel_dists.pdf}
+% \includegraphics[width=\textwidth]{/Users/adrian/projects/streams/plots/paper2/rel_dists.pdf}
 \caption{ TODO }\label{fig:reldist}
@@
 \begin{center}
-%\includegraphics[width=0.45\textwidth]{fig4.eps}
-\includegraphics[width=\textwidth]{/Users/adrian/projects/streams/plots/paper2/potentials.pdf}
+% \includegraphics[width=\textwidth]{/Users/adrian/projects/streams/plots/paper2/potentials.pdf}
 \caption{ TODO }\label{fig:potential}
@@
 
+[Another prediction of hierarchical structure formation in $\Lambda$CDM is the presence of substructure. In principle, this method would work on any potential that we can integrate orbits in.]
+
 \bibliographystyle{apj}

### 2014-02-13 text/paper2/ms.tex
affiliations

@@
 \title{Inferring the gravitational potential of the Milky Way with a few precisely measured stars}
-\author{Adrian M. Price-Whelan, David W. Hogg, Kathryn V. Johnston}
+\author{Adrian M. Price-Whelan\altaffilmark{\colum,\adrn}, 
+	    David W. Hogg\altaffilmark{\nyu,\mpia}, 
+	    Kathryn V. Johnston\altaffilmark{\colum}, 
+	    David Hendel\altaffilmark{\colum}}
+
+% Affiliations
+\newcommand{\colum}{1}
+\newcommand{\adrn}{2}
+\newcommand{\nyu}{3}
+\newcommand{\mpia}{4}
+\altaffiltext{\colum}{Department of Astronomy, 
+		              Columbia University, 
+		              550 W 120th St., 
+		              New York, NY 10027, USA}
+\altaffiltext{\adrn}{To whom correspondence should be addressed: \email{adrn@astro.columbia.edu}}
+\altaffiltext{\nyu}{Center for Cosmology and Particle Physics,
+                      Department of Physics, New York University,
+                      4 Washington Place, New York, NY, 10003, USA}
+\altaffiltext{\mpia}{Max-Planck-Institut f\"ur Astronomie,
+                     K\"onigstuhl 17, D-69117 Heidelberg, Germany}
 
@@
 This has the advantage of naturally incorporating observational errors, but is known to be invalid as debris does not exactly follow an orbit \citep[e.g.,][]{eyre11, sanders13a}.
-To account for this, methods have been proposed that model the debris as an orbit plus some dispersion or offset, either in phase space \citep[e.g.,][]{eyre09a, varghese11, kuepper12} or action-angle coordinates \citep{eyre11, sanders13b, bovy14}.
+To account for this, methods have been proposed that model the debris as an orbit plus some dispersion or offset, either in phase space \citep[e.g.,][]{eyre09a, varghese11, kuepper12} or action-angle coordinates \citep{eyre11, sanders13b, bovy14, sanders14}.
 Other statistical methods have been proposed \citep[][]{penarrubia12, sanderson14} that may prove powerful when applied to, e.g., data from the \gaia mission, where full 6D coordinates will be known for large samples of stars in the halo --- and therefore many debris structures --- but stream membership is not known for all stars (however these statistical methods do not account for observational errors).
@@
 
-In \citet[][hereafter \paperone]{apw13}, we introduced a simple method for using individual stars associated with tidal debris combined with knowledge about the mass and orbit of the progenitor to constrain properties of the host galaxy potential (\emph{Rewinder}).
-[...]
+In \citet[][hereafter \paperone]{apw13}, we introduced a simple method (\emph{Rewinder}) for using individual stars associated with tidal debris combined with knowledge about the mass and orbit of the progenitor to constrain properties of the host galaxy potential.
+The method exploits the relationship between the phase-space distribution of debris and (measurable) properties of the progenitor system (e.g., the tidal radius and escape velocity).
+In this paper, we investigate how these relationships scale with the mass and orbit of the progenitor and the quality of the data, then present a new probabilistic framework for simultaneously modeling stream debris and the parent potential based on the same arguments.
 
-[We properly handle observational errors]
+In Section~\ref{sec:sims} we describe a suite of N-body simulations that span a range of progenitor masses over two characteristic orbits. In Section~\ref{sec:}
 

### 2014-02-14 text/paper2/ms.tex
added bit on scalings

@@
 \newcommand{\rtide}{r_{\rm tide}}
-
-\newcommand{\bsr}{\boldsymbol r}
-\newcommand{\bsv}{\boldsymbol v}
+\newcommand{\bs}{\boldsymbol}
 
@@
 \begin{abstract}
-Context:
-Aims: 
+Context: Connecting cosmological simulations to observed galaxies is difficult because baryonic processes not well understood/modeled but also because they are gravitationally dominated by unobservable dark matter.
+Aims: Develop a method for inferring higher order properties of the Galactic potential using debris from disrupted satellites.
 Methods: 
@@
 
-In Section~\ref{sec:sims} we describe a suite of N-body simulations that span a range of progenitor masses over two characteristic orbits. In Section~\ref{sec:}
+In Section~\ref{sec:sims} we describe a suite of N-body simulations that span a range of progenitor masses over two characteristic orbits. In Section~\ref{sec:} we
 
@@
 
-We [...] N-body simulations run with an SCF code [...]. [KVJ: describe simulations here?]
+We [...] N-body simulations run with an SCF code [...]. [KVJ: describe simulations here?] [Describe tub]
 
-\section{Method}
+\section{[Scalings?]}
 
-\subsection{Notation}
-Take $\D$ to be the observed, heliocentric 6D position of a star --- e.g., the measured position on the sky, ($l$, $b$); distance, $D$; proper motions, ($\mu_l$, $\mu_b$); and line-of-sight velocity, $v_r$ --- and take $\W$ to be the corresponding \emph{true}, 6D position of the star in the same coordinate and reference frame.
-Assume that we have determined that this star was once part of a progenitor system (e.g., satellite galaxy) with mass $M_p(t)$ and velocity dispersion $\sigma_p$ tidally disrupting and forming a cold debris structure in the potential, $\Phi$, of some parent galaxy.
-The position of the progenitor is observed to be at heliocentric 6D position $\D_p$, with true position $\W_p$ (where any subscript $p$ refers to the progenitor).
-The true 6D heliocentric spherical positions of the star and progenitor -- $\W$ and $\W_p$ -- have corresponding 6D Galactocentric cartesian coordinate vectors $\X$ and $\X_p$ comprised of 3D position and velocity vectors ($\bsr$,$\bsv$) and ($\bsr_p$,$\bsv_p$).
+In \paperone, we define a set of coordinates relative to the instantaneous phase-space position of the progenitor, $(\bs{x}_{p},\bs{v}_{p})$, normalized by the instantaneous tidal radius and escape velocity.
+Here we (1) instead scale the velocity coordinates by the velocity dispersion of the satellite instead of the escape velocity, (2) reduce the dimensionality by considering only the magnitude of these vectors, and (3) take the logarithm of the magnitude,
+\begin{equation}
+  R = \ln\left(\frac{\left\vert\bs{x} - \bs{x}_{p}\right\vert}{R_{\rm tide}}\right)\,\,\,\,,\,\,\,\,
+  V = \ln\left(\frac{\left\vert\bs{v} - \bs{v}_{p}\right\vert}{\sigma_v}\right).
+\end{equation}
 
-\subsection{Motivation}
-
-[Cold vs. Hot streams in the context of observational errors]
-
-In \paperone, we define a phase-space distance, $D_{ps}$, as the sum of the normalized, relative position, $\frac{\bsr-\bsr_p}{\rtide}$, and velocity, $\frac{\bsv-\bsv_p}{v_{\rm esc}}$, for a stream star.
+a phase-space distance, $D_{ps}$, as the sum of the normalized, relative position, $\frac{\bs{r}-\bs{r}_p}{\rtide}$, and velocity, $\frac{\bs{v}-\bs{v}_p}{v_{\rm esc}}$, for a stream star.
 We minimized the generalized variance of the 6D distribution over each normalized coordinate evaluated at the time that each star has minimum $D_ps$.
@@
 \begin{center}
-% \includegraphics[width=\textwidth]{/Users/adrian/projects/streams/plots/paper2/rel_dists.pdf}
+\includegraphics[width=\textwidth]{/Users/adrian/projects/streams/plots/paper2/rel_dists.pdf}
 \caption{ TODO }\label{fig:reldist}
@@
 %The core of this method provides a basis for the probabilistic model presented below.
+
+[Cold vs. Hot streams in the context of observational errors]
+
+\section{Method}
+
+\subsection{Notation}
+Take $\D$ to be the observed, heliocentric 6D position of a star --- e.g., the measured position on the sky, ($l$, $b$); distance, $D$; proper motions, ($\mu_l$, $\mu_b$); and line-of-sight velocity, $v_r$ --- and take $\W$ to be the corresponding \emph{true}, 6D position of the star in the same coordinate and reference frame.
+Assume that we have determined that this star was once part of a progenitor system (e.g., satellite galaxy) with mass $M_p(t)$ and velocity dispersion $\sigma_p$ tidally disrupting and forming a cold debris structure in the potential, $\Phi$, of some parent galaxy.
+The position of the progenitor is observed to be at heliocentric 6D position $\D_p$, with true position $\W_p$ (where any subscript $p$ refers to the progenitor).
+The true 6D heliocentric spherical positions of the star and progenitor -- $\W$ and $\W_p$ -- have corresponding 6D Galactocentric cartesian coordinate vectors $\X$ and $\X_p$ comprised of 3D position and velocity vectors ($\bs{r}$,$\bs{v}$) and ($\bs{r}_p$,$\bs{v}_p$).
 
@@
 
-The disruption of a satellite galaxy due to tidal shocking [...connect to Choi work some how?] We model the present-day position and velocity of a disrupted star as the result of sampling from isotropic, log-normal distributions in relative distance, $R=\left\vert \bsr-\bsr_p \right\vert$, and relative velocity, $V=\left\vert \bsv-\bsv_p \right\vert$, at the time the star comes unbound from the progenitor, evolved to present day in the parent potential, $\Phi$.
+The disruption of a satellite galaxy due to tidal shocking [...connect to Choi work some how?] We model the present-day position and velocity of a disrupted star as the result of sampling from isotropic, log-normal distributions in relative distance, $R=\left\vert \bs{r}-\bs{r}_p \right\vert$, and relative velocity, $V=\left\vert \bs{v}-\bs{v}_p \right\vert$, at the time the star comes unbound from the progenitor, evolved to present day in the parent potential, $\Phi$.
 The distributions over $\ln R$ and $\ln V$ are centered on the logarithm of the instantaneous tidal radius, $\rtide$, and velocity dispersion, $\sigma_v$, of the progenitor at time $\tau$: $\mathcal{N}(\ln R \given \ln\rtide)\vert_{t=\tau}$ and $\mathcal{N}(\ln V \given \ln \sigma_v)\vert_{t=\tau}$.
@@
 \acknowledgements
+APW: NSF GRFP, Sigma Xi
 

### 2014-02-15 text/paper2/ms.tex
references and intro text

@@
 Context: Connecting cosmological simulations to observed galaxies is difficult because baryonic processes not well understood/modeled but also because they are gravitationally dominated by unobservable dark matter.
-Aims: Develop a method for inferring higher order properties of the Galactic potential using debris from disrupted satellites.
+One powerful measure is the shape of the potential -- triaxial, density profile, etc.
+Tidal streams are excellent tracers of mass.
+
+Aims: Present a method for inferring higher order properties of the Galactic potential using debris from disrupted satellites.
+
 Methods:
-Results: 
+
+Results: Can simultaneously constrain properties of progenitor, missing dimensions, and potential with a simple model for disrupted debris in phase-space.
+The method works well with small numbers of stars and with finite or large observational uncertainties or missing dimensions and generalizes naturally to multiple independent debris structures.
+
 Conclusions:
@@
 For example, the rotation curves of disk galaxies are excellent probes of the inner mass, especially since matter in disks can be assumed to move on near circular orbits.
-The inner potential is likely well-described by a closely-integrable potential, thus allowing [...] Conversely, measuring the dark matter distribution at large radii is complicated by the low density of visible tracers and [...something about assumptions about integrability and smoothness].
+Conversely, measuring the dark matter distribution at large radii is complicated by the low density of visible tracers, observational difficulties, and [unclear what assumptions can be made about the integrability and smoothness of the mass distribution].
 Around external galaxies, the extended mass distribution has been probed using a variety of approaches \citep[see][for a a complete and detailed review]{courteau13}.
@@
 Tidal streams are dynamically cold systems --- debris typically have small distributions of energy and angular momentum --- and thus require orders of magnitude fewer tracers than a random sample to get constraints of comparable accuracy.
-For example, in the simplest case we might assume that debris stars are actually still on the same orbit as their progenitor system \citep[a \emph{wrong} assumption, see e.g.,][]{}.
+For example, in the simplest case we might assume that debris stars are actually still on the same orbit as their progenitor system \citep[a \emph{wrong} assumption, see e.g.,][]{eyre2011}.
 Then we can imagine measuring the full-space velocities ${\bf v}$ at different points ${\bf x}$ in the structure (e.g., along a stream), which would give us a direct measure of differences in a potential, $\Phi$.
@@
 
-[Here we need to talk about integrable potentials and shortcomings of the methods above?] Methods that rely on action-angle coordinates, [...].
-Any generative model that relies on matching densities --- whether in phase-space, action-space, or observed coordinates --- must also include details about the progenitor and a model for the background of stars.
-[...]
-- Then a short note on worries\\
-\indent\indent- for generative models\\
-\indent\indent- for any methods based on action/angle \\
-\indent\indent- GD-1 was fit as if it highlighted an orbit.  That is known to be wrong.\\
-\indent\indent- Densities are hard to match: need a background model, way to go from luminosity to density, needs assumptions about the mass function and structure of the satellite\\
-\indent\indent- Triaxiality / radially-dependent shapes and substructure present problems for integral-based methods\\
-- Then how rewinder gets round these????\\
-\indent\indent- Reference Deg \& Widrow to say they illustrate the idea of the power of multiple streams to back up the bit where you say that's the long-term aim
-
-[Actions, angles, integrable, not, etc.  Can integrate in any
-  potential, even a time-varying one. We use this to
-  ``run the clock back''.  Why is an objective based on past collisions not the
-  same as assuming that the stars are on the same orbit?]
+The potential of the inner galaxy is likely well-described by a (set of) closely-integrable potentials, thus the dynamics of collisionless systems is plausibly captured by a transformation to some integral-of-motion-space, e.g., action-angle coordinates.
+However there is uncertainty in the exact shape, radial profile, and graininess of the outer halo.
+To test the extent to which the outer halo is well-fit by integrable potentials, we cannot use methods that \emph{assume} integrability.
+In this article, we present a fully probabilistic method for measuring properties of a potential that only relies on numerical orbit integration in ordinary phase-space.
+The model uses individual stars as constraints on the potential, and can therefore handle very small samples of stars.
+Generative models that instead rely on matching densities (in phase-space, action-space, or observed coordinates) must also include fine details about the progenitor (e.g., mass function or structure) and a model for the background of stars.
+Our method also naturally extends to using multiple debris structures \citep[see][who illustrate the power of using multiple streams to simultaneously constrain the potential]{deg14}.
 
@@
 
-\section{[Scalings?]}
+\section{Method}
+
+\subsection{Notation}
+Take $\D$ to be the observed, heliocentric 6D position of a star --- e.g., the measured position on the sky, ($l$, $b$); distance, $D$; proper motions, ($\mu_l$, $\mu_b$); and line-of-sight velocity, $v_r$ --- and take $\W$ to be the corresponding \emph{true}, 6D position of the star in the same coordinate and reference frame.
+Assume that we have determined that this star was once part of a progenitor system (e.g., satellite galaxy) with mass $M_p(t)$ and velocity dispersion $\sigma_p$ tidally disrupting and forming a cold debris structure in the potential, $\Phi$, of some parent galaxy.
+The position of the progenitor is observed to be at heliocentric 6D position $\D_p$, with true position $\W_p$ (where any subscript $p$ refers to the progenitor).
+The true 6D heliocentric spherical positions of the star and progenitor -- $\W$ and $\W_p$ -- have corresponding 6D Galactocentric cartesian coordinate vectors $\X$ and $\X_p$ comprised of 3D position and velocity vectors ($\bs{r}$,$\bs{v}$) and ($\bs{r}_p$,$\bs{v}_p$).
+
+\section{Scalings}
 
 In \paperone, we define a set of coordinates relative to the instantaneous phase-space position of the progenitor, $(\bs{x}_{p},\bs{v}_{p})$, normalized by the instantaneous tidal radius and escape velocity.
-Here we (1) instead scale the velocity coordinates by the velocity dispersion of the satellite instead of the escape velocity, (2) reduce the dimensionality by considering only the magnitude of these vectors, and (3) take the logarithm of the magnitude,
+Here we (1) scale the velocity coordinates by the velocity dispersion of the satellite instead of the escape velocity, (2) reduce the dimensionality by considering only the magnitude of these vectors, and (3) take the logarithm of the magnitude,
 \begin{equation}
@@
 [Cold vs. Hot streams in the context of observational errors]
-
-\section{Method}
-
-\subsection{Notation}
-Take $\D$ to be the observed, heliocentric 6D position of a star --- e.g., the measured position on the sky, ($l$, $b$); distance, $D$; proper motions, ($\mu_l$, $\mu_b$); and line-of-sight velocity, $v_r$ --- and take $\W$ to be the corresponding \emph{true}, 6D position of the star in the same coordinate and reference frame.
-Assume that we have determined that this star was once part of a progenitor system (e.g., satellite galaxy) with mass $M_p(t)$ and velocity dispersion $\sigma_p$ tidally disrupting and forming a cold debris structure in the potential, $\Phi$, of some parent galaxy.
-The position of the progenitor is observed to be at heliocentric 6D position $\D_p$, with true position $\W_p$ (where any subscript $p$ refers to the progenitor).
-The true 6D heliocentric spherical positions of the star and progenitor -- $\W$ and $\W_p$ -- have corresponding 6D Galactocentric cartesian coordinate vectors $\X$ and $\X_p$ comprised of 3D position and velocity vectors ($\bs{r}$,$\bs{v}$) and ($\bs{r}_p$,$\bs{v}_p$).
 

### 2014-02-15 text/paper2/ms.tex
x to r

@@
 \section{Simulations}\label{sec:sims}
-
-We [...] N-body simulations run with an SCF code [...]. [KVJ: describe simulations here?] [Describe tub]
+[David H.]
+We [...] N-body simulations run with an SCF code [...]. [describe simulations here] [Describe tub]
 
@@
 
-In \paperone, we define a set of coordinates relative to the instantaneous phase-space position of the progenitor, $(\bs{x}_{p},\bs{v}_{p})$, normalized by the instantaneous tidal radius and escape velocity.
+In \paperone, we define a set of coordinates relative to the instantaneous phase-space position of the progenitor, $(\bs{r}_{p},\bs{v}_{p})$, normalized by the instantaneous tidal radius and escape velocity.
 Here we (1) scale the velocity coordinates by the velocity dispersion of the satellite instead of the escape velocity, (2) reduce the dimensionality by considering only the magnitude of these vectors, and (3) take the logarithm of the magnitude,
 \begin{equation}
-  R = \ln\left(\frac{\left\vert\bs{x} - \bs{x}_{p}\right\vert}{R_{\rm tide}}\right)\,\,\,\,,\,\,\,\,
+  R = \ln\left(\frac{\left\vert\bs{r} - \bs{r}_{p}\right\vert}{R_{\rm tide}}\right)\,\,\,\,,\,\,\,\,
   V = \ln\left(\frac{\left\vert\bs{v} - \bs{v}_{p}\right\vert}{\sigma_v}\right).

### 2014-02-15 text/paper2/ms.tex
comment out figures
total abstract rewrite

@@
 		              New York, NY 10027, USA}
-\altaffiltext{\adrn}{To whom correspondence should be addressed: \email{adrn@astro.columbia.edu}}
+\altaffiltext{\adrn}{To whom correspondence should be addressed: adrn@astro.columbia.edu}
 \altaffiltext{\nyu}{Center for Cosmology and Particle Physics,
@@
 \begin{abstract}
-Context: Connecting cosmological simulations to observed galaxies is difficult because baryonic processes not well understood/modeled but also because they are gravitationally dominated by unobservable dark matter.
-One powerful measure is the shape of the potential -- triaxial, density profile, etc.
-Tidal streams are excellent tracers of mass.
-
-Aims: Present a method for inferring higher order properties of the Galactic potential using debris from disrupted satellites.
-
-Methods:
-
-Results: Can simultaneously constrain properties of progenitor, missing dimensions, and potential with a simple model for disrupted debris in phase-space.
-The method works well with small numbers of stars and with finite or large observational uncertainties or missing dimensions and generalizes naturally to multiple independent debris structures.
-
-Conclusions:
-
-%Informative phase-space structures, and in particular those created by
-%disrupting stellar systems or revealed by chemical tags, may deliver
-%very precise measures of the gravitational potential in the Milky Way
-%Halo.  Here we show that even a single pair of stars---two stars that
-%are known (for some non-kinematic reason) to be likely to be
-%associated with one another at birth---could provide a significant
-%constraint on the potential.  The inference is based on a
-%probabilistic generative model---a likelihood function and priors over
-%nuisance parameters---that evolves a past putative origin disruption
-%event forward in time to the present day.  The time, six-volume, and
-%phase-space location of the origin event are nuisance parameters in
-%the model and marginalized away.  The method makes no assumption of
-%integrability, works with finite or even large observational
-%uncertainties, does not require all dimensions of phase space to be
-%observed, handles non-zero probability that the two stars are not in
-%fact associated, and generalizes naturally to larger numbers of stars
-%and multiple independent structures.  Applications to the GD-1 cold
-%stellar stream and current surveys for RR Lyrae stars in the Halo are
-%discussed.
-TODO...
+% Context
+ [this needs to get shorter, but]
+The dark-matter halo of the Milky Way is expected to be triaxial on large scales
+  and filled with substructure on all smaller scales.
+It is hoped that streams or shells of stars produced by tidal disruption
+  will produce very precise measures of the gravitational potential,
+  to test these predictions.
+% Aims
+We develop a justified method for inferring the gravitational potential
+  using precise measurements of stars in tidal streams
+  based on an approximate generative model.
+Our method does not rely on integrability of the potential;
+  we don't expect there to be any good approximation to the Milky Way potential that is integrable.
+% Methods: 
+Our model is probabilistic, with a likelihood function and priors on the parameters.
+It is based on the idea that the stars in a tidal structure
+  were close in phase space at some point in the past;
+  model parameters include a ``disruption time'' [APW adjust language] for each star,
+  and phase-space coordinates for the progenitor.
+It accounts properly for finite observational uncertainties and missing data.
+We test our method with simulated data in a non-integrable triaxial potential,
+  with realistic observational uncertainties for the next generation of surveys of RR Lyrae stars and the like.
+Importantly, the fake data were generated with an n-body simulation,
+  not with the approximate model underlying our likelihood function.
+% Results:
+We find that even with a small number (4 to 16) of well-measured stars,
+  we can precisely infer properties of the triaxial potential,
+  missing phase-space coordinates of the stars,
+  and the location of the progenitor object.
+The method generalizes to to encompass mixtures of independent debris structures,
+  or situations in which star membership in structures is unknown.
+[Get more quantitative here?]
+% Conclusions: 
 \end{abstract}
@@
 \begin{center}
-\includegraphics[width=\textwidth]{/Users/adrian/projects/streams/plots/paper2/rel_dists.pdf}
+% \includegraphics[width=\textwidth]{/Users/adrian/projects/streams/plots/paper2/rel_dists.pdf}
 \caption{ TODO }\label{fig:reldist}

### 2014-02-15 text/paper2/ms.tex
small samples text

@@
 
-The potential of the inner galaxy is likely well-described by a (set of) closely-integrable potentials, thus the dynamics of collisionless systems is plausibly captured by a transformation to some integral-of-motion-space, e.g., action-angle coordinates.
+The potential of the inner galaxy is likely well-described by a (set of) nearly-integrable potentials, thus the dynamics of collisionless systems is plausibly captured by a transformation to some integral-of-motion-space, e.g., action-angle coordinates.
 However there is uncertainty in the exact shape, radial profile, and graininess of the outer halo.
@@
 
-[Sometimes stellar samples might get extremely small.  For example,
-  \gaia\ might find some very low-mass cold stellar streams.  For
-  another, there might only be a couple RRL stars in a particular
-  stream, but those RRL stars might be individually extremely well
-  measured in six-space.]
+Although the \gaia\ project will deliver hundreds of millions of stars useful for dynamical inference,
+  it still might be that the most relevant stellar samples for inferring the Milky Way potentail might be very small in the end.
+One reason is that if distances turn out to be important, stars that produce good distance estimates might be much more
+  valuable than typical stars in the sample.
+At halo distances, \gaia's parallaxes are not expected to be better than the distance indicators produced by the best
+  variable stars observed at the best wavelengths; for example, near-infrared observations of RR Lyrae stars
+  can in principle deliver XX percent photometric distances at essentially all halo distances (CITATIONS).
+These valuable stars are rare (and age and metallicity dependent);
+  there could be many cold structures in the Milky Way halo that are highly constraining
+  on the potential in principle, but which contain only a few good distance-indicating members.
+That said, it is not yet known what the trade-offs are between having many stars at low precision and a few at high precision,
+  nor is it known how valuable distance information really is, when a structure contains many precisely observed members.
 
-[Even if you have many stars, some might be orders of magnitude more constraining (Bovy, ..., Hogg)]
+Another reason to think about small samples is more conceptual:
+In dynamical inferences that go beyond simple virial or Jeans arguments,
+  inference depends on the non-linear dynamical evolution of populations in phase space.
+The nonlinearities, combined with heteroskedastic data and requirements to marginalize over phase-space distribution functions,
+  lead naturally to situations in which some stars are \emph{far more informative} than others.
+For a simple example, in a toy study of the Solar System,
+  phase-space distribution marginalization was shown to make some planets are far more informative than others
+  about the gravitational potential (CITE Bovy et al).
+This is expected to be a generic feature in dynamical problems;
+  it might be that a very small fraction of \gaia's catalog entries
+  deliver a large fraction of the information about the Milky Way's potential and formation history.
 

### 2014-02-15 text/paper2/ms.tex
typos

@@
 Although the \gaia\ project will deliver hundreds of millions of stars useful for dynamical inference,
-  it still might be that the most relevant stellar samples for inferring the Milky Way potentail might be very small in the end.
+  it still might be that the most relevant stellar samples for inferring the Milky Way potential might be very small in the end.
 One reason is that if distances turn out to be important, stars that produce good distance estimates might be much more
@@
   can in principle deliver XX percent photometric distances at essentially all halo distances (CITATIONS).
-These valuable stars are rare (and age and metallicity dependent);
+These valuable stars are rare (and their abundances are age and metallicity dependent);
   there could be many cold structures in the Milky Way halo that are highly constraining

### 2014-02-16 text/paper2/ms.tex
figure stuff

--- 2014-02-15 text/paper2/ms.tex
@@
 		              New York, NY 10027, USA}
-\altaffiltext{\adrn}{To whom correspondence should be addressed: adrn@astro.columbia.edu}
+\altaffiltext{\adrn}{To whom correspondence should be addressed: \email{adrn@astro.columbia.edu}}
 \altaffiltext{\nyu}{Center for Cosmology and Particle Physics,
@@
 \begin{abstract}
-% Context
- [this needs to get shorter, but]
-The dark-matter halo of the Milky Way is expected to be triaxial on large scales
-  and filled with substructure on all smaller scales.
-It is hoped that streams or shells of stars produced by tidal disruption
-  will produce very precise measures of the gravitational potential,
-  to test these predictions.
-% Aims
-We develop a justified method for inferring the gravitational potential
-  using precise measurements of stars in tidal streams
-  based on an approximate generative model.
-Our method does not rely on integrability of the potential;
-  we don't expect there to be any good approximation to the Milky Way potential that is integrable.
-% Methods: 
-Our model is probabilistic, with a likelihood function and priors on the parameters.
-It is based on the idea that the stars in a tidal structure
-  were close in phase space at some point in the past;
-  model parameters include a ``disruption time'' [APW adjust language] for each star,
-  and phase-space coordinates for the progenitor.
-It accounts properly for finite observational uncertainties and missing data.
-We test our method with simulated data in a non-integrable triaxial potential,
-  with realistic observational uncertainties for the next generation of surveys of RR Lyrae stars and the like.
-Importantly, the fake data were generated with an n-body simulation,
-  not with the approximate model underlying our likelihood function.
-% Results:
-We find that even with a small number (4 to 16) of well-measured stars,
-  we can precisely infer properties of the triaxial potential,
-  missing phase-space coordinates of the stars,
-  and the location of the progenitor object.
-The method generalizes to to encompass mixtures of independent debris structures,
-  or situations in which star membership in structures is unknown.
-[Get more quantitative here?]
-% Conclusions: 
+Context: Connecting cosmological simulations to observed galaxies is difficult because baryonic processes not well understood/modeled but also because they are gravitationally dominated by unobservable dark matter.
+One powerful measure is the shape of the potential -- triaxial, density profile, etc.
+Tidal streams are excellent tracers of mass.
+
+Aims: Present a method for inferring higher order properties of the Galactic potential using debris from disrupted satellites.
+
+Methods:
+
+Results: Can simultaneously constrain properties of progenitor, missing dimensions, and potential with a simple model for disrupted debris in phase-space.
+The method works well with small numbers of stars and with finite or large observational uncertainties or missing dimensions and generalizes naturally to multiple independent debris structures.
+
+Conclusions:
+
+%Informative phase-space structures, and in particular those created by
+%disrupting stellar systems or revealed by chemical tags, may deliver
+%very precise measures of the gravitational potential in the Milky Way
+%Halo.  Here we show that even a single pair of stars---two stars that
+%are known (for some non-kinematic reason) to be likely to be
+%associated with one another at birth---could provide a significant
+%constraint on the potential.  The inference is based on a
+%probabilistic generative model---a likelihood function and priors over
+%nuisance parameters---that evolves a past putative origin disruption
+%event forward in time to the present day.  The time, six-volume, and
+%phase-space location of the origin event are nuisance parameters in
+%the model and marginalized away.  The method makes no assumption of
+%integrability, works with finite or even large observational
+%uncertainties, does not require all dimensions of phase space to be
+%observed, handles non-zero probability that the two stars are not in
+%fact associated, and generalizes naturally to larger numbers of stars
+%and multiple independent structures.  Applications to the GD-1 cold
+%stellar stream and current surveys for RR Lyrae stars in the Halo are
+%discussed.
+TODO...
 \end{abstract}
@@
 In \paperone, we define a set of coordinates relative to the instantaneous phase-space position of the progenitor, $(\bs{r}_{p},\bs{v}_{p})$, normalized by the instantaneous tidal radius and escape velocity.
-Here we (1) scale the velocity coordinates by the velocity dispersion of the satellite instead of the escape velocity, (2) reduce the dimensionality by considering only the magnitude of these vectors, and (3) take the logarithm of the magnitude,
+Here we (1) instead scale the velocity coordinates by the velocity dispersion of the satellite, (2) reduce the dimensionality by considering only the magnitude of these vectors, and (3) take the logarithm of the magnitude,
 \begin{equation}
@@
 \end{equation}
+Figure~\ref{fig:reldists} shows distributions of $R$ and $V$ at $t=t_{\rm ub}$ for all particles as a function of progenitor mass from the simulations described in Section~\ref{sec:sims}.
+In these scaled coordinates, the distributions
 

### 2014-02-16 text/paper2/ms.tex
fixed hoggs placeholders

@@
 Tidal streams are dynamically cold systems --- debris typically have small distributions of energy and angular momentum --- and thus require orders of magnitude fewer tracers than a random sample to get constraints of comparable accuracy.
-For example, in the simplest case we might assume that debris stars are actually still on the same orbit as their progenitor system \citep[a \emph{wrong} assumption, see e.g.,][]{eyre2011}.
+For example, in the simplest case we might assume that debris stars are actually still on the same orbit as their progenitor system \citep[a \emph{wrong} assumption, see e.g.,][]{eyre11}.
 Then we can imagine measuring the full-space velocities ${\bf v}$ at different points ${\bf x}$ in the structure (e.g., along a stream), which would give us a direct measure of differences in a potential, $\Phi$.
@@
   variable stars observed at the best wavelengths; for example, near-infrared observations of RR Lyrae stars
-  can in principle deliver XX percent photometric distances at essentially all halo distances (CITATIONS).
+  can in principle deliver $\sim$2\% percent photometric distances at essentially all halo distances \citep{madore12}.
 These valuable stars are rare (and their abundances are age and metallicity dependent);
@@
   phase-space distribution marginalization was shown to make some planets are far more informative than others
-  about the gravitational potential (CITE Bovy et al).
+  about the gravitational potential \citep{bovy10}.
 This is expected to be a generic feature in dynamical problems;
@@
 \section{Simulations}\label{sec:sims}
-[David H.]
+[David Hendel]
 We [...] N-body simulations run with an SCF code [...]. [describe simulations here] [Describe tub]

### 2014-02-16 text/paper2/ms.tex
tweaking the abstract

@@
 \newcommand{\project}[1]{\textsl{#1}}
-\newcommand{\gaia}{\project{Gaia}~}
+\newcommand{\gaia}{\project{Gaia}}
 \newcommand{\spitzer}{\project{Spitzer}~}
@@
 % Context
- [this needs to get shorter, but]
 The dark-matter halo of the Milky Way is expected to be triaxial on large scales
-  and filled with substructure on all smaller scales.
+  and filled with substructure on all scales.
 It is hoped that streams or shells of stars produced by tidal disruption
@@
 Our method does not rely on integrability of the potential;
-  we don't expect there to be any good approximation to the Milky Way potential that is integrable.
+  we don't expect there to be any good approximation to the large-scale 
+  Milky Way potential that is integrable.
 % Methods: 
@@
 It is based on the idea that the stars in a tidal structure
-  were close in phase space at some point in the past;
-  model parameters include a ``disruption time'' [APW adjust language] for each star,
-  and phase-space coordinates for the progenitor.
-It accounts properly for finite observational uncertainties and missing data.
-We test our method with simulated data in a non-integrable triaxial potential,
-  with realistic observational uncertainties for the next generation of surveys of RR Lyrae stars and the like.
+  were close in phase space at some point in the past.
+  Model parameters include the present-day phase-space position for each star
+  and the progenitor, and thus properly accounts for finite observational uncertainties 
+  and missing data.
+We test our method with simulated data in a non-integrable triaxial potential, with realistic observational uncertainties for RR Lyrae stars.
 Importantly, the fake data were generated with an n-body simulation,
@@
 To account for this, methods have been proposed that model the debris as an orbit plus some dispersion or offset, either in phase space \citep[e.g.,][]{eyre09a, varghese11, kuepper12} or action-angle coordinates \citep{eyre11, sanders13b, bovy14, sanders14}.
-Other statistical methods have been proposed \citep[][]{penarrubia12, sanderson14} that may prove powerful when applied to, e.g., data from the \gaia mission, where full 6D coordinates will be known for large samples of stars in the halo --- and therefore many debris structures --- but stream membership is not known for all stars (however these statistical methods do not account for observational errors).
+Other statistical methods have been proposed \citep[][]{penarrubia12, sanderson14} that may prove powerful when applied to, e.g., data from the \gaia\, mission, where full 6D coordinates will be known for large samples of stars in the halo --- and therefore many debris structures --- but stream membership is not known for all stars (however these statistical methods do not account for observational errors).
 
@@
 
-Although the \gaia\ project will deliver hundreds of millions of stars useful for dynamical inference,
+Although the \gaia\, project will deliver hundreds of millions of stars useful for dynamical inference,
   it still might be that the most relevant stellar samples for inferring the Milky Way potential might be very small in the end.
-One reason is that if distances turn out to be important, stars that produce good distance estimates might be much more
-  valuable than typical stars in the sample.
+For instance, stars that produce good distance estimates might be much more valuable than typical stars in the sample.
 At halo distances, \gaia's parallaxes are not expected to be better than the distance indicators produced by the best
@@
 Another reason to think about small samples is more conceptual:
-In dynamical inferences that go beyond simple virial or Jeans arguments,
+in dynamical inferences that go beyond simple virial or Jeans arguments,
   inference depends on the non-linear dynamical evolution of populations in phase space.
@@
   about the gravitational potential \citep{bovy10}.
-This is expected to be a generic feature in dynamical problems;
+This is expected to be a generic feature in dynamical problems:
   it might be that a very small fraction of \gaia's catalog entries

### 2014-02-16 text/paper2/ms.tex
some notes to self

@@
 
-\section{Scalings}
+\subsection{Scalings}
 
@@
 
-a phase-space distance, $D_{ps}$, as the sum of the normalized, relative position, $\frac{\bs{r}-\bs{r}_p}{\rtide}$, and velocity, $\frac{\bs{v}-\bs{v}_p}{v_{\rm esc}}$, for a stream star.
+- Point out constant $\sigma$ over mass\\
+- Good until 2.5e9\\
+- Proper description models 2D distribution, but [...]
+
+%a phase-space distance, $D_{ps}$, as the sum of the normalized, relative position, $\frac{\bs{r}-\bs{r}_p}{\rtide}$, and velocity, $\frac{\bs{v}-\bs{v}_p}{v_{\rm esc}}$, for a stream star.
 We minimized the generalized variance of the 6D distribution over each normalized coordinate evaluated at the time that each star has minimum $D_ps$.
@@
 However, this method does not properly incorporate observational uncertainties.
-
-Using the simulations described in Section~\ref{sec:sims}
 
@@
 
-The disruption of a satellite galaxy due to tidal shocking [...connect to Choi work some how?] We model the present-day position and velocity of a disrupted star as the result of sampling from isotropic, log-normal distributions in relative distance, $R=\left\vert \bs{r}-\bs{r}_p \right\vert$, and relative velocity, $V=\left\vert \bs{v}-\bs{v}_p \right\vert$, at the time the star comes unbound from the progenitor, evolved to present day in the parent potential, $\Phi$.
+We model the present-day position and velocity of a disrupted star as the result of sampling from isotropic, log-normal distributions in relative distance, $R=\left\vert \bs{r}-\bs{r}_p \right\vert$, and relative velocity, $V=\left\vert \bs{v}-\bs{v}_p \right\vert$, at the time the star comes unbound from the progenitor, evolved to present day in the parent potential, $\Phi$.
 The distributions over $\ln R$ and $\ln V$ are centered on the logarithm of the instantaneous tidal radius, $\rtide$, and velocity dispersion, $\sigma_v$, of the progenitor at time $\tau$: $\mathcal{N}(\ln R \given \ln\rtide)\vert_{t=\tau}$ and $\mathcal{N}(\ln V \given \ln \sigma_v)\vert_{t=\tau}$.

### 2014-02-16 scripts/figures/paper2_figures.py
before cleanup

@@ -435,6 +435,89 @@ def norm_dist(**kwargs):
     fig.tight_layout()
     fig.savefig(os.path.join(plot_path, "norm_dists.pdf".format(_m)))
 
+def twod_dist(**kwargs):
+
+    m = float(kwargs["m"])
+    sgr = SgrSimulation(kwargs["m"])
+    p = sgr.particles(N=5000, expr="tub!=0")
+    p_bound = sgr.particles(N=0, expr="tub==0")
+    v_disp = np.sqrt(np.var(p_bound["vx"]) + \
+                     np.var(p_bound["vy"]) + \
+                     np.var(p_bound["vz"])).value[0]
+    s = sgr.satellite()
+
+    lm10_potential = LawMajewski2010()
+
+    X = np.vstack((s._X[...,:3], p._X[...,:3].copy()))
+    V = np.vstack((s._X[...,3:], p._X[...,3:].copy()))
+    integrator = LeapfrogIntegrator(lm10_potential._acceleration_at,
+                                    np.array(X), np.array(V),
+                                    args=(X.shape[0], np.zeros_like(X)))
+    ts, rs, vs = integrator.run(t1=sgr.t1, t2=sgr.t2, dt=-1.)
+
+    s_orbit = np.vstack((rs[:,0][:,np.newaxis].T, vs[:,0][:,np.newaxis].T)).T
+    p_orbits = np.vstack((rs[:,1:].T, vs[:,1:].T)).T
+    t_idx = np.array([np.argmin(np.fabs(ts - t)) for t in p.tub])
+
+    p_x = np.array([p_orbits[jj,ii] for ii,jj in enumerate(t_idx)])
+    s_x = np.array([s_orbit[jj,0] for jj in t_idx])
+
+    diff = p_x-s_x
+    rel_r = np.sqrt(np.sum(diff[:,:3]**2, axis=-1))
+    rel_v = np.sqrt(np.sum(diff[:,3:]**2, axis=-1))
+
+    r_tide = lm10_potential._tidal_radius(m, s_x)*1.6
+    lnR = np.log(rel_r/r_tide)
+    lnV = np.log(rel_v/v_disp)
+
+    plt.clf()
+    gs = GridSpec(3,3)
+
+    fig = plt.figure(figsize=(12,12))
+
+    xlims = (-2,4)
+    ylims = (-2,4)
+    nbins = 50
+
+    plt.subplot(gs[0,:2])
[…]

### 2014-02-16 scripts/figures/paper2_figures.py
new figure

@@ -163,277 +163,51 @@ def potential_contours(**kwargs):
     fig.tight_layout(pad=1.5, h_pad=0.)
     fig.savefig(filename)
 
-def gaussian(x, mu, sigma):
-    return 1./np.sqrt(2*np.pi*sigma**2) * np.exp(-0.5 * ((x-mu)/sigma)**2)
+def simulated_streams(**kwargs):
 
-def rel_dist(**kwargs):
+    filename = os.path.join(plot_path, "simulated_streams.pdf")
+    fig,axes = plt.subplots(2,4,figsize=(14,7.5),
+                            sharex=True, sharey=True)
 
-    nbins = 50
-    fig,axes = plt.subplots(4,2,figsize=(8.5,11), sharex='col', sharey=True)
+    ticks = [-100,-50,0,50]
+    alpha = 0.25
+    rcparams = {'lines.linestyle' : 'none',
+                'lines.linewidth' : 1.,
+                'lines.marker' : '.',
+                'axes.facecolor' : '#ffffff'}
 
-    #_m = kwargs["mass"]
-    for kk,_m in enumerate(["2.5e6","2.5e7","2.5e8","2.5e9"]):
-        m = float(_m)
-        sgr = SgrSimulation(_m)
-        p = sgr.particles(N=5000, expr="tub!=0")
-        p_bound = sgr.particles(N=0, expr="tub==0")
-        v_disp = np.sqrt(np.var(p_bound["vx"]) + \
-                         np.var(p_bound["vy"]) + \
-                         np.var(p_bound["vz"])).value[0]
-        s = sgr.satellite()
+    for ii,_m in enumerate(range(6,9+1)):
+        mass = "2.5e{}".format(_m)
 
-        lm10_potential = LawMajewski2010()
+        m = float(mass)
+        sgr = SgrSimulation(mass)
+        p = sgr.particles(N=5000)
 
-        X = np.vstack((s._X[...,:3], p._X[...,:3].copy()))
-        V = np.vstack((s._X[...,3:], p._X[...,3:].copy()))
-        integrator = LeapfrogIntegrator(lm10_potential._acceleration_at,
-                                        np.array(X), np.array(V),
-                                        args=(X.shape[0], np.zeros_like(X)))
-        ts, rs, vs = integrator.run(t1=sgr.t1, t2=sgr.t2, dt=-1.)
[…]

### 2014-02-17 scripts/infer_potential.py
hot burn in

@@ -75,19 +75,19 @@ def main(config_file, mpi=False, threads=None, overwrite=False):
         # sample starting positions
         p0 = model.sample_priors(size=nwalkers)
 
-        # get the sampler
-        sampler = si.StreamModelSampler(model, nwalkers, pool=pool)
-
         if nburn > 0:
             time0 = time.time()
             logger.info("Burning in sampler for {} steps...".format(nburn))
 
             pos, xx, yy = sampler.run_mcmc(p0, nburn)
 
+            model.lnpargs[3] = 0.01
+            sampler = si.StreamModelSampler(model, nwalkers, pool=pool)
+
             best_idx = sampler.flatlnprobability.argmax()
             best_pos = sampler.flatchain[best_idx]
 
-            std = np.std(p0, axis=0) / 5.
+            std = np.std(p0, axis=0) / 10.
             pos = np.array([np.random.normal(best_pos, std) \
                             for kk in range(nwalkers)])
 
@@ -98,27 +98,15 @@ def main(config_file, mpi=False, threads=None, overwrite=False):
         else:
             pos = p0
 
+        model.lnpargs[3] = 1.
+        sampler = si.StreamModelSampler(model, nwalkers, pool=pool)
+
         logger.info("Running {} walkers for {} iterations of {} steps..."\
                     .format(nwalkers, niter, nsteps//niter))
 
         if nsteps > 0:
             time0 = time.time()
-            for ii in range(niter):
-                logger.debug("Iteration: {}".format(ii))
-                pos, prob, state = sampler.run_mcmc(pos, nsteps//niter)
-
-                # if any of the samplers have less than 5% acceptance,
-                #  start them from new positions sampled from the best position
-                acc_frac_test = sampler.acceptance_fraction < 0.05
-                if np.any(acc_frac_test):
-                    nbad = np.sum(acc_frac_test)
-                    med_pos = np.median(sampler.flatchain, axis=0)
-                    std = np.std(sampler.flatchain, axis=0)
-                    new_pos = sample_ball(med_pos, std, size=nwalkers)
-
-                    for jj in range(nwalkers):
-                        if acc_frac_test[jj]:
-                            pos[jj] = new_pos[jj]
+            pos, prob, state = sampler.run_mcmc(pos, nsteps)
 
             t = time.time() - time0
             logger.debug("Spent {} seconds on main sampling...".format(t))

### 2014-02-17 streams/inference/core.py

@@ -321,11 +321,11 @@ class StreamModel(object):
                                               self.true_satellite.mass,
                                               self.true_satellite.vdisp)
 
-        return np.sum(ln_like) + np.sum(ln_prior)
-        # try:
-        #     return (np.sum(ln_like) + np.sum(ln_prior))*args[3]
-        # except:
-        #     return np.sum(ln_like) + np.sum(ln_prior)
+        #return np.sum(ln_like) + np.sum(ln_prior)
+        try:
+            return (np.sum(ln_like) + np.sum(ln_prior))*args[3]
+        except:
+            return np.sum(ln_like) + np.sum(ln_prior)
 
     def __call__(self, p):
         # TODO: each call, adjust temperature according to self.annealing?

### 2014-02-17 text/paper2/ms.tex
new figure
added figure

@@
 
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=\textwidth]{../../plots/paper2/simulated_streams.pdf}
+\caption{ TODO }\label{fig:sims}
+\end{center}
+\end{figure}
+
 \section{Method}
@@
 \end{equation}
-Figure~\ref{fig:reldists} shows distributions of $R$ and $V$ at $t=t_{\rm ub}$ for all particles as a function of progenitor mass from the simulations described in Section~\ref{sec:sims}.
-In these scaled coordinates, the distributions
-
-- Point out constant $\sigma$ over mass\\
-- Good until 2.5e9\\
-- Proper description models 2D distribution, but [...]
-
-%a phase-space distance, $D_{ps}$, as the sum of the normalized, relative position, $\frac{\bs{r}-\bs{r}_p}{\rtide}$, and velocity, $\frac{\bs{v}-\bs{v}_p}{v_{\rm esc}}$, for a stream star.
-We minimized the generalized variance of the 6D distribution over each normalized coordinate evaluated at the time that each star has minimum $D_ps$.
-This simple idea proved quite powerful for constraining properties of the LM10 potential used simulated observations of their N-body simulation.
-However, this method does not properly incorporate observational uncertainties.
+Figure~\ref{fig:R_V} shows distributions of $R$ and $V$ --- computed at $t=t_{\rm ub}$ for each particle --- as a function of progenitor mass from the simulations described in Section~\ref{sec:sims}.
+In these scaled, log coordinates we model distributions in $R$ and $V$ as Gaussians.
+We find that for masses $\lesssim10^9M_\odot$, the variance ($\sigma^2$ in plot panels) remains approximately constant and the mean stays fixed at 0.
+For the largest mass simulation ($2.5\times10^9M_\odot$), the perturbation to the total potential due to the progenitor halo becomes significant and thus a large fraction of the stars are disrupted through interactions with the subhalo.
+For this article, we ignore covariances between $R$ and $V$.
+The
 
@@
 \begin{center}
-% \includegraphics[width=\textwidth]{/Users/adrian/projects/streams/plots/paper2/rel_dists.pdf}
-\caption{ TODO }\label{fig:reldist}
+\includegraphics[width=\textwidth]{../../plots/paper2/R_V.pdf}
+\caption{ TODO }\label{fig:R_V}
 \end{center}
 \end{figure}
-
-%The core of this method provides a basis for the probabilistic model presented below.
 

### 2014-02-17 scripts/figures/paper2_figures.py

@@ -75,6 +75,52 @@ def graphical_model(**kwargs):
     #pgm.figure.savefig("classic.pdf")
     pgm.figure.savefig(filename, dpi=150)
 
+def simulated_streams(**kwargs):
+
+    filename = os.path.join(plot_path, "simulated_streams.pdf")
+    fig,axes = plt.subplots(2,4,figsize=(14,7.5),
+                            sharex=True, sharey=True)
+
+    ticks = [-100,-50,0,50]
+    alpha = 0.25
+    rcparams = {'lines.linestyle' : 'none',
+                'lines.linewidth' : 1.,
+                'lines.marker' : '.',
+                'axes.facecolor' : '#ffffff'}
+
+    for ii,_m in enumerate(range(6,9+1)):
+        mass = "2.5e{}".format(_m)
+        m = float(mass)
+
+        sgr = SgrSimulation(mass)
+        p = sgr.particles(N=5000)
+
+        with rc_context(rc=rcparams):
+            axes[0,ii].text(0.5, 1.04, r"{}$M_\odot$".format(mass),
+                   horizontalalignment='center',
+                   fontsize=24,
+                   transform=axes[0,ii].transAxes)
+
+            axes[0,ii].plot(p["x"].value, p["y"].value,
+                            alpha=alpha)
+            axes[1,ii].plot(p["x"].value, p["z"].value,
+                            alpha=alpha)
+            axes[1,ii].set_xticks(ticks)
+            axes[1,ii].set_xlabel("$X$ [kpc]")
+
+    axes[0,0].set_ylabel("$Y$ [kpc]")
+    axes[1,0].set_ylabel("$Z$ [kpc]")
+
+    axes[0,0].set_yticks(ticks)
+    axes[1,0].set_yticks(ticks)
+    axes[-1,-1].set_xlim(-100,75)
+    axes[-1,-1].set_ylim(-100,75)
+
+    fig.tight_layout()
+    fig.subplots_adjust(top=0.92, hspace=0.025, wspace=0.1)
+    #fig.subplots_adjust(hspace=0., wspace=0.075)
+    fig.savefig(filename)
+
 def potential_contours(**kwargs):
 
     filename = os.path.join(plot_path, "potentials.pdf")
@@ -163,52 +209,6 @@ def potential_contours(**kwargs):
     fig.tight_layout(pad=1.5, h_pad=0.)
     fig.savefig(filename)
 
[…]

### 2014-02-17 text/paper2/ms.tex
switch to lnR

@@
 \begin{equation}
-  R = \ln\left(\frac{\left\vert\bs{r} - \bs{r}_{p}\right\vert}{R_{\rm tide}}\right)\,\,\,\,,\,\,\,\,
-  V = \ln\left(\frac{\left\vert\bs{v} - \bs{v}_{p}\right\vert}{\sigma_v}\right).
+  \ln R = \ln\left(\frac{\left\vert\bs{r} - \bs{r}_{p}\right\vert}{R_{\rm tide}}\right)\,\,\,\,,\,\,\,\,
+  \ln V = \ln\left(\frac{\left\vert\bs{v} - \bs{v}_{p}\right\vert}{\sigma_v}\right).
 \end{equation}
-Figure~\ref{fig:R_V} shows distributions of $R$ and $V$ --- computed at $t=t_{\rm ub}$ for each particle --- as a function of progenitor mass from the simulations described in Section~\ref{sec:sims}.
-In these scaled, log coordinates we model distributions in $R$ and $V$ as Gaussians.
+Figure~\ref{fig:R_V} shows distributions of $\ln R$ and $\ln V$ --- computed at $t=t_{\rm ub}$ for each particle --- as a function of progenitor mass from the simulations described in Section~\ref{sec:sims}.
+In these scaled coordinates we model distributions in $R$ and $V$ as log-normal and ignore covariances between $R$ and $V$.
 We find that for masses $\lesssim10^9M_\odot$, the variance ($\sigma^2$ in plot panels) remains approximately constant and the mean stays fixed at 0.
 For the largest mass simulation ($2.5\times10^9M_\odot$), the perturbation to the total potential due to the progenitor halo becomes significant and thus a large fraction of the stars are disrupted through interactions with the subhalo.
-For this article, we ignore covariances between $R$ and $V$.
-The
 
@@
 
+The scale-radius and velocity dispersion scale with the mass of the progenitor. [...observational errors] [ $D_{\rm ps}$]
 [Cold vs. Hot streams in the context of observational errors]
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=\textwidth]{../../plots/paper2/num_recombine.png}
+\caption{ TODO }\label{fig:R_V}
+\end{center}
+\end{figure}
 

### 2014-02-28 scripts/make_streams2.py
generative model

@@ -1,0 +1,199 @@
+# coding: utf-8
+
+""" Turn our model into a generative model. """
+
+from __future__ import division, print_function
+
+__author__ = "adrn <adrn@astro.columbia.edu>"
+
+# Standard library
+import os, sys
+
+# Third-party
+import astropy.units as u
+import astropy.coordinates as coord
+import matplotlib.pyplot as plt
+import numpy as np
+from scipy.signal import argrelmin
+
+# Project
+import streams.io as io
+from streams.coordinates import _gc_to_hel, _hel_to_gc
+from streams.coordinates.frame import heliocentric
+from streams.potential.lm10 import LawMajewski2010
+from streams.integrate import LeapfrogIntegrator
+from streams.integrate.stream_generator import StreamGenerator
+from streams.util import project_root
+
+plot_path = os.path.join(project_root, "plots")
+
+potential = LawMajewski2010()
+nparticles = 2000
+T = 4000.
+dt = 0.1
+_m = "2.e4"
+np.random.seed(42)
+
+# Set up a bit to specify leading or trailing
+tail_bit = np.ones(nparticles)
+tail_bit[:nparticles//2] = -1.
+
+mass = float(_m)
+# simulation = io.SgrSimulation(mass=_m)
+# particles = simulation.particles(N=1000, expr="tub!=0")
+# satellite = simulation.satellite()\
+#                       .to_frame(heliocentric)
+
+# s_hel = satellite._X.copy()
+# s_gc = _hel_to_gc(s_hel)
+s_gc = np.array([[8.363919011, 0.243352771, 16.864546659,
+                  -0.04468993,-0.12392801,-0.01664498]]) # Pal5
+s_hel = _gc_to_hel(s_gc)
+
+# First integrate the orbit of the satellite back to get initial conditions
+acc = np.zeros_like(s_gc[:,:3])
+integrator = LeapfrogIntegrator(potential._acceleration_at,
+                                np.array(s_gc[:,:3]), np.array(s_gc[:,3:]),
+                                args=(s_gc.shape[0], acc))
+
+t, rs, vs = integrator.run(t1=T, t2=0, dt=-dt)
+init_r,init_v = rs[-1], vs[-1]
+
[…]

### 2014-02-28 streams/integrate/stream_generator.py

@@ -1,0 +1,153 @@
+# coding: utf-8
+
+""" """
+
+from __future__ import division, print_function
+
+__author__ = "adrn <adrn@astro.columbia.edu>"
+
+# Standard library
+import os, sys
+
+# Third-party
+import numpy as np
+from astropy.utils import isiterable
+
+from .timespec import _parse_time_specification
+
+__all__ = ["StreamGenerator"]
+
+class StreamGenerator(object):
+
+    def __init__(self, potential, satellite_orbit, satellite_mass, acc_args=()):
+        """ Initialize a leapfrog integrator given a function to compute
+            the accelerations and initial conditions.
+
+            Naming convention for variables:
+                im1 -> i-1
+                im1_2 -> i-1/2
+                ip1 -> i+1
+                ip1_2 -> i+1/2
+
+            Initial position and velocity should have shape (ndim,nparticles)
+            e.g., for 100 particles in 3D cartesian coordinates, the position
+            array should have shape (100,3). For a single particle, (3,) is
+            accepted and converted to (1,3).
+
+            `acceleration_function` should accept the array of position(s) and
+            optionally a set of arguments specified by `acceleration_args`.
+
+            For details on the algorithm, see:
+                http://ursa.as.arizona.edu/~rad/phys305/ODE_III/node11.html
+
+            Parameters
+            ----------
+            acceleration_func : func
+            r_initial : array_like
+            v_initial : array_like
+            args : tuple (optional)
+                Any extra arguments for the acceleration function.
+
+        """
+
+        if not hasattr(potential, '_acceleration_at'):
+            raise ValueError("Potential must have an acceleration method.")
+
+        self.potential = potential
+        self._acc_args = acc_args
+
+        # Validate initial conditions
+        if not isinstance(satellite_orbit, np.ndarray):
+            satellite_orbit = np.array(satellite_orbit)
+
+        self.satellite_orbit = satellite_orbit
+        self.satellite_mass = satellite_mass
+
+    def step(self, dt):
+        """ Step forward the positions and velocities by the given timestep """
+
+        self.r_im1 += self.v_im1_2*dt
+        a_i = self.acc(self.r_im1, *self._acc_args)
+
[…]

### 2014-02-28 text/debris_properties/ms.tex
new paper

--- 2014-02-17 text/paper2/ms.tex
@@
 \newcommand{\project}[1]{\textsl{#1}}
-\newcommand{\gaia}{\project{Gaia}}
-\newcommand{\spitzer}{\project{Spitzer}~}
 \newcommand{\given}{\,|\,}
@@
 
-\newcommand{\D}{{\bf D}}
-\newcommand{\W}{{\bf W}}
-\newcommand{\X}{{\bf X}}
-\newcommand{\J}{{\boldsymbol J}}
-\newcommand{\bSigma}{{\bf \Sigma}}
-\newcommand{\bsigma}{\boldsymbol\sigma}
 \newcommand{\rtide}{r_{\rm tide}}
@@
 
-\newcommand{\paperone}{Paper 1}
-
 \begin{document}
 
-\title{Inferring the gravitational potential of the Milky Way with a few precisely measured stars}
+\title{The Phase-space Properties of Tidal Debris}
 \author{Adrian M. Price-Whelan\altaffilmark{\colum,\adrn}, 
-	    David W. Hogg\altaffilmark{\nyu,\mpia}, 
 	    Kathryn V. Johnston\altaffilmark{\colum}, 
@@
 \newcommand{\adrn}{2}
-\newcommand{\nyu}{3}
-\newcommand{\mpia}{4}
 \altaffiltext{\colum}{Department of Astronomy, 
@@
 \altaffiltext{\adrn}{To whom correspondence should be addressed: adrn@astro.columbia.edu}
-\altaffiltext{\nyu}{Center for Cosmology and Particle Physics,
-                      Department of Physics, New York University,
-                      4 Washington Place, New York, NY, 10003, USA}
-\altaffiltext{\mpia}{Max-Planck-Institut f\"ur Astronomie,
-                     K\"onigstuhl 17, D-69117 Heidelberg, Germany}
 
@@
 % Context
-The dark-matter halo of the Milky Way is expected to be triaxial on large scales
-  and filled with substructure on all scales.
-It is hoped that streams or shells of stars produced by tidal disruption
-  will produce very precise measures of the gravitational potential,
-  to test these predictions.
+The tidal disruption of collisionless systems of particles forms observable debris structures in galaxies.
+Cold debris structures are excellent probes of the underlying potential or mass distribution of galaxies and are detectable over a wide range of distances from the inner [...] out to the virial radius.
 % Aims
-We develop a justified method for inferring the gravitational potential
-  using precise measurements of stars in tidal streams
-  based on an approximate generative model.
-Our method does not rely on integrability of the potential;
-  we don't expect there to be any good approximation to the large-scale 
-  Milky Way potential that is integrable.
+We show that the properties of the debris can be thought of as generated from two distinct processes --- tidal leaking and tidal shocking --- and the extent of the debris scales in energy (or phase-space) with the mass-ratio between the system and the parent potential.
 % Methods: 
-Our model is probabilistic, with a likelihood function and priors on the parameters.
-It is based on the idea that the stars in a tidal structure
-  were close in phase space at some point in the past.
-  Model parameters include the present-day phase-space position for each star
-  and the progenitor, and thus properly accounts for finite observational uncertainties 
-  and missing data.
-We test our method with simulated data in a non-integrable triaxial potential, with realistic observational uncertainties for RR Lyrae stars.
-Importantly, the fake data were generated with an n-body simulation,
-  not with the approximate model underlying our likelihood function.
+We run N-body simulations of satellite galaxies and globular clusters on two qualitatively distinct orbits [...]
 % Results:
-We find that even with a small number (4 to 16) of well-measured stars,
-  we can precisely infer properties of the triaxial potential,
-  missing phase-space coordinates of the stars,
-  and the location of the progenitor object.
-The method generalizes to to encompass mixtures of independent debris structures,
-  or situations in which star membership in structures is unknown.
-[Get more quantitative here?]
 % Conclusions: 
@@
 
-\keywords{
-  cosmology: dark matter
-  ---
-  Galaxy: structure
-  ---
-  Galaxy: halo
-}
+\keywords{}
 
@@
 
-Cosmological simulations of galaxy formation in the $\Lambda$CDM paradigm predict dark matter halos that (1) are permeated with substructure on many scales, (2) are triaxial in shape, and (3) have density profiles and shapes that vary with radius \citep{dubinski91, jing02, kuhlen07, veraciro11}.
-Dark-matter-only simulations produce triaxial halos \citep{jing02} with large density fluctuations \citep{zemp09}.
-Including baryons tends to soften the triaxiality and graininess in the inner galaxy through a combination of dissipative infall \citep{dubinski94} or cooling \citep{bryan13}.
-These processes combined with the gravity from a baryonic disk (ellipsoid) can act to make the inner halo more oblate (spherical), however these processes do not erase the clumpy, triaxial nature of the outer halo \citep[e.g.,][]{pontzen12}.
-This leads to radially-dependent axis ratios and orientation, density profile, and smoothness, thus suggesting complicated, non-intregrable potential forms for Milky Way-like galaxies.
+\section{Scalings}
+The effective potential for the three-body problem for two point masses
 
-The bulk of the baryonic matter in galaxies spans roughly 5-10\% of the extent of the host dark matter halo.
-Hence, the major observable components of a galaxy are sensitive to the inner portion of its host halo's mass distribution.
-For example, the rotation curves of disk galaxies are excellent probes of the inner mass, especially since matter in disks can be assumed to move on near circular orbits.
-Conversely, measuring the dark matter distribution at large radii is complicated by the low density of visible tracers, observational difficulties, and [unclear what assumptions can be made about the integrability and smoothness of the mass distribution].
-Around external galaxies, the extended mass distribution has been probed using a variety of approaches \citep[see][for a a complete and detailed review]{courteau13}.
-For example, the kinematics of tracer populations such as globular clusters or planetary nebulae can be used to derive mass estimates under the assumptions that these satellite systems are relaxed and well-mixed \citep[early investigations include][]{mendez01,cote03}.
-Simple, parameterized models of both the mass and orbit distribution have been simultaneously constrained using such data \citep[e.g.][]{napolitano11,deason12c}.
-%\citep{romanowsky03,douglas07,romanowsky09,napolitano11,lee11}
-Alternatively, the statistical properties of gravitationally lensed background sources around a galaxy can be used to constrain the \emph{projected} shape, orientation, and radial profile of mass \citep[as done by the Lens Structure and Dynamics Survey described in][]{koopmans02}.
-Of course, lensing reconstructions can only be performed for galaxies which closely intersect our line of sight to background sources, but the advent of large photometric catalogues has allowed automatic searches for such chance alignments and significant increases in the number of objects studied in this way \citep[e.g.
-the Sloan Lens ACS Survey, see][]{bolton06}.
-
-Within the Milky Way our unique vantage point allows us a three-dimensional view of the structure in own dark matter halo.
-Our proximity allows us to use individual stars as kinematic tracers and hence build much larger samples that probe deeper into the halo than the globular cluster and planetary nebula studies of external galaxies.
-For example, \cite{deason12a} used halo BHB stars selected from the Sloan Digital Sky Survey \cite[SDSS;][]{york00} as a random tracer population to measure the mass and slope of a simple, power-law fit to the Galactic potential.
-Such studies assume that the tracer orbits are randomly sampled from a smooth distribution function and phase mixed, however the existence of large-scale substructure in stellar streams and clouds is known to bias mass and velocity inferences that ignore this fact.
-Another approach is to exploit the non-random nature of the stellar populations in the halo.
-Tidal streams are dynamically cold systems --- debris typically have small distributions of energy and angular momentum --- and thus require orders of magnitude fewer tracers than a random sample to get constraints of comparable accuracy.
-For example, in the simplest case we might assume that debris stars are actually still on the same orbit as their progenitor system \citep[a \emph{wrong} assumption, see e.g.,][]{eyre11}.
-Then we can imagine measuring the full-space velocities ${\bf v}$ at different points ${\bf x}$ in the structure (e.g., along a stream), which would give us a direct measure of differences in a potential, $\Phi$.
-%In practice, simple methods such as orbit fitting lead to systematic biases in inferred properties of the underlying potential because streams do not strictly delineate orbits [cite eyre binney work].
-
-Since the initial discovery of tidal debris from Sgr, myriad other streams and debris systems have been identified in large photometric surveys, probing Galactocentric distances from $\sim$15-100 kpc.
-The known debris structures are $\gtrsim$10 kpc from the Sun, thus while overdensities in the halo are less contaminated by background, the constituent stars often have incomplete or poorly measured phase-space information.
-Perhaps the most rigorous method of modeling streams and the potential is to run full N-body simulations of satellite disruption and compare the density of observed stars to simulated stars in whatever coordinates are measured.
-A generative N-body model could naturally incorporate observational uncertainties, missing dimensions, and would, in principle, allow for studying any arbitrary time-dependent or non-integrable potential.
-However, such a model is presently computationally intractable.
-\cite{law10} demonstrated the power of this idea by running a grid of N-body models of the disruption of the Sagittarius (Sgr) dwarf spheroidal galaxy and matched their simulated streams to all Sgr data available at that time.
-By comparing the phase-space positions of their stream particles with the data, they were able to constrain the shape of the Milky Way potential out to $\sim$70 kpc and found that the best-fitting halo has a prolate, triaxial shape elongated in the Galactic Z direction, orthogonal to the plane of the disk.
-% (though, see ... for possible complications in the Sgr debris belokurov 2012?).
-
-The difficulty and computational costs associated with proper N-body modeling has motivated numerous efforts for developing approximate methods for constraining properties of the Galactic potential with cold debris structures.
-The simplest approximation is to fit a single orbit to observed debris \citep[e.g.,][]{koposov10, deg13}.
-This has the advantage of naturally incorporating observational errors, but is known to be invalid as debris does not exactly follow an orbit \citep[e.g.,][]{eyre11, sanders13a}.
-To account for this, methods have been proposed that model the debris as an orbit plus some dispersion or offset, either in phase space \citep[e.g.,][]{eyre09a, varghese11, kuepper12} or action-angle coordinates \citep{eyre11, sanders13b, bovy14, sanders14}.
-Other statistical methods have been proposed \citep[][]{penarrubia12, sanderson14} that may prove powerful when applied to, e.g., data from the \gaia\, mission, where full 6D coordinates will be known for large samples of stars in the halo --- and therefore many debris structures --- but stream membership is not known for all stars (however these statistical methods do not account for observational errors).
-
-The potential of the inner galaxy is likely well-described by a (set of) nearly-integrable potentials, thus the dynamics of collisionless systems is plausibly captured by a transformation to some integral-of-motion-space, e.g., action-angle coordinates.
-However there is uncertainty in the exact shape, radial profile, and graininess of the outer halo.
-To test the extent to which the outer halo is well-fit by integrable potentials, we cannot use methods that \emph{assume} integrability.
-In this article, we present a fully probabilistic method for measuring properties of a potential that only relies on numerical orbit integration in ordinary phase-space.
-The model uses individual stars as constraints on the potential, and can therefore handle very small samples of stars.
-Generative models that instead rely on matching densities (in phase-space, action-space, or observed coordinates) must also include fine details about the progenitor (e.g., mass function or structure) and a model for the background of stars.
-Our method also naturally extends to using multiple debris structures \citep[see][who illustrate the power of using multiple streams to simultaneously constrain the potential]{deg14}.
-
-Although the \gaia\, project will deliver hundreds of millions of stars useful for dynamical inference,
-  it still might be that the most relevant stellar samples for inferring the Milky Way potential might be very small in the end.
-For instance, stars that produce good distance estimates might be much more valuable than typical stars in the sample.
-At halo distances, \gaia's parallaxes are not expected to be better than the distance indicators produced by the best
-  variable stars observed at the best wavelengths; for example, near-infrared observations of RR Lyrae stars
-  can in principle deliver $\sim$2\% percent photometric distances at essentially all halo distances \citep{madore12}.
-These valuable stars are rare (and their abundances are age and metallicity dependent);
-  there could be many cold structures in the Milky Way halo that are highly constraining
-  on the potential in principle, but which contain only a few good distance-indicating members.
-That said, it is not yet known what the trade-offs are between having many stars at low precision and a few at high precision,
-  nor is it known how valuable distance information really is, when a structure contains many precisely observed members.
-
-Another reason to think about small samples is more conceptual:
-in dynamical inferences that go beyond simple virial or Jeans arguments,
-  inference depends on the non-linear dynamical evolution of populations in phase space.
-The nonlinearities, combined with heteroskedastic data and requirements to marginalize over phase-space distribution functions,
-  lead naturally to situations in which some stars are \emph{far more informative} than others.
-For a simple example, in a toy study of the Solar System,
-  phase-space distribution marginalization was shown to make some planets are far more informative than others
-  about the gravitational potential \citep{bovy10}.
-This is expected to be a generic feature in dynamical problems:
-  it might be that a very small fraction of \gaia's catalog entries
-  deliver a large fraction of the information about the Milky Way's potential and formation history.
-
-In \citet[][hereafter \paperone]{apw13}, we introduced a simple method (\emph{Rewinder}) for using individual stars associated with tidal debris combined with knowledge about the mass and orbit of the progenitor to constrain properties of the host galaxy potential.
-The method exploits the relationship between the phase-space distribution of debris and (measurable) properties of the progenitor system (e.g., the tidal radius and escape velocity).
-In this paper, we investigate how these relationships scale with the mass and orbit of the progenitor and the quality of the data, then present a new probabilistic framework for simultaneously modeling stream debris and the parent potential based on the same arguments.
-
-In Section~\ref{sec:sims} we describe a suite of N-body simulations that span a range of progenitor masses over two characteristic orbits. In Section~\ref{sec:} we
-
-\section{Simulations}\label{sec:sims}
+\section{Simulations}
 [David Hendel]
@@
 
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=\textwidth]{../../plots/paper2/simulated_streams.pdf}
-\caption{ TODO }\label{fig:sims}
-\end{center}
-\end{figure}
-
-\section{Method}
-
-\subsection{Notation}
-Take $\D$ to be the observed, heliocentric 6D position of a star --- e.g., the measured position on the sky, ($l$, $b$); distance, $D$; proper motions, ($\mu_l$, $\mu_b$); and line-of-sight velocity, $v_r$ --- and take $\W$ to be the corresponding \emph{true}, 6D position of the star in the same coordinate and reference frame.
-Assume that we have determined that this star was once part of a progenitor system (e.g., satellite galaxy) with mass $M_p(t)$ and velocity dispersion $\sigma_p$ tidally disrupting and forming a cold debris structure in the potential, $\Phi$, of some parent galaxy.
-The position of the progenitor is observed to be at heliocentric 6D position $\D_p$, with true position $\W_p$ (where any subscript $p$ refers to the progenitor).
-The true 6D heliocentric spherical positions of the star and progenitor -- $\W$ and $\W_p$ -- have corresponding 6D Galactocentric cartesian coordinate vectors $\X$ and $\X_p$ comprised of 3D position and velocity vectors ($\bs{r}$,$\bs{v}$) and ($\bs{r}_p$,$\bs{v}_p$).
-
-\subsection{Scalings}
-
-In \paperone, we define a set of coordinates relative to the instantaneous phase-space position of the progenitor, $(\bs{r}_{p},\bs{v}_{p})$, normalized by the instantaneous tidal radius and escape velocity.
-Here we (1) instead scale the velocity coordinates by the velocity dispersion of the satellite, (2) reduce the dimensionality by considering only the magnitude of these vectors, and (3) take the logarithm of the magnitude,
-\begin{equation}
-  R = \ln\left(\frac{\left\vert\bs{r} - \bs{r}_{p}\right\vert}{R_{\rm tide}}\right)\,\,\,\,,\,\,\,\,
-  V = \ln\left(\frac{\left\vert\bs{v} - \bs{v}_{p}\right\vert}{\sigma_v}\right).
-\end{equation}
-Figure~\ref{fig:R_V} shows distributions of $R$ and $V$ --- computed at $t=t_{\rm ub}$ for each particle --- as a function of progenitor mass from the simulations described in Section~\ref{sec:sims}.
-In these scaled, log coordinates we model distributions in $R$ and $V$ as Gaussians.
-We find that for masses $\lesssim10^9M_\odot$, the variance ($\sigma^2$ in plot panels) remains approximately constant and the mean stays fixed at 0.
-For the largest mass simulation ($2.5\times10^9M_\odot$), the perturbation to the total potential due to the progenitor halo becomes significant and thus a large fraction of the stars are disrupted through interactions with the subhalo.
-For this article, we ignore covariances between $R$ and $V$.
-The
-
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=\textwidth]{../../plots/paper2/R_V.pdf}
-\caption{ TODO }\label{fig:R_V}
-\end{center}
-\end{figure}
-
-[Cold vs. Hot streams in the context of observational errors]
-
-\subsection{Probabilistic model}
-With this notation, we write the joint posterior probability for the observed position of a star, progenitor, and parent potential parameters as:
-\begin{align}
-	p(\Phi, \W, \W_p, \tau \given \D, \D_p) &= \frac{1}{\mathcal{Z}} p(\D \given \W) p(\D_p \given \W_p) 
-												       p(\W \given \W_p, \tau, \Phi) 
-												       p(\Phi) p(\tau)
-\end{align}
-where the factor $\mathcal{Z}$ only depends on the properties of the data (the \emph{evidence}), and is thus constant when varying the model parameters.
-We assume that the observational errors are Gaussian in heliocentric coordinates:
-\begin{align}
-	p(\D \given \W) &= \mathcal{N}(\W \given \D, \bSigma)\\
-	p(\D_p \given \W_p) &= \mathcal{N}(\W_p \given \D_p, \bSigma_p)
-\end{align}
-where the covariance matrices $\bSigma$ and $\bSigma_p$ specify the observational uncertainties on the observed 6D position of star and the progenitor, respectively.
-
-We model the present-day position and velocity of a disrupted star as the result of sampling from isotropic, log-normal distributions in relative distance, $R=\left\vert \bs{r}-\bs{r}_p \right\vert$, and relative velocity, $V=\left\vert \bs{v}-\bs{v}_p \right\vert$, at the time the star comes unbound from the progenitor, evolved to present day in the parent potential, $\Phi$.
-The distributions over $\ln R$ and $\ln V$ are centered on the logarithm of the instantaneous tidal radius, $\rtide$, and velocity dispersion, $\sigma_v$, of the progenitor at time $\tau$: $\mathcal{N}(\ln R \given \ln\rtide)\vert_{t=\tau}$ and $\mathcal{N}(\ln V \given \ln \sigma_v)\vert_{t=\tau}$.
-In this model, we neglect the effect of self-interaction with the halo of the progenitor.
-
-The likelihood of a star's position today is computed by first transforming the observed position to Galactocentric cartesian coordinates and integrating the orbits of the progenitor and star backwards for several Gigayears in the parent potential, $\Phi$, treating both as test particles.
-The distance, $R$, and velocity, $V$, of the star relative to the progenitor are computed at each time-step, and the likelihood is computed as the product of the isotropic, log-normal centered on the logarithm of the instantaneous tidal radius with the isotropic, log-normal centered on the logarithm of the velocity dispersion of the progenitor:
-\begin{align}
-	p(\W \given \W_p, \tau, \Phi) &= p(\X \given \X_p, \tau, \Phi) \jac{(l,b,D,\mu_l,\mu_b,v_r)}_{t=\tau}\\
-	p(\X \given \X_p, \tau, \Phi) &= p(\ln R, \ln V \given \tau, \Phi) \jac{(x,y,z,v_x,v_y,v_z)}_{t=\tau}\label{eq:int} \\
-	p(\ln R, \ln V \given \tau, \Phi) &= \left[\mathcal{N}(\ln R | \ln \rtide, \sigma_R)\mathcal{N}(\ln V | \ln \sigma_v, \sigma_V)\right]_{t=\tau}
-\end{align}
-% , but the vectors $\D$ and $\W$ [exist] in different coordinate systems --- heliocentric spherical and Galactocentric cartesian, respectively --- so we must also track the Jacobian that [describes] this transformation:
-where in eq.~\ref{eq:int} we have implicitly integrated over and neglected the two angle coordinates associated with $\ln R$ and the two angles associated with $\ln V$. 
-\begin{align}
-	\ln p(\ln R, \ln V \given \tau, \Phi) = -\ln(2\pi)-\ln(\sigma_R\sigma_V)-
-		\ln\left[\left(\frac{R}{\rtide}\right)^{1/\sigma_R} \left(\frac{V}{\sigma_v}\right)^{1/\sigma_V}\right]
-\end{align}
-We then compute the marginal likelihood, $p(\W \given \W_p, \Phi)$, by integrating over all possible unbinding times,
-\begin{align}
-	p(\W \given \W_p, \Phi) = \int p(\W \given \W_p, \tau, \Phi)p(\tau) d\tau.
-\end{align}
-In general, $p(\tau)$ is likely conditional on the orbit of the progenitor --- stars are preferentially stripped at pericentric passages --- but this is neglected and is taken to be uniform over the entire interaction history between the progenitor and parent.
-
-\section{Experiments}
-Initially we assume the present-day mass of the progenitor, $M_p$, and velocity dispersion, $\sigma_v$, have been measured perfectly and neglect mass-loss from the progenitor.
-
-For all tests in this article, we model the potential of the parent galaxy as a three-component sum of a Miyamoto-Nagai disk \citep{}, Hernquist bulge \citep[spheroid;][]{}, and a triaxial, logarithmic halo \citep[e.g.,][]{law10}:
-\begin{align}
-	&\Phi_{\rm disk}(R,z) = -\frac{GM_{\rm disk}}{\sqrt{R^2 + (a + \sqrt{z^2 + b^2})^2}}\\
-	&\Phi_{\rm spher}(r) = -\frac{GM_{\rm spher}}{r + c}\\
-	&\Phi_{halo}(x,y,z) = v_{h}^2 \ln(C_1 x^2 + C_2 y^2 + C_3 xy + (z/q_z)^2 + r_h^2)
-\end{align}
-where $C_1$, $C_2$, and $C_3$ are combinations of the $x$ and $y$ axis
-ratios ($q_1$, $q_2$) and orientation of the halo with respect to the
-baryonic disk ($\phi$):
-\begin{align}
-  C_1 &= \frac{\cos^2\phi}{q_1^2} + \frac{\sin^2\phi}{q_2^2}\\
-  C_2 &= \frac{\sin^2\phi}{q_1^2} + \frac{\cos^2\phi}{q_2^2}\\
-  C_3 &= 2\sin\phi\cos\phi \left(q_1^{-2} - q_2^{-2}\right).
-\end{align}
-
-\begin{figure}[h]
-\begin{center}
-% \includegraphics[width=\textwidth]{/Users/adrian/projects/streams/plots/paper2/potentials.pdf}
-\caption{ TODO }\label{fig:potential}
-\end{center}
-\end{figure}
-
-When recovering the potential, we hold fixed the disk and spheroid parameters (see Table~\ref{tbl:potential}), along with two halo parameters: the scale radius, $r_h$, and one of the flattening components in the x-y plane, $q_2$.
-These halo parameters are degenerate with combinations of the other parameters --- e.g., setting $q_1=2.0$, $q_2=1.0$, $\phi=0^\circ$ is equivalent to setting $q_1=1.0$, $q_2=2.0$, $\phi=90^\circ$.
-The priors on the remaining halo parameters are taken to be uniform over a conservative domain of realistic values: for $v_h$, 100-200~kms$^{-1}$ corresponds to a range in solar circular velocities from $\sim$210-250 kms$^{-1}$ (holding other parameters fixed); the range in axis ratios allow for prolate, oblate, and mixed triaxiality; and $\phi$ is restricted to $\pm45^\circ$ around the value measured by LM10 ($\phi = 97^\circ$).
-Figure~\ref{fig:potential} shows equipotential slices in the Galactic X-Z plane (Y=0) and Y-Z plane (X=0) for a few choices of  $q_1$,  $q_z$, and $\phi$ while holding all other parameters fixed, as described in the figure caption.
-
-\begin{table}[h]
-\begin{center}
-	\begin{tabular}{l c c c} \toprule
-		{\bf Component} & {\bf Parameter} & {\bf Value} & {\bf Prior} \\\toprule
-		disk & $M_{\rm disk}$ & $1.0\times10^{11}M_\odot$ & -- \\ 
-		& $a$ & 6.5 kpc & --\\
-		& $b$ & 0.26 kpc & --\\
-		\midrule
-		spheroid & $M_{\rm spher}$ & $3.4\times10^{10}M_\odot$ & --\\ 
-		& $c$ & 0.7 kpc & --\\
-		\midrule
-		halo & $v_h$ & -- & $\mathcal{U}(100,200)$ km $\mathrm{s}^{-1}$ \\
-		& $q_1$ & -- & $\mathcal{U}(1,2)$\\
-		& $q_2$ & 1.0 & --\\
-		& $q_z$ & -- & $\mathcal{U}(1,2)$\\
-		& $\phi$ & -- & $\mathcal{U}(52,142)$ deg\\
-		& $r_h$ & 12 kpc & --\\
-		\bottomrule
-		\hline
-	\end{tabular}
-	\caption{Parameter values used ...\label{tbl:potential}}
-\end{center}
-\end{table}
-
-\subsection{\gaia + \spitzer}
-4 stars in Sgr-like stream
-
-\subsection{\spitzer + current data}
-4 stars in Sgr-like stream
-
-\subsection{\spitzer + current data, no progenitor}
-4 stars in Orphan-like stream, XX stars in GD-1 -like stream
-
-\section{Discussion}
-
-[Cold vs. Hot streams]
-
-[bow-tie structure, Kuepper work, and so on]
-
-[What's unrealistic about what we have done?  The kinematic model of
-  the progenitor is exceedingly simplistic; why are we okay with it
-  nonetheless but what could we do to improve it?  Related to this,
-  the progenitor doesn't shrink with time (as it should).  What's up
-  with that?]
-
-[We did not assume the potential is static in time.  At least not in
-  general; nothing in the method requires or suggests this.]
-
-[Issues about computation and scaling.]
-
-[Nothing special about streams; will work on any structures, including
-  shells and the like.]
-
-[Don't eschew low-dimensional constraints on the potential!]
-
-[Another prediction of hierarchical structure formation in $\Lambda$CDM is the presence of substructure. In principle, this method would work on any potential that we can integrate orbits in.]
-
-\bibliographystyle{apj}
-\bibliography{refs}
-
-\acknowledgements
-APW: NSF GRFP, Sigma Xi
-
 \end{document}

### 2014-03-05 streams/inference/back_integrate.py
sweet model in cartesian

@@ -115,7 +115,7 @@ def back_integration_likelihood(t1, t2, dt, potential, p_hel, s_hel, logm0, logm
 
     #r_tide = potential._tidal_radius(s_mass, s_orbit)
     r_tide = potential._tidal_radius(m_t, s_orbit[...,:3])
-    v_disp = s_V * r_tide / s_R / 2.2
+    v_disp = s_V * r_tide / s_R
 
     # instantaneous cartesian basis to project into
     x_hat = s_orbit[...,:3] / np.sqrt(np.sum(s_orbit[...,:3]**2, axis=-1))[...,np.newaxis]
@@ -129,25 +129,63 @@ def back_integration_likelihood(t1, t2, dt, potential, p_hel, s_hel, logm0, logm
     rel_vel = rel_orbits[...,3:]
 
     # project onto each
-    X = np.sum(rel_pos * x_hat, axis=-1)
-    Y = np.sum(rel_pos * y_hat, axis=-1)
-    Z = np.sum(rel_pos * z_hat, axis=-1)
+    X = np.sum(rel_pos * x_hat, axis=-1) / r_tide
+    Y = np.sum(rel_pos * y_hat, axis=-1) / r_tide
+    Z = np.sum(rel_pos * z_hat, axis=-1) / r_tide
 
-    VX = np.sum(rel_vel * x_hat, axis=-1)
-    VY = np.sum(rel_vel * y_hat, axis=-1)
-    VZ = np.sum(rel_vel * z_hat, axis=-1)
+    VX = np.sum(rel_vel * x_hat, axis=-1) / v_disp
+    VY = np.sum(rel_vel * y_hat, axis=-1) / v_disp
+    VZ = np.sum(rel_vel * z_hat, axis=-1) / v_disp
 
     # position likelihood is gaussian at lagrange points
-    var_x = (np.median(r_tide)/5.)**2
-    r_term = -0.5*((np.log(var_x) + (X - tail_bit*r_tide)**2/var_x) + \
-                   (np.log(3*var_x) + (Y)**2/(3*var_x)) + \
-                   (np.log(var_x) + (Z)**2/var_x))
+    var_x = var_z = 0.1
+    var_y = 0.1
+    r_term = -0.5*((np.log(var_x) + (X-tail_bit)**2/var_x) + \
+                   (np.log(var_y) + (Y)**2/var_y) + \
+                   (np.log(var_z) + (Z)**2/var_z))
 
[…]

### 2014-03-05 text/paper2/ms.tex
new paper and minor tweaks to paper2

@@
 Assume that we have determined that this star was once part of a progenitor system (e.g., satellite galaxy) with mass $M_p(t)$ and velocity dispersion $\sigma_p$ tidally disrupting and forming a cold debris structure in the potential, $\Phi$, of some parent galaxy.
-The position of the progenitor is observed to be at heliocentric 6D position $\D_p$, with true position $\W_p$ (where any subscript $p$ refers to the progenitor).
+The present position of the progenitor is observed to be at heliocentric 6D position $\D_p$, with true position $\W_p$ (where any subscript $p$ refers to the progenitor).
 The true 6D heliocentric spherical positions of the star and progenitor -- $\W$ and $\W_p$ -- have corresponding 6D Galactocentric cartesian coordinate vectors $\X$ and $\X_p$ comprised of 3D position and velocity vectors ($\bs{r}$,$\bs{v}$) and ($\bs{r}_p$,$\bs{v}_p$).

### 2014-03-05 scripts/test_infer_potential.py
test case

@@ -38,9 +38,12 @@ logging.basicConfig(level=logging.DEBUG)
 minimum_config = """
 name: test
 data_file: data/observed_particles/2.5e8_N1024.hdf5
+# nparticles: 7
+# particle_idx: [758, 635, 503, 122, 961, 713, 360]
 nparticles: 8
-particle_idx: [62, 151, 247, 47, 9, 215, 137, 200]
+particle_idx: [1019, 606, 740, 301, 261, 39, 735, 448]
 # nparticles: 16
+# nparticles: 128
 
 potential:
     class_name: LawMajewski2010
@@ -63,20 +66,20 @@ ptc_params = """
 #    parameters: [d, mul, mub, vr]
 
 sat_params = """
-    parameters: [logmass, d, mul, mub, vr]
+    parameters: [logmass, logmdot, d, mul, mub, vr]
 """
-#    parameters: [logmass]
+#    parameters: [logmass, logmdot, d, mul, mub, vr]
 
 lm10_c = minimum_config.format(potential_params=pot_params,
                                particles_params="",
                                satellite_params="")
 
-_config = minimum_config.format(potential_params=pot_params,
-                                particles_params=ptc_params,
-                                satellite_params=sat_params)
-# _config = minimum_config.format(potential_params="",
+# _config = minimum_config.format(potential_params=pot_params,
 #                                 particles_params=ptc_params,
-#                                 satellite_params="")
+#                                 satellite_params=sat_params)
+_config = minimum_config.format(potential_params=pot_params,
+                                particles_params="",
+                                satellite_params=sat_params)
 
 # particles_params=ptc_params,
 # satellite_params=sat_params
@@ -142,73 +145,6 @@ class TestStreamModel(object):
         assert true_ln_p > ln_p
         assert true_ln_p == true_ln_p2
         assert ln_p == ln_p2
-
-    def test_likelihood(self):
-
-        simulation = SgrSimulation("2.5e8")
-        all_gc_particles = simulation.particles(N=1000, expr="tub!=0")\
-                                     .to_frame(galactocentric)
-
-        c = io.read_config(lm10_c)
-        model = si.StreamModel.from_config(c)
-        true_potential = model._potential_class(**model._given_potential_params)
-
-        test_path = os.path.join(output_path, "likelihood")
-        if not os.path.exists(test_path):
-            os.mkdir(test_path)
-
[…]

### 2014-03-05 text/paper2/ms.tex
working on method section

@@
 
-\subsection{Scalings}
+\subsection{Tidal debris}
 
-In \paperone, we define a set of coordinates relative to the instantaneous phase-space position of the progenitor, $(\bs{r}_{p},\bs{v}_{p})$, normalized by the instantaneous tidal radius and escape velocity.
-Here we (1) instead scale the velocity coordinates by the velocity dispersion of the satellite, (2) reduce the dimensionality by considering only the magnitude of these vectors, and (3) take the logarithm of the magnitude,
+For a satellite galaxy on a mildly eccentric orbit, stars are disrupted via tidal ``pruning'' due to the tidal field of the parent system or tidal shocking due to the rapidly changing tidal field as the progenitor moves through pericenter.
+The interplay between these two processes as a function of orbital properties will be studied in forthcoming work, and thus we focus on the slow disruption of stars due to steady tidal forcing.
+
+Consider a point-mass satellite of mass $m$ on a circular orbit with frequency $\Omega$ around a point-mass of mass $M\gg m$ \citep[the restricted three-body problem, e.g.,][]{binneytremaine}.
+In a frame rotating with the orbital frequency of the satellite, the effective potential around the satellite has two stationary, unstable points located around $\sim R_p \pm r_J$, where $R_p$ is the distance to the satellite and $r_J$ is the Jacobi or tidal radius,
 \begin{equation}
-  \ln R = \ln\left(\frac{\left\vert\bs{r} - \bs{r}_{p}\right\vert}{R_{\rm tide}}\right)\,\,\,\,,\,\,\,\,
-  \ln V = \ln\left(\frac{\left\vert\bs{v} - \bs{v}_{p}\right\vert}{\sigma_v}\right).
+	r_J = \left(\frac{m}{3M}\right)^{1/3}R_p.
 \end{equation}
-Figure~\ref{fig:R_V} shows distributions of $\ln R$ and $\ln V$ --- computed at $t=t_{\rm ub}$ for each particle --- as a function of progenitor mass from the simulations described in Section~\ref{sec:sims}.
-In these scaled coordinates we model distributions in $R$ and $V$ as log-normal and ignore covariances between $R$ and $V$.
-We find that for masses $\lesssim10^9M_\odot$, the variance ($\sigma^2$ in plot panels) remains approximately constant and the mean stays fixed at 0.
-For the largest mass simulation ($2.5\times10^9M_\odot$), the perturbation to the total potential due to the progenitor halo becomes significant and thus a large fraction of the stars are disrupted through interactions with the subhalo.
+Particles that would be bound to the satellite system in isolation may have enough (Jacobi) energy to overcome the effective potential barrier at these Lagrange points and thus will be preferentially stripped from the satellite at these points (Fig.~8.6 of BT).
+
+For a spherical, extended mass distribution [...] \label{eq:tidalradius}
+
+[Describe velocity dispersion...] \label{eq:velscale}
+
+In a triaxial potential (e.g., Equation~\ref{eq:lm10}), the orbital plane of the satellite is not fixed but we still expect there to be instantaneous Lagrange points along the line of centers connecting the origin of the parent potential to the satellite.
+We thus proceed by using coordinates relative to the position and velocity of the progenitor, ($\bs{r}_p$,$\bs{v}_p$), rotated into the instantaneous orbital plane, and normalized by the instantaneous tidal radius for positions (Eq.~\ref{eq:tidal radius}) and instantaneous velocity scale for velocity (Eq.~\ref{eq:velscale}).
+% where $\hat{\bs r}_p = {\bs r}_p / |{\bs r}_p|$ and $\hat{\bs v}_p = {\bs v}_p / |{\bs v}_p|$
+The (time-dependent) basis vectors are given by
+\begin{align}
+	\hat{x}_1 &= \bs{r}_p / \|\bs{r}_p\|\\
+	\hat{x}_2 &= -\hat{x}_1 \times \hat{x}_3\\
+	\hat{x}_3 &= (\bs{r}_p \times \bs{v}_p) / \|\bs{r}_p \times \bs{v}_p\|.
+\end{align}
+Figure~\ref{fig:lpts_r} shows sections of particle orbits in this coordinate system, symmetric around the unbinding time, $t_{\rm ub}$, (defined above), as a function of progenitor mass for each N-body simulation described in Section~\ref{sec:}.
+In these coordinates, the classical Lagrange points would be located around $\pm1$.
+Figure~\ref{fig:lpts_v} shows the same for the velocity of each star.
+The location of and dispersion about the effective Lagrange points in these scaled coordinates remain remarkably consistent across the range of progenitor masses explored.
+Furthermore, the dispersion in velocity is well-normalized by [...].
+Even in [...], the phase-space dispersion of tidal debris scales with the mass ratio $(m_{\rm sat} / M_{\rm enc})^{1/3}$.
 
@@
 \begin{center}
-\includegraphics[width=\textwidth]{../../plots/paper2/R_V.pdf}
-\caption{ TODO }\label{fig:R_V}
+\includegraphics[width=\textwidth]{../../plots/paper2/Lpts_r.png}
+\caption{ TODO }\label{fig:lpts_r}
 \end{center}
@@
 
-The scale-radius and velocity dispersion scale with the mass of the progenitor. [...observational errors] [ $D_{\rm ps}$]
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=\textwidth]{../../plots/paper2/Lpts_v.png}
+\caption{ TODO }\label{fig:lpts_v}
+\end{center}
+\end{figure}
+
 [Cold vs. Hot streams in the context of observational errors]
@@
 	&\Phi_{\rm spher}(r) = -\frac{GM_{\rm spher}}{r + c}\\
-	&\Phi_{halo}(x,y,z) = v_{h}^2 \ln(C_1 x^2 + C_2 y^2 + C_3 xy + (z/q_z)^2 + r_h^2)
+	&\Phi_{\rm halo}(x,y,z) = v_{h}^2 \ln(C_1 x^2 + C_2 y^2 + C_3 xy + (z/q_z)^2 + r_h^2)
 \end{align}
@@
 \end{align}
+The total potential then is just
+\begin{equation}
+	\Phi_{\rm tot} = \Phi_{\rm disk} + \Phi_{\rm spher} + \Phi_{\rm halo}\label{eq:lm10}.
+\end{equation}
 

### 2014-03-06 text/paper2/ms.tex
rewrote method section

@@
 \newcommand{\inverse}[1]{{#1}^{-1}}
-\newcommand{\jac}[1]{\left\vert \J{#1} \right\vert}
 
@@
 \newcommand{\bs}{\boldsymbol}
+\newcommand{\sat}{{\rm p}}
+\newcommand{\tub}{t_{\rm ub}}
+\newcommand{\rr}{\widetilde{\bs{r}}}
+\newcommand{\vv}{\widetilde{\bs{v}}}
 
@@
 
-\subsection{Notation}
-Take $\D$ to be the observed, heliocentric 6D position of a star --- e.g., the measured position on the sky, ($l$, $b$); distance, $D$; proper motions, ($\mu_l$, $\mu_b$); and line-of-sight velocity, $v_r$ --- and take $\W$ to be the corresponding \emph{true}, 6D position of the star in the same coordinate and reference frame.
-Assume that we have determined that this star was once part of a progenitor system (e.g., satellite galaxy) with mass $M_p(t)$ and velocity dispersion $\sigma_p$ tidally disrupting and forming a cold debris structure in the potential, $\Phi$, of some parent galaxy.
-The present position of the progenitor is observed to be at heliocentric 6D position $\D_p$, with true position $\W_p$ (where any subscript $p$ refers to the progenitor).
-The true 6D heliocentric spherical positions of the star and progenitor -- $\W$ and $\W_p$ -- have corresponding 6D Galactocentric cartesian coordinate vectors $\X$ and $\X_p$ comprised of 3D position and velocity vectors ($\bs{r}$,$\bs{v}$) and ($\bs{r}_p$,$\bs{v}_p$).
+\subsection{Tidal debris}\label{sec:debris}
 
-\subsection{Tidal debris}
-
-For a satellite galaxy on a mildly eccentric orbit, stars are disrupted via tidal ``pruning'' due to the tidal field of the parent system or tidal shocking due to the rapidly changing tidal field as the progenitor moves through pericenter.
-The interplay between these two processes as a function of orbital properties will be studied in forthcoming work, and thus we focus on the slow disruption of stars due to steady tidal forcing.
+For a satellite galaxy on a mildly eccentric orbit, mass-loss is driven via tidal stripping due to the tidal field of the parent system or tidal shocking due to the rapidly changing tidal field as the progenitor system moves through pericenter.
+The interplay between these two processes as a function of orbital properties will be studied in forthcoming work and focus below on the slow disruption of stars due to steady tidal forcing.
 
 Consider a point-mass satellite of mass $m$ on a circular orbit with frequency $\Omega$ around a point-mass of mass $M\gg m$ \citep[the restricted three-body problem, e.g.,][]{binneytremaine}.
-In a frame rotating with the orbital frequency of the satellite, the effective potential around the satellite has two stationary, unstable points located around $\sim R_p \pm r_J$, where $R_p$ is the distance to the satellite and $r_J$ is the Jacobi or tidal radius,
+In a frame rotating with the orbital frequency of the satellite, the effective potential around the satellite has (amongst others) two stationary, unstable points located around $\sim R \pm r_J$, where $R$ is the distance to the satellite and $r_J$ is the Jacobi or tidal radius,
 \begin{equation}
-	r_J = \left(\frac{m}{3M}\right)^{1/3}R_p.
+	r_J \sim R\left(\frac{m}{3M}\right)^{1/3}.\label{eq:ptmass}
 \end{equation}
 Particles that would be bound to the satellite system in isolation may have enough (Jacobi) energy to overcome the effective potential barrier at these Lagrange points and thus will be preferentially stripped from the satellite at these points (Fig.~8.6 of BT).
+For a spherical, extended parent mass distribution the tidal radius instead scales with the enclosed mass of the parent system at a given orbital position, $R$, with additional terms that account for the local slope of the density profile.
+Thus, in a more general parent potential the effective Lagrange points may not be symmetric about the center of the satellite and may deviate from the point-mass tidal radius (Eq.~\ref{eq:ptmass}) by factors of order unity.
+We thus simply adopt
+\begin{equation}
+	r_{\rm tide} = R\left(\frac{m_{\rm sat}}{3M_{\rm enc}}\right)^{1/3}\label{eq:tidalradius}
+\end{equation}
+as a normalizing distance scale.
 
-For a spherical, extended mass distribution [...] \label{eq:tidalradius}
+At pericenter, the velocity of the satellite will be of order $V\sim \sqrt{GM_{\rm enc}/R}$. If the velocity dispersion of the satellite is $\sigma_v \sim \sqrt{Gm/r_J}$, then
+\begin{equation}
+	\sigma_v \sim V\left(\frac{m_{\rm sat}}{M_{\rm enc}}\right)^{1/3}\label{eq:velscale}.
+\end{equation}
+Thus, we expect the particle velocities to scale with $(m/M)^{1/3}$.
 
-[Describe velocity dispersion...] \label{eq:velscale}
-
-In a triaxial potential (e.g., Equation~\ref{eq:lm10}), the orbital plane of the satellite is not fixed but we still expect there to be instantaneous Lagrange points along the line of centers connecting the origin of the parent potential to the satellite.
-We thus proceed by using coordinates relative to the position and velocity of the progenitor, ($\bs{r}_p$,$\bs{v}_p$), rotated into the instantaneous orbital plane, and normalized by the instantaneous tidal radius for positions (Eq.~\ref{eq:tidal radius}) and instantaneous velocity scale for velocity (Eq.~\ref{eq:velscale}).
+In a triaxial potential (e.g., Equation~\ref{eq:lm10}), the orbital plane of the satellite is not fixed but we still expect there to be effective Lagrange points along the line of centers connecting the origin of the parent potential to the satellite.
+We thus proceed by defining a coordinate system relative to the position and velocity of the progenitor, ($\bs{r}_p$,$\bs{v}_p$), rotated into the instantaneous orbital plane defined by $\bs{r}_p \times \bs{v}_p$, and normalized by the instantaneous tidal radius for positions (Eq.~\ref{eq:tidalradius}) and instantaneous velocity scale for velocity (Eq.~\ref{eq:velscale}).
 % where $\hat{\bs r}_p = {\bs r}_p / |{\bs r}_p|$ and $\hat{\bs v}_p = {\bs v}_p / |{\bs v}_p|$
@@
 \begin{align}
-	\hat{x}_1 &= \bs{r}_p / \|\bs{r}_p\|\\
+	\hat{x}_1 &= \frac{\bs{r}_p}{\|\bs{r}_p\|}\\
 	\hat{x}_2 &= -\hat{x}_1 \times \hat{x}_3\\
-	\hat{x}_3 &= (\bs{r}_p \times \bs{v}_p) / \|\bs{r}_p \times \bs{v}_p\|.
+	\hat{x}_3 &= \frac{\bs{r}_p \times \bs{v}_p}{\|\bs{r}_p \times \bs{v}_p\|}.
 \end{align}
-Figure~\ref{fig:lpts_r} shows sections of particle orbits in this coordinate system, symmetric around the unbinding time, $t_{\rm ub}$, (defined above), as a function of progenitor mass for each N-body simulation described in Section~\ref{sec:}.
-In these coordinates, the classical Lagrange points would be located around $\pm1$.
+Figure~\ref{fig:lpts_r} shows sections of particle orbits in this coordinate system, symmetric around the unbinding time, $\tub$ (Section~\ref{sec:sims}), as a function of progenitor mass.
+In these coordinates, the classical Lagrange points would be located around $x_1=\pm1,x_2=0,x_3=0$ (illustrated by the unit circle in top panels of Figure~\ref{fig:lpts_r}).
 Figure~\ref{fig:lpts_v} shows the same for the velocity of each star.
-The location of and dispersion about the effective Lagrange points in these scaled coordinates remain remarkably consistent across the range of progenitor masses explored.
-Furthermore, the dispersion in velocity is well-normalized by [...].
-Even in [...], the phase-space dispersion of tidal debris scales with the mass ratio $(m_{\rm sat} / M_{\rm enc})^{1/3}$.
+Though not centered on the point-mass Lagrange points, the location of and dispersion about the effective Lagrange points in these scaled coordinates remain remarkably consistent across the range of progenitor masses explored at $x_1\sim\pm2$.
+Furthermore, the dispersion in velocity is well-normalized by the velocity scale of Eq.~\ref{eq:velscale}, though it is clear that the velocity dispersion along $\hat{x}_1$, the radial vector to the satellite position, is larger than that in other dimensions.
+The phase-space dispersion of tidally stripped debris scales with the mass ratio $(m_{\rm sat} / M_{\rm enc})^{1/3}$, the orbital radius, and the orbital velocity of the progenitor system.
 
@@
 \includegraphics[width=\textwidth]{../../plots/paper2/Lpts_r.png}
-\caption{ TODO }\label{fig:lpts_r}
+\caption{ Orbits of 2000 particles around $t=\tub$.
+The orbits were integrated backwards from their present-day positions (final time step of the N-body simulations) without the potential of the progenitor.
+TODO: coordinates...}\label{fig:lpts_r}
 \end{center}
@@
 \includegraphics[width=\textwidth]{../../plots/paper2/Lpts_v.png}
-\caption{ TODO }\label{fig:lpts_v}
+\caption{ Same as Figure~\ref{fig:lpts_r} but for particle velocities. }\label{fig:lpts_v}
 \end{center}
@@
 
-[Cold vs. Hot streams in the context of observational errors]
-
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=\textwidth]{../../plots/paper2/num_recombine.png}
-\caption{ TODO }\label{fig:R_V}
-\end{center}
-\end{figure}
+%\begin{figure}[h]
+%\begin{center}
+%\includegraphics[width=\textwidth]{../../plots/paper2/num_recombine.png}
+%\caption{ TODO }\label{fig:R_V}
+%\end{center}
+%\end{figure}
 
 \subsection{Probabilistic model}
-With this notation, we write the joint posterior probability for the observed position of a star, progenitor, and parent potential parameters as:
-\begin{align}
-	p(\Phi, \W, \W_p, \tau \given \D, \D_p) &= \frac{1}{\mathcal{Z}} p(\D \given \W) p(\D_p \given \W_p) 
-												       p(\W \given \W_p, \tau, \Phi) 
-												       p(\Phi) p(\tau)
-\end{align}
-where the factor $\mathcal{Z}$ only depends on the properties of the data (the \emph{evidence}), and is thus constant when varying the model parameters.
-We assume that the observational errors are Gaussian in heliocentric coordinates:
+[ intro, based on conclusion from previous section, etc.]
+
+We observe the position of a star, $\D$, in heliocentric coordinates --- e.g., the measured position on the sky, ($l$, $b$); distance, $d$; proper motions, ($\mu_l$, $\mu_b$); and line-of-sight velocity, $v_r$ --- and have determined through some other means that this star was once part of a progenitor system (e.g., satellite galaxy) with mass $m(t)$ that is disrupting and forming a cold debris structure in the potential, $\Phi$, of the parent galaxy (e.g., the Milky Way).
+We assume that the mass of the parent system enclosed in the orbit of the satellite, $M_{\rm enc}$, is much larger than the initial mass of the satellite, $M_{\rm enc} \gg m(t=0)$.
+The present position of the progenitor is observed to be at heliocentric position $\D_\sat$ (where any subscript $\sat$ refers to the progenitor).
+In general, the data for the star and progenitor will have missing dimensions and significant uncertainties at distances typical to the Galactic halo, thus we define $\W$ and $\W_\sat$ as the true 6D, heliocentric positions of the star and progenitor and will include these in our model.
+By including the 6D position of the progenitor in the model, this model will work even when the progenitor location is unknown as in, e.g., the Orphan stream \citep{some orphan paper}.
+
+We assume that the observational uncertainties are Gaussian in heliocentric, spherical coordinates, such that
 \begin{align}
 	p(\D \given \W) &= \mathcal{N}(\W \given \D, \bSigma)\\
-	p(\D_p \given \W_p) &= \mathcal{N}(\W_p \given \D_p, \bSigma_p)
+	p(\D_\sat \given \W_\sat) &= \mathcal{N}(\W_\sat \given \D_\sat, \bSigma_\sat)
 \end{align}
-where the covariance matrices $\bSigma$ and $\bSigma_p$ specify the observational uncertainties on the observed 6D position of star and the progenitor, respectively.
+where $\mathcal{N}$ represents the normal distribution, and the covariance matrices $\bSigma$ and $\bSigma_\sat$ specify the observational uncertainties on the observed 6D position of the star and progenitor, respectively.
 
-We model the present-day position and velocity of a disrupted star as the result of sampling from isotropic, log-normal distributions in relative distance, $R=\left\vert \bs{r}-\bs{r}_p \right\vert$, and relative velocity, $V=\left\vert \bs{v}-\bs{v}_p \right\vert$, at the time the star comes unbound from the progenitor, evolved to present day in the parent potential, $\Phi$.
-The distributions over $\ln R$ and $\ln V$ are centered on the logarithm of the instantaneous tidal radius, $\rtide$, and velocity dispersion, $\sigma_v$, of the progenitor at time $\tau$: $\mathcal{N}(\ln R \given \ln\rtide)\vert_{t=\tau}$ and $\mathcal{N}(\ln V \given \ln \sigma_v)\vert_{t=\tau}$.
-In this model, we neglect the effect of self-interaction with the halo of the progenitor.
+To model the true, 6D, present-day position of the star, $\W$, we first transform to cartesian, Galactocentric coordinates where ($\bs{r},\bs{v}$) and ($\bs{r}_\sat,\bs{v}_\sat$) are the position and velocity of the star and progenitor, respectively.
+In cartesian coordinates, the present-day position of the star is modeled as having been sampled at $t=\tub$ from an isotropic Gaussian centered on one of the instantaneous, effective Lagrange points in position, and an isotropic Gaussian centered on the origin in velocity, and evolved to present day in the parent potential, $\Phi$.
+That is, we assume that once the star becomes ``unbound'' from the satellite the potential of the satellite can be ignored, and thus treat the star as a test particle.
+\begin{align}
+	p(\W \given \W_\sat, \tub, \Phi, K, c) &= p(\bs{r} \given \bs{r}_\sat, \tub, \Phi, K, c)p(\bs{v} \given \bs{v}_\sat, \tub, \Phi) \left\vert\J\right\vert_{t=\tub}\\
+	p(\bs{r} \given \bs{r}_\sat, \tub, \Phi, K, c) &= \mathcal{N}(\bs{r} \given \bs{r}_\sat - cK\rtide \hat{x}_1, \sigma_1)_{t=\tub}\\
+	p(\bs{v} \given \bs{v}_\sat, \tub, \Phi) &= \mathcal{N}(\bs{v} \given \bs{v}_\sat, \sigma_v)_{t=\tub}
+\end{align}
+where $\rtide$ is given by Eq.~\ref{eq:tidalradius}, $\sigma_v$ by Eq.~\ref{eq:velscale}, $K$ is either $+1$ or $-1$ depending on whether the star is in the leading or trailing tail, and $c$ is a hyper-parameter that specifies the exact position of the effective Lagrange points.
+$\J$ is the Jacobian that defines the transformation from heliocentric, spherical to Galactocentric, cartesian coordinates.
+This model could trivially allow for mass loss by incorporating the initial satellite mass, $m$ and, e.g., a constant mass-loss term, $\dot{m}$, but for now we neglect this.
 
-The likelihood of a star's position today is computed by first transforming the observed position to Galactocentric cartesian coordinates and integrating the orbits of the progenitor and star backwards for several Gigayears in the parent potential, $\Phi$, treating both as test particles.
-The distance, $R$, and velocity, $V$, of the star relative to the progenitor are computed at each time-step, and the likelihood is computed as the product of the isotropic, log-normal centered on the logarithm of the instantaneous tidal radius with the isotropic, log-normal centered on the logarithm of the velocity dispersion of the progenitor:
-\begin{align}
-	p(\W \given \W_p, \tau, \Phi) &= p(\X \given \X_p, \tau, \Phi) \jac{(l,b,D,\mu_l,\mu_b,v_r)}_{t=\tau}\\
-	p(\X \given \X_p, \tau, \Phi) &= p(\ln R, \ln V \given \tau, \Phi) \jac{(x,y,z,v_x,v_y,v_z)}_{t=\tau}\label{eq:int} \\
-	p(\ln R, \ln V \given \tau, \Phi) &= \left[\mathcal{N}(\ln R | \ln \rtide, \sigma_R)\mathcal{N}(\ln V | \ln \sigma_v, \sigma_V)\right]_{t=\tau}
-\end{align}
-% , but the vectors $\D$ and $\W$ [exist] in different coordinate systems --- heliocentric spherical and Galactocentric cartesian, respectively --- so we must also track the Jacobian that [describes] this transformation:
-where in eq.~\ref{eq:int} we have implicitly integrated over and neglected the two angle coordinates associated with $\ln R$ and the two angles associated with $\ln V$. 
-\begin{align}
-	\ln p(\ln R, \ln V \given \tau, \Phi) = -\ln(2\pi)-\ln(\sigma_R\sigma_V)-
-		\ln\left[\left(\frac{R}{\rtide}\right)^{1/\sigma_R} \left(\frac{V}{\sigma_v}\right)^{1/\sigma_V}\right]
-\end{align}
-We then compute the marginal likelihood, $p(\W \given \W_p, \Phi)$, by integrating over all possible unbinding times,
-\begin{align}
-	p(\W \given \W_p, \Phi) = \int p(\W \given \W_p, \tau, \Phi)p(\tau) d\tau.
-\end{align}
-In general, $p(\tau)$ is likely conditional on the orbit of the progenitor --- stars are preferentially stripped at pericentric passages --- but this is neglected and is taken to be uniform over the entire interaction history between the progenitor and parent.
+The full posterior probability for the observed position of a star, progenitor, and parent potential is
+\begin{equation}
+	p(\Phi, \W, \W_\sat, \tub, K, c \given \D, \D_\sat) = \frac{1}{\mathcal{Z}} p(\D \given \W) p(\D_\sat \given \W_\sat) 
+												       p(\W \given \W_\sat, \tub, \Phi, K, c)
+												       p(\Phi) p(\tau)p(c)
+\end{equation}
+where the factor $\mathcal{Z}$ only depends on the properties of the data (the \emph{evidence}), and is thus constant when varying the model parameters.
+
+In general, $p(\tub)$ is likely conditional on the orbit of the progenitor --- stars are preferentially stripped at pericentric passages --- but this is neglected and is taken to be uniform over the entire interaction history between the progenitor and parent.
 

### 2014-03-07 text/paper2/ms.tex
intro work

@@
 Our method does not rely on integrability of the potential;
-  we don't expect there to be any good approximation to the large-scale 
-  Milky Way potential that is integrable.
+  we do not know that any good approximation to the large-scale 
+  Milky Way potential is integrable.
 % Methods: 
@@
   and missing data.
+This method advances in a few key points that have not been jointly addressed in prior work: assumptions of integrability, proper accounting for observational uncertainties, and allowing for missing dimensions.
 We test our method with simulated data in a non-integrable triaxial potential, with realistic observational uncertainties for RR Lyrae stars.
@@
 
-Cosmological simulations of galaxy formation in the $\Lambda$CDM paradigm predict dark matter halos that (1) are permeated with substructure on many scales, (2) are triaxial in shape, and (3) have density profiles and shapes that vary with radius \citep{dubinski91, jing02, kuhlen07, veraciro11}.
+Early large-scale, cosmological simulations of galaxy formation in the $\Lambda$CDM paradigm suggested that the spherically-averaged density profiles of dark matter halos follow a universal profile across a huge dynamic range in mass \citep{navarro96}.
+Since then, higher resolution simulations --- both with and without baryons --- have produced dark matter halos that (1) are permeated with substructure on many scales, (2) are triaxial in shape, and (3) have shapes and orientations that vary with radius \citep{dubinski91, jing02, kuhlen07, veraciro11}.
 Dark-matter-only simulations produce triaxial halos \citep{jing02} with large density fluctuations \citep{zemp09}.
 Including baryons tends to soften the triaxiality and graininess in the inner galaxy through a combination of dissipative infall \citep{dubinski94} or cooling \citep{bryan13}.
-These processes combined with the gravity from a baryonic disk (ellipsoid) can act to make the inner halo more oblate (spherical), however these processes do not erase the clumpy, triaxial nature of the outer halo \citep[e.g.,][]{pontzen12}.
-This leads to radially-dependent axis ratios and orientation, density profile, and smoothness, thus suggesting complicated, non-intregrable potential forms for Milky Way-like galaxies.
+These processes combined with the gravity from a baryonic disk or ellipsoid can act to make the inner halo more oblate or spherical, however they do not seem to erase the clumpy, triaxial nature of the outer halo \citep[e.g.,][]{pontzen12}.
+This leads to radially-dependent axis ratios, orientation, and smoothness, thus suggesting complicated mass distributions for Milky Way-like galaxies that are not easily represented by integrable potentials.
 
@@
 For example, the rotation curves of disk galaxies are excellent probes of the inner mass, especially since matter in disks can be assumed to move on near circular orbits.
-Conversely, measuring the dark matter distribution at large radii is complicated by the low density of visible tracers, observational difficulties, and [unclear what assumptions can be made about the integrability and smoothness of the mass distribution].
+Conversely, measuring the dark matter distribution at large radii is complicated by the low density of visible tracers, observational difficulties, and assumptions about the smoothness and analytic representation.
 Around external galaxies, the extended mass distribution has been probed using a variety of approaches \citep[see][for a a complete and detailed review]{courteau13}.
@@
 Our proximity allows us to use individual stars as kinematic tracers and hence build much larger samples that probe deeper into the halo than the globular cluster and planetary nebula studies of external galaxies.
-For example, \cite{deason12a} used halo BHB stars selected from the Sloan Digital Sky Survey \cite[SDSS;][]{york00} as a random tracer population to measure the mass and slope of a simple, power-law fit to the Galactic potential.
-Such studies assume that the tracer orbits are randomly sampled from a smooth distribution function and phase mixed, however the existence of large-scale substructure in stellar streams and clouds is known to bias mass and velocity inferences that ignore this fact.
-Another approach is to exploit the non-random nature of the stellar populations in the halo.
+For example, \cite{deason12a} used halo BHB stars selected from the Sloan Digital Sky Survey \cite[SDSS;][]{york00} as a random tracer population to measure the mass and slope of a power-law fit.
+Such studies assume that the tracer orbits are randomly sampled from a smooth distribution function and are fully phase mixed.
+However, large photometric surveys such as the SDSS have discovered large amounts of substructure --- in streams and associations of stars --- in the Milky Way halo, thus demonstrating that the stellar distribution is not phase-mixed.
+Since the initial discovery of tidal debris from the Sagittarius (Sgr) dwarf spheroidal galaxy, myriad other streams and debris systems have been identified probing Galactocentric distances from $\sim$15-100 kpc.
+The existence of large-scale substructure in stellar streams and clouds is known to bias mass and velocity inferences by several tens of percent \citep{yencho06}.
+
+Another approach to using halo stars as potential measures is to exploit the non-random nature of halo.
+For example, \cite{law10} used N-body simulations of the disruption of the Sgr dwarf to simultaneously model the debris and underlying Milky Way potential.
+By varying parameters of the Milky Way potential, they found ``best-fit'' parameters by comparing the properties of observed Sgr stars to their simulated debris over a grid of potential parameters.
+The computational costs of running N-body simulations limited their search to a grid of values and forced the authors to fix many other parameters (e.g., properties of the disk and bulge).
+Nevertheless, they were able to constrain the shape of the Milky Way potential out to $\sim$70 kpc and found that the best-fitting halo has a prolate, triaxial shape elongated in the Galactic Z direction, orthogonal to the plane of the disk.
+Though an unlikely orientation for the halo --- e.g., \cite{debattista13} find that the disk of the Milky Way would not remain stable in such a configuration --- LM10 showed that the data is at a state where such inference can be attempted.
+The computational costs associated with N-body modeling has motivated many methods that approximately model tidal streams.
+
 Tidal streams are dynamically cold systems --- debris typically have small distributions of energy and angular momentum --- and thus require orders of magnitude fewer tracers than a random sample to get constraints of comparable accuracy.
 For example, in the simplest case we might assume that debris stars are actually still on the same orbit as their progenitor system \citep[a \emph{wrong} assumption, see e.g.,][]{eyre11}.
-Then we can imagine measuring the full-space velocities ${\bf v}$ at different points ${\bf x}$ in the structure (e.g., along a stream), which would give us a direct measure of differences in a potential, $\Phi$.
-%In practice, simple methods such as orbit fitting lead to systematic biases in inferred properties of the underlying potential because streams do not strictly delineate orbits [cite eyre binney work].
+Then we can imagine measuring the full-space velocities ${\bf v}$ at different points ${\bf x}$ in the structure (e.g., along a stream), which would give us a direct measure of differences in a potential, $\Phi$, without running a full model of the disruption.
+An extension of this idea is to fit a single orbit to observed debris \citep[e.g.,][]{koposov10, deg13}.
+Though this is known to be incorrect and leads to biases in inferred properties of the underlying potential \citep[e.g.,][]{eyre11, lux13, sanders13a}, \cite{deg14} and \cite{lux13} have used orbit fitting to demonstrate the power of combining multiple streams in dynamical inference.
+To account for the offset between the orbit of the progenitor and the orbits of the debris stars, methods have been proposed that model the debris as an orbit plus some dispersion or offset, either in phase space \citep[e.g.,][]{eyre09a, varghese11, kuepper12} or action-angle coordinates \citep{eyre11, sanders13b, bovy14, sanders14}.
+Other statistical methods have been proposed \citep[][]{penarrubia12, sanderson14} that may prove powerful when applied to, e.g., data from the \gaia\, mission, where full 6D coordinates will be known for large samples of stars in the halo --- and therefore many debris structures --- but stream membership is not known for all stars.
 
-Since the initial discovery of tidal debris from Sgr, myriad other streams and debris systems have been identified in large photometric surveys, probing Galactocentric distances from $\sim$15-100 kpc.
-The known debris structures are $\gtrsim$10 kpc from the Sun, thus while overdensities in the halo are less contaminated by background, the constituent stars often have incomplete or poorly measured phase-space information.
-Perhaps the most rigorous method of modeling streams and the potential is to run full N-body simulations of satellite disruption and compare the density of observed stars to simulated stars in whatever coordinates are measured.
-A generative N-body model could naturally incorporate observational uncertainties, missing dimensions, and would, in principle, allow for studying any arbitrary time-dependent or non-integrable potential.
-However, such a model is presently computationally intractable.
-\cite{law10} demonstrated the power of this idea by running a grid of N-body models of the disruption of the Sagittarius (Sgr) dwarf spheroidal galaxy and matched their simulated streams to all Sgr data available at that time.
-By comparing the phase-space positions of their stream particles with the data, they were able to constrain the shape of the Milky Way potential out to $\sim$70 kpc and found that the best-fitting halo has a prolate, triaxial shape elongated in the Galactic Z direction, orthogonal to the plane of the disk.
-% (though, see ... for possible complications in the Sgr debris belokurov 2012?).
+We identify XXX problems with the approximate models mentioned above; each method solves only a subset of these issues:
+\begin{enumerate}
+	\item The known debris structures are $\gtrsim$10 kpc from the Sun, thus while overdensities in the halo are less contaminated by background, the constituent stars often have incomplete or poorly measured phase-space information.
+Thus it is critical for any approximate model of tidal debris to incorporate observational uncertainties and missing dimensions in a consistent and justified way.
+	\item The potential of the inner galaxy is plausibly well-described by a (set of) nearly-integrable potentials, thus the dynamics of collisionless systems is conceivably captured by a transformation to some integral-of-motion-space, e.g., action-angle coordinates.
+However there is large uncertainty in the radial profile, shape, orientation, and graininess of the outer halo and the constancy of these parameters over distance.
+To test the degree to which the outer halo is well-fit by integrable potentials, it is problematic to use methods that \emph{assume} integrability.
+	\item Matching densities, small numbers of stars in Gaia %Generative models that instead rely on matching densities (in phase-space, action-space, or observed coordinates) must also include fine details about the progenitor (e.g., mass function or structure) and a model for the background of stars.
+	\item Multiple streams and other information
+	\item Computational costs [...]
+\end{enumerate}
 
-The difficulty and computational costs associated with proper N-body modeling has motivated numerous efforts for developing approximate methods for constraining properties of the Galactic potential with cold debris structures.
-The simplest approximation is to fit a single orbit to observed debris \citep[e.g.,][]{koposov10, deg13}.
-This has the advantage of naturally incorporating observational errors, but is known to be invalid as debris does not exactly follow an orbit \citep[e.g.,][]{eyre11, sanders13a}.
-To account for this, methods have been proposed that model the debris as an orbit plus some dispersion or offset, either in phase space \citep[e.g.,][]{eyre09a, varghese11, kuepper12} or action-angle coordinates \citep{eyre11, sanders13b, bovy14, sanders14}.
-Other statistical methods have been proposed \citep[][]{penarrubia12, sanderson14} that may prove powerful when applied to, e.g., data from the \gaia\, mission, where full 6D coordinates will be known for large samples of stars in the halo --- and therefore many debris structures --- but stream membership is not known for all stars (however these statistical methods do not account for observational errors).
+%[][hereafter \paperone]
+In \citet{apw13}, we introduced a simple method (\emph{Rewinder}) for using individual stars associated with tidal debris combined with knowledge about the mass and orbit of the progenitor to constrain properties of the host galaxy potential.
+The method exploits the relationship between the phase-space distribution of debris and (measurable) properties of the progenitor system (e.g., the tidal radius and escape velocity).
+In this article, we use these simple scalings --- which depend only on the mass and orbit of the progenitor and parent potential --- to present a fully probabilistic method for modeling tidal streams in any underlying potential given data with realistic uncertainties.
+The method relies on numerical orbit integration in ordinary phase-space and can thus incorporate arbitrarily complex models for the parent potential.
+The model uses individual stars as constraints --- not density matching --- and can therefore handle very small samples of stars.
+The model trivially extends to incorporating multiple streams, debris structures, and other kinematic information.
+[...something about computational cost]
 
-The potential of the inner galaxy is likely well-described by a (set of) nearly-integrable potentials, thus the dynamics of collisionless systems is plausibly captured by a transformation to some integral-of-motion-space, e.g., action-angle coordinates.
-However there is uncertainty in the exact shape, radial profile, and graininess of the outer halo.
-To test the extent to which the outer halo is well-fit by integrable potentials, we cannot use methods that \emph{assume} integrability.
-In this article, we present a fully probabilistic method for measuring properties of a potential that only relies on numerical orbit integration in ordinary phase-space.
-The model uses individual stars as constraints on the potential, and can therefore handle very small samples of stars.
-Generative models that instead rely on matching densities (in phase-space, action-space, or observed coordinates) must also include fine details about the progenitor (e.g., mass function or structure) and a model for the background of stars.
-Our method also naturally extends to using multiple debris structures \citep[see][who illustrate the power of using multiple streams to simultaneously constrain the potential]{deg14}.
-
-Although the \gaia\, project will deliver hundreds of millions of stars useful for dynamical inference,
-  it still might be that the most relevant stellar samples for inferring the Milky Way potential might be very small in the end.
-For instance, stars that produce good distance estimates might be much more valuable than typical stars in the sample.
-At halo distances, \gaia's parallaxes are not expected to be better than the distance indicators produced by the best
-  variable stars observed at the best wavelengths; for example, near-infrared observations of RR Lyrae stars
-  can in principle deliver $\sim$2\% percent photometric distances at essentially all halo distances \citep{madore12}.
-These valuable stars are rare (and their abundances are age and metallicity dependent);
-  there could be many cold structures in the Milky Way halo that are highly constraining
-  on the potential in principle, but which contain only a few good distance-indicating members.
-That said, it is not yet known what the trade-offs are between having many stars at low precision and a few at high precision,
-  nor is it known how valuable distance information really is, when a structure contains many precisely observed members.
-
-Another reason to think about small samples is more conceptual:
-in dynamical inferences that go beyond simple virial or Jeans arguments,
-  inference depends on the non-linear dynamical evolution of populations in phase space.
-The nonlinearities, combined with heteroskedastic data and requirements to marginalize over phase-space distribution functions,
-  lead naturally to situations in which some stars are \emph{far more informative} than others.
-For a simple example, in a toy study of the Solar System,
-  phase-space distribution marginalization was shown to make some planets are far more informative than others
-  about the gravitational potential \citep{bovy10}.
-This is expected to be a generic feature in dynamical problems:
-  it might be that a very small fraction of \gaia's catalog entries
-  deliver a large fraction of the information about the Milky Way's potential and formation history.
-
-In \citet[][hereafter \paperone]{apw13}, we introduced a simple method (\emph{Rewinder}) for using individual stars associated with tidal debris combined with knowledge about the mass and orbit of the progenitor to constrain properties of the host galaxy potential.
-The method exploits the relationship between the phase-space distribution of debris and (measurable) properties of the progenitor system (e.g., the tidal radius and escape velocity).
-In this paper, we investigate how these relationships scale with the mass and orbit of the progenitor and the quality of the data, then present a new probabilistic framework for simultaneously modeling stream debris and the parent potential based on the same arguments.
-
-In Section~\ref{sec:sims} we describe a suite of N-body simulations that span a range of progenitor masses over two characteristic orbits. In Section~\ref{sec:} we
+In Section~\ref{sec:sims} we describe a suite of N-body simulations that span a range of progenitor masses on a characteristic, mildly eccentric orbit. In Section~\ref{sec:} we describe [...]
 
@@
 We [...] N-body simulations run with an SCF code [...]. [describe simulations here] [Describe tub]
+
+The potential is taken to be a three-component sum of a Miyamoto-Nagai disk \citep{}, Hernquist bulge \citep[spheroid;][]{}, and a triaxial, logarithmic halo \citep[e.g.,][]{law10}:
+\begin{align}
+	&\Phi_{\rm disk}(R,z) = -\frac{GM_{\rm disk}}{\sqrt{R^2 + (a + \sqrt{z^2 + b^2})^2}}\\
+	&\Phi_{\rm spher}(r) = -\frac{GM_{\rm spher}}{r + c}\\
+	&\Phi_{\rm halo}(x,y,z) = v_{h}^2 \ln(C_1 x^2 + C_2 y^2 + C_3 xy + (z/q_z)^2 + r_h^2)
+\end{align}
+where $C_1$, $C_2$, and $C_3$ are combinations of the $x$ and $y$ axis
+ratios ($q_1$, $q_2$) and orientation of the halo with respect to the
+baryonic disk ($\phi$):
+\begin{align}
+  C_1 &= \frac{\cos^2\phi}{q_1^2} + \frac{\sin^2\phi}{q_2^2}\\
+  C_2 &= \frac{\sin^2\phi}{q_1^2} + \frac{\cos^2\phi}{q_2^2}\\
+  C_3 &= 2\sin\phi\cos\phi \left(q_1^{-2} - q_2^{-2}\right).
+\end{align}
+The total potential then is just
+\begin{equation}
+	\Phi_{\rm tot} = \Phi_{\rm disk} + \Phi_{\rm spher} + \Phi_{\rm halo}\label{eq:lm10}.
+\end{equation}
 
@@
 
-For a satellite galaxy on a mildly eccentric orbit, mass-loss is driven via tidal stripping due to the tidal field of the parent system or tidal shocking due to the rapidly changing tidal field as the progenitor system moves through pericenter.
-The interplay between these two processes as a function of orbital properties will be studied in forthcoming work and focus below on the slow disruption of stars due to steady tidal forcing.
+For a satellite galaxy on a mildly eccentric orbit, mass-loss is driven by a combination of tidal stripping due to the tidal field of the parent system and tidal shocking due to the rapidly changing tidal field as the progenitor system moves through pericenter.
+The interplay between these two processes as a function of orbital properties will be studied in forthcoming work; we focus below on the slow disruption of stars due to steady tidal forcing.
 
-Consider a point-mass satellite of mass $m$ on a circular orbit with frequency $\Omega$ around a point-mass of mass $M\gg m$ \citep[the restricted three-body problem, e.g.,][]{binneytremaine}.
+Consider a point-mass satellite of mass $m$ on a circular orbit with frequency $\Omega$ around a point-mass of mass $M\gg m$ \citep[the restricted three-body problem; e.g., \S 8.3][]{binneytremaine}.
 In a frame rotating with the orbital frequency of the satellite, the effective potential around the satellite has (amongst others) two stationary, unstable points located around $\sim R \pm r_J$, where $R$ is the distance to the satellite and $r_J$ is the Jacobi or tidal radius,
@@
 \begin{equation}
-	r_{\rm tide} = R\left(\frac{m_{\rm sat}}{3M_{\rm enc}}\right)^{1/3}\label{eq:tidalradius}
+	r_{\rm tide} = R\left(\frac{m}{3M_{\rm enc}}\right)^{1/3}\label{eq:tidalradius}
 \end{equation}
-as a normalizing distance scale.
+as a normalizing distance scale, where $M_{\rm enc}$ is the instantaneous enclosed mass of the parent system within orbital radius $R$.
+At pericenter, the velocity of the satellite will be of order $V\sim \sqrt{GM_{\rm enc}/R}$.
+If the velocity dispersion of the satellite is $\sigma_v \sim \sqrt{Gm/r_J}$, then
+\begin{equation}
+	\sigma_v \sim V\left(\frac{m}{M_{\rm enc}}\right)^{1/3}\label{eq:velscale}.
+\end{equation}
+Thus, we expect the particle velocities to scale with $(m/M_{\rm enc})^{1/3}$.
 
-At pericenter, the velocity of the satellite will be of order $V\sim \sqrt{GM_{\rm enc}/R}$. If the velocity dispersion of the satellite is $\sigma_v \sim \sqrt{Gm/r_J}$, then
-\begin{equation}
-	\sigma_v \sim V\left(\frac{m_{\rm sat}}{M_{\rm enc}}\right)^{1/3}\label{eq:velscale}.
-\end{equation}
-Thus, we expect the particle velocities to scale with $(m/M)^{1/3}$.
-
-In a triaxial potential (e.g., Equation~\ref{eq:lm10}), the orbital plane of the satellite is not fixed but we still expect there to be effective Lagrange points along the line of centers connecting the origin of the parent potential to the satellite.
+In a triaxial potential, the orbital plane of the satellite is not fixed but we still expect there to be effective Lagrange points along the line of centers connecting the origin of the parent potential to the satellite.
 We thus proceed by defining a coordinate system relative to the position and velocity of the progenitor, ($\bs{r}_p$,$\bs{v}_p$), rotated into the instantaneous orbital plane defined by $\bs{r}_p \times \bs{v}_p$, and normalized by the instantaneous tidal radius for positions (Eq.~\ref{eq:tidalradius}) and instantaneous velocity scale for velocity (Eq.~\ref{eq:velscale}).
@@
 \end{align}
-Figure~\ref{fig:lpts_r} shows sections of particle orbits in this coordinate system, symmetric around the unbinding time, $\tub$ (Section~\ref{sec:sims}), as a function of progenitor mass.
+%(Section~\ref{sec:sims})
+Figure~\ref{fig:lpts_r} shows sections of particle orbits in this coordinate system, symmetric around the unbinding time, $\tub$, as a function of progenitor mass.
 In these coordinates, the classical Lagrange points would be located around $x_1=\pm1,x_2=0,x_3=0$ (illustrated by the unit circle in top panels of Figure~\ref{fig:lpts_r}).
 Figure~\ref{fig:lpts_v} shows the same for the velocity of each star.
-Though not centered on the point-mass Lagrange points, the location of and dispersion about the effective Lagrange points in these scaled coordinates remain remarkably consistent across the range of progenitor masses explored at $x_1\sim\pm2$.
-Furthermore, the dispersion in velocity is well-normalized by the velocity scale of Eq.~\ref{eq:velscale}, though it is clear that the velocity dispersion along $\hat{x}_1$, the radial vector to the satellite position, is larger than that in other dimensions.
-The phase-space dispersion of tidally stripped debris scales with the mass ratio $(m_{\rm sat} / M_{\rm enc})^{1/3}$, the orbital radius, and the orbital velocity of the progenitor system.
+Though not centered on the point-mass Lagrange points, the location of and dispersion about the effective Lagrange points in these scaled coordinates remain remarkably consistent across the range of progenitor masses explored.
+Furthermore, the dispersion in velocity is well-normalized by the velocity scale of Eq.~\ref{eq:velscale}, though it is clear that the velocity dispersion along $\hat{x}_1$, the radial vector to the satellite position, is larger than that in other dimensions, possibly due to mild tidal shocking.
+Even in a triaxial potential, the phase-space dispersion of tidally stripped debris scales with the mass ratio $(m / M_{\rm enc})^{1/3}$.
+%, the orbital radius, and the orbital velocity of the progenitor system.
 
@@
 To model the true, 6D, present-day position of the star, $\W$, we first transform to cartesian, Galactocentric coordinates where ($\bs{r},\bs{v}$) and ($\bs{r}_\sat,\bs{v}_\sat$) are the position and velocity of the star and progenitor, respectively.
-In cartesian coordinates, the present-day position of the star is modeled as having been sampled at $t=\tub$ from an isotropic Gaussian centered on one of the instantaneous, effective Lagrange points in position, and an isotropic Gaussian centered on the origin in velocity, and evolved to present day in the parent potential, $\Phi$.
+In cartesian coordinates, the present-day position of the star is taken as having been sampled at $t=\tub$ from an isotropic Gaussian centered on one of the instantaneous, effective Lagrange points in position, and an isotropic Gaussian centered on the origin in velocity, and evolved to present day in the parent potential, $\Phi$.
 That is, we assume that once the star becomes ``unbound'' from the satellite the potential of the satellite can be ignored, and thus treat the star as a test particle.
+We add two other parameters: $K$, a ``bit'' that is either $+1$ or $-1$ depending on whether the star is in the leading or trailing tail, and a constant offset, $c$, that specifies the exact position of the effective Lagrange points relative to the classical tidal radius.
 \begin{align}
 	p(\W \given \W_\sat, \tub, \Phi, K, c) &= p(\bs{r} \given \bs{r}_\sat, \tub, \Phi, K, c)p(\bs{v} \given \bs{v}_\sat, \tub, \Phi) \left\vert\J\right\vert_{t=\tub}\\
-	p(\bs{r} \given \bs{r}_\sat, \tub, \Phi, K, c) &= \mathcal{N}(\bs{r} \given \bs{r}_\sat - cK\rtide \hat{x}_1, \sigma_1)_{t=\tub}\\
+	p(\bs{r} \given \bs{r}_\sat, \tub, \Phi, K, c) &= \mathcal{N}(\bs{r} \given \bs{r}_\sat - cK\rtide \hat{x}_1, \sigma_r)_{t=\tub}\\
 	p(\bs{v} \given \bs{v}_\sat, \tub, \Phi) &= \mathcal{N}(\bs{v} \given \bs{v}_\sat, \sigma_v)_{t=\tub}
 \end{align}
-where $\rtide$ is given by Eq.~\ref{eq:tidalradius}, $\sigma_v$ by Eq.~\ref{eq:velscale}, $K$ is either $+1$ or $-1$ depending on whether the star is in the leading or trailing tail, and $c$ is a hyper-parameter that specifies the exact position of the effective Lagrange points.
+where $\rtide$ is given by Eq.~\ref{eq:tidalradius}, $\sigma_v$ by Eq.~\ref{eq:velscale}.
 $\J$ is the Jacobian that defines the transformation from heliocentric, spherical to Galactocentric, cartesian coordinates.
-This model could trivially allow for mass loss by incorporating the initial satellite mass, $m$ and, e.g., a constant mass-loss term, $\dot{m}$, but for now we neglect this.
+This model could trivially allow for mass loss by incorporating the initial satellite mass, $m$ and, e.g., a constant mass-loss term, $\dot{m}$.
 
-The full posterior probability for the observed position of a star, progenitor, and parent potential is
+The full posterior probability for the position of a star, progenitor, and properties of the parent potential, given the data, is then
 \begin{equation}
@@
 
-In general, $p(\tub)$ is likely conditional on the orbit of the progenitor --- stars are preferentially stripped at pericentric passages --- but this is neglected and is taken to be uniform over the entire interaction history between the progenitor and parent.
+[...]
+
+% In general, $p(\tub)$ is likely conditional on the orbit of the progenitor --- stars are preferentially stripped at pericentric passages --- but this is neglected and is taken to be uniform over the entire interaction history between the progenitor and parent.
 
 \section{Experiments}
+In what follows, we apply this model to ``observations'' of the N-body simulations described in Section~\ref{sec:sims}.
+For all tests in this article, we model the potential of the parent galaxy with the same functional form used in the simulations (Equation~\ref{}).
+
 Initially we assume the present-day mass of the progenitor, $M_p$, and velocity dispersion, $\sigma_v$, have been measured perfectly and neglect mass-loss from the progenitor.
-
-For all tests in this article, we model the potential of the parent galaxy as a three-component sum of a Miyamoto-Nagai disk \citep{}, Hernquist bulge \citep[spheroid;][]{}, and a triaxial, logarithmic halo \citep[e.g.,][]{law10}:
-\begin{align}
-	&\Phi_{\rm disk}(R,z) = -\frac{GM_{\rm disk}}{\sqrt{R^2 + (a + \sqrt{z^2 + b^2})^2}}\\
-	&\Phi_{\rm spher}(r) = -\frac{GM_{\rm spher}}{r + c}\\
-	&\Phi_{\rm halo}(x,y,z) = v_{h}^2 \ln(C_1 x^2 + C_2 y^2 + C_3 xy + (z/q_z)^2 + r_h^2)
-\end{align}
-where $C_1$, $C_2$, and $C_3$ are combinations of the $x$ and $y$ axis
-ratios ($q_1$, $q_2$) and orientation of the halo with respect to the
-baryonic disk ($\phi$):
-\begin{align}
-  C_1 &= \frac{\cos^2\phi}{q_1^2} + \frac{\sin^2\phi}{q_2^2}\\
-  C_2 &= \frac{\sin^2\phi}{q_1^2} + \frac{\cos^2\phi}{q_2^2}\\
-  C_3 &= 2\sin\phi\cos\phi \left(q_1^{-2} - q_2^{-2}\right).
-\end{align}
-The total potential then is just
-\begin{equation}
-	\Phi_{\rm tot} = \Phi_{\rm disk} + \Phi_{\rm spher} + \Phi_{\rm halo}\label{eq:lm10}.
-\end{equation}
 
@@
 
+Our method also naturally extends to using multiple debris structures \citep[see][who illustrate the power of using multiple streams to simultaneously constrain the potential]{deg14}.
+
+Although the \gaia\, project will deliver hundreds of millions of stars useful for dynamical inference,
+  it still might be that the most relevant stellar samples for inferring the Milky Way potential might be very small in the end.
+For instance, stars that produce good distance estimates might be much more valuable than typical stars in the sample.
+At halo distances, \gaia's parallaxes are not expected to be better than the distance indicators produced by the best
+  variable stars observed at the best wavelengths; for example, near-infrared observations of RR Lyrae stars
+  can in principle deliver $\sim$2\% percent photometric distances at essentially all halo distances \citep{madore12}.
+These valuable stars are rare (and their abundances are age and metallicity dependent);
+  there could be many cold structures in the Milky Way halo that are highly constraining
+  on the potential in principle, but which contain only a few good distance-indicating members.
+That said, it is not yet known what the trade-offs are between having many stars at low precision and a few at high precision,
+  nor is it known how valuable distance information really is, when a structure contains many precisely observed members.
+
+Another reason to think about small samples is more conceptual:
+in dynamical inferences that go beyond simple virial or Jeans arguments,
+  inference depends on the non-linear dynamical evolution of populations in phase space.
+The nonlinearities, combined with heteroskedastic data and requirements to marginalize over phase-space distribution functions,
+  lead naturally to situations in which some stars are \emph{far more informative} than others.
+For a simple example, in a toy study of the Solar System,
+  phase-space distribution marginalization was shown to make some planets are far more informative than others
+  about the gravitational potential \citep{bovy10}.
+This is expected to be a generic feature in dynamical problems:
+  it might be that a very small fraction of \gaia's catalog entries
+  deliver a large fraction of the information about the Milky Way's potential and formation history.
+
 [Cold vs. Hot streams]

### 2014-03-07 text/paper2/ms.tex
moved potential stuff to sims

@@
 Another approach to using halo stars as potential measures is to exploit the non-random nature of halo.
-For example, \cite{law10} used N-body simulations of the disruption of the Sgr dwarf to simultaneously model the debris and underlying Milky Way potential.
+For example, \citet[][LM10]{law10} used N-body simulations of the disruption of the Sgr dwarf to simultaneously model the debris and underlying Milky Way potential.
 By varying parameters of the Milky Way potential, they found ``best-fit'' parameters by comparing the properties of observed Sgr stars to their simulated debris over a grid of potential parameters.
@@
 \end{equation}
+Figure~\ref{fig:potential} shows equipotential slices in the Galactic X-Z plane (Y=0) and Y-Z plane (X=0) for a few choices of  $q_1$,  $q_z$, and $\phi$ while holding all other parameters fixed, as described in the figure caption.
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=\textwidth]{../../plots/paper2/potentials.pdf}
+\caption{ TODO }\label{fig:potential}
+\end{center}
+\end{figure}
 
@@
 \section{Experiments}
-In what follows, we apply this model to ``observations'' of the N-body simulations described in Section~\ref{sec:sims}.
-For all tests in this article, we model the potential of the parent galaxy with the same functional form used in the simulations (Equation~\ref{}).
-
-Initially we assume the present-day mass of the progenitor, $M_p$, and velocity dispersion, $\sigma_v$, have been measured perfectly and neglect mass-loss from the progenitor.
-
-\begin{figure}[h]
-\begin{center}
-% \includegraphics[width=\textwidth]{/Users/adrian/projects/streams/plots/paper2/potentials.pdf}
-\caption{ TODO }\label{fig:potential}
-\end{center}
-\end{figure}
-
+In what follows, we apply the model described above to ``observations'' of the N-body simulations described in Section~\ref{sec:sims}.
+For all tests in this article, we model the potential of the parent galaxy with the same functional form used in the simulations (Equation~\ref{eq:lm10}).
 When recovering the potential, we hold fixed the disk and spheroid parameters (see Table~\ref{tbl:potential}), along with two halo parameters: the scale radius, $r_h$, and one of the flattening components in the x-y plane, $q_2$.
 These halo parameters are degenerate with combinations of the other parameters --- e.g., setting $q_1=2.0$, $q_2=1.0$, $\phi=0^\circ$ is equivalent to setting $q_1=1.0$, $q_2=2.0$, $\phi=90^\circ$.
-The priors on the remaining halo parameters are taken to be uniform over a conservative domain of realistic values: for $v_h$, 100-200~kms$^{-1}$ corresponds to a range in solar circular velocities from $\sim$210-250 kms$^{-1}$ (holding other parameters fixed); the range in axis ratios allow for prolate, oblate, and mixed triaxiality; and $\phi$ is restricted to $\pm45^\circ$ around the value measured by LM10 ($\phi = 97^\circ$).
-Figure~\ref{fig:potential} shows equipotential slices in the Galactic X-Z plane (Y=0) and Y-Z plane (X=0) for a few choices of  $q_1$,  $q_z$, and $\phi$ while holding all other parameters fixed, as described in the figure caption.
+The priors on the remaining halo parameters are taken to be uniform over a conservative domain of realistic values: for $v_h$, 100-200~kms$^{-1}$ corresponds to a range in solar circular velocities from $\sim$210-250 kms$^{-1}$ (holding other parameters fixed); the range in axis ratios allow for prolate, oblate, and generic triaxiality; and $\phi$ is restricted to $\pm45^\circ$ around the value measured by LM10 ($\phi = 97^\circ$).
 
@@
 
+\subsection{Perfect data}
+
 \subsection{\gaia + \spitzer}
-4 stars in Sgr-like stream
+8 stars in Sgr-like stream
+
+We assume the present-day mass of the progenitor, $M_p$, and velocity dispersion, $\sigma_v$, have been measured perfectly and neglect mass-loss from the progenitor.
 

### 2014-03-07 scripts/test_infer_potential.py
fix swtich to p_shocked

@@ -38,10 +38,10 @@ logging.basicConfig(level=logging.DEBUG)
 minimum_config = """
 name: test
 data_file: data/observed_particles/2.5e8_N1024.hdf5
-# nparticles: 7
-# particle_idx: [758, 635, 503, 122, 961, 713, 360]
 nparticles: 8
 particle_idx: [1019, 606, 740, 301, 261, 39, 735, 448]
+# nparticles: 4
+# particle_idx: [1019, 606, 740, 301]
 # nparticles: 16
 # nparticles: 128
 
@@ -61,7 +61,7 @@ pot_params = """
 """
 
 ptc_params = """
-    parameters: [shocked_bit]
+    parameters: [p_shocked]
 """
 #    parameters: [d, mul, mub, vr]
 
@@ -77,9 +77,9 @@ lm10_c = minimum_config.format(potential_params=pot_params,
 # _config = minimum_config.format(potential_params=pot_params,
 #                                 particles_params=ptc_params,
 #                                 satellite_params=sat_params)
-_config = minimum_config.format(potential_params=pot_params,
-                                particles_params="",
-                                satellite_params=sat_params)
+_config = minimum_config.format(potential_params="",
+                                particles_params=ptc_params,
+                                satellite_params="")
 
 # particles_params=ptc_params,
 # satellite_params=sat_params
@@ -193,7 +193,7 @@ class TestStreamModel(object):
                             plt.close('all')
                             idx += 1
 
-                    elif param_name == 'shocked_bit':
+                    elif param_name == 'p_shocked':
                         for jj in range(param.value.shape[0]):
                             vals1 = np.linspace(param._prior.a[jj],
                                                 param._prior.b[jj],

### 2014-03-07 streams/inference/core.py

@@ -398,8 +398,7 @@ class StreamModel(object):
         try:
             p_shocked = param_dict['particles']['p_shocked']
         except KeyError:
-            #p_shocked = self.particles.p_shocked.truth
-            p_shocked = np.array([0,0,0,0,0,1,0,0.])
+            p_shocked = self.particles.p_shocked.truth
 
         # heliocentric satellite positions
         s_hel = []

### 2014-03-07 streams/io/core.py

@@ -137,8 +137,10 @@ def read_hdf5(h5file, nparticles=None, particle_idx=None):
         tail_bit = ModelParameter(name="tail_bit",
                                   truth=true_tail_bit,
                                   prior=LogUniformPrior([-2.]*nparticles,[2.]*nparticles))
+
+        # truth=[0.]*nparticles,
         p_shocked = ModelParameter(name="p_shocked",
-                                   truth=[0.]*p_shocked,
+                                   truth=[0,0,0,0,0,1,0,0.],
                                    prior=LogUniformPrior([0.]*nparticles,[1.]*nparticles))
 
         if "error" in ptcl.keys():

### 2014-03-10 text/paper2/ms.tex
add note about mixture model
more writing

@@
   and the location of the progenitor object.
-The method generalizes to to encompass mixtures of independent debris structures,
+The method generalizes to encompass mixtures of independent debris structures,
   or situations in which star membership in structures is unknown.
-[Get more quantitative here?]
+% [Get more quantitative here?]
 % Conclusions: 
@@
 Another approach to using halo stars as potential measures is to exploit the non-random nature of halo.
-For example, \citet[][LM10]{law10} used N-body simulations of the disruption of the Sgr dwarf to simultaneously model the debris and underlying Milky Way potential.
+For example, \citet[][LM10]{law10} used n-body simulations of the disruption of the Sgr dwarf to simultaneously model the debris and underlying Milky Way potential.
 By varying parameters of the Milky Way potential, they found ``best-fit'' parameters by comparing the properties of observed Sgr stars to their simulated debris over a grid of potential parameters.
-The computational costs of running N-body simulations limited their search to a grid of values and forced the authors to fix many other parameters (e.g., properties of the disk and bulge).
+The computational costs of running n-body simulations limited their search to a grid of values and forced the authors to fix many other parameters (e.g., properties of the disk and bulge).
 Nevertheless, they were able to constrain the shape of the Milky Way potential out to $\sim$70 kpc and found that the best-fitting halo has a prolate, triaxial shape elongated in the Galactic Z direction, orthogonal to the plane of the disk.
 Though an unlikely orientation for the halo --- e.g., \cite{debattista13} find that the disk of the Milky Way would not remain stable in such a configuration --- LM10 showed that the data is at a state where such inference can be attempted.
-The computational costs associated with N-body modeling has motivated many methods that approximately model tidal streams.
+The computational costs associated with n-body modeling has motivated many methods that approximately model tidal streams.
 
@@
 
-We identify XXX problems with the approximate models mentioned above; each method solves only a subset of these issues:
+We identify a number of problems with approximate models; each method  mentioned above solves only a subset of these issues:
 \begin{enumerate}
@@
 To test the degree to which the outer halo is well-fit by integrable potentials, it is problematic to use methods that \emph{assume} integrability.
-	\item Matching densities, small numbers of stars in Gaia %Generative models that instead rely on matching densities (in phase-space, action-space, or observed coordinates) must also include fine details about the progenitor (e.g., mass function or structure) and a model for the background of stars.
-	\item Multiple streams and other information
-	\item Computational costs [...]
+	\item Near-future photometric surveys such as \gaia\, and LSST will likely discover many new streams and kinematic associations of stars.
+Potential recovery methods should be able to simultaneously model multiple streams and incorporate other dynamical constraints.
+	\item Matching generated models to observed stream densities is difficult: models that rely on this must account for observational biases and properly model the underlying background of halo stars. 
+	\item Full N-body models are expensive to run.
+Incorporating an n-body model into a general model and then performing a parameter search would be computationally intensive and, presently, intractable.
 \end{enumerate}
@@
 The method exploits the relationship between the phase-space distribution of debris and (measurable) properties of the progenitor system (e.g., the tidal radius and escape velocity).
+Specifically, a better model for the potential is one in which orbits of test particle stars (integrated backwards from their present position) came close (within the tidal radius in position, and escape velocity in velocity) to the orbit of the progenitor.
 In this article, we use these simple scalings --- which depend only on the mass and orbit of the progenitor and parent potential --- to present a fully probabilistic method for modeling tidal streams in any underlying potential given data with realistic uncertainties.
 The method relies on numerical orbit integration in ordinary phase-space and can thus incorporate arbitrarily complex models for the parent potential.
-The model uses individual stars as constraints --- not density matching --- and can therefore handle very small samples of stars.
+The model uses individual stars as constraints --- not density matching --- and can therefore handle very small samples of well-measured stars (e.g., 4-16).
 The model trivially extends to incorporating multiple streams, debris structures, and other kinematic information.
-[...something about computational cost]
 
-In Section~\ref{sec:sims} we describe a suite of N-body simulations that span a range of progenitor masses on a characteristic, mildly eccentric orbit. In Section~\ref{sec:} we describe [...]
+In Section~\ref{sec:sims} we describe a suite of n-body simulations that span a range of progenitor masses on a characteristic, mildly eccentric orbit. In Section~\ref{sec:} we describe [...]
 
@@
 [David Hendel]
-We [...] N-body simulations run with an SCF code [...]. [describe simulations here] [Describe tub]
+We [...] n-body simulations run with an SCF code [...]. [describe simulations here] [Describe tub]
 
@@
 \includegraphics[width=\textwidth]{../../plots/paper2/potentials.pdf}
-\caption{ TODO }\label{fig:potential}
+\caption{ Equipotential contours for the LM10 potential (Eq.~\ref{eq:lm10}) in Galactocentric, cartesian coordinates for various halo parameter choices.
+For all panels, $v_h=121.858~\mathrm{km}/\mathrm{s}$, $r_h=12~\mathrm{kpc}$.
+Left to right, each column represents a new choice of parameters.
+If not specified, other parameters are fixed to $q_1=q_2=q_z=1$ and $\phi=0^\circ$ (far left panels).
+}\label{fig:potential}
 \end{center}
@@
 \caption{ Orbits of 2000 particles around $t=\tub$.
-The orbits were integrated backwards from their present-day positions (final time step of the N-body simulations) without the potential of the progenitor.
+The orbits were integrated backwards from their present-day positions (final time step of the n-body simulations) without the potential of the progenitor.
 TODO: coordinates...}\label{fig:lpts_r}
@@
 
+This model works well when all stars are lost due to tidal pruning.
+Even on a mildly eccentric orbit, stars become preferentially unbound around the satellite's pericenter because of tidal shocking (cite?).
+We will explore properly modeling this process in future work.
+For this article, we modify the above to be a mixture model: we add a new parameter for each star, $P_{\rm shock}$, which is the probability that the star is drawn from a broader, ``shocked'' distribution instead of a tight Gaussian centered on an effective Lagrange point:
+\begin{align}
+	p_{\rm strip} &= \left [ \mathcal{N}(\bs{r} \given \bs{r}_\sat - cK\rtide \hat{x}_1, \sigma_r)\mathcal{N}(\bs{v} \given \bs{v}_\sat, \sigma_v) \left\vert\J\right\vert\right]_{t=\tub}\\
+	p_{\rm shock} &= \left [ \mathcal{N}(\bs{r} \given \bs{r}_\sat, 3\sigma_r)\mathcal{N}(\bs{v} \given \bs{v}_\sat, 3\sigma_v) \left\vert\J\right\vert\right]_{t=\tub}\\
+	p(\W \given \W_\sat, \tub, \Phi, K, c, P_{\rm shock}) &= P_{\rm shock}p_{\rm shock} + (1-P_{\rm shock})p_{\rm strip}
+\end{align}
 The full posterior probability for the position of a star, progenitor, and properties of the parent potential, given the data, is then
@@
 
-[...]
-
 % In general, $p(\tub)$ is likely conditional on the orbit of the progenitor --- stars are preferentially stripped at pericentric passages --- but this is neglected and is taken to be uniform over the entire interaction history between the progenitor and parent.
@@
 \section{Experiments}
-In what follows, we apply the model described above to ``observations'' of the N-body simulations described in Section~\ref{sec:sims}.
+In what follows, we apply the model described above to ``observations'' of the n-body simulations described in Section~\ref{sec:sims}.
 For all tests in this article, we model the potential of the parent galaxy with the same functional form used in the simulations (Equation~\ref{eq:lm10}).
@@
 These halo parameters are degenerate with combinations of the other parameters --- e.g., setting $q_1=2.0$, $q_2=1.0$, $\phi=0^\circ$ is equivalent to setting $q_1=1.0$, $q_2=2.0$, $\phi=90^\circ$.
-The priors on the remaining halo parameters are taken to be uniform over a conservative domain of realistic values: for $v_h$, 100-200~kms$^{-1}$ corresponds to a range in solar circular velocities from $\sim$210-250 kms$^{-1}$ (holding other parameters fixed); the range in axis ratios allow for prolate, oblate, and generic triaxiality; and $\phi$ is restricted to $\pm45^\circ$ around the value measured by LM10 ($\phi = 97^\circ$).
+The priors on the remaining halo parameters are taken to be uniform over a conservative domain of realistic values: for $v_h$, 100-200~km/s corresponds to a range in solar circular velocities from $\sim$210-250 km/s (holding other parameters fixed); the range in axis ratios allow for prolate, oblate, and generic triaxiality; and $\phi$ is restricted to $\pm45^\circ$ around the value measured by LM10 ($\phi = 97^\circ$).
 

### 2014-03-11 text/paper2/ms.tex
updated parameter table

@@
 \newcommand{\tub}{t_{\rm ub}}
+\newcommand{\tint}{t_{\rm int}}
 \newcommand{\rr}{\widetilde{\bs{r}}}
 \newcommand{\vv}{\widetilde{\bs{v}}}
+\newcommand{\tailbit}{\beta}
+\newcommand{\Loffset}{\alpha}
 
@@
 These processes combined with the gravity from a baryonic disk or ellipsoid can act to make the inner halo more oblate or spherical, however they do not seem to erase the clumpy, triaxial nature of the outer halo \citep[e.g.,][]{pontzen12}.
-This leads to radially-dependent axis ratios, orientation, and smoothness, thus suggesting complicated mass distributions for Milky Way-like galaxies that are not easily represented by integrable potentials.
+This can lead to radially-dependent axis ratios, orientation, and smoothness, thus suggesting complicated mass distributions for Milky Way-like galaxies that are not easily represented by integrable potentials.
 
@@
 \includegraphics[width=\textwidth]{../../plots/paper2/simulated_streams.pdf}
-\caption{ TODO }\label{fig:sims}
+\caption{ Particle positions in the final timestep of four n-body simulations with the same progenitor orbit initial conditions over a range of progenitor masses.
+The simulations were run [...].
+}\label{fig:sims}
 \end{center}
@@
 \subsection{Probabilistic model}
-[ intro, based on conclusion from previous section, etc.]
-
-We observe the position of a star, $\D$, in heliocentric coordinates --- e.g., the measured position on the sky, ($l$, $b$); distance, $d$; proper motions, ($\mu_l$, $\mu_b$); and line-of-sight velocity, $v_r$ --- and have determined through some other means that this star was once part of a progenitor system (e.g., satellite galaxy) with mass $m(t)$ that is disrupting and forming a cold debris structure in the potential, $\Phi$, of the parent galaxy (e.g., the Milky Way).
+We observe the 6D position of a star, $\D$, in heliocentric coordinates --- e.g., the measured position on the sky, ($l$, $b$); distance, $d$; proper motions, ($\mu_l$, $\mu_b$); and line-of-sight velocity, $v_r$ --- and have determined through some other means that this star was once part of a progenitor system (e.g., satellite galaxy) with mass $m(t)$ that is disrupting and forming a cold debris structure in the potential, $\Phi$, of the parent galaxy (e.g., the Milky Way).
 We assume that the mass of the parent system enclosed in the orbit of the satellite, $M_{\rm enc}$, is much larger than the initial mass of the satellite, $M_{\rm enc} \gg m(t=0)$.
@@
 In general, the data for the star and progenitor will have missing dimensions and significant uncertainties at distances typical to the Galactic halo, thus we define $\W$ and $\W_\sat$ as the true 6D, heliocentric positions of the star and progenitor and will include these in our model.
-By including the 6D position of the progenitor in the model, this model will work even when the progenitor location is unknown as in, e.g., the Orphan stream \citep{some orphan paper}.
+By including these positions in the model, this model will work even when the star has missing dimensions or the progenitor location is unknown as in, e.g., the Orphan stream \citep{some orphan paper}.
 
@@
 \begin{align}
-	p(\D \given \W) &= \mathcal{N}(\W \given \D, \bSigma)\\
-	p(\D_\sat \given \W_\sat) &= \mathcal{N}(\W_\sat \given \D_\sat, \bSigma_\sat)
+	p(\D \given \W) &= \mathcal{N}(\D \given \W, \bSigma)\label{eq:obsstar}\\
+	p(\D_\sat \given \W_\sat) &= \mathcal{N}(\D_\sat \given \W_\sat, \bSigma_\sat)\label{eq:obsprog}
 \end{align}
@@
 
-To model the true, 6D, present-day position of the star, $\W$, we first transform to cartesian, Galactocentric coordinates where ($\bs{r},\bs{v}$) and ($\bs{r}_\sat,\bs{v}_\sat$) are the position and velocity of the star and progenitor, respectively.
+To model the true, 6D, present-day position of the star, $\W$, we first transform to cartesian, Galactocentric coordinates where ($\bs{r},\bs{v}$) and ($\bs{r}_\sat,\bs{v}_\sat$) are the position and velocity of the star and progenitor.
 In cartesian coordinates, the present-day position of the star is taken as having been sampled at $t=\tub$ from an isotropic Gaussian centered on one of the instantaneous, effective Lagrange points in position, and an isotropic Gaussian centered on the origin in velocity, and evolved to present day in the parent potential, $\Phi$.
-That is, we assume that once the star becomes ``unbound'' from the satellite the potential of the satellite can be ignored, and thus treat the star as a test particle.
-We add two other parameters: $K$, a ``bit'' that is either $+1$ or $-1$ depending on whether the star is in the leading or trailing tail, and a constant offset, $c$, that specifies the exact position of the effective Lagrange points relative to the classical tidal radius.
+That is, we assume that once the star becomes ``unbound'' from the satellite the potential of the satellite can be ignored, and thus treat the star as a test particle.\footnote{This assumption breaks down for sufficiently large-mass progenitors, somewhere between $\sim10^8-10^9~$M$_\odot$.} We add two other parameters: a constant offset, $\Loffset$, that specifies the position of the effective Lagrange points relative to the classical tidal radius, and a binary parameter $\tailbit$, that is either $-1$ or $+1$ depending on whether the star is in the leading or trailing tail.
 \begin{align}
-	p(\W \given \W_\sat, \tub, \Phi, K, c) &= p(\bs{r} \given \bs{r}_\sat, \tub, \Phi, K, c)p(\bs{v} \given \bs{v}_\sat, \tub, \Phi) \left\vert\J\right\vert_{t=\tub}\\
-	p(\bs{r} \given \bs{r}_\sat, \tub, \Phi, K, c) &= \mathcal{N}(\bs{r} \given \bs{r}_\sat - cK\rtide \hat{x}_1, \sigma_r)_{t=\tub}\\
-	p(\bs{v} \given \bs{v}_\sat, \tub, \Phi) &= \mathcal{N}(\bs{v} \given \bs{v}_\sat, \sigma_v)_{t=\tub}
+	p(\W \given \W_\sat, \tub, \Phi, \Loffset, \tailbit) &= p(\bs{r} \given \bs{r}_\sat, \tub, \Phi, \Loffset, \tailbit)\,p(\bs{v} \given \bs{v}_\sat, \tub, \Phi)\,\left\vert\J\right\vert_{t=\tub}\\
+	p(\bs{r} \given \bs{r}_\sat, \tub, \Phi, \Loffset, \tailbit) &= \mathcal{N}(\bs{r} \given \bs{r}_\sat - \Loffset\,\tailbit\,\rtide\,\hat{x}_1, \sigma^2_r)_{t=\tub}\\
+	p(\bs{v} \given \bs{v}_\sat, \tub, \Phi) &= \mathcal{N}(\bs{v} \given \bs{v}_\sat, \sigma^2_v)_{t=\tub}
 \end{align}
-where $\rtide$ is given by Eq.~\ref{eq:tidalradius}, $\sigma_v$ by Eq.~\ref{eq:velscale}.
-$\J$ is the Jacobian that defines the transformation from heliocentric, spherical to Galactocentric, cartesian coordinates.
+where $\rtide$ is given by Eq.~\ref{eq:tidalradius}, and $\sigma_v$ by Eq.~\ref{eq:velscale}.
+$\left\vert\J\right\vert$ is the absolute value of the determinant of the Jacobian that defines the transformation from heliocentric, spherical to Galactocentric, cartesian coordinates.
 This model could trivially allow for mass loss by incorporating the initial satellite mass, $m$ and, e.g., a constant mass-loss term, $\dot{m}$.
+For a single star, in detail, the model works as follows:
+\begin{enumerate}
+	\item sample from the observed distributions in heliocentric coordinates (Eq.~\ref{eq:obsstar} and Eq.~\ref{eq:obsprog}) to get present-day positions for both the star and progenitor;
+	\item transform to Galactocentric, cartesian coordinates, and integrate both orbits backwards (as test particles) to $t=t_{\rm ub}$;
+	\item evaluate the likelihood.
+\end{enumerate}
 
@@
 For this article, we modify the above to be a mixture model: we add a new parameter for each star, $P_{\rm shock}$, which is the probability that the star is drawn from a broader, ``shocked'' distribution instead of a tight Gaussian centered on an effective Lagrange point:
+\begin{multline}
+	p_{\rm strip} = [ \mathcal{N}(\bs{r} \given \bs{r}_\sat - \Loffset\,\tailbit\,\rtide\,\hat{x}_1, \sigma^2_r)\\
+		\times \mathcal{N}(\bs{v} \given \bs{v}_\sat, \sigma^2_v) \left\vert\J\right\vert ]_{t=\tub}
+\end{multline}
 \begin{align}
-	p_{\rm strip} &= \left [ \mathcal{N}(\bs{r} \given \bs{r}_\sat - cK\rtide \hat{x}_1, \sigma_r)\mathcal{N}(\bs{v} \given \bs{v}_\sat, \sigma_v) \left\vert\J\right\vert\right]_{t=\tub}\\
-	p_{\rm shock} &= \left [ \mathcal{N}(\bs{r} \given \bs{r}_\sat, 3\sigma_r)\mathcal{N}(\bs{v} \given \bs{v}_\sat, 3\sigma_v) \left\vert\J\right\vert\right]_{t=\tub}\\
-	p(\W \given \W_\sat, \tub, \Phi, K, c, P_{\rm shock}) &= P_{\rm shock}p_{\rm shock} + (1-P_{\rm shock})p_{\rm strip}
+	p_{\rm shock} = \left [ \mathcal{N}(\bs{r} \given \bs{r}_\sat, 9\sigma^2_r)\mathcal{N}(\bs{v} \given \bs{v}_\sat, 9\sigma^2_v) \left\vert\J\right\vert\right]_{t=\tub}
 \end{align}
+\begin{multline}
+	p(\W \given \W_\sat, \tub, \Phi, \tailbit, c, P_{\rm shock}) = \\
+		P_{\rm shock}p_{\rm shock} + (1-P_{\rm shock})p_{\rm strip}
+\end{multline}
 The full posterior probability for the position of a star, progenitor, and properties of the parent potential, given the data, is then
-\begin{equation}
-	p(\Phi, \W, \W_\sat, \tub, K, c \given \D, \D_\sat) = \frac{1}{\mathcal{Z}} p(\D \given \W) p(\D_\sat \given \W_\sat) 
-												       p(\W \given \W_\sat, \tub, \Phi, K, c)
-												       p(\Phi) p(\tau)p(c)
-\end{equation}
+\begin{multline}
+	p(\Phi, \W, \W_\sat, \tub, \tailbit, c, P_{\rm shock} \given \D, \D_\sat) = \\
+		\shoveleft{\hspace{3em}\frac{1}{\mathcal{Z}} \, p(\D \given \W) \, p(\D_\sat \given \W_\sat)}\\
+			\times p(\W \given \W_\sat, \tub, \Phi, \Loffset, \tailbit) \, 
+			p(\Phi) \, p(\tub) \, p(c)
+\end{multline}
 where the factor $\mathcal{Z}$ only depends on the properties of the data (the \emph{evidence}), and is thus constant when varying the model parameters.
+All model parameters are summarized in Table~\ref{tbl:params}.
 
@@
 For all tests in this article, we model the potential of the parent galaxy with the same functional form used in the simulations (Equation~\ref{eq:lm10}).
-When recovering the potential, we hold fixed the disk and spheroid parameters (see Table~\ref{tbl:potential}), along with two halo parameters: the scale radius, $r_h$, and one of the flattening components in the x-y plane, $q_2$.
+When recovering the potential, we hold fixed the disk and spheroid parameters (see Table~\ref{tbl:params}), along with two halo parameters: the scale radius, $r_h$, and one of the flattening components in the x-y plane, $q_2$.
 These halo parameters are degenerate with combinations of the other parameters --- e.g., setting $q_1=2.0$, $q_2=1.0$, $\phi=0^\circ$ is equivalent to setting $q_1=1.0$, $q_2=2.0$, $\phi=90^\circ$.
@@
 \begin{center}
-	\begin{tabular}{l c c c} \toprule
-		{\bf Component} & {\bf Parameter} & {\bf Value} & {\bf Prior} \\\toprule
-		disk & $M_{\rm disk}$ & $1.0\times10^{11}M_\odot$ & -- \\ 
-		& $a$ & 6.5 kpc & --\\
-		& $b$ & 0.26 kpc & --\\
+	\begin{tabular}{l c l} \toprule
+		\multicolumn{3}{l}{{\bf \emph{Global parameters}}} \\
+		\toprule
+		Component & Parameter & Prior \\\toprule
+		disk & $M_{\rm disk}$ & $1.0\times10^{11}M_\odot$ (fixed) \\ 
+		& $a$ & 6.5 kpc (fixed)\\
+		& $b$ & 0.26 kpc (fixed)\\
 		\midrule
-		spheroid & $M_{\rm spher}$ & $3.4\times10^{10}M_\odot$ & --\\ 
-		& $c$ & 0.7 kpc & --\\
+		spheroid & $M_{\rm spher}$ & $3.4\times10^{10}M_\odot$ (fixed)\\ 
+		& $c$ & 0.7 kpc (fixed)\\
 		\midrule
-		halo & $v_h$ & -- & $\mathcal{U}(100,200)$ km $\mathrm{s}^{-1}$ \\
-		& $q_1$ & -- & $\mathcal{U}(1,2)$\\
-		& $q_2$ & 1.0 & --\\
-		& $q_z$ & -- & $\mathcal{U}(1,2)$\\
-		& $\phi$ & -- & $\mathcal{U}(52,142)$ deg\\
-		& $r_h$ & 12 kpc & --\\
+		halo & $v_h$ & $\mathcal{U}(100,200)$ km/s \\
+		& $q_1$ & $\mathcal{U}(1,2)$\\
+		& $q_2$ & 1.0 (fixed)\\
+		& $q_z$ & $\mathcal{U}(1,2)$\\
+		& $\phi$ & $\mathcal{U}(52,142)$ deg\\
+		& $r_h$ & 12 kpc (fixed)\\
+		\toprule
+		\multicolumn{3}{l}{{\bf \emph{Progenitor parameters}}} \\
+		\toprule
+		position & $\cos l$ & $\mathcal{U}(-1,1)$ \\
+		& $\cos b$ & $\mathcal{U}(-1,1)$\\
+		& $d$ & $\mathcal{U}(10,100)$~kpc\\
+		velocity & $\mu_l$ & $\mathcal{U}(-100,100)$~mas/yr\\
+		& $\mu_b$ & $\mathcal{U}(-100,100)$~mas/yr\\
+		& $v_r$ & $\mathcal{U}(-1000,1000)$~km/s\\
+		Lagrange pt. offset & $\alpha$ & $\mathcal{N}(0.5,5)$\\
+		tail assignment & $\beta$ & $\{-1, 1\}$\\ % TODO: how to say equal weight to -1 or 1?
+		\toprule
+		\multicolumn{3}{l}{{\bf \emph{Parameters per star}}} \\
+		\toprule
+		position & $l$ & fixed at observed val.\\
+		& $b$ & fixed at observed val.\\
+		& $d$ & $\mathcal{U}(10,100)$~kpc\\
+		velocity & $\mu_l$ & $\mathcal{U}(-100,100)$~mas/yr\\
+		& $\mu_b$ & $\mathcal{U}(-100,100)$~mas/yr\\
+		& $v_r$ & $\mathcal{U}(-1000,1000)$~km/s\\
+		prob. shocked & $P_{\rm shocked}$ & $\mathcal{U}(0,1)$\\
+		time unbound & $\tub$ & $\mathcal{U}(0,\tint)$\\ % TODO: talk about \tint (integration time) in text!
 		\bottomrule
-		\hline
 		\end{tabular}
-	\caption{Parameter values used ...\label{tbl:potential}}
+	\caption{Parameter values used ...\label{tbl:params}}
 \end{center}

### 2014-03-11 text/paper2/ms.tex
minor notes

@@
 	\item sample from the observed distributions in heliocentric coordinates (Eq.~\ref{eq:obsstar} and Eq.~\ref{eq:obsprog}) to get present-day positions for both the star and progenitor;
-	\item transform to Galactocentric, cartesian coordinates, and integrate both orbits backwards (as test particles) to $t=t_{\rm ub}$;
-	\item evaluate the likelihood.
+	\item transform to Galactocentric, cartesian coordinates, and integrate both orbits backwards (as test particles) for the total interaction time, $\tint$;
+	\item evaluate the likelihood at $t=\tub$.
 \end{enumerate}
@@
 \begin{align}
-	p_{\rm shock} = \left [ \mathcal{N}(\bs{r} \given \bs{r}_\sat, 9\sigma^2_r)\mathcal{N}(\bs{v} \given \bs{v}_\sat, 9\sigma^2_v) \left\vert\J\right\vert\right]_{t=\tub}
+	p_{\rm shock} = \left [ \mathcal{N}(\bs{r} \given \bs{r}_\sat, 10\sigma^2_r)\mathcal{N}(\bs{v} \given \bs{v}_\sat, 10\sigma^2_v) \left\vert\J\right\vert\right]_{t=\tub}
 \end{align}
@@
 	p(\W \given \W_\sat, \tub, \Phi, \tailbit, c, P_{\rm shock}) = \\
-		P_{\rm shock}p_{\rm shock} + (1-P_{\rm shock})p_{\rm strip}
+		P_{\rm shock}p_{\rm shock} + (1-P_{\rm shock})p_{\rm strip}.
 \end{multline}
+Here we assume that the ``shocked'' distribution has a much larger variance, and arbitrarily set this to 10 times the variance of the stripped debris.
 The full posterior probability for the position of a star, progenitor, and properties of the parent potential, given the data, is then
@@
 \subsection{Perfect data}
+
+We first
 

### 2014-03-11 scripts/infer_potential.py
added mdot and mass to labels

@@ -48,7 +48,8 @@ _label_map['d'] = "$d$ [kpc]"
 _label_map['mul'] = r"$\mu_l$ [mas/yr]"
 _label_map['mub'] = r"$\mu_b$ [mas/yr]"
 _label_map['vr'] = r"$v_r$ [km/s]"
-_label_map['logmass'] = r"$\log M$"
+_label_map['logmass'] = r"$\log m$ [$M_\odot$]"
+_label_map['logmdot'] = r"$\log \dot{m}$ [$M_\odot/$Myr]"
 
 _unit_transform = defaultdict(lambda: lambda x: x)
 _unit_transform['phi'] = lambda x: (x*u.rad).to(u.degree).value

### 2014-03-11 scripts/test_infer_potential.py
test params

@@ -66,7 +66,7 @@ ptc_params = """
 #    parameters: [d, mul, mub, vr]
 
 sat_params = """
-    parameters: [c]
+    parameters: [c, d, mul, mub, vr]
 """
 #    parameters: [logmass, logmdot, d, mul, mub, vr]
 
@@ -77,7 +77,7 @@ lm10_c = minimum_config.format(potential_params=pot_params,
 # _config = minimum_config.format(potential_params=pot_params,
 #                                 particles_params=ptc_params,
 #                                 satellite_params=sat_params)
-_config = minimum_config.format(potential_params="",
+_config = minimum_config.format(potential_params=pot_params,
                                 particles_params="",
                                 satellite_params=sat_params)
 

### 2014-03-12 text/paper2/ms.tex
added note about sims

@@
 In what follows, we apply the model described above to ``observations'' of the n-body simulations described in Section~\ref{sec:sims}.
-For all tests in this article, we model the potential of the parent galaxy with the same functional form used in the simulations (Equation~\ref{eq:lm10}).
+For all tests in this article, we model the potential of the parent galaxy with the same functional form used in the simulations (Equation~\ref{eq:lm10}) --- we stress that the simulations are n-body simulations and not generated with the same generative model presented here.
 When recovering the potential, we hold fixed the disk and spheroid parameters (see Table~\ref{tbl:params}), along with two halo parameters: the scale radius, $r_h$, and one of the flattening components in the x-y plane, $q_2$.
@@
 
-We first
+We first test the model on small samples of stars sampled from the simulations above assuming perfect observational data.
 
-\subsection{\gaia + \spitzer}
+\subsection{\gaia\,+ \spitzer}
 8 stars in Sgr-like stream

### 2014-03-13 text/paper2/ms.tex
updates

@@
 \newcommand{\Loffset}{\alpha}
-
+\newcommand{\pshock}{P_{\rm shock}}
 \newcommand{\paperone}{Paper 1}
+\newcommand{\vhalo}{v_{\rm h}}
 
@@
 	&\Phi_{\rm spher}(r) = -\frac{GM_{\rm spher}}{r + c}\\
-	&\Phi_{\rm halo}(x,y,z) = v_{h}^2 \ln(C_1 x^2 + C_2 y^2 + C_3 xy + (z/q_z)^2 + r_h^2)
+	&\Phi_{\rm halo}(x,y,z) = \vhalo^2 \ln(C_1 x^2 + C_2 y^2 + C_3 xy + (z/q_z)^2 + r_h^2)
 \end{align}
@@
 \caption{ Equipotential contours for the LM10 potential (Eq.~\ref{eq:lm10}) in Galactocentric, cartesian coordinates for various halo parameter choices.
-For all panels, $v_h=121.858~\mathrm{km}/\mathrm{s}$, $r_h=12~\mathrm{kpc}$.
+For all panels, $\vhalo=121.858~\mathrm{km}/\mathrm{s}$, $r_h=12~\mathrm{kpc}$.
 Left to right, each column represents a new choice of parameters.
@@
 \subsection{Probabilistic model}
+The above scalings motivate a model for tidal debris in which stars were ``released'' at the instantaneous, effective Lagrange points with some dispersion in position and velocity, all of which depend only on the mass of the progenitor, the parent potential properties, and the orbit of the progenitor.
+We present this model in detail below.
+
 We observe the 6D position of a star, $\D$, in heliocentric coordinates --- e.g., the measured position on the sky, ($l$, $b$); distance, $d$; proper motions, ($\mu_l$, $\mu_b$); and line-of-sight velocity, $v_r$ --- and have determined through some other means that this star was once part of a progenitor system (e.g., satellite galaxy) with mass $m(t)$ that is disrupting and forming a cold debris structure in the potential, $\Phi$, of the parent galaxy (e.g., the Milky Way).
@@
 In cartesian coordinates, the present-day position of the star is taken as having been sampled at $t=\tub$ from an isotropic Gaussian centered on one of the instantaneous, effective Lagrange points in position, and an isotropic Gaussian centered on the origin in velocity, and evolved to present day in the parent potential, $\Phi$.
-That is, we assume that once the star becomes ``unbound'' from the satellite the potential of the satellite can be ignored, and thus treat the star as a test particle.\footnote{This assumption breaks down for sufficiently large-mass progenitors, somewhere between $\sim10^8-10^9~$M$_\odot$.} We add two other parameters: a constant offset, $\Loffset$, that specifies the position of the effective Lagrange points relative to the classical tidal radius, and a binary parameter $\tailbit$, that is either $-1$ or $+1$ depending on whether the star is in the leading or trailing tail.
+That is, we assume that once the star becomes ``unbound'' from the satellite the potential of the satellite can be ignored, and thus treat the star as a test particle.\footnote{This assumption breaks down for sufficiently large-mass progenitors, somewhere between $\sim10^8-10^9~$M$_\odot$.} The satellite mass does enter the model through the tidal radius; a changing satellite mass will skew the position of the effective Lagrange points.
+We allow for mass-loss by incorporating the initial satellite mass, $m$, and a constant mass-loss term, $\dot{m}$ into the model.
+Finally, we add two additional parameters: a constant, global offset, $\Loffset$, that scales the position of the effective Lagrange points relative to the classical tidal radius, and a binary parameter $\tailbit$, that is either $-1$ or $+1$ depending on whether the star is in the leading or trailing tail.
 \begin{align}
-	p(\W \given \W_\sat, \tub, \Phi, \Loffset, \tailbit) &= p(\bs{r} \given \bs{r}_\sat, \tub, \Phi, \Loffset, \tailbit)\,p(\bs{v} \given \bs{v}_\sat, \tub, \Phi)\,\left\vert\J\right\vert_{t=\tub}\\
-	p(\bs{r} \given \bs{r}_\sat, \tub, \Phi, \Loffset, \tailbit) &= \mathcal{N}(\bs{r} \given \bs{r}_\sat - \Loffset\,\tailbit\,\rtide\,\hat{x}_1, \sigma^2_r)_{t=\tub}\\
-	p(\bs{v} \given \bs{v}_\sat, \tub, \Phi) &= \mathcal{N}(\bs{v} \given \bs{v}_\sat, \sigma^2_v)_{t=\tub}
+	p(\W \given m, \dot{m}, \W_\sat, \tub, \Phi, \Loffset, \tailbit) &= p(\bs{r} \given \bs{r}_\sat, m, \dot{m}, \tub, \Phi, \Loffset, \tailbit)\,p(\bs{v} \given \bs{v}_\sat, m, \dot{m}, \tub, \Phi)\,\left\vert\J\right\vert_{t=\tub}\\
+	p(\bs{r} \given \bs{r}_\sat, m, \dot{m}, \tub, \Phi, \Loffset, \tailbit) &= \mathcal{N}(\bs{r} \given \bs{r}_\sat - \Loffset\,\tailbit\,\rtide\,\hat{x}_1, \sigma^2_r)_{t=\tub}\\
+	p(\bs{v} \given \bs{v}_\sat, m, \dot{m}, \tub, \Phi) &= \mathcal{N}(\bs{v} \given \bs{v}_\sat, \sigma^2_v)_{t=\tub}
 \end{align}
@@
 $\left\vert\J\right\vert$ is the absolute value of the determinant of the Jacobian that defines the transformation from heliocentric, spherical to Galactocentric, cartesian coordinates.
-This model could trivially allow for mass loss by incorporating the initial satellite mass, $m$ and, e.g., a constant mass-loss term, $\dot{m}$.
 For a single star, in detail, the model works as follows:
@@
 We will explore properly modeling this process in future work.
-For this article, we modify the above to be a mixture model: we add a new parameter for each star, $P_{\rm shock}$, which is the probability that the star is drawn from a broader, ``shocked'' distribution instead of a tight Gaussian centered on an effective Lagrange point:
+For this article, we modify the above to be a mixture model: we add a new parameter for each star, $\pshock$, which is the probability that the star is drawn from a broader, ``shocked'' distribution instead of a tight Gaussian centered on an effective Lagrange point:
 \begin{multline}
@@
 \begin{multline}
-	p(\W \given \W_\sat, \tub, \Phi, \tailbit, c, P_{\rm shock}) = \\
-		P_{\rm shock}p_{\rm shock} + (1-P_{\rm shock})p_{\rm strip}.
+	p(\W \given \W_\sat, \tub, \Phi, \tailbit, c, \pshock) = \\
+		\pshock p_{\rm shock} + (1-\pshock)p_{\rm strip}.
 \end{multline}
 Here we assume that the ``shocked'' distribution has a much larger variance, and arbitrarily set this to 10 times the variance of the stripped debris.
-The full posterior probability for the position of a star, progenitor, and properties of the parent potential, given the data, is then
+The posterior probability for the position of a star, progenitor, and properties of the parent potential, given the data, is then
 \begin{multline}
-	p(\Phi, \W, \W_\sat, \tub, \tailbit, c, P_{\rm shock} \given \D, \D_\sat) = \\
+	p(\Phi, \W, \W_\sat, \tub, \tailbit, c, \pshock \given \D, \D_\sat) = \\
 		\shoveleft{\hspace{3em}\frac{1}{\mathcal{Z}} \, p(\D \given \W) \, p(\D_\sat \given \W_\sat)}\\
-			\times p(\W \given \W_\sat, \tub, \Phi, \Loffset, \tailbit) \, 
+			\times p(\W \given \W_\sat, \tub, \Phi, \Loffset, \tailbit, \pshock) \, 
 			p(\Phi) \, p(\tub) \, p(c)
@@
 All model parameters are summarized in Table~\ref{tbl:params}.
-
-% In general, $p(\tub)$ is likely conditional on the orbit of the progenitor --- stars are preferentially stripped at pericentric passages --- but this is neglected and is taken to be uniform over the entire interaction history between the progenitor and parent.
+Assuming each star is an independent tracer, the full posterior for many stars is then just the product of the individual likelihoods with the priors:
+\begin{multline}
+	p(\Phi, \W, \W_\sat, \tub, \tailbit, c, \pshock \given \D, \D_\sat) = \\
+		\shoveleft{\hspace{1em}\frac{1}{\mathcal{Z}} \, \prod_i \, p(\D^{(i)} \given \W^{(i)}) \, p(\D_\sat \given \W_\sat)}\\
+			\times p(\W^{(i)} \given \W_\sat, \tub^{(i)}, \Phi, \Loffset, \tailbit^{(i)}, \pshock^{(i)}) \, 
+			\, p(\tub^{(i)}) p(\Phi) \, p(c)
+\end{multline}
 
@@
 These halo parameters are degenerate with combinations of the other parameters --- e.g., setting $q_1=2.0$, $q_2=1.0$, $\phi=0^\circ$ is equivalent to setting $q_1=1.0$, $q_2=2.0$, $\phi=90^\circ$.
-The priors on the remaining halo parameters are taken to be uniform over a conservative domain of realistic values: for $v_h$, 100-200~km/s corresponds to a range in solar circular velocities from $\sim$210-250 km/s (holding other parameters fixed); the range in axis ratios allow for prolate, oblate, and generic triaxiality; and $\phi$ is restricted to $\pm45^\circ$ around the value measured by LM10 ($\phi = 97^\circ$).
+The priors on the remaining halo parameters are taken to be uniform over a conservative domain of realistic values: for $\vhalo$, 100-200~km/s corresponds to a range in solar circular velocities from $\sim$210-250 km/s (holding other parameters fixed); the range in axis ratios allow for prolate, oblate, and generic triaxiality; and $\phi$ is restricted to $\pm45^\circ$ around the value measured by LM10 ($\phi = 97^\circ$).
 
@@
 		\midrule
-		halo & $v_h$ & $\mathcal{U}(100,200)$ km/s \\
+		halo & $\vhalo$ & $\mathcal{U}(100,200)$ km/s \\
 		& $q_1$ & $\mathcal{U}(1,2)$\\
@@
 		\toprule
-		position & $\cos l$ & $\mathcal{U}(-1,1)$ \\
-		& $\cos b$ & $\mathcal{U}(-1,1)$\\
-		& $d$ & $\mathcal{U}(10,100)$~kpc\\
-		velocity & $\mu_l$ & $\mathcal{U}(-100,100)$~mas/yr\\
-		& $\mu_b$ & $\mathcal{U}(-100,100)$~mas/yr\\
-		& $v_r$ & $\mathcal{U}(-1000,1000)$~km/s\\
-		Lagrange pt. offset & $\alpha$ & $\mathcal{N}(0.5,5)$\\
-		tail assignment & $\beta$ & $\{-1, 1\}$\\ % TODO: how to say equal weight to -1 or 1?
+		position & $\bs{r}_\sat$ & $\|\bs{r}_\sat\|\sim\mathcal{U}(0,200)$~kpc \\
+		velocity & $\bs{v}_\sat$ & $\|\bs{v}_\sat\|\sim\mathcal{N}(\mu=0,\sigma^2=4000)$~km/s\\
+		Lagrange pt. offset & $\alpha$ & $\mathcal{N}(\mu=2,\sigma^2=1)$\\
 		\toprule
@@
 		\toprule
-		position & $l$ & fixed at observed val.\\
-		& $b$ & fixed at observed val.\\
-		& $d$ & $\mathcal{U}(10,100)$~kpc\\
-		velocity & $\mu_l$ & $\mathcal{U}(-100,100)$~mas/yr\\
-		& $\mu_b$ & $\mathcal{U}(-100,100)$~mas/yr\\
-		& $v_r$ & $\mathcal{U}(-1000,1000)$~km/s\\
-		prob. shocked & $P_{\rm shocked}$ & $\mathcal{U}(0,1)$\\
+		position & $\bs{r}$ & $\|\bs{r}\|\sim\mathcal{U}(0,200)$~kpc \\
+		velocity & $\bs{v}$ & $\|\bs{v}\|\sim\mathcal{N}(\mu=0,\sigma^2=4000)$~km/s\\
+		tail assignment & $\beta$ & $\{-1, 1\}$\\ % TODO: how to say equal weight to -1 or 1?
+		prob. shocked & $\pshock$ & $\mathcal{U}(0,1)$\\
 		time unbound & $\tub$ & $\mathcal{U}(0,\tint)$\\ % TODO: talk about \tint (integration time) in text!
@@
 
-We first test the model on small samples of stars sampled from the simulations above assuming perfect observational data.
+We first test the model using eight stars sampled from the $2.5\times10^8~M_\odot$ simulation (Section~\ref{sec:sims}) --- four from the leading trail, and four from the trailing tail --- assuming perfect observational data for both the stars and the progenitor.
+The stars are additionally required to be within $50~kpc$ of the Sun.
+Figure~\ref{fig:} (left panel) shows a sample of stream particles (black) and the eight randomly chosen stars (blue) in the Galactic X-Z plane.
+We leave the potential parameters --- $q_1$, $q_z$, $\phi$, and $\vhalo$ --- free to vary, along with the Lagrange point offset, $\Loffset$, the tail assignment for each star, $\tailbit^{(i)}$, and probability that each star was shocked, $\pshock^{(i)}$.
+Figure~\ref{fig:} shows the posterior distribution for the four potential parameters, marginalized over the other parameters.
 
-\subsection{\gaia\,+ \spitzer}
-8 stars in Sgr-like stream
+\subsection{\spitzer + \gaia}
+We now take the same eight stars used in the previous experiment and ``observe'' them with optimistic observational uncertainties.
+We take these stars to be RR Lyrae variables which are known to be excellent distance indicators using the mid-infrared period-luminosity relation \citep[as shown in, e.g.,][]{madore12}; $\sim$100 RR Lyrae associated with the Sgr stream and $\sim$30 associated with the Orphan stream will be observed with \spitzer as part of the SMASH survey \citep{smashprop} with expected fractional distance uncertainties around $\sim$2\%.
+These stars will also be included in the \gaia\, proper motion catalog: at a distance of $\sim$50~kpc, a typical RR Lyrae will have a tangential velocity error around $\sim$20~km/s.
 
-We assume the present-day mass of the progenitor, $M_p$, and velocity dispersion, $\sigma_v$, have been measured perfectly and neglect mass-loss from the progenitor.
+For this experiment, we (1) assume the stars are bright distance indicators --- RR Lyrae --- such that the fractional distance uncertainty is $2\%$; (2) neglect the uncertainty in angular position (for a typical RR Lyrae at 50~kpc this is $\sim$$10^{-7}$~deg); (3) assume we can measure radial velocities to these stars with 5~km/s uncertainty; and (4) compute the sky-averaged \gaia\, proper-motion uncertainty for each star assuming an F0V spectral type using the \texttt{PyGaia}\footnote{https://github.com/agabrown/PyGaia} code and use this uncertainty for both components of proper motion.
+% TODO: acknowledge Gaia science team, Anthony Brown, Gaia data challenge?
+The observed positions and velocities of these stars are shown in Figure~\ref{fig:} (red, compare to the unobserved black points).
 
@@
 
+[Different streams will constraint different properties of the potential -- multiple streams are crucial.
+Here we use a stream roughly in the X-Z plane, thus we expect it to be more sensitive to flattening in Z direction]
+
 [Don't eschew low-dimensional constraints on the potential!]

### 2014-03-15 text/paper2/ms.tex
printed edits

@@
 
-For a satellite galaxy on a mildly eccentric orbit, mass-loss is driven by a combination of tidal stripping due to the tidal field of the parent system and tidal shocking due to the rapidly changing tidal field as the progenitor system moves through pericenter.
+For a satellite galaxy on a mildly eccentric orbit, mass-loss is driven by a combination of tidal stripping due to the steady tidal field of the parent system and tidal shocking due to the rapidly changing tidal field as the progenitor system moves through pericenter.
 The interplay between these two processes as a function of orbital properties will be studied in forthcoming work; we focus below on the slow disruption of stars due to steady tidal forcing.
 
-Consider a point-mass satellite of mass $m$ on a circular orbit with frequency $\Omega$ around a point-mass of mass $M\gg m$ \citep[the restricted three-body problem; e.g., \S 8.3][]{binneytremaine}.
+Consider a point-mass satellite of mass $m$ on a circular orbit with frequency $\Omega$ around a more massive point-mass, $M\gg m$ \citep[the restricted three-body problem; e.g., \S 8.3][]{binneytremaine}.
 In a frame rotating with the orbital frequency of the satellite, the effective potential around the satellite has (amongst others) two stationary, unstable points located around $\sim R \pm r_J$, where $R$ is the distance to the satellite and $r_J$ is the Jacobi or tidal radius,
@@
 \end{equation}
-Particles that would be bound to the satellite system in isolation may have enough (Jacobi) energy to overcome the effective potential barrier at these Lagrange points and thus will be preferentially stripped from the satellite at these points (Fig.~8.6 of BT).
-For a spherical, extended parent mass distribution the tidal radius instead scales with the enclosed mass of the parent system at a given orbital position, $R$, with additional terms that account for the local slope of the density profile.
-Thus, in a more general parent potential the effective Lagrange points may not be symmetric about the center of the satellite and may deviate from the point-mass tidal radius (Eq.~\ref{eq:ptmass}) by factors of order unity.
-We thus simply adopt
+Particles that would be bound to the satellite system in isolation may have enough (Jacobi) energy to overcome the effective potential barrier at these Lagrange points (Fig.~8.6 of BT) and thus will be preferentially stripped from the satellite at these points.
+For a spherical, extended parent mass distribution the tidal radius instead scales with the enclosed mass at a given orbital position, $R$, with additional terms that account for the local slope of the density profile.
+In general, the effective Lagrange points may not be symmetric about the center of the satellite and may deviate from the point-mass tidal radius (Eq.~\ref{eq:ptmass}) by factors of order unity.
+We thus adopt
 \begin{equation}
@@
 \end{equation}
-as a normalizing distance scale, where $M_{\rm enc}$ is the instantaneous enclosed mass of the parent system within orbital radius $R$.
+as a distance scale for the debris, where $M_{\rm enc}$ is the instantaneous enclosed mass of the parent system within orbital radius $R$.
 At pericenter, the velocity of the satellite will be of order $V\sim \sqrt{GM_{\rm enc}/R}$.
@@
 \end{equation}
-Thus, we expect the particle velocities to scale with $(m/M_{\rm enc})^{1/3}$.
+We expect the debris star velocities to also scale with $(m/M_{\rm enc})^{1/3}$.
 
@@
 %(Section~\ref{sec:sims})
-Figure~\ref{fig:lpts_r} shows sections of particle orbits in this coordinate system, symmetric around the unbinding time, $\tub$, as a function of progenitor mass.
+Figure~\ref{fig:lpts_r} shows sections of particle orbits from the n-body simulations described above projected into this coordinate system, symmetric around their unbinding times, $\tub$, for the four progenitor masses.
 In these coordinates, the classical Lagrange points would be located around $x_1=\pm1,x_2=0,x_3=0$ (illustrated by the unit circle in top panels of Figure~\ref{fig:lpts_r}).
-Figure~\ref{fig:lpts_v} shows the same for the velocity of each star.
+Figure~\ref{fig:lpts_v} shows the velocity of each star also projected into these coordinates and normalized.
 Though not centered on the point-mass Lagrange points, the location of and dispersion about the effective Lagrange points in these scaled coordinates remain remarkably consistent across the range of progenitor masses explored.
-Furthermore, the dispersion in velocity is well-normalized by the velocity scale of Eq.~\ref{eq:velscale}, though it is clear that the velocity dispersion along $\hat{x}_1$, the radial vector to the satellite position, is larger than that in other dimensions, possibly due to mild tidal shocking.
+Furthermore, the dispersion in velocity is well-normalized by the velocity scale of Eq.~\ref{eq:velscale}, though it is clear that the velocity dispersion along $\hat{x}_1$, the radial vector to the satellite position, is larger than that in other dimensions, possibly due to tidal shocking.
+
 Even in a triaxial potential, the phase-space dispersion of tidally stripped debris scales with the mass ratio $(m / M_{\rm enc})^{1/3}$.
-%, the orbital radius, and the orbital velocity of the progenitor system.
+This  motivates a model for tidal debris in which stars were ``released'' at the instantaneous, effective Lagrange points with some dispersion in position and velocity, all of which depend only on the mass and orbit of the progenitor, and the parent potential.
+We present this model in detail below.
 
@@
 \subsection{Probabilistic model}
-The above scalings motivate a model for tidal debris in which stars were ``released'' at the instantaneous, effective Lagrange points with some dispersion in position and velocity, all of which depend only on the mass of the progenitor, the parent potential properties, and the orbit of the progenitor.
-We present this model in detail below.
-
 We observe the 6D position of a star, $\D$, in heliocentric coordinates --- e.g., the measured position on the sky, ($l$, $b$); distance, $d$; proper motions, ($\mu_l$, $\mu_b$); and line-of-sight velocity, $v_r$ --- and have determined through some other means that this star was once part of a progenitor system (e.g., satellite galaxy) with mass $m(t)$ that is disrupting and forming a cold debris structure in the potential, $\Phi$, of the parent galaxy (e.g., the Milky Way).
@@
 In cartesian coordinates, the present-day position of the star is taken as having been sampled at $t=\tub$ from an isotropic Gaussian centered on one of the instantaneous, effective Lagrange points in position, and an isotropic Gaussian centered on the origin in velocity, and evolved to present day in the parent potential, $\Phi$.
-That is, we assume that once the star becomes ``unbound'' from the satellite the potential of the satellite can be ignored, and thus treat the star as a test particle.\footnote{This assumption breaks down for sufficiently large-mass progenitors, somewhere between $\sim10^8-10^9~$M$_\odot$.} The satellite mass does enter the model through the tidal radius; a changing satellite mass will skew the position of the effective Lagrange points.
+We assume that once the star becomes ``unbound'' from the satellite the potential of the satellite can be ignored, and thus treat the star as a test particle.\footnote{This assumption breaks down for sufficiently large-mass progenitors, somewhere between $\sim10^8-10^9~$M$_\odot$.} The satellite mass enters the model through the tidal radius; a changing satellite mass will skew the positions of the effective Lagrange points and velocity scale.
 We allow for mass-loss by incorporating the initial satellite mass, $m$, and a constant mass-loss term, $\dot{m}$ into the model.
@@
 \begin{multline}
-	p(\W \given \W_\sat, \tub, \Phi, \tailbit, c, \pshock) = \\
+	p(\W \given \W_\sat, \tub, \Phi, \Loffset, \tailbit, \pshock) = \\
 		\pshock p_{\rm shock} + (1-\pshock)p_{\rm strip}.
@@
 \begin{multline}
-	p(\Phi, \W, \W_\sat, \tub, \tailbit, c, \pshock \given \D, \D_\sat) = \\
+	p(\Phi, \W, \W_\sat, \tub, \Loffset, \tailbit, \pshock \given \D, \D_\sat) = \\
 		\shoveleft{\hspace{3em}\frac{1}{\mathcal{Z}} \, p(\D \given \W) \, p(\D_\sat \given \W_\sat)}\\
 			\times p(\W \given \W_\sat, \tub, \Phi, \Loffset, \tailbit, \pshock) \, 
-			p(\Phi) \, p(\tub) \, p(c)
+			p(\Phi) \, p(\tub) \, p(\Loffset)
 \end{multline}
@@
 \begin{multline}
-	p(\Phi, \W, \W_\sat, \tub, \tailbit, c, \pshock \given \D, \D_\sat) = \\
+	p(\Phi, \W, \W_\sat, \tub, \Loffset, \tailbit, \pshock \given \D, \D_\sat) = \\
 		\shoveleft{\hspace{1em}\frac{1}{\mathcal{Z}} \, \prod_i \, p(\D^{(i)} \given \W^{(i)}) \, p(\D_\sat \given \W_\sat)}\\
 			\times p(\W^{(i)} \given \W_\sat, \tub^{(i)}, \Phi, \Loffset, \tailbit^{(i)}, \pshock^{(i)}) \, 
-			\, p(\tub^{(i)}) p(\Phi) \, p(c)
+			\, p(\tub^{(i)}) p(\Phi) \, p(\Loffset)
 \end{multline}

### 2014-03-15 text/paper2/ms.tex
more minor changes to text

@@
 
-We first test the model using eight stars sampled from the $2.5\times10^8~M_\odot$ simulation (Section~\ref{sec:sims}) --- four from the leading trail, and four from the trailing tail --- assuming perfect observational data for both the stars and the progenitor.
-The stars are additionally required to be within $50~kpc$ of the Sun.
+We first test the model using eight stars --- four from the leading trail, and four from the trailing tail --- sampled from the $2.5\times10^8~M_\odot$ simulation (Section~\ref{sec:sims}) assuming perfect observational data for both the stars and the progenitor.
+The stars were required to be within $50$ kpc of the Sun.
 Figure~\ref{fig:} (left panel) shows a sample of stream particles (black) and the eight randomly chosen stars (blue) in the Galactic X-Z plane.
@@
 Figure~\ref{fig:} shows the posterior distribution for the four potential parameters, marginalized over the other parameters.
+
+[Interpret the posterior]
 

### 2014-03-15 scripts/observe_simulation.py
only take very nearby stars

@@ -81,7 +81,7 @@ def observe_simulation(class_name, particle_error_model=None, satellite_error_mo
     selection_expr = "(tub!=0) & (tub>{}) & (tub<{}) & (sqrt((x+8)**2 + y**2 + z**2)<{})"\
                      .format((1800*u.Myr).to(sim_time).value,
                              (5500*u.Myr).to(sim_time).value,
-                             (50*u.kpc).to(sim_leng).value)
+                             (35*u.kpc).to(sim_leng).value)
     particles = simulation.particles(N=N, expr=selection_expr, tail_bit=True)
 
     logger.debug("Read in {} particles with expr='{}'"\

### 2014-03-15 scripts/test_infer_potential.py
tests

@@ -37,13 +37,9 @@ logging.basicConfig(level=logging.DEBUG)
 
 minimum_config = """
 name: test
-data_file: data/observed_particles/2.5e8_N1024.hdf5
+data_file: data/observed_particles/2.5e8_N1024_DH.hdf5
 nparticles: 8
-particle_idx: [1019, 606, 740, 301, 261, 39, 735, 448]
-# nparticles: 4
-# particle_idx: [1019, 606, 740, 301]
-# nparticles: 16
-# nparticles: 128
+particle_idx: [255, 241, 447, 339, 213, 183, 643, 530]
 
 potential:
     class_name: LawMajewski2010
@@ -61,12 +57,12 @@ pot_params = """
 """
 
 ptc_params = """
-    parameters: [p_shocked]
+    parameters: [p_shocked, beta, d]
 """
 #    parameters: [d, mul, mub, vr]
 
 sat_params = """
-    parameters: [c, d, mul, mub, vr]
+    parameters: [d, mul, mub, vr]
 """
 #    parameters: [logmass, logmdot, d, mul, mub, vr]
 
@@ -79,7 +75,7 @@ lm10_c = minimum_config.format(potential_params=pot_params,
 #                                 satellite_params=sat_params)
 _config = minimum_config.format(potential_params=pot_params,
                                 particles_params="",
-                                satellite_params=sat_params)
+                                satellite_params="")
 
 # particles_params=ptc_params,
 # satellite_params=sat_params
@@ -208,6 +204,21 @@ class TestStreamModel(object):
                             plt.close('all')
                             idx += 1
 
+                    elif param_name == 'beta':
+                        for jj in range(param.value.shape[0]):
+                            vals1 = np.linspace(param._prior.a[jj],
+                                                param._prior.b[jj],
+                                                Ncoarse)
+                            vals2 = np.linspace(param._prior.a[jj],
+                                                param._prior.b[jj],
+                                                Nfine)
+
+                            fig = make_plot(model, idx, vals1, vals2)
+                            fig.savefig(os.path.join(test_path,
+                                    "ptcl{}_{}.png".format(idx,param_name)))
+                            plt.close('all')
+                            idx += 1
+
                     elif param_name == 'tub':
                         for jj in range(param.value.shape[0]):
                             vals1 = np.linspace(param._prior.a[jj],
@@ -241,26 +252,6 @@ class TestStreamModel(object):
                             plt.close('all')
                             idx += 1
 
-                    elif param_name == "fac_R":
[…]

### 2014-03-15 text/paper2/ms.tex
describe emcee add refs

--- 2014-03-15 text/paper2/ms.tex
@@
 
-We first test the model using eight stars sampled from the $2.5\times10^8~M_\odot$ simulation (Section~\ref{sec:sims}) --- four from the leading trail, and four from the trailing tail --- assuming perfect observational data for both the stars and the progenitor.
-The stars are additionally required to be within $50~kpc$ of the Sun.
+We first test the model using eight particles (stars) --- four from the leading trail, and four from the trailing tail --- randomly sampled from the $2.5\times10^8~M_\odot$ simulation (Section~\ref{sec:sims}), assuming perfect observational data for both the stars and the progenitor.
+The stars were required to have been stripped after the first pericentric passage and have a present-day distance within $50$ kpc of the Sun.
 Figure~\ref{fig:} (left panel) shows a sample of stream particles (black) and the eight randomly chosen stars (blue) in the Galactic X-Z plane.
-We leave the potential parameters --- $q_1$, $q_z$, $\phi$, and $\vhalo$ --- free to vary, along with the Lagrange point offset, $\Loffset$, the tail assignment for each star, $\tailbit^{(i)}$, and probability that each star was shocked, $\pshock^{(i)}$.
-Figure~\ref{fig:} shows the posterior distribution for the four potential parameters, marginalized over the other parameters.
+
+We leave the potential parameters --- $q_1$, $q_z$, $\phi$, and $\vhalo$ --- free to vary, along with the Lagrange point offset, $\Loffset$, and use Markov Chain Monte Carlo (MCMC) to sample from the posterior probability distribution over these five parameters.
+In detail, we use an affine-invariant, ensemble sampler \citep{goodman10} implemented in the Python programming language \citep{foremanmackey13} with an ensemble of 64 walkers initialized by sampling from the priors summarized in Table~\ref{tbl:params}.
+We run the walkers in parallel (using MPI todo) for 5000 steps to burn-in, then restart the sampling starting from the final position of the burn-in phase and run for another 10000 steps.
+We compute the autocorrelation time for each parameter [todo: reference] and thin the chains by taking every XXX$\times \mathrm{median}(t_{\rm acor})$ sample to ensure they are close to independent.
+Figure~\ref{fig:} shows the walker positions over the 10000 post-burn-in steps for each parameter for the $2.5\times10^8~M_\odot$ simulation; the autocorrelation time for each parameter is displayed on each corresponding panel.
+
+Figures~\ref{fig:}-\ref{fig:} show the posterior distributions for these five parameters for each of the the four mass progenitors.
 
@@
 4 stars in Orphan-like stream, XX stars in GD-1 -like stream
+
+\subsection{Mass comparison}
+[...]
 
@@
 \acknowledgements
-APW: NSF GRFP, Sigma Xi
+Aspen?\\
+APW: NSF GRFP, Sigma Xi\\
+Columbia HPC\\
 

### 2014-03-15 text/paper2/ms.tex
hide pshock stuff

@@
 Finally, we add two additional parameters: a constant, global offset, $\Loffset$, that scales the position of the effective Lagrange points relative to the classical tidal radius, and a binary parameter $\tailbit$, that is either $-1$ or $+1$ depending on whether the star is in the leading or trailing tail.
+The likelihood is then:
 \begin{align}
@@
 \end{enumerate}
-
+All model parameters are summarized in Table~\ref{tbl:params}.
+Assuming each star is an independent tracer, the full posterior for many stars is then just the product of the individual likelihoods with the priors:
+\begin{multline}
+	p(\Phi, \W, \W_\sat, \tub, \Loffset, \tailbit, \given \D, \D_\sat) = \\
+		\shoveleft{\hspace{1em}\frac{1}{\mathcal{Z}} \, \prod_i \, p(\D^{(i)} \given \W^{(i)}) \, p(\D_\sat \given \W_\sat)}\\
+			\times p(\W^{(i)} \given \W_\sat, \tub^{(i)}, \Phi, \Loffset, \tailbit^{(i)}) \, 
+			\, p(\tub^{(i)}) p(\Phi) \, p(\Loffset)
+\end{multline}
 This model works well when all stars are lost due to tidal pruning.
@@
 We will explore properly modeling this process in future work.
-For this article, we modify the above to be a mixture model: we add a new parameter for each star, $\pshock$, which is the probability that the star is drawn from a broader, ``shocked'' distribution instead of a tight Gaussian centered on an effective Lagrange point:
-\begin{multline}
-	p_{\rm strip} = [ \mathcal{N}(\bs{r} \given \bs{r}_\sat - \Loffset\,\tailbit\,\rtide\,\hat{x}_1, \sigma^2_r)\\
-		\times \mathcal{N}(\bs{v} \given \bs{v}_\sat, \sigma^2_v) \left\vert\J\right\vert ]_{t=\tub}
-\end{multline}
-\begin{align}
-	p_{\rm shock} = \left [ \mathcal{N}(\bs{r} \given \bs{r}_\sat, 10\sigma^2_r)\mathcal{N}(\bs{v} \given \bs{v}_\sat, 10\sigma^2_v) \left\vert\J\right\vert\right]_{t=\tub}
-\end{align}
-\begin{multline}
-	p(\W \given \W_\sat, \tub, \Phi, \Loffset, \tailbit, \pshock) = \\
-		\pshock p_{\rm shock} + (1-\pshock)p_{\rm strip}.
-\end{multline}
-Here we assume that the ``shocked'' distribution has a much larger variance, and arbitrarily set this to 10 times the variance of the stripped debris.
+
+%For this article, we modify the above to be a mixture model: we add a new parameter for each star, $\pshock$, which is the probability that the star is drawn from a broader, ``shocked'' distribution instead of a tight Gaussian centered on an effective Lagrange point:
+%\begin{multline}
+%	p_{\rm strip} = [ \mathcal{N}(\bs{r} \given \bs{r}_\sat - \Loffset\,\tailbit\,\rtide\,\hat{x}_1, \sigma^2_r)\\
+%		\times \mathcal{N}(\bs{v} \given \bs{v}_\sat, \sigma^2_v) \left\vert\J\right\vert ]_{t=\tub}
+%\end{multline}
+%\begin{align}
+%	p_{\rm shock} = \left [ \mathcal{N}(\bs{r} \given \bs{r}_\sat, 10\sigma^2_r)\mathcal{N}(\bs{v} \given \bs{v}_\sat, 10\sigma^2_v) \left\vert\J\right\vert\right]_{t=\tub}
+%\end{align}
+%\begin{multline}
+%	p(\W \given \W_\sat, \tub, \Phi, \Loffset, \tailbit, \pshock) = \\
+%		\pshock p_{\rm shock} + (1-\pshock)p_{\rm strip}.
+%\end{multline}
+%Here we assume that the ``shocked'' distribution has a much larger variance, and arbitrarily set this to 10 times the variance of the stripped debris.
 The posterior probability for the position of a star, progenitor, and properties of the parent potential, given the data, is then
-\begin{multline}
-	p(\Phi, \W, \W_\sat, \tub, \Loffset, \tailbit, \pshock \given \D, \D_\sat) = \\
-		\shoveleft{\hspace{3em}\frac{1}{\mathcal{Z}} \, p(\D \given \W) \, p(\D_\sat \given \W_\sat)}\\
-			\times p(\W \given \W_\sat, \tub, \Phi, \Loffset, \tailbit, \pshock) \, 
-			p(\Phi) \, p(\tub) \, p(\Loffset)
-\end{multline}
-where the factor $\mathcal{Z}$ only depends on the properties of the data (the \emph{evidence}), and is thus constant when varying the model parameters.
+%\begin{multline}
+%	p(\Phi, \W, \W_\sat, \tub, \Loffset, \tailbit, \pshock \given \D, \D_\sat) = \\
+%		\shoveleft{\hspace{3em}\frac{1}{\mathcal{Z}} \, p(\D \given \W) \, p(\D_\sat \given \W_\sat)}\\
+%			\times p(\W \given \W_\sat, \tub, \Phi, \Loffset, \tailbit, \pshock) \, 
+%			p(\Phi) \, p(\tub) \, p(\Loffset)
+%\end{multline}
+%where the factor $\mathcal{Z}$ only depends on the properties of the data (the \emph{evidence}), and is thus constant when varying the model parameters.
 All model parameters are summarized in Table~\ref{tbl:params}.
 Assuming each star is an independent tracer, the full posterior for many stars is then just the product of the individual likelihoods with the priors:
-\begin{multline}
-	p(\Phi, \W, \W_\sat, \tub, \Loffset, \tailbit, \pshock \given \D, \D_\sat) = \\
-		\shoveleft{\hspace{1em}\frac{1}{\mathcal{Z}} \, \prod_i \, p(\D^{(i)} \given \W^{(i)}) \, p(\D_\sat \given \W_\sat)}\\
-			\times p(\W^{(i)} \given \W_\sat, \tub^{(i)}, \Phi, \Loffset, \tailbit^{(i)}, \pshock^{(i)}) \, 
-			\, p(\tub^{(i)}) p(\Phi) \, p(\Loffset)
-\end{multline}
+%\begin{multline}
+%	p(\Phi, \W, \W_\sat, \tub, \Loffset, \tailbit, \pshock \given \D, \D_\sat) = \\
+%		\shoveleft{\hspace{1em}\frac{1}{\mathcal{Z}} \, \prod_i \, p(\D^{(i)} \given \W^{(i)}) \, p(\D_\sat \given \W_\sat)}\\
+%			\times p(\W^{(i)} \given \W_\sat, \tub^{(i)}, \Phi, \Loffset, \tailbit^{(i)}, \pshock^{(i)}) \, 
+%			\, p(\tub^{(i)}) p(\Phi) \, p(\Loffset)
+%\end{multline}
 
@@
 
-We first test the model using eight particles (stars) --- four from the leading trail, and four from the trailing tail --- randomly sampled from the $2.5\times10^8~M_\odot$ simulation (Section~\ref{sec:sims}), assuming perfect observational data for both the stars and the progenitor.
+We first test the model using eight particles (stars) --- four from the leading tail, and four from the trailing tail --- randomly sampled from the $2.5\times10^8~M_\odot$ simulation (Section~\ref{sec:sims}), assuming perfect observational data for both the stars and the progenitor.
 The stars were required to have been stripped after the first pericentric passage and have a present-day distance within $50$ kpc of the Sun.

### 2014-03-15 streams/inference/core.py
sample tightly around truth

@@ -271,8 +271,8 @@ class StreamModel(object):
 
                             p0[ii,ix1:ix1+param.size] = np.ravel(rvs)
                         elif hasattr(prior, 'mu'):
-                            p0[ii,ix1:ix1+param.size] = np.ravel(param.sample())
-
+                            v = np.random.normal(param.truth, prior.sigma/10.)
+                            p0[ii,ix1:ix1+param.size] = np.ravel(v)
                         else:
                             p0[ii,ix1:ix1+param.size] = np.ravel(param.truth)
                     else:

### 2014-03-15 scripts/infer_potential.py
only plot if more than one param to show

@@ -314,7 +314,7 @@ def main(config_file, mpi=False, threads=None, overwrite=False):
             flatchains['potential'] = this_flatchain
 
         nparticles = model.true_particles.nparticles
-        if particles_group:
+        if particles_group and len(particles_group) > 1:
             for jj in range(nparticles):
                 this_flatchain = np.zeros((len(thin_flatchain),len(particles_group)))
                 this_p0 = np.zeros((len(p0),len(particles_group)))
@@ -344,7 +344,7 @@ def main(config_file, mpi=False, threads=None, overwrite=False):
                 fig.savefig(os.path.join(output_path, "particle{}.{}".format(jj,plot_ext)))
 
         # plot the posterior for the satellite parameters
-        if satellite_group:
+        if satellite_group and len(satellite_group) > 1:
             jj = 0
             this_flatchain = np.zeros((len(thin_flatchain),len(satellite_group)))
             this_p0 = np.zeros((len(p0),len(satellite_group)))

### 2014-03-15 text/paper2/ms.tex
added more notes on experiments

@@
 
-Figures~\ref{fig:}-\ref{fig:} show the posterior distributions for these five parameters for each of the the four mass progenitors.
+Figures~\ref{fig:}-\ref{fig:} show the posterior distributions for these five parameters for each of the the four mass progenitors...
 
 \subsection{\spitzer + \gaia}
-We now take the same eight stars used in the previous experiment and ``observe'' them with optimistic observational uncertainties.
-We take these stars to be RR Lyrae variables which are known to be excellent distance indicators using the mid-infrared period-luminosity relation \citep[as shown in, e.g.,][]{madore12}; $\sim$100 RR Lyrae associated with the Sgr stream and $\sim$30 associated with the Orphan stream will be observed with \spitzer as part of the SMASH survey \citep{smashprop} with expected fractional distance uncertainties around $\sim$2\%.
+For this and subsequent experiments, we use only the $m=2.5\times10^8~M_\odot$ simulation.
+We take the same eight stars used in the previous experiment and ``observe'' them with optimistic observational uncertainties.
+We take these stars to be RR Lyrae variables which are known to be excellent distance indicators via the mid-infrared period-luminosity relation \citep[as shown in, e.g.,][]{madore12}.
+$\sim$100 RR Lyrae associated with the Sgr stream and $\sim$30 associated with the Orphan stream will be observed with \spitzer as part of the SMASH survey \citep{smashprop} with expected fractional distance uncertainties around $\sim$2\%.
 These stars will also be included in the \gaia\, proper motion catalog: at a distance of $\sim$50~kpc, a typical RR Lyrae will have a tangential velocity error around $\sim$20~km/s.
 
-For this experiment, we (1) assume the stars are bright distance indicators --- RR Lyrae --- such that the fractional distance uncertainty is $2\%$; (2) neglect the uncertainty in angular position (for a typical RR Lyrae at 50~kpc this is $\sim$$10^{-7}$~deg); (3) assume we can measure radial velocities to these stars with 5~km/s uncertainty; and (4) compute the sky-averaged \gaia\, proper-motion uncertainty for each star assuming an F0V spectral type using the \texttt{PyGaia}\footnote{https://github.com/agabrown/PyGaia} code and use this uncertainty for both components of proper motion.
+For this experiment, we (1) assume the stars are bright distance indicators --- RR Lyrae --- such that the fractional distance uncertainty is $2\%$; (2) neglect the uncertainty in angular position (for a typical RR Lyrae at 50~kpc this is $\sim$$10^{-7}$~deg for \gaia); (3) assume we can measure radial velocities to these stars with 5~km/s uncertainty; and (4) compute the sky-averaged \gaia\, proper-motion uncertainty for each star assuming an F0V spectral type using the \texttt{PyGaia}\footnote{https://github.com/agabrown/PyGaia} code and use this uncertainty for both components of proper motion.
 % TODO: acknowledge Gaia science team, Anthony Brown, Gaia data challenge?
 The observed positions and velocities of these stars are shown in Figure~\ref{fig:} (red, compare to the unobserved black points).
+We further assume that we know the tail assignment for each star, $\beta$.
+We observe the position of the progenitor with the same observational uncertainties, though in reality some coordinates will have higher or lower uncertainties.
+
+This experiment has the four potential parameters, the Lagrange point offset, four phase-space coordinate parameters for each star (32 total), and four phase-space coordinate parameters for the progenitor --- 41 parameters in total.
 
@@
 Aspen?\\
+Gaia data challenge / Surrey\\
 APW: NSF GRFP, Sigma Xi\\

### 2014-03-15 scripts/observe_simulation.py
add more comments

@@ -29,7 +29,7 @@ from streams import usys
 from streams.coordinates.frame import heliocentric, galactocentric
 from streams.dynamics import ObservedParticle
 import streams.io as s_io
-from streams.observation.gaia import gaia_spitzer_errors
+from streams.observation.gaia import gaia_spitzer_errors, current_data_errors
 from streams.potential.lm10 import LawMajewski2010
 from streams.util import _parse_quantity
 from streams.integrate import LeapfrogIntegrator
@@ -221,11 +221,17 @@ python scripts/observe_simulation.py -v --class_name=SgrSimulationDH --expr='tub
 python scripts/observe_simulation.py -v --class_name=SgrSimulationDH --expr='tub!=0' \
 --N=1024 --file=/Users/adrian/projects/streams/data/observed_particles/2.5e9_N1024_DH.hdf5 \
 --seed=42 --mass="2.5e9" --overwrite
+
+OR
+
+python scripts/observe_simulation.py -v --class_name=SgrSimulationDH --expr='tub!=0' \
+--N=1024 --file=/Users/adrian/projects/streams/data/observed_particles/2.5e8_N1024_DH_exp3.hdf5 \
+--seed=42 --mass="2.5e8" --overwrite
     """
 
     # TODO: class kwargs
+    # particle_error_model=current_data_errors, satellite_error_model=current_data_errors,
     # particle_error_model=gaia_spitzer_errors, satellite_error_model=gaia_spitzer_errors,
-    # particle_error_model=gaia_spitzer_errors, satellite_error_model=None,
     observe_simulation(args.class_name,
         particle_error_model=gaia_spitzer_errors, satellite_error_model=gaia_spitzer_errors,
         selection_expr=args.expr, N=args.N, output_file=args.output_file,

### 2014-03-15 text/paper2/ms.tex
experiments

@@
 We leave the potential parameters --- $q_1$, $q_z$, $\phi$, and $\vhalo$ --- free to vary, along with the Lagrange point offset, $\Loffset$, and use Markov Chain Monte Carlo (MCMC) to sample from the posterior probability distribution over these five parameters.
-In detail, we use an affine-invariant, ensemble sampler \citep{goodman10} implemented in the Python programming language \citep{foremanmackey13} with an ensemble of 64 walkers initialized by sampling from the priors summarized in Table~\ref{tbl:params}.
-We run the walkers in parallel (using MPI todo) for 5000 steps to burn-in, then restart the sampling starting from the final position of the burn-in phase and run for another 10000 steps.
+We use an affine-invariant ensemble sampler \citep{goodman10} implemented in the Python programming language \citep{foremanmackey13} with an ensemble of 64 walkers initialized by sampling from the priors summarized in Table~\ref{tbl:params}.
+We run the walkers in parallel (using MPI todo) for 5000 steps to burn-in, then restart the sampler starting from the final position of the burn-in phase and run for another 10000 steps.
 We compute the autocorrelation time for each parameter [todo: reference] and thin the chains by taking every XXX$\times \mathrm{median}(t_{\rm acor})$ sample to ensure they are close to independent.
 Figure~\ref{fig:} shows the walker positions over the 10000 post-burn-in steps for each parameter for the $2.5\times10^8~M_\odot$ simulation; the autocorrelation time for each parameter is displayed on each corresponding panel.
-
-Figures~\ref{fig:}-\ref{fig:} show the posterior distributions for these five parameters for each of the the four mass progenitors...
+Figure~\ref{fig:}c shows the posterior probability for these five parameters for the same simulation.
 

### 2014-03-15 streams/observation/gaia.py
don't multiply by 100

@@ -112,9 +112,8 @@ def gaia_spitzer_errors(particles):
     dmu = proper_motion_error(V, rrl_V_minus_I)
 
     # angular position error is negligible
-    # HACK: extra factor of 100 here
-    l = dlb.to(particles["l"].unit)*100
-    b = dlb.to(particles["b"].unit)*100
+    l = dlb.to(particles["l"].unit)
+    b = dlb.to(particles["b"].unit)
 
     # proper motion
     mul = dmu.to(particles["mul"].unit)

### 2014-03-16 scripts/infer_potential.py
added continue option

@@ -64,7 +64,18 @@ _unit_transform["mul"] = lambda x: (x*u.rad/u.Myr).to(u.mas/u.yr).value
 _unit_transform["mub"] = lambda x: (x*u.rad/u.Myr).to(u.mas/u.yr).value
 _unit_transform["vr"] = lambda x: (x*u.kpc/u.Myr).to(u.km/u.s).value
 
-def main(config_file, mpi=False, threads=None, overwrite=False):
+def fix_whack_walkers(acc_frac, pos, threshold=0.01):
+    if np.any(acc_frac < 0.02):
+        ix = acc_frac < 0.02
+
+        # resample positions for bad walkers
+        pos[ix] = np.random.normal(np.median(pos[~ix], axis=0),
+                                   np.median(pos[~ix], axis=0)/10.,
+                                   size=(sum(ix),pos.shape[1]))
+
+    return pos
+
+def main(config_file, mpi=False, threads=None, overwrite=False, continue_sampler=False):
     """ TODO: """
 
     # get a pool object given the configuration parameters
@@ -99,17 +110,12 @@ def main(config_file, mpi=False, threads=None, overwrite=False):
     ncool_down = config.get("cool_down", 100)
     burn_beta = config.get("burn_beta", 1.)
 
-    if not os.path.exists(output_file):
+    if not os.path.exists(output_file) and not continue_sampler:
         logger.info("Output file '{}' doesn't exist, running inference...".format(output_file))
 
         # sample starting positions
         p0 = model.sample_priors(size=nwalkers,
                                  start_truth=config.get("start_truth", False))
-        # if config.get("start_truth", False):
-        #     p0 = np.random.normal(model.truths,
-        #                           np.std(p0,axis=0)/10.,
-        #                           size=p0.shape)
-
         logger.debug("Priors sampled...")
 
         if nburn > 0:
@@ -120,14 +126,7 @@ def main(config_file, mpi=False, threads=None, overwrite=False):
             logger.info("Burning in sampler for {} steps...".format(nburn))
 
             pos, xx, yy = sampler.run_mcmc(p0, nburn)
-
-            if np.any(sampler.acceptance_fraction < 0.01):
-                ix = sampler.acceptance_fraction < 0.01
-
-                # resample positions for bad walkers
-                pos[ix] = np.random.normal(np.median(pos[~ix], axis=0),
-                                           np.std(pos[~ix], axis=0)/10.,
-                                           size=(sum(ix),pos.shape[1]))
[…]

### 2014-03-16 text/paper2/ms.tex
more experiments

@@
 		tail assignment & $\beta$ & $\{-1, 1\}$\\ % TODO: how to say equal weight to -1 or 1?
-		prob. shocked & $\pshock$ & $\mathcal{U}(0,1)$\\
+%		prob. shocked & $\pshock$ & $\mathcal{U}(0,1)$\\
 		time unbound & $\tub$ & $\mathcal{U}(0,\tint)$\\ % TODO: talk about \tint (integration time) in text!
@@
 \subsection{\spitzer + \gaia}
-For this and subsequent experiments, we use only the $m=2.5\times10^8~M_\odot$ simulation.
+For this and subsequent experiments (unless specified), we use only the $2.5\times10^8~M_\odot$ simulation.
 We take the same eight stars used in the previous experiment and ``observe'' them with optimistic observational uncertainties.
@@
 For this experiment, we (1) assume the stars are bright distance indicators --- RR Lyrae --- such that the fractional distance uncertainty is $2\%$; (2) neglect the uncertainty in angular position (for a typical RR Lyrae at 50~kpc this is $\sim$$10^{-7}$~deg for \gaia); (3) assume we can measure radial velocities to these stars with 5~km/s uncertainty; and (4) compute the sky-averaged \gaia\, proper-motion uncertainty for each star assuming an F0V spectral type using the \texttt{PyGaia}\footnote{https://github.com/agabrown/PyGaia} code and use this uncertainty for both components of proper motion.
-% TODO: acknowledge Gaia science team, Anthony Brown, Gaia data challenge?
 The observed positions and velocities of these stars are shown in Figure~\ref{fig:} (red, compare to the unobserved black points).
@@
 
-This experiment has the four potential parameters, the Lagrange point offset, four phase-space coordinate parameters for each star (32 total), and four phase-space coordinate parameters for the progenitor --- 41 parameters in total.
+This experiment samples over the four potential parameters, the Lagrange point offset, four phase-space coordinate parameters for each star (32 total), and four phase-space coordinate parameters for the progenitor --- 41 parameters in total.
+We use an ensemble of 256 walkers and draw initial conditions for the coordinate parameters by sampling from Gaussian's centered on the ``observed'' value with variances specified by the observational uncertainties.
+Other parameters --- potential parameters, $\Loffset$, and $\tub$ --- are initialized by drawing from the priors summarized in Table~\ref{tbl:params}.
+We again burn in the walkers for 5000 steps and run for 10000 steps post-burn-in.
+The chains are also thinned by taking every XXX$\times \mathrm{median}(t_{\rm acor})$ sample.
+Figures~\ref{fig:}-\ref{fig:} show the marginalized posterior for the potential parameters, satellite, and a single particle.
+[....]
 
 \subsection{\spitzer + current data}
-4 stars in Sgr-like stream
+The SMASH survey \citep{smashprop} will be completed by XXX, long before the final \gaia\, data release.
+Thus, we will soon have very precise distance measurements for stars in Sgr and Orphan but poor proper motion constraints.
+[some comments on state of data for Sgr core, proper motions of] Here we study how well we can constrain the potential with this model given stars with missing data.
+We assume the same observational uncertainties for the stars as mentioned above, but assume the proper motion's have not been measured.
 
@@
 Aspen?\\
-Gaia data challenge / Surrey\\
+Gaia data challenge / Surrey / Gaia data team / Anthony Brown \\
 APW: NSF GRFP, Sigma Xi\\

### 2014-03-16 text/paper2/ms.tex
one sim

@@
 Figure~\ref{fig:} shows the walker positions over the 10000 post-burn-in steps for each parameter for the $2.5\times10^8~M_\odot$ simulation; the autocorrelation time for each parameter is displayed on each corresponding panel.
-Figure~\ref{fig:}c shows the posterior probability for these five parameters for the same simulation.
+Figure~\ref{fig:} shows projections of the posterior probability distribution for the parameters.
 
 \subsection{\spitzer + \gaia}
-For this and subsequent experiments (unless specified), we use only the $2.5\times10^8~M_\odot$ simulation.
 We take the same eight stars used in the previous experiment and ``observe'' them with optimistic observational uncertainties.

### 2014-03-16 scripts/infer_potential.py
fixed prior on tub

@@ -70,7 +70,7 @@ def fix_whack_walkers(acc_frac, pos, threshold=0.01):
 
         # resample positions for bad walkers
         pos[ix] = np.random.normal(np.median(pos[~ix], axis=0),
-                                   np.median(pos[~ix], axis=0)/10.,
+                                   np.fabs(np.median(pos[~ix], axis=0)/10.),
                                    size=(sum(ix),pos.shape[1]))
 
     return pos

### 2014-03-16 streams/inference/prior.py

@@ -23,6 +23,9 @@ class LogPrior(object):
     def __call__(self, value):
         return 0.
 
+    def sample(self, size=None):
+        return np.nan
+
 class LogUniformPrior(LogPrior):
 
     def __call__(self, value):

### 2014-03-16 streams/io/core.py

@@ -136,7 +136,7 @@ def read_hdf5(h5file, nparticles=None, particle_idx=None):
 
         tub = ModelParameter(name="tub",
                              truth=true_tub,
-                             prior=LogUniformPrior([t1]*nparticles,[t2]*nparticles))
+                             prior=LogUniformPrior([t2]*nparticles,[t1]*nparticles))
         beta = ModelParameter(name="beta",
                               truth=true_tail_bit,
                               prior=LogUniformPrior([-2.]*nparticles,[2.]*nparticles))

### 2014-03-16 text/paper2/ms.tex
added choi reference
add figure references

@@
 \usepackage{booktabs}
+\usepackage[caption=false]{subfig}
 
@@
 	p(\W \given m, \dot{m}, \W_\sat, \tub, \Phi, \Loffset, \tailbit) &= p(\bs{r} \given \bs{r}_\sat, m, \dot{m}, \tub, \Phi, \Loffset, \tailbit)\,p(\bs{v} \given \bs{v}_\sat, m, \dot{m}, \tub, \Phi)\,\left\vert\J\right\vert_{t=\tub}\\
-	p(\bs{r} \given \bs{r}_\sat, m, \dot{m}, \tub, \Phi, \Loffset, \tailbit) &= \mathcal{N}(\bs{r} \given \bs{r}_\sat - \Loffset\,\tailbit\,\rtide\,\hat{x}_1, \sigma^2_r)_{t=\tub}\\
+	p(\bs{r} \given \bs{r}_\sat, m, \dot{m}, \tub, \Phi, \Loffset, \tailbit) &= \mathcal{N}(\bs{r} \given \bs{r}_\sat - \Loffset\,\tailbit\,\rtide\,\hat{x}_1, \rtide^2)_{t=\tub}\\
 	p(\bs{v} \given \bs{v}_\sat, m, \dot{m}, \tub, \Phi) &= \mathcal{N}(\bs{v} \given \bs{v}_\sat, \sigma^2_v)_{t=\tub}
@@
 \begin{multline}
-	p(\Phi, \W, \W_\sat, \tub, \Loffset, \tailbit, \given \D, \D_\sat) = \\
+	p(\Phi, \{\W^{(i)}\}, \W_\sat, \tub, \Loffset, \tailbit, \given \{\D^{(i)}\}, \D_\sat) = \\
 		\shoveleft{\hspace{1em}\frac{1}{\mathcal{Z}} \, \prod_i \, p(\D^{(i)} \given \W^{(i)}) \, p(\D_\sat \given \W_\sat)}\\
@@
 This model works well when all stars are lost due to tidal pruning.
-Even on a mildly eccentric orbit, stars become preferentially unbound around the satellite's pericenter because of tidal shocking (cite?).
+Even on a mildly eccentric orbit, stars become preferentially unbound around the satellite's pericenter because of tidal shocking \citep[e.g.,][]{choi09}.
 We will explore properly modeling this process in future work.
@@
 The stars were required to have been stripped after the first pericentric passage and have a present-day distance within $50$ kpc of the Sun.
-Figure~\ref{fig:} (left panel) shows a sample of stream particles (black) and the eight randomly chosen stars (blue) in the Galactic X-Z plane.
+Figure~\ref{fig:2.5e8_xz} (left panel) shows a sample of stream particles (grey) and the eight randomly chosen stars (blue) in the Galactic X-Z plane.
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=\textwidth]{../../plots/paper2/e8_xz.png}
+\caption{ TODO  }\label{fig:2.5e8_xz}
+\end{center}
+\end{figure}
 
@@
 We compute the autocorrelation time for each parameter [todo: reference] and thin the chains by taking every XXX$\times \mathrm{median}(t_{\rm acor})$ sample to ensure they are close to independent.
-Figure~\ref{fig:} shows the walker positions over the 10000 post-burn-in steps for each parameter for the $2.5\times10^8~M_\odot$ simulation; the autocorrelation time for each parameter is displayed on each corresponding panel.
-Figure~\ref{fig:} shows projections of the posterior probability distribution for the parameters.
+Figure~\ref{fig:trace} shows the walker positions over the 10000 post-burn-in steps for each parameter for the $2.5\times10^8~M_\odot$ simulation; the autocorrelation time for each parameter is displayed on each corresponding panel.
+Figure~\ref{fig:exp1_posterior} shows projections of the posterior probability distribution for the parameters.
+
+\begin{figure*}[!ht]
+	\captionsetup[subfigure]{labelformat=empty}
+        \centering
+        \subfloat[]{
+                \includegraphics[width=0.5\textwidth]{../../plots/paper2/trace_0.png}
+                \label{fig:trace0}
+	}
+        \subfloat[]{
+                \includegraphics[width=0.5\textwidth]{../../plots/paper2/trace_1.png}
+                \label{fig:trace0}
+	}
+
+        \subfloat[]{
+                \includegraphics[width=0.5\textwidth]{../../plots/paper2/trace_2.png}
+                \label{fig:trace0}
+	}
+        \subfloat[]{
+                \includegraphics[width=0.5\textwidth]{../../plots/paper2/trace_3.png}
+                \label{fig:trace0}
+	}
+
+        \subfloat[]{
+                \includegraphics[width=0.5\textwidth]{../../plots/paper2/trace_4.png}
+                \label{fig:trace0}
+	}
+        \caption{TODO: mcmc trace plots}\label{fig:trace}
+\end{figure*}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[width=\textwidth]{../../plots/paper2/exp1_posterior.png}
+\caption{ TODO  }\label{fig:exp1_posterior}
+\end{center}
+\end{figure}
 
@@
 For this experiment, we (1) assume the stars are bright distance indicators --- RR Lyrae --- such that the fractional distance uncertainty is $2\%$; (2) neglect the uncertainty in angular position (for a typical RR Lyrae at 50~kpc this is $\sim$$10^{-7}$~deg for \gaia); (3) assume we can measure radial velocities to these stars with 5~km/s uncertainty; and (4) compute the sky-averaged \gaia\, proper-motion uncertainty for each star assuming an F0V spectral type using the \texttt{PyGaia}\footnote{https://github.com/agabrown/PyGaia} code and use this uncertainty for both components of proper motion.
-The observed positions and velocities of these stars are shown in Figure~\ref{fig:} (red, compare to the unobserved black points).
+The observed positions and velocities of these stars are shown in Figure~\ref{fig:2.5e8_xz} (red, compare to the unobserved black points).
 We further assume that we know the tail assignment for each star, $\beta$.
@@
 The chains are also thinned by taking every XXX$\times \mathrm{median}(t_{\rm acor})$ sample.
-Figures~\ref{fig:}-\ref{fig:} show the marginalized posterior for the potential parameters, satellite, and a single particle.
+Figures~\ref{fig:exp2_potential}, \ref{fig:exp2_satellite}, \ref{fig:exp2_particle0} show the marginalized posteriors for the potential, satellite, and a single particle parameters.
 [....]
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[width=\textwidth]{../../plots/paper2/exp2_potential.png}
+\caption{ TODO  }\label{fig:exp2_potential}
+\end{center}
+\end{figure}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[width=\textwidth]{../../plots/paper2/exp2_satellite.png}
+\caption{ TODO  }\label{fig:exp2_satellite}
+\end{center}
+\end{figure}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[width=\textwidth]{../../plots/paper2/exp2_particle0.png}
+\caption{ TODO  }\label{fig:exp2_particle0}
+\end{center}
+\end{figure}
 

### 2014-03-18 scripts/figures/paper2_figures.py
stylistic tweaks

@@ -26,8 +26,10 @@ import scipy.optimize as so
 from scipy.stats import norm
 
 from streams import usys
-from streams.util import project_root
+from streams.util import streamspath
+from streams.coordinates.frame import galactocentric
 from streams.io.sgr import SgrSimulation
+from streams.io import read_hdf5, read_config
 from streams.integrate import LeapfrogIntegrator
 from streams.potential.lm10 import LawMajewski2010
 
@@ -45,45 +47,66 @@ snapfile = 'SNAP113'
 # sgr_path = 'sgr_plummer/2.5e{}'
 # snapfile = 'SNAP'
 
-plot_path = "plots/paper2/"
+plot_path = os.path.join(streamspath, "plots/paper2/")
 if not os.path.exists(plot_path):
     os.mkdir(plot_path)
 
-ext = 'png'
+ext = 'pdf'
+grid_figsize = (14,7.5)
 
 def simulated_streams(**kwargs):
 
     filename = os.path.join(plot_path, "simulated_streams.{}".format(ext))
-    fig,axes = plt.subplots(2,4,figsize=(14,7.5),
+    fig,axes = plt.subplots(2,4,figsize=grid_figsize,
                             sharex=True, sharey=True)
 
     ticks = [-100,-50,0,50]
-    alpha = 0.3
+    alphas = [0.2, 0.27, 0.34, 0.4]
     rcparams = {'lines.linestyle' : 'none',
-                'lines.marker' : '.',
-                'lines.markersize' : 3}
+                'lines.marker' : ','}
 
     with rc_context(rc=rcparams):
         for ii,_m in enumerate(range(6,9+1)):
+            alpha = alphas[ii]
             mass = "2.5e{}".format(_m)
             print(mass)
             m = float(mass)
 
+            data_filename = os.path.join(streamspath, "data", "observed_particles",
+                                         "2.5e{}_N1024_DH.hdf5".format(_m))
+            cfg_filename = os.path.join(streamspath, "config", "exp1_{}.yml".format(_m))
+            data = read_hdf5(data_filename)
+            true_particles = data["true_particles"].to_frame(galactocentric)
+            config = read_config(cfg_filename)
+            idx = config['particle_idx']
+
             sgr = SgrSimulation(sgr_path.format(_m),snapfile)
[…]

### 2014-03-18 text/paper2/ms.tex
updates

@@
 \includegraphics[width=\textwidth]{../../plots/paper2/simulated_streams.pdf}
-\caption{ Particle positions in the final timestep of four n-body simulations with the same progenitor orbit initial conditions over a range of progenitor masses.
-The simulations were run [...].
-}\label{fig:sims}
+\caption{ Particle positions (grey dots) in Galactocentric cartesian coordinates from the final timestep of four n-body simulations with the same progenitor orbit initial conditions over a range of progenitor masses (columns).
+The progenitors were initialized as [TODO: DAH].
+Black crosses indicate eight particles chosen from each mass simulation and used in the experiments described in Section~\ref{sec:experiments}.}\label{fig:sims}
 \end{center}
@@
 
-For a satellite galaxy on a mildly eccentric orbit, mass-loss is driven by a combination of tidal stripping due to the steady tidal field of the parent system and tidal shocking due to the rapidly changing tidal field as the progenitor system moves through pericenter.
+For a satellite galaxy on a mildly eccentric orbit, mass-loss is driven by a combination of tidal stripping due to the steady tidal field of the parent system and tidal shocking due to the rapidly changing tidal field as the progenitor system moves through pericenter \citep[e.g.,][]{choi09}.
 The interplay between these two processes as a function of orbital properties will be studied in forthcoming work; we focus below on the slow disruption of stars due to steady tidal forcing.
@@
 Consider a point-mass satellite of mass $m$ on a circular orbit with frequency $\Omega$ around a more massive point-mass, $M\gg m$ \citep[the restricted three-body problem; e.g., \S 8.3][]{binneytremaine}.
-In a frame rotating with the orbital frequency of the satellite, the effective potential around the satellite has (amongst others) two stationary, unstable points located around $\sim R \pm r_J$, where $R$ is the distance to the satellite and $r_J$ is the Jacobi or tidal radius,
+In a frame rotating with the orbital frequency of the satellite, the effective potential around the satellite has (amongst others) two stationary, unstable points located around $\sim R \pm r_J$, where $R$ is the orbital radius of the satellite and $r_J$ is the Jacobi or tidal radius,
 \begin{equation}
@@
 \end{equation}
-Particles that would be bound to the satellite system in isolation may have enough (Jacobi) energy to overcome the effective potential barrier at these Lagrange points (Fig.~8.6 of BT) and thus will be preferentially stripped from the satellite at these points.
-For a spherical, extended parent mass distribution the tidal radius instead scales with the enclosed mass at a given orbital position, $R$, with additional terms that account for the local slope of the density profile.
+Particles that would be bound to the satellite mass in isolation may have enough (Jacobi) energy to overcome the effective potential barrier at these Lagrange points (Fig.~8.6 of BT) and thus will be preferentially stripped from the satellite at these points.
+For a spherical, extended parent mass distribution the tidal radius instead scales with the enclosed mass at the instantaneous orbital radius, $R(t)$, with additional terms that account for the local slope of the density profile.
+The satellite may also be losing mass, in which case the satellite mass also depends on time.
 In general, the effective Lagrange points may not be symmetric about the center of the satellite and may deviate from the point-mass tidal radius (Eq.~\ref{eq:ptmass}) by factors of order unity.
-We thus adopt
+We adopt
 \begin{equation}
-	r_{\rm tide} = R\left(\frac{m}{3M_{\rm enc}}\right)^{1/3}\label{eq:tidalradius}
+	r_{\rm tide}(t) = R(t)\left(\frac{m(t)}{3M_{\rm enc}(R)}\right)^{1/3}\label{eq:tidalradius}
 \end{equation}
@@
 We expect the debris star velocities to also scale with $(m/M_{\rm enc})^{1/3}$.
+These scalings assume that the satellite is spherical with isotropic velocities; correlations amongst intra-satellite orbits from, e.g., a disk will break these assumptions.
 
-In a triaxial potential, the orbital plane of the satellite is not fixed but we still expect there to be effective Lagrange points along the line of centers connecting the origin of the parent potential to the satellite.
+In a triaxial potential, the orbital plane of the satellite is not fixed but we still expect there to be \emph{effective} Lagrange points for a spherical satellite system along the line of centers connecting the origin of the parent potential to the satellite.
 We thus proceed by defining a coordinate system relative to the position and velocity of the progenitor, ($\bs{r}_p$,$\bs{v}_p$), rotated into the instantaneous orbital plane defined by $\bs{r}_p \times \bs{v}_p$, and normalized by the instantaneous tidal radius for positions (Eq.~\ref{eq:tidalradius}) and instantaneous velocity scale for velocity (Eq.~\ref{eq:velscale}).
@@
 %(Section~\ref{sec:sims})
-Figure~\ref{fig:lpts_r} shows sections of particle orbits from the n-body simulations described above projected into this coordinate system, symmetric around their unbinding times, $\tub$, for the four progenitor masses.
-In these coordinates, the classical Lagrange points would be located around $x_1=\pm1,x_2=0,x_3=0$ (illustrated by the unit circle in top panels of Figure~\ref{fig:lpts_r}).
+Figure~\ref{fig:lpts_r} shows sections of particle orbits from the n-body simulations described above projected into this coordinate system and normalized, symmetric around their unbinding times, $\tub$, for each of the four progenitor masses.
+In these coordinates, the classical Lagrange points would be located at $x_1\approx\pm1,x_2\approx0,x_3\approx0$ (illustrated by the unit circle in top panels of Figure~\ref{fig:lpts_r}).
 Figure~\ref{fig:lpts_v} shows the velocity of each star also projected into these coordinates and normalized.
@@
 
-Even in a triaxial potential, the phase-space dispersion of tidally stripped debris scales with the mass ratio $(m / M_{\rm enc})^{1/3}$.
-This  motivates a model for tidal debris in which stars were ``released'' at the instantaneous, effective Lagrange points with some dispersion in position and velocity, all of which depend only on the mass and orbit of the progenitor, and the parent potential.
+We find that even in a triaxial potential, the phase-space dispersion of tidally stripped debris scales with the mass ratio $(m / M_{\rm enc})^{1/3}$.
+This motivates a model for tidal debris in each star was ``released'' at the instantaneous, effective Lagrange point at its unbinding time, $\tub$, with some dispersion in position and velocity, all of which depend only on the mass and orbit of the progenitor and the parent potential.
 We present this model in detail below.
@@
 We observe the 6D position of a star, $\D$, in heliocentric coordinates --- e.g., the measured position on the sky, ($l$, $b$); distance, $d$; proper motions, ($\mu_l$, $\mu_b$); and line-of-sight velocity, $v_r$ --- and have determined through some other means that this star was once part of a progenitor system (e.g., satellite galaxy) with mass $m(t)$ that is disrupting and forming a cold debris structure in the potential, $\Phi$, of the parent galaxy (e.g., the Milky Way).
-We assume that the mass of the parent system enclosed in the orbit of the satellite, $M_{\rm enc}$, is much larger than the initial mass of the satellite, $M_{\rm enc} \gg m(t=0)$.
+We assume that the mass of the parent system enclosed within the pericenter of the orbit of the satellite, $M_{\rm enc}(R_{\rm peri})$, is much larger than the initial mass of the satellite, $M_{\rm enc}(R_{\rm peri})\gg m(t=0)$.
 The present position of the progenitor is observed to be at heliocentric position $\D_\sat$ (where any subscript $\sat$ refers to the progenitor).
@@
 To model the true, 6D, present-day position of the star, $\W$, we first transform to cartesian, Galactocentric coordinates where ($\bs{r},\bs{v}$) and ($\bs{r}_\sat,\bs{v}_\sat$) are the position and velocity of the star and progenitor.
-In cartesian coordinates, the present-day position of the star is taken as having been sampled at $t=\tub$ from an isotropic Gaussian centered on one of the instantaneous, effective Lagrange points in position, and an isotropic Gaussian centered on the origin in velocity, and evolved to present day in the parent potential, $\Phi$.
+In cartesian coordinates, the present-day position of the star is taken as having been sampled at $t=\tub$ from an isotropic Gaussian centered on one of the instantaneous, effective Lagrange points in position, and an isotropic Gaussian centered on the origin in velocity, evolved to present day in the parent potential, $\Phi$.
 We assume that once the star becomes ``unbound'' from the satellite the potential of the satellite can be ignored, and thus treat the star as a test particle.\footnote{This assumption breaks down for sufficiently large-mass progenitors, somewhere between $\sim10^8-10^9~$M$_\odot$.} The satellite mass enters the model through the tidal radius; a changing satellite mass will skew the positions of the effective Lagrange points and velocity scale.
-We allow for mass-loss by incorporating the initial satellite mass, $m$, and a constant mass-loss term, $\dot{m}$ into the model.
+We allow for mass-loss by incorporating the initial satellite mass, $m_0$, and a constant mass-loss term, $\dot{m}$ into the model: $m(t) = m_0 - \dot{m}t$.
 Finally, we add two additional parameters: a constant, global offset, $\Loffset$, that scales the position of the effective Lagrange points relative to the classical tidal radius, and a binary parameter $\tailbit$, that is either $-1$ or $+1$ depending on whether the star is in the leading or trailing tail.
@@
 \begin{align}
-	p(\W \given m, \dot{m}, \W_\sat, \tub, \Phi, \Loffset, \tailbit) &= p(\bs{r} \given \bs{r}_\sat, m, \dot{m}, \tub, \Phi, \Loffset, \tailbit)\,p(\bs{v} \given \bs{v}_\sat, m, \dot{m}, \tub, \Phi)\,\left\vert\J\right\vert_{t=\tub}\\
-	p(\bs{r} \given \bs{r}_\sat, m, \dot{m}, \tub, \Phi, \Loffset, \tailbit) &= \mathcal{N}(\bs{r} \given \bs{r}_\sat - \Loffset\,\tailbit\,\rtide\,\hat{x}_1, \rtide^2)_{t=\tub}\\
-	p(\bs{v} \given \bs{v}_\sat, m, \dot{m}, \tub, \Phi) &= \mathcal{N}(\bs{v} \given \bs{v}_\sat, \sigma^2_v)_{t=\tub}
+	p(\W \given \W_\sat, m_0, \dot{m}, \tub, \Phi, \Loffset, \tailbit) &= p(\bs{r} \given \bs{r}_\sat, m_0, \dot{m}, \tub, \Phi, \Loffset, \tailbit)\,p(\bs{v} \given \bs{v}_\sat, m_0, \dot{m}, \tub, \Phi)\,\left\vert\J\right\vert_{t=\tub}\\
+	p(\bs{r} \given \bs{r}_\sat, m_0, \dot{m}, \tub, \Phi, \Loffset, \tailbit) &= \mathcal{N}(\bs{r} \given \bs{r}_\sat - \Loffset\,\tailbit\,\rtide\,\hat{x}_1, \rtide^2)_{t=\tub}\\
+	p(\bs{v} \given \bs{v}_\sat, m_0, \dot{m}, \tub, \Phi) &= \mathcal{N}(\bs{v} \given \bs{v}_\sat, \sigma^2_v)_{t=\tub}
 \end{align}
@@
 \begin{multline}
-	p(\Phi, \{\W^{(i)}\}, \W_\sat, \tub, \Loffset, \tailbit, \given \{\D^{(i)}\}, \D_\sat) = \\
+	p(\Phi, \{\W^{(i)}\}, \W_\sat, m_0, \dot{m}, \tub, \Loffset, \tailbit, \given \{\D^{(i)}\}, \D_\sat) = \\
 		\shoveleft{\hspace{1em}\frac{1}{\mathcal{Z}} \, \prod_i \, p(\D^{(i)} \given \W^{(i)}) \, p(\D_\sat \given \W_\sat)}\\
-			\times p(\W^{(i)} \given \W_\sat, \tub^{(i)}, \Phi, \Loffset, \tailbit^{(i)}) \, 
+			\times p(\W^{(i)} \given \W_\sat, m_0, \dot{m}, \tub^{(i)}, \Phi, \Loffset, \tailbit^{(i)}) \, 
 			\, p(\tub^{(i)}) p(\Phi) \, p(\Loffset)
@@
 This model works well when all stars are lost due to tidal pruning.
-Even on a mildly eccentric orbit, stars become preferentially unbound around the satellite's pericenter because of tidal shocking \citep[e.g.,][]{choi09}.
+Even on a mildly eccentric orbit, stars become preferentially unbound around the satellite's pericenter because of tidal shocking.
 We will explore properly modeling this process in future work.
@@
 
-\section{Experiments}
+\section{Experiments} \label{sec:experiments}
 In what follows, we apply the model described above to ``observations'' of the n-body simulations described in Section~\ref{sec:sims}.
-For all tests in this article, we model the potential of the parent galaxy with the same functional form used in the simulations (Equation~\ref{eq:lm10}) --- we stress that the simulations are n-body simulations and not generated with the same generative model presented here.
+For all tests in this article, we model the potential of the parent galaxy with the same functional form used in the simulations (Equation~\ref{eq:lm10}) --- we stress that the data are generated with n-body simulations and are not generated with the same model presented here.
 When recovering the potential, we hold fixed the disk and spheroid parameters (see Table~\ref{tbl:params}), along with two halo parameters: the scale radius, $r_h$, and one of the flattening components in the x-y plane, $q_2$.
 These halo parameters are degenerate with combinations of the other parameters --- e.g., setting $q_1=2.0$, $q_2=1.0$, $\phi=0^\circ$ is equivalent to setting $q_1=1.0$, $q_2=2.0$, $\phi=90^\circ$.
-The priors on the remaining halo parameters are taken to be uniform over a conservative domain of realistic values: for $\vhalo$, 100-200~km/s corresponds to a range in solar circular velocities from $\sim$210-250 km/s (holding other parameters fixed); the range in axis ratios allow for prolate, oblate, and generic triaxiality; and $\phi$ is restricted to $\pm45^\circ$ around the value measured by LM10 ($\phi = 97^\circ$).
+The priors on the remaining halo parameters are taken to be uniform over a conservative domain of realistic values: for $\vhalo$, 100-200~km/s corresponds to a range in solar circular velocities from $\sim$210-250 km/s (holding other parameters fixed); the range in axis ratios allow for prolate, oblate, and generic triaxiality; and $\phi$ is restricted to $\pm45^\circ$ around the true simulation value, $\phi = 97^\circ$.
+
+For all experiments below, we use a Markov Chain Monte Carlo (MCMC) algorithm to sample from the posterior probability distribution given by our model.
+We use an affine-invariant ensemble sampler \citep{goodman10} implemented in the Python programming language \citep{foremanmackey13} and run the walkers in parallel using the message passing interface (MPI).
+We compute the autocorrelation time for each sampled parameter (using \texttt{ACOR}\footnote{www.math.nyu.edu/faculty/goodman/software/acor/}$^{,}$\footnote{https://github.com/dfm/acor}) and thin the chains by taking every $\mathrm{median}(t_{\rm acor})$ sample to ensure the samples are close to independent.
 
@@
 The stars were required to have been stripped after the first pericentric passage and have a present-day distance within $50$ kpc of the Sun.
-Figure~\ref{fig:2.5e8_xz} (left panel) shows a sample of stream particles (grey) and the eight randomly chosen stars (blue) in the Galactic X-Z plane.
+Figure~\ref{fig:sims} (second column from right) shows the eight randomly chosen stars (black crosses) in Galactic coordinates, over-plotted on all other simulated particles (grey points).
 
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=\textwidth]{../../plots/paper2/e8_xz.png}
-\caption{ TODO  }\label{fig:2.5e8_xz}
-\end{center}
-\end{figure}
-
-We leave the potential parameters --- $q_1$, $q_z$, $\phi$, and $\vhalo$ --- free to vary, along with the Lagrange point offset, $\Loffset$, and use Markov Chain Monte Carlo (MCMC) to sample from the posterior probability distribution over these five parameters.
-We use an affine-invariant ensemble sampler \citep{goodman10} implemented in the Python programming language \citep{foremanmackey13} with an ensemble of 64 walkers initialized by sampling from the priors summarized in Table~\ref{tbl:params}.
-We run the walkers in parallel (using MPI todo) for 5000 steps to burn-in, then restart the sampler starting from the final position of the burn-in phase and run for another 10000 steps.
-We compute the autocorrelation time for each parameter [todo: reference] and thin the chains by taking every XXX$\times \mathrm{median}(t_{\rm acor})$ sample to ensure they are close to independent.
-Figure~\ref{fig:trace} shows the walker positions over the 10000 post-burn-in steps for each parameter for the $2.5\times10^8~M_\odot$ simulation; the autocorrelation time for each parameter is displayed on each corresponding panel.
-Figure~\ref{fig:exp1_posterior} shows projections of the posterior probability distribution for the parameters.
+We leave the potential parameters --- $q_1$, $q_z$, $\phi$, and $\vhalo$ --- free to vary, along with the Lagrange point offset, $\Loffset$, and initialize an ensemble of 64 walkers by sampling from the priors summarized in Table~\ref{tbl:params}.
+We run the walkers for 10000 steps to burn-in, then restart the sampler starting from the final position of the burn-in phase and run for another 10000 steps.
+Figure~\ref{fig:trace} shows the walker positions over the 10000 post-burn-in steps for each of the five parameters.
+The autocorrelation time for each parameter is displayed on its corresponding panel and the horizontal lines indicate the true simulation values for the potential parameters.
+Figure~\ref{fig:exp1_posterior} shows projections of the posterior probability distribution for the parameters; with perfect data, the uncertainties on the potential parameters are all $<1\%$.
 
@@
 \subsection{\spitzer + \gaia}
-We take the same eight stars used in the previous experiment and ``observe'' them with optimistic observational uncertainties.
+We next take the same eight stars used in the previous experiment and ``observe'' them with optimistic observational uncertainties.
 We take these stars to be RR Lyrae variables which are known to be excellent distance indicators via the mid-infrared period-luminosity relation \citep[as shown in, e.g.,][]{madore12}.
@@
 Other parameters --- potential parameters, $\Loffset$, and $\tub$ --- are initialized by drawing from the priors summarized in Table~\ref{tbl:params}.
-We again burn in the walkers for 5000 steps and run for 10000 steps post-burn-in.
-The chains are also thinned by taking every XXX$\times \mathrm{median}(t_{\rm acor})$ sample.
+We again burn in the walkers for 10000 steps and run for 10000 steps post-burn-in.
 Figures~\ref{fig:exp2_potential}, \ref{fig:exp2_satellite}, \ref{fig:exp2_particle0} show the marginalized posteriors for the potential, satellite, and a single particle parameters.
@@
 The SMASH survey \citep{smashprop} will be completed by XXX, long before the final \gaia\, data release.
-Thus, we will soon have very precise distance measurements for stars in Sgr and Orphan but poor proper motion constraints.
-[some comments on state of data for Sgr core, proper motions of] Here we study how well we can constrain the potential with this model given stars with missing data.
-We assume the same observational uncertainties for the stars as mentioned above, but assume the proper motion's have not been measured.
+Thus, we will soon have precise distance measurements for stars in the Sgr and Orphan streams, but poor proper motion constraints.
+[some comments on state of data for Sgr core, proper motions of] Here we study how well we can constrain the potential with our model given stars with missing data.
+We use the same observational uncertainties for the stars as mentioned in the previous experiment, but assume the proper motion's have not been measured.
 
 \subsection{\spitzer + current data, no progenitor}
+Most of the streams and debris structures discovered thus far have undetected or uncertain progenitors. [...]
 4 stars in Orphan-like stream, XX stars in GD-1 -like stream
 
-\subsection{Mass comparison}
-[...]
+%\subsection{Mass comparison}
+%[...]
 
@@
 APW: NSF GRFP, Sigma Xi\\
+Astropy\\
 Columbia HPC\\

### 2014-03-18 streams/inference/back_integrate.py
update to use smarter tests

@@ -147,6 +147,36 @@ def back_integration_likelihood(t1, t2, dt, potential, p_gc, s_gc, logm0, logmdo
     r_term = -0.5*((np.log(sigma_r) + (X-alpha*K)**2/sigma_r**2) + \
                    (np.log(sigma_r) + Y**2/sigma_r**2) + \
                    (np.log(sigma_r) + Z**2/sigma_r**2))
+    # r_term_trai = -0.5*((np.log(sigma_r) + (X-2*alpha)**2/sigma_r**2) + \
+    #                     (np.log(sigma_r) + Y**2/sigma_r**2) + \
+    #                     (np.log(sigma_r) + Z**2/sigma_r**2))
+    # r_term = np.zeros_like(r_term_lead)
+    # r_term[K<1] = r_term_lead[K<1]
+    # r_term[K>1] = r_term_trai[K>1]
+
+    # import matplotlib.pyplot as plt
+    # fig,axes = plt.subplots(2,2,figsize=(12,12),sharex=True,sharey=True)
+
+    # XX = np.concatenate((np.random.normal(-alpha,0.5,size=len(X)//2),
+    #                      np.random.normal(alpha,0.5,size=len(X)//2)))
+    # YY = np.random.normal(0.,0.5,size=len(X))
+    # ZZ = np.random.normal(0.,0.5,size=len(X))
+    # alpha = 0.1
+    # axes[0,0].plot(XX, YY, linestyle='none', marker='.', alpha=alpha)
+    # axes[0,1].plot(YY, YY, linestyle='none', marker='.', alpha=alpha)
+    # axes[1,0].plot(XX, ZZ, linestyle='none', marker='.', alpha=alpha)
+    # axes[1,1].plot(YY, ZZ, linestyle='none', marker='.', alpha=alpha)
+
+    # alpha = 0.4
+    # axes[0,0].plot(X, Y, linestyle='none', marker='.', alpha=alpha)
+    # axes[0,1].plot(Y, Y, linestyle='none', marker='.', alpha=alpha)
+    # axes[1,0].plot(X, Z, linestyle='none', marker='.', alpha=alpha)
+    # axes[1,1].plot(Y, Z, linestyle='none', marker='.', alpha=alpha)
+
+    # axes[1,1].set_xlim(-5,5)
+    # axes[1,1].set_ylim(-5,5)
+    # fig.savefig("/Users/adrian/Desktop/derp.png")
+    # sys.exit(0)
 
     sigma_v = 0.5
[…]

### 2014-03-18 streams/inference/core.py

@@ -423,10 +423,11 @@ class StreamModel(object):
             beta = self.particles.beta.truth
 
         # probability the particle was shocked
-        try:
-            p_shocked = param_dict['particles']['p_shocked']
-        except KeyError:
-            p_shocked = self.particles.p_shocked.truth
+        # try:
+        #     p_shocked = param_dict['particles']['p_shocked']
+        # except KeyError:
+        #     p_shocked = self.particles.p_shocked.truth
+        p_shocked = 0.
 
         # heliocentric satellite positions
         s_hel = []

### 2014-03-18 streams/inference/util.py

@@ -16,10 +16,9 @@ import astropy.units as u
 # Project
 from ..integrate import LeapfrogIntegrator
 
-__all__ = ["guess_tail_bit"]
+__all__ = ["particles_x1x2x3", "guess_tail_bit"]
 
-def guess_tail_bit(particles, satellite, potential, t1, t2, dt):
-    """ Guess the tail assigment for each particle. """
+def particles_x1x2x3(particles, satellite, potential, t1, t2, dt, at_tub=True):
     s = satellite
     p = particles
 
@@ -32,9 +31,10 @@ def guess_tail_bit(particles, satellite, potential, t1, t2, dt):
     s_orbit = np.vstack((rs[:,0][:,np.newaxis].T, vs[:,0][:,np.newaxis].T)).T
     p_orbits = np.vstack((rs[:,1:].T, vs[:,1:].T)).T
 
-    t_idx = np.array([np.argmin(np.fabs(ts - t)) for t in p.tub])
-    s_orbit = np.array([s_orbit[jj,0] for jj in t_idx])
-    p_orbits = np.array([p_orbits[jj,ii] for ii,jj in enumerate(t_idx)])
+    if at_tub:
+        t_idx = np.array([np.argmin(np.fabs(ts - t)) for t in p.tub])
+        s_orbit = np.array([s_orbit[jj,0] for jj in t_idx])
+        p_orbits = np.array([p_orbits[jj,ii] for ii,jj in enumerate(t_idx)])
 
     # instantaneous cartesian basis to project into
     x_hat = s_orbit[...,:3] / np.sqrt(np.sum(s_orbit[...,:3]**2, axis=-1))[...,np.newaxis]
@@ -55,8 +55,13 @@ def guess_tail_bit(particles, satellite, potential, t1, t2, dt):
     VY = np.sum(rel_vel * y_hat, axis=-1)
     VZ = np.sum(rel_vel * z_hat, axis=-1)
 
-    Phi = np.arctan2(Y, X)
-    tail_bit = np.ones(p.nparticles)
+    return (X,Y,Z,VX,VY,VZ)
+
+def guess_tail_bit(x1,x2):
+    """ Guess the tail assigment for each particle. """
+
+    Phi = np.arctan2(x2, x1)
+    tail_bit = np.ones(len(x1))
     tail_bit[:] = np.nan
     tail_bit[np.cos(Phi) < -0.5] = -1.
     tail_bit[np.cos(Phi) > 0.5] = 1.

### 2014-03-18 streams/io/core.py
compute cleanness

@@ -141,7 +141,6 @@ def read_hdf5(h5file, nparticles=None, particle_idx=None):
                               truth=true_tail_bit,
                               prior=LogUniformPrior([-2.]*nparticles,[2.]*nparticles))
 
-        # truth=[0.]*nparticles,
         p_shocked = ModelParameter(name="p_shocked",
                                    truth=[0.]*nparticles,
                                    prior=LogUniformPrior([0.]*nparticles,[1.]*nparticles))
@@ -194,8 +193,8 @@ def read_hdf5(h5file, nparticles=None, particle_idx=None):
 
         # TODO HACK? Check tidal radius offset -- will vary?
         s.alpha = ModelParameter(name="alpha",
-                                 truth=1.7,
-                                 prior=LogNormalPrior(1.7,0.5))
+                                 truth=1.5,
+                                 prior=LogNormalPrior(1.5,0.5))
 
         return_dict["satellite"] = s
 

### 2014-03-18 streams/io/sgr.py

@@ -28,7 +28,7 @@ from ..dynamics import Particle, Orbit
 from ..util import streamspath
 from ..coordinates.frame import galactocentric
 from ..potential.lm10 import LawMajewski2010
-from ..inference.util import guess_tail_bit
+from ..inference.util import guess_tail_bit, particles_x1x2x3
 
 __all__ = ["SgrSimulation"]
 
@@ -59,7 +59,7 @@ class SgrSimulation(object):
         self.t1 = self.particle_table.meta["timestep"]
         self.t2 = 0.
 
-    def particles(self, n=None, expr=None, tail_bit=False):
+    def particles(self, n=None, expr=None, tail_bit=False, clean=False):
         """ Return a Particle object with N particles selected from the
             simulation with expression expr.
 
@@ -100,11 +100,33 @@ class SgrSimulation(object):
 
         # guess whether in leading or trailing tail
         if tail_bit:
-            p.meta["tail_bit"] = p.tail_bit = guess_tail_bit(p, self.satellite(),
-                                                             self.potential,
-                                                             self.t1, self.t2, -1.)
+            x1,x2,x3,vx1,vx2,vx3 = particles_x1x2x3(p, self.satellite(),
+                                                    self.potential,
+                                                    self.t1, self.t2, -1)
+            p.meta["tail_bit"] = p.tail_bit = guess_tail_bit(x1,x2)
         else:
             tail_bit = np.ones(p.nparticles)*np.nan
+
+        if clean:
+            if not tail_bit:
+                x1,x2,x3,vx1,vx2,vx3 = particles_x1x2x3(p, self.satellite(),
+                                                        self.potential,
+                                                        self.t1, self.t2, -1)
+                tail_bit = guess_tail_bit(x1,x2)
+            else:
+                tail_bit = p.tail_bit
+
+            # reject any with nan tail_bit
+            idx = ~np.isnan(tail_bit)
+
+            # reject any with |x1| > 3  or |x2| > 2
+            idx &= np.fabs(x1) < 3
+            idx &= np.fabs(x2) < 2
+
+            _X = p._X[idx]
+            meta["tub"] = p.tub[idx]
+            meta["tail_bit"] = tail_bit[idx]
+            p = Particle(_X.T.copy(), frame=p.frame, units=p._internal_units, meta=meta)
 
         return p
 

### 2014-03-18 text/paper2/ms.tex
notes from hogg

@@
 	    David Hendel\altaffilmark{\colum}}
+% TODO: check in with DH about middle initial
 
@@
 			\times p(\W^{(i)} \given \W_\sat, m_0, \dot{m}, \tub^{(i)}, \Phi, \Loffset, \tailbit^{(i)}) \, 
-			\, p(\tub^{(i)}) p(\Phi) \, p(\Loffset)
+			\, p(\tub^{(i)}) \, p(\Phi) \, p(\Loffset)
 \end{multline}
@@
 In what follows, we apply the model described above to ``observations'' of the n-body simulations described in Section~\ref{sec:sims}.
-For all tests in this article, we model the potential of the parent galaxy with the same functional form used in the simulations (Equation~\ref{eq:lm10}) --- we stress that the data are generated with n-body simulations and are not generated with the same model presented here.
+For all tests in this article, we model the potential of the parent galaxy with the same functional form used in the simulations (Equation~\ref{eq:lm10}) --- we stress that the data are generated with n-body simulations and are not generated with anything resembling the likelihood function presented above.
 When recovering the potential, we hold fixed the disk and spheroid parameters (see Table~\ref{tbl:params}), along with two halo parameters: the scale radius, $r_h$, and one of the flattening components in the x-y plane, $q_2$.
@@
 
+% TODO: define walkers and burn-in
+
 \begin{table}[h]
@@
 	\begin{tabular}{l c l} \toprule
-		\multicolumn{3}{l}{{\bf \emph{Global parameters}}} \\
+		\multicolumn{3}{l}{{\bf \emph{Milky Way parameters}}} \\
 		\toprule
@@
 		\toprule
-		\multicolumn{3}{l}{{\bf \emph{Parameters per star}}} \\
+		\multicolumn{3}{l}{{\bf \emph{Star parameters}}} \\
 		\toprule
@@
 		velocity & $\bs{v}$ & $\|\bs{v}\|\sim\mathcal{N}(\mu=0,\sigma^2=4000)$~km/s\\
-		tail assignment & $\beta$ & $\{-1, 1\}$\\ % TODO: how to say equal weight to -1 or 1?
+		tail assignment (binary) & $\beta$ & $\pm1$~equally likely\\ % TODO: how to say equal weight to -1 or 1?
 %		prob. shocked & $\pshock$ & $\mathcal{U}(0,1)$\\
@@
 \end{table}
+% TODO: in caption, redefine N is normal U is uniform. Star parameters are repeated for each star. summarize number of parameters for milky way, progenitor, etc. -- how to read the parameter
 
-\subsection{Perfect data}
+\subsection{Data with negligible uncertainties}
 
@@
 
-We leave the potential parameters --- $q_1$, $q_z$, $\phi$, and $\vhalo$ --- free to vary, along with the Lagrange point offset, $\Loffset$, and initialize an ensemble of 64 walkers by sampling from the priors summarized in Table~\ref{tbl:params}.
+We leave the potential parameters ($q_1$, $q_z$, $\phi$, $\vhalo$) free to vary, along with the Lagrange point offset, $\Loffset$, and initialize an ensemble of 64 MCMC walkers by sampling from the priors summarized in Table~\ref{tbl:params}.
 We run the walkers for 10000 steps to burn-in, then restart the sampler starting from the final position of the burn-in phase and run for another 10000 steps.
-Figure~\ref{fig:trace} shows the walker positions over the 10000 post-burn-in steps for each of the five parameters.
+Figure~\ref{fig:trace} shows the walker positions over the 10000 inference steps for each of the five parameters.
 The autocorrelation time for each parameter is displayed on its corresponding panel and the horizontal lines indicate the true simulation values for the potential parameters.
@@
 
+\acknowledgements
+Rix, Bovy\\
+Aspen?\\
+Gaia data challenge / Surrey / Gaia data team / Anthony Brown \\ % check with AB about citations
+APW: NSF GRFP, Sigma Xi\\
+Astropy\\
+Columbia HPC\\
+DWH was partially supported by NASA (Grant NNX12AI50G) and the NSF (Grant IIS-1124794).
+
 \bibliographystyle{apj}
@@
 
-\acknowledgements
-Aspen?\\
-Gaia data challenge / Surrey / Gaia data team / Anthony Brown \\
-APW: NSF GRFP, Sigma Xi\\
-Astropy\\
-Columbia HPC\\
-
 \end{document}

### 2014-03-18 scripts/figures/paper2_figures.py
trace plots

@@ -15,6 +15,7 @@ import inspect
 import astropy.units as u
 from astropy.constants import G
 from astropy.io.misc import fnpickle, fnunpickle
+import h5py
 import numpy as np
 import matplotlib
 matplotlib.use("agg")
@@ -30,6 +31,7 @@ from streams.util import streamspath
 from streams.coordinates.frame import galactocentric
 from streams.io.sgr import SgrSimulation
 from streams.io import read_hdf5, read_config
+from streams.inference import StreamModel
 from streams.integrate import LeapfrogIntegrator
 from streams.potential.lm10 import LawMajewski2010
 
@@ -51,10 +53,10 @@ plot_path = os.path.join(streamspath, "plots/paper2/")
 if not os.path.exists(plot_path):
     os.mkdir(plot_path)
 
-ext = 'pdf'
+ext = 'png'
 grid_figsize = (14,7.5)
 
-def simulated_streams(**kwargs):
+def simulated_streams():
 
     filename = os.path.join(plot_path, "simulated_streams.{}".format(ext))
     fig,axes = plt.subplots(2,4,figsize=grid_figsize,
@@ -120,7 +122,7 @@ def simulated_streams(**kwargs):
     fig.subplots_adjust(top=0.92, hspace=0.025, wspace=0.1)
     fig.savefig(filename, dpi=200)
 
-def potentials(**kwargs):
+def potentials():
 
     filename = os.path.join(plot_path, "potentials.{}".format(ext))
     fig,axes = plt.subplots(2,4,figsize=grid_figsize)
@@ -208,7 +210,7 @@ def potentials(**kwargs):
     fig.tight_layout(pad=1.5, h_pad=0.)
     fig.savefig(filename)
 
-def Lpts(**kwargs):
+def Lpts():
 
     potential = LawMajewski2010()
     filename = os.path.join(plot_path, "Lpts_r.{}".format(ext))
@@ -356,7 +358,7 @@ def Lpts(**kwargs):
     fig2.subplots_adjust(top=0.92, hspace=0.025, wspace=0.1)
     fig2.savefig(filename2)
 
-def phasespace(**kwargs):
+def phasespace():
 
     filename = os.path.join(plot_path, "Lpts_rv.png")
     fig,axes = plt.subplots(3,4,figsize=(14,12),
@@ -447,7 +449,7 @@ def phasespace(**kwargs):
[…]

### 2014-03-19 text/paper2/ms.tex
hogg comments

@@
 \newcommand{\inverse}[1]{{#1}^{-1}}
+%\newcommand{\project}[1]{{\sffamily #1}}
 
@@
 The dark-matter halo of the Milky Way is expected to be triaxial on large scales
-  and filled with substructure on all scales.
+  and filled with substructure.
 It is hoped that streams or shells of stars produced by tidal disruption
-  will produce very precise measures of the gravitational potential,
+  will provide very precise measures of the gravitational potential,
   to test these predictions.
@@
   and missing data.
-This method advances in a few key points that have not been jointly addressed in prior work: assumptions of integrability, proper accounting for observational uncertainties, and allowing for missing dimensions.
-We test our method with simulated data in a non-integrable triaxial potential, with realistic observational uncertainties for RR Lyrae stars.
+This method is new in that it makes no assumption about the form of the Milky Way 
+  potential, accounts properly for observational uncertainties, and allows for missing data
+  dimensions.
+We test our method with simulated data in a non-integrable triaxial potential, with realistic 
+  observational uncertainties for RR Lyrae stars.
 Importantly, the fake data were generated with an n-body simulation,
@@
 % Results:
-We find that even with a small number (4 to 16) of well-measured stars,
+We find that even with a small number (8 to 16) of well-measured stream stars,
   we can precisely infer properties of the triaxial potential,
@@
 
-Early large-scale, cosmological simulations of galaxy formation in the $\Lambda$CDM paradigm suggested that the spherically-averaged density profiles of dark matter halos follow a universal profile across a huge dynamic range in mass \citep{navarro96}.
+Early large-scale, cosmological simulations of galaxy formation in the $\Lambda$CDM paradigm suggested that the spherically-averaged density profiles of dark matter halos follow a universal profile across a large dynamic range in mass \citep{navarro96}.
 Since then, higher resolution simulations --- both with and without baryons --- have produced dark matter halos that (1) are permeated with substructure on many scales, (2) are triaxial in shape, and (3) have shapes and orientations that vary with radius \citep{dubinski91, jing02, kuhlen07, veraciro11}.
 Dark-matter-only simulations produce triaxial halos \citep{jing02} with large density fluctuations \citep{zemp09}.
-Including baryons tends to soften the triaxiality and graininess in the inner galaxy through a combination of dissipative infall \citep{dubinski94} or cooling \citep{bryan13}.
+Inclusion of baryons tends to soften the triaxiality and graininess in the inner galaxy through a combination of dissipative infall \citep{dubinski94} or cooling \citep{bryan13}.
 These processes combined with the gravity from a baryonic disk or ellipsoid can act to make the inner halo more oblate or spherical, however they do not seem to erase the clumpy, triaxial nature of the outer halo \citep[e.g.,][]{pontzen12}.
@@
 The bulk of the baryonic matter in galaxies spans roughly 5-10\% of the extent of the host dark matter halo.
-Hence, the major observable components of a galaxy are sensitive to the inner portion of its host halo's mass distribution.
-For example, the rotation curves of disk galaxies are excellent probes of the inner mass, especially since matter in disks can be assumed to move on near circular orbits.
-Conversely, measuring the dark matter distribution at large radii is complicated by the low density of visible tracers, observational difficulties, and assumptions about the smoothness and analytic representation.
-Around external galaxies, the extended mass distribution has been probed using a variety of approaches \citep[see][for a a complete and detailed review]{courteau13}.
+Hence, the most observable components of a galaxy are sensitive to the inner portion of the host halos mass distribution.
+For example, the rotation curves of disk galaxies trace the inner mass, especially since matter in disks can be assumed to move on nearly circular orbits.
+Conversely, measurements of the dark matter distribution at large radii is complicated by the low density of visible tracers, observational difficulties, and assumptions about the smoothness and analytic form.
+Around external galaxies, the extended mass distribution has been studied using a variety of approaches \citep[see][for a a complete and detailed review]{courteau13}.
 For example, the kinematics of tracer populations such as globular clusters or planetary nebulae can be used to derive mass estimates under the assumptions that these satellite systems are relaxed and well-mixed \citep[early investigations include][]{mendez01,cote03}.
@@
 %\citep{romanowsky03,douglas07,romanowsky09,napolitano11,lee11}
-Alternatively, the statistical properties of gravitationally lensed background sources around a galaxy can be used to constrain the \emph{projected} shape, orientation, and radial profile of mass \citep[as done by the Lens Structure and Dynamics Survey described in][]{koopmans02}.
+Alternatively, the statistical properties of gravitationally lensed background sources around a galaxy can be used to constrain the \emph{projected} shape, orientation, and radial profile of mass \citep[see, for example, the Lens Structure and Dynamics Survey described in][]{koopmans02}.
 Of course, lensing reconstructions can only be performed for galaxies which closely intersect our line of sight to background sources, but the advent of large photometric catalogues has allowed automatic searches for such chance alignments and significant increases in the number of objects studied in this way \citep[e.g.
@@
 
-Tidal streams are dynamically cold systems --- debris typically have small distributions of energy and angular momentum --- and thus require orders of magnitude fewer tracers than a random sample to get constraints of comparable accuracy.
+Tidal streams are dynamically cold systems --- debris typically have small distributions of energy and angular momentum --- and thus require orders of magnitude fewer tracers than a random sample to get constraints of comparable accuracy to, e.g., Jeans analysis.
 For example, in the simplest case we might assume that debris stars are actually still on the same orbit as their progenitor system \citep[a \emph{wrong} assumption, see e.g.,][]{eyre11}.
@@
 To account for the offset between the orbit of the progenitor and the orbits of the debris stars, methods have been proposed that model the debris as an orbit plus some dispersion or offset, either in phase space \citep[e.g.,][]{eyre09a, varghese11, kuepper12} or action-angle coordinates \citep{eyre11, sanders13b, bovy14, sanders14}.
-Other statistical methods have been proposed \citep[][]{penarrubia12, sanderson14} that may prove powerful when applied to, e.g., data from the \gaia\, mission, where full 6D coordinates will be known for large samples of stars in the halo --- and therefore many debris structures --- but stream membership is not known for all stars.
+We will \emph{not} assume that streams follow the orbit of the progenitor.
+Other statistical methods have been proposed \citep[][]{penarrubia12, apw13, sanderson14} that may prove powerful when applied to, e.g., data from the \gaia\, mission, where full 6D coordinates will be known for large samples of stars in the halo --- and therefore many debris structures --- but stream membership is not known for all stars.
 
-We identify a number of problems with approximate models; each method  mentioned above solves only a subset of these issues:
+We identify a number of problems with existing approximate models; each method mentioned above solves only a subset of these issues:
 \begin{enumerate}
@@
 Thus it is critical for any approximate model of tidal debris to incorporate observational uncertainties and missing dimensions in a consistent and justified way.
-	\item The potential of the inner galaxy is plausibly well-described by a (set of) nearly-integrable potentials, thus the dynamics of collisionless systems is conceivably captured by a transformation to some integral-of-motion-space, e.g., action-angle coordinates.
+	\item The potential of the inner galaxy is conceivably well-described by a (set of) nearly-integrable potentials, thus the dynamics of collisionless systems is conceivably captured by a transformation to some integral-of-motion-space, e.g., action-angle coordinates.
 However there is large uncertainty in the radial profile, shape, orientation, and graininess of the outer halo and the constancy of these parameters over distance.
 To test the degree to which the outer halo is well-fit by integrable potentials, it is problematic to use methods that \emph{assume} integrability.
-	\item Near-future photometric surveys such as \gaia\, and LSST will likely discover many new streams and kinematic associations of stars.
+	\item Near-future photometric surveys such as \gaia\, and the \project{LSST} will likely discover many new streams and kinematic associations of stars.
 Potential recovery methods should be able to simultaneously model multiple streams and incorporate other dynamical constraints.
@@
 For all experiments below, we use a Markov Chain Monte Carlo (MCMC) algorithm to sample from the posterior probability distribution given by our model.
-We use an affine-invariant ensemble sampler \citep{goodman10} implemented in the Python programming language \citep{foremanmackey13} and run the walkers in parallel using the message passing interface (MPI).
+Standard MCMC algorithms (e.g., Metropolis-Hastings) update a single chain while exploring parameter space.
+We instead use an affine-invariant ``ensemble'' sampler \citep{goodman10} that, each step in parameter space, updates the positions of many ``walkers''.
+This algorithm is implemented in the \project{Python} programming language \citep{foremanmackey13} and runs naturally in a parallelized environment, e.g., the message passing interface (MPI).
 We compute the autocorrelation time for each sampled parameter (using \texttt{ACOR}\footnote{www.math.nyu.edu/faculty/goodman/software/acor/}$^{,}$\footnote{https://github.com/dfm/acor}) and thin the chains by taking every $\mathrm{median}(t_{\rm acor})$ sample to ensure the samples are close to independent.
@@
 
-We first test the model using eight particles (stars) --- four from the leading tail, and four from the trailing tail --- randomly sampled from the $2.5\times10^8~M_\odot$ simulation (Section~\ref{sec:sims}), assuming perfect observational data for both the stars and the progenitor.
+We first test the model using eight particles (stars) --- four from the leading tail, and four from the trailing tail --- randomly sampled from the $2.5\times10^8~M_\odot$ simulation (Section~\ref{sec:sims}), assuming perfect knowledge of the observed 6D positions for both the stars and the progenitor.
 The stars were required to have been stripped after the first pericentric passage and have a present-day distance within $50$ kpc of the Sun.

### 2014-03-19 text/paper2/ms.tex
compress notation

@@
 To model the true, 6D, present-day position of the star, $\W$, we first transform to cartesian, Galactocentric coordinates where ($\bs{r},\bs{v}$) and ($\bs{r}_\sat,\bs{v}_\sat$) are the position and velocity of the star and progenitor.
-In cartesian coordinates, the present-day position of the star is taken as having been sampled at $t=\tub$ from an isotropic Gaussian centered on one of the instantaneous, effective Lagrange points in position, and an isotropic Gaussian centered on the origin in velocity, evolved to present day in the parent potential, $\Phi$.
+In cartesian coordinates, the present-day position of the star is taken as having been sampled at $t=\tub$ from an isotropic Gaussian centered on one of the instantaneous, effective Lagrange points in position, and an isotropic Gaussian centered on the origin in velocity, evolved to present day in the parent potential, $\Phi$, parametrized by the vector $\bs{\Theta}_\Phi$.
 We assume that once the star becomes ``unbound'' from the satellite the potential of the satellite can be ignored, and thus treat the star as a test particle.\footnote{This assumption breaks down for sufficiently large-mass progenitors, somewhere between $\sim10^8-10^9~$M$_\odot$.} The satellite mass enters the model through the tidal radius; a changing satellite mass will skew the positions of the effective Lagrange points and velocity scale.
@@
 Finally, we add two additional parameters: a constant, global offset, $\Loffset$, that scales the position of the effective Lagrange points relative to the classical tidal radius, and a binary parameter $\tailbit$, that is either $-1$ or $+1$ depending on whether the star is in the leading or trailing tail.
+To compress notation, we pack all progenitor parameters ($\W_\sat, m_0, \dot{m}, \Loffset$) into the vector $\bs{\Theta}_\sat$, and all star parameters ($\W, \tailbit, \tub$) into the vector $\bs{\Theta}$.
 The likelihood is then:
 \begin{align}
-	p(\W \given \W_\sat, m_0, \dot{m}, \tub, \Phi, \Loffset, \tailbit) &= p(\bs{r} \given \bs{r}_\sat, m_0, \dot{m}, \tub, \Phi, \Loffset, \tailbit)\,p(\bs{v} \given \bs{v}_\sat, m_0, \dot{m}, \tub, \Phi)\,\left\vert\J\right\vert_{t=\tub}\\
-	p(\bs{r} \given \bs{r}_\sat, m_0, \dot{m}, \tub, \Phi, \Loffset, \tailbit) &= \mathcal{N}(\bs{r} \given \bs{r}_\sat - \Loffset\,\tailbit\,\rtide\,\hat{x}_1, \rtide^2)_{t=\tub}\\
-	p(\bs{v} \given \bs{v}_\sat, m_0, \dot{m}, \tub, \Phi) &= \mathcal{N}(\bs{v} \given \bs{v}_\sat, \sigma^2_v)_{t=\tub}
+	p(\bs{\Theta} \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi) &= p(\bs{r} \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi)\,p(\bs{v} \given\bs{\Theta}_\sat, \bs{\Theta}_\Phi)\,\left\vert\J\right\vert_{t=\tub}\\
+	p(\bs{r} \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi) &= \mathcal{N}(\bs{r} \given \bs{r}_\sat - \Loffset\,\tailbit\,\rtide\,\hat{x}_1, \rtide^2)_{t=\tub}\\
+	p(\bs{v} \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi) &= \mathcal{N}(\bs{v} \given \bs{v}_\sat, \sigma^2_v)_{t=\tub}
 \end{align}
+%\begin{align}
+%	p(\W \given \W_\sat, m_0, \dot{m}, \tub, \Phi, \Loffset, \tailbit) &= p(\bs{r} \given \bs{r}_\sat, m_0, \dot{m}, \tub, \Phi, \Loffset, \tailbit)\,p(\bs{v} \given \bs{v}_\sat, m_0, \dot{m}, \tub, \Phi)\,\left\vert\J\right\vert_{t=\tub}\\
+%	p(\bs{r} \given \bs{r}_\sat, m_0, \dot{m}, \tub, \Phi, \Loffset, \tailbit) &= \mathcal{N}(\bs{r} \given \bs{r}_\sat - \Loffset\,\tailbit\,\rtide\,\hat{x}_1, \rtide^2)_{t=\tub}\\
+%	p(\bs{v} \given \bs{v}_\sat, m_0, \dot{m}, \tub, \Phi) &= \mathcal{N}(\bs{v} \given \bs{v}_\sat, \sigma^2_v)_{t=\tub}
+%\end{align}
 where $\rtide$ is given by Eq.~\ref{eq:tidalradius}, and $\sigma_v$ by Eq.~\ref{eq:velscale}.
@@
 Assuming each star is an independent tracer, the full posterior for many stars is then just the product of the individual likelihoods with the priors:
+%\begin{multline}
+%	p(\Phi, \{\W^{(i)}\}, \W_\sat, m_0, \dot{m}, \tub, \Loffset, \tailbit, \given \{\D^{(i)}\}, \D_\sat) = \\
+%		\shoveleft{\hspace{1em}\frac{1}{\mathcal{Z}} \, \prod_i \, p(\D^{(i)} \given \W^{(i)}) \, p(\D_\sat \given \W_\sat)}\\
+%			\times p(\W^{(i)} \given \W_\sat, m_0, \dot{m}, \tub^{(i)}, \Phi, \Loffset, \tailbit^{(i)}) \, 
+%			\, p(\tub^{(i)}) \, p(\Phi) \, p(\Loffset)
+%\end{multline}
 \begin{multline}
-	p(\Phi, \{\W^{(i)}\}, \W_\sat, m_0, \dot{m}, \tub, \Loffset, \tailbit, \given \{\D^{(i)}\}, \D_\sat) = \\
-		\shoveleft{\hspace{1em}\frac{1}{\mathcal{Z}} \, \prod_i \, p(\D^{(i)} \given \W^{(i)}) \, p(\D_\sat \given \W_\sat)}\\
-			\times p(\W^{(i)} \given \W_\sat, m_0, \dot{m}, \tub^{(i)}, \Phi, \Loffset, \tailbit^{(i)}) \, 
-			\, p(\tub^{(i)}) \, p(\Phi) \, p(\Loffset)
+	p(\{\bs{\Theta}^{(i)}\}, \bs{\Theta}_\sat, \bs{\Theta}_\Phi \given \{\D^{(i)}\}, \D_\sat) = \\
+		\shoveleft{\frac{1}{\mathcal{Z}} \, \prod_i \, p(\D^{(i)} \given \W^{(i)}) \, p(\D_\sat \given \W_\sat)} 
+			p(\bs{\Theta}^{(i)} \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi)) \, 
+			\, p(\bs{\Theta}_\sat)\,p(\bs{\Theta}_\Phi)
 \end{multline}

### 2014-03-19 text/paper2/ms.tex
added thetas to table
changed experiment names

@@
 \subsection{Probabilistic model}
+% TODO: define what we mean by "Model"
 We observe the 6D position of a star, $\D$, in heliocentric coordinates --- e.g., the measured position on the sky, ($l$, $b$); distance, $d$; proper motions, ($\mu_l$, $\mu_b$); and line-of-sight velocity, $v_r$ --- and have determined through some other means that this star was once part of a progenitor system (e.g., satellite galaxy) with mass $m(t)$ that is disrupting and forming a cold debris structure in the potential, $\Phi$, of the parent galaxy (e.g., the Milky Way).
@@
 		\shoveleft{\frac{1}{\mathcal{Z}} \, \prod_i \, p(\D^{(i)} \given \W^{(i)}) \, p(\D_\sat \given \W_\sat)} 
-			p(\bs{\Theta}^{(i)} \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi)) \, 
-			\, p(\bs{\Theta}_\sat)\,p(\bs{\Theta}_\Phi)
+			p(\bs{\Theta}^{(i)} \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi) \, 
+			p(\bs{\Theta}) \, p(\bs{\Theta}_\sat)\,p(\bs{\Theta}_\Phi)
 \end{multline}
@@
 	\begin{tabular}{l c l} \toprule
-		\multicolumn{3}{l}{{\bf \emph{Milky Way parameters}}} \\
+		\multicolumn{3}{l}{{\bf \emph{Milky Way parameters} ($\bs{\Theta}_\Phi$)}} \\
 		\toprule
@@
 		\toprule
-		\multicolumn{3}{l}{{\bf \emph{Progenitor parameters}}} \\
+		\multicolumn{3}{l}{{\bf \emph{Progenitor parameters} ($\bs{\Theta}_\sat$)}} \\
 		\toprule
@@
 		\toprule
-		\multicolumn{3}{l}{{\bf \emph{Star parameters}}} \\
+		\multicolumn{3}{l}{{\bf \emph{Star parameters} ($\bs{\Theta}^{(i)}$)}} \\
 		\toprule
@@
 
-\subsection{\spitzer + \gaia}
+\subsection{Data with near-future uncertainties}
 We next take the same eight stars used in the previous experiment and ``observe'' them with optimistic observational uncertainties.
@@
 
-\subsection{\spitzer + current data}
+\subsection{Precise distance measurements with missing proper motions}
 The SMASH survey \citep{smashprop} will be completed by XXX, long before the final \gaia\, data release.
@@
 
-\subsection{\spitzer + current data, no progenitor}
+\subsection{Precise distance measurements of stars with no observed progenitor}
 Most of the streams and debris structures discovered thus far have undetected or uncertain progenitors. [...]

### 2014-03-19 text/paper2/ms.tex
changed bases

@@
 \begin{center}
-\includegraphics[width=\textwidth]{../../plots/paper2/simulated_streams.pdf}
+\includegraphics[width=\textwidth]{../../plots/paper2/simulated_streams.png}
 \caption{ Particle positions (grey dots) in Galactocentric cartesian coordinates from the final timestep of four n-body simulations with the same progenitor orbit initial conditions over a range of progenitor masses (columns).
@@
 In a triaxial potential, the orbital plane of the satellite is not fixed but we still expect there to be \emph{effective} Lagrange points for a spherical satellite system along the line of centers connecting the origin of the parent potential to the satellite.
-We thus proceed by defining a coordinate system relative to the position and velocity of the progenitor, ($\bs{r}_p$,$\bs{v}_p$), rotated into the instantaneous orbital plane defined by $\bs{r}_p \times \bs{v}_p$, and normalized by the instantaneous tidal radius for positions (Eq.~\ref{eq:tidalradius}) and instantaneous velocity scale for velocity (Eq.~\ref{eq:velscale}).
+We thus proceed by defining a coordinate system relative to the position and velocity of the progenitor, ($\bs{r}_p$,$\bs{v}_p$), rotated into the instantaneous orbital plane defined by $\bs{L} = \bs{r}_p \times \bs{v}_p$, and normalized by the instantaneous tidal radius for positions (Eq.~\ref{eq:tidalradius}) and instantaneous velocity scale for velocity (Eq.~\ref{eq:velscale}).
 % where $\hat{\bs r}_p = {\bs r}_p / |{\bs r}_p|$ and $\hat{\bs v}_p = {\bs v}_p / |{\bs v}_p|$
@@
 \begin{align}
-	\hat{x}_1 &= \frac{\bs{r}_p}{\|\bs{r}_p\|}\\
-	\hat{x}_2 &= -\hat{x}_1 \times \hat{x}_3\\
-	\hat{x}_3 &= \frac{\bs{r}_p \times \bs{v}_p}{\|\bs{r}_p \times \bs{v}_p\|}.
+	\hat{\bs{x}}_1 &= \frac{\bs{r}_p}{\|\bs{r}_p\|}\\
+	\hat{\bs{x}}_2 &= \frac{\bs{L} \times \hat{\bs{x}}_1}{\|\bs{L} \times \hat{\bs{x}}_1\|} = \hat{\bs{x}}_3 \times \hat{\bs{x}}_1\\
+	\hat{\bs{x}}_3 &= \hat{\bs{L}} = \frac{\bs{r}_p \times \bs{v}_p}{\|\bs{r}_p \times \bs{v}_p\|}.
 \end{align}
@@
 
-\begin{figure*}[!ht]
-	\captionsetup[subfigure]{labelformat=empty}
-        \centering
-        \subfloat[]{
-                \includegraphics[width=0.5\textwidth]{../../plots/paper2/trace_0.png}
-                \label{fig:trace0}
-	}
-        \subfloat[]{
-                \includegraphics[width=0.5\textwidth]{../../plots/paper2/trace_1.png}
-                \label{fig:trace0}
-	}
-
-        \subfloat[]{
-                \includegraphics[width=0.5\textwidth]{../../plots/paper2/trace_2.png}
-                \label{fig:trace0}
-	}
-        \subfloat[]{
-                \includegraphics[width=0.5\textwidth]{../../plots/paper2/trace_3.png}
-                \label{fig:trace0}
-	}
-
-        \subfloat[]{
-                \includegraphics[width=0.5\textwidth]{../../plots/paper2/trace_4.png}
-                \label{fig:trace0}
-	}
-        \caption{TODO: mcmc trace plots}\label{fig:trace}
-\end{figure*}
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[width=\textwidth]{../../plots/paper2/mcmc_trace.png}
+\caption{ TODO: mcmc trace plots  }\label{fig:trace}
+\end{center}
+\end{figure}
 
@@
 For this experiment, we (1) assume the stars are bright distance indicators --- RR Lyrae --- such that the fractional distance uncertainty is $2\%$; (2) neglect the uncertainty in angular position (for a typical RR Lyrae at 50~kpc this is $\sim$$10^{-7}$~deg for \gaia); (3) assume we can measure radial velocities to these stars with 5~km/s uncertainty; and (4) compute the sky-averaged \gaia\, proper-motion uncertainty for each star assuming an F0V spectral type using the \texttt{PyGaia}\footnote{https://github.com/agabrown/PyGaia} code and use this uncertainty for both components of proper motion.
-The observed positions and velocities of these stars are shown in Figure~\ref{fig:2.5e8_xz} (red, compare to the unobserved black points).
 We further assume that we know the tail assignment for each star, $\beta$.

### 2014-03-19 scripts/test_infer_potential.py
changed sample

@@ -38,7 +38,7 @@ minimum_config = """
 name: test
 data_file: data/observed_particles/2.5e8.hdf5
 nparticles: 8
-particle_idx: [611, 352, 67, 177, 681, 1762, 986, 721]
+particle_idx: [582, 585, 231, 300, 637, 508, 404, 1460]
 
 potential:
     class_name: LawMajewski2010

### 2014-03-19 text/paper2/ms.tex
change step number

@@
 We leave the potential parameters ($q_1$, $q_z$, $\phi$, $\vhalo$) free to vary, along with the Lagrange point offset, $\Loffset$, and initialize an ensemble of 64 MCMC walkers by sampling from the priors summarized in Table~\ref{tbl:params}.
-We run the walkers for 10000 steps to burn-in, then restart the sampler starting from the final position of the burn-in phase and run for another 10000 steps.
-Figure~\ref{fig:trace} shows the walker positions over the 10000 inference steps for each of the five parameters.
+We run the walkers for 10000 steps to burn-in, then restart the sampler starting from the final position of the burn-in phase and run for another 5000 steps.
+Figure~\ref{fig:trace} shows the walker positions over the 5000 inference steps for each of the five parameters.
 The autocorrelation time for each parameter is displayed on its corresponding panel and the horizontal lines indicate the true simulation values for the potential parameters.

### 2014-03-20 text/paper2/ms.tex
acknowledgements

@@
 Rix, Bovy\\
-Aspen?\\
 Gaia data challenge / Surrey / Gaia data team / Anthony Brown \\ % check with AB about citations
-APW: NSF GRFP, Sigma Xi\\
-Astropy\\
+APW is supported by a National Science Foundation Graduate Research
+Fellowship under Grant No.\ 11-44155. This work was supported in part by 
+the National Science Foundation under Grant No. PHYS-1066293 and the 
+hospitality of the Aspen Center for Physics.\\
+DWH was partially supported by NASA (Grant NNX12AI50G) and the NSF (Grant IIS-1124794).\\
+This research made use of Astropy, a community-developed core Python package for Astronomy \citep{astropy13}. \\
 Columbia HPC\\
-DWH was partially supported by NASA (Grant NNX12AI50G) and the NSF (Grant IIS-1124794).
 

### 2014-03-20 streams/inference/core.py
nan and inf check

@@ -217,6 +217,30 @@ class StreamModel(object):
             ix1 = 0
             for group_name,group in self.parameters.items():
                 for param_name,param in group.items():
+
+                    if param_name in heliocentric.coord_names:
+                        err = getattr(self, group_name).errors[param_name].decompose(usys).value
+                        if start_truth:
+                            val = getattr(self, "true_"+group_name)[param_name]\
+                                        .decompose(usys).value
+                            err /= 2.
+                        else:
+                            val = getattr(self, group_name)[param_name]\
+                                        .decompose(usys).value
+
+                        if np.any(np.isinf(err)):
+                            err[np.isinf(err)] = 0.001*val
+
+                        try:
+                            prior = self._prior_cache[(group_name,param_name)]
+                        except KeyError:
+                            prior = LogNormalPrior(val, err)
+                            self._prior_cache[(group_name,param_name)] = prior
+
+                        p0[ii,ix1:ix1+param.size] = np.ravel(prior.sample().T)
+                        ix1 += param.size
+                        continue
+
                     if start_truth:
                         prior = param._prior
                         if hasattr(prior, 'a') and hasattr(prior, 'b'):
@@ -407,8 +431,9 @@ class StreamModel(object):
         W_ij = np.array(W_ij)
         D_ij = np.array(D_ij)
         sig_ij = np.array(sig_ij)
-        data_like = -W_ij.shape[0]/2.*np.log(2*np.pi) - \
-                     np.sum(0.5*(np.log(sig_ij**2) + ((W_ij - D_ij)/sig_ij)**2), axis=0)
+        data_like = 0.5*(np.log(sig_ij**2) + ((W_ij - D_ij)/sig_ij)**2)
+        data_like[~np.isfinite(data_like)] = 0.
+        data_like = -W_ij.shape[0]/2.*np.log(2*np.pi) - np.sum(data_like, axis=0)
 
         # particle unbinding time
         try:
@@ -453,8 +478,9 @@ class StreamModel(object):
         W_j = np.array(W_j)
         D_j = np.squeeze(D_j)
         sig_j = np.squeeze(sig_j)
-        sat_like = -W_j.shape[0]/2.*np.log(2*np.pi) - \
-                    np.sum(0.5*(np.log(sig_j**2) + ((W_j - D_j)/sig_j)**2), axis=0)
+        sat_like = 0.5*(np.log(sig_j**2) + ((W_j - D_j)/sig_j)**2)
[…]

### 2014-03-20 text/paper2/ms.tex
data quality

@@
 Thus, we will soon have precise distance measurements for stars in the Sgr and Orphan streams, but poor proper motion constraints.
-[some comments on state of data for Sgr core, proper motions of] Here we study how well we can constrain the potential with our model given stars with missing data.
+[some comments on state of data for Sgr core, proper motions of] Here we study how well we can constrain the potential with our model given stars with missing proper motions.
 We use the same observational uncertainties for the stars as mentioned in the previous experiment, but assume the proper motion's have not been measured.

### 2014-03-20 scripts/observe_simulation.py
missing dims

@@ -229,13 +229,18 @@ OR
 python scripts/observe_simulation.py -v --class_name=SgrSimulation --expr='tub!=0' \
 --file=/Users/adrian/projects/streams/data/observed_particles/2.5e8_exp3.hdf5 \
 --seed=42 --path="sgr_nfw/M2.5e+08" --snapfile="SNAP113" --overwrite
+
+python scripts/observe_simulation.py -v --class_name=SgrSimulation --expr='tub!=0' \
+--file=/Users/adrian/projects/streams/data/observed_particles/2.5e8_exp4.hdf5 \
+--seed=42 --path="sgr_nfw/M2.5e+08" --snapfile="SNAP113" --overwrite
     """
 
     # TODO: class kwargs
     # particle_error_model=current_data_errors, satellite_error_model=current_data_errors,
     # particle_error_model=gaia_spitzer_errors, satellite_error_model=gaia_spitzer_errors,
     observe_simulation(args.class_name,
-        particle_error_model=current_data_errors, satellite_error_model=current_data_errors,
+        particle_error_model=lambda p: current_data_errors(p, missing_dims=["mul","mub"]),
+        satellite_error_model=lambda p: current_data_errors(p, missing_dims=["l","b","d","mul","mub", "vr"]),
         selection_expr=args.expr, output_file=args.output_file,
         overwrite=args.overwrite, seed=args.seed,
         class_kwargs=dict(path=args.path,snapfile=args.snapfile))

### 2014-03-20 streams/observation/gaia.py
msising dims

@@ -127,7 +127,7 @@ def gaia_spitzer_errors(particles):
 
     return dict(l=l, b=b, d=d, mul=mul,mub=mub,vr=vr)
 
-def current_data_errors(particles):
+def current_data_errors(particles, missing_dims=[]):
     """ Given particles in heliocentric frame, return an array of
         Gaia + Spitzer errors (same shape as particles._X).
     """
@@ -157,4 +157,9 @@ def current_data_errors(particles):
     # radial velocity
     vr = 10.*u.km/u.s * np.ones_like(particles["vr"].value)
 
-    return dict(l=l, b=b, d=d, mul=mul,mub=mub,vr=vr)
+    errs = dict(l=l, b=b, d=d, mul=mul,mub=mub,vr=vr)
+
+    for dim in missing_dims:
+        errs[dim] = np.inf*errs[dim].unit
+
+    return errs

### 2014-03-20 text/paper2/ms.tex
citation

@@
 In general, the data for the star and progenitor will have missing dimensions and significant uncertainties at distances typical to the Galactic halo, thus we define $\W$ and $\W_\sat$ as the true 6D, heliocentric positions of the star and progenitor and will include these in our model.
-By including these positions in the model, this model will work even when the star has missing dimensions or the progenitor location is unknown as in, e.g., the Orphan stream \citep{some orphan paper}.
+By including these positions in the model, this model will work even when the star has missing dimensions or the progenitor location is unknown as in, e.g., the Orphan stream \citep{belokurov07}.
 

### 2014-03-21 text/paper2/ms.tex
acknowledgements

@@
 \newcommand{\vhalo}{v_{\rm h}}
+\newcommand{\rewinder}{\emph{Rewinder}}
+\newcommand{\potp}{$q_1,q_z,\phi,\vhalo$}
 
@@
 %[][hereafter \paperone]
-In \citet{apw13}, we introduced a simple method (\emph{Rewinder}) for using individual stars associated with tidal debris combined with knowledge about the mass and orbit of the progenitor to constrain properties of the host galaxy potential.
+In \citet{apw13}, we introduced a simple method for using individual stars associated with tidal debris combined with knowledge about the mass and orbit of the progenitor to constrain properties of the host galaxy potential.
 The method exploits the relationship between the phase-space distribution of debris and (measurable) properties of the progenitor system (e.g., the tidal radius and escape velocity).
 Specifically, a better model for the potential is one in which orbits of test particle stars (integrated backwards from their present position) came close (within the tidal radius in position, and escape velocity in velocity) to the orbit of the progenitor.
-In this article, we use these simple scalings --- which depend only on the mass and orbit of the progenitor and parent potential --- to present a fully probabilistic method for modeling tidal streams in any underlying potential given data with realistic uncertainties.
-The method relies on numerical orbit integration in ordinary phase-space and can thus incorporate arbitrarily complex models for the parent potential.
-The model uses individual stars as constraints --- not density matching --- and can therefore handle very small samples of well-measured stars (e.g., 4-16).
-The model trivially extends to incorporating multiple streams, debris structures, and other kinematic information.
+In this article, we use these simple scalings --- which depend only on the mass and orbit of the progenitor and parent potential --- to present a fully probabilistic method (\rewinder) for modeling tidal streams in any underlying potential given data with realistic uncertainties.
+By ``model'' here we mean a justified, parametrized likelihood function with priors on the parameters.
+\rewinder\, relies on numerical orbit integration in ordinary phase-space and can thus incorporate arbitrarily complex models for the parent potential.
+Individual stars are used as constraints, thus \rewinder\, can handle very small samples of well-measured stars (e.g., 4-16).
+This model trivially extends to incorporating multiple streams, debris structures, and other kinematic information.
 
@@
 
-We leave the potential parameters ($q_1$, $q_z$, $\phi$, $\vhalo$) free to vary, along with the Lagrange point offset, $\Loffset$, and initialize an ensemble of 64 MCMC walkers by sampling from the priors summarized in Table~\ref{tbl:params}.
+We leave the potential parameters ($q_1,q_z,\phi,\vhalo$) free to vary, along with the Lagrange point offset, $\Loffset$, and initialize an ensemble of 64 MCMC walkers by sampling from the priors summarized in Table~\ref{tbl:params}.
 We run the walkers for 10000 steps to burn-in, then restart the sampler starting from the final position of the burn-in phase and run for another 5000 steps.
@@
 \begin{center}
-\includegraphics[width=\textwidth]{../../plots/paper2/mcmc_trace.png}
-\caption{ TODO: mcmc trace plots  }\label{fig:trace}
+\includegraphics[width=0.9\textwidth]{../../plots/paper2/mcmc_trace.png}
+\caption{ Positions of all 64 MCMC walkers (see Section~\ref{sec:experiments}) for each parameter at each step of the inference (black lines).
+Horizontal green lines show the input values of the potential parameters (\potp).
+The MCMC chains do not display any bulk movement or stray walkers, implying, by eye, that the chains have converged.
+Projections of the sample densities are shown in Figure~\ref{fig:exp1_posterior}.
+}\label{fig:trace}
 \end{center}
@@
 \includegraphics[width=\textwidth]{../../plots/paper2/exp1_posterior.png}
-\caption{ TODO  }\label{fig:exp1_posterior}
+\caption{ Projections of the posterior probability distribution over the four potential parameters (\potp) and Lagrange point offset ($\alpha$) assuming negligible uncertainties on the observed phase-space coordinates of eight stars and the progenitor, visualized as two-dimensional histograms of MCMC samples.
+Solid contours (black lines) show approximately 1$\sigma$ and 2$\sigma$ levels of the distribution.
+Vertical and horizontal lines (blue) show the true, input values for the potential parameters used in the n-body simulations.
+For the potential parameters, the axis ranges were chosen to be $\pm$5\% around the true values.
+Note that any biases in the inferred values are $\lesssim1\%$.
+}\label{fig:exp1_posterior}
 \end{center}
@@
 We take these stars to be RR Lyrae variables which are known to be excellent distance indicators via the mid-infrared period-luminosity relation \citep[as shown in, e.g.,][]{madore12}.
-$\sim$100 RR Lyrae associated with the Sgr stream and $\sim$30 associated with the Orphan stream will be observed with \spitzer as part of the SMASH survey \citep{smashprop} with expected fractional distance uncertainties around $\sim$2\%.
+Nearly 100 RR Lyrae associated with the Sgr stream and $\sim$30 associated with the Orphan stream will be observed with \spitzer as part of the SMASH survey \citep{smashprop} with expected fractional distance uncertainties around $\sim$2\%.
 These stars will also be included in the \gaia\, proper motion catalog: at a distance of $\sim$50~kpc, a typical RR Lyrae will have a tangential velocity error around $\sim$20~km/s.
@@
 \begin{center}
-\includegraphics[width=\textwidth]{../../plots/paper2/exp2_particle0.png}
-\caption{ TODO  }\label{fig:exp2_particle0}
+\includegraphics[width=\textwidth]{../../plots/paper2/exp2_particle.png}
+\caption{ TODO  }\label{fig:exp2_particle}
 \end{center}
@@
 \acknowledgements
-Rix, Bovy\\
-Gaia data challenge / Surrey / Gaia data team / Anthony Brown \\ % check with AB about citations
-APW is supported by a National Science Foundation Graduate Research
-Fellowship under Grant No.\ 11-44155. This work was supported in part by 
-the National Science Foundation under Grant No. PHYS-1066293 and the 
-hospitality of the Aspen Center for Physics.\\
-DWH was partially supported by NASA (Grant NNX12AI50G) and the NSF (Grant IIS-1124794).\\
-This research made use of Astropy, a community-developed core Python package for Astronomy \citep{astropy13}. \\
-Columbia HPC\\
+We thank Anthony Brown (Leiden) for insight on the \gaia\, data quality and for providing the open source \texttt{PyGaia} code for computing predicted uncertainties.
+We thank the organizers of the Gaia Challenge (2013) and the University of Surrey for hospitality during this workshop.
+We thank Hans-Walter Rix (MPIA) and Jo Bovy (IAS) for ...
+[TODO: HOGG].
+APW is supported by a National Science Foundation Graduate Research Fellowship under Grant No.\ 11-44155.
+This work was supported in part by the National Science Foundation under Grant No.
+PHYS-1066293.
+DWH was partially supported by NASA (Grant NNX12AI50G) and the NSF (Grant IIS-1124794).
+This research made use of Astropy, a community-developed core Python package for Astronomy \citep{astropy13}.
+This work additionally relied on Columbia University's \emph{hotfoot} and \emph{yeti} compute clusters, and we acknowledge the Columbia HPC support staff for assistance. \\
 

### 2014-03-21 text/paper2/ms.tex
conclusions

@@
 Other parameters --- potential parameters, $\Loffset$, and $\tub$ --- are initialized by drawing from the priors summarized in Table~\ref{tbl:params}.
-We again burn in the walkers for 10000 steps and run for 10000 steps post-burn-in.
-Figures~\ref{fig:exp2_potential}, \ref{fig:exp2_satellite}, \ref{fig:exp2_particle0} show the marginalized posteriors for the potential, satellite, and a single particle parameters.
-[....]
+We burn in the walkers for 10000 steps and run for 10000 inference steps.
+Figures~\ref{fig:exp2_potential}, \ref{fig:exp2_satellite}, \ref{fig:exp2_particle} show the marginalized posteriors for the potential parameters, satellite parameters, and parameters for a single particle.
 
@@
 \subsection{Precise distance measurements with missing proper motions}
-The SMASH survey \citep{smashprop} will be completed by XXX, long before the final \gaia\, data release.
+The SMASH survey \citep{smashprop} will be completed in 2014, long before the final \gaia\, data release.
 Thus, we will soon have precise distance measurements for stars in the Sgr and Orphan streams, but poor proper motion constraints.
-[some comments on state of data for Sgr core, proper motions of] Here we study how well we can constrain the potential with our model given stars with missing proper motions.
-We use the same observational uncertainties for the stars as mentioned in the previous experiment, but assume the proper motion's have not been measured.
+SMASH also targets several RR Lyrae in the Sgr core, which enable a high-precision distance measurement of the progenitor.
+Though proper motions have been measured for the Sgr progenitor, we now consider a case in which we have high-precision distances to stream stars and the progenitor, 15~km/s radial velocity uncertainty, and missing proper motions for all stars and progenitor.
+
+As with the previous experiment, this experiment includes 41 model parameters in total.
+We again use an ensemble of 256 walkers, but this time initialize the walkers by sampling the coordinate parameters from Gaussians centered on the true positions with variances set to 1\% of the parameter values.
+Other parameters --- potential parameters, $\Loffset$, and $\tub$ --- are initialized by drawing from the priors summarized in Table~\ref{tbl:params}.
+We again burn in the walkers for 10000 steps and run for 10000 inference steps.
+Figures~\ref{fig:exp3_potential}, \ref{fig:exp3_satellite}, \ref{fig:exp3_particle} show the marginalized posteriors for the potential parameters, satellite parameters, and parameters for a single particle.
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[width=\textwidth]{../../plots/paper2/exp3_potential.png}
+\caption{ Same as Figure~\ref{fig:exp2_potential} but for data with no proper motion measurements. }\label{fig:exp3_potential}
+\end{center}
+\end{figure}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[width=\textwidth]{../../plots/paper2/exp3_satellite.png}
+\caption{ Same as Figure~\ref{fig:exp2_satellite} but for data with no proper motion measurements.  }\label{fig:exp3_satellite}
+\end{center}
+\end{figure}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[width=\textwidth]{../../plots/paper2/exp3_particle.png}
+\caption{ Same as Figure~\ref{fig:exp2_particle} but for data with no proper motion measurements.  }\label{fig:exp3_particle}
+\end{center}
+\end{figure}
 
 \subsection{Precise distance measurements of stars with no observed progenitor}
-Most of the streams and debris structures discovered thus far have undetected or uncertain progenitors. [...]
+Most of the streams and debris structures discovered thus far have undetected or unconfirmed progenitors. [...]
 4 stars in Orphan-like stream, XX stars in GD-1 -like stream
@@
 
+\section{Conclusions}
+We have presented a probabilistic model (\rewinder) --- a likelihood function and with priors on the parameters --- for using stars observed in tidal streams to constrain properties of any underlying gravitational potential.
+\rewinder\, relies on direct orbit integration and not on computing conserved quantities (e.g., actions) and can thus be used with arbitrarily complex (e.g., time-dependent or clumpy) forms of the potential.
+We have performed several experiments ...
+
 \acknowledgements

### 2014-03-21 text/paper2/ms.tex
burn-in

@@
 	    David Hendel\altaffilmark{\colum}}
-% TODO: check in with DH about middle initial
 
@@
 These processes combined with the gravity from a baryonic disk or ellipsoid can act to make the inner halo more oblate or spherical, however they do not seem to erase the clumpy, triaxial nature of the outer halo \citep[e.g.,][]{pontzen12}.
-This can lead to radially-dependent axis ratios, orientation, and smoothness, thus suggesting complicated mass distributions for Milky Way-like galaxies that are not easily represented by integrable potentials.
+This can lead to radially-dependent axis ratios, orientation, and smoothness, thus suggesting that the true mass distributions around Milky Way-like galaxies are not easily represented by integrable potentials.
 
@@
 Thus it is critical for any approximate model of tidal debris to incorporate observational uncertainties and missing dimensions in a consistent and justified way.
-	\item The potential of the inner galaxy is conceivably well-described by a (set of) nearly-integrable potentials, thus the dynamics of collisionless systems is conceivably captured by a transformation to some integral-of-motion-space, e.g., action-angle coordinates.
+	\item The potential of the inner galaxy is conceivably well-described by a (set of) nearly-integrable potentials, thus the dynamics of collisionless systems is plausibly captured by a transformation to some integral-of-motion-space, e.g., action-angle coordinates.
 However there is large uncertainty in the radial profile, shape, orientation, and graininess of the outer halo and the constancy of these parameters over distance.
-To test the degree to which the outer halo is well-fit by integrable potentials, it is problematic to use methods that \emph{assume} integrability.
+Properly describing the potential may require complex models where computing actions is non-trivial or even impossible if it is not representable by perturbing a nearby set of canonical coordinates.
 	\item Near-future photometric surveys such as \gaia\, and the \project{LSST} will likely discover many new streams and kinematic associations of stars.
@@
 \begin{align}
-	\hat{\bs{x}}_1 &= \frac{\bs{r}_p}{\|\bs{r}_p\|}\\
+	\hat{\bs{x}}_1 &= \frac{\bs{r}_p}{\|\bs{r}_p\|}\label{eq:x1}\\
 	\hat{\bs{x}}_2 &= \frac{\bs{L} \times \hat{\bs{x}}_1}{\|\bs{L} \times \hat{\bs{x}}_1\|} = \hat{\bs{x}}_3 \times \hat{\bs{x}}_1\\
-	\hat{\bs{x}}_3 &= \hat{\bs{L}} = \frac{\bs{r}_p \times \bs{v}_p}{\|\bs{r}_p \times \bs{v}_p\|}.
+	\hat{\bs{x}}_3 &= \hat{\bs{L}} = \frac{\bs{r}_p \times \bs{v}_p}{\|\bs{r}_p \times \bs{v}_p\|}\label{eq:x3}.
 \end{align}
@@
 \includegraphics[width=\textwidth]{../../plots/paper2/Lpts_r.png}
-\caption{ Orbits of 2000 particles around $t=\tub$.
-The orbits were integrated backwards from their present-day positions (final time step of the n-body simulations) without the potential of the progenitor.
-TODO: coordinates...}\label{fig:lpts_r}
+\caption{ Orbits of 2000 randomly drawn particles projected into the instantaneous orbital plane coordinates (Eqs.~\ref{eq:x1}-\ref{eq:x3}), normalized by the tidal radius (Eq.~\ref{eq:tidalradius}), and only shown within half a satellite crossing time around $t=\tub$ for each of the four progenitor masses.
+The orbits were integrated backwards from their present-day positions (final time step of the n-body simulations) as test particles without the potential of the progenitor.
+Horizontal black line shows $x_2=0$ (top panels) and $x_3=0$ (bottom panels), and the unit circle (black circle) shows the classical disruption radius in these coordinates.
+}\label{fig:lpts_r}
 \end{center}
@@
 \subsection{Probabilistic model}
-% TODO: define what we mean by "Model"
 We observe the 6D position of a star, $\D$, in heliocentric coordinates --- e.g., the measured position on the sky, ($l$, $b$); distance, $d$; proper motions, ($\mu_l$, $\mu_b$); and line-of-sight velocity, $v_r$ --- and have determined through some other means that this star was once part of a progenitor system (e.g., satellite galaxy) with mass $m(t)$ that is disrupting and forming a cold debris structure in the potential, $\Phi$, of the parent galaxy (e.g., the Milky Way).
@@
 Standard MCMC algorithms (e.g., Metropolis-Hastings) update a single chain while exploring parameter space.
-We instead use an affine-invariant ``ensemble'' sampler \citep{goodman10} that, each step in parameter space, updates the positions of many ``walkers''.
+We instead use an affine-invariant ``ensemble'' sampler \citep{goodman10} that, each step in parameter space, updates the positions of many ``walkers'' (the ensemble).
 This algorithm is implemented in the \project{Python} programming language \citep{foremanmackey13} and runs naturally in a parallelized environment, e.g., the message passing interface (MPI).
+In each experiment we run the walkers for a large number of steps from starting positions described below, then throw away these samples and take the positions from the final step of this ``burn-in'' phase as a starting point for the samples used for inference.
 We compute the autocorrelation time for each sampled parameter (using \texttt{ACOR}\footnote{www.math.nyu.edu/faculty/goodman/software/acor/}$^{,}$\footnote{https://github.com/dfm/acor}) and thin the chains by taking every $\mathrm{median}(t_{\rm acor})$ sample to ensure the samples are close to independent.
-
-% TODO: define walkers and burn-in
 

### 2014-03-22 scripts/figures/paper2_figures.py
pretty up mcmc trace plot

@@ -569,10 +569,9 @@ def trace_plots():
         p0 = f["p0"].value
         acor = f["acor"].value
 
-    bounds = (-7,7)
     labels = ["$q_1$", "$q_z$", r"$\phi$", "$v_h$", r"$\alpha$"]
-    ticks = (-5, 0, 5)
-    ticklabels = ["-5%", "truth", "5%"]
+    bounds = [(1.3,1.5),(1.25,1.45),(85,115),(118,132),(0.5,2.5)]
+    ticks = [(1.35,1.4,1.45),(1.3,1.35,1.4),(90,100,110),(120,125,130),(1.,1.5,2.)]
 
     # plot individual walkers
     fig,axes = plt.subplots(5,1,figsize=(8.5,11),sharex=True)
@@ -580,34 +579,45 @@ def trace_plots():
     k = 0
     for gname,group in model.parameters.items():
         for pname,p in group.items():
-
-            if gname == "potential":
-                thischain = (chain[...,k] - p.truth) / p.truth*100.
-            else:
-                thischain = chain[...,k]
+            thischain = _unit_transform[pname](chain[...,k])
 
             for ii in range(config['walkers']):
                 axes.flat[k].plot(thischain[ii,5000:],
                                   alpha=0.1, marker=None,
                                   drawstyle='steps', color='k')
-            axes.flat[k].set_ylabel(labels[k], rotation='horizontal')
+            #axes.flat[k].set_ylabel(labels[k], rotation='horizontal')
+            axes[k].text(-0.02, 0.5, labels[k],
+                         horizontalalignment='right',
+                         fontsize=22,
+                         transform=axes[k].transAxes)
 
-            if gname == "potential":
-                axes.flat[k].set_ylim(bounds)
-                axes.flat[k].axhline(0., color='#31a354', lw=2., linestyle='-', alpha=0.9)
-                axes.flat[k].set_yticks(ticks)
-                axes.flat[k].set_yticklabels(ticklabels)
-            else:
-                axes.flat[k].set_yticks([1.,1.5,2.])
-                axes.flat[k].set_ylim(0.75,2.25)
-            #axes.flat[k].yaxis.tick_right()
-            axes.flat[k].yaxis.set_label_position("right")
+            if pname == "phi":
+                axes[k].text(1.07, 0.475, "deg",
+                         horizontalalignment='left',
+                         fontsize=18,
+                         transform=axes[k].transAxes)
+
[…]

### 2014-03-22 text/paper2/ms.tex
comments from my readthrough

@@
 However there is large uncertainty in the radial profile, shape, orientation, and graininess of the outer halo and the constancy of these parameters over distance.
-Properly describing the potential may require complex models where computing actions is non-trivial or even impossible if it is not representable by perturbing a nearby set of canonical coordinates.
+Properly describing the potential may require complex models where computing actions is non-trivial or even impossible.
 	\item Near-future photometric surveys such as \gaia\, and the \project{LSST} will likely discover many new streams and kinematic associations of stars.
@@
 
-Consider a point-mass satellite of mass $m$ on a circular orbit with frequency $\Omega$ around a more massive point-mass, $M\gg m$ \citep[the restricted three-body problem; e.g., \S 8.3][]{binneytremaine}.
-In a frame rotating with the orbital frequency of the satellite, the effective potential around the satellite has (amongst others) two stationary, unstable points located around $\sim R \pm r_J$, where $R$ is the orbital radius of the satellite and $r_J$ is the Jacobi or tidal radius,
+Consider a point-mass satellite of mass $m$ on a circular orbit with frequency $\Omega$ around a more massive mass, $M\gg m$ \citep[the ``restricted three-body problem''; e.g., \S 8.3][]{binneytremaine}.
+In a frame rotating with the orbital frequency of the satellite, the satellite remains fixed and the static, effective potential around the satellite has (amongst others) two stationary, unstable points located around $\sim R \pm r_J$, where $R$ is the orbital radius of the satellite and $r_J$ is the Jacobi or tidal radius,
 \begin{equation}
@@
 For a spherical, extended parent mass distribution the tidal radius instead scales with the enclosed mass at the instantaneous orbital radius, $R(t)$, with additional terms that account for the local slope of the density profile.
-The satellite may also be losing mass, in which case the satellite mass also depends on time.
-In general, the effective Lagrange points may not be symmetric about the center of the satellite and may deviate from the point-mass tidal radius (Eq.~\ref{eq:ptmass}) by factors of order unity.
-We adopt
+The satellite may also be losing mass so that in general the satellite mass also depends on time.
+The distance scale of the debris then follows
 \begin{equation}
@@
 \end{equation}
-as a distance scale for the debris, where $M_{\rm enc}$ is the instantaneous enclosed mass of the parent system within orbital radius $R$.
-At pericenter, the velocity of the satellite will be of order $V\sim \sqrt{GM_{\rm enc}/R}$.
+where $M_{\rm enc}$ is the instantaneous enclosed mass of the parent system within orbital radius $R$.
+In general, the effective Lagrange points may not be symmetric about the center of the satellite and may deviate from the tidal radius by factors of order unity.
+
+At pericenter, where a large fraction of the stars are unbound, the velocity of the satellite will be of order $V\sim \sqrt{GM_{\rm enc}/R}$.
 If the velocity dispersion of the satellite is $\sigma_v \sim \sqrt{Gm/r_J}$, then
@@
 
-In a triaxial potential, the orbital plane of the satellite is not fixed but we still expect there to be \emph{effective} Lagrange points for a spherical satellite system along the line of centers connecting the origin of the parent potential to the satellite.
-We thus proceed by defining a coordinate system relative to the position and velocity of the progenitor, ($\bs{r}_p$,$\bs{v}_p$), rotated into the instantaneous orbital plane defined by $\bs{L} = \bs{r}_p \times \bs{v}_p$, and normalized by the instantaneous tidal radius for positions (Eq.~\ref{eq:tidalradius}) and instantaneous velocity scale for velocity (Eq.~\ref{eq:velscale}).
-% where $\hat{\bs r}_p = {\bs r}_p / |{\bs r}_p|$ and $\hat{\bs v}_p = {\bs v}_p / |{\bs v}_p|$
+In a triaxial potential, the orbital plane of the satellite is not fixed but we still expect there to be \emph{effective} Lagrange points for a spherical satellite system along the line of centers connecting the origin of the parent potential to the satellite --- that is, we expect the stars to be stripped at some characteristic radius with some dispersion about this point.
+We proceed by defining a coordinate system relative to the position and velocity of the progenitor, ($\bs{r}_p$,$\bs{v}_p$), rotated into the instantaneous orbital plane defined by $\bs{L} = \bs{r}_p \times \bs{v}_p$.
+The positions are normalized by the instantaneous tidal radius (Eq.~\ref{eq:tidalradius}) and velocities are normalized by the instantaneous velocity scale (Eq.~\ref{eq:velscale}).
 The (time-dependent) basis vectors are given by
@@
 	\hat{\bs{x}}_2 &= \frac{\bs{L} \times \hat{\bs{x}}_1}{\|\bs{L} \times \hat{\bs{x}}_1\|} = \hat{\bs{x}}_3 \times \hat{\bs{x}}_1\\
-	\hat{\bs{x}}_3 &= \hat{\bs{L}} = \frac{\bs{r}_p \times \bs{v}_p}{\|\bs{r}_p \times \bs{v}_p\|}\label{eq:x3}.
+	\hat{\bs{x}}_3 &= \hat{\bs{L}} = \frac{\bs{L}}{\|\bs{L}\|} = \frac{\bs{r}_p \times \bs{v}_p}{\|\bs{r}_p \times \bs{v}_p\|}\label{eq:x3}.
 \end{align}
@@
 Figure~\ref{fig:lpts_r} shows sections of particle orbits from the n-body simulations described above projected into this coordinate system and normalized, symmetric around their unbinding times, $\tub$, for each of the four progenitor masses.
-In these coordinates, the classical Lagrange points would be located at $x_1\approx\pm1,x_2\approx0,x_3\approx0$ (illustrated by the unit circle in top panels of Figure~\ref{fig:lpts_r}).
-Figure~\ref{fig:lpts_v} shows the velocity of each star also projected into these coordinates and normalized.
+In these coordinates, the classical Lagrange points would be located at $x_1\approx\pm1,x_2\approx0,x_3\approx0$ (illustrated by the unit circle in black in Figure~\ref{fig:lpts_r}).
+Figure~\ref{fig:lpts_v} shows the velocity orbits of each star also projected into these coordinates and normalized.
 Though not centered on the point-mass Lagrange points, the location of and dispersion about the effective Lagrange points in these scaled coordinates remain remarkably consistent across the range of progenitor masses explored.
@@
 We find that even in a triaxial potential, the phase-space dispersion of tidally stripped debris scales with the mass ratio $(m / M_{\rm enc})^{1/3}$.
-This motivates a model for tidal debris in each star was ``released'' at the instantaneous, effective Lagrange point at its unbinding time, $\tub$, with some dispersion in position and velocity, all of which depend only on the mass and orbit of the progenitor and the parent potential.
+This motivates a model for tidal debris in which each star was ``released'' at the instantaneous, effective Lagrange point at its unbinding time, $\tub$, with some dispersion in position and velocity, all of which depend only on the mass and orbit of the progenitor and the parent potential.
 We present this model in detail below.
@@
 \includegraphics[width=\textwidth]{../../plots/paper2/Lpts_v.png}
-\caption{ Same as Figure~\ref{fig:lpts_r} but for particle velocities. }\label{fig:lpts_v}
+\caption{ Same as Figure~\ref{fig:lpts_r} but for particle velocities normalized by the velocity scale (Eq.~\ref{eq:velscale}). }\label{fig:lpts_v}
 \end{center}
@@
 
-\subsection{Probabilistic model}
+\subsection{Probabilistic model (\rewinder)}
 We observe the 6D position of a star, $\D$, in heliocentric coordinates --- e.g., the measured position on the sky, ($l$, $b$); distance, $d$; proper motions, ($\mu_l$, $\mu_b$); and line-of-sight velocity, $v_r$ --- and have determined through some other means that this star was once part of a progenitor system (e.g., satellite galaxy) with mass $m(t)$ that is disrupting and forming a cold debris structure in the potential, $\Phi$, of the parent galaxy (e.g., the Milky Way).
@@
 To model the true, 6D, present-day position of the star, $\W$, we first transform to cartesian, Galactocentric coordinates where ($\bs{r},\bs{v}$) and ($\bs{r}_\sat,\bs{v}_\sat$) are the position and velocity of the star and progenitor.
-In cartesian coordinates, the present-day position of the star is taken as having been sampled at $t=\tub$ from an isotropic Gaussian centered on one of the instantaneous, effective Lagrange points in position, and an isotropic Gaussian centered on the origin in velocity, evolved to present day in the parent potential, $\Phi$, parametrized by the vector $\bs{\Theta}_\Phi$.
+In cartesian coordinates, the present-day position of the star is taken as having been sampled at $t=\tub$ from an isotropic Gaussian centered on one of the instantaneous, effective Lagrange points in position, and an isotropic Gaussian centered on the origin in velocity, evolved to present day in the parent potential, $\Phi$, parametrized by the parameter vector $\bs{\Theta}_\Phi$.
 We assume that once the star becomes ``unbound'' from the satellite the potential of the satellite can be ignored, and thus treat the star as a test particle.\footnote{This assumption breaks down for sufficiently large-mass progenitors, somewhere between $\sim10^8-10^9~$M$_\odot$.} The satellite mass enters the model through the tidal radius; a changing satellite mass will skew the positions of the effective Lagrange points and velocity scale.
@@
 \begin{align}
-	p(\bs{\Theta} \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi) &= p(\bs{r} \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi)\,p(\bs{v} \given\bs{\Theta}_\sat, \bs{\Theta}_\Phi)\,\left\vert\J\right\vert_{t=\tub}\\
+	p(\bs{\Theta} \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi) &= p(\bs{r} \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi)\,p(\bs{v} \given\bs{\Theta}_\sat, \bs{\Theta}_\Phi)\,\left\vert\J\right\vert_{t=\tub}\label{eq:starlike}\\
 	p(\bs{r} \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi) &= \mathcal{N}(\bs{r} \given \bs{r}_\sat - \Loffset\,\tailbit\,\rtide\,\hat{x}_1, \rtide^2)_{t=\tub}\\
@@
 \begin{enumerate}
-	\item sample from the observed distributions in heliocentric coordinates (Eq.~\ref{eq:obsstar} and Eq.~\ref{eq:obsprog}) to get present-day positions for both the star and progenitor;
-	\item transform to Galactocentric, cartesian coordinates, and integrate both orbits backwards (as test particles) for the total interaction time, $\tint$;
-	\item evaluate the likelihood at $t=\tub$.
+	\item sample from observed distributions over the heliocentric positions (Eqs.~\ref{eq:obsstar},\ref{eq:obsprog}), or priors if positions are unknown, for the star and progenitor;
+	\item transform star and progenitor positions in heliocentric coordinates to Galactocentric, cartesian coordinates;
+	\item integrate both orbits backwards (as test particles) for the total interaction time, $t_{\rm int}\approx6$~Gyr;
+	\item at the timestep closest to $\tub$, transform the particle position to the relative, normalized coordinates defined in Section~\ref{sec:debris};
+	\item evaluate the likelihood according to which tail the star belongs to, $\tailbit$, (Eq.~\ref{eq:starlike}).
 \end{enumerate}
-All model parameters are summarized in Table~\ref{tbl:params}.
 Assuming each star is an independent tracer, the full posterior for many stars is then just the product of the individual likelihoods with the priors:
@@
 	p(\{\bs{\Theta}^{(i)}\}, \bs{\Theta}_\sat, \bs{\Theta}_\Phi \given \{\D^{(i)}\}, \D_\sat) = \\
-		\shoveleft{\frac{1}{\mathcal{Z}} \, \prod_i \, p(\D^{(i)} \given \W^{(i)}) \, p(\D_\sat \given \W_\sat)} 
+		\frac{1}{\mathcal{Z}} \, \prod_i \, p(\D^{(i)} \given \W^{(i)}) \, p(\D_\sat \given \W_\sat)\,
 			p(\bs{\Theta}^{(i)} \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi) \, 
@@
 \end{multline}
+All parameters are summarized in Table~\ref{tbl:params}.
+
 This model works well when all stars are lost due to tidal pruning.
-Even on a mildly eccentric orbit, stars become preferentially unbound around the satellite's pericenter because of tidal shocking.
+Even on a mildly eccentric orbit, stars are preferentially stripped around the satellite's pericenter because of tidal shocking.
 We will explore properly modeling this process in future work.
@@
 \section{Experiments} \label{sec:experiments}
-In what follows, we apply the model described above to ``observations'' of the n-body simulations described in Section~\ref{sec:sims}.
+In what follows, we use \rewinder\, to model ``observations'' of the n-body simulations described in Section~\ref{sec:sims}.
 For all tests in this article, we model the potential of the parent galaxy with the same functional form used in the simulations (Equation~\ref{eq:lm10}) --- we stress that the data are generated with n-body simulations and are not generated with anything resembling the likelihood function presented above.
@@
 
-\subsection{Data with negligible uncertainties}
+\subsection{Data with negligible uncertainties}\label{sec:exp1}
 
-We first test the model using eight particles (stars) --- four from the leading tail, and four from the trailing tail --- randomly sampled from the $2.5\times10^8~M_\odot$ simulation (Section~\ref{sec:sims}), assuming perfect knowledge of the observed 6D positions for both the stars and the progenitor.
+We first test the model using eight particles (stars) --- four from the leading tail, and four from the trailing tail --- randomly sampled from the $2.5\times10^8~M_\odot$ simulation (Section~\ref{sec:sims}), assuming the observed 6D positions for both the stars and the progenitor have negligible uncertainties.
 The stars were required to have been stripped after the first pericentric passage and have a present-day distance within $50$ kpc of the Sun.
@@
 We leave the potential parameters ($q_1,q_z,\phi,\vhalo$) free to vary, along with the Lagrange point offset, $\Loffset$, and initialize an ensemble of 64 MCMC walkers by sampling from the priors summarized in Table~\ref{tbl:params}.
-We run the walkers for 10000 steps to burn-in, then restart the sampler starting from the final position of the burn-in phase and run for another 5000 steps.
+We run the walkers for 10000 steps to burn-in, then restart the sampler starting from the final position of the burn-in phase and run for another 5000 inference steps.
 Figure~\ref{fig:trace} shows the walker positions over the 5000 inference steps for each of the five parameters.
-The autocorrelation time for each parameter is displayed on its corresponding panel and the horizontal lines indicate the true simulation values for the potential parameters.
+The autocorrelation time for each parameter is displayed on its corresponding panel.
 Figure~\ref{fig:exp1_posterior} shows projections of the posterior probability distribution for the parameters; with perfect data, the uncertainties on the potential parameters are all $<1\%$.
@@
 
-\subsection{Data with near-future uncertainties}
+\subsection{Data with near-future uncertainties}\label{sec:exp2}
 We next take the same eight stars used in the previous experiment and ``observe'' them with optimistic observational uncertainties.
@@
 
-This experiment samples over the four potential parameters, the Lagrange point offset, four phase-space coordinate parameters for each star (32 total), and four phase-space coordinate parameters for the progenitor --- 41 parameters in total.
+This experiment samples over the four potential parameters, the Lagrange point offset, four phase-space coordinate parameters plus the unbinding time for each star (40 total), and four phase-space coordinate parameters for the progenitor --- 49 parameters in total.
 We use an ensemble of 256 walkers and draw initial conditions for the coordinate parameters by sampling from Gaussian's centered on the ``observed'' value with variances specified by the observational uncertainties.
@@
 \includegraphics[width=\textwidth]{../../plots/paper2/exp2_potential.png}
-\caption{ TODO  }\label{fig:exp2_potential}
+\caption{ Projections of the marginal posterior over the triaxial potential parameters for observed stars and progenitor with near-future uncertainties (Section~\ref{sec:exp2}).
+Axis ranges show the lower and upper bounds on the uniform priors over these parameters.
+}\label{fig:exp2_potential}
 \end{center}
@@
 \includegraphics[width=\textwidth]{../../plots/paper2/exp2_satellite.png}
-\caption{ TODO  }\label{fig:exp2_satellite}
+\caption{ Projections of the marginal posterior over the progenitor parameters for observed stars and progenitor with near-future uncertainties (Section~\ref{sec:exp2}). }\label{fig:exp2_satellite}
 \end{center}
@@
 \includegraphics[width=\textwidth]{../../plots/paper2/exp2_particle.png}
-\caption{ TODO  }\label{fig:exp2_particle}
+\caption{ Projections of the marginal posterior over parameters for one of the stars for observed stars and progenitor with near-future uncertainties (Section~\ref{sec:exp2}).
+}\label{fig:exp2_particle}
 \end{center}
@@
 
-\subsection{Precise distance measurements with missing proper motions}
+\subsection{Precise distance measurements with missing proper motions}\label{sec:exp3}
 The SMASH survey \citep{smashprop} will be completed in 2014, long before the final \gaia\, data release.
@@
 SMASH also targets several RR Lyrae in the Sgr core, which enable a high-precision distance measurement of the progenitor.
-Though proper motions have been measured for the Sgr progenitor, we now consider a case in which we have high-precision distances to stream stars and the progenitor, 15~km/s radial velocity uncertainty, and missing proper motions for all stars and progenitor.
+Though proper motions have been measured for the Sgr progenitor, we now consider a case in which we have high-precision (2\%) distances to stream stars and the progenitor, 10~km/s radial velocity uncertainty, and missing proper motions for all stars and progenitor.
 
-As with the previous experiment, this experiment includes 41 model parameters in total.
+As with the previous experiment, this experiment includes 49 model parameters in total.
 We again use an ensemble of 256 walkers, but this time initialize the walkers by sampling the coordinate parameters from Gaussians centered on the true positions with variances set to 1\% of the parameter values.
@@
 Figures~\ref{fig:exp3_potential}, \ref{fig:exp3_satellite}, \ref{fig:exp3_particle} show the marginalized posteriors for the potential parameters, satellite parameters, and parameters for a single particle.
+Accounting for observational uncertainties does increase the uncertainties in the constraints on the potential parameters (as expected), but this demonstrates that a small sample of well-measured stream stars provide an enormous amount of information about the triaxiality of the potential.
 
@@
 \subsection{Precise distance measurements of stars with no observed progenitor}
-Most of the streams and debris structures discovered thus far have undetected or unconfirmed progenitors. [...]
-4 stars in Orphan-like stream, XX stars in GD-1 -like stream
+Most of the streams and debris structures discovered thus far have undetected or unconfirmed progenitors.
+We expect the phase-space positions of the stars to provide constraints on the position of the progenitor even if it is unobserved.
+For this experiment, we consider a situation where we have precise (2\%) distance measurements, 10~km/s radial velocity uncertainties, and no proper motion measurements to eight stream stars (the same used in previous experiments), and no knowledge of the 6D position of the progenitor object.
 
-%\subsection{Mass comparison}
-%[...]
+Figures~\ref{fig:exp4_potential}, \ref{fig:exp4_satellite}, \ref{fig:exp4_particle} show the marginalized posteriors for the potential parameters, satellite parameters, and parameters for a single particle.
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[width=\textwidth]{../../plots/paper2/exp4_potential.png}
+\caption{Same as Figure~\ref{fig:exp2_potential} but with no a priori knowledge of the 6D position of the progenitor.}\label{fig:exp4_potential}
+\end{center}
+\end{figure}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[width=\textwidth]{../../plots/paper2/exp4_satellite.png}
+\caption{ Same as Figure~\ref{fig:exp2_satellite} but with no a priori knowledge of the 6D position of the progenitor.}\label{fig:exp4_satellite}
+\end{center}
+\end{figure}
+
+\begin{figure}[!ht]
+\begin{center}
+\includegraphics[width=\textwidth]{../../plots/paper2/exp4_particle.png}
+\caption{ Same as Figure~\ref{fig:exp2_particle} but with no a priori knowledge of the 6D position of the progenitor.}\label{fig:exp4_particle}
+\end{center}
+\end{figure}
 
@@
 
+[Computational...]
+
+[Respond to each point in the list from intro:]
+\begin{enumerate}
+	\item The known debris structures are $\gtrsim$10 kpc from the Sun, thus while overdensities in the halo are less contaminated by background, the constituent stars often have incomplete or poorly measured phase-space information.
+Thus it is critical for any approximate model of tidal debris to incorporate observational uncertainties and missing dimensions in a consistent and justified way.
+	\item The potential of the inner galaxy is conceivably well-described by a (set of) nearly-integrable potentials, thus the dynamics of collisionless systems is plausibly captured by a transformation to some integral-of-motion-space, e.g., action-angle coordinates.
+However there is large uncertainty in the radial profile, shape, orientation, and graininess of the outer halo and the constancy of these parameters over distance.
+Properly describing the potential may require complex models where computing actions is non-trivial or even impossible if it is not representable by perturbing a nearby set of canonical coordinates.
+	\item Near-future photometric surveys such as \gaia\, and the \project{LSST} will likely discover many new streams and kinematic associations of stars.
+Potential recovery methods should be able to simultaneously model multiple streams and incorporate other dynamical constraints.
+	\item Matching generated models to observed stream densities is difficult: models that rely on this must account for observational biases and properly model the underlying background of halo stars. 
+	\item Full N-body models are expensive to run.
+Incorporating an n-body model into a general model and then performing a parameter search would be computationally intensive and, presently, intractable.
+\end{enumerate}
+
 \section{Conclusions}
@@
 \rewinder\, relies on direct orbit integration and not on computing conserved quantities (e.g., actions) and can thus be used with arbitrarily complex (e.g., time-dependent or clumpy) forms of the potential.
-We have performed several experiments ...
+We have performed several experiments to show that \rewinder\, simultaneously constrains the potential and models the stream given simulated data with a range of realistic uncertainties.
 

### 2014-03-22 streams/dynamics/particles.py
properly return errors based on error_X

@@ -165,7 +165,6 @@ class ObservedParticle(Particle):
         super(ObservedParticle, self).__init__(coords, frame, units=units, meta=meta)
 
         # parse errors same as coords
-        self.errors = dict()
         self._error_X = np.zeros_like(self._X)
         for ii in range(self.ndim):
             name = self.frame.coord_names[ii]
@@ -185,9 +184,17 @@ class ObservedParticle(Particle):
 
             self._error_X[...,ii] = value
 
+    @property
+    def errors(self):
+        errors = dict()
+
+        for ii in range(self.ndim):
+            name = self.frame.coord_names[ii]
             # set error dictionary
-            self.errors[name] = (self._error_X[...,ii]*self._internal_units[ii])\
-                                    .to(self._repr_units[ii])
+            errors[name] = (self._error_X[...,ii]*self._internal_units[ii])\
+                            .to(self._repr_units[ii])
+
+        return errors
 
     @property
     def _repr_error_X(self):

### 2014-03-22 streams/inference/back_integrate.py
compute data likelihood properly

@@ -34,7 +34,7 @@ def xyz_sph_jac(hel):
     return deet
 
 def back_integration_likelihood(t1, t2, dt, potential, p_gc, s_gc, logm0, logmdot,
-                                tub, beta, p_shocked, alpha):
+                                tub, beta, alpha):
 
     """ Compute the likelihood of 6D heliocentric star positions today given the
         potential and position of the satellite.
@@ -59,15 +59,12 @@ def back_integration_likelihood(t1, t2, dt, potential, p_gc, s_gc, logm0, logmdo
             Array of unbinding times for each star.
         beta : array_like
             Array of tail bits (K) to specify leading/trailing tail for each star.
-        p_shocked : array_like
-            Probability the star was shocked.
         alpha : float
             Position of effective tidal radius.
 
     """
 
-    K = np.sign(beta)
-    p_shocked = np.array(p_shocked)
+    beta = np.sign(beta)
 
     gc = np.vstack((s_gc,p_gc)).copy()
     acc = np.zeros_like(gc[:,:3])
@@ -145,7 +142,7 @@ def back_integration_likelihood(t1, t2, dt, potential, p_gc, s_gc, logm0, logmdo
     #                (np.log(v_disp) + VZ**2))
 
     sigma_r = 0.5
-    r_term = -0.5*((np.log(sigma_r) + (X-alpha*K)**2/sigma_r**2) + \
+    r_term = -0.5*((np.log(sigma_r) + (X-alpha*beta)**2/sigma_r**2) + \
                    (np.log(sigma_r) + Y**2/sigma_r**2) + \
                    (np.log(sigma_r) + Z**2/sigma_r**2))
     # r_term_trai = -0.5*((np.log(sigma_r) + (X-2*alpha)**2/sigma_r**2) + \

### 2014-03-22 streams/inference/core.py

@@ -98,6 +98,7 @@ class StreamModel(object):
         t1 = config.get("t1", float(d["t1"]))
         t2 = config.get("t2", float(d["t2"]))
         dt = config.get("dt", config.get('dt', -1.))
+        logger.debug("Integration from {} to {}, dt={} Myr".format(t1,t2,dt))
 
         # get the potential object specified from the potential subpackage
         from .. import potential as sp
@@ -105,7 +106,7 @@ class StreamModel(object):
         potential = Potential()
         logger.info("Using potential '{}'...".format(config["potential"]["class_name"]))
 
-        # Define the empty model to add parameters to
+        # Initialize the empty model to add parameters to
         model = cls(potential, lnpargs=[t1,t2,dt],
                     particles=particles,
                     satellite=satellite,
@@ -116,7 +117,7 @@ class StreamModel(object):
         if config["potential"]["parameters"]:
             for ii,name in enumerate(config["potential"]["parameters"]):
                 p = potential.parameters[name].copy()
-                logger.debug("Prior on {}: Uniform({}, {})".format(name, p._prior.a, p._prior.b))
+                logger.debug("Prior on {}: {}".format(name, p._prior))
                 model.add_parameter('potential', p)
 
         # Particle parameters
@@ -134,9 +135,13 @@ class StreamModel(object):
                     logger.debug("\t\t{}".format(name))
                 else:
                     p = getattr(particles,name)
-                    logger.debug("Prior on {}: Uniform({}, {})".format(name, p._prior.a,
-                                                                             p._prior.b))
+                    logger.debug("Prior on {}: {}".format(name, p._prior))
                     model.add_parameter('particles', p)
+
+            missing_dims = config["particles"].get("missing_dims", list())
+            for missing_dim in missing_dims:
+                ix = heliocentric.coord_names.index(missing_dim)
+                model.particles._error_X[:,ix] = np.inf
 
         # Satellite parameters
         if config['satellite']['parameters']:
@@ -158,9 +163,13 @@ class StreamModel(object):
 
                 else:
                     p = getattr(satellite,name)
-                    logger.debug("Prior on {}: Uniform({}, {})".format(name, p._prior.a,
-                                                                             p._prior.b))
[…]

### 2014-03-22 streams/inference/prior.py

@@ -28,6 +28,12 @@ class LogPrior(object):
 
 class LogUniformPrior(LogPrior):
 
+    def __str__(self):
+        return "Uniform({}, {})".format(self.a, self.b)
+
+    def __repr__(self):
+        return self.__str__()
+
     def __call__(self, value):
         if np.any((value < self.a) | (value > self.b)):
             return -np.inf

### 2014-03-22 text/paper2/ms.tex
discussion

@@
 \section{Discussion}
+\rewinder\, addresses each of the issues identified with other stream modeling methods: 
+\begin{enumerate}
+	\item The true 6D positions of stars and the progenitor are included as parameters in the model and thus observational uncertainties and missing data dimensions are naturally incorporated in the inferences.
+	\item The orbits of the stars are directly integrated and only require a function that evaluates the acceleration due to the potential at any given time and position.
+In this article we only consider one parametrized model for the potential but [...]
+	\item Incorporating multiple streams only requires multiplying likelihoods.
+	\item The inference is much faster than full n-body modeling because the orbits of the stars and progenitor are integrated as test particles.
+\end{enumerate}
 
-Our method also naturally extends to using multiple debris structures \citep[see][who illustrate the power of using multiple streams to simultaneously constrain the potential]{deg14}.
+% Our method also naturally extends to using multiple debris structures \citep[see][who illustrate the power of using multiple streams to simultaneously constrain the potential]{deg14}.
 
@@
 
+[Haven't explored parameter space...]
+For experiments in this article with realistic uncertainties we have not performed a search in parameter space for the optimal values, and instead started the MCMC walkers with a small dispersion around the true parameter values.
+
 [Computational...]
-
-[Respond to each point in the list from intro:]
-\begin{enumerate}
-	\item The known debris structures are $\gtrsim$10 kpc from the Sun, thus while overdensities in the halo are less contaminated by background, the constituent stars often have incomplete or poorly measured phase-space information.
-Thus it is critical for any approximate model of tidal debris to incorporate observational uncertainties and missing dimensions in a consistent and justified way.
-	\item The potential of the inner galaxy is conceivably well-described by a (set of) nearly-integrable potentials, thus the dynamics of collisionless systems is plausibly captured by a transformation to some integral-of-motion-space, e.g., action-angle coordinates.
-However there is large uncertainty in the radial profile, shape, orientation, and graininess of the outer halo and the constancy of these parameters over distance.
-Properly describing the potential may require complex models where computing actions is non-trivial or even impossible if it is not representable by perturbing a nearby set of canonical coordinates.
-	\item Near-future photometric surveys such as \gaia\, and the \project{LSST} will likely discover many new streams and kinematic associations of stars.
-Potential recovery methods should be able to simultaneously model multiple streams and incorporate other dynamical constraints.
-	\item Matching generated models to observed stream densities is difficult: models that rely on this must account for observational biases and properly model the underlying background of halo stars. 
-	\item Full N-body models are expensive to run.
-Incorporating an n-body model into a general model and then performing a parameter search would be computationally intensive and, presently, intractable.
-\end{enumerate}
+Each step in parameter space with \rewinder\, requires integrating the orbits of stars and the progenitor: we presently integrate with a fixed timestep using leapfrog integration.
+Computing the acceleration due to the given potential at each step is implemented in \project{Cython} (and approaches \project{ANSI-C}-like speed), but the rest of the code is written in pure-\project{Python}.
+Though already significantly less computationally intensive than running a full n-body simulation for every parameter step, running each MCMC walker for 10000 steps requires parallelization and a few hours of CPU time on a compute cluster.
+Further optimization of the integration method (e.g., using an adaptive method or implementing in \project{Cython}) could speed up the inference significantly.
 

### 2014-03-23 scripts/figures/paper2_figures.py
duh

@@ -658,7 +658,7 @@ def exp1_posterior():
                           truths=truths, extents=bounds, labels=labels)
     fig.savefig(os.path.join(plot_path, "exp1_posterior.png"))
 
-def exp_posteriors(exp_num):
+def exp_posteriors(exp_num, slicey=-5000):
     cfg_filename = os.path.join(streamspath, "config", "exp{}.yml".format(exp_num))
     config = read_config(cfg_filename)
     model = StreamModel.from_config(config)
@@ -672,7 +672,13 @@ def exp_posteriors(exp_num):
         acor = f["acor"].value
 
     print("median acor: {}".format(int(np.median(acor))))
-    _flatchain = np.vstack(chain[acceptance_fraction>0.03,5000::int(np.median(acor))])
+
+    if slicey is None:
+        slicey = slice(-5000,None,int(np.median(acor)))
+    else:
+        slicey = slice(slicey,None,int(np.median(acor)))
+
+    _flatchain = np.vstack(chain[acceptance_fraction>0.03,slicey])
     d = model.label_flatchain(_flatchain)
     p0 = model.label_flatchain(_p0)
 
@@ -732,10 +738,9 @@ def exp_posteriors(exp_num):
     truths = []
     bounds = []
     labels = []
-    # HACK BECAUSE I SUCK
-    keys = ["d","mul","mub","vr","alpha"]
-    #for ii,pname in enumerate(d["satellite"].keys()):
-    for ii,pname in enumerate(keys):
+
+    #for ii,pname in enumerate(keys):
+    for ii,pname in enumerate(d["satellite"].keys()):
         this_flatchain[:,ii] = _unit_transform[pname](d["satellite"][pname][:,0])
         this_p0[:,ii] = _unit_transform[pname](p0["satellite"][pname][:,0])
 
@@ -755,6 +760,7 @@ def exp_posteriors(exp_num):
 
     # HACK
     bounds = [(29,32), (-2,-1.75), (1.4,1.7), (130,180), bounds[-1]]
+    #bounds = None
     fig = triangle.corner(this_flatchain, plot_datapoints=False,
                           truths=truths, labels=labels, extents=bounds)
     fig.savefig(os.path.join(plot_path, "exp{}_satellite.{}".format(exp_num, ext)))
@@ -763,7 +769,7 @@ def exp2_posteriors():
     exp_posteriors(2)
 
 def exp3_posteriors():
-    exp_posteriors(3)
[…]

### 2014-03-23 text/paper2/ms.tex

@@
 
-\section{Introduction}
+\section{Introduction}\label{sec:intro}
 
@@
 
+% TODO: m and mdot!!
 \begin{table}[h]
@@
 
-\subsection{Precise distance measurements of stars with no observed progenitor}
+\subsection{Near-future uncertainties but no observed progenitor}
 Most of the streams and debris structures discovered thus far have undetected or unconfirmed progenitors.
@@
 \section{Discussion}
-\rewinder\, addresses each of the issues identified with other stream modeling methods: 
-\begin{enumerate}
-	\item The true 6D positions of stars and the progenitor are included as parameters in the model and thus observational uncertainties and missing data dimensions are naturally incorporated in the inferences.
-	\item The orbits of the stars are directly integrated and only require a function that evaluates the acceleration due to the potential at any given time and position.
-In this article we only consider one parametrized model for the potential but [...]
-	\item Incorporating multiple streams only requires multiplying likelihoods.
-	\item The inference is much faster than full n-body modeling because the orbits of the stars and progenitor are integrated as test particles.
-\end{enumerate}
+In Section~\ref{sec:intro} we identified several issues that have only been partially solved by other stream modeling methods; \rewinder\, addresses each of these.
+%\begin{enumerate}
+%	\item Observational uncertainties and missing data dimensions are naturally handled in the inferences by including the true 6D positions of stars and the progenitor as parameters in the model.
+%	\item The orbits of the stars and progenitor are directly integrated and only require a function that evaluates the acceleration due to the potential at a given position (and time). 
+%	\item Incorporating multiple streams into the inference only requires multiplying likelihoods computed for each stream and progenitor.
+%	\item Each star adds constraints on the potential parameters and does not require matching a simulated density to the observed density of stars.
+%	\item The inference is much faster than full n-body modeling because the orbits of the stars and progenitor are integrated as test particles.
+%\end{enumerate}
 
-% Our method also naturally extends to using multiple debris structures \citep[see][who illustrate the power of using multiple streams to simultaneously constrain the potential]{deg14}.
+Observational uncertainties and missing data dimensions are naturally handled in the inferences by including the true 6D positions of stars and the progenitor as parameters in the model.
+This is a very important aspect [...because errors will be a function of distance, to properly model debris requires accounting for the uncertainties].
+In this article we have tested \rewinder\, using data of varied quality for a small sample of stars.
+In future work, we will investigate the effect of reducing the data quality but increasing the sample size.
+%[the power of this method is in small samples with high-quality data]
+
+No assumptions were made about the the form of the Galactic potential used in \rewinder: the potential could be time-dependent, clumpy, or have properties that vary with radius.
+The orbits of the stars and progenitor are directly integrated and only require a function that evaluates the acceleration due to the potential at a given position (and time).
+The experiments in this article infer potential parameters from the same functional potential used in the n-body simulations that generated the fake data.
+In future work, we will (1) try fitting incorrect potential models to see if we can still recover global properties, e.g., flattening; (2) run simulations with smoothly changing potentials \citep[e.g.,][]{buist14} and model the time dependence; and (3) explore using a generic, non-parametric potential form (e.g., a basis function expansion) as the recovery potential model.
+
+Each streams will provide constraints on different aspects of the potential.
+For example, more eccentric streams may better constrain the radial profile \citep[see][who illustrate the power of using multiple streams to simultaneously constrain the potential using orbit fitting]{deg14}.
+In this article, we only consider a stream on a Sgr-like orbit, which lies nearly in the Galactic $x$-$z$ plane.
+We find that the stream puts better constraints on the $z$-axis flattening, $q_z$, than on the flattening in the $x$-$y$ plane, $q_1$, as expected.
+The best measurements of the detailed shape of the Galactic potential will come from combining constraints from multiple streams.
+We have defined above a proper likelihood function for observed stars in a given stream, thus incorporating multiple streams into the inference only requires multiplying the likelihoods computed for each stream and progenitor pair.
+
+Each star adds constraints on the potential parameters and does not require matching a simulated density to the observed density of stars.
+
+[Issues about computation and scaling.]
+The inference is much faster than full n-body modeling because the orbits of the stars and progenitor are integrated as test particles.
+Each step in parameter space with \rewinder\, requires integrating the orbits of stars and the progenitor: we presently integrate with a fixed timestep using leapfrog integration.
+Computing the acceleration due to the given potential at each step is implemented in \project{Cython} (and approaches \project{ANSI-C}-like speed), but the rest of the code is written in pure-\project{Python}.
+Though already significantly less computationally intensive than running a full n-body simulation for every parameter step, running each MCMC walker for 10000 steps requires parallelization and a few hours of CPU time on a compute cluster.
+Further optimization of the integration method (e.g., using an adaptive method or implementing in \project{Cython}) could speed up the inference significantly.
+
+% TODO above: now address each point in more detail, shortcomings, what we'll do in future work
 
@@
 
-[We did not assume the potential is static in time.  At least not in
-  general; nothing in the method requires or suggests this.]
-
-[Issues about computation and scaling.]
+[Haven't explored parameter space...]
+For experiments in this article with realistic uncertainties we have not performed a search in parameter space for the optimal values, and instead started the MCMC walkers with a small dispersion around the true parameter values.
 
@@
 
-[Different streams will constraint different properties of the potential -- multiple streams are crucial.
-Here we use a stream roughly in the X-Z plane, thus we expect it to be more sensitive to flattening in Z direction]
-
-[Don't eschew low-dimensional constraints on the potential!]
-
-[Another prediction of hierarchical structure formation in $\Lambda$CDM is the presence of substructure. In principle, this method would work on any potential that we can integrate orbits in.]
-
-[Haven't explored parameter space...]
-For experiments in this article with realistic uncertainties we have not performed a search in parameter space for the optimal values, and instead started the MCMC walkers with a small dispersion around the true parameter values.
-
-[Computational...]
-Each step in parameter space with \rewinder\, requires integrating the orbits of stars and the progenitor: we presently integrate with a fixed timestep using leapfrog integration.
-Computing the acceleration due to the given potential at each step is implemented in \project{Cython} (and approaches \project{ANSI-C}-like speed), but the rest of the code is written in pure-\project{Python}.
-Though already significantly less computationally intensive than running a full n-body simulation for every parameter step, running each MCMC walker for 10000 steps requires parallelization and a few hours of CPU time on a compute cluster.
-Further optimization of the integration method (e.g., using an adaptive method or implementing in \project{Cython}) could speed up the inference significantly.
-
 \section{Conclusions}
@@
 \rewinder\, relies on direct orbit integration and not on computing conserved quantities (e.g., actions) and can thus be used with arbitrarily complex (e.g., time-dependent or clumpy) forms of the potential.
-We have performed several experiments to show that \rewinder\, simultaneously constrains the potential and models the stream given simulated data with a range of realistic uncertainties.
+We have performed several experiments to show that \rewinder\, simultaneously constrains the potential and models a tidal stream given simulated data with a range of realistic uncertainties.
 

### 2014-03-26 text/paper2/ms.tex
changed exp 4

@@
 % Results:
-We find that even with a small number (8 to 16) of well-measured stream stars,
-  we can precisely infer properties of the triaxial potential,
-  missing phase-space coordinates of the stars,
-  and the location of the progenitor object.
+We find that even with a small number (8) of well-measured stream stars,
+  we can precisely infer properties of the triaxial potential and
+  missing phase-space coordinates of the stars.
 The method generalizes to encompass mixtures of independent debris structures,
@@
 
+% TODO: audit all uses of the word 'model'
+
 Early large-scale, cosmological simulations of galaxy formation in the $\Lambda$CDM paradigm suggested that the spherically-averaged density profiles of dark matter halos follow a universal profile across a large dynamic range in mass \citep{navarro96}.
@@
 Simple, parameterized models of both the mass and orbit distribution have been simultaneously constrained using such data \citep[e.g.][]{napolitano11,deason12c}.
-%\citep{romanowsky03,douglas07,romanowsky09,napolitano11,lee11}
 Alternatively, the statistical properties of gravitationally lensed background sources around a galaxy can be used to constrain the \emph{projected} shape, orientation, and radial profile of mass \citep[see, for example, the Lens Structure and Dynamics Survey described in][]{koopmans02}.
@@
 	\item Matching generated models to observed stream densities is difficult: models that rely on this must account for observational biases and properly model the underlying background of halo stars. 
-	\item Full N-body models are expensive to run.
+	\item Full N-body simulations are expensive to run.
 Incorporating an n-body model into a general model and then performing a parameter search would be computationally intensive and, presently, intractable.
@@
 
-%[][hereafter \paperone]
 In \citet{apw13}, we introduced a simple method for using individual stars associated with tidal debris combined with knowledge about the mass and orbit of the progenitor to constrain properties of the host galaxy potential.
@@
 In this article, we use these simple scalings --- which depend only on the mass and orbit of the progenitor and parent potential --- to present a fully probabilistic method (\rewinder) for modeling tidal streams in any underlying potential given data with realistic uncertainties.
-By ``model'' here we mean a justified, parametrized likelihood function with priors on the parameters.
+By ``model'' we mean a justified, parametrized likelihood function with priors on the parameters.
 \rewinder\, relies on numerical orbit integration in ordinary phase-space and can thus incorporate arbitrarily complex models for the parent potential.
@@
 
-\subsection{Tidal debris}\label{sec:debris}
+\subsection{Motivation: tidal debris}\label{sec:debris}
 
-For a satellite galaxy on a mildly eccentric orbit, mass-loss is driven by a combination of tidal stripping due to the steady tidal field of the parent system and tidal shocking due to the rapidly changing tidal field as the progenitor system moves through pericenter \citep[e.g.,][]{choi09}.
-The interplay between these two processes as a function of orbital properties will be studied in forthcoming work; we focus below on the slow disruption of stars due to steady tidal forcing.
+For a satellite galaxy on a mildly eccentric orbit, mass loss is driven by a combination of tidal stripping by the steady tidal field of the parent system and tidal shocking by the rapidly changing tidal field as the progenitor system moves through pericenter \citep[e.g.,][]{choi09}.
+The interplay between these two processes as a function of orbital properties is outside our scope; we focus here on the slow removal of stars by steady tidal forcing.
 
-Consider a point-mass satellite of mass $m$ on a circular orbit with frequency $\Omega$ around a more massive mass, $M\gg m$ \citep[the ``restricted three-body problem''; e.g., \S 8.3][]{binneytremaine}.
-In a frame rotating with the orbital frequency of the satellite, the satellite remains fixed and the static, effective potential around the satellite has (amongst others) two stationary, unstable points located around $\sim R \pm r_J$, where $R$ is the orbital radius of the satellite and $r_J$ is the Jacobi or tidal radius,
+Consider a point-mass satellite of mass $m$ on a circular orbit with frequency $\Omega$ around a more massive mass $M\gg m$ \citep[the ``restricted three-body problem''; e.g., \S 8.3][]{binneytremaine}.
+In a frame rotating with the orbital frequency of the satellite, the satellite remains fixed and the static, effective potential around the satellite has (amongst others) two stationary, unstable points located around galactocentric radii $\sim R \pm r_J$, where $R$ is the orbital radius of the satellite and $r_J$ is the Jacobi or tidal radius
 \begin{equation}
@@
 \end{equation}
-Particles that would be bound to the satellite mass in isolation may have enough (Jacobi) energy to overcome the effective potential barrier at these Lagrange points (Fig.~8.6 of BT) and thus will be preferentially stripped from the satellite at these points.
+Particles that would be bound to the satellite in isolation may have enough (Jacobi) energy to overcome the effective potential barrier at these Lagrange points (Fig.~8.6 of BT) and thus will be preferentially stripped from the satellite at these points.
 For a spherical, extended parent mass distribution the tidal radius instead scales with the enclosed mass at the instantaneous orbital radius, $R(t)$, with additional terms that account for the local slope of the density profile.
-The satellite may also be losing mass so that in general the satellite mass also depends on time.
+In general the satellite mass also depends on time.
 The distance scale of the debris then follows
@@
 \end{equation}
-We expect the debris star velocities to also scale with $(m/M_{\rm enc})^{1/3}$.
-These scalings assume that the satellite is spherical with isotropic velocities; correlations amongst intra-satellite orbits from, e.g., a disk will break these assumptions.
+We expect the debris-star velocities also to scale with $(m/M_{\rm enc})^{1/3}$.
+These scalings assume that the satellite is spherical with isotropic velocities; correlations among intra-satellite orbits from, e.g., a disk will break these assumptions.
 
@@
 \end{align}
-%(Section~\ref{sec:sims})
-Figure~\ref{fig:lpts_r} shows sections of particle orbits from the n-body simulations described above projected into this coordinate system and normalized, symmetric around their unbinding times, $\tub$, for each of the four progenitor masses.
-In these coordinates, the classical Lagrange points would be located at $x_1\approx\pm1,x_2\approx0,x_3\approx0$ (illustrated by the unit circle in black in Figure~\ref{fig:lpts_r}).
+Figure~\ref{fig:lpts_r} shows sections of particle orbits (for particles that are stripped) from the n-body simulations described above projected into this coordinate system and normalized, symmetric around their unbinding times, $\tub$, for each of the four progenitor masses.
+In these coordinates, the classical Lagrange points would be located at $x_1\approx\pm\rtide,x_2\approx0,x_3\approx0$ (illustrated by the intersection of the black unit circles and horizontal lines in Figure~\ref{fig:lpts_r}).
 Figure~\ref{fig:lpts_v} shows the velocity orbits of each star also projected into these coordinates and normalized.
 Though not centered on the point-mass Lagrange points, the location of and dispersion about the effective Lagrange points in these scaled coordinates remain remarkably consistent across the range of progenitor masses explored.
-Furthermore, the dispersion in velocity is well-normalized by the velocity scale of Eq.~\ref{eq:velscale}, though it is clear that the velocity dispersion along $\hat{x}_1$, the radial vector to the satellite position, is larger than that in other dimensions, possibly due to tidal shocking.
+Furthermore, the dispersion in velocity is well-normalized by the velocity scale of Eq.~\ref{eq:velscale}, though it is clear that the velocity dispersion along $\hat{x}_1$, the radial vector to the satellite position, is larger than that in other dimensions, possibly because of tidal shocking.
 
-We find that even in a triaxial potential, the phase-space dispersion of tidally stripped debris scales with the mass ratio $(m / M_{\rm enc})^{1/3}$.
-This motivates a model for tidal debris in which each star was ``released'' at the instantaneous, effective Lagrange point at its unbinding time, $\tub$, with some dispersion in position and velocity, all of which depend only on the mass and orbit of the progenitor and the parent potential.
+We find that even in a triaxial potential, the dispersion --- in position and velocity --- of tidally stripped debris as it comes unbound from the satellite scales with the mass ratio $(m / M_{\rm enc})^{1/3}$.
+This motivates a model for tidal debris in which each star was ``released'' at the instantaneous, effective Lagrange point at its unbinding time, $\tub$, with a dispersion in position and velocity, all of which depend only on the mass and orbit of the progenitor and the parent potential.
 We present this model in detail below.
@@
 In general, the data for the star and progenitor will have missing dimensions and significant uncertainties at distances typical to the Galactic halo, thus we define $\W$ and $\W_\sat$ as the true 6D, heliocentric positions of the star and progenitor and will include these in our model.
-By including these positions in the model, this model will work even when the star has missing dimensions or the progenitor location is unknown as in, e.g., the Orphan stream \citep{belokurov07}.
+Since we include these positions as parameters, this model will work even when the star has missing dimensions or the progenitor location is unknown as in, e.g., the Orphan stream \citep{belokurov07}.
 
@@
 In cartesian coordinates, the present-day position of the star is taken as having been sampled at $t=\tub$ from an isotropic Gaussian centered on one of the instantaneous, effective Lagrange points in position, and an isotropic Gaussian centered on the origin in velocity, evolved to present day in the parent potential, $\Phi$, parametrized by the parameter vector $\bs{\Theta}_\Phi$.
-We assume that once the star becomes ``unbound'' from the satellite the potential of the satellite can be ignored, and thus treat the star as a test particle.\footnote{This assumption breaks down for sufficiently large-mass progenitors, somewhere between $\sim10^8-10^9~$M$_\odot$.} The satellite mass enters the model through the tidal radius; a changing satellite mass will skew the positions of the effective Lagrange points and velocity scale.
+We assume that once the star becomes ``unbound'' from the satellite the potential of the satellite can be ignored,\footnote{This assumption breaks down for sufficiently large-mass progenitors, somewhere between $\sim10^8-10^9~$M$_\odot$.} and thus treat the star as a test particle.
+The satellite mass enters the model through the tidal radius; a changing satellite mass will skew the positions of the effective Lagrange points and velocity scale.
 We allow for mass-loss by incorporating the initial satellite mass, $m_0$, and a constant mass-loss term, $\dot{m}$ into the model: $m(t) = m_0 - \dot{m}t$.
@@
 We instead use an affine-invariant ``ensemble'' sampler \citep{goodman10} that, each step in parameter space, updates the positions of many ``walkers'' (the ensemble).
-This algorithm is implemented in the \project{Python} programming language \citep{foremanmackey13} and runs naturally in a parallelized environment, e.g., the message passing interface (MPI).
+This algorithm is implemented in the \project{Python} programming language \citep{foremanmackey13} and runs naturally in a parallelized environment such as the message passing interface (MPI).
 In each experiment we run the walkers for a large number of steps from starting positions described below, then throw away these samples and take the positions from the final step of this ``burn-in'' phase as a starting point for the samples used for inference.
@@
 
-% TODO: m and mdot!!
 \begin{table}[h]
@@
 		Lagrange pt. offset & $\alpha$ & $\mathcal{N}(\mu=2,\sigma^2=1)$\\
+		initial mass & $m_0$ & $2.5\times10^8~M_\odot$ (fixed)\\
+		mass loss & $\dot{m}$ & $3.2\times10^4~M_\odot$/Myr (fixed)\\
 		\toprule
@@
 		velocity & $\bs{v}$ & $\|\bs{v}\|\sim\mathcal{N}(\mu=0,\sigma^2=4000)$~km/s\\
-		tail assignment (binary) & $\beta$ & $\pm1$~equally likely\\ % TODO: how to say equal weight to -1 or 1?
+		tail assignment (binary) & $\beta$ & $\pm1$~equally likely (fixed at truth)\\ % TODO: how to say equal weight to -1 or 1?
 %		prob. shocked & $\pshock$ & $\mathcal{U}(0,1)$\\
-		time unbound & $\tub$ & $\mathcal{U}(0,\tint)$\\ % TODO: talk about \tint (integration time) in text!
+		time unbound & $\tub$ & $\mathcal{U}(0,6.2)~Gyr$\\ % TODO: talk about \tint (integration time) in text!
 		\bottomrule
 		\end{tabular}
-	\caption{Parameter values used ...\label{tbl:params}}
+	\caption{Parameter values used in the experiments of Section~\ref{sec:experiments}.
+$\mathcal{N}$ is the normal (Gaussian) distribution, and $\mathcal{U}$ the uniform distribution.
+There are 11 parameters for the Milky Way potential, but only four are left free to vary; some parameters are fixed (denoted by ``(fixed)'') at the true values used in the n-body simulations that generated the fake data.
+The progenitor has nine parameters --- the position, $\bs{r}_{\rm p}$, and velocity, $\bs{v}_{\rm p}$, vectors each contain three components --- but only five are left free to vary.
+The sky coordinates (e.g., Galactic $l$, $b$) are assumed to be known with negligible uncertainty.
+Each star has eight associated parameters, five of which are allowed to vary.
+Sky coordinates are fixed, along with the tail assignment (whether the star belongs to the leading or trailing tail).
+For inference with eight stars, there are $4+5+8\times5=49$ free parameters.
+\label{tbl:params}}
 \end{center}
 \end{table}
-% TODO: in caption, redefine N is normal U is uniform. Star parameters are repeated for each star. summarize number of parameters for milky way, progenitor, etc. -- how to read the parameter
 
@@
 
-We first test the model using eight particles (stars) --- four from the leading tail, and four from the trailing tail --- randomly sampled from the $2.5\times10^8~M_\odot$ simulation (Section~\ref{sec:sims}), assuming the observed 6D positions for both the stars and the progenitor have negligible uncertainties.
+We first test the model using only eight stars (particles) --- four from the leading tail, and four from the trailing tail --- randomly sampled from the $2.5\times10^8~M_\odot$ simulation (Section~\ref{sec:sims}), assuming the observed 6D positions for both the stars and the progenitor have negligible uncertainties.
 The stars were required to have been stripped after the first pericentric passage and have a present-day distance within $50$ kpc of the Sun.
@@
 
+% TODO: slight ~1% bias in values due to particular sample of stars?
+
 \begin{figure}[!ht]
@@
 \subsection{Precise distance measurements with missing proper motions}\label{sec:exp3}
-The SMASH survey \citep{smashprop} will be completed in 2014, long before the final \gaia\, data release.
+The SMASH survey \citep{smashprop} will be completed within a year(early 2015), long before the final \gaia\, data release.
 Thus, we will soon have precise distance measurements for stars in the Sgr and Orphan streams, but poor proper motion constraints.
@@
 
-\subsection{Near-future uncertainties but no observed progenitor}
-Most of the streams and debris structures discovered thus far have undetected or unconfirmed progenitors.
+\subsection{Current data}\label{sec:exp4}
+% Most of the streams and debris structures discovered thus far have undetected or unconfirmed progenitors.
 We expect the phase-space positions of the stars to provide constraints on the position of the progenitor even if it is unobserved.
 For this experiment, we consider a situation where we have precise (2\%) distance measurements, 10~km/s radial velocity uncertainties, and no proper motion measurements to eight stream stars (the same used in previous experiments), and no knowledge of the 6D position of the progenitor object.
+Without measured proper motions, the extremely precise distance measurements combined with moderate precision radial velocities are sufficient for constraining the properties of the debris and the potential.
+How important are these distance measurements?
+We now test \rewinder\, on just eight stars assuming the best uncertainties available for existing data of the Sgr stream.
+For the stream stars, we adopt 15\% distance uncertainties \citep[typical for M giants, e.g.,][]{majewski04}, 0.1~mas/yr proper motion uncertainties \citep{koposov13}, and 5~km/s radial velocity uncertainties \citep{majewski04}.
+For the progenitor (Sgr core), we adopt a 1~kpc distance uncertainty \citep{kunder09}, 0.2~mas/yr proper motion uncertainties \citep{pryor10}, and 10~km/s radial velocity uncertainty \citep{frinchaboy12}.
+We again neglect uncertainties in the sky coordinates.
 
-Figures~\ref{fig:exp4_potential}, \ref{fig:exp4_satellite}, \ref{fig:exp4_particle} show the marginalized posteriors for the potential parameters, satellite parameters, and parameters for a single particle.
+Again, this experiment includes 49 model parameters in total.
+We use an ensemble of 256 walkers and also initialize the walkers by sampling the coordinate parameters from Gaussians centered on the true positions with variances set to 1\% of the parameter values.
+The potential parameters, $\Loffset$, and $\tub$ are initialized by drawing from the priors summarized in Table~\ref{tbl:params}.
+We again burn in the walkers for 10000 steps and run for 10000 inference steps.
+Figures~\ref{fig:exp4_potential}, \ref{fig:exp4_satellite}, \ref{fig:exp4_particle} show the marginalized posteriors for the potential parameters, satellite parameters, and parameters for the same star shown in previous figures.
+%[TODO: ..statement about results]
 
@@
 \includegraphics[width=\textwidth]{../../plots/paper2/exp4_potential.png}
-\caption{Same as Figure~\ref{fig:exp2_potential} but with no a priori knowledge of the 6D position of the progenitor.}\label{fig:exp4_potential}
+\caption{Same as Figure~\ref{fig:exp2_potential} but with realistic, present-day uncertainties. }\label{fig:exp4_potential}
 \end{center}
@@
 \includegraphics[width=\textwidth]{../../plots/paper2/exp4_satellite.png}
-\caption{ Same as Figure~\ref{fig:exp2_satellite} but with no a priori knowledge of the 6D position of the progenitor.}\label{fig:exp4_satellite}
+\caption{ Same as Figure~\ref{fig:exp2_satellite} but with realistic, present-day uncertainties. }\label{fig:exp4_satellite}
 \end{center}
@@
 \includegraphics[width=\textwidth]{../../plots/paper2/exp4_particle.png}
-\caption{ Same as Figure~\ref{fig:exp2_particle} but with no a priori knowledge of the 6D position of the progenitor.}\label{fig:exp4_particle}
+\caption{ Same as Figure~\ref{fig:exp2_particle} but with realistic, present-day uncertainties. }\label{fig:exp4_particle}
 \end{center}
@@
 \section{Discussion}
-In Section~\ref{sec:intro} we identified several issues that have only been partially solved by other stream modeling methods; \rewinder\, addresses each of these.
+% In Section~\ref{sec:intro} we identified several issues that have only been partially solved by other stream modeling methods; \rewinder\, addresses each of these.
 %\begin{enumerate}
@@
 %\end{enumerate}
-
-Observational uncertainties and missing data dimensions are naturally handled in the inferences by including the true 6D positions of stars and the progenitor as parameters in the model.
-This is a very important aspect [...because errors will be a function of distance, to properly model debris requires accounting for the uncertainties].
+% are naturally handled in the inferences 
+Any method that uses tidal debris as a potential measure must also model the uncertainties on the kinematic measurements because the observational uncertainties will be significant.
+\rewinder\, handles observational uncertainties and missing data dimensions by including the true 6D positions of stars and the progenitor as parameters in the model.
 In this article we have tested \rewinder\, using data of varied quality for a small sample of stars.
 In future work, we will investigate the effect of reducing the data quality but increasing the sample size.
-%[the power of this method is in small samples with high-quality data]
 
@@
 The orbits of the stars and progenitor are directly integrated and only require a function that evaluates the acceleration due to the potential at a given position (and time).
-The experiments in this article infer potential parameters from the same functional potential used in the n-body simulations that generated the fake data.
-In future work, we will (1) try fitting incorrect potential models to see if we can still recover global properties, e.g., flattening; (2) run simulations with smoothly changing potentials \citep[e.g.,][]{buist14} and model the time dependence; and (3) explore using a generic, non-parametric potential form (e.g., a basis function expansion) as the recovery potential model.
+The simple experiments in this article infer potential parameters from the same functional potential used in the n-body simulations that generated the fake data.
+In future work, we will (1) try fitting incorrect potential models to see if we can still recover global properties, e.g., flattening; (2) run simulations with smoothly changing potentials \citep[e.g.,][]{buist14} and attempt to model the time dependence; and (3) explore using a generic, non-parametric potential form (e.g., a basis function expansion) as the recovery potential model.
 
-Each streams will provide constraints on different aspects of the potential.
+Each stream will provide constraints on different properties of the potential.
 For example, more eccentric streams may better constrain the radial profile \citep[see][who illustrate the power of using multiple streams to simultaneously constrain the potential using orbit fitting]{deg14}.
@@
 
-Each star adds constraints on the potential parameters and does not require matching a simulated density to the observed density of stars.
-
-[Issues about computation and scaling.]
-The inference is much faster than full n-body modeling because the orbits of the stars and progenitor are integrated as test particles.
+The inference is much faster than full n-body modeling because the stars and progenitor are treated as test particles.
 Each step in parameter space with \rewinder\, requires integrating the orbits of stars and the progenitor: we presently integrate with a fixed timestep using leapfrog integration.
-Computing the acceleration due to the given potential at each step is implemented in \project{Cython} (and approaches \project{ANSI-C}-like speed), but the rest of the code is written in pure-\project{Python}.
+Computing the acceleration due to the given potential at each step is implemented in \project{Cython} (and approaches \project{C}-like speed), but the rest of the code is written in pure-\project{Python}.
 Though already significantly less computationally intensive than running a full n-body simulation for every parameter step, running each MCMC walker for 10000 steps requires parallelization and a few hours of CPU time on a compute cluster.
@@
 
-% TODO above: now address each point in more detail, shortcomings, what we'll do in future work
-
-Although the \gaia\, project will deliver hundreds of millions of stars useful for dynamical inference,
-  it still might be that the most relevant stellar samples for inferring the Milky Way potential might be very small in the end.
-For instance, stars that produce good distance estimates might be much more valuable than typical stars in the sample.
-At halo distances, \gaia's parallaxes are not expected to be better than the distance indicators produced by the best
-  variable stars observed at the best wavelengths; for example, near-infrared observations of RR Lyrae stars
-  can in principle deliver $\sim$2\% percent photometric distances at essentially all halo distances \citep{madore12}.
-These valuable stars are rare (and their abundances are age and metallicity dependent);
-  there could be many cold structures in the Milky Way halo that are highly constraining
-  on the potential in principle, but which contain only a few good distance-indicating members.
-That said, it is not yet known what the trade-offs are between having many stars at low precision and a few at high precision,
-  nor is it known how valuable distance information really is, when a structure contains many precisely observed members.
+One point of concern is that the number of parameters scales with eight times the number of stars: each star has six coordinate parameters (the true position), the unbinding time, and the tail assignment.
+Computational constraints limit the sample size, although it might be that the most relevant stellar samples for inferring the Milky Way potential are small.
+For instance, stars that produce good distance estimates might be much more valuable than typical stars in the sample so that we may limit to only variable stars, e.g., RR Lyraes.
+These valuable stars are rare (and their abundances are age and metallicity dependent); there could be many cold structures in the Milky Way halo that are highly constraining on the potential in principle, but which contain only a few good distance-indicating members.
+With \rewinder, each star adds constraints on the potential parameters and does not require matching a simulated density to the observed density of stars, thus this method works well even for small samples of well-measured stream stars.
+That said, it is not yet known what the trade-offs are between having many stars at low precision and a few at high precision, nor is it known how valuable distance information really is, when a structure contains many precisely observed members.
 
@@
   inference depends on the non-linear dynamical evolution of populations in phase space.
-The nonlinearities, combined with heteroskedastic data and requirements to marginalize over phase-space distribution functions,
+The nonlinearities, 
+  combined with heteroskedastic data and requirements to marginalize over phase-space distribution functions,
   lead naturally to situations in which some stars are \emph{far more informative} than others.
@@
 
-[Cold vs. Hot streams]
+In this article, we have only applied \rewinder\, to simulated observations of a kinematically ``hot'' stream; a large mass satellite will have a large dispersion in both position and velocity space.
+It has been shown previously that simulated cold streams are better than hot streams at recovering potential properties (cite?), however it is not clear whether this remains true with realistic observational uncertainties on the stream stars.
+For instance, when the uncertainties ``blur'' out the kinematic overdensity that is the stream, it is not known whether a cold stream would still be more informative than an equal sample of stars in a hot stream such as Sgr.
+This blurring could also make it more difficult to determine stream membership.
 
-[bow-tie structure, Kuepper work, and so on]
+We have focused on modeling streams of stars formed from a disrupting satellite on a mildly eccentric orbit.
+On more eccentric orbits, the debris tends to form more shell or cloud-like structures.
+The morphological distinction between stellar streams and stellar clouds is just related to the orbit of the disrupting satellite within the parent galaxy.
+With a proper model for the tidal shocking that occurs on an eccentric orbit, \rewinder\, could also be applied to structures such as the Triangulum-Andromeda (cite) or Hercules-Aquila (cite) stellar clouds, both possible shell-like structures viewed from the inside.
 
-[What's unrealistic about what we have done?  The kinematic model of
-  the progenitor is exceedingly simplistic; why are we okay with it
-  nonetheless but what could we do to improve it?  Related to this,
-  the progenitor doesn't shrink with time (as it should).  What's up
-  with that?]
+We model tidal debris as having been stripped from the progenitor with some dispersion about effective Lagrange points.
+We assume that the distributions in position and velocity at these points are isotropic Gaussians.
+It is clear from Figures~\ref{fig:lpts_r} and \ref{fig:lpts_v} that there is more structure to the positions and velocities of stars around the time they are stripped.
+For instance, the distribution of positions of the trailing debris is more extended and possibly offset from zero in the $\hat{\bs{x}}_2$ direction.
+We have shown that using this simple model, \emph{precise} measurements of the Milky Way potential are attainable with near-future data, however we have not discussed the \emph{accuracy} of such measurements.
+Bonaca et al.
+(in prep.) show that fitting a rigid potential model to the potential of the \project{Via Lactea} halo ...
+%\cite{bonaca14}
 
-[Haven't explored parameter space...]
-For experiments in this article with realistic uncertainties we have not performed a search in parameter space for the optimal values, and instead started the MCMC walkers with a small dispersion around the true parameter values.
-
-[Nothing special about streams; will work on any structures, including
-  shells and the like.]
+%[Haven't explored parameter space...]
+%For the second two experiments (Section~\ref{sec:exp3} and \ref{sec:exp4}), we have not performed a proper search in parameter space for the optimal values.
+Instead, the MCMC walkers were started with a small dispersion around the true parameter values.
+We have shown that if started from truth, the walkers will [...].
 

### 2014-03-28 text/paper2/ms.tex
started on kathryns comments

@@
 These processes combined with the gravity from a baryonic disk or ellipsoid can act to make the inner halo more oblate or spherical, however they do not seem to erase the clumpy, triaxial nature of the outer halo \citep[e.g.,][]{pontzen12}.
-This can lead to radially-dependent axis ratios, orientation, and smoothness, thus suggesting that the true mass distributions around Milky Way-like galaxies are not easily represented by integrable potentials.
+This can lead to radially-dependent axis ratios, orientation, and smoothness, thus suggesting that the true mass distributions around Milky Way-like galaxies are not easily represented by simple potentials.
+[Methods that seek to measure the potential of the Milky Way must be flexible enough to handle generic potential forms.
+We take a first step towards this goal with the likelihood function presented below; it works in phase-space and relies on orbit integration, not on computing actions or approximate potentials...]
 
 The bulk of the baryonic matter in galaxies spans roughly 5-10\% of the extent of the host dark matter halo.
-Hence, the most observable components of a galaxy are sensitive to the inner portion of the host halos mass distribution.
-For example, the rotation curves of disk galaxies trace the inner mass, especially since matter in disks can be assumed to move on nearly circular orbits.
-Conversely, measurements of the dark matter distribution at large radii is complicated by the low density of visible tracers, observational difficulties, and assumptions about the smoothness and analytic form.
+Hence, the brightest and most easily observable components of a galaxy are sensitive to the inner portion of the host halos mass distribution.
+For example, the rotation curves of disk galaxies trace the inner mass with exquisite sensitivity since matter in disks can be assumed to move on nearly circular orbits.
+Measurements of the dark matter distribution at large radii is complicated by the low density of visible tracers, observational difficulties of measuring kinematics of stars at large distances, and unknown orbits.
 Around external galaxies, the extended mass distribution has been studied using a variety of approaches \citep[see][for a a complete and detailed review]{courteau13}.
-For example, the kinematics of tracer populations such as globular clusters or planetary nebulae can be used to derive mass estimates under the assumptions that these satellite systems are relaxed and well-mixed \citep[early investigations include][]{mendez01,cote03}.
+For example, the kinematics of tracer populations such as globular clusters or planetary nebulae can be used to derive mass estimates under the assumptions that these satellite systems are on random orbits and are well-mixed in orbital phase \citep[early investigations include][]{mendez01,cote03}.
 Simple, parameterized models of both the mass and orbit distribution have been simultaneously constrained using such data \citep[e.g.][]{napolitano11,deason12c}.
@@
 Such studies assume that the tracer orbits are randomly sampled from a smooth distribution function and are fully phase mixed.
-However, large photometric surveys such as the SDSS have discovered large amounts of substructure --- in streams and associations of stars --- in the Milky Way halo, thus demonstrating that the stellar distribution is not phase-mixed.
-Since the initial discovery of tidal debris from the Sagittarius (Sgr) dwarf spheroidal galaxy, myriad other streams and debris systems have been identified probing Galactocentric distances from $\sim$15-100 kpc.
+However, large photometric surveys such as the SDSS and 2MASS \citep{TODO} have discovered large amounts of substructure --- in streams and associations of stars --- in the Milky Way halo \citep{TODO many}, thus demonstrating that the stellar distribution is not fully phase-mixed.
 The existence of large-scale substructure in stellar streams and clouds is known to bias mass and velocity inferences by several tens of percent \citep{yencho06}.
@@
 Another approach to using halo stars as potential measures is to exploit the non-random nature of halo.
-For example, \citet[][LM10]{law10} used n-body simulations of the disruption of the Sgr dwarf to simultaneously model the debris and underlying Milky Way potential.
-By varying parameters of the Milky Way potential, they found ``best-fit'' parameters by comparing the properties of observed Sgr stars to their simulated debris over a grid of potential parameters.
-The computational costs of running n-body simulations limited their search to a grid of values and forced the authors to fix many other parameters (e.g., properties of the disk and bulge).
-Nevertheless, they were able to constrain the shape of the Milky Way potential out to $\sim$70 kpc and found that the best-fitting halo has a prolate, triaxial shape elongated in the Galactic Z direction, orthogonal to the plane of the disk.
-Though an unlikely orientation for the halo --- e.g., \cite{debattista13} find that the disk of the Milky Way would not remain stable in such a configuration --- LM10 showed that the data is at a state where such inference can be attempted.
-The computational costs associated with n-body modeling has motivated many methods that approximately model tidal streams.
-
 Tidal streams are dynamically cold systems --- debris typically have small distributions of energy and angular momentum --- and thus require orders of magnitude fewer tracers than a random sample to get constraints of comparable accuracy to, e.g., Jeans analysis.
@@
 Then we can imagine measuring the full-space velocities ${\bf v}$ at different points ${\bf x}$ in the structure (e.g., along a stream), which would give us a direct measure of differences in a potential, $\Phi$, without running a full model of the disruption.
-An extension of this idea is to fit a single orbit to observed debris \citep[e.g.,][]{koposov10, deg13}.
+
+\citet[][LM10]{law10} used n-body simulations of the disruption of the Sgr dwarf to simultaneously model the debris and underlying Milky Way potential.
+By varying parameters of the Milky Way potential, they found ``best-fit'' parameters by comparing the properties of observed Sgr stars to their simulated debris.
+The computational costs of running n-body simulations limited their search to a grid of potential parameters and forced the authors to fix many other parameters (e.g., properties of the disk and bulge).
+Nevertheless, they were able to constrain the shape of the Milky Way potential out to $\sim$70 kpc and found that the best-fitting halo has a prolate, triaxial shape elongated in the Galactic Z direction, orthogonal to the plane of the disk.
+Though an unlikely orientation for the halo --- e.g., \cite{debattista13} find that the disk of the Milky Way would not remain stable in such a configuration --- LM10 showed that the data is at a state where such inference can be attempted.
+
+The computational costs associated with n-body modeling has motivated many methods that approximately model tidal streams.
+The simplest alternative is to fit a single orbit to observed debris \citep[e.g.,][]{koposov10, deg13}.
 Though this is known to be incorrect and leads to biases in inferred properties of the underlying potential \citep[e.g.,][]{eyre11, lux13, sanders13a}, \cite{deg14} and \cite{lux13} have used orbit fitting to demonstrate the power of combining multiple streams in dynamical inference.
@@
 We will \emph{not} assume that streams follow the orbit of the progenitor.
-Other statistical methods have been proposed \citep[][]{penarrubia12, apw13, sanderson14} that may prove powerful when applied to, e.g., data from the \gaia\, mission, where full 6D coordinates will be known for large samples of stars in the halo --- and therefore many debris structures --- but stream membership is not known for all stars.
+Other statistical methods have been proposed \citep[][]{penarrubia12, johnston99, sanderson14} that may prove powerful when applied to, e.g., data from the \gaia\, mission, where full 6D coordinates will be known for large samples of stars in the halo --- and therefore many debris structures --- but stream membership is not known for all stars.
 
-We identify a number of problems with existing approximate models; each method mentioned above solves only a subset of these issues:
+Motivated by the strengths of previous work, we identify a minimum set of issues that any approximate stream modeling or potential recovery method should address (see Section~\ref{sec:discussion} for a more detailed discussion of these points in the context of this work):
 \begin{enumerate}
-	\item The known debris structures are $\gtrsim$10 kpc from the Sun, thus while overdensities in the halo are less contaminated by background, the constituent stars often have incomplete or poorly measured phase-space information.
-Thus it is critical for any approximate model of tidal debris to incorporate observational uncertainties and missing dimensions in a consistent and justified way.
-	\item The potential of the inner galaxy is conceivably well-described by a (set of) nearly-integrable potentials, thus the dynamics of collisionless systems is plausibly captured by a transformation to some integral-of-motion-space, e.g., action-angle coordinates.
-However there is large uncertainty in the radial profile, shape, orientation, and graininess of the outer halo and the constancy of these parameters over distance.
+	\item \textbf{Observational uncertainties:} The known debris structures are $\gtrsim$10 kpc from the Sun, thus it is critical for any approximate model of tidal debris to incorporate observational uncertainties and missing dimensions in a consistent and justified way.
+	\item \textbf{Form of the potential:} There is large uncertainty in the radial profile, shape, orientation, and graininess of the outer halo and the constancy of these parameters over distance.
 Properly describing the potential may require complex models where computing actions is non-trivial or even impossible.
-	\item Near-future photometric surveys such as \gaia\, and the \project{LSST} will likely discover many new streams and kinematic associations of stars.
+	\item \textbf{Multiple debris structures:} Near-future photometric surveys such as \gaia\, and the \project{LSST} will likely discover many new streams and kinematic associations of stars.
 Potential recovery methods should be able to simultaneously model multiple streams and incorporate other dynamical constraints.
-	\item Matching generated models to observed stream densities is difficult: models that rely on this must account for observational biases and properly model the underlying background of halo stars. 
-	\item Full N-body simulations are expensive to run.
+	\item \textbf{Comparing models to data:} Matching generated models to observed stream densities is difficult; models that rely on this must account for observational biases and properly model the underlying background of halo stars.
+	\item \textbf{Computational expense:} Full N-body simulations are expensive to run.
 Incorporating an n-body model into a general model and then performing a parameter search would be computationally intensive and, presently, intractable.
@@
 
-In \citet{apw13}, we introduced a simple method for using individual stars associated with tidal debris combined with knowledge about the mass and orbit of the progenitor to constrain properties of the host galaxy potential.
+In \citet{apw13}, we introduced a simple method based on the work of \citet{johnston99} for using individual stars associated with tidal debris combined with knowledge about the mass and orbit of the progenitor to constrain properties of the host galaxy potential.
 The method exploits the relationship between the phase-space distribution of debris and (measurable) properties of the progenitor system (e.g., the tidal radius and escape velocity).
 Specifically, a better model for the potential is one in which orbits of test particle stars (integrated backwards from their present position) came close (within the tidal radius in position, and escape velocity in velocity) to the orbit of the progenitor.
-In this article, we use these simple scalings --- which depend only on the mass and orbit of the progenitor and parent potential --- to present a fully probabilistic method (\rewinder) for modeling tidal streams in any underlying potential given data with realistic uncertainties.
+
+%TODO: this paragraph...
+In this paper, we use these simple scalings --- which depend only on the mass and orbit of the progenitor and parent potential --- to present a fully probabilistic method (\rewinder) for modeling tidal streams in any underlying potential given data with realistic uncertainties.
 By ``model'' we mean a justified, parametrized likelihood function with priors on the parameters.
@@
 
-\section{Method}
+\section{Method (\rewinder)}
 
@@
 
-At pericenter, where a large fraction of the stars are unbound, the velocity of the satellite will be of order $V\sim \sqrt{GM_{\rm enc}/R}$.
-If the velocity dispersion of the satellite is $\sigma_v \sim \sqrt{Gm/r_J}$, then
+The velocity of the satellite will be of order $V\sim \sqrt{GM_{\rm enc}/R}$. If the velocity dispersion of the satellite is $\sigma_v \sim \sqrt{Gm/r_J}$, then
 \begin{equation}
@@
 We expect the debris-star velocities also to scale with $(m/M_{\rm enc})^{1/3}$.
-These scalings assume that the satellite is spherical with isotropic velocities; correlations among intra-satellite orbits from, e.g., a disk will break these assumptions.
+These scalings assume that the satellite is spherical with isotropic velocities; non-random, internal satellite orbits (e.g., a disk) will break these assumptions.
 
@@
 
-\subsection{Probabilistic model (\rewinder)}
-We observe the 6D position of a star, $\D$, in heliocentric coordinates --- e.g., the measured position on the sky, ($l$, $b$); distance, $d$; proper motions, ($\mu_l$, $\mu_b$); and line-of-sight velocity, $v_r$ --- and have determined through some other means that this star was once part of a progenitor system (e.g., satellite galaxy) with mass $m(t)$ that is disrupting and forming a cold debris structure in the potential, $\Phi$, of the parent galaxy (e.g., the Milky Way).
+\subsection{Probabilistic model}
+Suppose we observe the 6D position of a star, $\D$, in heliocentric coordinates --- e.g., the measured position on the sky, ($l$, $b$); distance, $d$; proper motions, ($\mu_l$, $\mu_b$); and line-of-sight velocity, $v_r$ --- and have determined through some other means that this star was once part of a progenitor system (e.g., satellite galaxy) with mass $m(t)$ that is disrupting and forming a cold debris structure in the potential, $\Phi$, of the parent galaxy (e.g., the Milky Way).
 We assume that the mass of the parent system enclosed within the pericenter of the orbit of the satellite, $M_{\rm enc}(R_{\rm peri})$, is much larger than the initial mass of the satellite, $M_{\rm enc}(R_{\rm peri})\gg m(t=0)$.
@@
 
-To model the true, 6D, present-day position of the star, $\W$, we first transform to cartesian, Galactocentric coordinates where ($\bs{r},\bs{v}$) and ($\bs{r}_\sat,\bs{v}_\sat$) are the position and velocity of the star and progenitor.
-In cartesian coordinates, the present-day position of the star is taken as having been sampled at $t=\tub$ from an isotropic Gaussian centered on one of the instantaneous, effective Lagrange points in position, and an isotropic Gaussian centered on the origin in velocity, evolved to present day in the parent potential, $\Phi$, parametrized by the parameter vector $\bs{\Theta}_\Phi$.
+To model the true, 6D, present-day position of the star, $\W$, we first transform to cartesian, Galactocentric coordinates where ($\bs{r}_0,\bs{v}_0$) and ($\bs{r}_{\sat,0},\bs{v}_{\sat,0}$) are the position and velocity of the star and progenitor today.
+The star is taken as having been sampled at $t=\tub$ (the unbinding time) from an isotropic Gaussian centered on one of the effective Lagrange points in position, and an isotropic Gaussian centered on the origin in velocity.
+The present-day position is a result of integration in the parent potential, $\Phi$, whose form is parametrized by the vector $\bs{\Theta}_\Phi$.
 We assume that once the star becomes ``unbound'' from the satellite the potential of the satellite can be ignored,\footnote{This assumption breaks down for sufficiently large-mass progenitors, somewhere between $\sim10^8-10^9~$M$_\odot$.} and thus treat the star as a test particle.
@@
 \begin{align}
-	p(\bs{\Theta} \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi) &= p(\bs{r} \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi)\,p(\bs{v} \given\bs{\Theta}_\sat, \bs{\Theta}_\Phi)\,\left\vert\J\right\vert_{t=\tub}\label{eq:starlike}\\
-	p(\bs{r} \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi) &= \mathcal{N}(\bs{r} \given \bs{r}_\sat - \Loffset\,\tailbit\,\rtide\,\hat{x}_1, \rtide^2)_{t=\tub}\\
-	p(\bs{v} \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi) &= \mathcal{N}(\bs{v} \given \bs{v}_\sat, \sigma^2_v)_{t=\tub}
+	p(\bs{\Theta} \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi) &= p(\bs{r}_0,\bs{v}_0, \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi)\,\left\vert\J\right\vert_{t=\tub}\label{eq:starlike}\\
+	p(\bs{r}_0,\bs{v}_0, \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi) &= \mathcal{N}(\bs{r} \given \bs{r}_\sat - \Loffset\,\tailbit\,\rtide\,\hat{x}_1, \rtide^2)\,\mathcal{N}(\bs{v} \given \bs{v}_\sat, \sigma^2_v)_{t=\tub}
 \end{align}
-%\begin{align}
-%	p(\W \given \W_\sat, m_0, \dot{m}, \tub, \Phi, \Loffset, \tailbit) &= p(\bs{r} \given \bs{r}_\sat, m_0, \dot{m}, \tub, \Phi, \Loffset, \tailbit)\,p(\bs{v} \given \bs{v}_\sat, m_0, \dot{m}, \tub, \Phi)\,\left\vert\J\right\vert_{t=\tub}\\
-%	p(\bs{r} \given \bs{r}_\sat, m_0, \dot{m}, \tub, \Phi, \Loffset, \tailbit) &= \mathcal{N}(\bs{r} \given \bs{r}_\sat - \Loffset\,\tailbit\,\rtide\,\hat{x}_1, \rtide^2)_{t=\tub}\\
-%	p(\bs{v} \given \bs{v}_\sat, m_0, \dot{m}, \tub, \Phi) &= \mathcal{N}(\bs{v} \given \bs{v}_\sat, \sigma^2_v)_{t=\tub}
-%\end{align}
 where $\rtide$ is given by Eq.~\ref{eq:tidalradius}, and $\sigma_v$ by Eq.~\ref{eq:velscale}.
 $\left\vert\J\right\vert$ is the absolute value of the determinant of the Jacobian that defines the transformation from heliocentric, spherical to Galactocentric, cartesian coordinates.
-For a single star, in detail, the model works as follows:
+All quantities in this expression are evaluated at $t=\tub$.
+For a single star, the evaluation of the likelihood works as follows: given present-day coordinates for the star and progenitor and nuisance parameters ($\tub,\Loffset,\tailbit$),
 \begin{enumerate}
-	\item sample from observed distributions over the heliocentric positions (Eqs.~\ref{eq:obsstar},\ref{eq:obsprog}), or priors if positions are unknown, for the star and progenitor;
 	\item transform star and progenitor positions in heliocentric coordinates to Galactocentric, cartesian coordinates;
-	\item integrate both orbits backwards (as test particles) for the total interaction time, $t_{\rm int}\approx6$~Gyr;
+	\item integrate star and progenitor orbits backwards (as test particles) for the total interaction time, $t_{\rm int}\approx6$~Gyr;
 	\item at the timestep closest to $\tub$, transform the particle position to the relative, normalized coordinates defined in Section~\ref{sec:debris};
-	\item evaluate the likelihood according to which tail the star belongs to, $\tailbit$, (Eq.~\ref{eq:starlike}).
+	\item compute the likelihood given by Equation~\ref{eq:starlike}.
 \end{enumerate}
@@
 All parameters are summarized in Table~\ref{tbl:params}.
-
-This model works well when all stars are lost due to tidal pruning.
-Even on a mildly eccentric orbit, stars are preferentially stripped around the satellite's pericenter because of tidal shocking.
-We will explore properly modeling this process in future work.
 
@@
 In what follows, we use \rewinder\, to model ``observations'' of the n-body simulations described in Section~\ref{sec:sims}.
-For all tests in this article, we model the potential of the parent galaxy with the same functional form used in the simulations (Equation~\ref{eq:lm10}) --- we stress that the data are generated with n-body simulations and are not generated with anything resembling the likelihood function presented above.
-When recovering the potential, we hold fixed the disk and spheroid parameters (see Table~\ref{tbl:params}), along with two halo parameters: the scale radius, $r_h$, and one of the flattening components in the x-y plane, $q_2$.
-These halo parameters are degenerate with combinations of the other parameters --- e.g., setting $q_1=2.0$, $q_2=1.0$, $\phi=0^\circ$ is equivalent to setting $q_1=1.0$, $q_2=2.0$, $\phi=90^\circ$.
+For all tests in this article, we model the potential of the parent galaxy with the same functional form used in the simulations (Equation~\ref{eq:lm10}).
+In particular, we consider this a strong test of the method because the data are generated with n-body simulations and are not generated with anything resembling the likelihood function presented above.
+When recovering the potential, we hold fixed the disk and spheroid parameters (see Table~\ref{tbl:params}), along with one halo parameters: the scale radius, $r_h$.
 The priors on the remaining halo parameters are taken to be uniform over a conservative domain of realistic values: for $\vhalo$, 100-200~km/s corresponds to a range in solar circular velocities from $\sim$210-250 km/s (holding other parameters fixed); the range in axis ratios allow for prolate, oblate, and generic triaxiality; and $\phi$ is restricted to $\pm45^\circ$ around the true simulation value, $\phi = 97^\circ$.
@@
 		\toprule
-		position & $\bs{r}_\sat$ & $\|\bs{r}_\sat\|\sim\mathcal{U}(0,200)$~kpc \\
-		velocity & $\bs{v}_\sat$ & $\|\bs{v}_\sat\|\sim\mathcal{N}(\mu=0,\sigma^2=4000)$~km/s\\
+		position & $\bs{r}_{\sat,0}$ & $\|\bs{r}_{\sat,0}\|\sim\mathcal{U}(0,200)$~kpc \\
+		velocity & $\bs{v}_{\sat,0}$ & $\|\bs{v}_{\sat,0}\|\sim\mathcal{N}(\mu=0,\sigma^2=4000)$~km/s\\
 		Lagrange pt. offset & $\alpha$ & $\mathcal{N}(\mu=2,\sigma^2=1)$\\
@@
 		\toprule
-		position & $\bs{r}$ & $\|\bs{r}\|\sim\mathcal{U}(0,200)$~kpc \\
-		velocity & $\bs{v}$ & $\|\bs{v}\|\sim\mathcal{N}(\mu=0,\sigma^2=4000)$~km/s\\
+		position & $\bs{r}_0$ & $\|\bs{r}_0\|\sim\mathcal{U}(0,200)$~kpc \\
+		velocity & $\bs{v}_0$ & $\|\bs{v}_0\|\sim\mathcal{N}(\mu=0,\sigma^2=4000)$~km/s\\
 		tail assignment (binary) & $\beta$ & $\pm1$~equally likely (fixed at truth)\\ % TODO: how to say equal weight to -1 or 1?
-%		prob. shocked & $\pshock$ & $\mathcal{U}(0,1)$\\
 		time unbound & $\tub$ & $\mathcal{U}(0,6.2)~Gyr$\\ % TODO: talk about \tint (integration time) in text!
@@
 We now test \rewinder\, on just eight stars assuming the best uncertainties available for existing data of the Sgr stream.
-For the stream stars, we adopt 15\% distance uncertainties \citep[typical for M giants, e.g.,][]{majewski04}, 0.1~mas/yr proper motion uncertainties \citep{koposov13}, and 5~km/s radial velocity uncertainties \citep{majewski04}.
+For the stream stars, we adopt 15\% distance uncertainties \citep[typical for M-giants, e.g.,][]{majewski04}, 0.1~mas/yr proper motion uncertainties \citep{koposov13}, and 5~km/s radial velocity uncertainties \citep{majewski04}.
+Note that we neglect any systematic uncertainties or trends with metallicity which are known to effect M-giant distance measurement accuracy.
 For the progenitor (Sgr core), we adopt a 1~kpc distance uncertainty \citep{kunder09}, 0.2~mas/yr proper motion uncertainties \citep{pryor10}, and 10~km/s radial velocity uncertainty \citep{frinchaboy12}.
@@
 %	\item The inference is much faster than full n-body modeling because the orbits of the stars and progenitor are integrated as test particles.
+%\end{enumerate}
+%\begin{enumerate}
+%	\item The known debris structures are $\gtrsim$10 kpc from the Sun, thus while overdensities in the halo are less contaminated by background, the constituent stars often have incomplete or poorly measured phase-space information.
+Thus it is critical for any approximate model of tidal debris to incorporate observational uncertainties and missing dimensions in a consistent and justified way.
+%	\item The potential of the inner galaxy is conceivably well-described by a (set of) nearly-integrable potentials, thus the dynamics of collisionless systems is plausibly captured by a transformation to some integral-of-motion-space, e.g., action-angle coordinates.
+However there is large uncertainty in the radial profile, shape, orientation, and graininess of the outer halo and the constancy of these parameters over distance.
+Properly describing the potential may require complex models where computing actions is non-trivial or even impossible.
+%	\item Near-future photometric surveys such as \gaia\, and the \project{LSST} will likely discover many new streams and kinematic associations of stars.
+Potential recovery methods should be able to simultaneously model multiple streams and incorporate other dynamical constraints.
+%	\item Matching generated models to observed stream densities is difficult: models that rely on this must account for observational biases and properly model the underlying background of halo stars. 
+%	\item Full N-body simulations are expensive to run.
+Incorporating an n-body model into a general model and then performing a parameter search would be computationally intensive and, presently, intractable.
 %\end{enumerate}
@@
 
+This model works well when all stars are lost due to tidal pruning.
+Even on a mildly eccentric orbit, stars are preferentially stripped around the satellite's pericenter because of tidal shocking.
+We will explore properly modeling this process in future work.
 We have focused on modeling streams of stars formed from a disrupting satellite on a mildly eccentric orbit.

### 2014-03-28 text/paper2/ms.tex
more references

@@
 %\newcommand{\project}[1]{{\sffamily #1}}
+\newcommand{\msun}{\mathrm{M}_\odot}
 
@@
 % Aims
-We develop a justified method for inferring the gravitational potential
+We develop a statistically justified method for inferring the gravitational potential
   using precise measurements of stars in tidal streams
-  based on an approximate generative model.
-Our method does not rely on integrability of the potential;
+  based the idea that these stars
+  were close in phase space at some point in the past.
+Our method is flexible enough to adapt to any form for the Galactic potential.
+In particular, it works in phase-space rather than action-space and hence relies neither on
+our ability to derive actions nor on the 
+assumption of integrability of the potential;
   we do not know that any good approximation to the large-scale 
@@
 Our model is probabilistic, with a likelihood function and priors on the parameters.
-It is based on the idea that the stars in a tidal structure
-  were close in phase space at some point in the past.
-  Model parameters include the present-day phase-space position for each star
-  and the progenitor, and thus properly accounts for finite observational uncertainties 
-  and missing data.
-This method is new in that it makes no assumption about the form of the Milky Way 
-  potential, accounts properly for observational uncertainties, and allows for missing data
-  dimensions.
-We test our method with simulated data in a non-integrable triaxial potential, with realistic 
-  observational uncertainties for RR Lyrae stars.
-Importantly, the fake data were generated with an n-body simulation,
-  not with the approximate model underlying our likelihood function.
+  These parameters include the present-day phase-space position for each star
+  and the progenitor, and thus the method can properly account for both finite observational uncertainties 
+  and missing data dimensions.
+We test our method on synthetic data sets generated from an N-body simulation
+of satellite disruption in a multi-component Milky Way including a triaxial dark matter halo. 
+The data sets were generated with
+  observational uncertainties to mimic current and near-future surveys of various stars.
 % Results:
 We find that even with a small number (8) of well-measured stream stars,
-  we can precisely infer properties of the triaxial potential and
-  missing phase-space coordinates of the stars.
+  we can infer properties of the triaxial potential ???? with accuracies of order ??? percent ??? and
+  missing phase-space coordinates of the stars ??? with accuracies of order ????.
 The method generalizes to encompass mixtures of independent debris structures,
@@
 Such studies assume that the tracer orbits are randomly sampled from a smooth distribution function and are fully phase mixed.
-However, large photometric surveys such as the SDSS and 2MASS \citep{TODO} have discovered large amounts of substructure --- in streams and associations of stars --- in the Milky Way halo \citep{TODO many}, thus demonstrating that the stellar distribution is not fully phase-mixed.
+However, large photometric surveys such as the SDSS and 2MASS \citep{skrutskie06} have discovered large amounts of substructure --- in streams and associations of stars --- in the Milky Way halo \citep[e.g.,][]{belokurov06, TODO}, thus demonstrating that the stellar distribution is not fully phase-mixed.
 The existence of large-scale substructure in stellar streams and clouds is known to bias mass and velocity inferences by several tens of percent \citep{yencho06}.
@@
 We will \emph{not} assume that streams follow the orbit of the progenitor.
-Other statistical methods have been proposed \citep[][]{penarrubia12, johnston99, sanderson14} that may prove powerful when applied to, e.g., data from the \gaia\, mission, where full 6D coordinates will be known for large samples of stars in the halo --- and therefore many debris structures --- but stream membership is not known for all stars.
+Other statistical methods have been proposed \citep[][]{penarrubia12, johnston99a, sanderson14} that may prove powerful when applied to, e.g., data from the \gaia\, mission, where full 6D coordinates will be known for large samples of stars in the halo --- and therefore many debris structures --- but stream membership is not known for all stars.
 
@@
 
-In \citet{apw13}, we introduced a simple method based on the work of \citet{johnston99} for using individual stars associated with tidal debris combined with knowledge about the mass and orbit of the progenitor to constrain properties of the host galaxy potential.
+In \citet{apw13}, we introduced a simple method based on the work of \citet{johnston99a} for using individual stars associated with tidal debris combined with knowledge about the mass and orbit of the progenitor to constrain properties of the host galaxy potential.
 The method exploits the relationship between the phase-space distribution of debris and (measurable) properties of the progenitor system (e.g., the tidal radius and escape velocity).
@@
 By ``model'' we mean a justified, parametrized likelihood function with priors on the parameters.
-\rewinder\, relies on numerical orbit integration in ordinary phase-space and can thus incorporate arbitrarily complex models for the parent potential.
-Individual stars are used as constraints, thus \rewinder\, can handle very small samples of well-measured stars (e.g., 4-16).
+\rewinder\ relies on numerical orbit integration in ordinary phase-space and can thus incorporate arbitrarily complex models for the parent potential.
+Individual stars are used as constraints, thus \rewinder\ can handle very small samples of well-measured stars (e.g., 4-16).
 This model trivially extends to incorporating multiple streams, debris structures, and other kinematic information.
@@
 \section{Simulations}\label{sec:sims}
-[David Hendel]
-We [...] n-body simulations run with an SCF code [...]. [describe simulations here] [Describe tub]
+We performed a set of N-body simulations with the SCF basis function expansion code \citep{hernquist92} to build realistic models of streams for studying the phase-space distribution of debris as it is stripped from its progenitor.
+We later use one of these simulations to test \rewinder.
+In each, a $10^5$ particle NFW-profile satellite was inserted at the apogalacticon of its orbit in a static, triaxial host halo, described below.
+The satellite was evolved first in isolation, then the host potential was turned on slowly over 10 satellite internal dynamical times to reduce artificial gravitational shocking.
+The initial position and velocity was obtained by integrating the orbit of the Sagittarius dwarf galaxy \citep[${\bf x}  =(19.0, 2.7, -6.9)\ \mathrm{kpc}, \ {\bf v} = (230., -35., 195.) \ \mathrm{km\ s^{-1}}$;][]{law10} backwards for 6 Gyr.
+The N-body satellite was then reintegrated from that initial position for the same interaction time and number of timesteps.
+This orbit has its apogalacticon at approximately 59 kpc and perigalacticon near 12 kpc with an average orbital period of 930 Myr, although all these quantities vary over the course of the simulation due to the halo's triaxiality.
+Total energy is conserved to $\sim$2\% of the satellite internal potential energy.
 
-The potential is taken to be a three-component sum of a Miyamoto-Nagai disk \citep{}, Hernquist bulge \citep[spheroid;][]{}, and a triaxial, logarithmic halo \citep[e.g.,][]{law10}:
+To capture the character of streams across a range of likely merger mass ratios, four satellites with masses $\mathrm{m} = \{2.5 \times 10^6,\ 2.5 \times 10^7,\ 2.5 \times 10^8,\ 2.5 \times 10^9\}\ \msun$, where m is the mass enclosed within 35 NFW scale radii, were evolved in the setup described above.
+Their scale radii, $r_0$ were adjusted to maintain a constant density across all simulations which results in identical fractional mass loss rates, removing 74\% of the initial mass by the end of the simulation; the base value is $\mathrm{r_0}=0.01565$ kpc at $\mathrm{m} = 2.5 \times 10^6\ \msun$.
+
+\rewinder\ [todo...] the phase-space positions of satellite particles at their unbinding time, therefore we take care to record the true unbinding times accurately in the simulations.
+We locate the position of the remnant iteratively by first calculating the satellite potential with all particles, removing particles with kinetic energies sufficient to escape, then recalculating the potential without those particles until the system converges.
+There is no spatial restriction on where a particle may become unbound and all particles (both bound and unbound) contribute their gravity during normal timesteps, presuming that they are resolved by the basis functions.
+
+The host potential is taken to be a three-component sum of a Miyamoto-Nagai disk \citep{miyamoto75}, Hernquist spheroid , and a triaxial, logarithmic halo \citep[e.g.,][]{law10}:
 \begin{align}
@@
 \includegraphics[width=\textwidth]{../../plots/paper2/simulated_streams.png}
-\caption{ Particle positions (grey dots) in Galactocentric cartesian coordinates from the final timestep of four n-body simulations with the same progenitor orbit initial conditions over a range of progenitor masses (columns).
-The progenitors were initialized as [TODO: DAH].
+\caption{ Particle positions (grey dots) in Galactocentric cartesian coordinates from the final timestep of four N-body simulations (Section~\ref{sec:sims}) with the same progenitor orbit initial conditions over a range of progenitor masses (columns).
 Black crosses indicate eight particles chosen from each mass simulation and used in the experiments described in Section~\ref{sec:experiments}.}\label{fig:sims}
@@
 \end{align}
-Figure~\ref{fig:lpts_r} shows sections of particle orbits (for particles that are stripped) from the n-body simulations described above projected into this coordinate system and normalized, symmetric around their unbinding times, $\tub$, for each of the four progenitor masses.
+Figure~\ref{fig:lpts_r} shows sections of particle orbits (for particles that are stripped) from the N-body simulations described above projected into this coordinate system and normalized, symmetric around their unbinding times, $\tub$, for each of the four progenitor masses.
 In these coordinates, the classical Lagrange points would be located at $x_1\approx\pm\rtide,x_2\approx0,x_3\approx0$ (illustrated by the intersection of the black unit circles and horizontal lines in Figure~\ref{fig:lpts_r}).
@@
 \caption{ Orbits of 2000 randomly drawn particles projected into the instantaneous orbital plane coordinates (Eqs.~\ref{eq:x1}-\ref{eq:x3}), normalized by the tidal radius (Eq.~\ref{eq:tidalradius}), and only shown within half a satellite crossing time around $t=\tub$ for each of the four progenitor masses.
-The orbits were integrated backwards from their present-day positions (final time step of the n-body simulations) as test particles without the potential of the progenitor.
+The orbits were integrated backwards from their present-day positions (final time step of the N-body simulations) as test particles without the potential of the progenitor.
 Horizontal black line shows $x_2=0$ (top panels) and $x_3=0$ (bottom panels), and the unit circle (black circle) shows the classical disruption radius in these coordinates.
@@
 The present-day position is a result of integration in the parent potential, $\Phi$, whose form is parametrized by the vector $\bs{\Theta}_\Phi$.
-We assume that once the star becomes ``unbound'' from the satellite the potential of the satellite can be ignored,\footnote{This assumption breaks down for sufficiently large-mass progenitors, somewhere between $\sim10^8-10^9~$M$_\odot$.} and thus treat the star as a test particle.
+We assume that once the star becomes ``unbound'' from the satellite the potential of the satellite can be ignored,\footnote{This assumption breaks down for sufficiently large-mass progenitors, somewhere between $\sim10^8-10^9~\msun$.} and thus treat the star as a test particle.
 The satellite mass enters the model through the tidal radius; a changing satellite mass will skew the positions of the effective Lagrange points and velocity scale.
@@
 \section{Experiments} \label{sec:experiments}
-In what follows, we use \rewinder\, to model ``observations'' of the n-body simulations described in Section~\ref{sec:sims}.
+In what follows, we use \rewinder\ to model ``observations'' of the N-body simulations described in Section~\ref{sec:sims}.
 For all tests in this article, we model the potential of the parent galaxy with the same functional form used in the simulations (Equation~\ref{eq:lm10}).
-In particular, we consider this a strong test of the method because the data are generated with n-body simulations and are not generated with anything resembling the likelihood function presented above.
+In particular, we consider this a strong test of the method because the data are generated with N-body simulations and are not generated with anything resembling the likelihood function presented above.
 When recovering the potential, we hold fixed the disk and spheroid parameters (see Table~\ref{tbl:params}), along with one halo parameters: the scale radius, $r_h$.
@@
 		Component & Parameter & Prior \\\toprule
-		disk & $M_{\rm disk}$ & $1.0\times10^{11}M_\odot$ (fixed) \\ 
+		disk & $M_{\rm disk}$ & $1.0\times10^{11}\,\msun$ (fixed) \\ 
 		& $a$ & 6.5 kpc (fixed)\\
@@
 		\midrule
-		spheroid & $M_{\rm spher}$ & $3.4\times10^{10}M_\odot$ (fixed)\\ 
+		spheroid & $M_{\rm spher}$ & $3.4\times10^{10}\,\msun$ (fixed)\\ 
 		& $c$ & 0.7 kpc (fixed)\\
@@
 		Lagrange pt. offset & $\alpha$ & $\mathcal{N}(\mu=2,\sigma^2=1)$\\
-		initial mass & $m_0$ & $2.5\times10^8~M_\odot$ (fixed)\\
-		mass loss & $\dot{m}$ & $3.2\times10^4~M_\odot$/Myr (fixed)\\
+		initial mass & $m_0$ & $2.5\times10^8\,\msun$ (fixed)\\
+		mass loss & $\dot{m}$ & $3.2\times10^4\,\msun$/Myr (fixed)\\
 		\toprule
@@
 $\mathcal{N}$ is the normal (Gaussian) distribution, and $\mathcal{U}$ the uniform distribution.
-There are 11 parameters for the Milky Way potential, but only four are left free to vary; some parameters are fixed (denoted by ``(fixed)'') at the true values used in the n-body simulations that generated the fake data.
+There are 11 parameters for the Milky Way potential, but only four are left free to vary; some parameters are fixed (denoted by ``(fixed)'') at the true values used in the N-body simulations that generated the fake data.
 The progenitor has nine parameters --- the position, $\bs{r}_{\rm p}$, and velocity, $\bs{v}_{\rm p}$, vectors each contain three components --- but only five are left free to vary.
@@
 
-We first test the model using only eight stars (particles) --- four from the leading tail, and four from the trailing tail --- randomly sampled from the $2.5\times10^8~M_\odot$ simulation (Section~\ref{sec:sims}), assuming the observed 6D positions for both the stars and the progenitor have negligible uncertainties.
+We first test the model using only eight stars (particles) --- four from the leading tail, and four from the trailing tail --- randomly sampled from the $2.5\times10^8\,\msun$ simulation (Section~\ref{sec:sims}), assuming the observed 6D positions for both the stars and the progenitor have negligible uncertainties.
 The stars were required to have been stripped after the first pericentric passage and have a present-day distance within $50$ kpc of the Sun.
@@
 Solid contours (black lines) show approximately 1$\sigma$ and 2$\sigma$ levels of the distribution.
-Vertical and horizontal lines (blue) show the true, input values for the potential parameters used in the n-body simulations.
+Vertical and horizontal lines (blue) show the true, input values for the potential parameters used in the N-body simulations.
 For the potential parameters, the axis ranges were chosen to be $\pm$5\% around the true values.
@@
 How important are these distance measurements?
-We now test \rewinder\, on just eight stars assuming the best uncertainties available for existing data of the Sgr stream.
+We now test \rewinder\ on just eight stars assuming the best uncertainties available for existing data of the Sgr stream.
 For the stream stars, we adopt 15\% distance uncertainties \citep[typical for M-giants, e.g.,][]{majewski04}, 0.1~mas/yr proper motion uncertainties \citep{koposov13}, and 5~km/s radial velocity uncertainties \citep{majewski04}.
@@
 \section{Discussion}
-% In Section~\ref{sec:intro} we identified several issues that have only been partially solved by other stream modeling methods; \rewinder\, addresses each of these.
+% In Section~\ref{sec:intro} we identified several issues that have only been partially solved by other stream modeling methods; \rewinder\ addresses each of these.
 %\begin{enumerate}
@@
 Any method that uses tidal debris as a potential measure must also model the uncertainties on the kinematic measurements because the observational uncertainties will be significant.
-\rewinder\, handles observational uncertainties and missing data dimensions by including the true 6D positions of stars and the progenitor as parameters in the model.
-In this article we have tested \rewinder\, using data of varied quality for a small sample of stars.
+\rewinder\ handles observational uncertainties and missing data dimensions by including the true 6D positions of stars and the progenitor as parameters in the model.
+In this article we have tested \rewinder\ using data of varied quality for a small sample of stars.
 In future work, we will investigate the effect of reducing the data quality but increasing the sample size.
@@
 The orbits of the stars and progenitor are directly integrated and only require a function that evaluates the acceleration due to the potential at a given position (and time).
-The simple experiments in this article infer potential parameters from the same functional potential used in the n-body simulations that generated the fake data.
+The simple experiments in this article infer potential parameters from the same functional potential used in the N-body simulations that generated the fake data.
 In future work, we will (1) try fitting incorrect potential models to see if we can still recover global properties, e.g., flattening; (2) run simulations with smoothly changing potentials \citep[e.g.,][]{buist14} and attempt to model the time dependence; and (3) explore using a generic, non-parametric potential form (e.g., a basis function expansion) as the recovery potential model.
@@
 
-The inference is much faster than full n-body modeling because the stars and progenitor are treated as test particles.
-Each step in parameter space with \rewinder\, requires integrating the orbits of stars and the progenitor: we presently integrate with a fixed timestep using leapfrog integration.
+The inference is much faster than full N-body modeling because the stars and progenitor are treated as test particles.
+Each step in parameter space with \rewinder\ requires integrating the orbits of stars and the progenitor: we presently integrate with a fixed timestep using leapfrog integration.
 Computing the acceleration due to the given potential at each step is implemented in \project{Cython} (and approaches \project{C}-like speed), but the rest of the code is written in pure-\project{Python}.
-Though already significantly less computationally intensive than running a full n-body simulation for every parameter step, running each MCMC walker for 10000 steps requires parallelization and a few hours of CPU time on a compute cluster.
+Though already significantly less computationally intensive than running a full N-body simulation for every parameter step, running each MCMC walker for 10000 steps requires parallelization and a few hours of CPU time on a compute cluster.
 Further optimization of the integration method (e.g., using an adaptive method or implementing in \project{Cython}) could speed up the inference significantly.
@@
 
-In this article, we have only applied \rewinder\, to simulated observations of a kinematically ``hot'' stream; a large mass satellite will have a large dispersion in both position and velocity space.
+In this article, we have only applied \rewinder\ to simulated observations of a kinematically ``hot'' stream; a large mass satellite will have a large dispersion in both position and velocity space.
 It has been shown previously that simulated cold streams are better than hot streams at recovering potential properties (cite?), however it is not clear whether this remains true with realistic observational uncertainties on the stream stars.
@@
 The morphological distinction between stellar streams and stellar clouds is just related to the orbit of the disrupting satellite within the parent galaxy.
-With a proper model for the tidal shocking that occurs on an eccentric orbit, \rewinder\, could also be applied to structures such as the Triangulum-Andromeda (cite) or Hercules-Aquila (cite) stellar clouds, both possible shell-like structures viewed from the inside.
+With a proper model for the tidal shocking that occurs on an eccentric orbit, \rewinder\ could also be applied to structures such as the Triangulum-Andromeda (cite) or Hercules-Aquila (cite) stellar clouds, both possible shell-like structures viewed from the inside.
 
@@
 We have presented a probabilistic model (\rewinder) --- a likelihood function and with priors on the parameters --- for using stars observed in tidal streams to constrain properties of any underlying gravitational potential.
-\rewinder\, relies on direct orbit integration and not on computing conserved quantities (e.g., actions) and can thus be used with arbitrarily complex (e.g., time-dependent or clumpy) forms of the potential.
-We have performed several experiments to show that \rewinder\, simultaneously constrains the potential and models a tidal stream given simulated data with a range of realistic uncertainties.
+\rewinder\ relies on direct orbit integration and not on computing conserved quantities (e.g., actions) and can thus be used with arbitrarily complex (e.g., time-dependent or clumpy) forms of the potential.
+We have performed several experiments to show that \rewinder\ simultaneously constrains the potential and models a tidal stream given simulated data with a range of realistic uncertainties.
 

### 2014-03-28 streams/inference/core.py
change start truth for uniform

@@ -259,7 +259,7 @@ class StreamModel(object):
                         prior = param._prior
                         if hasattr(prior, 'a') and hasattr(prior, 'b'):
                             mu = np.ravel(param.truth)
-                            sigma = (prior.b-prior.a)/2.
+                            sigma = mu/100.
                             a = (prior.a - mu) / sigma
                             b = (prior.b - mu) / sigma
                             if param.size > 1:

### 2014-03-31 text/paper2/ms.tex
spacing
remove notes

@@
 % packages
-\usepackage{amssymb,amsmath, amsbsy}
+\usepackage{amssymb,amsmath,amsbsy}
 \usepackage{booktabs}
@@
 		velocity & $\bs{v}_0$ & $\|\bs{v}_0\|\sim\mathcal{N}(\mu=0,\sigma^2=4000)$~km/s\\
-		tail assignment (binary) & $\beta$ & $\pm1$~equally likely (fixed at truth)\\ % TODO: how to say equal weight to -1 or 1?
-		time unbound & $\tub$ & $\mathcal{U}(0,6.2)~Gyr$\\ % TODO: talk about \tint (integration time) in text!
+		tail assignment (binary) & $\beta$ & $\pm1$~equally likely (fixed at truth)\\ 
+		time unbound & $\tub$ & $\mathcal{U}(0,6.2)~Gyr$\\ 
 		\bottomrule

### 2014-04-07 text/paper2/ms.tex
added todo
comments to self

@@
 These processes combined with the gravity from a baryonic disk or ellipsoid can act to make the inner halo more oblate or spherical, however they do not seem to erase the clumpy, triaxial nature of the outer halo \citep[e.g.,][]{pontzen12}.
-This can lead to radially-dependent axis ratios, orientation, and smoothness, thus suggesting that the true mass distributions around Milky Way-like galaxies are not easily represented by simple potentials.
+This can lead to radially-dependent axis ratios, orientation, and smoothness, and suggests that the true mass distributions around Milky Way-like galaxies are not easily represented by simple potentials.
 [Methods that seek to measure the potential of the Milky Way must be flexible enough to handle generic potential forms.
@@
 For example, the kinematics of tracer populations such as globular clusters or planetary nebulae can be used to derive mass estimates under the assumptions that these satellite systems are on random orbits and are well-mixed in orbital phase \citep[early investigations include][]{mendez01,cote03}.
-Simple, parameterized models of both the mass and orbit distribution have been simultaneously constrained using such data \citep[e.g.][]{napolitano11,deason12c}.
+Simple, parameterized fits to both the mass and orbit distribution have been simultaneously constrained using such data \citep[e.g.][]{napolitano11,deason12c}.
 Alternatively, the statistical properties of gravitationally lensed background sources around a galaxy can be used to constrain the \emph{projected} shape, orientation, and radial profile of mass \citep[see, for example, the Lens Structure and Dynamics Survey described in][]{koopmans02}.
@@
 
-Within the Milky Way our unique vantage point allows us a three-dimensional view of the structure in own dark matter halo.
+Within the Milky Way our unique vantage point allows us a three-dimensional view of stellar substructure in our own dark matter halo.
 Our proximity allows us to use individual stars as kinematic tracers and hence build much larger samples that probe deeper into the halo than the globular cluster and planetary nebula studies of external galaxies.
-For example, \cite{deason12a} used halo BHB stars selected from the Sloan Digital Sky Survey \cite[SDSS;][]{york00} as a random tracer population to measure the mass and slope of a power-law fit.
+For example, \cite{deason12a} used halo BHB stars selected from the Sloan Digital Sky Survey \cite[SDSS;][]{york00} as a random tracer population to measure the mass and slope of a power-law fit to the potential.
 Such studies assume that the tracer orbits are randomly sampled from a smooth distribution function and are fully phase mixed.
 However, large photometric surveys such as the SDSS and 2MASS \citep{skrutskie06} have discovered large amounts of substructure --- in streams and associations of stars --- in the Milky Way halo \citep[e.g.,][]{belokurov06, TODO}, thus demonstrating that the stellar distribution is not fully phase-mixed.
-The existence of large-scale substructure in stellar streams and clouds is known to bias mass and velocity inferences by several tens of percent \citep{yencho06}.
+Substructure in the form of stellar streams and clouds is known to bias mass and velocity inferences by several tens of percent \citep{yencho06}.
 
 Another approach to using halo stars as potential measures is to exploit the non-random nature of halo.
-Tidal streams are dynamically cold systems --- debris typically have small distributions of energy and angular momentum --- and thus require orders of magnitude fewer tracers than a random sample to get constraints of comparable accuracy to, e.g., Jeans analysis.
+Tidal streams are dynamically cold systems --- debris typically have small distributions of energy and angular momentum --- and thus require orders of magnitude fewer tracers than a random sample to get constraints of comparable accuracy to Jeans analysis.
 For example, in the simplest case we might assume that debris stars are actually still on the same orbit as their progenitor system \citep[a \emph{wrong} assumption, see e.g.,][]{eyre11}.
@@
 
-\citet[][LM10]{law10} used n-body simulations of the disruption of the Sgr dwarf to simultaneously model the debris and underlying Milky Way potential.
+\citet[][LM10]{law10} used N-body simulations of the disruption of the Sgr dwarf galaxy to simultaneously model the debris and underlying Milky Way potential.
 By varying parameters of the Milky Way potential, they found ``best-fit'' parameters by comparing the properties of observed Sgr stars to their simulated debris.
-The computational costs of running n-body simulations limited their search to a grid of potential parameters and forced the authors to fix many other parameters (e.g., properties of the disk and bulge).
-Nevertheless, they were able to constrain the shape of the Milky Way potential out to $\sim$70 kpc and found that the best-fitting halo has a prolate, triaxial shape elongated in the Galactic Z direction, orthogonal to the plane of the disk.
-Though an unlikely orientation for the halo --- e.g., \cite{debattista13} find that the disk of the Milky Way would not remain stable in such a configuration --- LM10 showed that the data is at a state where such inference can be attempted.
+The computational costs of running N-body simulations limited their search to a grid of potential parameters and forced the authors to fix many other parameters (e.g., properties of the disk and bulge).
+Nevertheless, they were able to constrain the 3D shape of the Milky Way potential out to $\sim$70 kpc and found that the best-fitting halo has a prolate, triaxial shape elongated in the Galactic Z direction, orthogonal to the plane of the disk.
+Though an unlikely orientation for the halo --- \cite{debattista13} find that the disk of the Milky Way would not remain stable in such a configuration --- LM10 showed that the data is at a state where such inference can be attempted.
 
-The computational costs associated with n-body modeling has motivated many methods that approximately model tidal streams.
+The computational costs associated with N-body modeling has motivated the development of many methods that approximately model tidal streams.
 The simplest alternative is to fit a single orbit to observed debris \citep[e.g.,][]{koposov10, deg13}.
@@
 To account for the offset between the orbit of the progenitor and the orbits of the debris stars, methods have been proposed that model the debris as an orbit plus some dispersion or offset, either in phase space \citep[e.g.,][]{eyre09a, varghese11, kuepper12} or action-angle coordinates \citep{eyre11, sanders13b, bovy14, sanders14}.
-We will \emph{not} assume that streams follow the orbit of the progenitor.
-Other statistical methods have been proposed \citep[][]{penarrubia12, johnston99a, sanderson14} that may prove powerful when applied to, e.g., data from the \gaia\, mission, where full 6D coordinates will be known for large samples of stars in the halo --- and therefore many debris structures --- but stream membership is not known for all stars.
+In this work we do \emph{not} assume that streams follow the orbit of the progenitor.
+Other statistical methods have been proposed \citep[][]{penarrubia12, johnston99a, sanderson14} that may prove powerful when applied to, for example, data from the \gaia\, mission, where full 6D coordinates will be known for large samples of stars in the halo --- and therefore many debris structures --- but stream membership is not known for all stars.
 
@@
 	\item \textbf{Form of the potential:} There is large uncertainty in the radial profile, shape, orientation, and graininess of the outer halo and the constancy of these parameters over distance.
-Properly describing the potential may require complex models where computing actions is non-trivial or even impossible.
+Properly describing the potential may require complex analytic functions where computing actions is non-trivial or even impossible.
 	\item \textbf{Multiple debris structures:} Near-future photometric surveys such as \gaia\, and the \project{LSST} will likely discover many new streams and kinematic associations of stars.
-Potential recovery methods should be able to simultaneously model multiple streams and incorporate other dynamical constraints.
-	\item \textbf{Comparing models to data:} Matching generated models to observed stream densities is difficult; models that rely on this must account for observational biases and properly model the underlying background of halo stars.
+Potential recovery methods should be able to simultaneously use multiple streams and incorporate other dynamical constraints.
+	\item \textbf{Comparing models to data:} Matching generated streams to observed stellar densities is difficult; models that rely on this must account for observational biases and properly model the underlying background of halo stars.
 	\item \textbf{Computational expense:} Full N-body simulations are expensive to run.
-Incorporating an n-body model into a general model and then performing a parameter search would be computationally intensive and, presently, intractable.
+Incorporating an N-body model into a general model and then performing a parameter search would be computationally intensive and, presently, intractable.
 \end{enumerate}
@@
 The method exploits the relationship between the phase-space distribution of debris and (measurable) properties of the progenitor system (e.g., the tidal radius and escape velocity).
-Specifically, a better model for the potential is one in which orbits of test particle stars (integrated backwards from their present position) came close (within the tidal radius in position, and escape velocity in velocity) to the orbit of the progenitor.
+Specifically, a better potential is one in which orbits of test particle stars (integrated backwards from their present position) came close (within the tidal radius in position, and escape velocity in velocity) to the orbit of the progenitor at some time in the past.
 
-%TODO: this paragraph...
-In this paper, we use these simple scalings --- which depend only on the mass and orbit of the progenitor and parent potential --- to present a fully probabilistic method (\rewinder) for modeling tidal streams in any underlying potential given data with realistic uncertainties.
-By ``model'' we mean a justified, parametrized likelihood function with priors on the parameters.
+In this paper, we present a fully probabilistic model (\rewinder) for modeling tidal streams in any underlying potential given data with realistic uncertainties by building on these simple scalings, which depend only on the mass and orbit of the progenitor and parent potential.
+By ``probabilistic model'' we mean a justified, parametrized likelihood function with priors on the parameters.
 \rewinder\ relies on numerical orbit integration in ordinary phase-space and can thus incorporate arbitrarily complex models for the parent potential.
@@
 
-In Section~\ref{sec:sims} we describe a suite of n-body simulations that span a range of progenitor masses on a characteristic, mildly eccentric orbit. In Section~\ref{sec:} we describe [...]
+In Section~\ref{sec:sims} we describe a suite of N-body simulations that span a range of progenitor masses on a characteristic, mildly eccentric orbit. In Section~\ref{sec:} we describe [...]
 
@@
 We later use one of these simulations to test \rewinder.
-In each, a $10^5$ particle NFW-profile satellite was inserted at the apogalacticon of its orbit in a static, triaxial host halo, described below.
+In each simulation, a $10^5$ particle NFW-profile satellite was inserted at the apogalacticon of its orbit in a static, triaxial host halo, described below.
 The satellite was evolved first in isolation, then the host potential was turned on slowly over 10 satellite internal dynamical times to reduce artificial gravitational shocking.
-The initial position and velocity was obtained by integrating the orbit of the Sagittarius dwarf galaxy \citep[${\bf x}  =(19.0, 2.7, -6.9)\ \mathrm{kpc}, \ {\bf v} = (230., -35., 195.) \ \mathrm{km\ s^{-1}}$;][]{law10} backwards for 6 Gyr.
+The initial position and velocity was obtained by integrating the orbit of the Sagittarius dwarf galaxy from present-day coordinates ${\bf r}  =(19.0, 2.7, -6.9)\ \mathrm{kpc}, \ {\bf v} = (230., -35., 195.) \ \mathrm{km\ s^{-1}}$ \citep{law10} backwards for $\sim$6 Gyr.
 The N-body satellite was then reintegrated from that initial position for the same interaction time and number of timesteps.
-This orbit has its apogalacticon at approximately 59 kpc and perigalacticon near 12 kpc with an average orbital period of 930 Myr, although all these quantities vary over the course of the simulation due to the halo's triaxiality.
+This orbit has its apogalacticon at approximately 59 kpc and perigalacticon near 12 kpc with an average orbital period of 930 Myr, although all of these quantities vary over the course of the simulation due to the halo's triaxiality.
 Total energy is conserved to $\sim$2\% of the satellite internal potential energy.
 
-To capture the character of streams across a range of likely merger mass ratios, four satellites with masses $\mathrm{m} = \{2.5 \times 10^6,\ 2.5 \times 10^7,\ 2.5 \times 10^8,\ 2.5 \times 10^9\}\ \msun$, where m is the mass enclosed within 35 NFW scale radii, were evolved in the setup described above.
-Their scale radii, $r_0$ were adjusted to maintain a constant density across all simulations which results in identical fractional mass loss rates, removing 74\% of the initial mass by the end of the simulation; the base value is $\mathrm{r_0}=0.01565$ kpc at $\mathrm{m} = 2.5 \times 10^6\ \msun$.
+To capture the character of streams across a range of merger mass ratios, four satellites with masses $\mathrm{m} = 2.5 \times 10^6,\ 2.5 \times 10^7,\ 2.5 \times 10^8,\ 2.5 \times 10^9\ \msun$, where m is the mass enclosed within 35 NFW scale radii, were evolved in the setup described above.
+Their scale radii, $r_0$, were adjusted to maintain a constant density across all simulations which results in identical fractional mass loss rates: 74\% of the initial mass is lost by the end of each simulation.
+The base value is $\mathrm{r_0}=0.01565$ kpc at $\mathrm{m} = 2.5 \times 10^6\ \msun$.
 
-\rewinder\ [todo...] the phase-space positions of satellite particles at their unbinding time, therefore we take care to record the true unbinding times accurately in the simulations.
-We locate the position of the remnant iteratively by first calculating the satellite potential with all particles, removing particles with kinetic energies sufficient to escape, then recalculating the potential without those particles until the system converges.
+We take care to record the time at which each particle is unbound from the satellite in the simulations: we locate the position of the remnant iteratively by first calculating the satellite potential with all particles, remove particles with kinetic energies sufficient to escape, then recalculate the potential without those particles until the system converges.
 There is no spatial restriction on where a particle may become unbound and all particles (both bound and unbound) contribute their gravity during normal timesteps, presuming that they are resolved by the basis functions.
 
-The host potential is taken to be a three-component sum of a Miyamoto-Nagai disk \citep{miyamoto75}, Hernquist spheroid , and a triaxial, logarithmic halo \citep[e.g.,][]{law10}:
+In all simulations, the host potential is taken to be a three-component sum of a Miyamoto-Nagai disk \citep{miyamoto75}, Hernquist spheroid , and a triaxial, logarithmic halo \citep[e.g.,][]{law10}:
 \begin{align}
@@
 Figure~\ref{fig:potential} shows equipotential slices in the Galactic X-Z plane (Y=0) and Y-Z plane (X=0) for a few choices of  $q_1$,  $q_z$, and $\phi$ while holding all other parameters fixed, as described in the figure caption.
+The parameters used for the simulations are the same as used in \citet{law10}: [TODO].
 
@@
 
-For a satellite galaxy on a mildly eccentric orbit, mass loss is driven by a combination of tidal stripping by the steady tidal field of the parent system and tidal shocking by the rapidly changing tidal field as the progenitor system moves through pericenter \citep[e.g.,][]{choi09}.
+For a satellite galaxy orbiting within a static host potential, mass loss is driven by a combination of tidal stripping by the steady tidal field of the parent system and tidal shocking by the rapidly changing tidal field as the progenitor system moves through pericenter \citep[e.g.,][]{choi09}.
+On a mildly eccentric orbit, the tidal shocking is subdominant to the steady disruption.
 The interplay between these two processes as a function of orbital properties is outside our scope; we focus here on the slow removal of stars by steady tidal forcing.
 
-Consider a point-mass satellite of mass $m$ on a circular orbit with frequency $\Omega$ around a more massive mass $M\gg m$ \citep[the ``restricted three-body problem''; e.g., \S 8.3][]{binneytremaine}.
-In a frame rotating with the orbital frequency of the satellite, the satellite remains fixed and the static, effective potential around the satellite has (amongst others) two stationary, unstable points located around galactocentric radii $\sim R \pm r_J$, where $R$ is the orbital radius of the satellite and $r_J$ is the Jacobi or tidal radius
+Consider a point-mass satellite with mass $m$ on a circular orbit with frequency $\Omega$ around a more massive ``host'' mass $M\gg m$ \citep[the ``restricted three-body problem''; e.g., \S 8.3][]{binneytremaine}.
+In a frame rotating with the orbital frequency of the satellite, the satellite remains fixed and the static, effective potential around the satellite has (amongst others) two unstable optima located around galactocentric radii $\sim R \pm r_J$, where $R$ is the orbital radius of the satellite and $r_J$ is the Jacobi or tidal radius
 \begin{equation}
@@
 In general the satellite mass also depends on time.
-The distance scale of the debris then follows
+The tidal radius of the debris then follows
 \begin{equation}
@@
 \end{equation}
-where $M_{\rm enc}$ is the instantaneous enclosed mass of the parent system within orbital radius $R$.
+where $M_{\rm enc}$ is the instantaneous enclosed mass of the parent system within orbital radius $R$.\footnote{Note that this is the instantaneous orbital radius, \emph{not} the orbital radius at pericenter.}
 In general, the effective Lagrange points may not be symmetric about the center of the satellite and may deviate from the tidal radius by factors of order unity.
 
-The velocity of the satellite will be of order $V\sim \sqrt{GM_{\rm enc}/R}$. If the velocity dispersion of the satellite is $\sigma_v \sim \sqrt{Gm/r_J}$, then
+The velocity of the satellite will be of order $V\sim \sqrt{GM_{\rm enc}/R}$. If the velocity dispersion of the satellite is $\sigma_v \sim \sqrt{Gm/r_J}$, then it follows
 \begin{equation}
-	\sigma_v \sim V\left(\frac{m}{M_{\rm enc}}\right)^{1/3}\label{eq:velscale}.
+	\sigma_v(t) \sim V\left(\frac{m(t)}{M_{\rm enc}(t)}\right)^{1/3}\label{eq:velscale}.
 \end{equation}
@@
 
-In a triaxial potential, the orbital plane of the satellite is not fixed but we still expect there to be \emph{effective} Lagrange points for a spherical satellite system along the line of centers connecting the origin of the parent potential to the satellite --- that is, we expect the stars to be stripped at some characteristic radius with some dispersion about this point.
+In a triaxial potential, the orbital plane of the satellite is not fixed but we still expect there to be \emph{effective} Lagrange points for a spherical satellite system along the line of centers connecting the origin of the parent potential to the satellite --- that is, we expect the stars to be stripped at some characteristic radius near the tidal radius with some dispersion about this point.
 We proceed by defining a coordinate system relative to the position and velocity of the progenitor, ($\bs{r}_p$,$\bs{v}_p$), rotated into the instantaneous orbital plane defined by $\bs{L} = \bs{r}_p \times \bs{v}_p$.
-The positions are normalized by the instantaneous tidal radius (Eq.~\ref{eq:tidalradius}) and velocities are normalized by the instantaneous velocity scale (Eq.~\ref{eq:velscale}).
 The (time-dependent) basis vectors are given by
@@
 \end{align}
-Figure~\ref{fig:lpts_r} shows sections of particle orbits (for particles that are stripped) from the N-body simulations described above projected into this coordinate system and normalized, symmetric around their unbinding times, $\tub$, for each of the four progenitor masses.
-In these coordinates, the classical Lagrange points would be located at $x_1\approx\pm\rtide,x_2\approx0,x_3\approx0$ (illustrated by the intersection of the black unit circles and horizontal lines in Figure~\ref{fig:lpts_r}).
+Figure~\ref{fig:lpts_r} shows sections of particle orbits (for particles that are stripped) from the N-body simulations described above projected into this coordinate system.
+The positions are normalized by the instantaneous tidal radius (Eq.~\ref{eq:tidalradius}) and velocities are normalized by the instantaneous velocity scale (Eq.~\ref{eq:velscale}).
+Sections of the orbits symmetric around their unbinding times, $\tub$, are shown for each of the four progenitor masses.
+In these coordinates, the classical Lagrange points would be located at $x_1\approx\pm\rtide,x_2=0,x_3=0$ (illustrated by the intersection of the black unit circles and horizontal lines in Figure~\ref{fig:lpts_r}).
+Though not centered on the point-mass Lagrange points, the location of and dispersion about the effective Lagrange points in these scaled coordinates remain remarkably consistent across the range of progenitor masses explored.
 Figure~\ref{fig:lpts_v} shows the velocity orbits of each star also projected into these coordinates and normalized.
-Though not centered on the point-mass Lagrange points, the location of and dispersion about the effective Lagrange points in these scaled coordinates remain remarkably consistent across the range of progenitor masses explored.
-Furthermore, the dispersion in velocity is well-normalized by the velocity scale of Eq.~\ref{eq:velscale}, though it is clear that the velocity dispersion along $\hat{x}_1$, the radial vector to the satellite position, is larger than that in other dimensions, possibly because of tidal shocking.
+The dispersion in velocity is well-normalized by the velocity scale of Eq.~\ref{eq:velscale}, though it is clear that the velocity dispersion along $\hat{x}_1$, the radial vector to the satellite position, is larger than that in other dimensions, possibly because of tidal shocking.
 
@@
 The present position of the progenitor is observed to be at heliocentric position $\D_\sat$ (where any subscript $\sat$ refers to the progenitor).
-In general, the data for the star and progenitor will have missing dimensions and significant uncertainties at distances typical to the Galactic halo, thus we define $\W$ and $\W_\sat$ as the true 6D, heliocentric positions of the star and progenitor and will include these in our model.
-Since we include these positions as parameters, this model will work even when the star has missing dimensions or the progenitor location is unknown as in, e.g., the Orphan stream \citep{belokurov07}.
+In general, the data for the star and progenitor will have significant uncertainties or missing dimensions at distances typical to the Galactic halo, thus we define $\W$ and $\W_\sat$ as the true 6D, heliocentric positions of the star and progenitor and will include these in our model.
+Since we include these positions as parameters, this model will work even when the star has missing dimensions or the progenitor location is unknown \citep[as in the Orphan stream,][]{belokurov07}.
 
@@
 
+%TODO: remove exp 3 and explain that it is possible but computationally difficult...
+
 \subsection{Current data}\label{sec:exp4}
@@
 %	\item Each star adds constraints on the potential parameters and does not require matching a simulated density to the observed density of stars.
-%	\item The inference is much faster than full n-body modeling because the orbits of the stars and progenitor are integrated as test particles.
+%	\item The inference is much faster than full N-body modeling because the orbits of the stars and progenitor are integrated as test particles.
 %\end{enumerate}
@@
 %	\item Full N-body simulations are expensive to run.
-Incorporating an n-body model into a general model and then performing a parameter search would be computationally intensive and, presently, intractable.
+Incorporating an N-body model into a general model and then performing a parameter search would be computationally intensive and, presently, intractable.
 %\end{enumerate}

### 2014-04-07 streams/inference/back_integrate.py
new jacobian

@@ -29,8 +29,13 @@ def xyz_sph_jac(hel):
     cosb, sinb = np.cos(b), np.sin(b)
 
     Rsun = 8.
-    dtmnt = d**2*(Rsun**2*cosb + Rsun*d*sinb**2*cosl - 2*Rsun*d*cosl + d**2*sinb**4*cosb - d**2*cosb**5 + 2*d**2*cosb**3)*cosb
-    deet = np.log(np.abs(dtmnt))
+    #dtmnt = d**2*(Rsun**2*cosb + Rsun*d*sinb**2*cosl - 2*Rsun*d*cosl + d**2*sinb**4*cosb - d**2*cosb**5 + 2*d**2*cosb**3)*cosb
+    #dtmnt = np.log(np.abs(d**2*cosb)) + np.log(np.abs(Rsun**2*cosb - Rsun*d*cosb**2*cosl - Rsun*d*cosl + d**2*cosb))
+    dtmnt2 = np.log(np.abs(d**4*cosb))
+
+    #deet = np.log(np.abs(dtmnt))
+    deet = dtmnt
+
     return deet
 
 def back_integration_likelihood(t1, t2, dt, potential, p_gc, s_gc, logm0, logmdot,

### 2014-04-07 text/paper2/ms.tex
added discussion section label

@@
 The star is taken as having been sampled at $t=\tub$ (the unbinding time) from an isotropic Gaussian centered on one of the effective Lagrange points in position, and an isotropic Gaussian centered on the origin in velocity.
-The present-day position is a result of integration in the parent potential, $\Phi$, whose form is parametrized by the vector $\bs{\Theta}_\Phi$.
+The present-day phase-space position is the result of integrating this sample forward from $\tub$ in the parent potential, $\Phi$, whose form is parametrized by the vector $\bs{\Theta}_\Phi$.
 We assume that once the star becomes ``unbound'' from the satellite the potential of the satellite can be ignored,\footnote{This assumption breaks down for sufficiently large-mass progenitors, somewhere between $\sim10^8-10^9~\msun$.} and thus treat the star as a test particle.
@@
 We allow for mass-loss by incorporating the initial satellite mass, $m_0$, and a constant mass-loss term, $\dot{m}$ into the model: $m(t) = m_0 - \dot{m}t$.
-Finally, we add two additional parameters: a constant, global offset, $\Loffset$, that scales the position of the effective Lagrange points relative to the classical tidal radius, and a binary parameter $\tailbit$, that is either $-1$ or $+1$ depending on whether the star is in the leading or trailing tail.
+Finally, we add two additional parameters: a constant, global factor, $\Loffset$, that scales the position of the effective Lagrange points relative to the classical tidal radius, and a binary parameter $\tailbit$, that is either $-1$ or $+1$ depending on whether the star is in the leading or trailing tail.
 To compress notation, we pack all progenitor parameters ($\W_\sat, m_0, \dot{m}, \Loffset$) into the vector $\bs{\Theta}_\sat$, and all star parameters ($\W, \tailbit, \tub$) into the vector $\bs{\Theta}$.
@@
 
-\section{Discussion}
+\section{Discussion}\label{sec:discussion}
 % In Section~\ref{sec:intro} we identified several issues that have only been partially solved by other stream modeling methods; \rewinder\ addresses each of these.

### 2014-04-07 text/paper2/ms.tex
more notes to self

@@
 This can lead to radially-dependent axis ratios, orientation, and smoothness, and suggests that the true mass distributions around Milky Way-like galaxies are not easily represented by simple potentials.
-[Methods that seek to measure the potential of the Milky Way must be flexible enough to handle generic potential forms.
-We take a first step towards this goal with the likelihood function presented below; it works in phase-space and relies on orbit integration, not on computing actions or approximate potentials...]
+Methods that seek to measure the potential of the Milky Way must be flexible enough to handle generic potentials where finding approximate analytic forms or computing actions is not possible.
 
@@
 In this work we do \emph{not} assume that streams follow the orbit of the progenitor.
-Other statistical methods have been proposed \citep[][]{penarrubia12, johnston99a, sanderson14} that may prove powerful when applied to, for example, data from the \gaia\, mission, where full 6D coordinates will be known for large samples of stars in the halo --- and therefore many debris structures --- but stream membership is not known for all stars.
+Other statistical methods have been proposed \citep[][]{penarrubia12, johnston99a} that may prove powerful when applied to, for example, data from the \gaia\, mission, where full 6D coordinates will be known for large samples of stars in the halo --- and therefore many debris structures --- but stream membership is not known for all stars.
+%sanderson14
 
@@
 
-In Section~\ref{sec:sims} we describe a suite of N-body simulations that span a range of progenitor masses on a characteristic, mildly eccentric orbit. In Section~\ref{sec:} we describe [...]
+In Section~\ref{sec:sims} we describe a suite of N-body simulations that span a range of progenitor masses on a characteristic, mildly eccentric orbit.
+In Section~\ref{sec:method} we use these simulations to motivate a new, flexible model for tidal streams that works entirely in phase-space (\rewinder).
+In Section~\ref{sec:experiments} we demonstrate how \rewinder\ can be used to measure properties of a non-trivial Galactic potential by performing several experiments with simulated observations of data from N-body simulations.
+We discuss the results of these experiments in Section~\ref{sec:discussion}.
+We conclude in Section~\ref{sec:conclusion}.
 
@@
 Figure~\ref{fig:potential} shows equipotential slices in the Galactic X-Z plane (Y=0) and Y-Z plane (X=0) for a few choices of  $q_1$,  $q_z$, and $\phi$ while holding all other parameters fixed, as described in the figure caption.
-The parameters used for the simulations are the same as used in \citet{law10}: [TODO].
+The potential parameters used for the simulations are shown in the ``Truth'' column of Table~\ref{tbl:params}.
 
@@
 
-\section{Method (\rewinder)}
+\section{Method (\rewinder)}\label{sec:method}
 
@@
 All quantities in this expression are evaluated at $t=\tub$.
-For a single star, the evaluation of the likelihood works as follows: given present-day coordinates for the star and progenitor and nuisance parameters ($\tub,\Loffset,\tailbit$),
+For a single star, the evaluation of the likelihood works as follows: given present-day coordinates for the star and progenitor, potential parameter values ($\bs{\Theta}_\Phi$), and nuisance parameter values ($\tub,\Loffset,\tailbit$),
 \begin{enumerate}
 	\item transform star and progenitor positions in heliocentric coordinates to Galactocentric, cartesian coordinates;
-	\item integrate star and progenitor orbits backwards (as test particles) for the total interaction time, $t_{\rm int}\approx6$~Gyr;
+	\item integrate star and progenitor orbits backwards as test particles in the potential given by $\bs{\Theta}_\Phi$ for the total interaction time, $t_{\rm int}\approx6$~Gyr;
 	\item at the timestep closest to $\tub$, transform the particle position to the relative, normalized coordinates defined in Section~\ref{sec:debris};
@@
 \end{enumerate}
-Assuming each star is an independent tracer, the full posterior for many stars is then just the product of the individual likelihoods with the priors:
+Assuming each star is an independent tracer, the full likelihood for many stars is then just the product of the individual likelihoods:
 %\begin{multline}
-%	p(\Phi, \{\W^{(i)}\}, \W_\sat, m_0, \dot{m}, \tub, \Loffset, \tailbit, \given \{\D^{(i)}\}, \D_\sat) = \\
-%		\shoveleft{\hspace{1em}\frac{1}{\mathcal{Z}} \, \prod_i \, p(\D^{(i)} \given \W^{(i)}) \, p(\D_\sat \given \W_\sat)}\\
-%			\times p(\W^{(i)} \given \W_\sat, m_0, \dot{m}, \tub^{(i)}, \Phi, \Loffset, \tailbit^{(i)}) \, 
-%			\, p(\tub^{(i)}) \, p(\Phi) \, p(\Loffset)
+%	p(\{\bs{\Theta}^{(i)}\}, \bs{\Theta}_\sat, \bs{\Theta}_\Phi \given \{\D^{(i)}\}, \D_\sat) = \\
+%		\frac{1}{\mathcal{Z}} \, \prod_i \, p(\D^{(i)} \given \W^{(i)}) \, p(\D_\sat \given \W_\sat)\,
+%			p(\bs{\Theta}^{(i)} \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi) \, 
+%			p(\bs{\Theta}) \, p(\bs{\Theta}_\sat)\,p(\bs{\Theta}_\Phi)
 %\end{multline}
 \begin{multline}
-	p(\{\bs{\Theta}^{(i)}\}, \bs{\Theta}_\sat, \bs{\Theta}_\Phi \given \{\D^{(i)}\}, \D_\sat) = \\
-		\frac{1}{\mathcal{Z}} \, \prod_i \, p(\D^{(i)} \given \W^{(i)}) \, p(\D_\sat \given \W_\sat)\,
-			p(\bs{\Theta}^{(i)} \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi) \, 
-			p(\bs{\Theta}) \, p(\bs{\Theta}_\sat)\,p(\bs{\Theta}_\Phi)
+	p(\{\D^{(i)}\}, \D_\sat \given \{\bs{\Theta}^{(i)}\}, \bs{\Theta}_\sat, \bs{\Theta}_\Phi ) = \\
+		p(\D_\sat \given \W_\sat) \, \prod_i \, p(\D^{(i)} \given \W^{(i)}) \,
+			p(\bs{\Theta}^{(i)} \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi)
 \end{multline}
@@
 \begin{center}
-	\begin{tabular}{l c l} \toprule
-		\multicolumn{3}{l}{{\bf \emph{Milky Way parameters} ($\bs{\Theta}_\Phi$)}} \\
+	\begin{tabular}{l c c l} \toprule
+		\multicolumn{4}{l}{{\bf \emph{Milky Way parameters} ($\bs{\Theta}_\Phi$)}} \\
 		\toprule
-		Component & Parameter & Prior \\\toprule
-		disk & $M_{\rm disk}$ & $1.0\times10^{11}\,\msun$ (fixed) \\ 
-		& $a$ & 6.5 kpc (fixed)\\
-		& $b$ & 0.26 kpc (fixed)\\
+		Component & Parameter & Truth & Prior \\\toprule
+		disk & $M_{\rm disk}$ & $1.0\times10^{11}\,\msun$ & fixed \\ 
+		& $a$ & 6.5 kpc & fixed\\
+		& $b$ & 0.26 kpc & fixed\\
 		\midrule
-		spheroid & $M_{\rm spher}$ & $3.4\times10^{10}\,\msun$ (fixed)\\ 
-		& $c$ & 0.7 kpc (fixed)\\
+		spheroid & $M_{\rm spher}$ & $3.4\times10^{10}\,\msun$ & fixed\\ 
+		& $c$ & 0.7 kpc & fixed\\
 		\midrule
-		halo & $\vhalo$ & $\mathcal{U}(100,200)$ km/s \\
-		& $q_1$ & $\mathcal{U}(1,2)$\\
-		& $q_2$ & 1.0 (fixed)\\
-		& $q_z$ & $\mathcal{U}(1,2)$\\
-		& $\phi$ & $\mathcal{U}(52,142)$ deg\\
-		& $r_h$ & 12 kpc (fixed)\\
+		halo & $\vhalo$ & 121.858 & $\mathcal{U}(100,200)$ km/s \\
+		& $q_1$ & 1.38 & $\mathcal{U}(1,2)$\\
+		& $q_2$ & 1.0 & fixed\\
+		& $q_z$ & 1.36 & $\mathcal{U}(1,2)$\\
+		& $\phi$ & 97 & $\mathcal{U}(52,142)$ deg\\
+		& $r_h$ & 12 kpc & fixed\\
 		\toprule
-		\multicolumn{3}{l}{{\bf \emph{Progenitor parameters} ($\bs{\Theta}_\sat$)}} \\
+		\multicolumn{4}{l}{{\bf \emph{Progenitor parameters} ($\bs{\Theta}_\sat$)}} \\
 		\toprule
-		position & $\bs{r}_{\sat,0}$ & $\|\bs{r}_{\sat,0}\|\sim\mathcal{U}(0,200)$~kpc \\
-		velocity & $\bs{v}_{\sat,0}$ & $\|\bs{v}_{\sat,0}\|\sim\mathcal{N}(\mu=0,\sigma^2=4000)$~km/s\\
-		Lagrange pt. offset & $\alpha$ & $\mathcal{N}(\mu=2,\sigma^2=1)$\\
-		initial mass & $m_0$ & $2.5\times10^8\,\msun$ (fixed)\\
-		mass loss & $\dot{m}$ & $3.2\times10^4\,\msun$/Myr (fixed)\\
+		position & $\bs{r}_{\sat,0}$ & -- & $\|\bs{r}_{\sat,0}\|\sim\mathcal{U}(0,200)$~kpc \\
+		velocity & $\bs{v}_{\sat,0}$ & -- & $\|\bs{v}_{\sat,0}\|\sim\mathcal{U}(0,500)$~km/s\\
+		Lagrange pt. offset & $\alpha$ & -- & $\mathcal{U}(0.5, 2.5)$\\
+		initial mass & $m_0$ & $2.5\times10^8\,\msun$ & fixed\\
+		mass loss & $\dot{m}$ & -- & $3.2\times10^4\,\msun$/Myr (fixed)\\
 		\toprule
-		\multicolumn{3}{l}{{\bf \emph{Star parameters} ($\bs{\Theta}^{(i)}$)}} \\
+		\multicolumn{4}{l}{{\bf \emph{Star parameters} ($\bs{\Theta}^{(i)}$)}} \\
 		\toprule
-		position & $\bs{r}_0$ & $\|\bs{r}_0\|\sim\mathcal{U}(0,200)$~kpc \\
-		velocity & $\bs{v}_0$ & $\|\bs{v}_0\|\sim\mathcal{N}(\mu=0,\sigma^2=4000)$~km/s\\
-		tail assignment (binary) & $\beta$ & $\pm1$~equally likely (fixed at truth)\\ 
-		time unbound & $\tub$ & $\mathcal{U}(0,6.2)~Gyr$\\ 
+		position & $\bs{r}_0$ & -- & $\|\bs{r}_0\|\sim\mathcal{U}(0,200)$~kpc \\
+		velocity & $\bs{v}_0$ & -- & $\|\bs{v}_0\|\sim\mathcal{U}(0,500)$~km/s\\
+		tail assignment (binary) & $\beta$ & -- & $\pm1$~equally likely (fixed at truth)\\ 
+		time unbound & $\tub$ & -- & $\mathcal{U}(0,6.2)$ Gyr\\ 
 		\bottomrule
@@
 
-\section{Conclusions}
+\section{Conclusions}\label{sec:conclusion}
 We have presented a probabilistic model (\rewinder) --- a likelihood function and with priors on the parameters --- for using stars observed in tidal streams to constrain properties of any underlying gravitational potential.

### 2014-04-07 text/paper2/ms.tex
wording

@@
 Such studies assume that the tracer orbits are randomly sampled from a smooth distribution function and are fully phase mixed.
-However, large photometric surveys such as the SDSS and 2MASS \citep{skrutskie06} have discovered large amounts of substructure --- in streams and associations of stars --- in the Milky Way halo \citep[e.g.,][]{belokurov06, TODO}, thus demonstrating that the stellar distribution is not fully phase-mixed.
+However, large photometric surveys such as the SDSS and 2MASS \citep{skrutskie06} have discovered large amounts of substructure --- in streams and associations of stars --- in the Milky Way halo \citep[e.g.,][]{belokurov06, rochapinto04}, thus demonstrating that the stellar distribution is not fully phase-mixed.
 Substructure in the form of stellar streams and clouds is known to bias mass and velocity inferences by several tens of percent \citep{yencho06}.
@@
 For example, in the simplest case we might assume that debris stars are actually still on the same orbit as their progenitor system \citep[a \emph{wrong} assumption, see e.g.,][]{eyre11}.
-Then we can imagine measuring the full-space velocities ${\bf v}$ at different points ${\bf x}$ in the structure (e.g., along a stream), which would give us a direct measure of differences in a potential, $\Phi$, without running a full model of the disruption.
+This information about the orbits combined with measurements of the full-space velocities ${\bf v}$ at different points ${\bf x}$ in the structure (e.g., along a stream) would give us a direct measure of differences in a potential, $\Phi$.
 
-\citet[][LM10]{law10} used N-body simulations of the disruption of the Sgr dwarf galaxy to simultaneously model the debris and underlying Milky Way potential.
+\citet[][LM10]{law10} used N-body simulations of the disruption of the Sgr dwarf galaxy to simultaneously fit a model to the available data on the debris stars and a triaxial, analytic Milky Way potential.
 By varying parameters of the Milky Way potential, they found ``best-fit'' parameters by comparing the properties of observed Sgr stars to their simulated debris.
@@
 
-The computational costs associated with N-body modeling has motivated the development of many methods that approximately model tidal streams.
+The computational costs associated with N-body simulations has motivated the development of many methods that approximately model tidal streams.
 The simplest alternative is to fit a single orbit to observed debris \citep[e.g.,][]{koposov10, deg13}.
 Though this is known to be incorrect and leads to biases in inferred properties of the underlying potential \citep[e.g.,][]{eyre11, lux13, sanders13a}, \cite{deg14} and \cite{lux13} have used orbit fitting to demonstrate the power of combining multiple streams in dynamical inference.
-To account for the offset between the orbit of the progenitor and the orbits of the debris stars, methods have been proposed that model the debris as an orbit plus some dispersion or offset, either in phase space \citep[e.g.,][]{eyre09a, varghese11, kuepper12} or action-angle coordinates \citep{eyre11, sanders13b, bovy14, sanders14}.
+To account for the offset between the orbit of the progenitor and the orbits of the debris stars, methods have been proposed that add some dispersion or offset around a single orbit either in phase space \citep[e.g.,][]{eyre09a, varghese11, kuepper12} or action-angle coordinates \citep{eyre11, sanders13b, bovy14, sanders14}.
 In this work we do \emph{not} assume that streams follow the orbit of the progenitor.
@@
 
-Motivated by the strengths of previous work, we identify a minimum set of issues that any approximate stream modeling or potential recovery method should address (see Section~\ref{sec:discussion} for a more detailed discussion of these points in the context of this work):
+Motivated by the strengths of previous work, we identify a minimum set of issues that any approximate stream model or potential recovery method should address (see Section~\ref{sec:discussion} for a more detailed discussion of these points in the context of this work):
 \begin{enumerate}
-	\item \textbf{Observational uncertainties:} The known debris structures are $\gtrsim$10 kpc from the Sun, thus it is critical for any approximate model of tidal debris to incorporate observational uncertainties and missing dimensions in a consistent and justified way.
+	\item \textbf{Observational uncertainties:} The known debris structures are $\gtrsim$10 kpc from the Sun, thus it is critical for any method that uses tidal debris to incorporate observational uncertainties and missing dimensions in a consistent and justified way.
 	\item \textbf{Form of the potential:} There is large uncertainty in the radial profile, shape, orientation, and graininess of the outer halo and the constancy of these parameters over distance.
@@
 Potential recovery methods should be able to simultaneously use multiple streams and incorporate other dynamical constraints.
-	\item \textbf{Comparing models to data:} Matching generated streams to observed stellar densities is difficult; models that rely on this must account for observational biases and properly model the underlying background of halo stars.
-	\item \textbf{Computational expense:} Full N-body simulations are expensive to run.
-Incorporating an N-body model into a general model and then performing a parameter search would be computationally intensive and, presently, intractable.
+	\item \textbf{Comparing models to data:} Matching generated streams to observed stellar densities is difficult; models that rely on this must account for observational biases, internal properties of the progenitor, and background halo stars.
+	\item \textbf{Computational expense:} Full N-body simulations are expensive to run; incorporating an N-body simulation into a likelihood function evaluation and then performing a parameter search would be computationally intensive and, presently, intractable.
 \end{enumerate}
@@
 
-In this paper, we present a fully probabilistic model (\rewinder) for modeling tidal streams in any underlying potential given data with realistic uncertainties by building on these simple scalings, which depend only on the mass and orbit of the progenitor and parent potential.
+In this paper, we present a fully probabilistic model (\rewinder) for tidal streams that builds on these simple scalings, which depend only on the mass and orbit of the progenitor and parent potential.
 By ``probabilistic model'' we mean a justified, parametrized likelihood function with priors on the parameters.
-\rewinder\ relies on numerical orbit integration in ordinary phase-space and can thus incorporate arbitrarily complex models for the parent potential.
-Individual stars are used as constraints, thus \rewinder\ can handle very small samples of well-measured stars (e.g., 4-16).
-This model trivially extends to incorporating multiple streams, debris structures, and other kinematic information.
+\rewinder\ relies on numerical orbit integration in ordinary phase-space and can thus incorporate arbitrarily complex forms for the parent potential.
+Individual stars are constraints on the potential, thus \rewinder\ can handle very small samples of well-measured stars (e.g., 4-16).
+Incorporating multiple streams, debris structures, and other kinematic information is trivial but will be explored in future work.
 
 In Section~\ref{sec:sims} we describe a suite of N-body simulations that span a range of progenitor masses on a characteristic, mildly eccentric orbit.
-In Section~\ref{sec:method} we use these simulations to motivate a new, flexible model for tidal streams that works entirely in phase-space (\rewinder).
+In Section~\ref{sec:method} we use these simulations to motivate a new, flexible model for tidal streams that works entirely in phase-space.
 In Section~\ref{sec:experiments} we demonstrate how \rewinder\ can be used to measure properties of a non-trivial Galactic potential by performing several experiments with simulated observations of data from N-body simulations.
@@
 In general, the data for the star and progenitor will have significant uncertainties or missing dimensions at distances typical to the Galactic halo, thus we define $\W$ and $\W_\sat$ as the true 6D, heliocentric positions of the star and progenitor and will include these in our model.
-Since we include these positions as parameters, this model will work even when the star has missing dimensions or the progenitor location is unknown \citep[as in the Orphan stream,][]{belokurov07}.
+Since we include these positions as parameters, this will work even when the star has missing dimensions or the progenitor location is unknown \citep[as in the Orphan stream,][]{belokurov07}.
 
@@
 We assume that once the star becomes ``unbound'' from the satellite the potential of the satellite can be ignored,\footnote{This assumption breaks down for sufficiently large-mass progenitors, somewhere between $\sim10^8-10^9~\msun$.} and thus treat the star as a test particle.
-The satellite mass enters the model through the tidal radius; a changing satellite mass will skew the positions of the effective Lagrange points and velocity scale.
+The satellite mass enters this prescription through the tidal radius; a changing satellite mass will skew the positions of the effective Lagrange points and velocity scale.
 We allow for mass-loss by incorporating the initial satellite mass, $m_0$, and a constant mass-loss term, $\dot{m}$ into the model: $m(t) = m_0 - \dot{m}t$.
@@
 \section{Experiments} \label{sec:experiments}
-In what follows, we use \rewinder\ to model ``observations'' of the N-body simulations described in Section~\ref{sec:sims}.
-For all tests in this article, we model the potential of the parent galaxy with the same functional form used in the simulations (Equation~\ref{eq:lm10}).
-In particular, we consider this a strong test of the method because the data are generated with N-body simulations and are not generated with anything resembling the likelihood function presented above.
+In what follows, we use \rewinder\ to model ``observations'' of a small sample of stars from one of the N-body simulations described in Section~\ref{sec:sims}.
+We consider this a strong test of the method because the data are generated with N-body simulations and are not generated with anything resembling the likelihood function presented above.
+For all tests in this article, we use the same functional form for the potential as used in the simulations (Equation~\ref{eq:lm10}).
 When recovering the potential, we hold fixed the disk and spheroid parameters (see Table~\ref{tbl:params}), along with one halo parameters: the scale radius, $r_h$.
@@
 \subsection{Precise distance measurements with missing proper motions}\label{sec:exp3}
-The SMASH survey \citep{smashprop} will be completed within a year(early 2015), long before the final \gaia\, data release.
+The SMASH survey \citep{smashprop} will be completed within a year (early 2015), long before the final \gaia\, data release.
 Thus, we will soon have precise distance measurements for stars in the Sgr and Orphan streams, but poor proper motion constraints.

### 2014-04-08 text/paper2/ms.tex
discussion

@@
 
-\subsection{Motivation: tidal debris}\label{sec:debris}
-
+\rewinder\ integrates stars and the progenitor back from their present-day, observed positions to the time at which they become unbound from the satellite, where we evaluate the likelihood for each star.
+Below we (1) motivate a simple model for the debris just as it disrupts (near $\tub$ for each star) and (2) present this idea within a probabilistic framework for incorporating observational uncertainties and missing data.
 For a satellite galaxy orbiting within a static host potential, mass loss is driven by a combination of tidal stripping by the steady tidal field of the parent system and tidal shocking by the rapidly changing tidal field as the progenitor system moves through pericenter \citep[e.g.,][]{choi09}.
@@
 The interplay between these two processes as a function of orbital properties is outside our scope; we focus here on the slow removal of stars by steady tidal forcing.
+
+\subsection{Motivation: tidal debris}\label{sec:debris}
 
@@
 
-We first test the model using only eight stars (particles) --- four from the leading tail, and four from the trailing tail --- randomly sampled from the $2.5\times10^8\,\msun$ simulation (Section~\ref{sec:sims}), assuming the observed 6D positions for both the stars and the progenitor have negligible uncertainties.
+We first test \rewinder\ using only eight stars (particles) --- four from the leading tail, and four from the trailing tail --- randomly sampled from the $2.5\times10^8\,\msun$ simulation (Section~\ref{sec:sims}), assuming the observed 6D positions for both the stars and the progenitor have negligible uncertainties.
 The stars were required to have been stripped after the first pericentric passage and have a present-day distance within $50$ kpc of the Sun.
@@
 We next take the same eight stars used in the previous experiment and ``observe'' them with optimistic observational uncertainties.
-We take these stars to be RR Lyrae variables which are known to be excellent distance indicators via the mid-infrared period-luminosity relation \citep[as shown in, e.g.,][]{madore12}.
+We require these stars to be RR Lyrae variables which are known to be excellent distance indicators via the mid-infrared period-luminosity relation \citep[as shown in, e.g.,][]{madore12}.
 Nearly 100 RR Lyrae associated with the Sgr stream and $\sim$30 associated with the Orphan stream will be observed with \spitzer as part of the SMASH survey \citep{smashprop} with expected fractional distance uncertainties around $\sim$2\%.
@@
 We further assume that we know the tail assignment for each star, $\beta$.
-We observe the position of the progenitor with the same observational uncertainties, though in reality some coordinates will have higher or lower uncertainties.
+We observe the position of the progenitor with the same observational uncertainties though in reality some coordinates will have even better measurements.
 
@@
 Other parameters --- potential parameters, $\Loffset$, and $\tub$ --- are initialized by drawing from the priors summarized in Table~\ref{tbl:params}.
-We burn in the walkers for 10000 steps and run for 10000 inference steps.
+[TODO: start from truth...] We burn in the walkers for 10000 steps and run for 10000 inference steps.
 Figures~\ref{fig:exp2_potential}, \ref{fig:exp2_satellite}, \ref{fig:exp2_particle} show the marginalized posteriors for the potential parameters, satellite parameters, and parameters for a single particle.
 
-% TODO: slight ~1% bias in values due to particular sample of stars?
+% TODO: slight bias in values due to particular sample of stars?
 
@@
 For the progenitor (Sgr core), we adopt a 1~kpc distance uncertainty \citep{kunder09}, 0.2~mas/yr proper motion uncertainties \citep{pryor10}, and 10~km/s radial velocity uncertainty \citep{frinchaboy12}.
-We again neglect uncertainties in the sky coordinates.
+We  neglect uncertainties in the sky coordinates.
 
@@
 We use an ensemble of 256 walkers and also initialize the walkers by sampling the coordinate parameters from Gaussians centered on the true positions with variances set to 1\% of the parameter values.
-The potential parameters, $\Loffset$, and $\tub$ are initialized by drawing from the priors summarized in Table~\ref{tbl:params}.
+The potential parameters, $\Loffset$, and $\tub$ [TODO: truth] are initialized by drawing from the priors summarized in Table~\ref{tbl:params}.
 We again burn in the walkers for 10000 steps and run for 10000 inference steps.
@@
 \section{Discussion}\label{sec:discussion}
-% In Section~\ref{sec:intro} we identified several issues that have only been partially solved by other stream modeling methods; \rewinder\ addresses each of these.
-%\begin{enumerate}
-%	\item Observational uncertainties and missing data dimensions are naturally handled in the inferences by including the true 6D positions of stars and the progenitor as parameters in the model.
-%	\item The orbits of the stars and progenitor are directly integrated and only require a function that evaluates the acceleration due to the potential at a given position (and time). 
-%	\item Incorporating multiple streams into the inference only requires multiplying likelihoods computed for each stream and progenitor.
-%	\item Each star adds constraints on the potential parameters and does not require matching a simulated density to the observed density of stars.
-%	\item The inference is much faster than full N-body modeling because the orbits of the stars and progenitor are integrated as test particles.
-%\end{enumerate}
-%\begin{enumerate}
-%	\item The known debris structures are $\gtrsim$10 kpc from the Sun, thus while overdensities in the halo are less contaminated by background, the constituent stars often have incomplete or poorly measured phase-space information.
-Thus it is critical for any approximate model of tidal debris to incorporate observational uncertainties and missing dimensions in a consistent and justified way.
-%	\item The potential of the inner galaxy is conceivably well-described by a (set of) nearly-integrable potentials, thus the dynamics of collisionless systems is plausibly captured by a transformation to some integral-of-motion-space, e.g., action-angle coordinates.
-However there is large uncertainty in the radial profile, shape, orientation, and graininess of the outer halo and the constancy of these parameters over distance.
-Properly describing the potential may require complex models where computing actions is non-trivial or even impossible.
-%	\item Near-future photometric surveys such as \gaia\, and the \project{LSST} will likely discover many new streams and kinematic associations of stars.
-Potential recovery methods should be able to simultaneously model multiple streams and incorporate other dynamical constraints.
-%	\item Matching generated models to observed stream densities is difficult: models that rely on this must account for observational biases and properly model the underlying background of halo stars. 
-%	\item Full N-body simulations are expensive to run.
-Incorporating an N-body model into a general model and then performing a parameter search would be computationally intensive and, presently, intractable.
-%\end{enumerate}
-% are naturally handled in the inferences 
-Any method that uses tidal debris as a potential measure must also model the uncertainties on the kinematic measurements because the observational uncertainties will be significant.
-\rewinder\ handles observational uncertainties and missing data dimensions by including the true 6D positions of stars and the progenitor as parameters in the model.
+{\bf 1.
+Observational uncertainties:} Any method that uses tidal debris as a potential measure must also model the (significant) uncertainties on the kinematic measurements; \rewinder\ handles observational uncertainties and missing data dimensions by including the true 6D positions of stars and the progenitor as model parameters.
 In this article we have tested \rewinder\ using data of varied quality for a small sample of stars.
-In future work, we will investigate the effect of reducing the data quality but increasing the sample size.
 
-No assumptions were made about the the form of the Galactic potential used in \rewinder: the potential could be time-dependent, clumpy, or have properties that vary with radius.
+{\bf 2.
+Form of the potential:} No assumptions were made about the the form of the Galactic potential used in \rewinder: the potential could be time-dependent, clumpy, or have properties that vary with radius.
 The orbits of the stars and progenitor are directly integrated and only require a function that evaluates the acceleration due to the potential at a given position (and time).
 The simple experiments in this article infer potential parameters from the same functional potential used in the N-body simulations that generated the fake data.
-In future work, we will (1) try fitting incorrect potential models to see if we can still recover global properties, e.g., flattening; (2) run simulations with smoothly changing potentials \citep[e.g.,][]{buist14} and attempt to model the time dependence; and (3) explore using a generic, non-parametric potential form (e.g., a basis function expansion) as the recovery potential model.
+In future work, we will (1) try fitting incorrect potential models to see if we can still recover global properties (e.g., flattening); (2) run simulations with smoothly changing potentials \citep[e.g.,][]{buist14} and attempt to model the time dependence; and (3) explore using a generic, non-parametric potential form (e.g., a basis function expansion) as the recovery potential model.
 
-Each stream will provide constraints on different properties of the potential.
+{\bf 3.
+Multiple debris structures:} Each stream will provide constraints on different properties of the potential.
 For example, more eccentric streams may better constrain the radial profile \citep[see][who illustrate the power of using multiple streams to simultaneously constrain the potential using orbit fitting]{deg14}.
@@
 
-The inference is much faster than full N-body modeling because the stars and progenitor are treated as test particles.
+{\bf 4.
+Comparing models to data:} With \rewinder, each star adds constraints on the potential parameters and does not require matching a simulated density to the observed density of stars, thus this method works well even for small samples of well-measured stream stars.
+It might be that the most relevant stellar samples for inferring the Milky Way potential are small.
+For instance, stars that produce good distance estimates might be much more valuable than typical stars in the sample so that we may limit to only variable stars, e.g., RR Lyraes.
+These valuable stars are rare (and their abundances are age and metallicity dependent); there could be many cold structures in the Milky Way halo that are highly constraining on the potential in principle, but which contain only a few good distance-indicating members.
+It is not yet known what the trade-offs are between having many stars at low precision and a few at high precision, nor is it known how valuable distance information really is, when a structure contains many precisely observed members.
+
+{\bf 5.
+Computational expense:} One point of concern is that the number of parameters scales with eight times the number of stars: each star has six coordinate parameters (the true position), the unbinding time, and the tail assignment.
+Computational constraints limit the sample size, but even so, the inference is much faster than full N-body modeling because the stars and progenitor are treated as test particles.
 Each step in parameter space with \rewinder\ requires integrating the orbits of stars and the progenitor: we presently integrate with a fixed timestep using leapfrog integration.
@@
 
-One point of concern is that the number of parameters scales with eight times the number of stars: each star has six coordinate parameters (the true position), the unbinding time, and the tail assignment.
-Computational constraints limit the sample size, although it might be that the most relevant stellar samples for inferring the Milky Way potential are small.
-For instance, stars that produce good distance estimates might be much more valuable than typical stars in the sample so that we may limit to only variable stars, e.g., RR Lyraes.
-These valuable stars are rare (and their abundances are age and metallicity dependent); there could be many cold structures in the Milky Way halo that are highly constraining on the potential in principle, but which contain only a few good distance-indicating members.
-With \rewinder, each star adds constraints on the potential parameters and does not require matching a simulated density to the observed density of stars, thus this method works well even for small samples of well-measured stream stars.
-That said, it is not yet known what the trade-offs are between having many stars at low precision and a few at high precision, nor is it known how valuable distance information really is, when a structure contains many precisely observed members.
+%Another reason to think about small samples is more conceptual:
+%in dynamical inferences that go beyond simple virial or Jeans arguments,
+%  inference depends on the non-linear dynamical evolution of populations in phase space.
+%The nonlinearities, 
+%  combined with heteroskedastic data and requirements to marginalize over phase-space distribution functions,
+%  lead naturally to situations in which some stars are \emph{far more informative} than others.
+%For a simple example, in a toy study of the Solar System,
+%  phase-space distribution marginalization was shown to make some planets are far more informative than others
+%  about the gravitational potential \citep{bovy10}.
+%This is expected to be a generic feature in dynamical problems:
+%  it might be that a very small fraction of \gaia's catalog entries
+%  deliver a large fraction of the information about the Milky Way's potential and formation history.
 
-Another reason to think about small samples is more conceptual:
-in dynamical inferences that go beyond simple virial or Jeans arguments,
-  inference depends on the non-linear dynamical evolution of populations in phase space.
-The nonlinearities, 
-  combined with heteroskedastic data and requirements to marginalize over phase-space distribution functions,
-  lead naturally to situations in which some stars are \emph{far more informative} than others.
-For a simple example, in a toy study of the Solar System,
-  phase-space distribution marginalization was shown to make some planets are far more informative than others
-  about the gravitational potential \citep{bovy10}.
-This is expected to be a generic feature in dynamical problems:
-  it might be that a very small fraction of \gaia's catalog entries
-  deliver a large fraction of the information about the Milky Way's potential and formation history.
-
-In this article, we have only applied \rewinder\ to simulated observations of a kinematically ``hot'' stream; a large mass satellite will have a large dispersion in both position and velocity space.
+%In this article, we have only applied \rewinder\ to simulated observations of a kinematically ``hot'' stream; a large mass satellite will have a large dispersion in both position and velocity space.
 It has been shown previously that simulated cold streams are better than hot streams at recovering potential properties (cite?), however it is not clear whether this remains true with realistic observational uncertainties on the stream stars.
@@
 
-This model works well when all stars are lost due to tidal pruning.
+\rewinder works well when all stars are lost due to tidal pruning.
 Even on a mildly eccentric orbit, stars are preferentially stripped around the satellite's pericenter because of tidal shocking.

### 2014-04-14 text/paper2/ms.tex
kvj comments

@@
 It is hoped that streams or shells of stars produced by tidal disruption
-  will provide very precise measures of the gravitational potential,
+  of smaller galaxies or globular clusters will provide very precise measures of 
+  the gravitational potential, 
   to test these predictions.
@@
   using precise measurements of stars in tidal streams
-  based the idea that these stars
+  based on the idea that these stars
   were close in phase space at some point in the past.
@@
 This can lead to radially-dependent axis ratios, orientation, and smoothness, and suggests that the true mass distributions around Milky Way-like galaxies are not easily represented by simple potentials.
-Methods that seek to measure the potential of the Milky Way must be flexible enough to handle generic potentials where finding approximate analytic forms or computing actions is not possible.
+Methods that seek to measure the potential of the Milky Way must be flexible enough to handle generic potentials where finding simple analytic forms or computing actions may not be possible.
 
-The bulk of the baryonic matter in galaxies spans roughly 5-10\% of the extent of the host dark matter halo.
+The bulk of the baryonic matter in galaxies spans roughly 5-10\% of the spatial extent of the host dark matter halo.
 Hence, the brightest and most easily observable components of a galaxy are sensitive to the inner portion of the host halos mass distribution.
@@
 
-Within the Milky Way our unique vantage point allows us a three-dimensional view of stellar substructure in our own dark matter halo.
+Within the Milky Way our unique vantage point allows us a three-dimensional view of stars within in our own dark matter halo.
 Our proximity allows us to use individual stars as kinematic tracers and hence build much larger samples that probe deeper into the halo than the globular cluster and planetary nebula studies of external galaxies.
@@
 Such studies assume that the tracer orbits are randomly sampled from a smooth distribution function and are fully phase mixed.
-However, large photometric surveys such as the SDSS and 2MASS \citep{skrutskie06} have discovered large amounts of substructure --- in streams and associations of stars --- in the Milky Way halo \citep[e.g.,][]{belokurov06, rochapinto04}, thus demonstrating that the stellar distribution is not fully phase-mixed.
+However, large photometric surveys such as the SDSS and 2MASS \citep{skrutskie06} have discovered large amounts of substructure --- in streams and associations of stars --- in the Milky Way halo \citep[e.g.,][]{belokurov06, rochapinto04}, thus demonstrating that the stellar distribution is neither on random orbits nor fully phase-mixed.
 Substructure in the form of stellar streams and clouds is known to bias mass and velocity inferences by several tens of percent \citep{yencho06}.
@@
 \begin{enumerate}
-	\item \textbf{Observational uncertainties:} The known debris structures are $\gtrsim$10 kpc from the Sun, thus it is critical for any method that uses tidal debris to incorporate observational uncertainties and missing dimensions in a consistent and justified way.
+	\item \textbf{Observational uncertainties:} The known debris structures are $\gtrsim$10 kpc from the Sun, where distance and proper motion measurement errors are significant.
+Thus it is critical for any method that uses tidal debris to incorporate observational uncertainties and missing dimensions in a consistent and justified way.
 	\item \textbf{Form of the potential:} There is large uncertainty in the radial profile, shape, orientation, and graininess of the outer halo and the constancy of these parameters over distance.
-Properly describing the potential may require complex analytic functions where computing actions is non-trivial or even impossible.
+Properly describing the potential may require non-parametric techniques or complex analytic functions so the method should not rely on the existence of or ability to compute simple conserved orbital properties, such as actions.
 	\item \textbf{Multiple debris structures:} Near-future photometric surveys such as \gaia\, and the \project{LSST} will likely discover many new streams and kinematic associations of stars.
@@
 	\item \textbf{Comparing models to data:} Matching generated streams to observed stellar densities is difficult; models that rely on this must account for observational biases, internal properties of the progenitor, and background halo stars.
+[TODO: are we matching densities?]
 	\item \textbf{Computational expense:} Full N-body simulations are expensive to run; incorporating an N-body simulation into a likelihood function evaluation and then performing a parameter search would be computationally intensive and, presently, intractable.
+[TODO: KVJ?]
 \end{enumerate}
@@
 By ``probabilistic model'' we mean a justified, parametrized likelihood function with priors on the parameters.
-\rewinder\ relies on numerical orbit integration in ordinary phase-space and can thus incorporate arbitrarily complex forms for the parent potential.
+\rewinder\ relies on numerical orbit integration in ordinary phase-space and can thus incorporate arbitrarily complex forms for the parent potential including time dependence and significant substructure.
 Individual stars are constraints on the potential, thus \rewinder\ can handle very small samples of well-measured stars (e.g., 4-16).
@@
 We later use one of these simulations to test \rewinder.
-In each simulation, a $10^5$ particle NFW-profile satellite was inserted at the apogalacticon of its orbit in a static, triaxial host halo, described below.
+In each simulation, a $10^5$ particle NFW-profile satellite was inserted at the apogalacticon of its orbit in a static, multi-component galaxy with a triaxial host halo, described below.
 The satellite was evolved first in isolation, then the host potential was turned on slowly over 10 satellite internal dynamical times to reduce artificial gravitational shocking.
@@
 \rewinder\ integrates stars and the progenitor back from their present-day, observed positions to the time at which they become unbound from the satellite, where we evaluate the likelihood for each star.
-Below we (1) motivate a simple model for the debris just as it disrupts (near $\tub$ for each star) and (2) present this idea within a probabilistic framework for incorporating observational uncertainties and missing data.
+Below we (1) motivate a simple model for the debris just as it disrupts and (2) present this idea within a probabilistic framework for incorporating observational uncertainties and missing data.
 For a satellite galaxy orbiting within a static host potential, mass loss is driven by a combination of tidal stripping by the steady tidal field of the parent system and tidal shocking by the rapidly changing tidal field as the progenitor system moves through pericenter \citep[e.g.,][]{choi09}.
@@
 where $\rtide$ is given by Eq.~\ref{eq:tidalradius}, and $\sigma_v$ by Eq.~\ref{eq:velscale}.
-$\left\vert\J\right\vert$ is the absolute value of the determinant of the Jacobian that defines the transformation from heliocentric, spherical to Galactocentric, cartesian coordinates.
+$\left\vert\J\right\vert = \left\vert\frac{\partial(x,y,z,v_x,v_y,v_z)}{\partial(l,b,d,\mu_l,\mu_b,v_r)}\right\vert$ is the absolute value of the determinant of the Jacobian that defines the transformation from heliocentric, spherical to Galactocentric, cartesian coordinates.
 All quantities in this expression are evaluated at $t=\tub$.
@@
 {\bf 5.
-Computational expense:} One point of concern is that the number of parameters scales with eight times the number of stars: each star has six coordinate parameters (the true position), the unbinding time, and the tail assignment.
+Computational expense:} One point of concern for \rewinder\ is that the number of parameters scales with eight times the number of stars: each star has six coordinate parameters (the true position), the unbinding time, and the tail assignment.
 Computational constraints limit the sample size, but even so, the inference is much faster than full N-body modeling because the stars and progenitor are treated as test particles.

### 2014-04-16 text/paper2/ms.tex

@@
 The morphological distinction between stellar streams and stellar clouds is just related to the orbit of the disrupting satellite within the parent galaxy.
-With a proper model for the tidal shocking that occurs on an eccentric orbit, \rewinder\ could also be applied to structures such as the Triangulum-Andromeda (cite) or Hercules-Aquila (cite) stellar clouds, both possible shell-like structures viewed from the inside.
+With a proper model for the tidal shocking that occurs on an eccentric orbit, \rewinder\ could also be applied to structures such as the Triangulum-Andromeda \citep{rochapinto04} or Hercules-Aquila \citep{belokurov07b} stellar clouds, both possible shell-like structures viewed from the inside.
 

### 2014-04-28 streams/potential/core.py
line break

@@ -28,8 +28,7 @@ __all__ = ["Potential", "CartesianPotential", "CompositePotential", "PotentialPa
 
 class PotentialParameter(object):
 
-    def __init__(self, value=None, truth=None, range=(),
-                 latex="", units=usys):
+    def __init__(self, value=None, truth=None, range=(), latex="", units=usys):
 
         if value is None and truth is None:
             raise ValueError("If value not specified, must specify truth.")

### 2014-04-28 text/paper2/ms.tex
start describing tub marg

@@
 All quantities in this expression are evaluated at $t=\tub$.
+
+We have found that leaving the unbinding time as a free parameter makes the likelihood severely multi-modal.
+
 For a single star, the evaluation of the likelihood works as follows: given present-day coordinates for the star and progenitor, potential parameter values ($\bs{\Theta}_\Phi$), and nuisance parameter values ($\tub,\Loffset,\tailbit$),
@@
 \end{enumerate}
+
 Assuming each star is an independent tracer, the full likelihood for many stars is then just the product of the individual likelihoods:

### 2014-04-28 scripts/figures/paper2_figures.py
remove some old funcs

@@ -330,200 +330,6 @@ def Lpts():
     fig2.subplots_adjust(top=0.92, hspace=0.025, wspace=0.1)
     fig2.savefig(filename2)
 
-def Lpts_tub():
-
-    potential = LawMajewski2010()
-    filename = os.path.join(plot_path, "Lpts_r_tub.{}".format(ext))
-    filename2 = os.path.join(plot_path, "Lpts_v_tub.{}".format(ext))
-
-    fig,axes = plt.subplots(2,4,figsize=grid_figsize,
-                            sharex=True, sharey=True)
-    fig2,axes2 = plt.subplots(2,4,figsize=grid_figsize,
-                              sharex=True, sharey=True)
-
-    bins = np.linspace(-3,3,50)
-    nparticles = 2000
-    for k,_m in enumerate(range(6,9+1)):
-        mass = "2.5e{}".format(_m)
-        m = float(mass)
-        print(mass)
-
-        sgr = SgrSimulation(sgr_path.format(_m),snapfile)
-        p = sgr.particles(n=nparticles, expr=expr)
-        s = sgr.satellite()
-
-        coord, r_tide, v_disp = particles_x1x2x3(p, s,
-                                                 sgr.potential,
-                                                 sgr.t1, sgr.t2, -1.,
-                                                 at_tub=True)
-        (x1,x2,x3,vx1,vx2,vx3) = coord
-
-        axes[0,k].plot(x1/r_tide,
-                       x2/r_tide,
-                       linestyle='none', alpha=0.1, marker='.',
-                       color='#555555', zorder=-1)
-        axes[1,k].plot(x1/r_tide,
-                       x3/r_tide,
-                       linestyle='none', alpha=0.1, marker='.',
-                       color='#555555', zorder=-1)
-
-        circ = Circle((0,0), radius=1., fill=False, alpha=0.75,
-                      edgecolor='k', linestyle='solid')
-        axes[0,k].add_patch(circ)
-        circ = Circle((0,0), radius=1., fill=False, alpha=0.75,
-                     edgecolor='k', linestyle='solid')
-        axes[1,k].add_patch(circ)
-
-        axes[0,k].axhline(0., color='k', alpha=0.75)
[…]

### 2014-04-29 scripts/master_inference.py
only walkers with >2% acc frac

@@ -36,9 +36,11 @@ def master_inference(filename):
                 chain = np.hstack((chain,f["chain"].value))
             except NameError:
                 chain = f["chain"].value
-    
+
+            accfrac = f["acceptance_fraction"].value
+
     #taur = [acor.acor(chain[:,:,i])[0] for i in range(chain.shape[2])]
-    tau,mm,xx = acor.acor(np.mean(chain,axis=0).T)
+    tau,mm,xx = acor.acor(np.mean(chain[accfrac > 0.02],axis=0).T)
     acor_time = int(2*np.max(tau))
     print("Autocorrelation time: ", acor_time)
 

### 2014-04-29 scripts/master_inference.py
print path

@@ -28,7 +28,7 @@ def master_inference(filename):
 
     cache_path = os.path.join(streamspath, "plots", "infer_potential",
                               filename, "cache")
-
+    print(cache_path)
     for filename in sorted(glob.glob(os.path.join(cache_path,"inference_*.hdf5"))):
         print(filename)
         with h5py.File(filename, "r") as f:

### 2014-04-29 text/paper2/ms.tex

--- 2014-04-16 text/paper2/ms.tex
@@
 % Results:
-We find that even with a small number (8) of well-measured stream stars,
+We find that even with four well-measured stream stars,
   we can infer properties of the triaxial potential ???? with accuracies of order ??? percent ??? and
@@
 \section{Introduction}\label{sec:intro}
-
-% TODO: audit all uses of the word 'model'
 
@@
 	\item \textbf{Computational expense:} Full N-body simulations are expensive to run; incorporating an N-body simulation into a likelihood function evaluation and then performing a parameter search would be computationally intensive and, presently, intractable.
-[TODO: KVJ?]
 \end{enumerate}
@@
 
-In all simulations, the host potential is taken to be a three-component sum of a Miyamoto-Nagai disk \citep{miyamoto75}, Hernquist spheroid , and a triaxial, logarithmic halo \citep[e.g.,][]{law10}:
+In all simulations, the host potential is taken to be a three-component sum of a Miyamoto-Nagai disk \citep{miyamoto75}, Hernquist spheroid, and a triaxial, logarithmic halo \citep[e.g.,][]{law10}:
 \begin{align}
@@
 \end{align}
-Figure~\ref{fig:lpts_r} shows sections of particle orbits (for particles that are stripped) from the N-body simulations described above projected into this coordinate system.
+Figure~\ref{fig:lpts_r} shows sections of particle orbits (for particles that are stripped) from the N-body simulations described above, projected into this coordinate system.
 The positions are normalized by the instantaneous tidal radius (Eq.~\ref{eq:tidalradius}) and velocities are normalized by the instantaneous velocity scale (Eq.~\ref{eq:velscale}).
@@
 
+To model the true, 6D, present-day position of the star, $\W$, we first transform to cartesian, Galactocentric coordinates where ($\bs{r}_0,\bs{v}_0$) and ($\bs{r}_{\sat,0},\bs{v}_{\sat,0}$) are the position and velocity of the star and progenitor today.
+The star is taken as having been sampled at $t=\tub$ (the unbinding time) from an isotropic Gaussian centered on one of the effective Lagrange points in position, and an isotropic Gaussian centered on the origin in velocity.
+The present-day phase-space position is the result of integrating this sample forward from $\tub$ in the parent potential, $\Phi$, whose form is parametrized by the vector $\bs{\Theta}_\Phi$.
+We assume that once the star becomes ``unbound'' from the satellite the potential of the satellite can be ignored,\footnote{This assumption breaks down for sufficiently large-mass progenitors, somewhere between $\sim10^8-10^9~\msun$.} and thus treat the star as a test particle.
+The satellite mass enters this prescription through the tidal radius; a changing satellite mass will skew the positions of the effective Lagrange points and velocity scale.
+We allow for mass-loss by incorporating the initial satellite mass, $m_0$, and a constant mass-loss term, $\dot{m}$ into the model: $m(t) = m_0 - \dot{m}t$.
+Finally, we add two additional parameters: a constant, global factor, $\Loffset$, that scales the position of the effective Lagrange points relative to the classical tidal radius, and a binary parameter $\tailbit$, that is either $-1$ or $+1$ depending on whether the star is in the leading or trailing tail.
+To compress notation, we pack all progenitor parameters ($\W_\sat, m_0, \dot{m}, \Loffset$) into the vector $\bs{\Theta}_\sat$.
+The likelihood is then: %, and all star parameters ($\W, \tailbit, \tub$) into the vector $\bs{\Theta}$
+\begin{align}
+	p(\W \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi, \tailbit, \tub) &= p(\bs{r}_0,\bs{v}_0, \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi, \tailbit, \tub)\,\left\vert\J\right\vert\label{eq:starlike}\\
+	p(\bs{r}_0,\bs{v}_0, \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi) &= \mathcal{N}(\bs{r} \given \bs{r}_\sat - \Loffset\,\tailbit\,\rtide\,\hat{x}_1, \rtide^2)\,\mathcal{N}(\bs{v} \given \bs{v}_\sat, \sigma^2_v)
+\end{align}
+where $\rtide$ is given by Eq.~\ref{eq:tidalradius}, and $\sigma_v$ by Eq.~\ref{eq:velscale} and the positions and velocities ($\bs{r},\bs{v}, \bs{r}_{\sat},\bs{v}_{\sat}$) are evaluated at the unbinding time, $\tub$, by integrating the orbits backwards from ($\bs{r}_0,\bs{v}_0, \bs{r}_{\sat,0},\bs{v}_{\sat,0}$).
+$\left\vert\J\right\vert = \left\vert\frac{\partial(x,y,z,v_x,v_y,v_z)}{\partial(l,b,d,\mu_l,\mu_b,v_r)}\right\vert$ is the absolute value of the determinant of the Jacobian that defines the transformation from heliocentric, spherical to Galactocentric, cartesian coordinates.
+
+The likelihood above is evaluated for each star at its unbinding time, however this is not a priori known for any star.
+Furthermore, the idea of a single-valued unbinding time is not well defined as (1) it depends strongly on the particular definition of ``unbound'' and (2) for debris disrupted on near-circular orbits, the debris may pass close to the progenitor several times.
+We marginalize over the unbinding time, assuming a uniform distribution of unbinding times over the entire interaction time:
+\begin{align}
+	p(\W \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi, \tailbit) &= \int p(\W \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi, \tailbit, \tub)\,p(\tub)\,d\tub
+\end{align}
+
 We assume that the observational uncertainties are Gaussian in heliocentric, spherical coordinates, such that
@@
 
-To model the true, 6D, present-day position of the star, $\W$, we first transform to cartesian, Galactocentric coordinates where ($\bs{r}_0,\bs{v}_0$) and ($\bs{r}_{\sat,0},\bs{v}_{\sat,0}$) are the position and velocity of the star and progenitor today.
-The star is taken as having been sampled at $t=\tub$ (the unbinding time) from an isotropic Gaussian centered on one of the effective Lagrange points in position, and an isotropic Gaussian centered on the origin in velocity.
-The present-day phase-space position is the result of integrating this sample forward from $\tub$ in the parent potential, $\Phi$, whose form is parametrized by the vector $\bs{\Theta}_\Phi$.
-We assume that once the star becomes ``unbound'' from the satellite the potential of the satellite can be ignored,\footnote{This assumption breaks down for sufficiently large-mass progenitors, somewhere between $\sim10^8-10^9~\msun$.} and thus treat the star as a test particle.
-The satellite mass enters this prescription through the tidal radius; a changing satellite mass will skew the positions of the effective Lagrange points and velocity scale.
-We allow for mass-loss by incorporating the initial satellite mass, $m_0$, and a constant mass-loss term, $\dot{m}$ into the model: $m(t) = m_0 - \dot{m}t$.
-Finally, we add two additional parameters: a constant, global factor, $\Loffset$, that scales the position of the effective Lagrange points relative to the classical tidal radius, and a binary parameter $\tailbit$, that is either $-1$ or $+1$ depending on whether the star is in the leading or trailing tail.
-To compress notation, we pack all progenitor parameters ($\W_\sat, m_0, \dot{m}, \Loffset$) into the vector $\bs{\Theta}_\sat$, and all star parameters ($\W, \tailbit, \tub$) into the vector $\bs{\Theta}$.
-The likelihood is then:
-\begin{align}
-	p(\bs{\Theta} \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi) &= p(\bs{r}_0,\bs{v}_0, \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi)\,\left\vert\J\right\vert_{t=\tub}\label{eq:starlike}\\
-	p(\bs{r}_0,\bs{v}_0, \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi) &= \mathcal{N}(\bs{r} \given \bs{r}_\sat - \Loffset\,\tailbit\,\rtide\,\hat{x}_1, \rtide^2)\,\mathcal{N}(\bs{v} \given \bs{v}_\sat, \sigma^2_v)_{t=\tub}
-\end{align}
-where $\rtide$ is given by Eq.~\ref{eq:tidalradius}, and $\sigma_v$ by Eq.~\ref{eq:velscale}.
-$\left\vert\J\right\vert = \left\vert\frac{\partial(x,y,z,v_x,v_y,v_z)}{\partial(l,b,d,\mu_l,\mu_b,v_r)}\right\vert$ is the absolute value of the determinant of the Jacobian that defines the transformation from heliocentric, spherical to Galactocentric, cartesian coordinates.
-All quantities in this expression are evaluated at $t=\tub$.
 For a single star, the evaluation of the likelihood works as follows: given present-day coordinates for the star and progenitor, potential parameter values ($\bs{\Theta}_\Phi$), and nuisance parameter values ($\tub,\Loffset,\tailbit$),
@@
 	\item integrate star and progenitor orbits backwards as test particles in the potential given by $\bs{\Theta}_\Phi$ for the total interaction time, $t_{\rm int}\approx6$~Gyr;
-	\item at the timestep closest to $\tub$, transform the particle position to the relative, normalized coordinates defined in Section~\ref{sec:debris};
-	\item compute the likelihood given by Equation~\ref{eq:starlike}.
+	\item transform the particle position to the relative, normalized coordinates defined in Section~\ref{sec:debris};
+	\item compute the likelihood given by Equation~\ref{eq:starlike} at each timestep of the integration and marginalize over the unbinding time, $\tub$.
 \end{enumerate}
+
 Assuming each star is an independent tracer, the full likelihood for many stars is then just the product of the individual likelihoods:
@@
 		p(\D_\sat \given \W_\sat) \, \prod_i \, p(\D^{(i)} \given \W^{(i)}) \,
-			p(\bs{\Theta}^{(i)} \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi)
+			p(\W^{(i)} \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi, \tailbit^{(i)})
 \end{multline}
-All parameters are summarized in Table~\ref{tbl:params}.
+where $\bs{\Theta}^{(i)} = (\W^{(i)}, \tailbit^{(i)})$. All parameters are summarized in Table~\ref{tbl:params}.
 
@@
 
-We first test \rewinder\ using only eight stars (particles) --- four from the leading tail, and four from the trailing tail --- randomly sampled from the $2.5\times10^8\,\msun$ simulation (Section~\ref{sec:sims}), assuming the observed 6D positions for both the stars and the progenitor have negligible uncertainties.
+We first test \rewinder\ using only four stars (particles) --- two from the leading tail, and two from the trailing tail --- randomly sampled from the $2.5\times10^8\,\msun$ simulation (Section~\ref{sec:sims}), assuming the observed 6D positions for both the stars and the progenitor have negligible uncertainties.
 The stars were required to have been stripped after the first pericentric passage and have a present-day distance within $50$ kpc of the Sun.
@@
 \subsection{Data with near-future uncertainties}\label{sec:exp2}
-We next take the same eight stars used in the previous experiment and ``observe'' them with optimistic observational uncertainties.
+We next take the same four stars used in the previous experiment and ``observe'' them with optimistic observational uncertainties.
 We require these stars to be RR Lyrae variables which are known to be excellent distance indicators via the mid-infrared period-luminosity relation \citep[as shown in, e.g.,][]{madore12}.
@@
 
-For this experiment, we (1) assume the stars are bright distance indicators --- RR Lyrae --- such that the fractional distance uncertainty is $2\%$; (2) neglect the uncertainty in angular position (for a typical RR Lyrae at 50~kpc this is $\sim$$10^{-7}$~deg for \gaia); (3) assume we can measure radial velocities to these stars with 5~km/s uncertainty; and (4) compute the sky-averaged \gaia\, proper-motion uncertainty for each star assuming an F0V spectral type using the \texttt{PyGaia}\footnote{https://github.com/agabrown/PyGaia} code and use this uncertainty for both components of proper motion.
+For this experiment, we (1) assume the stars are RR Lyrae stars (e.g., bright distance indicators) such that the fractional distance uncertainty is $2\%$; (2) neglect the uncertainty in angular position, $l,b$ (for a typical RR Lyrae at 50~kpc this is $\sim$$10^{-7}$~deg for \gaia); (3) assume we can measure radial velocities to these stars with 5~km/s uncertainty; and (4) compute the sky-averaged \gaia\, proper-motion uncertainty for each star assuming an F0V spectral type using the \texttt{PyGaia}\footnote{https://github.com/agabrown/PyGaia} code and use this uncertainty for both components of proper motion.
 We further assume that we know the tail assignment for each star, $\beta$.
@@
 
-This experiment samples over the four potential parameters, the Lagrange point offset, four phase-space coordinate parameters plus the unbinding time for each star (40 total), and four phase-space coordinate parameters for the progenitor --- 49 parameters in total.
+This experiment samples over the four potential parameters, the Lagrange point offset, four phase-space coordinate parameters for each star (16 total), and four phase-space coordinate parameters for the progenitor --- 25 parameters in total.
 We use an ensemble of 256 walkers and draw initial conditions for the coordinate parameters by sampling from Gaussian's centered on the ``observed'' value with variances specified by the observational uncertainties.
-Other parameters --- potential parameters, $\Loffset$, and $\tub$ --- are initialized by drawing from the priors summarized in Table~\ref{tbl:params}.
-[TODO: start from truth...] We burn in the walkers for 10000 steps and run for 10000 inference steps.
+Other parameters --- potential parameters and $\Loffset$ --- are initialized by drawing from the priors summarized in Table~\ref{tbl:params}.
+We burn in the walkers for 50000 steps and run for 100000 inference steps, but thin the chains by taking every $2\times \max(t_{\rm acor})$ sample, where $\max(t_{\rm acor}) \approx 4000$~steps.
 Figures~\ref{fig:exp2_potential}, \ref{fig:exp2_satellite}, \ref{fig:exp2_particle} show the marginalized posteriors for the potential parameters, satellite parameters, and parameters for a single particle.
-
-% TODO: slight bias in values due to particular sample of stars?
 
@@
 The SMASH survey \citep{smashprop} will be completed within a year (early 2015), long before the final \gaia\, data release.
-Thus, we will soon have precise distance measurements for stars in the Sgr and Orphan streams, but poor proper motion constraints.
-SMASH also targets several RR Lyrae in the Sgr core, which enable a high-precision distance measurement of the progenitor.
-Though proper motions have been measured for the Sgr progenitor, we now consider a case in which we have high-precision (2\%) distances to stream stars and the progenitor, 10~km/s radial velocity uncertainty, and missing proper motions for all stars and progenitor.
+Thus, we will soon have precise distance measurements for stars in the Sgr and Orphan streams, but poor or no proper motion constraints.
+SMASH also targets several RR Lyrae in the Sgr core, which will enable a high-precision distance measurement of the progenitor.
+Though proper motions have been measured for the Sgr progenitor, we now consider a case in which we have high-precision (2\%) distances to stream stars and the progenitor, 10~km/s radial velocity uncertainty, but missing proper motions for all stars and progenitor.
 
-As with the previous experiment, this experiment includes 49 model parameters in total.
-We again use an ensemble of 256 walkers, but this time initialize the walkers by sampling the coordinate parameters from Gaussians centered on the true positions with variances set to 1\% of the parameter values.
-Other parameters --- potential parameters, $\Loffset$, and $\tub$ --- are initialized by drawing from the priors summarized in Table~\ref{tbl:params}.
-We again burn in the walkers for 10000 steps and run for 10000 inference steps.
+As with the previous experiment, this experiment includes 25 model parameters in total.
+We again use an ensemble of 256 walkers.
+Other parameters --- potential parameters and $\Loffset$ --- are initialized by drawing from the priors summarized in Table~\ref{tbl:params}.
+We burn in the walkers for 50000 steps and run for 500000 inference steps.
+We again thin the chains by taking every $2\times \max(t_{\rm acor})$ sample, but here the autocorrelation time is found to be very long, $\max(t_{\rm acor}) \approx 25000$~steps.
 Figures~\ref{fig:exp3_potential}, \ref{fig:exp3_satellite}, \ref{fig:exp3_particle} show the marginalized posteriors for the potential parameters, satellite parameters, and parameters for a single particle.
-Accounting for observational uncertainties does increase the uncertainties in the constraints on the potential parameters (as expected), but this demonstrates that a small sample of well-measured stream stars provide an enormous amount of information about the triaxiality of the potential.
+Even a small sample of stars with precise distance measurements provide an enormous amount of information about the triaxiality of the potential.
 
@@
 
-%TODO: remove exp 3 and explain that it is possible but computationally difficult...
-
-\subsection{Current data}\label{sec:exp4}
-% Most of the streams and debris structures discovered thus far have undetected or unconfirmed progenitors.
+%\subsection{Current data}\label{sec:exp4}
+%% Most of the streams and debris structures discovered thus far have undetected or unconfirmed progenitors.
 We expect the phase-space positions of the stars to provide constraints on the position of the progenitor even if it is unobserved.
 For this experiment, we consider a situation where we have precise (2\%) distance measurements, 10~km/s radial velocity uncertainties, and no proper motion measurements to eight stream stars (the same used in previous experiments), and no knowledge of the 6D position of the progenitor object.
-Without measured proper motions, the extremely precise distance measurements combined with moderate precision radial velocities are sufficient for constraining the properties of the debris and the potential.
+%Without measured proper motions, the extremely precise distance measurements combined with moderate precision radial velocities are sufficient for constraining the properties of the debris and the potential.
 How important are these distance measurements?
@@
 We  neglect uncertainties in the sky coordinates.
-
-Again, this experiment includes 49 model parameters in total.
+%
+%Again, this experiment includes 49 model parameters in total.
 We use an ensemble of 256 walkers and also initialize the walkers by sampling the coordinate parameters from Gaussians centered on the true positions with variances set to 1\% of the parameter values.
@@
 %[TODO: ..statement about results]
-
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[width=\textwidth]{../../plots/paper2/exp4_potential.png}
-\caption{Same as Figure~\ref{fig:exp2_potential} but with realistic, present-day uncertainties. }\label{fig:exp4_potential}
-\end{center}
-\end{figure}
-
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[width=\textwidth]{../../plots/paper2/exp4_satellite.png}
-\caption{ Same as Figure~\ref{fig:exp2_satellite} but with realistic, present-day uncertainties. }\label{fig:exp4_satellite}
-\end{center}
-\end{figure}
-
-\begin{figure}[!ht]
-\begin{center}
-\includegraphics[width=\textwidth]{../../plots/paper2/exp4_particle.png}
-\caption{ Same as Figure~\ref{fig:exp2_particle} but with realistic, present-day uncertainties. }\label{fig:exp4_particle}
-\end{center}
-\end{figure}
+%
+%\begin{figure}[!ht]
+%\begin{center}
+%\includegraphics[width=\textwidth]{../../plots/paper2/exp4_potential.png}
+%\caption{Same as Figure~\ref{fig:exp2_potential} but with realistic, present-day uncertainties. }\label{fig:exp4_potential}
+%\end{center}
+%\end{figure}
+%
+%\begin{figure}[!ht]
+%\begin{center}
+%\includegraphics[width=\textwidth]{../../plots/paper2/exp4_satellite.png}
+%\caption{ Same as Figure~\ref{fig:exp2_satellite} but with realistic, present-day uncertainties. }\label{fig:exp4_satellite}
+%\end{center}
+%\end{figure}
+%
+%\begin{figure}[!ht]
+%\begin{center}
+%\includegraphics[width=\textwidth]{../../plots/paper2/exp4_particle.png}
+%\caption{ Same as Figure~\ref{fig:exp2_particle} but with realistic, present-day uncertainties. }\label{fig:exp4_particle}
+%\end{center}
+%\end{figure}
 
 \section{Discussion}\label{sec:discussion}
+
 {\bf 1.
@@
 
-\rewinder works well when all stars are lost due to tidal pruning.
+\rewinder\, works well when all stars are lost due to tidal pruning.
 Even on a mildly eccentric orbit, stars are preferentially stripped around the satellite's pericenter because of tidal shocking.
@@
 
+% TODO: We have tested leaving the unbinding time as a free parameter but have found that this makes the likelihood severely multi-modal and thus very difficult to sample
+
 \section{Conclusions}\label{sec:conclusion}
@@
 This research made use of Astropy, a community-developed core Python package for Astronomy \citep{astropy13}.
-This work additionally relied on Columbia University's \emph{hotfoot} and \emph{yeti} compute clusters, and we acknowledge the Columbia HPC support staff for assistance. \\
+This work additionally relied on Columbia University's \emph{Hotfoot} and \emph{Yeti} compute clusters, and we acknowledge the Columbia HPC support staff for assistance. \\
 

### 2014-04-30 text/paper2/ms.tex
final draft???

@@
 % Results:
-We find that even with four well-measured stream stars,
-  we can infer properties of the triaxial potential ???? with accuracies of order ??? percent ??? and
-  missing phase-space coordinates of the stars ??? with accuracies of order ????.
+We find that with just four well-measured stream stars,
+  we can infer properties of a triaxial potential with precisions of order 5-7 percent.
+Without proper motions we find that these same stars provide 15 percent constraints on potential parameters
+  and precisions around 25 percent for recovering missing phase-space coordinates.
 The method generalizes to encompass mixtures of independent debris structures,
   or situations in which star membership in structures is unknown.
-% [Get more quantitative here?]
 % Conclusions: 
@@
 \begin{center}
-\includegraphics[width=\textwidth]{../../plots/paper2/potentials.pdf}
+\includegraphics[width=\textwidth]{../../plots/paper2/potentials.png}
 \caption{ Equipotential contours for the LM10 potential (Eq.~\ref{eq:lm10}) in Galactocentric, cartesian coordinates for various halo parameter choices.
-For all panels, $\vhalo=121.858~\mathrm{km}/\mathrm{s}$, $r_h=12~\mathrm{kpc}$.
+For all panels, $\vhalo=121.858~\mathrm{km}/\mathrm{s}$, $r_h=12~\mathrm{kpc}$, and $q_2=1$.
 Left to right, each column represents a new choice of parameters.
-If not specified, other parameters are fixed to $q_1=q_2=q_z=1$ and $\phi=0^\circ$ (far left panels).
+If not specified, other parameters are fixed to $q_1=q_z=1$ and $\phi=0^\circ$ (far left panels).
+Panels on far right show the best-fit parameter values from LM10.
 }\label{fig:potential}
@@
 The simple experiments in this article infer potential parameters from the same functional potential used in the N-body simulations that generated the fake data.
+We have shown that using this simple model, \emph{precise} measurements of the Milky Way potential are attainable with near-future data, however we have not discussed the \emph{accuracy} of such measurements.
+Bonaca et al.
+(in prep.) show that fitting a rigid potential model to the live potential of the \project{Via Lactea} halo ...TODO.
 In future work, we will (1) try fitting incorrect potential models to see if we can still recover global properties (e.g., flattening); (2) run simulations with smoothly changing potentials \citep[e.g.,][]{buist14} and attempt to model the time dependence; and (3) explore using a generic, non-parametric potential form (e.g., a basis function expansion) as the recovery potential model.
@@
 For instance, the distribution of positions of the trailing debris is more extended and possibly offset from zero in the $\hat{\bs{x}}_2$ direction.
-We have shown that using this simple model, \emph{precise} measurements of the Milky Way potential are attainable with near-future data, however we have not discussed the \emph{accuracy} of such measurements.
-Bonaca et al.
-(in prep.) show that fitting a rigid potential model to the potential of the \project{Via Lactea} halo ...
-%\cite{bonaca14}
 
@@
 We have performed several experiments to show that \rewinder\ simultaneously constrains the potential and models a tidal stream given simulated data with a range of realistic uncertainties.
+We find that with realistic but future high-quality data --- that is, high precision distance measurements from RR Lyrae and proper motions from \gaia --- a small sample of just four stars in a tidal stream modeled with \rewinder\, provide measurements of potential parameters that rival present-day constraints from running full N-body simulations.
+For this high-quality data, we recover the input potential parameter values with uncertainties of order 5-7 percent.
+Without proper motion data the uncertainties are around 15 percent.
+TODO: concluding sentence.
 
@@
 [TODO: HOGG].
+[TODO: thank Marla, Ana?]
 APW is supported by a National Science Foundation Graduate Research Fellowship under Grant No.\ 11-44155.

### 2014-04-30 scripts/figures/paper2_figures.py
tweaks

@@ -79,7 +79,7 @@ def simulated_streams():
 
             data_filename = os.path.join(streamspath, "data", "observed_particles",
                                          "2.5e{}.hdf5".format(_m))
-            cfg_filename = os.path.join(streamspath, "config", "exp1_{}.yml".format(_m))
+            cfg_filename = os.path.join(streamspath, "config", "exp2.yml".format(_m))
             data = read_hdf5(data_filename)
             true_particles = data["true_particles"].to_frame(galactocentric)
             config = read_config(cfg_filename)
@@ -133,26 +133,28 @@ def potentials():
     potentials.append(LawMajewski2010(**base_params))
 
     pp = base_params.copy()
-    pp['q1'] = 1.5
+    pp['qz'] = 1.5
     potentials.append(LawMajewski2010(**pp))
-    axes[0,1].text(0.5, 1.05, r"$q_1=1.5$",
+    axes[0,1].text(0.5, 1.05, r"$q_z=1.5$",
                    horizontalalignment='center',
                    fontsize=20,
                    transform=axes[0,1].transAxes)
 
     pp = base_params.copy()
-    pp['qz'] = 1.5
+    pp['phi'] = (45*u.degree).to(u.radian).value
+    pp['q1'] = 1.5
     potentials.append(LawMajewski2010(**pp))
-    axes[0,2].text(0.5, 1.05, r"$q_z=1.5$",
+    axes[0,2].text(0.5, 1.05, r"$q_1=1.5$, $\phi=45^\circ$",
                    horizontalalignment='center',
                    fontsize=20,
                    transform=axes[0,2].transAxes)
 
     pp = base_params.copy()
-    pp['phi'] = 45*u.degree
-    pp['q1'] = 1.5
+    pp['q1'] = 1.38
+    pp['qz'] = 1.36
+    pp['phi'] = (97*u.degree).to(u.radian).value
     potentials.append(LawMajewski2010(**pp))
-    axes[0,3].text(0.5, 1.05, r"$q_1=1.5$, $\phi=45^\circ$",
+    axes[0,3].text(0.5, 1.05, r"$q_1=1.38$, $q_z=1.36$, $\phi=97^\circ$",
                    horizontalalignment='center',
                    fontsize=20,
                    transform=axes[0,3].transAxes)
@@ -165,13 +167,15 @@ def potentials():
[…]

### 2014-04-30 text/paper2/ms.tex
notes to self

@@
 This can lead to radially-dependent axis ratios, orientation, and smoothness, and suggests that the true mass distributions around Milky Way-like galaxies are not easily represented by simple potentials.
-Methods that seek to measure the potential of the Milky Way must be flexible enough to handle generic potentials where finding simple analytic forms or computing actions may not be possible.
+Methods that seek to measure the gravitational potentials around such galaxies must be flexible enough to handle generic potential forms where finding simple analytic approximations or computing actions may not be possible.
 
@@
 Such studies assume that the tracer orbits are randomly sampled from a smooth distribution function and are fully phase mixed.
-However, large photometric surveys such as the SDSS and 2MASS \citep{skrutskie06} have discovered large amounts of substructure --- in streams and associations of stars --- in the Milky Way halo \citep[e.g.,][]{belokurov06, rochapinto04}, thus demonstrating that the stellar distribution is neither on random orbits nor fully phase-mixed.
-Substructure in the form of stellar streams and clouds is known to bias mass and velocity inferences by several tens of percent \citep{yencho06}.
+However, large photometric surveys such as the SDSS and 2MASS \citep{skrutskie06} have discovered large amounts of substructure --- in streams and kinematic associations of stars --- in the Milky Way halo \citep[e.g.,][]{belokurov06, rochapinto04}, thus demonstrating that the stellar distribution is neither on random orbits nor fully phase-mixed.
+Substructure in the form of stellar streams and clouds is known to bias mass and velocity inferences from random tracer methods by several tens of percent \citep{yencho06}.
 
@@
 Tidal streams are dynamically cold systems --- debris typically have small distributions of energy and angular momentum --- and thus require orders of magnitude fewer tracers than a random sample to get constraints of comparable accuracy to Jeans analysis.
-For example, in the simplest case we might assume that debris stars are actually still on the same orbit as their progenitor system \citep[a \emph{wrong} assumption, see e.g.,][]{eyre11}.
+For example, in the simplest case we might assume that debris stars are actually still on the same orbit as their progenitor system (a \emph{wrong} assumption, see below).
 This information about the orbits combined with measurements of the full-space velocities ${\bf v}$ at different points ${\bf x}$ in the structure (e.g., along a stream) would give us a direct measure of differences in a potential, $\Phi$.
@@
 Nevertheless, they were able to constrain the 3D shape of the Milky Way potential out to $\sim$70 kpc and found that the best-fitting halo has a prolate, triaxial shape elongated in the Galactic Z direction, orthogonal to the plane of the disk.
-Though an unlikely orientation for the halo --- \cite{debattista13} find that the disk of the Milky Way would not remain stable in such a configuration --- LM10 showed that the data is at a state where such inference can be attempted.
+Though an unlikely orientation for the halo --- \cite{debattista13} find that the disk of the Milky Way would not remain stable in such a configuration --- LM10 showed that the data is at a state where such inference is possible.
 
@@
 \begin{enumerate}
-	\item \textbf{Observational uncertainties:} The known debris structures are $\gtrsim$10 kpc from the Sun, where distance and proper motion measurement errors are significant.
+	\item \textbf{Observational uncertainties:} The known debris structures are $\gtrsim$10 kpc from the Sun where distance and proper motion measurement errors are significant.
 Thus it is critical for any method that uses tidal debris to incorporate observational uncertainties and missing dimensions in a consistent and justified way.
@@
 By ``probabilistic model'' we mean a justified, parametrized likelihood function with priors on the parameters.
-\rewinder\ relies on numerical orbit integration in ordinary phase-space and can thus incorporate arbitrarily complex forms for the parent potential including time dependence and significant substructure.
+\rewinder\ relies on numerical orbit integration in ordinary phase-space and can thus incorporate arbitrarily complex forms for the parent potential (such as time dependence and significant substructure).
 Individual stars are constraints on the potential, thus \rewinder\ can handle very small samples of well-measured stars (e.g., 4-16).
@@
 We conclude in Section~\ref{sec:conclusion}.
+
+\begin{figure*}[!ht]
+\begin{center}
+\includegraphics[width=\textwidth]{../../plots/paper2/potentials.png}
+\caption{ Equipotential contours for the LM10 potential (Eq.~\ref{eq:lm10}) in Galactocentric, cartesian coordinates for various halo parameter choices.
+For all panels, $\vhalo=121.858~\mathrm{km}/\mathrm{s}$, $r_h=12~\mathrm{kpc}$, and $q_2=1$.
+Left to right, each column represents a new choice of parameters.
+If not specified, other parameters are fixed to $q_1=q_z=1$ and $\phi=0^\circ$ (far left panels).
+Panels on far right show the best-fit parameter values from LM10.
+}\label{fig:potential}
+\end{center}
+\end{figure*}
 
@@
 \end{equation}
+We use this potential not because we think it is a realistic representation of the Galactic potential, but because successful inference with this potential demonstrates that it is possible to recover information about non-trivial potential forms.
 Figure~\ref{fig:potential} shows equipotential slices in the Galactic X-Z plane (Y=0) and Y-Z plane (X=0) for a few choices of  $q_1$,  $q_z$, and $\phi$ while holding all other parameters fixed, as described in the figure caption.
@@
 
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=\textwidth]{../../plots/paper2/potentials.png}
-\caption{ Equipotential contours for the LM10 potential (Eq.~\ref{eq:lm10}) in Galactocentric, cartesian coordinates for various halo parameter choices.
-For all panels, $\vhalo=121.858~\mathrm{km}/\mathrm{s}$, $r_h=12~\mathrm{kpc}$, and $q_2=1$.
-Left to right, each column represents a new choice of parameters.
-If not specified, other parameters are fixed to $q_1=q_z=1$ and $\phi=0^\circ$ (far left panels).
-Panels on far right show the best-fit parameter values from LM10.
-}\label{fig:potential}
-\end{center}
-\end{figure}
-
-\begin{figure}[h]
+\begin{figure*}[!ht]
 \begin{center}
@@
 \end{center}
-\end{figure}
+\end{figure*}
 
@@
 On a mildly eccentric orbit, the tidal shocking is subdominant to the steady disruption.
-The interplay between these two processes as a function of orbital properties is outside our scope; we focus here on the slow removal of stars by steady tidal forcing.
+The interplay between these two processes as a function of orbital properties is outside the scope of this paper and will be explored in future work; we focus here on the slow removal of stars by steady tidal forcing.
 
@@
 
-In a triaxial potential, the orbital plane of the satellite is not fixed but we still expect there to be \emph{effective} Lagrange points for a spherical satellite system along the line of centers connecting the origin of the parent potential to the satellite --- that is, we expect the stars to be stripped at some characteristic radius near the tidal radius with some dispersion about this point.
+In a triaxial potential, the orbital plane of the satellite is not fixed but we still expect there to be \emph{effective} Lagrange points for a spherical satellite system along the line of centers connecting the origin of the parent potential to the satellite --- that is, we expect the stars to be stripped at some characteristic radius near the tidal radius with some dispersion about these points.
 We proceed by defining a coordinate system relative to the position and velocity of the progenitor, ($\bs{r}_p$,$\bs{v}_p$), rotated into the instantaneous orbital plane defined by $\bs{L} = \bs{r}_p \times \bs{v}_p$.
@@
 	\hat{\bs{x}}_1 &= \frac{\bs{r}_p}{\|\bs{r}_p\|}\label{eq:x1}\\
-	\hat{\bs{x}}_2 &= \frac{\bs{L} \times \hat{\bs{x}}_1}{\|\bs{L} \times \hat{\bs{x}}_1\|} = \hat{\bs{x}}_3 \times \hat{\bs{x}}_1\\
+	\hat{\bs{x}}_2 &= \frac{\bs{L} \times \hat{\bs{x}}_1}{\|\bs{L} \times \hat{\bs{x}}_1\|}\\
 	\hat{\bs{x}}_3 &= \hat{\bs{L}} = \frac{\bs{L}}{\|\bs{L}\|} = \frac{\bs{r}_p \times \bs{v}_p}{\|\bs{r}_p \times \bs{v}_p\|}\label{eq:x3}.
@@
 Figure~\ref{fig:lpts_v} shows the velocity orbits of each star also projected into these coordinates and normalized.
-The dispersion in velocity is well-normalized by the velocity scale of Eq.~\ref{eq:velscale}, though it is clear that the velocity dispersion along $\hat{x}_1$, the radial vector to the satellite position, is larger than that in other dimensions, possibly because of tidal shocking.
+The dispersion in velocity is well-normalized by the velocity scale of Eq.~\ref{eq:velscale}, though it is clear that the velocity dispersion along $\hat{x}_1$, the radial vector to the satellite position, is larger than that in other dimensions, possibly because of mild tidal shocking.
 
@@
 
-\begin{figure}[h]
+\begin{figure*}[!ht]
 \begin{center}
@@
 \end{center}
-\end{figure}
+\end{figure*}
 
-\begin{figure}[h]
+\begin{figure*}[!ht]
 \begin{center}
@@
 \end{center}
-\end{figure}
+\end{figure*}
+
+\clearpage
 
@@
 \subsection{Probabilistic model}
-Suppose we observe the 6D position of a star, $\D$, in heliocentric coordinates --- e.g., the measured position on the sky, ($l$, $b$); distance, $d$; proper motions, ($\mu_l$, $\mu_b$); and line-of-sight velocity, $v_r$ --- and have determined through some other means that this star was once part of a progenitor system (e.g., satellite galaxy) with mass $m(t)$ that is disrupting and forming a cold debris structure in the potential, $\Phi$, of the parent galaxy (e.g., the Milky Way).
+Suppose we observe the 6D position of a star, $\D = (l,b,d,\mu_l,\mu_b,v_r)$, in heliocentric coordinates --- e.g., the measured position on the sky, ($l$, $b$); distance, $d$; proper motions, ($\mu_l$, $\mu_b$); and line-of-sight velocity, $v_r$ --- and have determined through some other means that this star was once part of a progenitor system (e.g., satellite galaxy) with mass $m(t)$ that is disrupting and forming a cold debris structure in the potential, $\Phi$, of the parent galaxy (e.g., the Milky Way).
 We assume that the mass of the parent system enclosed within the pericenter of the orbit of the satellite, $M_{\rm enc}(R_{\rm peri})$, is much larger than the initial mass of the satellite, $M_{\rm enc}(R_{\rm peri})\gg m(t=0)$.
@@
 	p(\W \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi, \tailbit, \tub) &= p(\bs{r}_0,\bs{v}_0, \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi, \tailbit, \tub)\,\left\vert\J\right\vert\label{eq:starlike}\\
-	p(\bs{r}_0,\bs{v}_0, \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi) &= \mathcal{N}(\bs{r} \given \bs{r}_\sat - \Loffset\,\tailbit\,\rtide\,\hat{x}_1, \rtide^2)\,\mathcal{N}(\bs{v} \given \bs{v}_\sat, \sigma^2_v)
+	p(\bs{r}_0,\bs{v}_0, \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi,\tailbit,\tub) &= \left[\mathcal{N}(\bs{r} \given \bs{\Theta}_\sat, \tailbit)\,\mathcal{N}(\bs{v} \given \bs{\Theta}_\sat)\right]_{\tub}
 \end{align}
@@
 Furthermore, the idea of a single-valued unbinding time is not well defined as (1) it depends strongly on the particular definition of ``unbound'' and (2) for debris disrupted on near-circular orbits, the debris may pass close to the progenitor several times.
-We marginalize over the unbinding time, assuming a uniform distribution of unbinding times over the entire interaction time:
+We marginalize the likelihood over all possible unbinding times, assuming a uniform distribution of unbinding times over the entire interaction time:
 \begin{align}
-	p(\W \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi, \tailbit) &= \int p(\W \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi, \tailbit, \tub)\,p(\tub)\,d\tub
+	p(\W \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi, \tailbit) &= \int^{t_{\rm int}}_0 p(\W \given \bs{\Theta}_\sat, \bs{\Theta}_\Phi, \tailbit, \tub)\,p(\tub)\,d\tub
 \end{align}
+where the interaction time is taken to be $t_{\rm int}=6.2$~Gyr.
 
@@
 
-For a single star, the evaluation of the likelihood works as follows: given present-day coordinates for the star and progenitor, potential parameter values ($\bs{\Theta}_\Phi$), and nuisance parameter values ($\tub,\Loffset,\tailbit$),
+For a single star, the evaluation of the likelihood works as follows: given observed, present-day coordinates for the star and progenitor ($\D,\D_\sat$), potential parameter values ($\bs{\Theta}_\Phi$), and nuisance parameter values ($\Loffset,\tailbit$),
 \begin{enumerate}
@@
 In each experiment we run the walkers for a large number of steps from starting positions described below, then throw away these samples and take the positions from the final step of this ``burn-in'' phase as a starting point for the samples used for inference.
-We compute the autocorrelation time for each sampled parameter (using \texttt{ACOR}\footnote{www.math.nyu.edu/faculty/goodman/software/acor/}$^{,}$\footnote{https://github.com/dfm/acor}) and thin the chains by taking every $\mathrm{median}(t_{\rm acor})$ sample to ensure the samples are close to independent.
+We compute the autocorrelation time for each sampled parameter (using \texttt{ACOR}\footnote{\url{http://www.math.nyu.edu/faculty/goodman/software/acor/}}$^{,}$\footnote{\url{https://github.com/dfm/acor}}) and thin the chains by taking every $\mathrm{max}(t_{\rm acor})$ sample to ensure the samples are close to independent.
 
@@
 
-We first test \rewinder\ using only four stars (particles) --- two from the leading tail, and two from the trailing tail --- randomly sampled from the $2.5\times10^8\,\msun$ simulation (Section~\ref{sec:sims}), assuming the observed 6D positions for both the stars and the progenitor have negligible uncertainties.
+We test \rewinder\ using only four stars (particles) --- two from the leading tail, and two from the trailing tail --- randomly sampled from the $2.5\times10^8\,\msun$ simulation (Section~\ref{sec:sims}), assuming the observed 6D positions for both the stars and the progenitor have negligible uncertainties.
 The stars were required to have been stripped after the first pericentric passage and have a present-day distance within $50$ kpc of the Sun.
-Figure~\ref{fig:sims} (second column from right) shows the eight randomly chosen stars (black crosses) in Galactic coordinates, over-plotted on all other simulated particles (grey points).
+Figure~\ref{fig:sims} (second column from right) shows the randomly chosen stars (black crosses) in Galactic coordinates, over-plotted on all other simulated particles (grey points).
 
 We leave the potential parameters ($q_1,q_z,\phi,\vhalo$) free to vary, along with the Lagrange point offset, $\Loffset$, and initialize an ensemble of 64 MCMC walkers by sampling from the priors summarized in Table~\ref{tbl:params}.
-We run the walkers for 10000 steps to burn-in, then restart the sampler starting from the final position of the burn-in phase and run for another 5000 inference steps.
+We run the walkers for 5000 steps to burn-in, then restart the sampler starting from the final position of the burn-in phase and run for another 5000 inference steps.
 Figure~\ref{fig:trace} shows the walker positions over the 5000 inference steps for each of the five parameters.
@@
 
-\begin{figure}[!ht]
+\begin{figure}[!h]
 \begin{center}
-\includegraphics[width=0.9\textwidth]{../../plots/paper2/mcmc_trace.png}
+\includegraphics[width=0.5\textwidth]{../../plots/paper2/mcmc_trace.png}
 \caption{ Positions of all 64 MCMC walkers (see Section~\ref{sec:experiments}) for each parameter at each step of the inference (black lines).
@@
 
-\begin{figure}[!ht]
+\begin{figure*}[!h]
 \begin{center}
@@
 Nearly 100 RR Lyrae associated with the Sgr stream and $\sim$30 associated with the Orphan stream will be observed with \spitzer as part of the SMASH survey \citep{smashprop} with expected fractional distance uncertainties around $\sim$2\%.
-These stars will also be included in the \gaia\, proper motion catalog: at a distance of $\sim$50~kpc, a typical RR Lyrae will have a tangential velocity error around $\sim$20~km/s.
+These stars will also be included in the \gaia\, proper motion catalog: at a distance of $\sim$50~kpc, a typical RR Lyrae will have a tangential velocity error around $\sim$20~km/s, though our sample of test stars are all within 40~kpc.
 
-For this experiment, we (1) assume the stars are RR Lyrae stars (e.g., bright distance indicators) such that the fractional distance uncertainty is $2\%$; (2) neglect the uncertainty in angular position, $l,b$ (for a typical RR Lyrae at 50~kpc this is $\sim$$10^{-7}$~deg for \gaia); (3) assume we can measure radial velocities to these stars with 5~km/s uncertainty; and (4) compute the sky-averaged \gaia\, proper-motion uncertainty for each star assuming an F0V spectral type using the \texttt{PyGaia}\footnote{https://github.com/agabrown/PyGaia} code and use this uncertainty for both components of proper motion.
+For this experiment, we (1) assume the stars are RR Lyrae stars (e.g., bright distance indicators) such that the fractional distance uncertainty is $2\%$; (2) neglect the uncertainty in angular position, $l,b$ (for a typical RR Lyrae at 50~kpc this is $\sim$$10^{-7}$~deg for \gaia); (3) assume we can measure radial velocities to these stars with 5~km/s uncertainty; and (4) compute the sky-averaged \gaia\, proper-motion uncertainty for each star assuming an F0V spectral type using the \texttt{PyGaia}\footnote{\url{https://github.com/agabrown/PyGaia}} code and use this uncertainty for both components of proper motion.
 We further assume that we know the tail assignment for each star, $\beta$.
@@
 Other parameters --- potential parameters and $\Loffset$ --- are initialized by drawing from the priors summarized in Table~\ref{tbl:params}.
-We burn in the walkers for 50000 steps and run for 100000 inference steps, but thin the chains by taking every $2\times \max(t_{\rm acor})$ sample, where $\max(t_{\rm acor}) \approx 4000$~steps.
+We burn in the walkers for 50000 steps and run for 100000 inference steps, but thin the chains by taking every $\max(t_{\rm acor})$ sample, where $\max(t_{\rm acor}) \approx 4000$~steps.
 Figures~\ref{fig:exp2_potential}, \ref{fig:exp2_satellite}, \ref{fig:exp2_particle} show the marginalized posteriors for the potential parameters, satellite parameters, and parameters for a single particle.
 
-\begin{figure}[!ht]
+\begin{figure*}[!h]
 \begin{center}
@@
 \end{center}
-\end{figure}
+\end{figure*}
 
-\begin{figure}[!ht]
+\begin{figure*}[!h]
 \begin{center}
@@
 We burn in the walkers for 50000 steps and run for 500000 inference steps.
-We again thin the chains by taking every $2\times \max(t_{\rm acor})$ sample, but here the autocorrelation time is found to be very long, $\max(t_{\rm acor}) \approx 25000$~steps.
+We again thin the chains by taking every $\max(t_{\rm acor})$ sample, but here the autocorrelation time is found to be very long, $\max(t_{\rm acor}) \approx 15000$~steps.
 Figures~\ref{fig:exp3_potential}, \ref{fig:exp3_satellite}, \ref{fig:exp3_particle} show the marginalized posteriors for the potential parameters, satellite parameters, and parameters for a single particle.
@@
 Each step in parameter space with \rewinder\ requires integrating the orbits of stars and the progenitor: we presently integrate with a fixed timestep using leapfrog integration.
-Computing the acceleration due to the given potential at each step is implemented in \project{Cython} (and approaches \project{C}-like speed), but the rest of the code is written in pure-\project{Python}.
-Though already significantly less computationally intensive than running a full N-body simulation for every parameter step, running each MCMC walker for 10000 steps requires parallelization and a few hours of CPU time on a compute cluster.
+Computing the acceleration due to the given potential at each step is implemented in \project{Cython} (and approaches \project{C}-like speed), but the rest of the code is written in pure-\project{Python}.\footnote{All code use for this project is available at \url{https://github.com/adrn/streams}.} Though already significantly less computationally intensive than running a full N-body simulation for every parameter step, running each MCMC walker for $\gtrsim$100000 steps requires parallelization and many hours of CPU time on a compute cluster.
 Further optimization of the integration method (e.g., using an adaptive method or implementing in \project{Cython}) could speed up the inference significantly.
@@
 
-\rewinder\, works well when all stars are lost due to tidal pruning.
+%\rewinder\, works well when all stars are lost due to tidal pruning.
 Even on a mildly eccentric orbit, stars are preferentially stripped around the satellite's pericenter because of tidal shocking.
@@
 With a proper model for the tidal shocking that occurs on an eccentric orbit, \rewinder\ could also be applied to structures such as the Triangulum-Andromeda \citep{rochapinto04} or Hercules-Aquila \citep{belokurov07b} stellar clouds, both possible shell-like structures viewed from the inside.
-
-We model tidal debris as having been stripped from the progenitor with some dispersion about effective Lagrange points.
+%
+%We model tidal debris as having been stripped from the progenitor with some dispersion about effective Lagrange points.
 We assume that the distributions in position and velocity at these points are isotropic Gaussians.

### 2014-04-30 scripts/figures/paper2_figures.py
bounds

@@ -539,6 +539,7 @@ def exp1_posterior():
         bounds.append((0.95*truth, 1.05*truth))
         flatchain[:,ii] = _unit_transform[p.name](_flatchain[:,ii])
 
+    bounds = [(0.7,2.),(0.7,2.),(52,142),(100,200),(1.1,2.5)]
     fig = triangle.corner(flatchain, plot_datapoints=False,
                           truths=truths, extents=bounds, labels=labels)
     fig.savefig(os.path.join(plot_path, "exp1_posterior.png"))
@@ -560,7 +561,7 @@ def exp_posteriors(exp_num):
     # Potential
     this_flatchain = np.zeros((_flatchain.shape[0], len(d["potential"])))
     truths = []
-    bounds = [(0.7,2.),(0.7,2.),(52,142),(100,200)]
+    bounds = [(0.7,2.),(0.7,2.),(52,142),(100,200),(1.1,2.5)]
     labels = []
     for ii,pname in enumerate(d["potential"].keys()):
         this_flatchain[:,ii] = _unit_transform[pname](np.squeeze(d["potential"][pname]))

### 2014-05-07 text/paper2/ms.tex
remove one todo
comments from KVJ and DWH

@@
 \newcommand{\vhalo}{v_{\rm h}}
-\newcommand{\rewinder}{\emph{Rewinder}}
+\newcommand{\rewinder}{\texttt{Rewinder}}
 \newcommand{\potp}{$q_1,q_z,\phi,\vhalo$}
@@
   of smaller galaxies or globular clusters will provide very precise measures of 
-  the gravitational potential, 
+  the gravitational potential 
   to test these predictions.
@@
   were close in phase space at some point in the past.
-Our method is flexible enough to adapt to any form for the Galactic potential.
+Our method can flexibly adapt to any form for the Galactic potential.
 In particular, it works in phase-space rather than action-space and hence relies neither on
@@
   and precisions around 25 percent for recovering missing phase-space coordinates.
-The method generalizes to encompass mixtures of independent debris structures,
+These results are are encouraging for the eventual goal to use more flexible, time-dependent potential models.
+The method generalizes to encompass mixtures of independent debris structures
   or situations in which star membership in structures is unknown.
@@
 These processes combined with the gravity from a baryonic disk or ellipsoid can act to make the inner halo more oblate or spherical, however they do not seem to erase the clumpy, triaxial nature of the outer halo \citep[e.g.,][]{pontzen12}.
-This can lead to radially-dependent axis ratios, orientation, and smoothness, and suggests that the true mass distributions around Milky Way-like galaxies are not easily represented by simple potentials.
+This can lead to radially-dependent axis ratios, orientation, and smoothness, and suggests that the true mass distributions around Milky Way-like galaxies are not easily represented by simple, time-independent potentials.
 Methods that seek to measure the gravitational potentials around such galaxies must be flexible enough to handle generic potential forms where finding simple analytic approximations or computing actions may not be possible.
@@
 
-Within the Milky Way our unique vantage point allows us a three-dimensional view of stars within in our own dark matter halo.
+Within the Milky Way our unique vantage point allows us a three-dimensional view of stars within our own dark matter halo.
 Our proximity allows us to use individual stars as kinematic tracers and hence build much larger samples that probe deeper into the halo than the globular cluster and planetary nebula studies of external galaxies.
@@
 The computational costs of running N-body simulations limited their search to a grid of potential parameters and forced the authors to fix many other parameters (e.g., properties of the disk and bulge).
-Nevertheless, they were able to constrain the 3D shape of the Milky Way potential out to $\sim$70 kpc and found that the best-fitting halo has a prolate, triaxial shape elongated in the Galactic Z direction, orthogonal to the plane of the disk.
+Nevertheless, they were able to constrain the 3D shape that their assumed potential model must assume in order to best represent the Milky Way out to $\sim$70 kpc and found that the best-fitting halo has a prolate, triaxial shape flattened in the Galactic X direction.
 Though an unlikely orientation for the halo --- \cite{debattista13} find that the disk of the Milky Way would not remain stable in such a configuration --- LM10 showed that the data is at a state where such inference is possible.
@@
 To account for the offset between the orbit of the progenitor and the orbits of the debris stars, methods have been proposed that add some dispersion or offset around a single orbit either in phase space \citep[e.g.,][]{eyre09a, varghese11, kuepper12} or action-angle coordinates \citep{eyre11, sanders13b, bovy14, sanders14}.
-In this work we do \emph{not} assume that streams follow the orbit of the progenitor.
-Other statistical methods have been proposed \citep[][]{penarrubia12, johnston99a} that may prove powerful when applied to, for example, data from the \gaia\, mission, where full 6D coordinates will be known for large samples of stars in the halo --- and therefore many debris structures --- but stream membership is not known for all stars.
-%sanderson14
+Other statistical methods have been proposed \citep[][]{johnston99a, penarrubia12, sanderson14} that may prove powerful when applied to, for example, data from the \gaia\, mission, where full 6D coordinates will be known for large samples of stars in the halo --- and therefore many debris structures --- but stream membership is not known for all stars.
 
@@
 	\item \textbf{Form of the potential:} There is large uncertainty in the radial profile, shape, orientation, and graininess of the outer halo and the constancy of these parameters over distance.
-Properly describing the potential may require non-parametric techniques or complex analytic functions so the method should not rely on the existence of or ability to compute simple conserved orbital properties, such as actions.
+Properly describing the potential may require non-parametric techniques or complex analytic functions so the method should not rely on the existence of or ability to compute conserved orbital properties such as actions.
 	\item \textbf{Multiple debris structures:} Near-future photometric surveys such as \gaia\, and the \project{LSST} will likely discover many new streams and kinematic associations of stars.
@@
 	\item \textbf{Comparing models to data:} Matching generated streams to observed stellar densities is difficult; models that rely on this must account for observational biases, internal properties of the progenitor, and background halo stars.
-[TODO: are we matching densities?]
 	\item \textbf{Computational expense:} Full N-body simulations are expensive to run; incorporating an N-body simulation into a likelihood function evaluation and then performing a parameter search would be computationally intensive and, presently, intractable.
@@
 
-In Section~\ref{sec:sims} we describe a suite of N-body simulations that span a range of progenitor masses on a characteristic, mildly eccentric orbit.
-In Section~\ref{sec:method} we use these simulations to motivate a new, flexible model for tidal streams that works entirely in phase-space.
+In Section~\ref{sec:sims} we describe a suite of N-body simulations that span a range of progenitor masses on a characteristic, mildly eccentric orbit and then use them in Section~\ref{sec:method} to motivate a new, flexible model for tidal streams that works entirely in phase-space.
 In Section~\ref{sec:experiments} we demonstrate how \rewinder\ can be used to measure properties of a non-trivial Galactic potential by performing several experiments with simulated observations of data from N-body simulations.
@@
 \section{Simulations}\label{sec:sims}
-We performed a set of N-body simulations with the SCF basis function expansion code \citep{hernquist92} to build realistic models of streams for studying the phase-space distribution of debris as it is stripped from its progenitor.
+We performed a set of N-body simulations with the Self-Consistent Field (SCF) basis function expansion code \citep{hernquist92} to build realistic models of streams for studying the phase-space distribution of debris as it is stripped from its progenitor.
 We later use one of these simulations to test \rewinder.
@@
 
-We take care to record the time at which each particle is unbound from the satellite in the simulations: we locate the position of the remnant iteratively by first calculating the satellite potential with all particles, remove particles with kinetic energies sufficient to escape, then recalculate the potential without those particles until the system converges.
+We take care to record the time at which each particle is unbound from the satellite in the simulations: we locate the position of the remnant iteratively by first calculating the satellite potential with all particles, removing particles with kinetic energies sufficient to escape, then recalculate the potential without those particles until the system converges.
 There is no spatial restriction on where a particle may become unbound and all particles (both bound and unbound) contribute their gravity during normal timesteps, presuming that they are resolved by the basis functions.
@@
 \caption{ Particle positions (grey dots) in Galactocentric cartesian coordinates from the final timestep of four N-body simulations (Section~\ref{sec:sims}) with the same progenitor orbit initial conditions over a range of progenitor masses (columns).
-Black crosses indicate eight particles chosen from each mass simulation and used in the experiments described in Section~\ref{sec:experiments}.}\label{fig:sims}
+Black crosses indicate four particles chosen from each mass simulation and used in the experiments described in Section~\ref{sec:experiments}.}\label{fig:sims}
 \end{center}
@@
 
-The velocity of the satellite will be of order $V\sim \sqrt{GM_{\rm enc}/R}$. If the velocity dispersion of the satellite is $\sigma_v \sim \sqrt{Gm/r_J}$, then it follows
+The bulk velocity of the satellite will be of order $V\sim \sqrt{GM_{\rm enc}/R}$. If the velocity dispersion of the satellite is $\sigma_v \sim \sqrt{Gm/r_J}$, then it follows
 \begin{equation}
@@
 \end{equation}
-We expect the debris-star velocities also to scale with $(m/M_{\rm enc})^{1/3}$.
+\citep[as pointed out in][]{binneytremaine} and hence expect the debris-star velocities also to scale with $(m/M_{\rm enc})^{1/3}$.
 These scalings assume that the satellite is spherical with isotropic velocities; non-random, internal satellite orbits (e.g., a disk) will break these assumptions.
@@
 In these coordinates, the classical Lagrange points would be located at $x_1\approx\pm\rtide,x_2=0,x_3=0$ (illustrated by the intersection of the black unit circles and horizontal lines in Figure~\ref{fig:lpts_r}).
-Though not centered on the point-mass Lagrange points, the location of and dispersion about the effective Lagrange points in these scaled coordinates remain remarkably consistent across the range of progenitor masses explored.
+Though not exactly centered on the point-mass Lagrange points, the location of and dispersion about the effective Lagrange points in these scaled coordinates remain remarkably consistent across the range of progenitor masses explored.
 Figure~\ref{fig:lpts_v} shows the velocity orbits of each star also projected into these coordinates and normalized.
@@
 
-We find that even in a triaxial potential, the dispersion --- in position and velocity --- of tidally stripped debris as it comes unbound from the satellite scales with the mass ratio $(m / M_{\rm enc})^{1/3}$.
+We conclude that even on a non-circular orbit in a complex potential, the dispersion --- in position and velocity --- of tidally stripped debris as it comes unbound from the satellite scales with the mass ratio $(m / M_{\rm enc})^{1/3}$.
 This motivates a model for tidal debris in which each star was ``released'' at the instantaneous, effective Lagrange point at its unbinding time, $\tub$, with a dispersion in position and velocity, all of which depend only on the mass and orbit of the progenitor and the parent potential.
@@
 In general, the data for the star and progenitor will have significant uncertainties or missing dimensions at distances typical to the Galactic halo, thus we define $\W$ and $\W_\sat$ as the true 6D, heliocentric positions of the star and progenitor and will include these in our model.
-Since we include these positions as parameters, this will work even when the star has missing dimensions or the progenitor location is unknown \citep[as in the Orphan stream,][]{belokurov07}.
+Since we include these positions as parameters, this method will work even when the star has missing dimensions or the progenitor location is unknown \citep[as in the Orphan stream,][]{belokurov07}.
 
@@
 The present-day phase-space position is the result of integrating this sample forward from $\tub$ in the parent potential, $\Phi$, whose form is parametrized by the vector $\bs{\Theta}_\Phi$.
-We assume that once the star becomes ``unbound'' from the satellite the potential of the satellite can be ignored,\footnote{This assumption breaks down for sufficiently large-mass progenitors, somewhere between $\sim10^8-10^9~\msun$.} and thus treat the star as a test particle.
+We assume that once the star becomes ``unbound'' from the satellite the potential of the satellite can be ignored,\footnote{This assumption breaks down for sufficiently large-mass progenitors, somewhere between $\sim10^8-10^9~\msun$.} and thus we treat the star as a test particle.
 The satellite mass enters this prescription through the tidal radius; a changing satellite mass will skew the positions of the effective Lagrange points and velocity scale.
@@
 To compress notation, we pack all progenitor parameters ($\W_\sat, m_0, \dot{m}, \Loffset$) into the vector $\bs{\Theta}_\sat$.
-The likelihood is then: %, and all star parameters ($\W, \tailbit, \tub$) into the vector $\bs{\Theta}$
+The likelihood for this model is: %, and all star parameters ($\W, \tailbit, \tub$) into the vector $\bs{\Theta}$
 \begin{align}
@@
 \end{align}
-where the interaction time is taken to be $t_{\rm int}=6.2$~Gyr.
+where the interaction time is taken to be $t_{\rm int}=6.2$~Gyr, the total duration of the simulations of Section~\ref{sec:sims}.
 
@@
 In what follows, we use \rewinder\ to model ``observations'' of a small sample of stars from one of the N-body simulations described in Section~\ref{sec:sims}.
-We consider this a strong test of the method because the data are generated with N-body simulations and are not generated with anything resembling the likelihood function presented above.
+We consider this a strong test of the method because the data are generated with N-body simulations and not with anything resembling the likelihood function presented above.
 For all tests in this article, we use the same functional form for the potential as used in the simulations (Equation~\ref{eq:lm10}).
-When recovering the potential, we hold fixed the disk and spheroid parameters (see Table~\ref{tbl:params}), along with one halo parameters: the scale radius, $r_h$.
+When recovering the potential, we hold fixed the disk and spheroid parameters (see Table~\ref{tbl:params}), along with one of the halo parameters: the scale radius, $r_h$.
 The priors on the remaining halo parameters are taken to be uniform over a conservative domain of realistic values: for $\vhalo$, 100-200~km/s corresponds to a range in solar circular velocities from $\sim$210-250 km/s (holding other parameters fixed); the range in axis ratios allow for prolate, oblate, and generic triaxiality; and $\phi$ is restricted to $\pm45^\circ$ around the true simulation value, $\phi = 97^\circ$.
@@
 We instead use an affine-invariant ``ensemble'' sampler \citep{goodman10} that, each step in parameter space, updates the positions of many ``walkers'' (the ensemble).
-This algorithm is implemented in the \project{Python} programming language \citep{foremanmackey13} and runs naturally in a parallelized environment such as the message passing interface (MPI).
+This algorithm is implemented in the \texttt{Python} programming language \citep{foremanmackey13} and runs naturally in a parallelized environment such as the message passing interface (MPI).
 In each experiment we run the walkers for a large number of steps from starting positions described below, then throw away these samples and take the positions from the final step of this ``burn-in'' phase as a starting point for the samples used for inference.
@@
 		tail assignment (binary) & $\beta$ & -- & $\pm1$~equally likely (fixed at truth)\\ 
-		time unbound & $\tub$ & -- & $\mathcal{U}(0,6.2)$ Gyr\\ 
 		\bottomrule
@@
 $\mathcal{N}$ is the normal (Gaussian) distribution, and $\mathcal{U}$ the uniform distribution.
-There are 11 parameters for the Milky Way potential, but only four are left free to vary; some parameters are fixed (denoted by ``(fixed)'') at the true values used in the N-body simulations that generated the fake data.
+There are 11 parameters for the Milky Way potential, but only four are left free to vary; some parameters are fixed (denoted by ``(fixed)'') at the true values used in the N-body simulations that generated the fake test data.
 The progenitor has nine parameters --- the position, $\bs{r}_{\rm p}$, and velocity, $\bs{v}_{\rm p}$, vectors each contain three components --- but only five are left free to vary.
@@
 Sky coordinates are fixed, along with the tail assignment (whether the star belongs to the leading or trailing tail).
-For inference with eight stars, there are $4+5+8\times5=49$ free parameters.
+For inference with four stars, there are $4+5+4\times4=25$ free parameters.
 \label{tbl:params}}
@@
 Vertical and horizontal lines (blue) show the true, input values for the potential parameters used in the N-body simulations.
-For the potential parameters, the axis ranges were chosen to be $\pm$5\% around the true values.
-Note that any biases in the inferred values are $\lesssim1\%$.
+For the potential parameters, the axis ranges are chosen to be the same for this and the potential posterior plots to follow.
 }\label{fig:exp1_posterior}
@@
 Figures~\ref{fig:exp2_potential}, \ref{fig:exp2_satellite}, \ref{fig:exp2_particle} show the marginalized posteriors for the potential parameters, satellite parameters, and parameters for a single particle.
+The uncertainties on the potential parameters are only 5-7\%
 
@@
 Even a small sample of stars with precise distance measurements provide an enormous amount of information about the triaxiality of the potential.
+Fractional uncertainties on the potential parameters are $\sim$15\% and the initially missing proper-motions are recovered with $\sim$25\% uncertainties.
 
@@
 In this article, we only consider a stream on a Sgr-like orbit, which lies nearly in the Galactic $x$-$z$ plane.
-We find that the stream puts better constraints on the $z$-axis flattening, $q_z$, than on the flattening in the $x$-$y$ plane, $q_1$, as expected.
+We find that the stream puts better constraints on the $z$-axis flattening, $q_z$, than on the flattening in the $x$-$y$ plane, $q_1$, as one might expect.
 The best measurements of the detailed shape of the Galactic potential will come from combining constraints from multiple streams.
@@
 Each step in parameter space with \rewinder\ requires integrating the orbits of stars and the progenitor: we presently integrate with a fixed timestep using leapfrog integration.
-Computing the acceleration due to the given potential at each step is implemented in \project{Cython} (and approaches \project{C}-like speed), but the rest of the code is written in pure-\project{Python}.\footnote{All code use for this project is available at \url{https://github.com/adrn/streams}.} Though already significantly less computationally intensive than running a full N-body simulation for every parameter step, running each MCMC walker for $\gtrsim$100000 steps requires parallelization and many hours of CPU time on a compute cluster.
-Further optimization of the integration method (e.g., using an adaptive method or implementing in \project{Cython}) could speed up the inference significantly.
+Computing the acceleration due to the given potential at each step is implemented in \texttt{Cython} (and approaches \texttt{C}-like speed), but the rest of the code is written in pure-\texttt{Python}.\footnote{All code use for this project is available at \url{https://github.com/adrn/streams}.} Though already significantly less computationally intensive than running a full N-body simulation for every parameter step, running each MCMC walker for $\gtrsim$100000 steps requires parallelization and many hours of CPU time on a compute cluster.
+Further optimization of the integration method (e.g., using an adaptive method or implementing in \texttt{Cython}) could speed up the inference significantly.
 
@@
 Without proper motion data the uncertainties are around 15 percent.
-TODO: concluding sentence.
+We consider this work to be an encouraging first step towards the goal of recovering the (presumably) much more complex --- time-dependent, clumpy, with axis ratios and orientations that vary with radius --- Milky Way potential with larger data sets.
 
@@
 We thank Anthony Brown (Leiden) for insight on the \gaia\, data quality and for providing the open source \texttt{PyGaia} code for computing predicted uncertainties.
-We thank the organizers of the Gaia Challenge (2013) and the University of Surrey for hospitality during this workshop.
-We thank Hans-Walter Rix (MPIA) and Jo Bovy (IAS) for ...
-[TODO: HOGG].
-[TODO: thank Marla, Ana?]
+We thank the organizers of the Gaia Challenge (2013) and the University of Surrey for hospitality during the workshop.
+We thank Ana Bonaca (Yale), Jo Bovy (IAS), Marla Geha (Yale), Andreas K{\"u}pper (Columbia), and Hans-Walter Rix (MPIA) for useful discussions.
+KVJ thanks the Institute of Astronomy for hospitality.
 APW is supported by a National Science Foundation Graduate Research Fellowship under Grant No.\ 11-44155.
 This work was supported in part by the National Science Foundation under Grant No.
-PHYS-1066293.
-DWH was partially supported by NASA (Grant NNX12AI50G) and the NSF (Grant IIS-1124794).
-This research made use of Astropy, a community-developed core Python package for Astronomy \citep{astropy13}.
+PHYS-1066293 and AST-1312196.
+DWH was partially supported by the NSF (Grant IIS-1124794) and the Moore-Sloan Data Science Environment at NYU.
+This research made use of Astropy, a community-developed core \texttt{Python} package for Astronomy \citep{astropy13}. This research has made use of NASA's Astrophysics Data System.
 This work additionally relied on Columbia University's \emph{Hotfoot} and \emph{Yeti} compute clusters, and we acknowledge the Columbia HPC support staff for assistance. \\

### 2014-05-07 text/paper2/ms.tex
bonaca

@@
-%\documentclass{emulateapj}
-\documentclass[letterpaper,12pt,preprint]{aastex}
+\documentclass{emulateapj}
+%\documentclass[letterpaper,12pt,preprint]{aastex}
 
@@
 	\item transform star and progenitor positions in heliocentric coordinates to Galactocentric, cartesian coordinates;
-	\item integrate star and progenitor orbits backwards as test particles in the potential given by $\bs{\Theta}_\Phi$ for the total interaction time, $t_{\rm int}\approx6$~Gyr;
+	\item integrate star and progenitor orbits backwards as test particles in the potential given by $\bs{\Theta}_\Phi$ for the total interaction time, $t_{\rm int}\approx6.2$~Gyr;
 	\item transform the particle position to the relative, normalized coordinates defined in Section~\ref{sec:debris};
@@
 
-\begin{table}[h]
+\begin{table*}[ht]
 \begin{center}
@@
 		velocity & $\bs{v}_0$ & -- & $\|\bs{v}_0\|\sim\mathcal{U}(0,500)$~km/s\\
-		tail assignment (binary) & $\beta$ & -- & $\pm1$~equally likely (fixed at truth)\\ 
+		tail assignment & $\beta$ & -- & $\pm1$~equally likely (fixed at truth)\\ 
 		\bottomrule
@@
 \end{center}
-\end{table}
+\end{table*}
 
@@
 Bonaca et al.
-(in prep.) show that fitting a rigid potential model to the live potential of the \project{Via Lactea} halo ...TODO.
+(in prep.) show that using a static potential model to fit the live potential of the \project{Via Lactea} halo with tidal streams can introduce significant biases to measurements of the halo mass, even with perfect knowledge of the 6D coordinates of stars in the streams.
 In future work, we will (1) try fitting incorrect potential models to see if we can still recover global properties (e.g., flattening); (2) run simulations with smoothly changing potentials \citep[e.g.,][]{buist14} and attempt to model the time dependence; and (3) explore using a generic, non-parametric potential form (e.g., a basis function expansion) as the recovery potential model.
@@
 We thank Ana Bonaca (Yale), Jo Bovy (IAS), Marla Geha (Yale), Andreas K{\"u}pper (Columbia), and Hans-Walter Rix (MPIA) for useful discussions.
-KVJ thanks the Institute of Astronomy for hospitality.
+KVJ thanks the Institute of Astronomy at the University of Cambridge for hospitality.
 APW is supported by a National Science Foundation Graduate Research Fellowship under Grant No.\ 11-44155.

### 2014-05-10 text/paper2/ms.tex
minor changes and added ref

@@
-\documentclass{emulateapj}
-%\documentclass[letterpaper,12pt,preprint]{aastex}
+%\documentclass{emulateapj}
+\documentclass[letterpaper,12pt,preprint]{aastex}
 
@@
 our ability to derive actions nor on the 
-assumption of integrability of the potential;
-  we do not know that any good approximation to the large-scale 
-  Milky Way potential is integrable.
+assumption of integrability of the potential.
 % Methods: 
 Our model is probabilistic, with a likelihood function and priors on the parameters.
-  These parameters include the present-day phase-space position for each star
-  and the progenitor, and thus the method can properly account for both finite observational uncertainties 
+The method can properly account for both finite observational uncertainties 
   and missing data dimensions.
 We test our method on synthetic data sets generated from an N-body simulation
-of satellite disruption in a multi-component Milky Way including a triaxial dark matter halo. 
-The data sets were generated with
+of satellite disruption in a static, multi-component Milky Way including a triaxial dark matter halo. 
+The mock data sets were generated with
   observational uncertainties to mimic current and near-future surveys of various stars.
@@
   and precisions around 25 percent for recovering missing phase-space coordinates.
-These results are are encouraging for the eventual goal to use more flexible, time-dependent potential models.
-The method generalizes to encompass mixtures of independent debris structures
-  or situations in which star membership in structures is unknown.
+These results are encouraging for the eventual goal of using more flexible, time-dependent potential models 
+  combined with larger data sets to unravel the detailed shape of the dark matter distribution around the Milky Way.
 % Conclusions:
+% The method generalizes to encompass mixtures of independent debris structures
+%   or situations in which star membership in structures is unknown.
+
 \end{abstract}
@@
 The computational costs of running N-body simulations limited their search to a grid of potential parameters and forced the authors to fix many other parameters (e.g., properties of the disk and bulge).
-Nevertheless, they were able to constrain the 3D shape that their assumed potential model must assume in order to best represent the Milky Way out to $\sim$70 kpc and found that the best-fitting halo has a prolate, triaxial shape flattened in the Galactic X direction.
+Nevertheless, they were able to constrain the 3D shape that their assumed potential model must assume in order to best represent the Milky Way out to $\sim$70 kpc and found that the best-fitting halo has a nearly oblate (only mildly triaxial) shape flattened in a direction close to the Sun-Galactic center line.
 Though an unlikely orientation for the halo --- \cite{debattista13} find that the disk of the Milky Way would not remain stable in such a configuration --- LM10 showed that the data is at a state where such inference is possible.
@@
 In Section~\ref{sec:experiments} we demonstrate how \rewinder\ can be used to measure properties of a non-trivial Galactic potential by performing several experiments with simulated observations of data from N-body simulations.
-We discuss the results of these experiments in Section~\ref{sec:discussion}.
+In Section~\ref{sec:discussion}, we discuss the results of these experiments and the extent to which we address the above points 1-5.
 We conclude in Section~\ref{sec:conclusion}.
@@
 In these coordinates, the classical Lagrange points would be located at $x_1\approx\pm\rtide,x_2=0,x_3=0$ (illustrated by the intersection of the black unit circles and horizontal lines in Figure~\ref{fig:lpts_r}).
-Though not exactly centered on the point-mass Lagrange points, the location of and dispersion about the effective Lagrange points in these scaled coordinates remain remarkably consistent across the range of progenitor masses explored.
+Though not exactly centered on the point-mass Lagrange points derived for a circular orbit, the location of and dispersion about the effective Lagrange points in these scaled coordinates remain remarkably consistent across the range of progenitor masses explored.
 Figure~\ref{fig:lpts_v} shows the velocity orbits of each star also projected into these coordinates and normalized.
@@
 
-%\begin{figure}[h]
-%\begin{center}
-%\includegraphics[width=\textwidth]{../../plots/paper2/num_recombine.png}
-%\caption{ TODO }\label{fig:R_V}
-%\end{center}
-%\end{figure}
-
 \subsection{Probabilistic model}
@@
 Finally, we add two additional parameters: a constant, global factor, $\Loffset$, that scales the position of the effective Lagrange points relative to the classical tidal radius, and a binary parameter $\tailbit$, that is either $-1$ or $+1$ depending on whether the star is in the leading or trailing tail.
-To compress notation, we pack all progenitor parameters ($\W_\sat, m_0, \dot{m}, \Loffset$) into the vector $\bs{\Theta}_\sat$.
-The likelihood for this model is: %, and all star parameters ($\W, \tailbit, \tub$) into the vector $\bs{\Theta}$
+To compress notation, we pack all progenitor parameters into the vector $\bs{\Theta}_\sat = (\W_\sat, m_0, \dot{m}, \Loffset)$.
+The likelihood for this model is:
 \begin{align}
@@
 $\left\vert\J\right\vert = \left\vert\frac{\partial(x,y,z,v_x,v_y,v_z)}{\partial(l,b,d,\mu_l,\mu_b,v_r)}\right\vert$ is the absolute value of the determinant of the Jacobian that defines the transformation from heliocentric, spherical to Galactocentric, cartesian coordinates.
-
-The likelihood above is evaluated for each star at its unbinding time, however this is not a priori known for any star.
-Furthermore, the idea of a single-valued unbinding time is not well defined as (1) it depends strongly on the particular definition of ``unbound'' and (2) for debris disrupted on near-circular orbits, the debris may pass close to the progenitor several times.
-We marginalize the likelihood over all possible unbinding times, assuming a uniform distribution of unbinding times over the entire interaction time:
+We then marginalize the likelihood over all possible unbinding times, assuming a uniform distribution of unbinding times over the entire interaction time:
 \begin{align}
@@
 where the interaction time is taken to be $t_{\rm int}=6.2$~Gyr, the total duration of the simulations of Section~\ref{sec:sims}.
+
+The likelihood above is evaluated for each star at all time-steps of the simulation during the marginalization --- that is, we consider all possible unbinding times for all stars because there is no way to independently measure $\tub$ for a given star.
+Furthermore, the idea of a single-valued unbinding time is not well defined as (1) it depends strongly on the particular definition of ``unbound'' and (2) for debris disrupted on near-circular orbits, the debris may pass close to the progenitor several times.
+We are effectively modeling the distribution of debris as it comes unbound from the satellite, however the exact shape of this distribution in phase-space will depend on the same complicating factors as above.
+Our choice of using a Gaussian to model the distribution of debris as it is disrupted is simple and almost certainly wrong, but we have found that the exact shape of this distribution is neither well-defined nor significant for a spherical satellite.
+For example, we find that changing this distribution to a log-normal distribution does not affect our conclusions.
 
@@
 The MCMC chains do not display any bulk movement or stray walkers, implying, by eye, that the chains have converged.
-Projections of the sample densities are shown in Figure~\ref{fig:exp1_posterior}.
+We also compute the autocorrelation functions for each parameter and find that, for this experiment, the autocorrelation times are short (as shown on each panel in this figure).
+Projections of the binned samples (posterior densities) are shown in Figure~\ref{fig:exp1_posterior}.
 }\label{fig:trace}
@@
 {\bf 2.
-Form of the potential:} No assumptions were made about the the form of the Galactic potential used in \rewinder: the potential could be time-dependent, clumpy, or have properties that vary with radius.
+Form of the potential:} No assumptions were made about the the form of the Galactic potential in deriving the likelihood for \rewinder.
+The simple experiments in this article infer potential parameters from the same static, analytic potential used in the N-body simulations that generated the fake data.
+However, in principle the potential could be time-dependent, clumpy, or have properties that vary with radius.
 The orbits of the stars and progenitor are directly integrated and only require a function that evaluates the acceleration due to the potential at a given position (and time).
-The simple experiments in this article infer potential parameters from the same functional potential used in the N-body simulations that generated the fake data.
-We have shown that using this simple model, \emph{precise} measurements of the Milky Way potential are attainable with near-future data, however we have not discussed the \emph{accuracy} of such measurements.
-Bonaca et al.
+We have shown that with the correct form for the Galactic potential, \emph{precise} measurements of the potential parameters are attainable with near-future data, but we have not discussed the \emph{accuracy} of such measurements due to incorrect assumptions.
+For example, Bonaca et al.
 (in prep.) show that using a static potential model to fit the live potential of the \project{Via Lactea} halo with tidal streams can introduce significant biases to measurements of the halo mass, even with perfect knowledge of the 6D coordinates of stars in the streams.
@@
 We thank Ana Bonaca (Yale), Jo Bovy (IAS), Marla Geha (Yale), Andreas K{\"u}pper (Columbia), and Hans-Walter Rix (MPIA) for useful discussions.
-KVJ thanks the Institute of Astronomy at the University of Cambridge for hospitality.
+KVJ thanks the Institute of Astronomy at the University of Cambridge for hospitality and Vasily Belokurov (IoA) and his group for discussion of his work while there.
 APW is supported by a National Science Foundation Graduate Research Fellowship under Grant No.\ 11-44155.

### 2014-05-10 streams/potential/_lm10_acceleration.pyx
added jerk function

@@ -31,6 +31,62 @@ cdef extern from "math.h":
 #DTYPE = np.double
 #ctypedef np.double_t DTYPE_t
 
+@cython.boundscheck(False) # turn of bounds-checking for entire function
+@cython.cdivision(True)
+@cython.wraparound(False)
+@cython.nonecheck(False)
+def lm10_potential(np.ndarray[double, ndim=2] r, int n_particles,
+                   np.ndarray[double, ndim=1] pot,
+                   double q1, double qz, double phi, double v_halo,
+                   double q2, double R_halo):
+
+    cdef double fac1, fac2, fac3, _tmp, _tmp_1, _tmp_2, R_pl_c
+    cdef double G, a, b_sq, c, m_disk, m_bulge
+    G = 4.49975332435e-12 # kpc^3 / Myr^2 / M_sun
+
+    # Miyamoto-Nagai
+    a = 6.5 # kpc
+    b_sq = 0.26*0.26 # kpc^2
+    Gm_disk = G*1.E11 # M_sun
+
+    # Hernquist
+    c = 0.7 # kpc
+    Gm_bulge = G*3.4E10 # M_sun
+
+    # Halo
+    cdef double q1_sq, q2_sq, qz_sq, R_halo_sq, v_halo_sq, sinphi, cosphi, C1, C2, C3
+    q1_sq = q1*q1
+    q2_sq = q2*q2
+    qz_sq = qz*qz
+    R_halo_sq = R_halo*R_halo # kpc
+    v_halo_sq = v_halo*v_halo
+    sinphi = sin(phi)
+    cosphi = cos(phi)
+    C1 = cosphi**2/q1_sq + sinphi**2/q2_sq
+    C2 = cosphi**2/q2_sq + sinphi**2/q1_sq
+    C3 = 2.*sinphi*cosphi*(1./q1_sq - 1./q2_sq)
+
+    cdef double x, y, z
+    cdef double xx, yy, zz
+
+    for ii in range(n_particles):
+        x = r[ii,0]
+        y = r[ii,1]
+        z = r[ii,2]
+
+        xx = x*x
+        yy = y*y
+        zz = z*z
+        R = sqrt(xx + yy + zz)
+
[…]

### 2014-05-12 text/paper2/ms.tex
paragraph

@@
 
-The likelihood above is evaluated for each star at all time-steps of the simulation during the marginalization --- that is, we consider all possible unbinding times for all stars because there is no way to independently measure $\tub$ for a given star.
-Furthermore, the idea of a single-valued unbinding time is not well defined as (1) it depends strongly on the particular definition of ``unbound'' and (2) for debris disrupted on near-circular orbits, the debris may pass close to the progenitor several times.
-We are effectively modeling the distribution of debris as it comes unbound from the satellite, however the exact shape of this distribution in phase-space will depend on the same complicating factors as above.
-Our choice of using a Gaussian to model the distribution of debris as it is disrupted is simple and almost certainly wrong, but we have found that the exact shape of this distribution is neither well-defined nor significant for a spherical satellite.
-For example, we find that changing this distribution to a log-normal distribution does not affect our conclusions.
+The likelihood above is evaluated for each star at all time-steps of the simulation during the marginalization with a uniform prior over all possible unbinding times (no assumptions are made about the orbital phase dependence of mass-loss).
+We chose this approach because the phase-space position (and hence time) at which a star becomes unbound from a satellite can only be strictly defined for the case of a spherical satellite orbiting on a circular orbit in a spherical potential.
+Various authors have chosen to measure mass loss in simulations of satellite disruption on eccentric orbits in non-spherical potentials by either looking at the kinetic energy of the particles relative to the satellite's potential energy \citep{johnston96} or looking at the point at which a star crosses the instantaneous tidal radius \citep{kuepper12}.
+Neither choice is strictly valid and both have been found useful for representing the mass-loss process.
+Hence we make the simplest choice of using Gaussians to characterize the phase-space distribution of debris as it makes the transition between being bound and unbound.
+We find that changing this distribution to a log-normal distribution does not affect our conclusions.
 

### 2014-05-12 text/paper2/ms.tex
final minor revisions

@@
 Such studies assume that the tracer orbits are randomly sampled from a smooth distribution function and are fully phase mixed.
-However, large photometric surveys such as the SDSS and 2MASS \citep{skrutskie06} have discovered large amounts of substructure --- in streams and kinematic associations of stars --- in the Milky Way halo \citep[e.g.,][]{belokurov06, rochapinto04}, thus demonstrating that the stellar distribution is neither on random orbits nor fully phase-mixed.
+However, large photometric surveys such as the SDSS and 2MASS \citep{skrutskie06} have discovered copious amounts of substructure --- in streams and kinematic associations of stars --- in the Milky Way halo \citep[e.g.,][]{belokurov06, rochapinto04}, thus demonstrating that the stellar distribution is neither on random orbits nor fully phase-mixed.
 Substructure in the form of stellar streams and clouds is known to bias mass and velocity inferences from random tracer methods by several tens of percent \citep{yencho06}.
@@
 The computational costs of running N-body simulations limited their search to a grid of potential parameters and forced the authors to fix many other parameters (e.g., properties of the disk and bulge).
-Nevertheless, they were able to constrain the 3D shape that their assumed potential model must assume in order to best represent the Milky Way out to $\sim$70 kpc and found that the best-fitting halo has a nearly oblate (only mildly triaxial) shape flattened in a direction close to the Sun-Galactic center line.
+Nevertheless, they were able to constrain the 3D shape that their assumed potential model must take in order to best represent the Milky Way out to $\sim$70 kpc and found that the best-fitting halo has a nearly oblate (only mildly triaxial) shape flattened in a direction close to the Sun-Galactic center line.
 Though an unlikely orientation for the halo --- \cite{debattista13} find that the disk of the Milky Way would not remain stable in such a configuration --- LM10 showed that the data is at a state where such inference is possible.
@@
 where $M_{\rm enc}$ is the instantaneous enclosed mass of the parent system within orbital radius $R$.\footnote{Note that this is the instantaneous orbital radius, \emph{not} the orbital radius at pericenter.}
-In general, the effective Lagrange points may not be symmetric about the center of the satellite and may deviate from the tidal radius by factors of order unity.
+In general, the Lagrange points may not be symmetric about the center of the satellite and may deviate from the tidal radius by factors of order unity.
 
@@
 \end{equation}
-\citep[as pointed out in][]{binneytremaine} and hence expect the debris-star velocities also to scale with $(m/M_{\rm enc})^{1/3}$.
+\citep[as pointed out in][]{binneytremaine} and hence we expect the debris-star velocities also to scale with $(m/M_{\rm enc})^{1/3}$.
 These scalings assume that the satellite is spherical with isotropic velocities; non-random, internal satellite orbits (e.g., a disk) will break these assumptions.
 
-In a triaxial potential, the orbital plane of the satellite is not fixed but we still expect there to be \emph{effective} Lagrange points for a spherical satellite system along the line of centers connecting the origin of the parent potential to the satellite --- that is, we expect the stars to be stripped at some characteristic radius near the tidal radius with some dispersion about these points.
+In a triaxial potential, the orbital plane of the satellite is not fixed but we still expect there to be \emph{effective} Lagrange points for a spherical satellite system along the line of centers connecting the origin of the parent potential to the satellite --- that is, we expect the stars to be stripped at some characteristic distance from the satellite (near the tidal radius) along the instantaneous position vector of the satellite with some dispersion about these points.
 We proceed by defining a coordinate system relative to the position and velocity of the progenitor, ($\bs{r}_p$,$\bs{v}_p$), rotated into the instantaneous orbital plane defined by $\bs{L} = \bs{r}_p \times \bs{v}_p$.
@@
 \includegraphics[width=\textwidth]{../../plots/paper2/Lpts_r.png}
-\caption{ Orbits of 2000 randomly drawn particles projected into the instantaneous orbital plane coordinates (Eqs.~\ref{eq:x1}-\ref{eq:x3}), normalized by the tidal radius (Eq.~\ref{eq:tidalradius}), and only shown within half a satellite crossing time around $t=\tub$ for each of the four progenitor masses.
+\caption{ Orbits of 2000 randomly-drawn, disrupted particles projected into the instantaneous orbital plane coordinates (Eqs.~\ref{eq:x1}-\ref{eq:x3}), normalized by the tidal radius (Eq.~\ref{eq:tidalradius}), and only shown within half a satellite crossing time around $t=\tub$ for each of the four progenitor masses.
 The orbits were integrated backwards from their present-day positions (final time step of the N-body simulations) as test particles without the potential of the progenitor.
-Horizontal black line shows $x_2=0$ (top panels) and $x_3=0$ (bottom panels), and the unit circle (black circle) shows the classical disruption radius in these coordinates.
+Horizontal black line shows $x_2=0$ (top panels) and $x_3=0$ (bottom panels), and the unit circle (black circle) illustrates the classical disruption radius in these coordinates.
 }\label{fig:lpts_r}
@@
 We allow for mass-loss by incorporating the initial satellite mass, $m_0$, and a constant mass-loss term, $\dot{m}$ into the model: $m(t) = m_0 - \dot{m}t$.
-Finally, we add two additional parameters: a constant, global factor, $\Loffset$, that scales the position of the effective Lagrange points relative to the classical tidal radius, and a binary parameter $\tailbit$, that is either $-1$ or $+1$ depending on whether the star is in the leading or trailing tail.
+Finally, we add two additional parameters: a constant, global factor, $\Loffset$, that scales the position of the effective Lagrange points relative to the classical tidal radius, and a binary parameter $\tailbit$ (one for each star), that is either $-1$ or $+1$ depending on whether the star is in the leading or trailing tail.
 To compress notation, we pack all progenitor parameters into the vector $\bs{\Theta}_\sat = (\W_\sat, m_0, \dot{m}, \Loffset)$.
@@
 $\left\vert\J\right\vert = \left\vert\frac{\partial(x,y,z,v_x,v_y,v_z)}{\partial(l,b,d,\mu_l,\mu_b,v_r)}\right\vert$ is the absolute value of the determinant of the Jacobian that defines the transformation from heliocentric, spherical to Galactocentric, cartesian coordinates.
-We then marginalize the likelihood over all possible unbinding times, assuming a uniform distribution of unbinding times over the entire interaction time:
+We then marginalize the likelihood over all possible unbinding times, assuming a uniform prior of unbinding times over the entire interaction time:
 \begin{align}
@@
 We chose this approach because the phase-space position (and hence time) at which a star becomes unbound from a satellite can only be strictly defined for the case of a spherical satellite orbiting on a circular orbit in a spherical potential.
-Various authors have chosen to measure mass loss in simulations of satellite disruption on eccentric orbits in non-spherical potentials by either looking at the kinetic energy of the particles relative to the satellite's potential energy \citep{johnston96} or looking at the point at which a star crosses the instantaneous tidal radius \citep{kuepper12}.
+Various authors have chosen to measure mass loss in simulations of satellite disruption on eccentric orbits in non-spherical potentials by either looking at the kinetic energy of the particles relative to the satellite's potential energy \citep{johnston96 TODO} or looking at the point at which a star crosses the instantaneous tidal radius \citep{kuepper12}.
 Neither choice is strictly valid and both have been found useful for representing the mass-loss process.
@@
 	\item transform star and progenitor positions in heliocentric coordinates to Galactocentric, cartesian coordinates;
-	\item integrate star and progenitor orbits backwards as test particles in the potential given by $\bs{\Theta}_\Phi$ for the total interaction time, $t_{\rm int}\approx6.2$~Gyr;
+	\item integrate star and progenitor orbits backwards as test particles in the potential given by $\bs{\Theta}_\Phi$ for the total interaction time, $t_{\rm int} (in this case, \approx6.2$~Gyr);
 	\item transform the particle position to the relative, normalized coordinates defined in Section~\ref{sec:debris};
@@
 We test \rewinder\ using only four stars (particles) --- two from the leading tail, and two from the trailing tail --- randomly sampled from the $2.5\times10^8\,\msun$ simulation (Section~\ref{sec:sims}), assuming the observed 6D positions for both the stars and the progenitor have negligible uncertainties.
-The stars were required to have been stripped after the first pericentric passage and have a present-day distance within $50$ kpc of the Sun.
+The stars were required to have been stripped after the first pericentric passage and have a present-day distance within $40$ kpc of the Sun.
 Figure~\ref{fig:sims} (second column from right) shows the randomly chosen stars (black crosses) in Galactic coordinates, over-plotted on all other simulated particles (grey points).
@@
 \caption{ Positions of all 64 MCMC walkers (see Section~\ref{sec:experiments}) for each parameter at each step of the inference (black lines).
-Horizontal green lines show the input values of the potential parameters (\potp).
 The MCMC chains do not display any bulk movement or stray walkers, implying, by eye, that the chains have converged.
@@
 Each step in parameter space with \rewinder\ requires integrating the orbits of stars and the progenitor: we presently integrate with a fixed timestep using leapfrog integration.
-Computing the acceleration due to the given potential at each step is implemented in \texttt{Cython} (and approaches \texttt{C}-like speed), but the rest of the code is written in pure-\texttt{Python}.\footnote{All code use for this project is available at \url{https://github.com/adrn/streams}.} Though already significantly less computationally intensive than running a full N-body simulation for every parameter step, running each MCMC walker for $\gtrsim$100000 steps requires parallelization and many hours of CPU time on a compute cluster.
+Computing the acceleration due to the given potential at each step is implemented in \texttt{Cython} (and approaches \texttt{C}-like speed), but the rest of the code is written in pure-\texttt{Python}.
+Though already significantly less computationally intensive than running a full N-body simulation for every parameter step, running each MCMC walker for $\gtrsim$100000 steps requires parallelization and many hours of CPU time on a compute cluster.
 Further optimization of the integration method (e.g., using an adaptive method or implementing in \texttt{Cython}) could speed up the inference significantly.
@@
 We have performed several experiments to show that \rewinder\ simultaneously constrains the potential and models a tidal stream given simulated data with a range of realistic uncertainties.
-We find that with realistic but future high-quality data --- that is, high precision distance measurements from RR Lyrae and proper motions from \gaia --- a small sample of just four stars in a tidal stream modeled with \rewinder\, provide measurements of potential parameters that rival present-day constraints from running full N-body simulations.
+We find that with future high-quality data --- that is, high precision distance measurements from RR Lyrae and proper motions from \gaia --- a small sample of just four stars in a tidal stream modeled with \rewinder\, provide measurements of potential parameters that rival present-day constraints from comparing full N-body simulations to large numbers of stellar tracers with poorly measured kinematics.
 For this high-quality data, we recover the input potential parameter values with uncertainties of order 5-7 percent.
